### 2016-05-11 bioc-workflow.Rmd
Added workflow manuscipt

---
title: "A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data"
author: "Lisa Breckels"
date: "11 May 2016"
output: html_document
---

# Forward

# Reading and handling mass-spectrometry based proteomics data

## The infrastructre: pRoloc and MSnbase in Bioconductor

## The use-case: prediction sub-cellular localisation in pluripotent embroyonic mouse stem cells

# Normalisation and Quality Control

# Markers

# Novelty Detection

# Supervised machine learning

### 2016-05-11 bioc-workflow.Rmd
Forward and use-case info

@@
 
+Quantitative mass spectrometry based spatial proteomics involves elaborate, expensive and time consuming experimental procedures and considerable effort is invested in the generation of such data.
+Multiple research groups have described a variety of approaches to establish high quality proteome-wide datasets.
+However, data analysis is as critical as data production for reliable and insightful biological interpretation.
+Here, we walk through a typical pipeline for the analysis of such data using the Bioconductor `pRoloc` package for the R statistical programming environment.
+
 # Reading and handling mass-spectrometry based proteomics data
@@
 ## The use-case: prediction sub-cellular localisation in pluripotent embroyonic mouse stem cells
+
+As a use-case we analyse a recent high-throughput spatial proteomics dataset from pluripotent mouse embryonic stem cells (E14TG2a) [@hyper].
+The data was generated using hyperplexed LOPIT (hyperLOPIT), a novel technique for robust classification of protein localisation across the whole cell.
+The method uses an elaborate sub-cellular fractionation scheme, enabled by the use of Tandem Mass Tag (TMT) 10-plex and application of the MS data acquisition technique named synchronous precursor selection MS^3 (SPS)-MS^3 [@McAlister:2014], for TMT quantification with high accuracy and precision.
+Three biological replicates were generated from the E14TG2a experiment, the first was to target low density fractions and the second and third were to emphasis seperation of the denser organelles.
+The intersect of replicates 1 and 2 was treated as a 20-plex dataset for the analysis discussed in the manuscipt [@hyper] as it has been shown that combining replicates across from different gradients can increase spatial resolution [@trotter].
+The combination of replicates resulted in 5032 proteins common in both experiments.
+Before combination, the two replicates were separately normalised by sum across the 10 channels (i.e.
+such that the sum of each protein's intensity is 1), for each replicate respectively.
+Normalisation is an essential part of data processing and several methods are available in `MSnbase`.
+The combination of data can also be performed effectively in `MSnbase` as detailed in the dedicated 'Combining MSnSet instances' section of the [`MSnbase` tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
 

### 2016-05-11 bioc-workflow.Rmd
Infrastructure for analyses

@@
 
-## The infrastructre: pRoloc and MSnbase in Bioconductor
-
 ## The use-case: prediction sub-cellular localisation in pluripotent embroyonic mouse stem cells
@@
 
+## The infrastructre: pRoloc and MSnbase in Bioconductor
+
+To make use of the full functionailty of the `pRoloc` software one is required to import their data into R as a `MSnSet` instance.
+The `MSnSet` is a dedicated data structure for the efficient manipulation and processing of mass spectrometry and proteomics data in R.
+Figure 1 illustrates a simplified view of the `MSnSet` structure; there exists 3 key slots (1) the `exprs` slot for storing the quantitation data, (2) the `fData` slot for storing the feature meta-data, and finally (3) the `pData` slot for storing the sample meta-data.
+![**Figure 1.** Simplified representation of the `MSnSet` data structure (reproduced with permission from the `MSnbase` vignette)](./Figures/msnset.png)
+
+There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the [`MSnbase` input/ouput capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
+One suggested simple method is to use the function `readMSnSet2` in `MSnbase`.
+The function takes a single spreadsheet as input and extracts the columns containing the quantitation data, as identified by the argument `ecol`, to create the expression data, while the other columns in the spreadsheet are appeneded to the feature meta-data slot.
+By example, in the code chunk below we read in the `csv` speadsheet containing the quantitation data from the intersect of replicates 1 and 2 of the mouse map [@hyper], using the `readMSnSet2` function.
+The data is as available online with the manuscipt (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `pRolocdata` data package.
+
+To use the `readMSnSet2` function, as a minimum one must specify the file path to the data and which columns of the spreadsheet contain quantitation data.
+The `getEcols` function exists to help users identify which columns of the spreadsheet that contain the quantitation data.
+The spreadsheet of E14TG2a data; "hyperLOPIT-SIData-ms3-rep12-intersect.csv", contains 5032 proteins common across the 2 biological replicates for the respective 2 x 10-plex reporter tags, along with assoiated feature meta-data such as protein markers, protein description, number of quantified peptides etc.
+The spreadsheet contains two headers, with the second header containing information about where the quantitation data is stored.
+We can display the names of the second header by calling the `getEcols` function with the argument `n = 2`, to specify that we wish to display the column names of the second header.
+It is now easy for one to identify that the quantitation data is located in columns 8 to 27.
+It is also possible to pass the optional argument `fnames` to indicate which column to use as the labels by which to identify each protein in the sample.
+Here, we use `fnames = 1` to use the Uniprot identifiers contained in the first column of the speadsheet.
+
+```{r makeMSnSet, echo=FALSE, message=FALSE, warning=FALSE}
+library("MSnbase")
+
+f0 <- dir(system.file("extdata", package = "pRolocdata"), full.names = TRUE, 
+          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
+
+getEcols(f0, split = ",", n = 2)
+
+lopit2015 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1)
+
+## Note: arguments can be passed to read.table and friends using ...
+## here we use skip = 1 to tell readMSnSet2 to skip 1 line before
+## beginning to read data.
+
+## Examine the quantitative information for first 5 proteins 
+exprs(lopit2015)[1:5, ]
+```
+
+As briefly mentioned above, the quantitation data is stored in the `exprs` slot of the `MSnSet` and can be accessed by `exprs(lopit2015)`.
+The feature meta-data is stored in the `fData` slot and can be accessed by `fData(lopit2015)`.
+When using `readMSnSet2`, automatically, everything that is not defined as quantitation data by `ecol` or the fetaure names by `fnames` is deposited to the `fData` slot.
+As we wish to demonstrate the complete analysis of this data we remove the results from the prior analysis described in [@hyper] in the code chunk below.
+We see the `fData` contains 25 columns describing information such as the number of peptides, associated markers, machine learning results etc.
+For demonstration in the code chunk below keep the two columns "phenoDisco.Input" and "SVM.marker.set" that contain two protein marker sets to use an input for machine learning analyses.
+
+```{r removefData}
+fvarLabels(lopit2015)
+fData(lopit2015) <- fData(lopit2015)[, c("phenoDisco.Input", "SVM.marker.set")]
+head(fData(lopit2015))
+```
+
 # Normalisation and Quality Control

### 2016-05-11 bioc-workflow.Rmd
Added .gitignore

@@
 
+Data quality is routinely examined through visualisation to verify that sub-cellular niches have been separated along the gradient.
+Based on De Duve's principle proteins that co-localise exhibit similar quantitation profiles across the fractions employed.
+One approach that has been widely used is principal components analysis (PCA) to inspect if structure, that is representative organellar separation, is present in the data.
+
+This is the first
+
+and by doing so, generated organelle-specific protein distributions along the gradient fractions.
+
+The most natural visualisation is shown on figure 2, obtained using the sub-setting functionality of MSnSet instances and the plotDist function, as illustrated below.
+
 # Markers

### 2016-05-12 bioc-workflow.Rmd
QC section

--- 2016-05-11 bioc-workflow.Rmd
@@
 
-```{r makeMSnSet, echo=FALSE, message=FALSE, warning=FALSE}
+```{r makeMSnSet, message=FALSE, warning=FALSE}
 library("MSnbase")
@@
 
+Data quality is routinely examined through visualisation to verify that sub-cellular niches have been separated along the gradient.
+Based on De Duve's principle [@DeDuve:1981] proteins that co-localise exhibit similar quantitation profiles across the gradient fractions employed.
+One approach that has been widely used to visualise and inspect highthrouput mass spectrometry-based proteomics data is principal components analysis (PCA).
+PCA is one of many dimensionality reduction methods, that allow one to effectively summarise multi-dimensional data in to 2 or 3 dimensions to enable visualisation.
+Very generally, the original continuous multi-dimensional data is transformed into a set of orthogonal components ordered according to the amount of variability that they describe.
+The `plot2D` method in `pRoloc` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the `x`- and `y`-axis, respectively (the `dims` argument can be used to plot other PCs).
+If distinct clusters are observed, we assume that there is organellar separation present in the data.
+In the code chunk below we produce a PCA plot of the mouse stem cell dataset.
+One point on the plot represents one protein.
+We can indeed see many distinct protein clusters.
+Although, plotting the PCs does not give us a hard quantitative measure of separation, it is extremely useful summarising complex experimental information in one figure, to get an overview of the data.
+
+```{r qcplot, message=FALSE, warning=FALSE}
+library('pRoloc')
+plot2D(lopit2015, fcol = NULL)
+```
+
 # Markers

### 2016-05-12 bioc-workflow.Rmd
Added paragraph on markers

@@
 
-# Normalisation and Quality Control
+# Quality Control
 
@@
 Very generally, the original continuous multi-dimensional data is transformed into a set of orthogonal components ordered according to the amount of variability that they describe.
-The `plot2D` method in `pRoloc` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the `x`- and `y`-axis, respectively (the `dims` argument can be used to plot other PCs).
+The `plot2D` method in `pRoloc` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the x- and y-axis, respectively (the `dims` argument can be used to plot other PCs).
 If distinct clusters are observed, we assume that there is organellar separation present in the data.
@@
 One point on the plot represents one protein.
-We can indeed see many distinct protein clusters.
-Although, plotting the PCs does not give us a hard quantitative measure of separation, it is extremely useful summarising complex experimental information in one figure, to get an overview of the data.
+We can indeed see several distinct protein clusters.
+Although, plotting the PCs does not give us a hard quantitative measure of separation, it is extremely useful summarising complex experimental information in one figure, to get an simplified overview of the data.
 
@@
 
+In the context of spatial proteomics, a marker protein is defined as a well-known resident of a specific sub-cellular niche in the species **and** condition of interest.
+Applying this to machine learning (ML), and specifically supervised learning, for the task of protein localisation prediction, markers constitute the labelled training data to use as input for a classification analyses.
+Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
+`pRoloc` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object.
+These can be accessed using the `pRolocmarkers()` function.
+The Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are an essential starting point for protein annotation and marker definition.
+It is important to note that automatic retrieval of sub-cellular localisation information is only the beginning, and expert curation is esssential to ensure annotation is in contect for the the biological question under investigation.
+
 # Novelty Detection

### 2016-05-12 bioc-workflow.Rmd
Update markers section

@@
 
-lopit2015 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1)
+lopit2015 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1, 
+                         stringsAsFactors = FALSE)
 
@@
 We see the `fData` contains 25 columns describing information such as the number of peptides, associated markers, machine learning results etc.
-For demonstration in the code chunk below keep the two columns "phenoDisco.Input" and "SVM.marker.set" that contain two protein marker sets to use an input for machine learning analyses.
+For demonstration in the code chunk below keep the 2nd, 8th and 11th columns which contain the Uniprot entry names and two different marker sets to use an input for machine learning analyses (see sections on markers and subsequent sections).
 
@@
 fvarLabels(lopit2015)
-fData(lopit2015) <- fData(lopit2015)[, c("phenoDisco.Input", "SVM.marker.set")]
+fData(lopit2015) <- fData(lopit2015)[, c(2, 8, 11)]
 head(fData(lopit2015))
@@
 
-In the context of spatial proteomics, a marker protein is defined as a well-known resident of a specific sub-cellular niche in the species **and** condition of interest.
+In the context of spatial proteomics, a marker protein is defined as a well-known resident of a specific sub-cellular niche in a species *and* condition of interest.
 Applying this to machine learning (ML), and specifically supervised learning, for the task of protein localisation prediction, markers constitute the labelled training data to use as input for a classification analyses.
 Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
-`pRoloc` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object.
-These can be accessed using the `pRolocmarkers()` function.
-The Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are an essential starting point for protein annotation and marker definition.
-It is important to note that automatic retrieval of sub-cellular localisation information is only the beginning, and expert curation is esssential to ensure annotation is in contect for the the biological question under investigation.
+`pRoloc` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object, as demonstrated in the code chunk below.
+These marker sets can be accessed using the `pRolocmarkers()` function.
+Marker sets are stored as a simple named vector in R, and originate from inhouse user-defined spreadsheets or a set of markers from previous published studies.
+The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
+The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2015))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
+It is then possible to match names between the markers and the `MSnSet` instance.
+
+```{r addmrkers}
+## List available marker sets
+pRolocmarkers()
+
+## Use mouse markers
+mrk <- pRolocmarkers(species = "mmus")
+head(mrk)
+
+## Change featureNames of the MSnSet to match the marker names 
+## which are named by Uniprot Entry Name
+featureNames(lopit2015) <- make.unique(fData(lopit2015)[, 1])
+
+## Add mouse markers
+lopit2015 <- addMarkers(lopit2015, mrk)
+
+## Remove marker sets with < 20 proteins
+lopit2015 <- minMarkers(lopit2015, n = 20)
+getMarkers(lopit2015, fcol = "markers20")
+plot2D(lopit2015, fcol = "markers20", main = "pRolocmarkers for mouse")
+
+## After expert curation
+plot2D(lopit2015, fcol = "SVM.marker.set", main = "Curated markers")
+```
+
+In general, the Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are a good starting point for protein annotation and marker definition.
+It is important to note however that automatic retrieval of sub-cellular localisation information is only the beginning, and expert curation is esssential to ensure that any annotation added is for the correct context for the the biological question under investigation.
 

### 2016-05-13 bioc-workflow.Rmd
More on markers

@@
 It is then possible to match names between the markers and the `MSnSet` instance.
+We see below that the markers cover many sub-cellular niches, with many niches only containing a few proteins.
+Depending on the biological question and downstream analyses we may wish the subset these marker classes, this can be done using the `minMarkers` function.
+In the code chunk below, we demonstrate how to add markers using `pRolocmarkers` function and then visualise these annotations using the `plot2D` function.
 
@@
 In general, the Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are a good starting point for protein annotation and marker definition.
-It is important to note however that automatic retrieval of sub-cellular localisation information is only the beginning, and expert curation is esssential to ensure that any annotation added is for the correct context for the the biological question under investigation.
+It is important to note however that automatic retrieval of sub-cellular localisation information, from `pRoloc` or elsewhere, is only the beginning in defining a marker set for downstream analyses.
+Expert curation is vital to check that any annotation added is in the correct context for the the biological question under investigation.
+In the code chunk above we show the PCA plot output of the mouse dataset with (i) the annotation for mouse pulled from `pRolocmarkers`, and (ii) annotation after expert curation (stored in the `featureData` column called `SVM.marker.set` that was used for a classification analyses in the original data analyses [@hyper]).
 

### 2016-05-13 bioc-workflow.Rmd
Added novelty detection section

@@
 
+The extraction of sub-cellular protein clusters can be difficult owing to the limited number of marker proteins that exist in databases and elsewhere.
+Furthermore, given the vast complexity of the cell, automatic annotation retrieval does not always give a full representation of the true sub-cellular diversity in the data.
+For downstream analyses, such as supervised machine learning, it is desirable to obtain reliable markers that cover as many sub-cellular niches as possible, as these markers are directly used in the training phase of the ML classification.
+We find that a lack of sub-cellular diveristy in the labelled training data leads to prediction errors, as unlabelled instance can only to predicted to a class that exists in the training data  [@Breckels:2013].
+In such scenarios novelty detection can be useful to identify data-specific sub-cellular groupings such as organelles and protein complexes.
+The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and available in `pRoloc`.
+It is an iterative semi-supervised learning method that combines the classification of proteins on existing labelled data with the detection of new clusters.
+Novelty detection methods are also useful for confirming the presence of known clusters in an unbiased fashion.
+For example, in [@hyper] the `phenoDisco` algorithm was used to detect the data-specific confirmation and presence of the nucleus and sub-compartments, of it was expected that nucleus associated clusters existed within the data as the protocol included a separate chromatin-rich fraction to enrich for nuclei.
+In the code chunk below, we show how to run a a series of novelty detection experiments using the `phenoDisco` algorithm on the mouse stem-cell dataset.
+There are several optional arguments that can be set dependent on the type of biological question of interest.
+
+- Explain how to use different group sizes for detecting small complexes etc.
+- ndims for using >=2 components
+- times -> important for convergence
+- input markers are different here -> explain choice of pd.markers
+
+```{r pd, eval = FALSE}
+## As per the hyperLOPIT paper
+pdRes <- phenoDisco(lopit2015, fcol = "phenoDisco.Input", times = 200, 
+                    GS = 20, p = 0.05)
+
+```
+
 # Supervised machine learning
+
+## Traditional approaches
+
+## Transfer learning

### 2016-05-18 bioc-workflow.Rmd
fixed some typos

@@
 
-## The use-case: prediction sub-cellular localisation in pluripotent embroyonic mouse stem cells
+## The use-case: prediction sub-cellular localisation in pluripotent embryonic mouse stem cells
 
@@
 The method uses an elaborate sub-cellular fractionation scheme, enabled by the use of Tandem Mass Tag (TMT) 10-plex and application of the MS data acquisition technique named synchronous precursor selection MS^3 (SPS)-MS^3 [@McAlister:2014], for TMT quantification with high accuracy and precision.
-Three biological replicates were generated from the E14TG2a experiment, the first was to target low density fractions and the second and third were to emphasis seperation of the denser organelles.
-The intersect of replicates 1 and 2 was treated as a 20-plex dataset for the analysis discussed in the manuscipt [@hyper] as it has been shown that combining replicates across from different gradients can increase spatial resolution [@trotter].
+Three biological replicates were generated from the E14TG2a experiment, the first was to target low density fractions and the second and third were to emphasis separation of the denser organelles.
+The intersect of replicates 1 and 2 was treated as a 20-plex dataset for the analysis discussed in the manuscript [@hyper] as it has been shown that combining replicates across from different gradients can increase spatial resolution [@trotter].
 The combination of replicates resulted in 5032 proteins common in both experiments.
@@
 
-## The infrastructre: pRoloc and MSnbase in Bioconductor
+## The infrastructure: pRoloc and MSnbase in Bioconductor
 
-To make use of the full functionailty of the `pRoloc` software one is required to import their data into R as a `MSnSet` instance.
+To make use of the full functionality of the `pRoloc` software one is required to import their data into R as a `MSnSet` instance.
 The `MSnSet` is a dedicated data structure for the efficient manipulation and processing of mass spectrometry and proteomics data in R.
@@
 
-There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the [`MSnbase` input/ouput capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
+There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the [`MSnbase` input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
 One suggested simple method is to use the function `readMSnSet2` in `MSnbase`.
-The function takes a single spreadsheet as input and extracts the columns containing the quantitation data, as identified by the argument `ecol`, to create the expression data, while the other columns in the spreadsheet are appeneded to the feature meta-data slot.
-By example, in the code chunk below we read in the `csv` speadsheet containing the quantitation data from the intersect of replicates 1 and 2 of the mouse map [@hyper], using the `readMSnSet2` function.
-The data is as available online with the manuscipt (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `pRolocdata` data package.
+The function takes a single spreadsheet as input and extracts the columns containing the quantitation data, as identified by the argument `ecol`, to create the expression data, while the other columns in the spreadsheet are appended to the feature meta-data slot.
+By example, in the code chunk below we read in the `csv` spreadsheet containing the quantitation data from the intersect of replicates 1 and 2 of the mouse map [@hyper], using the `readMSnSet2` function.
+The data is as available online with the manuscript (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `pRolocdata` data package.
 
@@
 The `getEcols` function exists to help users identify which columns of the spreadsheet that contain the quantitation data.
-The spreadsheet of E14TG2a data; "hyperLOPIT-SIData-ms3-rep12-intersect.csv", contains 5032 proteins common across the 2 biological replicates for the respective 2 x 10-plex reporter tags, along with assoiated feature meta-data such as protein markers, protein description, number of quantified peptides etc.
+The spreadsheet of E14TG2a data; "hyperLOPIT-SIData-ms3-rep12-intersect.csv", contains 5032 proteins common across the 2 biological replicates for the respective 2 x 10-plex reporter tags, along with associated feature meta-data such as protein markers, protein description, number of quantified peptides etc.
 The spreadsheet contains two headers, with the second header containing information about where the quantitation data is stored.
@@
 It is also possible to pass the optional argument `fnames` to indicate which column to use as the labels by which to identify each protein in the sample.
-Here, we use `fnames = 1` to use the Uniprot identifiers contained in the first column of the speadsheet.
+Here, we use `fnames = 1` to use the Uniprot identifiers contained in the first column of the spreadsheet.
 
@@
 The feature meta-data is stored in the `fData` slot and can be accessed by `fData(lopit2015)`.
-When using `readMSnSet2`, automatically, everything that is not defined as quantitation data by `ecol` or the fetaure names by `fnames` is deposited to the `fData` slot.
+When using `readMSnSet2`, automatically, everything that is not defined as quantitation data by `ecol` or the feature names by `fnames` is deposited to the `fData` slot.
 As we wish to demonstrate the complete analysis of this data we remove the results from the prior analysis described in [@hyper] in the code chunk below.
@@
 Based on De Duve's principle [@DeDuve:1981] proteins that co-localise exhibit similar quantitation profiles across the gradient fractions employed.
-One approach that has been widely used to visualise and inspect highthrouput mass spectrometry-based proteomics data is principal components analysis (PCA).
+One approach that has been widely used to visualise and inspect high throughput mass spectrometry-based proteomics data is principal components analysis (PCA).
 PCA is one of many dimensionality reduction methods, that allow one to effectively summarise multi-dimensional data in to 2 or 3 dimensions to enable visualisation.
@@
 These marker sets can be accessed using the `pRolocmarkers()` function.
-Marker sets are stored as a simple named vector in R, and originate from inhouse user-defined spreadsheets or a set of markers from previous published studies.
+Marker sets are stored as a simple named vector in R, and originate from in-house user-defined spreadsheets or a set of markers from previous published studies.
 The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
@@
 For downstream analyses, such as supervised machine learning, it is desirable to obtain reliable markers that cover as many sub-cellular niches as possible, as these markers are directly used in the training phase of the ML classification.
-We find that a lack of sub-cellular diveristy in the labelled training data leads to prediction errors, as unlabelled instance can only to predicted to a class that exists in the training data  [@Breckels:2013].
+We find that a lack of sub-cellular diversity in the labelled training data leads to prediction errors, as unlabelled instance can only to predicted to a class that exists in the training data  [@Breckels:2013].
 In such scenarios novelty detection can be useful to identify data-specific sub-cellular groupings such as organelles and protein complexes.

### 2016-05-31 bioc-workflow.Rmd
Started section on supervised ML

@@
 
-lopit2015 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1, 
+lopit2016 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1, 
                          stringsAsFactors = FALSE)
@@
 ## Examine the quantitative information for first 5 proteins 
-exprs(lopit2015)[1:5, ]
+exprs(lopit2016)[1:5, ]
 ```
 
-As briefly mentioned above, the quantitation data is stored in the `exprs` slot of the `MSnSet` and can be accessed by `exprs(lopit2015)`.
-The feature meta-data is stored in the `fData` slot and can be accessed by `fData(lopit2015)`.
+As briefly mentioned above, the quantitation data is stored in the `exprs` slot of the `MSnSet` and can be accessed by `exprs(lopit2016)`.
+The feature meta-data is stored in the `fData` slot and can be accessed by `fData(lopit2016)`.
 When using `readMSnSet2`, automatically, everything that is not defined as quantitation data by `ecol` or the feature names by `fnames` is deposited to the `fData` slot.
@@
 ```{r removefData}
-fvarLabels(lopit2015)
-fData(lopit2015) <- fData(lopit2015)[, c(2, 8, 11)]
-head(fData(lopit2015))
+fvarLabels(lopit2016)
+fData(lopit2016) <- fData(lopit2016)[, c(2, 8, 11)]
+head(fData(lopit2016))
 ```
@@
 library('pRoloc')
-plot2D(lopit2015, fcol = NULL)
+plot2D(lopit2016, fcol = NULL)
 ```
@@
 The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
-The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2015))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
+The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2016))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
 It is then possible to match names between the markers and the `MSnSet` instance.
@@
 ## which are named by Uniprot Entry Name
-featureNames(lopit2015) <- make.unique(fData(lopit2015)[, 1])
+featureNames(lopit2016) <- make.unique(fData(lopit2016)[, 1])
 
 ## Add mouse markers
-lopit2015 <- addMarkers(lopit2015, mrk)
+lopit2016 <- addMarkers(lopit2016, mrk)
 
 ## Remove marker sets with < 20 proteins
-lopit2015 <- minMarkers(lopit2015, n = 20)
-getMarkers(lopit2015, fcol = "markers20")
-plot2D(lopit2015, fcol = "markers20", main = "pRolocmarkers for mouse")
+lopit2016 <- minMarkers(lopit2016, n = 20)
+getMarkers(lopit2016, fcol = "markers20")
+plot2D(lopit2016, fcol = "markers20", main = "pRolocmarkers for mouse")
 
 ## After expert curation
-plot2D(lopit2015, fcol = "SVM.marker.set", main = "Curated markers")
+plot2D(lopit2016, fcol = "SVM.marker.set", main = "Curated markers")
 ```
@@
 For downstream analyses, such as supervised machine learning, it is desirable to obtain reliable markers that cover as many sub-cellular niches as possible, as these markers are directly used in the training phase of the ML classification.
-We find that a lack of sub-cellular diversity in the labelled training data leads to prediction errors, as unlabelled instance can only to predicted to a class that exists in the training data  [@Breckels:2013].
+We find that a lack of sub-cellular diversity in the labelled training data leads to prediction errors, as unlabelled instances can only be assigned to a class that exists in the training data [@Breckels:2013].
 In such scenarios novelty detection can be useful to identify data-specific sub-cellular groupings such as organelles and protein complexes.
-The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and available in `pRoloc`.
+The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and is available in `pRoloc`.
 It is an iterative semi-supervised learning method that combines the classification of proteins on existing labelled data with the detection of new clusters.
 Novelty detection methods are also useful for confirming the presence of known clusters in an unbiased fashion.
-For example, in [@hyper] the `phenoDisco` algorithm was used to detect the data-specific confirmation and presence of the nucleus and sub-compartments, of it was expected that nucleus associated clusters existed within the data as the protocol included a separate chromatin-rich fraction to enrich for nuclei.
-In the code chunk below, we show how to run a a series of novelty detection experiments using the `phenoDisco` algorithm on the mouse stem-cell dataset.
+For example, in [@hyper] the `phenoDisco` algorithm was used to detect the data-specific confirmation and presence of the nucleus and nucleus sub-compartments, as it was expected that nucleus associated clusters existed within the data as the protocol included a separate chromatin-rich fraction to enrich for nuclei.
+In the code chunk below, we show how to run a series of novelty detection experiments using the `phenoDisco` algorithm on the mouse stem-cell dataset.
 There are several optional arguments that can be set dependent on the type of biological question of interest.
+For example,
 
@@
 ## As per the hyperLOPIT paper
-pdRes <- phenoDisco(lopit2015, fcol = "phenoDisco.Input", times = 200, 
+pdRes <- phenoDisco(lopit2016, fcol = "phenoDisco.Input", times = 200, 
                     GS = 20, p = 0.05)
@@
 # Supervised machine learning
+Supervised machine learning, also known as classification, is an essential tool for the assignment of proteins to distinct sub-cellular niches.
+Using a set of labelled training examples i.e.
+markers, we can train a machine learning classifier to learn a mapping between the data i.e.
+the quantitative protein profiles, and the known localisation.
+The trained classifier can then be used to predict the localisation of a protein of unknown localisation, based on its observed protein profile.
+To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @trotter; @Hall:2009; @Dunkley:2006; @Tan:2009]
 

### 2016-05-31 bioc-workflow.Rmd
Add session info section

@@
 ## Transfer learning
+
+# Session information
+
+The function `sessionInfo` provides a summary of all packages and
+versions used to generate this document. This enables us to record the
+exact state of our session that lead to these exact
+results. Conversely, if the script stops working of if it returns
+different results, we are in a position to re-generate the original
+results using the adequate software versions and retrace changes in
+the software that lead to failure and/or different results.
+
+```{r si}
+sessionInfo()
+```
+
+It is always important to include session information details along
+with a
+[short reproducible example](http://adv-r.had.co.nz/Reproducibility.html)
+highlighting the problem or
+[question](https://support.bioconductor.org/) at hand.

### 2016-05-31 bioc-workflow.Rmd
code for classification

--- 2016-05-31 bioc-workflow.Rmd
@@
 # Supervised machine learning
+
 Supervised machine learning, also known as classification, is an essential tool for the assignment of proteins to distinct sub-cellular niches.
@@
 markers, we can train a machine learning classifier to learn a mapping between the data i.e.
-the quantitative protein profiles, and the known localisation.
+the quantitative protein profiles, and a known localisation.
 The trained classifier can then be used to predict the localisation of a protein of unknown localisation, based on its observed protein profile.
-To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @trotter; @Hall:2009; @Dunkley:2006; @Tan:2009]
+To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @trotter; @Hall:2009; @Dunkley:2006; @Tan:2009].
+In the below code chunk we demonstrate how to run a typical classification analysis.
+
+```{r loadpdRes, include=FALSE}
+pdRes <- lopit2016
+load("Data/fusionparams.rda")
+```
+```{r optimise, eval=FALSE}
+w <- table(fData(pdRes)[, "SVM.marker.set"])
+w <- 1/w[names(w) != "unknown"]
+
+params <- svmOptimisation(pdRes, fcol = "SVM.marker.set",
+                          times = 100, xval = 5,
+                          class.weights = w,
+                          verbose = TRUE)
+```
+```{r visualiseOpt}
+plot(params)
+levelPlot(params)
+(best <- getParams(params))
+```
+```{r classify, eval=TRUE}
+svmRes <- svmClassification(pdRes, params,
+                            class.weights = w,
+                            fcol = "SVM.marker.set")
+```
 

### 2016-05-31 bioc-workflow.Rmd

@@
 ## Transfer learning
+
+=======
+The function `sessionInfo` provides a summary of all packages and
+versions used to generate this document. This enables us to record the
+exact state of our session that lead to these exact
+results. Conversely, if the script stops working of if it returns
+different results, we are in a position to re-generate the original
+results using the adequate software versions and retrace changes in
+the software that lead to failure and/or different results.
+
+```{r si}
+sessionInfo()
+```
+
+It is always important to include session information details along
+with a
+[short reproducible example](http://adv-r.had.co.nz/Reproducibility.html)
+highlighting the problem or
+[question](https://support.bioconductor.org/) at hand.

### 2016-05-31 bioc-workflow.Rmd
update intro

--- 2016-05-31 bioc-workflow.Rmd
@@
 author: "Lisa Breckels"
-date: "11 May 2016"
 output: html_document
@@
 
-# Forward
+```{r env, echo=FALSE}
+library("BiocStyle")
+suppressPackageStartupMessages(library("MSnbase"))
+suppressPackageStartupMessages(library("pRoloc"))
+```
 
-Quantitative mass spectrometry based spatial proteomics involves elaborate, expensive and time consuming experimental procedures and considerable effort is invested in the generation of such data.
-Multiple research groups have described a variety of approaches to establish high quality proteome-wide datasets.
-However, data analysis is as critical as data production for reliable and insightful biological interpretation.
-Here, we walk through a typical pipeline for the analysis of such data using the Bioconductor `pRoloc` package for the R statistical programming environment.
+# Introduction
+
+Quantitative mass spectrometry based spatial proteomics involves
+elaborate, expensive and time consuming experimental procedures and
+considerable effort is invested in the generation of such
+data. Multiple research groups have described a variety of approaches
+to establish high quality proteome-wide datasets. However, data
+analysis is as critical as data production for reliable and insightful
+biological interpretation. Here, we walk the reader through a typical
+pipeline for the analysis of such data using several Bioconductor
+packages for the R statistical programming environment.
+
+The main package to analyse protein localisation data is
+`r Biocpkg("pRoloc")`, which offers a set of dedicated functions for the
+analysis of such data. `pRoloc` itself relies on `r Biocpkg("MSnbase")` to
+manipulate and process quantitative proteomics data. Many other
+packages are used by `pRoloc` for clustering, classification and
+visualisation.
+
+In this workflow, we will describe how to prepare the spatial
+proteomics data starting from a spreadsheet containing quantitative
+mass spectrometry data. We will focus on a recent pluripotent mouse
+embryonic stem cells experiment [@hyper]. Additional annotated and
+pre-formatted datasets from various species are readily available in
+the `r Biocexptpkg("pRolocdata")` package.
+
+Installation of Bioconductor package is documented in details on the
+[Bioconductor installation help page](http://bioconductor.org/install/#install-bioconductor-packages). This
+procedure is also applicable to any packages, from
+[CRAN](https://cran.r-project.org/) as well as GitHub. Once a package
+has been installed, it needs to be loaded for it functionality become
+available in the R session; this is done with the `library` function.
+
+If you have questions about this workflow in particular, or about
+other Bioconductor packages in general, they a best asked on the
+[Bioconductor support site](https://support.bioconductor.org/)
+following the
+[posting guidelines](http://www.bioconductor.org/help/support/posting-guide/). Questions
+can be tagged with specific package names or keywords. For more
+general information about mass spectrometry and proteomics, the
+readers are invited to read the `r Biocexptpkg("RforProteomics")`
+package vignettes and associated papers [@Gatto:2014;@Gatto2015].
 
@@
 
+# Interactive visualisation
+
 # Novelty Detection

### 2016-05-31 bioc-workflow.Rmd

@@
 
-# Session information
-
+=======
 The function `sessionInfo` provides a summary of all packages and

### 2016-05-31 bioc-workflow.Rmd
add lost si section title
Small updates to supervised ML

@@
 To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @trotter; @Hall:2009; @Dunkley:2006; @Tan:2009].
-In the below code chunk we demonstrate how to run a typical classification analysis.
+
+There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` Machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
+We find the general tendancy to be that it is not the choice of classifier but the proper optimisation of the algorithms parameters that limit the classification results.
+Before employing a classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
+For example, in the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
+In the example below, the training data is found in the `featureData` slot in the column called `SVM.marker.set`.
+
+Notes:
+- take care of properly setting the model parameters. 
+- Wrongly set parameters can have adverse effects 
+- how well they represent the multivariate data?
+- Parameter optimisation can be conducted in a number of ways. One of the most common is train/test plus 5-fold x-val
 
@@
 
-## Traditional approaches
-
 ## Transfer learning
 
-=======
+# Session information
+
 The function `sessionInfo` provides a summary of all packages and

### 2016-06-01 bioc-workflow.Rmd
small updates to use case section

--- 2016-05-31 bioc-workflow.Rmd
@@
 `r Biocpkg("pRoloc")`, which offers a set of dedicated functions for the
-analysis of such data. `pRoloc` itself relies on `r Biocpkg("MSnbase")` to
+analysis of such data. `r Biocpkg("pRoloc")` itself relies on `r Biocpkg("MSnbase")` to
 manipulate and process quantitative proteomics data. Many other
-packages are used by `pRoloc` for clustering, classification and
+packages are used by `r Biocpkg("pRoloc")` for clustering, classification and
 visualisation.
@@
 
-As a use-case we analyse a recent high-throughput spatial proteomics dataset from pluripotent mouse embryonic stem cells (E14TG2a) [@hyper].
-The data was generated using hyperplexed LOPIT (hyperLOPIT), a novel technique for robust classification of protein localisation across the whole cell.
-The method uses an elaborate sub-cellular fractionation scheme, enabled by the use of Tandem Mass Tag (TMT) 10-plex and application of the MS data acquisition technique named synchronous precursor selection MS^3 (SPS)-MS^3 [@McAlister:2014], for TMT quantification with high accuracy and precision.
-Three biological replicates were generated from the E14TG2a experiment, the first was to target low density fractions and the second and third were to emphasis separation of the denser organelles.
-The intersect of replicates 1 and 2 was treated as a 20-plex dataset for the analysis discussed in the manuscript [@hyper] as it has been shown that combining replicates across from different gradients can increase spatial resolution [@trotter].
-The combination of replicates resulted in 5032 proteins common in both experiments.
-Before combination, the two replicates were separately normalised by sum across the 10 channels (i.e.
-such that the sum of each protein's intensity is 1), for each replicate respectively.
-Normalisation is an essential part of data processing and several methods are available in `MSnbase`.
-The combination of data can also be performed effectively in `MSnbase` as detailed in the dedicated 'Combining MSnSet instances' section of the [`MSnbase` tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
+As a use-case we analyse a recent high-throughput spatial proteomics
+dataset from pluripotent mouse embryonic stem cells (E14TG2a)
+[@hyper]. The data was generated using hyperplexed LOPIT (hyperLOPIT),
+an improved method relying on better sub-cellular fractionation and
+more acurate quantitation, leading to more reliable classification of
+protein localisation across the whole sub-cellular space. The method
+uses an elaborate sub-cellular fractionation scheme, enabled by the
+use of Tandem Mass Tag (TMT) [Thompson:2003] 10-plex and application
+of the MS data acquisition technique named synchronous precursor
+selection MS$^3$ (SPS)-MS$^3$ [@McAlister:2014], for TMT
+quantification with high accuracy and precision. Three biological
+replicates were generated from the E14TG2a experiment, the first was
+to target low density fractions and the second and third were to
+emphasis separation of the denser organelles. The intersect of
+replicates 1 and 2 was treated as a 20-plex dataset for the analysis
+discussed in the manuscript [@hyper] as it has been shown that
+combining replicates across from different gradients can increase
+spatial resolution [@trotter]. The combination of replicates resulted
+in 5032 proteins common in both experiments. Before combination, the
+two replicates were separately normalised by sum across the 10
+channels (i.e. such that the sum of each protein's intensity is 1),
+for each replicate respectively. Normalisation is an essential part of
+data processing and several methods are available in 
+`r Biocpkg("MSnbase")`. The normalisation desired in this specific case
+would be obtained with the a call to the `normalise` method.
 
-## The infrastructure: pRoloc and MSnbase in Bioconductor
+The combination of data can also be performed effectively in 
+`r Biocpkg("MSnbase")` as detailed in the dedicated *Combining MSnSet
+instances* section of the `r Biocpkg("MSnbase")`
+[tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
 
-To make use of the full functionality of the `pRoloc` software one is required to import their data into R as a `MSnSet` instance.
+## The infrastructure: `r Biocpkg("pRoloc")` and `r Biocpkg("MSnbase")` in Bioconductor
+
+To make use of the full functionality of the `r Biocpkg("pRoloc")` software one is required to import their data into R as a `MSnSet` instance.
 The `MSnSet` is a dedicated data structure for the efficient manipulation and processing of mass spectrometry and proteomics data in R.
 Figure 1 illustrates a simplified view of the `MSnSet` structure; there exists 3 key slots (1) the `exprs` slot for storing the quantitation data, (2) the `fData` slot for storing the feature meta-data, and finally (3) the `pData` slot for storing the sample meta-data.
-![**Figure 1.** Simplified representation of the `MSnSet` data structure (reproduced with permission from the `MSnbase` vignette)](./Figures/msnset.png)
+![**Figure 1.** Simplified representation of the `MSnSet` data structure (reproduced with permission from the `r Biocpkg("MSnbase")` vignette)](./Figures/msnset.png)
 
-There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the [`MSnbase` input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
-One suggested simple method is to use the function `readMSnSet2` in `MSnbase`.
+There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the `r Biocpkg("MSnbase")` [input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
+One suggested simple method is to use the function `readMSnSet2` in `r Biocpkg("MSnbase")`.
 The function takes a single spreadsheet as input and extracts the columns containing the quantitation data, as identified by the argument `ecol`, to create the expression data, while the other columns in the spreadsheet are appended to the feature meta-data slot.
 By example, in the code chunk below we read in the `csv` spreadsheet containing the quantitation data from the intersect of replicates 1 and 2 of the mouse map [@hyper], using the `readMSnSet2` function.
-The data is as available online with the manuscript (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `pRolocdata` data package.
+The data is as available online with the manuscript (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `r Biocexptpkg("pRolocdata")` data package.
 
@@
 Very generally, the original continuous multi-dimensional data is transformed into a set of orthogonal components ordered according to the amount of variability that they describe.
-The `plot2D` method in `pRoloc` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the x- and y-axis, respectively (the `dims` argument can be used to plot other PCs).
+The `plot2D` method in `r Biocpkg("pRoloc")` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the x- and y-axis, respectively (the `dims` argument can be used to plot other PCs).
 If distinct clusters are observed, we assume that there is organellar separation present in the data.
@@
 Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
-`pRoloc` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object, as demonstrated in the code chunk below.
+`r Biocpkg("pRoloc")` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object, as demonstrated in the code chunk below.
 These marker sets can be accessed using the `pRolocmarkers()` function.
@@
 In general, the Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are a good starting point for protein annotation and marker definition.
-It is important to note however that automatic retrieval of sub-cellular localisation information, from `pRoloc` or elsewhere, is only the beginning in defining a marker set for downstream analyses.
+It is important to note however that automatic retrieval of sub-cellular localisation information, from `r Biocpkg("pRoloc")` or elsewhere, is only the beginning in defining a marker set for downstream analyses.
 Expert curation is vital to check that any annotation added is in the correct context for the the biological question under investigation.
@@
 In such scenarios novelty detection can be useful to identify data-specific sub-cellular groupings such as organelles and protein complexes.
-The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and is available in `pRoloc`.
+The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and is available in `r Biocpkg("pRoloc")`.
 It is an iterative semi-supervised learning method that combines the classification of proteins on existing labelled data with the detection of new clusters.
@@
 
-=======
+# Session information
+
 The function `sessionInfo` provides a summary of all packages and

### 2016-06-01 bioc-workflow.Rmd

@@
 
-## Traditional approaches
-
 ## Transfer learning

### 2016-06-01 bioc-workflow.Rmd
Data processing section and other suggestions
working on reading data into R as MSnSet

@@
 
-As a use-case we analyse a recent high-throughput spatial proteomics
+As a use-case, we analyse a recent high-throughput spatial proteomics
 dataset from pluripotent mouse embryonic stem cells (E14TG2a)
 [@hyper]. The data was generated using hyperplexed LOPIT (hyperLOPIT),
-an improved method relying on better sub-cellular fractionation and
+an improved method relying on improved sub-cellular fractionation and
 more acurate quantitation, leading to more reliable classification of
@@
 of the MS data acquisition technique named synchronous precursor
-selection MS$^3$ (SPS)-MS$^3$ [@McAlister:2014], for TMT
+selection MS$^3$ (SPS-MS$^3$) [@McAlister:2014], for TMT
 quantification with high accuracy and precision. Three biological
@@
 spatial resolution [@trotter]. The combination of replicates resulted
-in 5032 proteins common in both experiments. Before combination, the
-two replicates were separately normalised by sum across the 10
-channels (i.e. such that the sum of each protein's intensity is 1),
-for each replicate respectively. Normalisation is an essential part of
-data processing and several methods are available in 
-`r Biocpkg("MSnbase")`. The normalisation desired in this specific case
-would be obtained with the a call to the `normalise` method.
+in 5032 proteins common in both experiments.
 
-The combination of data can also be performed effectively in 
-`r Biocpkg("MSnbase")` as detailed in the dedicated *Combining MSnSet
-instances* section of the `r Biocpkg("MSnbase")`
-[tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
+These, as well as many other data are directly available as properly
+structured and annotated computational object from the 
+`r Biocexptpkg("pRolocdata")` experiment package. In this workflow, we
+will start with a description of how to generate these ad hoc objects
+starting from any arbitrary spreadsheets, as produced by many popular
+third-party applications.
+
+While we focus here on a LOPIT-type dataset, these analyses are
+relevant for any quantitative spatial proteomics data, irrespective of
+the fractionation or quantitation (i.e. labeled or label-free)
+methods.
 
@@
 
-To make use of the full functionality of the `r Biocpkg("pRoloc")` software one is required to import their data into R as a `MSnSet` instance.
-The `MSnSet` is a dedicated data structure for the efficient manipulation and processing of mass spectrometry and proteomics data in R.
-Figure 1 illustrates a simplified view of the `MSnSet` structure; there exists 3 key slots (1) the `exprs` slot for storing the quantitation data, (2) the `fData` slot for storing the feature meta-data, and finally (3) the `pData` slot for storing the sample meta-data.
-![**Figure 1.** Simplified representation of the `MSnSet` data structure (reproduced with permission from the `r Biocpkg("MSnbase")` vignette)](./Figures/msnset.png)
+To make use of the full functionality of the `r Biocpkg("pRoloc")`
+software one needs to import their data into R and prepare them as an
+`MSnSet`. The `MSnSet` is a dedicated data structure for the efficient
+manipulation and processing of mass spectrometry and proteomics data
+in R. Figure 1 illustrates a simplified view of the `MSnSet`
+structure; there exists 3 key sub-parts (termed slots) to such a data
+object: (1) the `exprs` slot for storing the quantitation data, (2)
+the `fData` slot for storing the feature meta-data, and finally (3)
+the `pData` slot for storing the sample meta-data.
 
-There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the `r Biocpkg("MSnbase")` [input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
-One suggested simple method is to use the function `readMSnSet2` in `r Biocpkg("MSnbase")`.
-The function takes a single spreadsheet as input and extracts the columns containing the quantitation data, as identified by the argument `ecol`, to create the expression data, while the other columns in the spreadsheet are appended to the feature meta-data slot.
-By example, in the code chunk below we read in the `csv` spreadsheet containing the quantitation data from the intersect of replicates 1 and 2 of the mouse map [@hyper], using the `readMSnSet2` function.
-The data is as available online with the manuscript (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `r Biocexptpkg("pRolocdata")` data package.
+![Simplified representation of the `MSnSet` data structure (reproduced with permission from the `r Biocpkg("MSnbase")` vignette)](./Figures/msnset.png)
 
-To use the `readMSnSet2` function, as a minimum one must specify the file path to the data and which columns of the spreadsheet contain quantitation data.
-The `getEcols` function exists to help users identify which columns of the spreadsheet that contain the quantitation data.
-The spreadsheet of E14TG2a data; "hyperLOPIT-SIData-ms3-rep12-intersect.csv", contains 5032 proteins common across the 2 biological replicates for the respective 2 x 10-plex reporter tags, along with associated feature meta-data such as protein markers, protein description, number of quantified peptides etc.
-The spreadsheet contains two headers, with the second header containing information about where the quantitation data is stored.
-We can display the names of the second header by calling the `getEcols` function with the argument `n = 2`, to specify that we wish to display the column names of the second header.
-It is now easy for one to identify that the quantitation data is located in columns 8 to 27.
-It is also possible to pass the optional argument `fnames` to indicate which column to use as the labels by which to identify each protein in the sample.
-Here, we use `fnames = 1` to use the Uniprot identifiers contained in the first column of the spreadsheet.
+There are a number of ways to import quantitation data and create an
+`MSnSet` instance and all methods are described in the `r
+Biocpkg("MSnbase")`
+[input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf). One
+suggested simple method is to use the function `readMSnSet2` in 
+`r Biocpkg("MSnbase")`. The function takes a single spreadsheet as input
+and extracts the columns containing the quantitation data, as
+identified by the argument `ecol`, to create the expression data,
+while the other columns in the spreadsheet are appended to the feature
+meta-data slot.  By example, in the code chunk below we read in the
+`csv` spreadsheet containing the quantitation data from the intersect
+of replicates 1 and 2 of the mouse map [@hyper], using the
+`readMSnSet2` function. The data is as available online with the
+manuscript (see tab 2 of the `xlsx` supplementary data set 1 in
+[@hyper], which should be exported as a text-based spreadsheet) and
+also as a `csv` in the Bioconductor `r Biocexptpkg("pRolocdata")` data
+package.
 
-```{r makeMSnSet, message=FALSE, warning=FALSE}
+To use the `readMSnSet2` function, as a minimum one must specify the
+file path to the data and which columns of the spreadsheet contain
+quantitation data. The `getEcols` function exists to help users
+identify which columns of the spreadsheet contain the
+quantitation data.
+
+We first locate the the spreadsheet of E14TG2a data distributed with
+the `r Biocexptpkg("pRolocdata")` data using `system.file`, that
+locates the `extdata` in the `r Biocexptpkg("pRolocdata")` package on
+the hard drive and `dir`, that displays the full path to the file
+matching the `hyperLOPIT-SIData-ms3-rep12-intersect.csv` pattern,
+which corresponds the the file of interest. In the last line, we print
+the filename (not the full path, which will vary from computer to
+computer).
+
+```{r getFilename}
 library("MSnbase")
+extdatadir <- system.file("extdata", package = "pRolocdata")
+csvfile <- dir(extdatadir, full.names = TRUE,
+          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
+basename(csvfile)
+```
 
-f0 <- dir(system.file("extdata", package = "pRolocdata"), full.names = TRUE, 
-          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
+Note that the file is compressed (as indicated by the `gz`, for
+`gzip`, extension), and will be decompressed on-the-fly when read into
+R later on.
 
-getEcols(f0, split = ",", n = 2)
+The spreadsheet that was deposited by the authors contains two
+headers, with the second header containing information about where the
+quantitation data is stored. We can display the names of the second
+header by calling the `getEcols` function with the argument `n = 2`
+(the default value is `n = 1`), to specify that we wish to display the
+column names of the second line.
 
-lopit2016 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1, 
+```{r getEcols}
+getEcols(csvfile, split = ",", n = 2)
+```
+
+It is now easy for one to
+identify that the quantitation data is located in columns 8 to 27. It
+is also possible to pass the optional argument `fnames` to indicate
+which column to use as the labels by which to identify each protein in
+the sample. Here, we use `fnames = 1` to use the Uniprot identifiers
+contained in the first column of the spreadsheet.
+
+```{r readMSnSet2}
+lopit2016 <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1, 
                          stringsAsFactors = FALSE)
@@
 ```
+
+contains 5032 proteins common across the 2 biological replicates for
+the respective 2 x 10-plex reporter tags, along with associated
+feature meta-data such as protein markers, protein description, number
+of quantified peptides etc.
 
@@
 ```
+
+## Data processing
+
+* normalisation
+* missing data
+* combining acquisistions
+
+Before combination, the two replicates were separately normalised by
+sum across the 10 channels (i.e. such that the sum of each protein's
+intensity is 1), for each replicate respectively. Normalisation is an
+essential part of data processing and several methods are available in
+`r Biocpkg("MSnbase")`. The normalisation desired in this specific
+case would be obtained with the a call to the `normalise` method.
+
+The combination of data can also be performed effectively in 
+`r Biocpkg("MSnbase")` as detailed in the dedicated *Combining MSnSet
+instances* section of the `r Biocpkg("MSnbase")`
+[tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
 

### 2016-06-01 bioc-workflow.Rmd
more details in readmsnset2

@@
 headers, with the second header containing information about where the
-quantitation data is stored. We can display the names of the second
-header by calling the `getEcols` function with the argument `n = 2`
-(the default value is `n = 1`), to specify that we wish to display the
-column names of the second line.
+quantitation data is stored.
+
+![A screenshot of the data in the spreadsheet.](./Figures/spreadsheet-screenshot.png)
+
+We can display the names of the second header by calling the
+`getEcols` function with the argument `n = 2` (the default value is `n
+= 1`), to specify that we wish to display the column names of the
+second line.
 
@@
 
-It is now easy for one to
-identify that the quantitation data is located in columns 8 to 27. It
-is also possible to pass the optional argument `fnames` to indicate
-which column to use as the labels by which to identify each protein in
-the sample. Here, we use `fnames = 1` to use the Uniprot identifiers
-contained in the first column of the spreadsheet.
+It is now easy for one to identify that the quantitation data,
+corresponding to the 10 TMT isobaric tags, is located in columns 8
+to 27. We now have the two mandatory arguments to `readMSnSet2`,
+namely the filename (stored in the `csvfile` variable) and the
+quantitation column indices. In addition to these, it is also possible
+to pass the optional argument `fnames` to indicate which column to use
+as the labels by which to identify each protein in the sample. Here,
+we use `fnames = 1` to use the Uniprot identifiers contained in the
+first (unnamed) column of the spreadsheet. We also need to specify to
+skip the first line of the file (for the same reason that we used 
+`n = 2` in `getEcols` above) to read the `csv` data and convert it to an
+`MSnSet` object.
 
 ```{r readMSnSet2}
-lopit2016 <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1, 
-                         stringsAsFactors = FALSE)
-
-## Note: arguments can be passed to read.table and friends using ...
-## here we use skip = 1 to tell readMSnSet2 to skip 1 line before
-## beginning to read data.
-
-## Examine the quantitative information for first 5 proteins 
-exprs(lopit2016)[1:5, ]
+lopit2016 <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1)
 ```
 
-contains 5032 proteins common across the 2 biological replicates for
-the respective 2 x 10-plex reporter tags, along with associated
-feature meta-data such as protein markers, protein description, number
-of quantified peptides etc.
+Below, we display a short summary of the data. The data contains 
+`r nrow(lopit2016)` proteins/features common across the 2 biological replicates
+for the respective 2 x 10-plex reporter tags (`r ncol(lopit2016)`
+columns/samples), along with associated feature meta-data such as
+protein markers, protein description, number of quantified peptides
+etc (see below).
 
-As briefly mentioned above, the quantitation data is stored in the `exprs` slot of the `MSnSet` and can be accessed by `exprs(lopit2016)`.
-The feature meta-data is stored in the `fData` slot and can be accessed by `fData(lopit2016)`.
-When using `readMSnSet2`, automatically, everything that is not defined as quantitation data by `ecol` or the feature names by `fnames` is deposited to the `fData` slot.
-As we wish to demonstrate the complete analysis of this data we remove the results from the prior analysis described in [@hyper] in the code chunk below.
-We see the `fData` contains 25 columns describing information such as the number of peptides, associated markers, machine learning results etc.
-For demonstration in the code chunk below keep the 2nd, 8th and 11th columns which contain the Uniprot entry names and two different marker sets to use an input for machine learning analyses (see sections on markers and subsequent sections).
+```{r showlopit2016}
+lopit2016
+```
+
+As briefly mentioned above, the quantitation data is stored in the
+`exprs` slot of the `MSnSet` and can be accessed by
+`exprs(lopit2016)`. Below, we examine the quantitative information for
+first 5 proteins. It is also possible to access specific rows and
+columns by naming the proteins and fractions of interest.
+
+```{r }
+exprs(lopit2016)[1:5, ]
+exprs(lopit2016)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
+```
+
+The feature meta-data is stored in the `fData`
+slot and can be accessed by `fData(lopit2016)`. When using
+`readMSnSet2`, automatically, everything that is not defined as
+quantitation data by `ecol` or the feature names by `fnames` is
+deposited to the `fData` slot. As we wish to demonstrate the complete
+analysis of this data we remove the results from the prior analysis
+described in [@hyper] in the code chunk below. We see the `fData`
+contains 25 columns describing information such as the number of
+peptides, associated markers, machine learning results etc. For
+demonstration in the code chunk below keep the 2nd, 8th and 11th
+columns which contain the Uniprot entry names and two different marker
+sets to use an input for machine learning analyses (see sections on
+markers and subsequent sections).
 

### 2016-06-01 bioc-workflow.Rmd
add short paragraph about pdata

@@
 contains 25 columns describing information such as the number of
-peptides, associated markers, machine learning results etc. For
-demonstration in the code chunk below keep the 2nd, 8th and 11th
+peptides, associated markers, machine learning results etc.
+
+For demonstration in the code chunk below keep the 2nd, 8th and 11th
 columns which contain the Uniprot entry names and two different marker
 sets to use an input for machine learning analyses (see sections on
-markers and subsequent sections).
+markers and subsequent sections) and rename the first feature variable
+names.
 
@@
 fData(lopit2016) <- fData(lopit2016)[, c(2, 8, 11)]
+fvarLabels(lopit2016)[1] <- "UniProtId"
 head(fData(lopit2016))
+```
+
+Note that when using the simple `readMSnSet2` procedure, sample
+meta-data is kept empty. It is advised to annotate the fractions as
+well. Below, we annotate the replicate from which the profiles
+originate and the TMT tag (extracted from the sample/fraction names).
+
+```{r pdata}
+pData(lopit2016)$replicate <- rep(1:2, each = 10)
+pData(lopit2016)$tag <- sub("\\.1$", "", sub("^X", "", sampleNames(lopit2016)))
+pData(lopit2016)
 ```
@@
 ## which are named by Uniprot Entry Name
-featureNames(lopit2016) <- make.unique(fData(lopit2016)[, 1])
+featureNames(lopit2016) <- make.unique(as.character(fData(lopit2016)[, 1]))
 

### 2016-06-01 bioc-workflow.Rmd
Updates to ML section

--- 2016-06-01 bioc-workflow.Rmd
@@
 
-As a use-case, we analyse a recent high-throughput spatial proteomics
+As a use-case we analyse a recent high-throughput spatial proteomics
 dataset from pluripotent mouse embryonic stem cells (E14TG2a)
 [@hyper]. The data was generated using hyperplexed LOPIT (hyperLOPIT),
-an improved method relying on improved sub-cellular fractionation and
+an improved method relying on better sub-cellular fractionation and
 more acurate quantitation, leading to more reliable classification of
@@
 of the MS data acquisition technique named synchronous precursor
-selection MS$^3$ (SPS-MS$^3$) [@McAlister:2014], for TMT
+selection MS$^3$ (SPS)-MS$^3$ [@McAlister:2014], for TMT
 quantification with high accuracy and precision. Three biological
@@
 
-These, as well as many other data are directly available as properly
-structured and annotated computational object from the 
-`r Biocexptpkg("pRolocdata")` experiment package. In this workflow, we
-will start with a description of how to generate these ad hoc objects
-starting from any arbitrary spreadsheets, as produced by many popular
-third-party applications.
+* Need to say that can be label-free or any other type of
+  quantitation, and, possibly, that features can be a peptide level.
 
-While we focus here on a LOPIT-type dataset, these analyses are
-relevant for any quantitative spatial proteomics data, irrespective of
-the fractionation or quantitation (i.e. labeled or label-free)
-methods.
+* I would like to show code for the below - maybe move to later
+  sections. See *Data processing* section.
+
+Before combination, the two replicates were separately normalised by
+sum across the 10 channels (i.e. such that the sum of each protein's
+intensity is 1), for each replicate respectively. Normalisation is an
+essential part of data processing and several methods are available in
+`r Biocpkg("MSnbase")`. The normalisation desired in this specific
+case would be obtained with the a call to the `normalise` method.
+
+The combination of data can also be performed effectively in 
+`r Biocpkg("MSnbase")` as detailed in the dedicated *Combining MSnSet
+instances* section of the `r Biocpkg("MSnbase")`
+[tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
 
@@
 
-To make use of the full functionality of the `r Biocpkg("pRoloc")`
-software one needs to import their data into R and prepare them as an
-`MSnSet`. The `MSnSet` is a dedicated data structure for the efficient
-manipulation and processing of mass spectrometry and proteomics data
-in R. Figure 1 illustrates a simplified view of the `MSnSet`
-structure; there exists 3 key sub-parts (termed slots) to such a data
-object: (1) the `exprs` slot for storing the quantitation data, (2)
-the `fData` slot for storing the feature meta-data, and finally (3)
-the `pData` slot for storing the sample meta-data.
+To make use of the full functionality of the `r Biocpkg("pRoloc")` software one is required to import their data into R as a `MSnSet` instance.
+The `MSnSet` is a dedicated data structure for the efficient manipulation and processing of mass spectrometry and proteomics data in R.
+Figure 1 illustrates a simplified view of the `MSnSet` structure; there exists 3 key slots (1) the `exprs` slot for storing the quantitation data, (2) the `fData` slot for storing the feature meta-data, and finally (3) the `pData` slot for storing the sample meta-data.
+![**Figure 1.** Simplified representation of the `MSnSet` data structure (reproduced with permission from the `r Biocpkg("MSnbase")` vignette)](./Figures/msnset.png)
 
-![Simplified representation of the `MSnSet` data structure (reproduced with permission from the `r Biocpkg("MSnbase")` vignette)](./Figures/msnset.png)
+There are a number of ways to import quantitation data and create a `MSnSet` instance and all methods are described in the `r Biocpkg("MSnbase")` [input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf).
+One suggested simple method is to use the function `readMSnSet2` in `r Biocpkg("MSnbase")`.
+The function takes a single spreadsheet as input and extracts the columns containing the quantitation data, as identified by the argument `ecol`, to create the expression data, while the other columns in the spreadsheet are appended to the feature meta-data slot.
+By example, in the code chunk below we read in the `csv` spreadsheet containing the quantitation data from the intersect of replicates 1 and 2 of the mouse map [@hyper], using the `readMSnSet2` function.
+The data is as available online with the manuscript (see tab 2 of the `xlsx` supplementary data set 1 in [@hyper]) and also as both a `csv` and `MSnSet` object in the Bioconductor `r Biocexptpkg("pRolocdata")` data package.
 
-There are a number of ways to import quantitation data and create an
-`MSnSet` instance and all methods are described in the `r
-Biocpkg("MSnbase")`
-[input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf). One
-suggested simple method is to use the function `readMSnSet2` in 
-`r Biocpkg("MSnbase")`. The function takes a single spreadsheet as input
-and extracts the columns containing the quantitation data, as
-identified by the argument `ecol`, to create the expression data,
-while the other columns in the spreadsheet are appended to the feature
-meta-data slot.  By example, in the code chunk below we read in the
-`csv` spreadsheet containing the quantitation data from the intersect
-of replicates 1 and 2 of the mouse map [@hyper], using the
-`readMSnSet2` function. The data is as available online with the
-manuscript (see tab 2 of the `xlsx` supplementary data set 1 in
-[@hyper], which should be exported as a text-based spreadsheet) and
-also as a `csv` in the Bioconductor `r Biocexptpkg("pRolocdata")` data
-package.
+To use the `readMSnSet2` function, as a minimum one must specify the file path to the data and which columns of the spreadsheet contain quantitation data.
+The `getEcols` function exists to help users identify which columns of the spreadsheet that contain the quantitation data.
+The spreadsheet of E14TG2a data; "hyperLOPIT-SIData-ms3-rep12-intersect.csv", contains 5032 proteins common across the 2 biological replicates for the respective 2 x 10-plex reporter tags, along with associated feature meta-data such as protein markers, protein description, number of quantified peptides etc.
+The spreadsheet contains two headers, with the second header containing information about where the quantitation data is stored.
+We can display the names of the second header by calling the `getEcols` function with the argument `n = 2`, to specify that we wish to display the column names of the second header.
+It is now easy for one to identify that the quantitation data is located in columns 8 to 27.
+It is also possible to pass the optional argument `fnames` to indicate which column to use as the labels by which to identify each protein in the sample.
+Here, we use `fnames = 1` to use the Uniprot identifiers contained in the first column of the spreadsheet.
 
-To use the `readMSnSet2` function, as a minimum one must specify the
-file path to the data and which columns of the spreadsheet contain
-quantitation data. The `getEcols` function exists to help users
-identify which columns of the spreadsheet contain the
-quantitation data.
+```{r makeMSnSet, message=FALSE, warning=FALSE}
+library("MSnbase")
 
-We first locate the the spreadsheet of E14TG2a data distributed with
-the `r Biocexptpkg("pRolocdata")` data using `system.file`, that
-locates the `extdata` in the `r Biocexptpkg("pRolocdata")` package on
-the hard drive and `dir`, that displays the full path to the file
-matching the `hyperLOPIT-SIData-ms3-rep12-intersect.csv` pattern,
-which corresponds the the file of interest. In the last line, we print
-the filename (not the full path, which will vary from computer to
-computer).
+f0 <- dir(system.file("extdata", package = "pRolocdata"), full.names = TRUE, 
+          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
 
-```{r getFilename}
-library("MSnbase")
-extdatadir <- system.file("extdata", package = "pRolocdata")
-csvfile <- dir(extdatadir, full.names = TRUE,
-          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
-basename(csvfile)
-```
+getEcols(f0, split = ",", n = 2)
 
-Note that the file is compressed (as indicated by the `gz`, for
-`gzip`, extension), and will be decompressed on-the-fly when read into
-R later on.
-
-The spreadsheet that was deposited by the authors contains two
-headers, with the second header containing information about where the
-quantitation data is stored. We can display the names of the second
-header by calling the `getEcols` function with the argument `n = 2`
-(the default value is `n = 1`), to specify that we wish to display the
-column names of the second line.
-
-```{r getEcols}
-getEcols(csvfile, split = ",", n = 2)
-```
-
-It is now easy for one to
-identify that the quantitation data is located in columns 8 to 27. It
-is also possible to pass the optional argument `fnames` to indicate
-which column to use as the labels by which to identify each protein in
-the sample. Here, we use `fnames = 1` to use the Uniprot identifiers
-contained in the first column of the spreadsheet.
-
-```{r readMSnSet2}
-lopit2016 <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1, 
+lopit2016 <- readMSnSet2(f0, ecol = c(8:27), fnames = 1, skip = 1, 
                          stringsAsFactors = FALSE)
@@
 ```
-
-contains 5032 proteins common across the 2 biological replicates for
-the respective 2 x 10-plex reporter tags, along with associated
-feature meta-data such as protein markers, protein description, number
-of quantified peptides etc.
 
@@
 * combining acquisistions
-
-Before combination, the two replicates were separately normalised by
-sum across the 10 channels (i.e. such that the sum of each protein's
-intensity is 1), for each replicate respectively. Normalisation is an
-essential part of data processing and several methods are available in
-`r Biocpkg("MSnbase")`. The normalisation desired in this specific
-case would be obtained with the a call to the `normalise` method.
-
-The combination of data can also be performed effectively in 
-`r Biocpkg("MSnbase")` as detailed in the dedicated *Combining MSnSet
-instances* section of the `r Biocpkg("MSnbase")`
-[tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
 
@@
 
-There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` Machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
-We find the general tendancy to be that it is not the choice of classifier but the proper optimisation of the algorithms parameters that limit the classification results.
-Before employing a classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
-For example, in the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
-In the example below, the training data is found in the `featureData` slot in the column called `SVM.marker.set`.
-
-Notes:
-- take care of properly setting the model parameters. 
-- Wrongly set parameters can have adverse effects 
-- how well they represent the multivariate data?
-- Parameter optimisation can be conducted in a number of ways. One of the most common is train/test plus 5-fold x-val
+There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
+We find the general tendancy to be that it is not the choice of classifier, but the optimisation of the algorithmic parameters, that limits classification accuracy.
+Before employing any classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
+In the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
+The training data is found in the `featureData` slot in the column called `SVM.marker.set`.
+As previously mentioned one first needs to train the classifiers parameters before an algorithm can be used to predict the class labels of the proteins with unknown location.
+One of the most common ways to optimise the parameters of a classifier is to partition the labelled data in to training and testing subsets.
+In this framework parameters are tested via a grid search using cross-validation on the training partition.
+The best parameters chosen from the cross-validation stage are then used to build a classifier to predict the class labels of the protein profiles on the test partition.
+Observed and expected classication results can be compared, and then used to assess how well a given model works by getting an estimate of the classiers ability to achieve a good generalisation i.e.
+that is given an unknown example predict its class label with high accuracy.
+In `pRoloc` algorithmic performance is estimated using stratified 80/20 partitioning for the training/testing subsets respectively, in conjuction with five-fold cross-validation in order to optimise the free parameters via a grid search.
+This procedure is usually repeated 100 times and then the best parameter(s) are selected upon investigation of classifier accuracy, here we use the harmonic mean of precision and recall; the macro F1 score.
+In the code chunk below we demonstrate how to optimise the free parameters; `sigma` and `cost`, of a classical SVM classifier with a Gaussian kernel using the function `svmOptimisation`.
+As the number of labelled instances per class varies from organelle to organelle, we can account for class imbalance by setting specific class weights when generating the SVM model.
+Below the weights, `w` are set to be inversely proportional to the class frequencies.
 
@@
 ```{r optimise, eval=FALSE}
-w <- table(fData(pdRes)[, "SVM.marker.set"])
+(w <- table(fData(pdRes)[, "SVM.marker.set"]))
 w <- 1/w[names(w) != "unknown"]
 
+## 100 rounds of optimisation with five-fold cross-validation
 params <- svmOptimisation(pdRes, fcol = "SVM.marker.set",
@@
                           class.weights = w,
-                          verbose = TRUE)
+                          verbose = FALSE)
 ```
+
+The output `params` is an object of class `GenRegRes`; a regularisation/optimisation framework container for the optimisation results.
+
+associated macro F1 scores. A high macro F1 score
+indicates that the marker proteins in the test dataset are consistently correctly assigned by
+17
+Spatial proteomics data analysis
+the algorithm. Often more than one parameter or set of parameters gives rise to the best
+generalisation accuracy. As such it is always important to investigate the model parameters
+and critically assess the best choice. The best choice may not be as simple as the parameter
+set that gives rise to the highest macro F1 score and one must be careful to avoid overtting
+and to choose parameters wisely.
+
 ```{r visualiseOpt}
@@
 ```
+
+Notes:
+- take care of properly setting the model parameters. 
+- Wrongly set parameters can have adverse effects 
+- how well they represent the multivariate data?
+- Parameter optimisation can be conducted in a number of ways. One of the most common is train/test plus 5-fold x-val
 

### 2016-06-01 bioc-workflow.Rmd
optimisation, classification sections

@@
 There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
-We find the general tendancy to be that it is not the choice of classifier, but the optimisation of the algorithmic parameters, that limits classification accuracy.
+We find the general tendancy to be that it is not the choice of classifier, but the improper optimisation of the algorithmic parameters, that limits classification accuracy.
 Before employing any classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
+
+## Optimisation
+
 In the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
@@
 
-The output `params` is an object of class `GenRegRes`; a regularisation/optimisation framework container for the optimisation results.
-
-associated macro F1 scores. A high macro F1 score
-indicates that the marker proteins in the test dataset are consistently correctly assigned by
-17
-Spatial proteomics data analysis
-the algorithm. Often more than one parameter or set of parameters gives rise to the best
-generalisation accuracy. As such it is always important to investigate the model parameters
-and critically assess the best choice. The best choice may not be as simple as the parameter
-set that gives rise to the highest macro F1 score and one must be careful to avoid overtting
-and to choose parameters wisely.
+The output `params` is an object of class `GenRegRes`; a dedicated container for the storage of the design and results from a machine learning optimisation.
+To assess classifier performance we can examine the macro F1 scores and the most frequently chosen parameters.
+A high macro F1 score indicates that the marker proteins in the test dataset are consistently and correctly assigned by the the algorithm.
+Often more than one parameter or set of parameters gives rise to the best generalisation accuracy.
+As such it is always important to investigate the model parameters and critically assess the best choice.
+The best choice may not be as simple as the parameter set that gives rise to the highest macro F1 score and one must be careful to avoid overfitting and to choose parameters wisely.
 
 ```{r visualiseOpt}
+(best <- getParams(params))
 plot(params)
 levelPlot(params)
-(best <- getParams(params))
 ```
+
+By using the function `getParams` we can extract the best set of parameters.
+Currently, `getParams` retrieves the first best set is automatically but users are encouraged to critically assess whether this is the most wise choice by visualising the results using the methods `plot` and `levelPlot`.
+The `plot` method for `GenRegRes` object shows the respective distributions of the 100 macro F1 scores for the best cost/sigma parameter pairs, and `levelPlot` shows the averaged macro F1 scores, for the full range of parameter values.
+Once we have selected the best parameters we can then use them to build a classifier from the labelled marker proteins.
+
+## Classification
+
+We can use the function `svmClassification` to return a classification result for all unlabelled instances in the dataset corresponding to their most likely sub-cellular location.
+The algorithm parameters are passed to the function, along with the class weights and the `fcol` to tell the function where the labelled training data is located, in our case, our marker proteins are located in the `featureData` with the label `"SVM.marker.set"`.
+
 ```{r classify, eval=TRUE}
@@
 
-Notes:
-- take care of properly setting the model parameters. 
-- Wrongly set parameters can have adverse effects 
-- how well they represent the multivariate data?
-- Parameter optimisation can be conducted in a number of ways. One of the most common is train/test plus 5-fold x-val
+Automatically, the output of the above classification; the organelle predictions and assignment scores, are stored in the `featureData` slot of the `MSnSet`.
+In this case, they are given the labels `svm` and `svm.scores` for the predictions and scores respectively.
+The resultant predictions can be visualised using `plot2D`.
+In the code chunk below `plot2D` is called to generate a PCA plot of the data and `fcol` is used to specify where the new assignments are located, here for example these are located in the column called `svm`.
+Additionally, when calling `plot2D` we use the `cex` argument to change the point size of each point on the plot (where one point represents one protein) to be inversely proportional to the SVM score.
+This gives an initial overview of the high scoring localisations from the SVM predictions.
+
+```{r setColors, include=FALSE}
+setStockcol(paste0(getStockcol(), 70))
+```
+```{r plotSVM, eval=TRUE}
+ptsze <- exp(fData(svmRes)$svm.scores) - 1
+plot2D(svmRes, fcol = "svm", cex = ptsze)
+addLegend(svmRes, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
+```
+
+## Thresholding
 

### 2016-06-01 bioc-workflow.Rmd
prepare data processing sub-sections

--- 2016-06-01 bioc-workflow.Rmd
@@
 
-* normalisation
-* missing data
-* combining acquisistions
+### Normalisation
 
@@
 case would be obtained with the a call to the `normalise` method.
+
+### Missing data
+
+### Combining acquisistions
 
@@
 
-There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
-We find the general tendancy to be that it is not the choice of classifier, but the optimisation of the algorithmic parameters, that limits classification accuracy.
-Before employing any classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
-In the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
-The training data is found in the `featureData` slot in the column called `SVM.marker.set`.
-As previously mentioned one first needs to train the classifiers parameters before an algorithm can be used to predict the class labels of the proteins with unknown location.
-One of the most common ways to optimise the parameters of a classifier is to partition the labelled data in to training and testing subsets.
-In this framework parameters are tested via a grid search using cross-validation on the training partition.
-The best parameters chosen from the cross-validation stage are then used to build a classifier to predict the class labels of the protein profiles on the test partition.
-Observed and expected classication results can be compared, and then used to assess how well a given model works by getting an estimate of the classiers ability to achieve a good generalisation i.e.
-that is given an unknown example predict its class label with high accuracy.
-In `pRoloc` algorithmic performance is estimated using stratified 80/20 partitioning for the training/testing subsets respectively, in conjuction with five-fold cross-validation in order to optimise the free parameters via a grid search.
-This procedure is usually repeated 100 times and then the best parameter(s) are selected upon investigation of classifier accuracy, here we use the harmonic mean of precision and recall; the macro F1 score.
-In the code chunk below we demonstrate how to optimise the free parameters; `sigma` and `cost`, of a classical SVM classifier with a Gaussian kernel using the function `svmOptimisation`.
-As the number of labelled instances per class varies from organelle to organelle, we can account for class imbalance by setting specific class weights when generating the SVM model.
-Below the weights, `w` are set to be inversely proportional to the class frequencies.
+There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` Machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
+We find the general tendancy to be that it is not the choice of classifier but the proper optimisation of the algorithms parameters that limit the classification results.
+Before employing a classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
+For example, in the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
+In the example below, the training data is found in the `featureData` slot in the column called `SVM.marker.set`.
+
+Notes:
+- take care of properly setting the model parameters. 
+- Wrongly set parameters can have adverse effects 
+- how well they represent the multivariate data?
+- Parameter optimisation can be conducted in a number of ways. One of the most common is train/test plus 5-fold x-val
 
@@
 ```{r optimise, eval=FALSE}
-(w <- table(fData(pdRes)[, "SVM.marker.set"]))
+w <- table(fData(pdRes)[, "SVM.marker.set"])
 w <- 1/w[names(w) != "unknown"]
 
-## 100 rounds of optimisation with five-fold cross-validation
 params <- svmOptimisation(pdRes, fcol = "SVM.marker.set",
@@
                           class.weights = w,
-                          verbose = FALSE)
+                          verbose = TRUE)
 ```
-
-The output `params` is an object of class `GenRegRes`; a regularisation/optimisation framework container for the optimisation results.
-
-associated macro F1 scores. A high macro F1 score
-indicates that the marker proteins in the test dataset are consistently correctly assigned by
-17
-Spatial proteomics data analysis
-the algorithm. Often more than one parameter or set of parameters gives rise to the best
-generalisation accuracy. As such it is always important to investigate the model parameters
-and critically assess the best choice. The best choice may not be as simple as the parameter
-set that gives rise to the highest macro F1 score and one must be careful to avoid overtting
-and to choose parameters wisely.
-
 ```{r visualiseOpt}
@@
 ```
-
-Notes:
-- take care of properly setting the model parameters. 
-- Wrongly set parameters can have adverse effects 
-- how well they represent the multivariate data?
-- Parameter optimisation can be conducted in a number of ways. One of the most common is train/test plus 5-fold x-val
 

### 2016-06-02 bioc-workflow.Rmd
data processing section

--- 2016-06-01 bioc-workflow.Rmd
@@
 suppressPackageStartupMessages(library("pRoloc"))
+suppressPackageStartupMessages(library("pRolocdata"))
 ```
@@
 
-# Reading and handling mass-spectrometry based proteomics data
+# Reading and processing spatial proteomics data
 
@@
 
+## Importing data
+
 There are a number of ways to import quantitation data and create an
-`MSnSet` instance and all methods are described in the `r
-Biocpkg("MSnbase")`
+`MSnSet` instance and all methods are described in the `r Biocpkg("MSnbase")`
 [input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf). One
@@
 
-* normalisation
-* missing data
-* combining acquisistions
+### Normalisation
 
-Before combination, the two replicates were separately normalised by
-sum across the 10 channels (i.e. such that the sum of each protein's
-intensity is 1), for each replicate respectively. Normalisation is an
-essential part of data processing and several methods are available in
-`r Biocpkg("MSnbase")`. The normalisation desired in this specific
-case would be obtained with the a call to the `normalise` method.
+There are two aspects related to data normalisation that are relevant
+to spatial proteomics data processing. The first one focuses on
+reducing purely technical variation between fractions without
+affecting biological variability (i.e. the shape of the quantitative
+profiles) too much, as generally focused on. This normalisation will
+depend on the underlying quantitative technology and the experimental
+design, and will not be addressed in this workflow. In addition, and
+more specific to spatial proteomics data, it is fundamental to scale
+all the organelle-specific profiles into a same intensity interval
+(typically 0 and 1) by, for example, diving each intensity by the sum
+of the intensities for that quantitative feature, as illustrated
+below.
 
-The combination of data can also be performed effectively in 
-`r Biocpkg("MSnbase")` as detailed in the dedicated *Combining MSnSet
-instances* section of the `r Biocpkg("MSnbase")`
+```{r normsum, eval = FALSE}
+lopit2016 <- normalise(lopit2016, method = "sum")
+```
+
+This transformation of the data assures to cancel the effect of the
+absolute intensities of the quantitive features along the rows, to
+focus subsequent analyses on the relative profiles along the
+sub-cellular fractions.
+
+The same `normalise` function (or `normalize`, both spellings are
+supported) can also be applied in the first case described
+above. Different normalisation methods such as mean or median scaling,
+variance stabilitation or quantile normalisation, to cite a few, can
+be applied to accomodation different needs.
+
+Before combination, the two replicates in the `lopit2016` data that we
+read into R were separately normalised by sum across the 10 channels
+for each replicate respectively. We can verify this by summing each
+rows for each replicate:
+
+```{r normcheck}
+summary(rowSums(exprs(lopit2016[, lopit2016$replicate == 1])))
+summary(rowSums(exprs(lopit2016[, lopit2016$replicate == 2])))
+```
+We see that some features do not add up exactly to 1 due to rounding
+errors after exporting to intermediate files. These small deviations
+do not bear any consequences here.
+
+### Combining acquisitions
+
+The spreadsheet that was used to create the `lopit2016` MSnSet
+included two replicates. We also provide individual replicates in the
+`r Biocexptpkg("pRolocdata")` package. Below, we show how to combine
+`MSnSet` objects and, subsequently, how to filter and hangle missing
+values. We start by loading the experimental data package and the
+equivalent replicates using the `load` function.
+
+```{r reps4combine}
+library("pRolocdata")
+data(hyperLOPIT2015ms3r1)
+data(hyperLOPIT2015ms3r2)
+```
+
+To list all data distributed with a package, one can use pass the
+package name as input to the `data` function. For instance
+
+```{r pkgdata, eval=FALSE}
+data(package = "pRolocdata")
+```
+
+will list the `r nrow(data(package = "pRolocdata")$results)` datasets
+that are available in `r Biocexptpkg("pRolocdata")`.
+
+Combining data is performed with the `combine` function. This function
+will inspect the feature and sample names to identify how to combine
+the data. As we want our replicates to be combined along the columns
+(same proteins, different sets of fractions), we need to assure that
+the respective sample names differ.
+
+```{r samplenames}
+identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
+hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
+hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
+sampleNames(hyperLOPIT2015ms3r1)
+sampleNames(hyperLOPIT2015ms3r2)
+```
+
+In addition, to matching names, the content of the feature metadata
+for identical feature annotations must match exactly across the data
+to be combined. In particular for these data, we expect the same
+proteins in each replicates to be annotated with the same UniProt
+entry names and descriptions, but not with the same coverage of number
+of peptides or peptide-spectrum matches (PSMs).
+
+```{r fvarnames}
+fvarLabels(hyperLOPIT2015ms3r1)
+fvarLabels(hyperLOPIT2015ms3r2)
+```
+
+Below, we update the replicate specific feature variable names and
+remove the shared annotation.
+
+```{r r fvarnames2}
+fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
+fvarLabels(hyperLOPIT2015ms3r2)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r2)[3:5], 2)
+fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
+```
+
+We can now combine the two experiments into a single `MSnSet`:
+
+```{r combine}
+hyperLOPIT <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
+hyperLOPIT
+```
+
+More details above combining data are given in the dedicated
+*Combining MSnSet instances* section of the `r Biocpkg("MSnbase")`
 [tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf).
+
+### Missing data
+
+Missing data are a recurrent issue in mass spectrometry applications,
+and should be addressed independently of this workflow
+[@Lazar:2016]. In [@Gatto:2014b], we have described how a high content
+in missing values in spatial proteomics data and their inappropriate
+handling leads to a reduction of sub-cellular resolution.
+
+```{r namap}
+image2(is.na(hyperLOPIT), col = c("black", "white"))
+```
+
+## Replication
+
+See issue #6.
 

### 2016-06-02 bioc-workflow.Rmd
Small updates to markers and ML

--- 2016-06-01 bioc-workflow.Rmd
@@
 In the context of spatial proteomics, a marker protein is defined as a well-known resident of a specific sub-cellular niche in a species *and* condition of interest.
-Applying this to machine learning (ML), and specifically supervised learning, for the task of protein localisation prediction, markers constitute the labelled training data to use as input for a classification analyses.
-Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
+Applying this to machine learning (ML), and specifically supervised learning, for the task of protein localisation prediction, markers constitute the labelled training data to use as input to a classification analyses.
+Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but it is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
 `r Biocpkg("pRoloc")` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object, as demonstrated in the code chunk below.
@@
 The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
-The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2016))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
-It is then possible to match names between the markers and the `MSnSet` instance.
-We see below that the markers cover many sub-cellular niches, with many niches only containing a few proteins.
-Depending on the biological question and downstream analyses we may wish the subset these marker classes, this can be done using the `minMarkers` function.
+The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2016))`) and the names of the vector of the mouse markers (`mmus` markers) are also Uniprot IDs (see `head(mrk)` in the code chunk below), so it is straightforward to match names between the markers and the `MSnSet` instance.
+If the naming between the marker sets and the `MSnSet` dataset are different, one will have to convert and match the proteins according to the appropriate identifier.
+Sometimes, we find the equivalent entry name, Uniprot ID or accession number is stored with the data, which makes conversion between identifers relatively straightforward.
+If this is not the case however, there are conversion softwares online available, for example XXX.
+
 In the code chunk below, we demonstrate how to add markers using `pRolocmarkers` function and then visualise these annotations using the `plot2D` function.
@@
 
-## Change featureNames of the MSnSet to match the marker names 
-## which are named by Uniprot Entry Name
-featureNames(lopit2016) <- make.unique(as.character(fData(lopit2016)[, 1]))
-
 ## Add mouse markers
@@
 
-## Remove marker sets with < 20 proteins
-lopit2016 <- minMarkers(lopit2016, n = 20)
-getMarkers(lopit2016, fcol = "markers20")
-plot2D(lopit2016, fcol = "markers20", main = "pRolocmarkers for mouse")
-
-## After expert curation
-plot2D(lopit2016, fcol = "SVM.marker.set", main = "Curated markers")
+## Plot mouse markers
+plot2D(lopit2016, fcol = "markers", main = "pRolocmarkers for mouse")
 ```
@@
 Expert curation is vital to check that any annotation added is in the correct context for the the biological question under investigation.
-In the code chunk above we show the PCA plot output of the mouse dataset with (i) the annotation for mouse pulled from `pRolocmarkers`, and (ii) annotation after expert curation (stored in the `featureData` column called `SVM.marker.set` that was used for a classification analyses in the original data analyses [@hyper]).
 
@@
 
-There are several classification algorithms available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
+There are several classification algorithms readily available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
 We find the general tendancy to be that it is not the choice of classifier, but the improper optimisation of the algorithmic parameters, that limits classification accuracy.
-Before employing any classification algorithm and generating a model on the training data with which to classify our set of unknown residents, one must find the optimal parameters for the algorithm of choice.
+Before employing any classification algorithm and generating a model on the training data, one must first find the optimal parameters for the algorithm of choice.
 
@@
 
-```{r visualiseOpt}
+```{r visualiseOpt, eval=FALSE}
 (best <- getParams(params))
@@
 levelPlot(params)
+```
+```{r visualiseOptHide, include=FALSE}
+grid.arrange(plot(params), levelPlot(params), ncol=2)
 ```

### 2016-06-02 bioc-workflow.Rmd
processing section

--- 2016-06-02 bioc-workflow.Rmd
@@
 combining replicates across from different gradients can increase
-spatial resolution [@trotter]. The combination of replicates resulted
+spatial resolution [@Trotter:2010]. The combination of replicates resulted
 in 5032 proteins common in both experiments.
@@
 `n = 2` in `getEcols` above) to read the `csv` data and convert it to an
-`MSnSet` object.
+`MSnSet` object, named `hl` (for hyperLOPIT).
 
 ```{r readMSnSet2}
-lopit2016 <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1)
+hl <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1)
 ```
@@
 Below, we display a short summary of the data. The data contains 
-`r nrow(lopit2016)` proteins/features common across the 2 biological replicates
-for the respective 2 x 10-plex reporter tags (`r ncol(lopit2016)`
+`r nrow(hl)` proteins/features common across the 2 biological replicates
+for the respective 2 x 10-plex reporter tags (`r ncol(hl)`
 columns/samples), along with associated feature meta-data such as
@@
 
-```{r showlopit2016}
-lopit2016
+```{r showhl}
+hl
 ```
@@
 `exprs` slot of the `MSnSet` and can be accessed by
-`exprs(lopit2016)`. Below, we examine the quantitative information for
+`exprs(hl)`. Below, we examine the quantitative information for
 first 5 proteins. It is also possible to access specific rows and
@@
 ```{r }
-exprs(lopit2016)[1:5, ]
-exprs(lopit2016)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
+exprs(hl)[1:5, ]
+exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
 ```
@@
 The feature meta-data is stored in the `fData`
-slot and can be accessed by `fData(lopit2016)`. When using
+slot and can be accessed by `fData(hl)`. When using
 `readMSnSet2`, automatically, everything that is not defined as
@@
 ```{r removefData}
-fvarLabels(lopit2016)
-fData(lopit2016) <- fData(lopit2016)[, c(2, 8, 11)]
-fvarLabels(lopit2016)[1] <- "UniProtId"
-head(fData(lopit2016))
+fvarLabels(hl)
+fData(hl) <- fData(hl)[, c(2, 8, 11)]
+fvarLabels(hl)[1] <- "UniProtId"
+head(fData(hl))
 ```
@@
 ```{r pdata}
-pData(lopit2016)$replicate <- rep(1:2, each = 10)
-pData(lopit2016)$tag <- sub("\\.1$", "", sub("^X", "", sampleNames(lopit2016)))
-pData(lopit2016)
+pData(hl)$replicate <- rep(1:2, each = 10)
+pData(hl)$tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
+pData(hl)
 ```
@@
 ```{r normsum, eval = FALSE}
-lopit2016 <- normalise(lopit2016, method = "sum")
+hl <- normalise(hl, method = "sum")
 ```
@@
 
-Before combination, the two replicates in the `lopit2016` data that we
+Before combination, the two replicates in the `hl` data that we
 read into R were separately normalised by sum across the 10 channels
@@
 ```{r normcheck}
-summary(rowSums(exprs(lopit2016[, lopit2016$replicate == 1])))
-summary(rowSums(exprs(lopit2016[, lopit2016$replicate == 2])))
+summary(rowSums(exprs(hl[, hl$replicate == 1])))
+summary(rowSums(exprs(hl[, hl$replicate == 2])))
 ```
@@
 
-The spreadsheet that was used to create the `lopit2016` MSnSet
+The spreadsheet that was used to create the `hl` MSnSet
 included two replicates. We also provide individual replicates in the
@@
 and should be addressed independently of this workflow
-[@Lazar:2016]. In [@Gatto:2014b], we have described how a high content
-in missing values in spatial proteomics data and their inappropriate
-handling leads to a reduction of sub-cellular resolution.
+[@Webb-Robertson:2015;@Lazar:2016]. In [@Gatto:2014b], we have
+described how a high content in missing values in spatial proteomics
+data and their inappropriate handling leads to a reduction of
+sub-cellular resolution. Missing data can be imputated using
+
+`r Biocpkg("MSnbase")`'s `impute` function. The method underlying the
+imputation method is then determined by a `methods` parameter. In our
+particular case, missing values are indicative of protein groups that
+were not profiles in both replicates.
 
 ```{r namap}
-image2(is.na(hyperLOPIT), col = c("black", "white"))
+image2(is.na(hyperLOPIT), col = c("black", "white"),
+       main = "Missing values (white cells) after combining replicates")
+```
+
+We prefer to remove proteins that were not assayed in our two
+replicated experiments. This is done with the `filterNA` function that
+removes features that contain more than a certain proportion (default
+is 0) missing values.
+
+```{r filterNA}
+hyperLOPIT <- filterNA(hyperLOPIT)
+hyperLOPIT
+```
+
+Which results in two identical matrices of quantitative protein
+profiles.
+
+```{r equaldata}
+all.equal(exprs(hl), exprs(hyperLOPIT), check.attributes = FALSE)
 ```
@@
 library('pRoloc')
-plot2D(lopit2016, fcol = NULL)
+plot2D(hl, fcol = NULL)
 ```
@@
 The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
-The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2016))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
+The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(hl))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
 It is then possible to match names between the markers and the `MSnSet` instance.
@@
 
-## Change featureNames of the MSnSet to match the marker names 
-## which are named by Uniprot Entry Name
-featureNames(lopit2016) <- make.unique(as.character(fData(lopit2016)[, 1]))
-
 ## Add mouse markers
-lopit2016 <- addMarkers(lopit2016, mrk)
+hl <- addMarkers(hl, mrk)
 
 ## Remove marker sets with < 20 proteins
-lopit2016 <- minMarkers(lopit2016, n = 20)
-getMarkers(lopit2016, fcol = "markers20")
-plot2D(lopit2016, fcol = "markers20", main = "pRolocmarkers for mouse")
+hl <- minMarkers(hl, n = 20)
+getMarkers(hl, fcol = "markers20")
+plot2D(hl, fcol = "markers20", main = "pRolocmarkers for mouse")
 
 ## After expert curation
-plot2D(lopit2016, fcol = "SVM.marker.set", main = "Curated markers")
+plot2D(hl, fcol = "SVM.marker.set", main = "Curated markers")
 ```
@@
 ## As per the hyperLOPIT paper
-pdRes <- phenoDisco(lopit2016, fcol = "phenoDisco.Input", times = 200, 
+pdRes <- phenoDisco(hl, fcol = "phenoDisco.Input", times = 200, 
                     GS = 20, p = 0.05)
@@
 The trained classifier can then be used to predict the localisation of a protein of unknown localisation, based on its observed protein profile.
-To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @trotter; @Hall:2009; @Dunkley:2006; @Tan:2009].
+To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006; @Tan:2009].
 
@@
 ```{r loadpdRes, include=FALSE}
-pdRes <- lopit2016
+pdRes <- hl
 load("Data/fusionparams.rda")
+(w <- table(fData(pdRes)[, "SVM.marker.set"]))
+w <- 1/w[names(w) != "unknown"]
 ```
+
 ```{r optimise, eval=FALSE}

### 2016-06-02 bioc-workflow.Rmd

@@
 library("BiocStyle")
+suppressPackageStartupMessages(library("gridExtra"))
 suppressPackageStartupMessages(library("MSnbase"))
@@
 
-```{r r fvarnames2}
+```{r fvarnames2}
 fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
@@
 The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
-The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(hl))`) and the names of the vector of the mouse markers (`mmus` markers) are Uniprot entry names (see `head(mrk)` in the code chunk below), therefore it is required we update the `MSnSet` to match the names of the markers, as demonstrated below.
-It is then possible to match names between the markers and the `MSnSet` instance.
-We see below that the markers cover many sub-cellular niches, with many niches only containing a few proteins.
-Depending on the biological question and downstream analyses we may wish the subset these marker classes, this can be done using the `minMarkers` function.
+The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2016))`) and the names of the vector of the mouse markers (`mmus` markers) are also Uniprot IDs (see `head(mrk)` in the code chunk below), so it is straightforward to match names between the markers and the `MSnSet` instance.
+If the naming between the marker sets and the `MSnSet` dataset are different, one will have to convert and match the proteins according to the appropriate identifier.
+Sometimes, we find the equivalent entry name, Uniprot ID or accession number is stored with the data, which makes conversion between identifers relatively straightforward.
+If this is not the case however, there are conversion softwares online available, for example XXX.
+
 In the code chunk below, we demonstrate how to add markers using `pRolocmarkers` function and then visualise these annotations using the `plot2D` function.
@@
 
-## Remove marker sets with < 20 proteins
-hl <- minMarkers(hl, n = 20)
-getMarkers(hl, fcol = "markers20")
-plot2D(hl, fcol = "markers20", main = "pRolocmarkers for mouse")
-
-## After expert curation
-plot2D(hl, fcol = "SVM.marker.set", main = "Curated markers")
+## Plot mouse markers
+plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
 ```
@@
 levelPlot(params)
+```
+```{r visualiseOptHide, include=FALSE}
+grid.arrange(plot(params), levelPlot(params), ncol = 2)
 ```

### 2016-06-05 bioc-workflow.Rmd
replication section stub

@@
 
-See issue #6.
+With the aim of maximising the sub-cellular resolution and,
+consequently, the reliability in protein sub-cellular assignments, we
+follow the advice in [@Trotter:2010] and combine replicated spatial
+proteomics experiments and described above. Trotter et al. have shown
+that ...
+
+Why is the below no a good indication: fraction, even if same index,
+are not necessarily same; replication of single fractions fractions is
+not really relevant, but rather complete profiles; more than single
+profiles, it's rather relative profiles in the frame of whole dataset.
+
+```{r eval = FALSE}
+par(mfrow = c(3, 4))
+for (i in 1:10) {
+    plot(exprs(hl[, c(i, i+10)]))
+    grid()
+    abline(0, 1, col = "red")
+}
+```
+
+```{r plot2Drep}
+par(mfrow = c(1, 2))
+plot2D(hl[hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
+plot2D(hl[hl$replicate == 2], fcol = "SVM.marker.set", mirrorY = TRUE, main = "Replicate 2")
+```
+
+Could compute squared differences between fractions over profile for
+each protein, and compare to Dunkley 2006, to show improvement in
+replicability. Or to have an internal replication measurement. Or what
+about wave?
+
+This might be better off in the QC section...
 
@@
 ```
-```{r visualiseOptHide, include=FALSE}
+```{r visualiseOptHide, echo=FALSE}
 grid.arrange(plot(params), levelPlot(params), ncol = 2)

### 2016-06-05 bioc-workflow.Rmd
adjust fig width

@@
 
-```{r plot2Drep}
+```{r plot2Drep, fig.width = 12}
 par(mfrow = c(1, 2))
@@
 ```
-```{r visualiseOptHide, echo=FALSE}
+```{r visualiseOptHide, echo=FALSE, fig.width = 12}
 grid.arrange(plot(params), levelPlot(params), ncol = 2)

### 2016-06-09 bioc-workflow.Rmd
TODO: other plot2D methods
replicates section and example

--- 2016-06-02 bioc-workflow.Rmd
@@
 library("BiocStyle")
+suppressPackageStartupMessages(library("knitr"))
 suppressPackageStartupMessages(library("gridExtra"))
@@
 
-## Replication
-
-See issue #6.
+When more than 2 data are to be combine and/or too many proteins were
+not consistently assayed, leading to too many proteins being filtered
+out, we sould suggest to implement an ensemble of classifiers voting
+on protein-sub-cellular niche membership over the output of several
+experiments (see section *Supervised machine learning* for the
+description of sub-cellular assingments).
 
@@
 ```
+
+**TODO** Mention that there are other dimensionality reduction
+techniques. Advantages, disadvantages? See
+[this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
+very interesting gist with intriguing tSNE examples.
 
@@
 Expert curation is vital to check that any annotation added is in the correct context for the the biological question under investigation.
+
+# Replication
+
+With the aim of maximising the sub-cellular resolution and,
+consequently, the reliability in protein sub-cellular assignments, we
+follow the advice in [@Trotter:2010] and combine replicated spatial
+proteomics experiments and described above.
+
+Trotter et al. have shown a significant improvement in
+protein–organelle association upon direct combination of single
+experiments, in particular when these resolve different subcellular
+niches.
+
+Direct comparisons of individual fractions in replicated experiments
+does not provide an good, goal-driven assessment of different
+experiments. Indeed, due to the nature of the experiment and fraction
+collection, the quantitative channels (loosely referred to as
+fractions) do not correspond to identical fractions along the
+gradient. As we see in the table below (taken from the `hyperLOPIT`'s
+`pData`), different sets of fractions are combined to obtain enough
+material and optimise acurate quantitation.
+
+```{r pdtab, echo=FALSE}
+i <- hyperLOPIT$Fraction.No %in% 7:10
+kable(pData(hyperLOPIT)[i, -3], row.names = FALSE)
+```
+
+On the figure below, we compare the relative intensities of fractions
+7 to 10 on scatter plots, highlighting different marker sets, in
+particular mitochondrion and chromatin. These differences result from
+different peak fractions for these niches.
+
+```{r repl1, eval = TRUE, echo=FALSE, fig.width = 12}
+sel0 <- fData(hyperLOPIT)$markers != "unknown"
+cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
+cls[!sel0] <- getUnknowncol()
+
+par(mfrow = c(2, 4))
+for (i in c(7, 8, 10, 9)) {
+    plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
+         main = paste("Fraction", pData(hyperLOPIT)[i, "Fraction.No"]))
+    points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
+    grid()
+    abline(0, 1, col = "red")
+}
+plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "n")
+addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
+
+i <- fData(hyperLOPIT)$markers == "Mitochondrion"
+plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
+         featureNames(hyperLOPIT)[i], type = "l",
+         mcol = "orange")
+title(main = "Mitochondrion")
+## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
+matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
+legend("topleft", paste("Replicate", 1:2),
+       col = c("#F781BF", "orange"),
+       lty = 1, bty = "n")
+
+i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
+plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
+         featureNames(hyperLOPIT)[i], type = "l")
+## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
+matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
+title(main = "Chromatin")
+legend("topleft", paste("Replicate", 1:2),
+       col = c("#984EA3", "steelblue"),
+       lty = 1, bty = "n")
+```
+
+The relevant comparison unit is not a single fraction, but rather
+complete protein occupancy profiles, which are best visualised as a
+whole on a PCA plot. As such, we prefer to focus on the direct
+comparison of individual replicate PCA plots, assuring that each
+displays acceptable sub-cellular resolution. Note that in the code
+chunk below, we mirror the y-axis to represent the two figures with
+the same orientation.
+
+```{r plot2Drep, fig.width = 12}
+par(mfrow = c(1, 2))
+plot2D(hl[hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
+plot2D(hl[hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
+       mirrorY = TRUE)
+```
 
@@
 ```
-```{r visualiseOptHide, include=FALSE}
+```{r visualiseOptHide, echo=FALSE, fig.width = 12}
 grid.arrange(plot(params), levelPlot(params), ncol = 2)

### 2016-06-09 bioc-workflow.Rmd
update repl section

@@
 not consistently assayed, leading to too many proteins being filtered
-out, we sould suggest to implement an ensemble of classifiers voting
-on protein-sub-cellular niche membership over the output of several
+out, we suggest to implement an ensemble of classifiers voting on
+protein-sub-cellular niche membership over the output of several
 experiments (see section *Supervised machine learning* for the
@@
 follow the advice in [@Trotter:2010] and combine replicated spatial
-proteomics experiments and described above.
-
-Trotter et al. have shown a significant improvement in
-protein–organelle association upon direct combination of single
-experiments, in particular when these resolve different subcellular
-niches.
+proteomics experiments and described above. Indeed, Trotter et
+al. have shown a significant improvement in protein–organelle
+association upon direct combination of single experiments, in
+particular when these resolve different subcellular niches.
 
 Direct comparisons of individual fractions in replicated experiments
-does not provide an good, goal-driven assessment of different
+does not provide an adequate, goal-driven assessment of different
 experiments. Indeed, due to the nature of the experiment and fraction
 collection, the quantitative channels (loosely referred to as
-fractions) do not correspond to identical fractions along the
-gradient. As we see in the table below (taken from the `hyperLOPIT`'s
-`pData`), different sets of fractions are combined to obtain enough
-material and optimise acurate quantitation.
+fractions) do not correspond to identical actual fractions along the
+gradient. As can be seen in the table below (taken from the
+`hyperLOPIT`'s `pData`, focusing on channels 7 to 10), different sets
+of fractions are combined to obtain enough material and optimise
+acurate quantitation.
 
@@
 
-On the figure below, we compare the relative intensities of fractions
-7 to 10 on scatter plots, highlighting different marker sets, in
-particular mitochondrion and chromatin. These differences result from
-different peak fractions for these niches.
+On the figure below, we compare the relative intensities of channels 7
+to 10 on scatter plots, highlighting different marker sets. These
+differences result from different peak fractions for these niches on
+the mitochondrial and chromatin occupancy profiles for the two
+replicates.
 
@@
     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
-         main = paste("Fraction", pData(hyperLOPIT)[i, "Fraction.No"]))
+         main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
@@
 }
-plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "n")
+plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
 addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
@@
 
-The relevant comparison unit is not a single fraction, but rather
-complete protein occupancy profiles, which are best visualised as a
-whole on a PCA plot. As such, we prefer to focus on the direct
-comparison of individual replicate PCA plots, assuring that each
-displays acceptable sub-cellular resolution. Note that in the code
-chunk below, we mirror the y-axis to represent the two figures with
-the same orientation.
+The more relevant comparison unit is not a single fraction, but rather
+the complete protein occupancy profiles, which are best visualised as
+experiment-wide on a PCA plot. As such, we prefer to focus on the
+direct, qualitative comparison of individual replicate PCA plots,
+assuring that each displays acceptable sub-cellular resolution. Note
+that in the code chunk below, we mirror the y-axis to represent the
+two figures with the same orientation.
 

### 2016-06-11 bioc-workflow.Rmd
fix a few typos

@@
 an improved method relying on improved sub-cellular fractionation and
-more acurate quantitation, leading to more reliable classification of
+more accurate quantitation, leading to more reliable classification of
 protein localisation across the whole sub-cellular space. The method
@@
 relevant for any quantitative spatial proteomics data, irrespective of
-the fractionation or quantitation (i.e. labeled or label-free)
+the fractionation or quantitation (i.e. labelled or label-free)
 methods.
@@
 which corresponds the the file of interest. In the last line, we print
-the filename (not the full path, which will vary from computer to
+the file name (not the full path, which will vary from computer to
 computer).
@@
 to 27. We now have the two mandatory arguments to `readMSnSet2`,
-namely the filename (stored in the `csvfile` variable) and the
+namely the file name (stored in the `csvfile` variable) and the
 quantitation column indices. In addition to these, it is also possible
@@
 as the labels by which to identify each protein in the sample. Here,
-we use `fnames = 1` to use the Uniprot identifiers contained in the
+we use `fnames = 1` to use the UniProt identifiers contained in the
 first (unnamed) column of the spreadsheet. We also need to specify to
@@
 For demonstration in the code chunk below keep the 2nd, 8th and 11th
-columns which contain the Uniprot entry names and two different marker
+columns which contain the UniProt entry names and two different marker
 sets to use an input for machine learning analyses (see sections on
@@
 This transformation of the data assures to cancel the effect of the
-absolute intensities of the quantitive features along the rows, to
+absolute intensities of the quantitative features along the rows, to
 focus subsequent analyses on the relative profiles along the
@@
 above. Different normalisation methods such as mean or median scaling,
-variance stabilitation or quantile normalisation, to cite a few, can
+variance stabilisation or quantile normalisation, to cite a few, can
 be applied to accomodation different needs.

### 2016-06-12 bioc-workflow.Rmd
some changes to QC section

@@
 
-Data quality is routinely examined through visualisation to verify that sub-cellular niches have been separated along the gradient.
-Based on De Duve's principle [@DeDuve:1981] proteins that co-localise exhibit similar quantitation profiles across the gradient fractions employed.
-One approach that has been widely used to visualise and inspect high throughput mass spectrometry-based proteomics data is principal components analysis (PCA).
-PCA is one of many dimensionality reduction methods, that allow one to effectively summarise multi-dimensional data in to 2 or 3 dimensions to enable visualisation.
-Very generally, the original continuous multi-dimensional data is transformed into a set of orthogonal components ordered according to the amount of variability that they describe.
-The `plot2D` method in `r Biocpkg("pRoloc")` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the x- and y-axis, respectively (the `dims` argument can be used to plot other PCs).
-If distinct clusters are observed, we assume that there is organellar separation present in the data.
-In the code chunk below we produce a PCA plot of the mouse stem cell dataset.
-One point on the plot represents one protein.
-We can indeed see several distinct protein clusters.
-Although, plotting the PCs does not give us a hard quantitative measure of separation, it is extremely useful summarising complex experimental information in one figure, to get an simplified overview of the data.
+Data quality is routinely examined through visualisation to verify
+that sub-cellular niches have been separated along the gradient. Based
+on De Duve's principle [@DeDuve:1981] proteins that co-localise
+exhibit similar quantitation profiles across the gradient fractions
+employed. One approach that has been widely used to visualise and
+inspect high throughput mass spectrometry-based proteomics data is
+principal components analysis (PCA). PCA is one of many dimensionality
+reduction methods, that allow one to effectively summarise
+multi-dimensional data in to 2 or 3 dimensions to enable
+visualisation. Very generally, the original continuous
+multi-dimensional data is transformed into a set of orthogonal
+components ordered according to the amount of variability that they
+describe. The `plot2D` method in `r Biocpkg("pRoloc")` allows one to
+plot the principal components (PCs) of a dataset against one another,
+by default the first two components are plotted on the x- and y-axis,
+respectively (the `dims` argument can be used to plot other PCs). If
+distinct clusters are observed, we assume that there is organellar
+separation present in the data. In the code chunk below we produce a
+PCA plot of the mouse stem cell dataset. One point on the plot
+represents one protein. We can indeed see several distinct protein
+clusters. Although, representing the multi-dimensional data along a
+limited set of PCs does not give us a hard quantitative measure of
+separation, it is extremely useful summarising complex experimental
+information in one figure, to get an simplified overview of the data.
+
+In the code chunk below, we specify `fcol = NULL`, which means not to
+consider any feature variable to annotate the features with
+colours. We will see later how to use this to annotate the PCA plot
+with prior information about sub-cellular markers.
 
@@
 
+We advise to visualise the without any annotation (i.e. with `fcol =
+NULL`) in a first place, before proceeding with data annotation. The
+identification of well resolved clusters in the data, constitutes an
+unbiased assessment of the data structure, demonstrating the
+successful separation of sub-cellular clusters.
+
+There are other dimensionality reduction methods available in the
+`plot2D` function, which can be parametrised with the `method`
+argument.
+
 **TODO** Mention that there are other dimensionality reduction
@@
 [this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
-very interesting gist with intriguing tSNE examples.
+very interesting gist with intriguing tSNE examples. This might be
+better to come after markers.
 
@@
 # Interactive visualisation
+
+A section about  `r Biocpkg("pRolocGUI")`.
 

### 2016-06-13 bioc-workflow.Rmd
replication: what about correlation? ; fix also bug in code

@@
 assuring that each displays acceptable sub-cellular resolution. Note
-that in the code chunk below, we mirror the y-axis to represent the
+that in the code chunk below, we mirror the x-axis to represent the
 two figures with the same orientation.
@@
 par(mfrow = c(1, 2))
-plot2D(hl[hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
-plot2D(hl[hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
-       mirrorY = TRUE)
+plot2D(hl[, hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
+plot2D(hl[, hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
+       mirrorX = TRUE)
 ```
+
+Peason correlation
+
+```{r}
+i1 <- exprs(hl[, hl$replicate == 1])
+i2 <- exprs(hl[, hl$replicate == 2])
+pearcor <- cor(t(i1), t(i2))
+mean(diag(pearcor))
+```
+
+- What about classification on different replicates?
 

### 2016-06-14 bioc-workflow.Rmd
plotDist example

@@
 
-**TODO** Mention that there are other dimensionality reduction
-techniques. Advantages, disadvantages? See
+**TODO** Advantages, disadvantages? See
 [this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
@@
 
-In the context of spatial proteomics, a marker protein is defined as a well-known resident of a specific sub-cellular niche in a species *and* condition of interest.
-Applying this to machine learning (ML), and specifically supervised learning, for the task of protein localisation prediction, markers constitute the labelled training data to use as input to a classification analyses.
-Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but it is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
-`r Biocpkg("pRoloc")` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object, as demonstrated in the code chunk below.
-These marker sets can be accessed using the `pRolocmarkers()` function.
-Marker sets are stored as a simple named vector in R, and originate from in-house user-defined spreadsheets or a set of markers from previous published studies.
-The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
-The mouse dataset used here has Uniprot IDs stored as the `featureNames` (see `head(featureNames(lopit2016))`) and the names of the vector of the mouse markers (`mmus` markers) are also Uniprot IDs (see `head(mrk)` in the code chunk below), so it is straightforward to match names between the markers and the `MSnSet` instance.
-If the naming between the marker sets and the `MSnSet` dataset are different, one will have to convert and match the proteins according to the appropriate identifier.
-Sometimes, we find the equivalent entry name, Uniprot ID or accession number is stored with the data, which makes conversion between identifers relatively straightforward.
-If this is not the case however, there are conversion softwares online available, for example XXX.
+In the context of spatial proteomics, a marker protein is defined as a
+well-known resident of a specific sub-cellular niche in a species
+*and* condition of interest. Applying this to machine learning (ML),
+and specifically supervised learning, for the task of protein
+localisation prediction, markers constitute the labelled training data
+to use as input to a classification analyses. Defining well-known
+residents, and obtaining labelled training data for ML analyses can be
+time consuming, but it is important to define markers that are
+representative of the multivariate data space and on which a
+classifier will be trained and generated. `r Biocpkg("pRoloc")`
+provides a convenience function, `addMarkers`, to directly add markers
+to a `MSnSet` object, as demonstrated in the code chunk below. These
+marker sets can be accessed using the `pRolocmarkers()`
+function. Marker sets are stored as a simple named vector in R, and
+originate from in-house user-defined sets of markers or from previous
+published studies [@Gatto:2014b].
 
-In the code chunk below, we demonstrate how to add markers using `pRolocmarkers` function and then visualise these annotations using the `plot2D` function.
+```{r markers}
+## List available marker sets
+pRolocmarkers()
+```
+
+These markers can then be mapped to an `MSnSet`'s `featureNames`. The
+mouse dataset used here has Uniprot IDs stored as the `featureNames`
+(see `head(featureNames(lopit2016))`) and the names of the vector of
+the mouse markers (`mmus` markers) are also Uniprot IDs (see
+`head(mrk)` in the code chunk below), so it is straightforward to
+match names between the markers and the `MSnSet` instance using the
+`addMarkers` function.
 
 ```{r addmrkers}
-## List available marker sets
-pRolocmarkers()
-
 ## Use mouse markers
@@
 hl <- addMarkers(hl, mrk)
+```
 
+If the naming between the marker sets and the `MSnSet` dataset are
+different, one will have to convert and match the proteins according
+to the appropriate identifier. Sometimes, we find the equivalent entry
+name, Uniprot ID or accession number is stored with the data, which
+makes conversion between identifers relatively straightforward. If
+this is not the case however, conversion can be performed using the 
+`r Biocpkg("biomaRt")`, the Bioconductor
+[annotation resouces](http://bioconductor.org/help/workflows/annotation/Annotation_Resources/)
+or any conversion softwares available online.
+
+We now visualise these annotations along the PCA plot using the
+`plot2D` function. We also use the `addLegend` function to map the
+marker classes to the pre-defined colours.
+
+```{r plotmarkers}
 ## Plot mouse markers
 plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
+addLegend(hl, fcol = "markers", where = "bottomleft", cex = .7)
 ```
 
-In general, the Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are a good starting point for protein annotation and marker definition.
-It is important to note however that automatic retrieval of sub-cellular localisation information, from `r Biocpkg("pRoloc")` or elsewhere, is only the beginning in defining a marker set for downstream analyses.
-Expert curation is vital to check that any annotation added is in the correct context for the the biological question under investigation.
+The colours have been defined so as to enable to differenciate up to
+30 classes. If more are provided, different character symbols
+(circles, squares, ... and empty and solid symbols) are used. The
+colours and the default plotting characters (solid dots for the
+markers and empty circles for the features of unknown localisation)
+can of course be changed, as described in the `setStockcol` manual
+page.
+
+In general, the Gene Ontology (GO) [@Ashburner:2000], and in
+particular the cellular compartment (CC) namespace are a good starting
+point for protein annotation and marker definition. It is important to
+note however that automatic retrieval of sub-cellular localisation
+information, from `r Biocpkg("pRoloc")` or elsewhere, is only the
+beginning in defining a marker set for downstream analyses. Expert
+curation is vital to check that any annotation added is in the correct
+context for the the biological question under investigation.
+
+Another useful visualisation that relies on marker annotation is the
+representation of the protein profiles occupancy along the gradient
+using the `plotDist` function. While the PCA plot enables to
+efficiently visualise the complete dataset and assess the relative
+separation of different sub-cellular niches, comparing profiles of a
+few marker clusters is useful to assess how exactly they differ (in
+terms of peak fractions, for example). Below, we plot the profile of
+the mitochondrial and peroxisome markers.
+
+```{r plotDist}
+par(mfrow = c(2, 1))
+o <- order(hyperLOPIT$Fraction.No)
+cls <- c(Peroxisome = "darkblue", Mitochondrion = "purple")
+for (k in c("Mitochondrion", "Peroxisome")) {
+    plotDist(hyperLOPIT[fData(hyperLOPIT)$markers == k, o],
+             pcol = cls[k])
+    title(main = k)
+}
+```
 
@@
 pearcor <- cor(t(i1), t(i2))
+pearcor0 <- cor(t(i1[, seq(1, 10, 2)]), t(i2[, seq(1, 10, 2)]))
 mean(diag(pearcor))
+mean(diag(pearcor0))
 ```
+
+```{r}
+data(hyperLOPIT2015ms3r1)
+data(hyperLOPIT2015ms3r2)
+data(hyperLOPIT2015ms3r3)
+
+hll <- MSnSetList(list(hyperLOPIT2015ms3r1,
+                       hyperLOPIT2015ms3r2,
+                       hyperLOPIT2015ms3r3))
+
+ij <- combn(3, 2)
+pc <- apply(ij, 2,
+            function(.ij) {
+                tmp <- suppressMessages(commonFeatureNames(hll[.ij]))
+                diag(cor(t(exprs(tmp[[1]])), t(exprs(tmp[[2]]))))
+            })
+names(pc) <- apply(ij, 2, paste, collapse = ".")
+sapply(pc, summary)
+```
+
+The third replicate was produce over 1 year after the first two by
+another operator.
 

### 2016-06-14 bioc-workflow.Rmd
plot2D with other PCs

@@
 
-There are other dimensionality reduction methods available in the
-`plot2D` function, which can be parametrised with the `method`
-argument.
-
-**TODO** Advantages, disadvantages? See
+**TODO** There are other dimensionality reduction methods available in
+the `plot2D` function, which can be parametrised with the `method`
+argument.  Advantages, disadvantages? See
 [this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
@@
 better to come after markers.
+
+It is also useful to visualise the relative intensities along the
+gradient to identify fractions displaying particularly low yield. This
+can be done using the `plotDist` and `boxplot` functions, that plot
+the protein profiles occupancy along the gradient (we also display the
+mean fraction intensities) and a `boxplot` of the column
+intensities. In the two plots displayed below, we re-order the
+fractions to pair corresponding fractions in the two replicates
+(rather than ordering the fractions by replicate).
+
+```{r qcbx, fig.width = 12}
+par(mfrow = c(1, 2))
+o <- order(hyperLOPIT$Fraction.No)
+plotDist(hyperLOPIT[, o], pcol = "#00000010")
+lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
+boxplot(exprs(hyperLOPIT[, o]))
+```
 
@@
 `plot2D` function. We also use the `addLegend` function to map the
-marker classes to the pre-defined colours.
+marker classes to the pre-defined colours. We also display the data
+along the first and seventh PCs using the `dims` argument. Note that
+in this second call to the `plot2D` function, we have omitted the
+`fcol` argument, as the `"markers"` feature variable name is the
+default value. We choose to display PCs 1 and 7 to illustrate that
+while upper principal components explain much less variability in the
+data (2.23% for PC7, as opposed to 48.41% for PC1), we see that the
+mitochondrial and peroxisome clusters can be differenciated, despite
+the apparent overlap in the two first PCs.
 
-```{r plotmarkers}
-## Plot mouse markers
+```{r plotmarkers, fig.width = 12}
+par(mfrow = c(1, 2))
 plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
 addLegend(hl, fcol = "markers", where = "bottomleft", cex = .7)
+plot2D(hl, dims = c(1, 7))
 ```
@@
 ```{r}
+
 data(hyperLOPIT2015ms3r1)
@@
 data(hyperLOPIT2015ms3r3)
+
+hyperLOPIT2015ms3r3 <- hyperLOPIT2015ms3r3[, -(2:4)]
 
@@
                 tmp <- suppressMessages(commonFeatureNames(hll[.ij]))
-                diag(cor(t(exprs(tmp[[1]])), t(exprs(tmp[[2]]))))
+                i <- intersect(tmp[[1]]$Fraction.No, tmp[[2]]$Fraction.No)
+                diag(cor(t(exprs(tmp[[1]][, match(i, tmp[[1]]$Fraction.No)])),
+                         t(exprs(tmp[[2]][, match(i, tmp[[2]]$Fraction.No)]))))
             })
@@
 sapply(pc, summary)
+
+data(hyperLOPIT2015ms3r3)
+tmp <- commonFeatureNames(hyperLOPIT2015ms3r1[, -(2:4)],
+                          hyperLOPIT2015ms3r3[, -(2:4)])
+summary(diag(cor(t(exprs(tmp[[1]])), t(exprs(tmp[[2]])))))
+
+data(hyperLOPIT2015ms3r3)
+tmp <- commonFeatureNames(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r3)
+summary(diag(cor(t(exprs(tmp[[1]])), t(exprs(tmp[[2]])))))
+
 ```

### 2016-06-14 bioc-workflow.Rmd
regenerate plotDist with fixed pRoloc

@@
 addLegend(hl, fcol = "markers", where = "bottomleft", cex = .7)
-plot2D(hl, dims = c(1, 7))
+plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
 ```
@@
 terms of peak fractions, for example). Below, we plot the profile of
-the mitochondrial and peroxisome markers.
+the mitochondrial and peroxisome markers to highlight the differences
+in profiles between these two sets of markers along the 6th and 7th
+fractions, as represented above along the 7th PC on the PCA plot.
 
 ```{r plotDist}
-par(mfrow = c(2, 1))
-o <- order(hyperLOPIT$Fraction.No)
-cls <- c(Peroxisome = "darkblue", Mitochondrion = "purple")
-for (k in c("Mitochondrion", "Peroxisome")) {
-    plotDist(hyperLOPIT[fData(hyperLOPIT)$markers == k, o],
-             pcol = cls[k])
-    title(main = k)
-}
+hlo <- hyperLOPIT[, order(hyperLOPIT$Fraction.No)]
+plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
+         pcol = "purple", fractions = "Fraction.No")
+title(main = "Marker occupancy profiles along the gradient")
+matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
+         lty = 1, col = "darkblue", type = "l")
+legend("topleft", c("Mitochondrion", "Peroxisome"),
+       lty = 1, col = c("purple", "blue"), bty = "n")
 ```

### 2016-06-14 bioc-workflow.Rmd
update superscript to mardown format

@@
 of the MS data acquisition technique named synchronous precursor
-selection MS$^3$ (SPS-MS$^3$) [@McAlister:2014], for TMT
+selection MS^3^ (SPS-MS^3^) [@McAlister:2014], for TMT
 quantification with high accuracy and precision. Three biological

### 2016-06-15 bioc-workflow.Rmd
clean up rep section

@@
 
-On the figure below, we compare the relative intensities of channels 7
-to 10 on scatter plots, highlighting different marker sets. These
-differences result from different peak fractions for these niches on
-the mitochondrial and chromatin occupancy profiles for the two
-replicates.
+<!-- On the figure below, we compare the relative intensities of channels 7 -->
+<!-- to 10 on scatter plots, highlighting different marker sets. These -->
+<!-- differences result from different peak fractions for these niches on -->
+<!-- the mitochondrial and chromatin occupancy profiles for the two -->
+<!-- replicates. -->
 
-```{r repl1, eval = TRUE, echo=FALSE, fig.width = 12}
-sel0 <- fData(hyperLOPIT)$markers != "unknown"
-cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
-cls[!sel0] <- getUnknowncol()
+<!-- ```{r repl1, eval = TRUE, echo=FALSE, fig.width = 12} -->
+<!-- sel0 <- fData(hyperLOPIT)$markers != "unknown" -->
+<!-- cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)] -->
+<!-- cls[!sel0] <- getUnknowncol() -->
 
-par(mfrow = c(2, 4))
-for (i in c(7, 8, 10, 9)) {
-    plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
-         main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
-    points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
-    grid()
-    abline(0, 1, col = "red")
-}
-plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
-addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
+<!-- par(mfrow = c(2, 4)) -->
+<!-- for (i in c(7, 8, 10, 9)) { -->
+<!--     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls, -->
+<!--          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"])) -->
+<!--     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19) -->
+<!--     grid() -->
+<!--     abline(0, 1, col = "red") -->
+<!-- } -->
+<!-- plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "") -->
+<!-- addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7) -->
 
-i <- fData(hyperLOPIT)$markers == "Mitochondrion"
-plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
-         featureNames(hyperLOPIT)[i], type = "l",
-         mcol = "orange")
-title(main = "Mitochondrion")
-## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
-matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
-legend("topleft", paste("Replicate", 1:2),
-       col = c("#F781BF", "orange"),
-       lty = 1, bty = "n")
+<!-- i <- fData(hyperLOPIT)$markers == "Mitochondrion" -->
+<!-- plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2], -->
+<!--          featureNames(hyperLOPIT)[i], type = "l", -->
+<!--          mcol = "orange") -->
+<!-- title(main = "Mitochondrion") -->
+<!-- ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2]) -->
+<!-- matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1) -->
+<!-- legend("topleft", paste("Replicate", 1:2), -->
+<!--        col = c("#F781BF", "orange"), -->
+<!--        lty = 1, bty = "n") -->
 
-i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
-plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
-         featureNames(hyperLOPIT)[i], type = "l")
-## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
-matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
-title(main = "Chromatin")
-legend("topleft", paste("Replicate", 1:2),
-       col = c("#984EA3", "steelblue"),
-       lty = 1, bty = "n")
-```
+<!-- i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin" -->
+<!-- plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2], -->
+<!--          featureNames(hyperLOPIT)[i], type = "l") -->
+<!-- ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2]) -->
+<!-- matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1) -->
+<!-- title(main = "Chromatin") -->
+<!-- legend("topleft", paste("Replicate", 1:2), -->
+<!--        col = c("#984EA3", "steelblue"), -->
+<!--        lty = 1, bty = "n") -->
+<!-- ``` -->
 
@@
 ```
-
-```{r}
-
-data(hyperLOPIT2015ms3r1)
-data(hyperLOPIT2015ms3r2)
-data(hyperLOPIT2015ms3r3)
-
-hyperLOPIT2015ms3r3 <- hyperLOPIT2015ms3r3[, -(2:4)]
-
-hll <- MSnSetList(list(hyperLOPIT2015ms3r1,
-                       hyperLOPIT2015ms3r2,
-                       hyperLOPIT2015ms3r3))
-
-ij <- combn(3, 2)
-pc <- apply(ij, 2,
-            function(.ij) {
-                tmp <- suppressMessages(commonFeatureNames(hll[.ij]))
-                i <- intersect(tmp[[1]]$Fraction.No, tmp[[2]]$Fraction.No)
-                diag(cor(t(exprs(tmp[[1]][, match(i, tmp[[1]]$Fraction.No)])),
-                         t(exprs(tmp[[2]][, match(i, tmp[[2]]$Fraction.No)]))))
-            })
-names(pc) <- apply(ij, 2, paste, collapse = ".")
-sapply(pc, summary)
-
-data(hyperLOPIT2015ms3r3)
-tmp <- commonFeatureNames(hyperLOPIT2015ms3r1[, -(2:4)],
-                          hyperLOPIT2015ms3r3[, -(2:4)])
-summary(diag(cor(t(exprs(tmp[[1]])), t(exprs(tmp[[2]])))))
-
-data(hyperLOPIT2015ms3r3)
-tmp <- commonFeatureNames(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r3)
-summary(diag(cor(t(exprs(tmp[[1]])), t(exprs(tmp[[2]])))))
-
-```
-
-The third replicate was produce over 1 year after the first two by
-another operator.
 

### 2016-06-15 bioc-workflow.Rmd
Lisa's updates

@@
 library("BiocStyle")
-suppressPackageStartupMessages(library("knitr"))
 suppressPackageStartupMessages(library("gridExtra"))
@@
 an improved method relying on improved sub-cellular fractionation and
-more accurate quantitation, leading to more reliable classification of
+more acurate quantitation, leading to more reliable classification of
 protein localisation across the whole sub-cellular space. The method
@@
 of the MS data acquisition technique named synchronous precursor
-selection MS^3^ (SPS-MS^3^) [@McAlister:2014], for TMT
+selection MS$^3$ (SPS-MS$^3$) [@McAlister:2014], for TMT
 quantification with high accuracy and precision. Three biological
@@
 relevant for any quantitative spatial proteomics data, irrespective of
-the fractionation or quantitation (i.e. labelled or label-free)
+the fractionation or quantitation (i.e. labeled or label-free)
 methods.
@@
 which corresponds the the file of interest. In the last line, we print
-the file name (not the full path, which will vary from computer to
+the filename (not the full path, which will vary from computer to
 computer).
@@
 to 27. We now have the two mandatory arguments to `readMSnSet2`,
-namely the file name (stored in the `csvfile` variable) and the
+namely the filename (stored in the `csvfile` variable) and the
 quantitation column indices. In addition to these, it is also possible
@@
 as the labels by which to identify each protein in the sample. Here,
-we use `fnames = 1` to use the UniProt identifiers contained in the
+we use `fnames = 1` to use the Uniprot identifiers contained in the
 first (unnamed) column of the spreadsheet. We also need to specify to
@@
 For demonstration in the code chunk below keep the 2nd, 8th and 11th
-columns which contain the UniProt entry names and two different marker
+columns which contain the Uniprot entry names and two different marker
 sets to use an input for machine learning analyses (see sections on
@@
 This transformation of the data assures to cancel the effect of the
-absolute intensities of the quantitative features along the rows, to
+absolute intensities of the quantitive features along the rows, to
 focus subsequent analyses on the relative profiles along the
@@
 above. Different normalisation methods such as mean or median scaling,
-variance stabilisation or quantile normalisation, to cite a few, can
+variance stabilitation or quantile normalisation, to cite a few, can
 be applied to accomodation different needs.
@@
 
-When more than 2 data are to be combine and/or too many proteins were
-not consistently assayed, leading to too many proteins being filtered
-out, we suggest to implement an ensemble of classifiers voting on
-protein-sub-cellular niche membership over the output of several
-experiments (see section *Supervised machine learning* for the
-description of sub-cellular assingments).
+## Replication
+
+With the aim of maximising the sub-cellular resolution and,
+consequently, the reliability in protein sub-cellular assignments, we
+follow the advice in [@Trotter:2010] and combine replicated spatial
+proteomics experiments and described above. Trotter et al. have shown
+that ...
+
+Why is the below no a good indication: fraction, even if same index,
+are not necessarily same; replication of single fractions fractions is
+not really relevant, but rather complete profiles; more than single
+profiles, it's rather relative profiles in the frame of whole dataset.
+
+```{r eval = FALSE}
+par(mfrow = c(3, 4))
+for (i in 1:10) {
+    plot(exprs(hl[, c(i, i+10)]))
+    grid()
+    abline(0, 1, col = "red")
+}
+```
+
+```{r plot2Drep, fig.width = 12}
+par(mfrow = c(1, 2))
+plot2D(hl[hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
+plot2D(hl[hl$replicate == 2], fcol = "SVM.marker.set", mirrorY = TRUE, main = "Replicate 2")
+```
+
+Could compute squared differences between fractions over profile for
+each protein, and compare to Dunkley 2006, to show improvement in
+replicability. Or to have an internal replication measurement. Or what
+about wave?
+
+This might be better off in the QC section...
 
@@
 
-Data quality is routinely examined through visualisation to verify
-that sub-cellular niches have been separated along the gradient. Based
-on De Duve's principle [@DeDuve:1981] proteins that co-localise
-exhibit similar quantitation profiles across the gradient fractions
-employed. One approach that has been widely used to visualise and
-inspect high throughput mass spectrometry-based proteomics data is
-principal components analysis (PCA). PCA is one of many dimensionality
-reduction methods, that allow one to effectively summarise
-multi-dimensional data in to 2 or 3 dimensions to enable
-visualisation. Very generally, the original continuous
-multi-dimensional data is transformed into a set of orthogonal
-components ordered according to the amount of variability that they
-describe. The `plot2D` method in `r Biocpkg("pRoloc")` allows one to
-plot the principal components (PCs) of a dataset against one another,
-by default the first two components are plotted on the x- and y-axis,
-respectively (the `dims` argument can be used to plot other PCs). If
-distinct clusters are observed, we assume that there is organellar
-separation present in the data. In the code chunk below we produce a
-PCA plot of the mouse stem cell dataset. One point on the plot
-represents one protein. We can indeed see several distinct protein
-clusters. Although, representing the multi-dimensional data along a
-limited set of PCs does not give us a hard quantitative measure of
-separation, it is extremely useful summarising complex experimental
-information in one figure, to get an simplified overview of the data.
-
-In the code chunk below, we specify `fcol = NULL`, which means not to
-consider any feature variable to annotate the features with
-colours. We will see later how to use this to annotate the PCA plot
-with prior information about sub-cellular markers.
+Data quality is routinely examined through visualisation to verify that sub-cellular niches have been separated along the gradient.
+Based on De Duve's principle [@DeDuve:1981] proteins that co-localise exhibit similar quantitation profiles across the gradient fractions employed.
+One approach that has been widely used to visualise and inspect high throughput mass spectrometry-based proteomics data is principal components analysis (PCA).
+PCA is one of many dimensionality reduction methods, that allow one to effectively summarise multi-dimensional data in to 2 or 3 dimensions to enable visualisation.
+Very generally, the original continuous multi-dimensional data is transformed into a set of orthogonal components ordered according to the amount of variability that they describe.
+The `plot2D` method in `r Biocpkg("pRoloc")` allows one to plot the principal components (PCs) of a dataset against one another, by default the first two components are plotted on the x- and y-axis, respectively (the `dims` argument can be used to plot other PCs).
+If distinct clusters are observed, we assume that there is organellar separation present in the data.
+In the code chunk below we produce a PCA plot of the mouse stem cell dataset.
+One point on the plot represents one protein.
+We can indeed see several distinct protein clusters.
+Although, plotting the PCs does not give us a hard quantitative measure of separation, it is extremely useful summarising complex experimental information in one figure, to get an simplified overview of the data.
 
@@
 
-We advise to visualise the without any annotation (i.e. with `fcol =
-NULL`) in a first place, before proceeding with data annotation. The
-identification of well resolved clusters in the data, constitutes an
-unbiased assessment of the data structure, demonstrating the
-successful separation of sub-cellular clusters.
-
-**TODO** There are other dimensionality reduction methods available in
-the `plot2D` function, which can be parametrised with the `method`
-argument.  Advantages, disadvantages? See
+**TODO** Mention that there are other dimensionality reduction
+techniques. Advantages, disadvantages? See
 [this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
-very interesting gist with intriguing tSNE examples. This might be
-better to come after markers.
-
-It is also useful to visualise the relative intensities along the
-gradient to identify fractions displaying particularly low yield. This
-can be done using the `plotDist` and `boxplot` functions, that plot
-the protein profiles occupancy along the gradient (we also display the
-mean fraction intensities) and a `boxplot` of the column
-intensities. In the two plots displayed below, we re-order the
-fractions to pair corresponding fractions in the two replicates
-(rather than ordering the fractions by replicate).
-
-```{r qcbx, fig.width = 12}
-par(mfrow = c(1, 2))
-o <- order(hyperLOPIT$Fraction.No)
-plotDist(hyperLOPIT[, o], pcol = "#00000010")
-lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
-boxplot(exprs(hyperLOPIT[, o]))
-```
+very interesting gist with intriguing tSNE examples.
 
@@
 
-In the context of spatial proteomics, a marker protein is defined as a
-well-known resident of a specific sub-cellular niche in a species
-*and* condition of interest. Applying this to machine learning (ML),
-and specifically supervised learning, for the task of protein
-localisation prediction, markers constitute the labelled training data
-to use as input to a classification analyses. Defining well-known
-residents, and obtaining labelled training data for ML analyses can be
-time consuming, but it is important to define markers that are
-representative of the multivariate data space and on which a
-classifier will be trained and generated. `r Biocpkg("pRoloc")`
-provides a convenience function, `addMarkers`, to directly add markers
-to a `MSnSet` object, as demonstrated in the code chunk below. These
-marker sets can be accessed using the `pRolocmarkers()`
-function. Marker sets are stored as a simple named vector in R, and
-originate from in-house user-defined sets of markers or from previous
-published studies [@Gatto:2014b].
+In the context of spatial proteomics, a marker protein is defined as a well-known resident of a specific sub-cellular niche in a species *and* condition of interest.
+Applying this to machine learning (ML), and specifically supervised learning, for the task of protein localisation prediction, markers constitute the labelled training data to use as input to a classification analyses.
+Defining well-known residents, and obtaining labelled training data for ML analyses can be time consuming, but it is important to define markers that are representative of the multivariate data space and on which a classifier will be trained and generated.
+`r Biocpkg("pRoloc")` provides a convenience function, `addMarkers`, to directly add markers to a `MSnSet` object, as demonstrated in the code chunk below.
+These marker sets can be accessed using the `pRolocmarkers()` function.
+Marker sets are stored as a simple named vector in R, and originate from in-house user-defined spreadsheets or a set of markers from previous published studies.
+The marker vectors that can be accessed from `pRolocmarkers` are named vectors and to enable mapping between the markers and the `MSnSet` instance it is required that the `featureNames` of the `MSnSet` instance match the `names` of the marker.
+The mouse dataset used here has Uniprot Accession Numbers (ACs) stored as the `featureNames` (see `head(featureNames(lopit2016))`) and mouse markers in `pRolocmarkers` are also ACs (see `head(mrk)` in the code chunk below), so it is straightforward to match the protein markers in `pRolocmarkers` to the proteins in the `MSnSet` instance.
+If the naming between the marker sets and the `MSnSet` dataset are different, there are several conversion softwares available online such as the [UniprotKB mapping tool](http://www.uniprot.org/uploadlists/).
 
-```{r markers}
+In the code chunk below, we demonstrate how to add markers using the `pRolocmarkers` function and then visualise these annotations using the `plot2D` function.
+
+```{r addmrkers}
 ## List available marker sets
 pRolocmarkers()
-```
 
-These markers can then be mapped to an `MSnSet`'s `featureNames`. The
-mouse dataset used here has Uniprot IDs stored as the `featureNames`
-(see `head(featureNames(lopit2016))`) and the names of the vector of
-the mouse markers (`mmus` markers) are also Uniprot IDs (see
-`head(mrk)` in the code chunk below), so it is straightforward to
-match names between the markers and the `MSnSet` instance using the
-`addMarkers` function.
-
-```{r addmrkers}
 ## Use mouse markers
@@
 hl <- addMarkers(hl, mrk)
+
+## Plot mouse markers
+plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
 ```
 
-If the naming between the marker sets and the `MSnSet` dataset are
-different, one will have to convert and match the proteins according
-to the appropriate identifier. Sometimes, we find the equivalent entry
-name, Uniprot ID or accession number is stored with the data, which
-makes conversion between identifers relatively straightforward. If
-this is not the case however, conversion can be performed using the 
-`r Biocpkg("biomaRt")`, the Bioconductor
-[annotation resouces](http://bioconductor.org/help/workflows/annotation/Annotation_Resources/)
-or any conversion softwares available online.
-
-We now visualise these annotations along the PCA plot using the
-`plot2D` function. We also use the `addLegend` function to map the
-marker classes to the pre-defined colours. We also display the data
-along the first and seventh PCs using the `dims` argument. Note that
-in this second call to the `plot2D` function, we have omitted the
-`fcol` argument, as the `"markers"` feature variable name is the
-default value. We choose to display PCs 1 and 7 to illustrate that
-while upper principal components explain much less variability in the
-data (2.23% for PC7, as opposed to 48.41% for PC1), we see that the
-mitochondrial and peroxisome clusters can be differenciated, despite
-the apparent overlap in the two first PCs.
-
-```{r plotmarkers, fig.width = 12}
-par(mfrow = c(1, 2))
-plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
-addLegend(hl, fcol = "markers", where = "bottomleft", cex = .7)
-plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
-```
-
-The colours have been defined so as to enable to differenciate up to
-30 classes. If more are provided, different character symbols
-(circles, squares, ... and empty and solid symbols) are used. The
-colours and the default plotting characters (solid dots for the
-markers and empty circles for the features of unknown localisation)
-can of course be changed, as described in the `setStockcol` manual
-page.
-
-In general, the Gene Ontology (GO) [@Ashburner:2000], and in
-particular the cellular compartment (CC) namespace are a good starting
-point for protein annotation and marker definition. It is important to
-note however that automatic retrieval of sub-cellular localisation
-information, from `r Biocpkg("pRoloc")` or elsewhere, is only the
-beginning in defining a marker set for downstream analyses. Expert
-curation is vital to check that any annotation added is in the correct
-context for the the biological question under investigation.
-
-Another useful visualisation that relies on marker annotation is the
-representation of the protein profiles occupancy along the gradient
-using the `plotDist` function. While the PCA plot enables to
-efficiently visualise the complete dataset and assess the relative
-separation of different sub-cellular niches, comparing profiles of a
-few marker clusters is useful to assess how exactly they differ (in
-terms of peak fractions, for example). Below, we plot the profile of
-the mitochondrial and peroxisome markers to highlight the differences
-in profiles between these two sets of markers along the 6th and 7th
-fractions, as represented above along the 7th PC on the PCA plot.
-
-```{r plotDist}
-hlo <- hyperLOPIT[, order(hyperLOPIT$Fraction.No)]
-plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
-         pcol = "purple", fractions = "Fraction.No")
-title(main = "Marker occupancy profiles along the gradient")
-matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
-         lty = 1, col = "darkblue", type = "l")
-legend("topleft", c("Mitochondrion", "Peroxisome"),
-       lty = 1, col = c("purple", "blue"), bty = "n")
-```
-
-# Replication
-
-With the aim of maximising the sub-cellular resolution and,
-consequently, the reliability in protein sub-cellular assignments, we
-follow the advice in [@Trotter:2010] and combine replicated spatial
-proteomics experiments and described above. Indeed, Trotter et
-al. have shown a significant improvement in protein–organelle
-association upon direct combination of single experiments, in
-particular when these resolve different subcellular niches.
-
-Direct comparisons of individual fractions in replicated experiments
-does not provide an adequate, goal-driven assessment of different
-experiments. Indeed, due to the nature of the experiment and fraction
-collection, the quantitative channels (loosely referred to as
-fractions) do not correspond to identical actual fractions along the
-gradient. As can be seen in the table below (taken from the
-`hyperLOPIT`'s `pData`, focusing on channels 7 to 10), different sets
-of fractions are combined to obtain enough material and optimise
-acurate quantitation.
-
-```{r pdtab, echo=FALSE}
-i <- hyperLOPIT$Fraction.No %in% 7:10
-kable(pData(hyperLOPIT)[i, -3], row.names = FALSE)
-```
-
-<!-- On the figure below, we compare the relative intensities of channels 7 -->
-<!-- to 10 on scatter plots, highlighting different marker sets. These -->
-<!-- differences result from different peak fractions for these niches on -->
-<!-- the mitochondrial and chromatin occupancy profiles for the two -->
-<!-- replicates. -->
-
-<!-- ```{r repl1, eval = TRUE, echo=FALSE, fig.width = 12} -->
-<!-- sel0 <- fData(hyperLOPIT)$markers != "unknown" -->
-<!-- cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)] -->
-<!-- cls[!sel0] <- getUnknowncol() -->
-
-<!-- par(mfrow = c(2, 4)) -->
-<!-- for (i in c(7, 8, 10, 9)) { -->
-<!--     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls, -->
-<!--          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"])) -->
-<!--     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19) -->
-<!--     grid() -->
-<!--     abline(0, 1, col = "red") -->
-<!-- } -->
-<!-- plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "") -->
-<!-- addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7) -->
-
-<!-- i <- fData(hyperLOPIT)$markers == "Mitochondrion" -->
-<!-- plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2], -->
-<!--          featureNames(hyperLOPIT)[i], type = "l", -->
-<!--          mcol = "orange") -->
-<!-- title(main = "Mitochondrion") -->
-<!-- ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2]) -->
-<!-- matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1) -->
-<!-- legend("topleft", paste("Replicate", 1:2), -->
-<!--        col = c("#F781BF", "orange"), -->
-<!--        lty = 1, bty = "n") -->
-
-<!-- i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin" -->
-<!-- plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2], -->
-<!--          featureNames(hyperLOPIT)[i], type = "l") -->
-<!-- ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2]) -->
-<!-- matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1) -->
-<!-- title(main = "Chromatin") -->
-<!-- legend("topleft", paste("Replicate", 1:2), -->
-<!--        col = c("#984EA3", "steelblue"), -->
-<!--        lty = 1, bty = "n") -->
-<!-- ``` -->
-
-The more relevant comparison unit is not a single fraction, but rather
-the complete protein occupancy profiles, which are best visualised as
-experiment-wide on a PCA plot. As such, we prefer to focus on the
-direct, qualitative comparison of individual replicate PCA plots,
-assuring that each displays acceptable sub-cellular resolution. Note
-that in the code chunk below, we mirror the x-axis to represent the
-two figures with the same orientation.
-
-```{r plot2Drep, fig.width = 12}
-par(mfrow = c(1, 2))
-plot2D(hl[, hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
-plot2D(hl[, hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
-       mirrorX = TRUE)
-```
-
-Peason correlation
-
-```{r}
-i1 <- exprs(hl[, hl$replicate == 1])
-i2 <- exprs(hl[, hl$replicate == 2])
-pearcor <- cor(t(i1), t(i2))
-pearcor0 <- cor(t(i1[, seq(1, 10, 2)]), t(i2[, seq(1, 10, 2)]))
-mean(diag(pearcor))
-mean(diag(pearcor0))
-```
-
-- What about classification on different replicates?
+In general, the Gene Ontology (GO) [@Ashburner:2000], and in particular the cellular compartment (CC) namespace are a good starting point for protein annotation and marker definition.
+It is important to note however that automatic retrieval of sub-cellular localisation information, from `r Biocpkg("pRoloc")` or elsewhere, is only the beginning in defining a marker set for downstream analyses.
+Expert curation is vital to check that any annotation added is in the correct context for the the biological question under investigation.
 
 # Interactive visualisation
-
-A section about  `r Biocpkg("pRolocGUI")`.
 
@@
 The resultant predictions can be visualised using `plot2D`.
-In the code chunk below `plot2D` is called to generate a PCA plot of the data and `fcol` is used to specify where the new assignments are located, here for example these are located in the column called `svm`.
-Additionally, when calling `plot2D` we use the `cex` argument to change the point size of each point on the plot (where one point represents one protein) to be inversely proportional to the SVM score.
+In the code chunk below `plot2D` is called to generate a PCA plot of the data and `fcol` is used to specify where the new assignments are located e.g.
+`fcol =  "svm"`.
+
+Additionally, when calling `plot2D` we can use the `cex` argument to change the size of each point on the plot (where one point represents one protein) to be inversely proportional to the SVM score.
 This gives an initial overview of the high scoring localisations from the SVM predictions.
@@
 ```{r plotSVM, eval=TRUE}
+## set point size of each protein to be inversely proportional to the SVM score
 ptsze <- exp(fData(svmRes)$svm.scores) - 1
+
+## plot new predictions
 plot2D(svmRes, fcol = "svm", cex = ptsze)
@@
 ## Thresholding
+
+It is common when applying a supervised classification algorithm to set a specific score cutoff on which to define new assignments, below which classifications are set to unknown/unassigned.
+This is because in a supervised learning setup proteins can only be predicted to be localised to one of the sub-cellular niches that appear in the labelled training data.
+We can not guaranette (and do not expect) that the whole class diversity to be represented in the training data as (1) finding markers that represent the whole diversity of the cell is challenging and (2) many sub-cellular niches contain too few proteins to train on (note: we recommend a minimum of 13 markers per sub-cellular class for stratified 80/20 partitioning and cross-validation).
+
+Deciding on a threshold is not trivial as classifier scores are heavily dependent upon the classifier used and different sub-cellular niches can exhibit different score distributions.
+We recommend one to set class-specific score distributions.
+In the code chunk below we display a boxplot of the score distributions per organelle.
+
+```{r thresholds}
+
+```
+
+To help examine these distributions there is also a dedicated interactive application in the `r Biocpkg("pRolocGUI")` package. To access this application:
+
+```{r classifyApp}
+
+```
+
+To launch the classify application:
+
+- Essential for protein localisation as we do not want to force all proteins in to one category (allow for multiple lcoalisations or localisations outside the training data set e.g.
+new novel compartments)
+
+- `getPredictions` function allows you to pass a score threshold. Also `classify` application available to interactively view output of thresholds
 

### 2016-06-22 bioc-workflow.Rmd
point size and thresholding

--- 2016-06-15 bioc-workflow.Rmd
@@
 
+The adjustment of the point size intuitively confers important
+information that is more difficult to defined formally, but that we
+will address in the next section. The classifier (SVM in our case, but
+this is also valid of other classifiers) defines boundaries based on
+the labelled marker proteins. These class/organelle boundaries define
+how non-assigned proteins are classifier and with what confidence.
+
 ## Thresholding

### 2016-06-29 bioc-workflow.Rmd
todo markers and boundaries
more todo
keep quotes consistent
Added Laurent as senior author

@@
 title: "A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data"
-author: "Lisa Breckels"
+author: "Lisa Breckels and Laurent Gatto"
 output: html_document
@@
 ```{r qcplot, message=FALSE, warning=FALSE}
-library('pRoloc')
+library("pRoloc")
 plot2D(hl, fcol = NULL)
@@
 the labelled marker proteins. These class/organelle boundaries define
-how non-assigned proteins are classifier and with what confidence.
+how non-assigned proteins are classifier and with what
+confidence. **TODO** explain how choosing markers
+positively/negatively affects this, and relation with thresholding.
 

### 2016-06-29 bioc-workflow.Rmd
Small updates to the text, typos etc

@@
 
-```{r }
+```{r assesshl}
 exprs(hl)[1:5, ]
@@
 `r Biocexptpkg("pRolocdata")` package. Below, we show how to combine
-`MSnSet` objects and, subsequently, how to filter and hangle missing
+`MSnSet` objects and, subsequently, how to filter and handle missing
 values. We start by loading the experimental data package and the
@@
 (same proteins, different sets of fractions), we need to assure that
-the respective sample names differ.
+the respective sample names differ so they can be identified from one
+another. The function `updateSampleNames` can be used do this.
 
@@
 
-We prefer to remove proteins that were not assayed in our two
+We prefer to remove proteins that were not assayed in both of our two
 replicated experiments. This is done with the `filterNA` function that
@@
 
-Which results in two identical matrices of quantitative protein
-profiles.
-
-```{r equaldata}
-all.equal(exprs(hl), exprs(hyperLOPIT), check.attributes = FALSE)
-```
-
-When more than 2 data are to be combine and/or too many proteins were
+When more than 2 data are to be combined and/or too many proteins were
 not consistently assayed, leading to too many proteins being filtered
@@
 experiments (see section *Supervised machine learning* for the
-description of sub-cellular assingments).
+description of sub-cellular assignments).
 
@@
 In the code chunk below, we specify `fcol = NULL`, which means not to
-consider any feature variable to annotate the features with
+consider any feature variable to annotate the features (proteins) with
 colours. We will see later how to use this to annotate the PCA plot
-with prior information about sub-cellular markers.
+with prior information about sub-cellular localisation.
 
@@
 
-We advise to visualise the without any annotation (i.e. with `fcol =
-NULL`) in a first place, before proceeding with data annotation. The
-identification of well resolved clusters in the data, constitutes an
-unbiased assessment of the data structure, demonstrating the
-successful separation of sub-cellular clusters.
+In the first instance we advise one to visualise their data without 
+any annotation (i.e. with `fcol = NULL`), before proceeding with data 
+annotation. The identification of well resolved clusters in the data, 
+constitutes an unbiased assessment of the data structure, demonstrating 
+the successful separation of sub-cellular clusters.
 
@@
 and specifically supervised learning, for the task of protein
-localisation prediction, markers constitute the labelled training data
-to use as input to a classification analyses. Defining well-known
+localisation prediction, these markers constitute the labelled training
+data to use as input to a classification analyses. Defining well-known
 residents, and obtaining labelled training data for ML analyses can be
@@
 originate from in-house user-defined sets of markers or from previous
-published studies [@Gatto:2014b].
+published studies [@Gatto:2014b], which are continuosly updated and 
+integrated.
 
@@
 (see `head(featureNames(lopit2016))`) and the names of the vector of
-the mouse markers (`mmus` markers) are also Uniprot IDs (see
-`head(mrk)` in the code chunk below), so it is straightforward to
-match names between the markers and the `MSnSet` instance using the
-`addMarkers` function.
+the mouse markers stored in `r Biocpkg("pRoloc")` (`mmus` markers) 
+are also Uniprot IDs (see `head(mrk)` in the code chunk below), so 
+it is straightforward to match names between the markers and 
+the `MSnSet` instance using the `addMarkers` function.
 
@@
 follow the advice in [@Trotter:2010] and combine replicated spatial
-proteomics experiments and described above. Indeed, Trotter et
+proteomics experiments as described above. Indeed, Trotter et
 al. have shown a significant improvement in protein–organelle
@@
 i <- hyperLOPIT$Fraction.No %in% 7:10
-kable(pData(hyperLOPIT)[i, -3], row.names = FALSE)
+pData(hyperLOPIT)[i, -3]
 ```

### 2016-07-04 bioc-workflow.Rmd
Started work on the pRolocGUI section

@@
 
-Peason correlation
+## Peason correlation
 
@@
 
-A section about  `r Biocpkg("pRolocGUI")`.
+Visualisation and data exploration is an important aspect of data analyses allowing one to shed light on data structure and patterns of interest.
+Using the `r Biocpkg("pRolocGUI")` package we can interactively visualise, explore and interrogate quantitative spatial proteomics data.
+The `r Biocpkg("pRolocGUI")` package is currently under active development and it relies on the `shiny` framework for reactivity and interactivity.
+The package currently distributes 3 different GUI's ("main" (default), "compare" or "compare") which are wrapped and launched by the `pRolocVis` function.
+
+The "main" application is designed for exploratory data analysis and is divied into 3 tabs: (1) PCA, (2) Profiles and (3) Table selection.
+The default view upon loading is the PCA tab, which features a clickable interface and zoomable PCA plot with an interactive data table for displaying the quantitation information.
+Particular proteins of interest can be highlighted using the text search box.
+There is also an alternate profiles tab for visualisation of the protein profiles, which can be used to examine the patterns of proteins of interest.
+The Table selection tab provides an interface to control data table column selection.
+In the below code chunk we lauch the main app.
+
+```{r mainapp, eval = FALSE, echo = TRUE}
+library("pRolocGUI")
+pRolocVis(hl)
+```
+
+![A screen shot of clickable interface and zoomable PCA plot of the main app in the `r Biocpkg("pRolocGUI")` package.](./Figures/mainapp.png)
+
+The "compare" application is useful for examining two replicate experiments, or two experiments from different conditions, treatments etc.
+Calling the argument `app = "compare"` as per the below code chunk launches the app wherein two PCA plots are loaded side-by-side.
+Common proteins between the two data sets can be searched for, identified and highlighted.
+As per the main application there is also a protein profiles tab to allow one to look at the patterns of protein profiles of interest in each dataset.
+One key difference between the main and compare app when launching the app is that the compare app requires a `MSnSetList` as input to the `pRolocVis` function.
+For example, in the code chunk below we first create a `MSnSetList` of replicates 1 and 2 of the hyperLOPIT data.
+
+```{r compareapp, eval = FALSE, echo = TRUE}
+data(hyperLOPIT2015ms3r1)
+data(hyperLOPIT2015ms3r2)
+mydata <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
+pRolocVis(mydata, app = "compare")
+```
+
+![The compare application, main panel](./Figures/SS_Compare1.jpg)
+
+The final application "classify", has been designed to view machine learning classification results according to user-specified thresholds for the assignment of proteins to its sub-cellular location.
 

### 2016-07-05 bioc-workflow.Rmd
Section on thresholding

@@
 library("pRoloc")
-plot2D(hl, fcol = NULL)
+plot2D(hyperLOPIT, fcol = NULL)
 ```
@@
 lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
-boxplot(exprs(hyperLOPIT[, o]))
+boxplot(exprs(hyperLOPIT[, o]), las = 2)
 ```
@@
 
-Visualisation and data exploration is an important aspect of data analyses allowing one to shed light on data structure and patterns of interest.
-Using the `r Biocpkg("pRolocGUI")` package we can interactively visualise, explore and interrogate quantitative spatial proteomics data.
-The `r Biocpkg("pRolocGUI")` package is currently under active development and it relies on the `shiny` framework for reactivity and interactivity.
-The package currently distributes 3 different GUI's ("main" (default), "compare" or "compare") which are wrapped and launched by the `pRolocVis` function.
-
-The "main" application is designed for exploratory data analysis and is divied into 3 tabs: (1) PCA, (2) Profiles and (3) Table selection.
-The default view upon loading is the PCA tab, which features a clickable interface and zoomable PCA plot with an interactive data table for displaying the quantitation information.
-Particular proteins of interest can be highlighted using the text search box.
-There is also an alternate profiles tab for visualisation of the protein profiles, which can be used to examine the patterns of proteins of interest.
-The Table selection tab provides an interface to control data table column selection.
-In the below code chunk we lauch the main app.
+Visualisation and data exploration is an important aspect of data
+analyses allowing one to shed light on data structure and patterns of
+interest. Using the `r Biocpkg("pRolocGUI")` package we can
+interactively visualise, explore and interrogate quantitative spatial
+proteomics data. The `r Biocpkg("pRolocGUI")` package is currently
+under active development and it relies on the `shiny` framework for
+reactivity and interactivity. The package currently distributes 3
+different GUI's ("main" (default), "compare" or "compare") which are
+wrapped and launched by the `pRolocVis` function. In the below code
+chunk we lauch the main app (note, we do not need to specify the
+argument, `app = "main"` as it is the default).
 
@@
 
-![A screen shot of clickable interface and zoomable PCA plot of the main app in the `r Biocpkg("pRolocGUI")` package.](./Figures/mainapp.png)
+![A screen shot of clickable interface and zoomable PCA plot of the
+main app in the `r Biocpkg("pRolocGUI")`
+package.](./Figures/mainapp.png)
 
-The "compare" application is useful for examining two replicate experiments, or two experiments from different conditions, treatments etc.
-Calling the argument `app = "compare"` as per the below code chunk launches the app wherein two PCA plots are loaded side-by-side.
-Common proteins between the two data sets can be searched for, identified and highlighted.
-As per the main application there is also a protein profiles tab to allow one to look at the patterns of protein profiles of interest in each dataset.
-One key difference between the main and compare app when launching the app is that the compare app requires a `MSnSetList` as input to the `pRolocVis` function.
-For example, in the code chunk below we first create a `MSnSetList` of replicates 1 and 2 of the hyperLOPIT data.
+As diplayed in the screenshot above, the "main" application is
+designed for exploratory data analysis and is divied into 3 tabs: (1)
+PCA, (2) Profiles and (3) Table selection. The default view upon
+loading is the PCA tab, which features a clickable interface and
+zoomable PCA plot with an interactive data table for displaying the
+quantitation information. Particular proteins of interest can be
+highlighted using the text search box. There is also an alternate
+profiles tab for visualisation of the protein profiles, which can be
+used to examine the patterns of proteins of interest. The Table
+selection tab provides an interface to control data table column
+selection.
+
+The "compare" application is useful for examining two replicate
+experiments, or two experiments from different conditions, treatments
+etc. The compare application is called by default if the input object
+to `pRolocVis` is a `MSnSetList` of 2 `MSnSets`, but it can also be
+specified by calling the argument `app = "compare"`. For example, in
+the code chunk below we first create a `MSnSetList` of replicates 1
+and 2 of the hyperLOPIT data, this is then passed to `pRolocVis`.
 
@@
 
-The final application "classify", has been designed to view machine learning classification results according to user-specified thresholds for the assignment of proteins to its sub-cellular location.
+The comparison app loads the two PCA plots side-by-side. Only common
+proteins between the two data sets are displayed. As per the main
+application proteins can be searched, identified and highlighted on
+both PCA plots and in the dedicated profiles tab. One key feature of
+the compare application is the ability to re-map the second dataset
+onto the PCA data space of the first (reference) data set (see
+`?pRolocVis` and the argument `remap = TRUE`). Using the first dataset
+as the reference set, PCA is carried out on the first dataset and the
+standard deviations of the principal components (i.e. the square roots
+of the eigenvalues of the covariance/correlation matrix) and the
+matrix of variable loadings (i.e. a matrix whose columns contain the
+eigenvectors) are stored and then used to calculate the principal
+components of the second dataset. Both datasets are scaled and
+centered in the usual way. The first dataset appears on the left, and
+the second re-mapped data appears on the right. The order of the first
+(the reference data for remapping) and second dataset can be changed
+through regeneration/re-ordering of the `MSnSetList` object.
+
+The final application "classify", has been designed to view machine
+learning classification results according to user-specified thresholds
+for the assignment of proteins to its sub-cellular location, this is
+discussed later in the subsection [thresholding](#thresholding) in the
+[supervised machine learning section](#supervised-machine-learning).
 
@@
 
-The extraction of sub-cellular protein clusters can be difficult owing to the limited number of marker proteins that exist in databases and elsewhere.
-Furthermore, given the vast complexity of the cell, automatic annotation retrieval does not always give a full representation of the true sub-cellular diversity in the data.
-For downstream analyses, such as supervised machine learning, it is desirable to obtain reliable markers that cover as many sub-cellular niches as possible, as these markers are directly used in the training phase of the ML classification.
-We find that a lack of sub-cellular diversity in the labelled training data leads to prediction errors, as unlabelled instances can only be assigned to a class that exists in the training data [@Breckels:2013].
-In such scenarios novelty detection can be useful to identify data-specific sub-cellular groupings such as organelles and protein complexes.
-The phenotype discovery (phenoDisco) algorithm [@Breckels:2013] is one such method and is available in `r Biocpkg("pRoloc")`.
-It is an iterative semi-supervised learning method that combines the classification of proteins on existing labelled data with the detection of new clusters.
-Novelty detection methods are also useful for confirming the presence of known clusters in an unbiased fashion.
-For example, in [@hyper] the `phenoDisco` algorithm was used to detect the data-specific confirmation and presence of the nucleus and nucleus sub-compartments, as it was expected that nucleus associated clusters existed within the data as the protocol included a separate chromatin-rich fraction to enrich for nuclei.
-In the code chunk below, we show how to run a series of novelty detection experiments using the `phenoDisco` algorithm on the mouse stem-cell dataset.
-There are several optional arguments that can be set dependent on the type of biological question of interest.
-For example,
+The extraction of sub-cellular protein clusters can be difficult owing
+to the limited number of marker proteins that exist in databases and
+elsewhere. Furthermore, given the vast complexity of the cell,
+automatic annotation retrieval does not always give a full
+representation of the true sub-cellular diversity in the data. For
+downstream analyses, such as supervised machine learning, it is
+desirable to obtain reliable markers that cover as many sub-cellular
+niches as possible, as these markers are directly used in the training
+phase of the ML classification. We find that a lack of sub-cellular
+diversity in the labelled training data leads to prediction errors, as
+unlabelled instances can only be assigned to a class that exists in
+the training data [@Breckels:2013]. In such scenarios novelty
+detection can be useful to identify data-specific sub-cellular
+groupings such as organelles and protein complexes. The phenotype
+discovery (phenoDisco) algorithm [@Breckels:2013] is one such method
+and is available in `r Biocpkg("pRoloc")`. It is an iterative
+semi-supervised learning method that combines the classification of
+proteins on existing labelled data with the detection of new clusters.
+Novelty detection methods are also useful for confirming the presence
+of known clusters in an unbiased fashion. For example, in [@hyper] the
+`phenoDisco` algorithm was used to detect the data-specific
+confirmation and presence of the nucleus and nucleus sub-compartments,
+as it was expected that nucleus associated clusters existed within the
+data as the protocol included a separate chromatin-rich fraction to
+enrich for nuclei. In the code chunk below, we show how to run a
+series of novelty detection experiments using the `phenoDisco`
+algorithm on the mouse stem-cell dataset. There are several optional
+arguments that can be set dependent on the type of biological question
+of interest. For example,
 
@@
 
-Supervised machine learning, also known as classification, is an essential tool for the assignment of proteins to distinct sub-cellular niches.
-Using a set of labelled training examples i.e.
-markers, we can train a machine learning classifier to learn a mapping between the data i.e.
-the quantitative protein profiles, and a known localisation.
-The trained classifier can then be used to predict the localisation of a protein of unknown localisation, based on its observed protein profile.
-To date, this method has been extensively used in spatial quantitative proteomics to assign thousands of proteins to distinct sub-cellular niches [@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006; @Tan:2009].
+Supervised machine learning, also known as classification, is an
+essential tool for the assignment of proteins to distinct sub-cellular
+niches. Using a set of labelled training examples i.e. markers, we can
+train a machine learning classifier to learn a mapping between the
+data i.e. the quantitative protein profiles, and a known localisation.
+The trained classifier can then be used to predict the localisation of
+a protein of unknown localisation, based on its observed protein
+profile. To date, this method has been extensively used in spatial
+quantitative proteomics to assign thousands of proteins to distinct
+sub-cellular niches [@hyper; @Groen:2014; @Trotter:2010; @Hall:2009;
+@Dunkley:2006; @Tan:2009].
 
-There are several classification algorithms readily available in `pRoloc`, which are documented in the dedicated [`pRoloc` machine learning techniques vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
-We find the general tendancy to be that it is not the choice of classifier, but the improper optimisation of the algorithmic parameters, that limits classification accuracy.
-Before employing any classification algorithm and generating a model on the training data, one must first find the optimal parameters for the algorithm of choice.
+There are several classification algorithms readily available in
+`pRoloc`, which are documented in the dedicated [`pRoloc` machine
+learning techniques
+vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf).
+We find the general tendancy to be that it is not the choice of
+classifier, but the improper optimisation of the algorithmic
+parameters, that limits classification accuracy. Before employing any
+classification algorithm and generating a model on the training data,
+one must first find the optimal parameters for the algorithm of
+choice.
 
@@
 
-In the code chunk below we employ the use of a Support Vector Machine (SVM) to learn a classifier on the labelled training data.
-The training data is found in the `featureData` slot in the column called `SVM.marker.set`.
-As previously mentioned one first needs to train the classifiers parameters before an algorithm can be used to predict the class labels of the proteins with unknown location.
-One of the most common ways to optimise the parameters of a classifier is to partition the labelled data in to training and testing subsets.
-In this framework parameters are tested via a grid search using cross-validation on the training partition.
-The best parameters chosen from the cross-validation stage are then used to build a classifier to predict the class labels of the protein profiles on the test partition.
-Observed and expected classication results can be compared, and then used to assess how well a given model works by getting an estimate of the classiers ability to achieve a good generalisation i.e.
-that is given an unknown example predict its class label with high accuracy.
-In `pRoloc` algorithmic performance is estimated using stratified 80/20 partitioning for the training/testing subsets respectively, in conjuction with five-fold cross-validation in order to optimise the free parameters via a grid search.
-This procedure is usually repeated 100 times and then the best parameter(s) are selected upon investigation of classifier accuracy, here we use the harmonic mean of precision and recall; the macro F1 score.
-In the code chunk below we demonstrate how to optimise the free parameters; `sigma` and `cost`, of a classical SVM classifier with a Gaussian kernel using the function `svmOptimisation`.
-As the number of labelled instances per class varies from organelle to organelle, we can account for class imbalance by setting specific class weights when generating the SVM model.
-Below the weights, `w` are set to be inversely proportional to the class frequencies.
+In the code chunk below we employ the use of a Support Vector Machine
+(SVM) to learn a classifier on the labelled training data. The
+training data is found in the `featureData` slot in the column called
+`SVM.marker.set`. As previously mentioned one first needs to train the
+classifiers parameters before an algorithm can be used to predict the
+class labels of the proteins with unknown location. One of the most
+common ways to optimise the parameters of a classifier is to partition
+the labelled data in to training and testing subsets. In this
+framework parameters are tested via a grid search using
+cross-validation on the training partition. The best parameters chosen
+from the cross-validation stage are then used to build a classifier to
+predict the class labels of the protein profiles on the test
+partition. Observed and expected classication results can be compared,
+and then used to assess how well a given model works by getting an
+estimate of the classiers ability to achieve a good generalisation
+i.e. that is given an unknown example predict its class label with
+high accuracy. In `pRoloc` algorithmic performance is estimated using
+stratified 80/20 partitioning for the training/testing subsets
+respectively, in conjuction with five-fold cross-validation in order
+to optimise the free parameters via a grid search. This procedure is
+usually repeated 100 times and then the best parameter(s) are selected
+upon investigation of classifier accuracy, here we use the harmonic
+mean of precision and recall; the macro F1 score. In the code chunk
+below we demonstrate how to optimise the free parameters; `sigma` and
+`cost`, of a classical SVM classifier with a Gaussian kernel using the
+function `svmOptimisation`. As the number of labelled instances per
+class varies from organelle to organelle, we can account for class
+imbalance by setting specific class weights when generating the SVM
+model. Below the weights, `w` are set to be inversely proportional to
+the class frequencies.
 
@@
 
-The output `params` is an object of class `GenRegRes`; a dedicated container for the storage of the design and results from a machine learning optimisation.
-To assess classifier performance we can examine the macro F1 scores and the most frequently chosen parameters.
-A high macro F1 score indicates that the marker proteins in the test dataset are consistently and correctly assigned by the the algorithm.
-Often more than one parameter or set of parameters gives rise to the best generalisation accuracy.
-As such it is always important to investigate the model parameters and critically assess the best choice.
-The best choice may not be as simple as the parameter set that gives rise to the highest macro F1 score and one must be careful to avoid overfitting and to choose parameters wisely.
+The output `params` is an object of class `GenRegRes`; a dedicated
+container for the storage of the design and results from a machine
+learning optimisation. To assess classifier performance we can examine
+the macro F1 scores and the most frequently chosen parameters. A high
+macro F1 score indicates that the marker proteins in the test dataset
+are consistently and correctly assigned by the the algorithm. Often
+more than one parameter or set of parameters gives rise to the best
+generalisation accuracy. As such it is always important to investigate
+the model parameters and critically assess the best choice. The best
+choice may not be as simple as the parameter set that gives rise to
+the highest macro F1 score and one must be careful to avoid
+overfitting and to choose parameters wisely.
 
@@
 
-By using the function `getParams` we can extract the best set of parameters.
-Currently, `getParams` retrieves the first best set is automatically but users are encouraged to critically assess whether this is the most wise choice by visualising the results using the methods `plot` and `levelPlot`.
-The `plot` method for `GenRegRes` object shows the respective distributions of the 100 macro F1 scores for the best cost/sigma parameter pairs, and `levelPlot` shows the averaged macro F1 scores, for the full range of parameter values.
-Once we have selected the best parameters we can then use them to build a classifier from the labelled marker proteins.
+By using the function `getParams` we can extract the best set of
+parameters. Currently, `getParams` retrieves the first best set is
+automatically but users are encouraged to critically assess whether
+this is the most wise choice by visualising the results using the
+methods `plot` and `levelPlot`. The `plot` method for `GenRegRes`
+object shows the respective distributions of the 100 macro F1 scores
+for the best cost/sigma parameter pairs, and `levelPlot` shows the
+averaged macro F1 scores, for the full range of parameter values. Once
+we have selected the best parameters we can then use them to build a
+classifier from the labelled marker proteins.
 
@@
 
-We can use the function `svmClassification` to return a classification result for all unlabelled instances in the dataset corresponding to their most likely sub-cellular location.
-The algorithm parameters are passed to the function, along with the class weights and the `fcol` to tell the function where the labelled training data is located, in our case, our marker proteins are located in the `featureData` with the label `"SVM.marker.set"`.
+We can use the function `svmClassification` to return a classification
+result for all unlabelled instances in the dataset corresponding to
+their most likely sub-cellular location. The algorithm parameters are
+passed to the function, along with the class weights and the `fcol` to
+tell the function where the labelled training data is located, in our
+case, our marker proteins are located in the `featureData` with the
+label `"SVM.marker.set"`.
 
 ```{r classify, eval=TRUE}
-svmRes <- svmClassification(pdRes, params,
+svmres <- svmClassification(pdRes, params,
                             class.weights = w,
@@
 
-Automatically, the output of the above classification; the organelle predictions and assignment scores, are stored in the `featureData` slot of the `MSnSet`.
-In this case, they are given the labels `svm` and `svm.scores` for the predictions and scores respectively.
-The resultant predictions can be visualised using `plot2D`.
-In the code chunk below `plot2D` is called to generate a PCA plot of the data and `fcol` is used to specify where the new assignments are located e.g.
+Automatically, the output of the above classification; the organelle
+predictions and assignment scores, are stored in the `featureData`
+slot of the `MSnSet`. In this case, they are given the labels `svm`
+and `svm.scores` for the predictions and scores respectively. The
+resultant predictions can be visualised using `plot2D`. In the code
+chunk below `plot2D` is called to generate a PCA plot of the data and
+`fcol` is used to specify where the new assignments are located e.g.
 `fcol =  "svm"`.
 
-Additionally, when calling `plot2D` we can use the `cex` argument to change the size of each point on the plot (where one point represents one protein) to be inversely proportional to the SVM score.
-This gives an initial overview of the high scoring localisations from the SVM predictions.
+Additionally, when calling `plot2D` we can use the `cex` argument to
+change the size of each point on the plot (where one point represents
+one protein) to be inversely proportional to the SVM score. This gives
+an initial overview of the high scoring localisations from the SVM
+predictions.
 
@@
 ```{r plotSVM, eval=TRUE}
-## set point size of each protein to be inversely proportional to the SVM score
-ptsze <- exp(fData(svmRes)$svm.scores) - 1
+## set point size of each protein to be inversely proportional to the
+ptsze <- exp(fData(svmres)$svm.scores) - 1
 
 ## plot new predictions
-plot2D(svmRes, fcol = "svm", cex = ptsze)
-addLegend(svmRes, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
+plot2D(svmres, fcol = "svm", cex = ptsze)
+addLegend(svmres, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
 ```
@@
 
-It is common when applying a supervised classification algorithm to set a specific score cutoff on which to define new assignments, below which classifications are set to unknown/unassigned.
-This is because in a supervised learning setup proteins can only be predicted to be localised to one of the sub-cellular niches that appear in the labelled training data.
-We can not guaranette (and do not expect) that the whole class diversity to be represented in the training data as (1) finding markers that represent the whole diversity of the cell is challenging and (2) many sub-cellular niches contain too few proteins to train on (note: we recommend a minimum of 13 markers per sub-cellular class for stratified 80/20 partitioning and cross-validation).
+It is common when applying a supervised classification algorithm to
+set a specific score cutoff on which to define new assignments, below
+which classifications are set to unknown/unassigned. This is important
+as in a supervised learning setup proteins can only be predicted to be
+localised to one of the sub-cellular niches that appear in the
+labelled training data. We can not guaranette (and do not expect) that
+the whole class diversity to be represented in the labelled training
+data as (1) finding markers that represent the whole diversity of the
+cell is challenging (especially obtaining dual- and multiply-localised
+protein markers) and (2) many sub-cellular niches contain too few
+proteins to train on (we recommend a minimum of 13 markers per
+sub-cellular class for stratified 80/20 partitioning and 5-fold
+cross-validation). <!--Conversely, with defining markers one must also
+be careful to not label too many markers, as this can lead to
+overfitting and mis-assignment.-->
 
-Deciding on a threshold is not trivial as classifier scores are heavily dependent upon the classifier used and different sub-cellular niches can exhibit different score distributions.
-We recommend one to set class-specific score distributions.
-In the code chunk below we display a boxplot of the score distributions per organelle.
+Deciding on a threshold is not trivial as classifier scores are
+heavily dependent upon the classifier used and different sub-cellular
+niches can exhibit different score distributions, as highlighted in
+the boxplot below. We recommend users to set class-specific thresholds.
+In the code chunk below we display a boxplot of the score distributions 
+per organelle.
 
-```{r thresholds}
+```{r thresholds, fig.height=8, fig.width=8}
+## First remove the markers
+preds <- unknownMSnSet(svmres, fcol = "SVM.marker.set")
 
+## Plot a boxplot of the scores of each organelle
+par(oma= c(10.5, 0, 0, 0))
+boxplot(svm.scores ~ svm, data = fData(preds), ylab = "SVM scores", las = 2)
 ```
 
-To help examine these distributions there is also a dedicated interactive application in the `r Biocpkg("pRolocGUI")` package. To access this application:
+There are many ways to set thresholds and the choice of method will depend
+on the biological question and experimental design at hand. One viable
+approach in the frame of the above experimetal design would be to manually
+set a FDR, say 5%, per organelle. To do this the user would examine 
+the top scoring predictions for each organelle, and then set a threshold at
+the score at which they achieve 5% of false assignments per organelle. 
+The definintion of a false assignment would depend on the information 
+available, for example, validity or lack of validity for the localisation
+from another experiment as reported in the literature or a reliable database. 
+If such information is not available, one crude method is to set a threshold  
+per organelle by extracting the median or 3rd quantile score per organelle. 
+For example, in the code chunk below, we use the `orgQuants` function
+to extract the median organelle scores and then pass these scores to the 
+`getPrediction` function to extract the new localisations that meet this
+scoring criteria. Any sub-cellular predictions that fall below the specified
+thresholds are labelled as unknown.
 
-```{r classifyApp}
+```{r orgquants}
+ts <- orgQuants(svmres, 
+                fcol = "svm", 
+                scol = "svm.scores",
+                mcol = "SVM.marker.set", 
+                t = .5)
 
+preds <- getPredictions(svmres, 
+                        fcol = "svm",
+                        scol = "svm.scores", 
+                        mcol = "SVM.marker.set",
+                        t = ts)
 ```
 
-To launch the classify application:
+The output of `getPredictons` is the original `MSnSet` dataset with a new feature variable appended to the feature data called `fcol.pred` (i.e.
+in our case `svm.pred`) containing the prediction results.
+The results can also be visualied using `plot2D` function.
 
-- Essential for protein localisation as we do not want to force all proteins in to one category (allow for multiple lcoalisations or localisations outside the training data set e.g.
-new novel compartments)
+```{r plotres}
+plot2D(preds, fcol = "svm.pred")
+```
 
-- `getPredictions` function allows you to pass a score threshold. Also `classify` application available to interactively view output of thresholds
+There is also a dedicated interactive application to help users examine these distributions in the `r Biocpkg("pRolocGUI")` package.
+This app can be launched via the `pRolocVis` function and specifying the argument `app = "classify"` along with the relevent `fcol`, `scol` and `mcol` which refer to the columns in the feature data that contain the new assignments, assignment scores and markers respectively (see also `fvarLabels(svmres)`).
+
+```{r classifyApp, eval=FALSE}
+library("pRolocGUI")
+pRolocVis(svmres, app = "classify", 
+          fcol = "svm", 
+	        scol = "svm.scores",
+          mcol = "SVM.marker.set")
+```
+
+![The classify application](./Figures/classify.png)
+
+The data is loaded and displayed on a PCA plot and a boxplot is used to
+display the classifier scores by data class. On the left there is
+a sidebar panel with sliders to control the thresholds upon which classifications 
+are made. There are two types of cut-off that the user can choose from: (1)
+"Quantile" and (2) "User-defined". By default, when the application is launched
+quatile scoring is selected and set to 0.5, the median. The 
+class-specific score thresholds that correspond to selecting the desired
+quantile are shown on as red dots on the boxplot. The assignments on the PCA
+plot are also updated according to the selected threshold. The quantile 
+threshold can be set by moving the corresponding quantile slider. If one
+wished to set their own cut-offs the "User-defined" radio button must be selected
+and then the sliders for defining user-specified scores become active and the
+scores and highlighted on the boxplot by blue dots. For more information we refer
+users to the [`r Biocpkg("pRolocGUI")` tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html).
 

### 2016-07-06 bioc-workflow.Rmd
Started section on TL

@@
 
-There is also a dedicated interactive application to help users examine these distributions in the `r Biocpkg("pRolocGUI")` package.
-This app can be launched via the `pRolocVis` function and specifying the argument `app = "classify"` along with the relevent `fcol`, `scol` and `mcol` which refer to the columns in the feature data that contain the new assignments, assignment scores and markers respectively (see also `fvarLabels(svmres)`).
+There is also a dedicated interactive application to help users
+examine these distributions in the `r Biocpkg("pRolocGUI")` package.
+This app can be launched via the `pRolocVis` function and specifying
+the argument `app = "classify"` along with the relevent `fcol`, `scol`
+and `mcol` which refer to the columns in the feature data that contain
+the new assignments, assignment scores and markers respectively (see
+also `fvarLabels(svmres)`).
 
@@
 
+In addition to high quality MS-based quantitative proteomics data there exist a number of other sources of information that are freely available in the public domain that may be useful to assign a protein to its sub-cellular niche.
+For example, imaging from immunofluorescence microscopy, protein
+annotations and sequences, and protein-protein interactions among others, which represent a rich and vast source of complementary information.
+The integration of data between different technologies is one of the biggest challenges in computational biology to date.
+A paradigm in machine learning, known as transfer learning (TL) is ideally suited to this task.
+The `r Biocpkg("pRoloc")` package provides functionality to do such analyses.
+
+In TL one typically has a primary task that they wish to solve, and some complementary (often heterogeneous) auxiliary information that is related to the primary learning objective, that can be used to help solve the primary goal.
+For example, here our primary task is to assign proteins to their sub-cellular niche with high generalisation accuracy from data collected from quantitative MS-based experiments.
+One useful source of information that may help us solve our task is Gene Ontology information, such as the cellular compartment namespace, for every protein found in the primary experiment.
+We have shown that using transfer learning in the context of spatial proteomics results in the assignment of proteins to sub-cellular niches with higher generalisation accuracy than using standard supervised machine learning [@Breckels:2016].
+
+In the code chunk below, we extract information from the Gene Ontology to use as an auxiliary data source to be intergrated with our high quality MS-based data.
+
+```{r go}
+
+```
+
 # Session information

### 2016-07-07 bioc-workflow.Rmd
Updates to TL section

--- 2016-07-05 bioc-workflow.Rmd
@@
 
+In addition to high quality MS-based quantitative proteomics data there exist a number of other sources of information that are freely available in the public domain that may be useful to assign a protein to its sub-cellular niche.
+For example, imaging from immunofluorescence microscopy, protein
+annotations and sequences, and protein-protein interactions among others, represent a rich and vast source of complementary information.
+We can integrate this auxiliary information with our primary MS-based quantitative data using a paradigm known as transfer learning (TL).
+The integration of data between different technologies is one of the biggest challenges in computational biology to date and the `r Biocpkg("pRoloc")` package provides functionality to do such analyses.
+We recently developed two transfer learning algorithms using a $k$-NN and SVM framework and applied them to the task of protein localisation prediction [@Breckels:2016].
+
+In TL one typically has a primary task that they wish to solve, and some complementary (often heterogeneous) auxiliary information that is related to the primary learning objective, that can be used to help solve the primary goal.
+For example, here our primary task is to assign proteins to their sub-cellular niche with high generalisation accuracy from data collected from quantitative MS-based experiments.
+In the example below we extract Gene Ontology (GO) information to use as an auxiliary data source to help solve our task of protein localisation prediction.
+
+Using the functions `setAnnotationParams` and `makeGoSet` we can contruct an auxiliary `MSnSet` of GO terms, from the primary data's features i.e.
+the protein accession numbers.
+All the GO terms associated to each accession number are retrieved and used to create a binary matrix where a 1 (0) at position (i; j) indicates that term j has (not) been used to annotate protein i.
+The GO terms are retrieved from an appropriate repository using the `r Biocpkg("biomaRt")` package.
+The specific `r Biocpkg("Biomart")` repository and query will depend on the species under study and the type of identifiers.
+The first step is to condtruct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`.
+Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study (below, left), then what type of identifier is used to annotate the proteins in your `MSnSet` (below, right).
+
+![]()
+
+It is also possible to pass patterns to match the species e.g. "Mus musculus", and the identifier used e.g. "Uniprot/Swissprot Accession", for the `Biocpkg("Biomart")` query.
+
+```{r setap}
+par <- setAnnotationParams(inputs = c("Mus musculus", 
+                                      "Uniprot/Swissprot Accession"))
+
+```
+
+Now we have contructed the query parameters we can use the `makeGoSet` function to retrieve and build an auxiliary GO `MSnSet` as described above.
+By default, the cellular component terms are downloaded, without any filtering on evidence codes.
+It is also possible to download terms from the molecular function and biological process GO namespaces, and also apply filtering based on evidence codes as desired, see `?makeGoSet` for more details.
+
+```{r}
+gocc <- makeGoSet(hl, params = par, 
+                  namespace = "cellular_component", 
+                  evidence = NULL)
+gocc
+```
+
+Now we have generated our auxiliary data we can use the $k$-NN implementation of transfer learning available in `Biocpkg("pRoloc")` to integrate this with our primary MS-based quantitative proteomics data using the functions `knntlOptimisation` to estimate the free-parameters for the integration, and `knntlClassification` for the predictions.
+
+The $k$-NN TL algorithm estimates optimal weights for the different data sources and the spatial niches described for the data at hand with the `knntlOptimisation` function.
+
+One useful source of information that may help us solve our task is Gene Ontology information, such as the cellular compartment namespace.
+We have shown that using transfer learning in the context of spatial proteomics results in the assignment of proteins to sub-cellular niches with a higher generalisation accuracy than using standard supervised machine learning with a single source of information [@Breckels:2016].
+
+In the code chunk below, we extract information from the Gene Ontology to use as an auxiliary data source to be intergrated with our high quality MS-based data.
+
 # Session information

### 2016-07-08 bioc-workflow.Rmd
Added TL reference

@@
 The first step is to condtruct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`.
-Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study (below, left), then what type of identifier is used to annotate the proteins in your `MSnSet` (below, right).
-
-![]()
-
-It is also possible to pass patterns to match the species e.g. "Mus musculus", and the identifier used e.g. "Uniprot/Swissprot Accession", for the `Biocpkg("Biomart")` query.
+Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study, and then what type of identifier is used to annotate the proteins in your `MSnSet`.
+It is also possible to pass patterns to match the species e.g.
+in the code chunk below we pass "Mus musculus", and the identifier used e.g.
+"Uniprot/Swissprot", for the `Biocpkg("Biomart")` query.
 
@@
 par <- setAnnotationParams(inputs = c("Mus musculus", 
-                                      "Uniprot/Swissprot Accession"))
+                                      "UniProt/Swissprot"))
 
@@
 
-```{r}
+```{r makego}
 gocc <- makeGoSet(hl, params = par, 
@@
 
-Now we have generated our auxiliary data we can use the $k$-NN implementation of transfer learning available in `Biocpkg("pRoloc")` to integrate this with our primary MS-based quantitative proteomics data using the functions `knntlOptimisation` to estimate the free-parameters for the integration, and `knntlClassification` for the predictions.
-
-The $k$-NN TL algorithm estimates optimal weights for the different data sources and the spatial niches described for the data at hand with the `knntlOptimisation` function.
-
-One useful source of information that may help us solve our task is Gene Ontology information, such as the cellular compartment namespace.
+Now we have generated our auxiliary data we can use the *k*-NN implementation of transfer learning available in `Biocpkg("pRoloc")` to integrate this with our primary MS-based quantitative proteomics data using the functions `knntlOptimisation` to estimate the free-parameters for the integration, and `knntlClassification` to do the predictions.
 We have shown that using transfer learning in the context of spatial proteomics results in the assignment of proteins to sub-cellular niches with a higher generalisation accuracy than using standard supervised machine learning with a single source of information [@Breckels:2016].
 
-In the code chunk below, we extract information from the Gene Ontology to use as an auxiliary data source to be intergrated with our high quality MS-based data.
+The *k*-NN TL algorithm estimates optimal weights for the different data sources and the spatial niches described for the data at hand with the `knntlOptimisation` function.
 

### 2016-07-08 bioc-workflow.Rmd
Added more on TL

@@
 
-## The use-case: prediction sub-cellular localisation in pluripotent embryonic mouse stem cells
+## The use-case: predicting sub-cellular localisation in pluripotent embryonic mouse stem cells
 
@@
 will start with a description of how to generate these ad hoc objects
-starting from any arbitrary spreadsheets, as produced by many popular
+starting from an arbitrary spreadsheets, as produced by many popular
 third-party applications.
@@
 sub-cellular resolution. Missing data can be imputated using 
-
 `r Biocpkg("MSnbase")`'s `impute` function. The method underlying the
@@
 The integration of data between different technologies is one of the biggest challenges in computational biology to date and the `r Biocpkg("pRoloc")` package provides functionality to do such analyses.
-We recently developed two transfer learning algorithms using a $k$-NN and SVM framework and applied them to the task of protein localisation prediction [@Breckels:2016].
+We recently developed two transfer learning algorithms using a *k*-NN and SVM framework and applied them to the task of protein localisation prediction [@Breckels:2016].
+In this section we will begin with explaining the concept of transfer learning and then show how to apply this in the frame of spatial proteomics and protein localisation prediction.
 
@@
 The specific `r Biocpkg("Biomart")` repository and query will depend on the species under study and the type of identifiers.
-The first step is to condtruct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`.
-Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study, and then what type of identifier is used to annotate the proteins in your `MSnSet`.
+The first step is to construct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`.
+Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study, and then what type of identifier you have used to annotate the proteins in your `MSnSet`.
 It is also possible to pass patterns to match the species e.g.
-in the code chunk below we pass "Mus musculus", and the identifier used e.g.
-"Uniprot/Swissprot", for the `Biocpkg("Biomart")` query.
-
-```{r setap}
+in the code chunk below we pass "Mus musculus", and the identifier type for our data (see `featureNames(hl)`) which is "Uniprot/Swissprot", for the `Biocpkg("Biomart")` query.
+
+```{r setap, eval=FALSE}
 par <- setAnnotationParams(inputs = c("Mus musculus", 
@@
 
-```{r makego}
+```{r makego, eval=FALSE}
 gocc <- makeGoSet(hl, params = par, 
@@
 
-The *k*-NN TL algorithm estimates optimal weights for the different data sources and the spatial niches described for the data at hand with the `knntlOptimisation` function.
+We begin by performing the optimisation. The `knntlOptimisation` function estimates the optimal weights for the different data sources and the spatial niches described for the data.
+
+- first need to do k-NN optimisation to get best k's for P and A
+- explain weight matrix of thetas, `by` and `length,out`
+- explain how to do this on the cluster, in practise too computationally heavy on a single processor
+
+```{r tlopt, eval=FALSE}
+
+getMarkers(hl, fcol = "phenoDisco.Input")
+
+opt <- knntlOptimisation(hl, gocc,
+                         fcol = "phenoDisco.Input",
+                         by = .5, 
+                         times = 1,
+                         xval = 5, k = c(3,3),
+                         BPPARAM = par
+  )
+```
 

### 2016-08-11 bioc-workflow.Rmd
Updated section on phenoDisco

@@
 [@hyper]. The data was generated using hyperplexed LOPIT (hyperLOPIT),
-an improved method relying on improved sub-cellular fractionation and
+a method relying on improved sub-cellular fractionation and
 more accurate quantitation, leading to more reliable classification of
@@
 proteins on existing labelled data with the detection of new clusters.
-Novelty detection methods are also useful for confirming the presence
-of known clusters in an unbiased fashion. For example, in [@hyper] the
-`phenoDisco` algorithm was used to detect the data-specific
-confirmation and presence of the nucleus and nucleus sub-compartments,
-as it was expected that nucleus associated clusters existed within the
-data as the protocol included a separate chromatin-rich fraction to
-enrich for nuclei. In the code chunk below, we show how to run a
-series of novelty detection experiments using the `phenoDisco`
-algorithm on the mouse stem-cell dataset. There are several optional
-arguments that can be set dependent on the type of biological question
-of interest. For example,
-
-- Explain how to use different group sizes for detecting small complexes etc.
-- ndims for using >=2 components
-- times -> important for convergence
-- input markers are different here -> explain choice of pd.markers
-
-```{r pd, eval = FALSE}
-## As per the hyperLOPIT paper
-pdRes <- phenoDisco(hl, fcol = "phenoDisco.Input", times = 200, 
-                    GS = 20, p = 0.05)
-
+
+In addition to extracting new phenotypes, novelty detection methods 
+are also useful for confirming the presence of known or postulated 
+clusters in an unbiased fashion. For example, in [@hyper] 
+the `phenoDisco` algorithm was used to confirm the data-specific
+presence of the nucleus and nucleus sub-compartments. In the code 
+chunk below, we re-run this analysis, highlighting some of the 
+optional arguments and parameters available for phenotype extraction 
+and give some advice on how to interpret the output.
+
+As the `phenoDisco` algorithm is semi-supervised it uses both 
+labelled (markers) and unlabelled data to explore the data structure 
+and find new sub-cellular data clusters. Thus the first step is to 
+define some input labelled data i.e. markers, that the algorithm 
+will use as input for the supervised learning aspect of the 
+algorithm. As described in [@hyper] we define a set of markers 
+to use as input for the analyses that cover well-known residents 
+from three distinct organelle structures; the mitochondria, plasma 
+membrane and ER, and from three well-known and abundant protein 
+complexes; the proteasome and two ribosomal subunits, 40S and 60S. 
+These input markers are stored in the `featureData` column of 
+`hl` where `fcol = "phenoDisoc.Input"`. We can use the convenience 
+accessor function `getMarkers` to print out a table of the markers 
+contained in this marker set. These initial markers were manually 
+curated using information from the UniProt database, the Gene 
+Ontology and the literature.
+
+```{r displayPDmarkers}
+getMarkers(hl, fcol = "phenoDisco.Input")
+```
+
+In the code chunk below we run the `phenoDisco` function and 
+return a novelty detection result stored in `pdRes`, according 
+to the specified parameters. The algorithm parameters `times`, 
+`GS` and `p` are passed to the function, along with the `fcol` 
+telling the algorithm where the input training data is contained.
+
+```{r runPD, eval = FALSE}
+## As per Christoforou et al (2016), 
+pdRes <- phenoDisco(hl, fcol = "phenoDisco.Input", 
+                    times = 200, 
+                    GS = 60)
+
+```
+
+The argument `times` defined the number of times we run unsupervied 
+Gaussian Mixture Modelling before defining a new phenotype cluster. 
+The recommended minimum and default value is 100. In the below 
+code chunk we increase the value to `times = 200` as we have found 
+for larger datasets (e.g. 5000+ proteins) a higher times is requried 
+for convergence. `GS` defines the minimum number of proteins allowed 
+per new data cluster and thus heavily influences what type of new 
+clusters are extracted. For example, if a user is interesed in the 
+detection of small complexes they may wish to use a small `GS = 10`, 
+or `=20` etc. If they wish to detect larger more abundant sub-cellular 
+niches a higher `GS` would be preferable. Specifying a small `GS` is be 
+more time consuming than a larger `GS`, and there is a trade off 
+between finding interesting small complexes and those that may not 
+be of interest as we find there is a tendancy to find more noise when 
+using a small `GS` compared to using a higher one.
+
+One may also consider increasing the search space for new data 
+clusters by increasing the value of the parameter `G`. This defines 
+the number of GMM components to test and fit; the default is `G = 1:9` 
+(`mclust`'s default). One should note that the choice of decreasing 
+`GS`, and increasing the values of the arguments `times`, `G` 
+(among other function arguments, see `?phenoDisco`) will heavily 
+influence (increase) the total time taken to run the algorithm. 
+`phenoDisco` supports parallelisation and we strongly suggest 
+you make use of a parallel processing to run these analyses (different 
+backends can be set with argument `BPPARAM`).
+
+The ouput of running the `phenoDisco` algorithm is a `MSnSet` containing 
+the new data clusters, appended to the `featureData` under the name 
+`pd`. We can see by typing `pdRes` directly into the console the 
+processing information has been updated to the `MSnSet` recording the 
+parameters that were used to run the analyses. This is handy for keeping 
+track of data analyses. The results can be displayed by using the 
+`getMarkers` function. We see that 5 new phenotype data clusters were 
+found.
+
+```{r loadPDres, include=FALSE}
+load("Data/pd_runs/pd.rda")
+setStockcol(NULL)
+setStockcol(paste0(getStockcol(), 70))
+```
+```{r showPDres}
+pdRes
+getMarkers(pdRes, fcol = "pd")
+```
+
+We can plot the results using the `plot2D` function.
+
+```{r plotPDres, fig.width=9}
+## Re-order the colours for the phenoDisco output
+cl <- getMarkerClasses(pdRes, "pd", verbose = FALSE)
+cols <- getStockcol()[seq(cl)]
+ind <- grep("Pheno", cl, invert = TRUE)
+cols[ind] <- getStockcol()[seq(cl)][1:length(ind)]
+cols[-ind] <- getStockcol()[seq(cl)][(length(ind) + 1):length(cl)]
+
+## Plot the input and output
+par(mfrow = c(1, 2))
+plot2D(pdRes, fcol = "phenodisco.input", 
+       main = "phenoDisco input markers", col = getStockcol()[1:6])
+addLegend(pdRes, fcol = "phenodisco.input", 
+          where = "bottomleft", cex = .7)
+plot2D(pdRes, fcol = "pd", main = "phenoDisco output", col = cols)
+addLegend(pdRes, fcol = "pd", where = "bottomleft", cex = .7, col = cols)
+```
+
+The five new phenotype data clusters can be extracted and examined. 
+In the code chunk below we write the results to a .csv file. We use 
+the argument `fDataCols` to specify which columns of the `featureData` 
+to write.
+
+```{r writePDres, eval=FALSE}
+fData(pdRes)$pd <- as.character(fData(pdRes)$pd)
+write.exprs(pdRes, fDataCols = "pd", file = "pd-results.csv", sep = ",")
+```
+
+We can also examine the each phenotype intercatively and visualise 
+their protein profiles by using the `pRolocVis` function in the 
+`pRolocGUI`.
+
+```{r visPDres, eval=FALSE}
+pRolocVis(pdRes, fcol = "pd")
 ```

### 2016-08-23 bioc-workflow.Rmd
A few updates and tl section

@@
 of peptides or peptide-spectrum matches (PSMs).
-
+vfn
 ```{r fvarnames}
@@
 par(mfrow = c(1, 2))
-o <- order(hyperLOPIT$Fraction.No)
+o <- order(hyperLOPIT$Iodixonal.Density)
 plotDist(hyperLOPIT[, o], pcol = "#00000010")
@@
 ```{r plotDist}
-hlo <- hyperLOPIT[, order(hyperLOPIT$Fraction.No)]
+hlo <- hyperLOPIT[, order(hyperLOPIT$Iodixonal.Density)]
 plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
@@
 
-```{r pdtab, echo=FALSE}
+```{r pdtab, echo=FALSE, eval = FALSE}
 i <- hyperLOPIT$Fraction.No %in% 7:10
@@
 ```{r setColors, include=FALSE}
+setStockcol(NULL)
 setStockcol(paste0(getStockcol(), 70))
@@
 sub-cellular class for stratified 80/20 partitioning and 5-fold
-cross-validation). <!--Conversely, with defining markers one must also
+cross-validation - this allows a minimum of ~10 examples for parameter optimisation on the training partition i.e.
+2 per fold for 5-fold cross-validation, and then 3 for testing the best parameters on the validation set).
+<!--Conversely, with defining markers one must also
 be careful to not label too many markers, as this can lead to
@@
 
-In addition to high quality MS-based quantitative proteomics data there exist a number of other sources of information that are freely available in the public domain that may be useful to assign a protein to its sub-cellular niche.
-For example, imaging from immunofluorescence microscopy, protein
-annotations and sequences, and protein-protein interactions among others, represent a rich and vast source of complementary information.
-We can integrate this auxiliary information with our primary MS-based quantitative data using a paradigm known as transfer learning (TL).
-The integration of data between different technologies is one of the biggest challenges in computational biology to date and the `r Biocpkg("pRoloc")` package provides functionality to do such analyses.
-We recently developed two transfer learning algorithms using a *k*-NN and SVM framework and applied them to the task of protein localisation prediction [@Breckels:2016].
-In this section we will begin with explaining the concept of transfer learning and then show how to apply this in the frame of spatial proteomics and protein localisation prediction.
-
-In TL one typically has a primary task that they wish to solve, and some complementary (often heterogeneous) auxiliary information that is related to the primary learning objective, that can be used to help solve the primary goal.
-For example, here our primary task is to assign proteins to their sub-cellular niche with high generalisation accuracy from data collected from quantitative MS-based experiments.
-In the example below we extract Gene Ontology (GO) information to use as an auxiliary data source to help solve our task of protein localisation prediction.
-
-Using the functions `setAnnotationParams` and `makeGoSet` we can contruct an auxiliary `MSnSet` of GO terms, from the primary data's features i.e.
-the protein accession numbers.
-All the GO terms associated to each accession number are retrieved and used to create a binary matrix where a 1 (0) at position (i; j) indicates that term j has (not) been used to annotate protein i.
-The GO terms are retrieved from an appropriate repository using the `r Biocpkg("biomaRt")` package.
-The specific `r Biocpkg("Biomart")` repository and query will depend on the species under study and the type of identifiers.
-The first step is to construct the annotation parameters that will enable to perform the query, this is done using `setAnnotationParams`.
-Typing into the R console `par <- setAnnotationParams()` will present two menus, firstly asking you to identify the species of study, and then what type of identifier you have used to annotate the proteins in your `MSnSet`.
-It is also possible to pass patterns to match the species e.g.
-in the code chunk below we pass "Mus musculus", and the identifier type for our data (see `featureNames(hl)`) which is "Uniprot/Swissprot", for the `Biocpkg("Biomart")` query.
+In addition to high quality MS-based quantitative proteomics data
+there exist a number of other sources of information that are freely
+available in the public domain that may be useful to assign a protein
+to its sub-cellular niche. For example, imaging from
+immunofluorescence microscopy, protein annotations and sequences, and
+protein-protein interactions among others, represent a rich and vast
+source of complementary information. We can integrate this auxiliary
+information with our primary MS-based quantitative data using a
+paradigm known as transfer learning (TL). The integration of data
+between different technologies is one of the biggest challenges in
+computational biology to date and the `r Biocpkg("pRoloc")` package
+provides functionality to do such analyses. We recently developed two
+transfer learning algorithms using a *k*-NN and SVM framework and
+applied them to the task of protein localisation prediction
+[@Breckels:2016]. In this section we will begin with explaining the
+concept of transfer learning and then show how to apply this in the
+frame of spatial proteomics and protein localisation prediction.
+
+In TL one typically has a primary task that they wish to solve, and
+some complementary (often heterogeneous) auxiliary information that is
+related to the primary learning objective, that can be used to help
+solve the primary goal. For example, here our primary task is to
+assign proteins to their sub-cellular niche with high generalisation
+accuracy from data collected from quantitative MS-based experiments.
+In the example below we extract Gene Ontology (GO) information to use
+as an auxiliary data source to help solve our task of protein
+localisation prediction.
+
+Using the functions `setAnnotationParams` and `makeGoSet` we can
+contruct an auxiliary `MSnSet` of GO terms, from the primary data's
+features i.e. the protein accession numbers. All the GO terms
+associated to each accession number are retrieved and used to create a
+binary matrix where a 1 (0) at position (i; j) indicates that term j
+has (not) been used to annotate protein i. The GO terms are retrieved
+from an appropriate repository using the `r Biocpkg("biomaRt")`
+package. The specific `r Biocpkg("Biomart")` repository and query will
+depend on the species under study and the type of identifiers. The
+first step is to construct the annotation parameters that will enable
+to perform the query, this is done using `setAnnotationParams`. Typing
+into the R console `par <- setAnnotationParams()` will present two
+menus, firstly asking you to identify the species of study, and then
+what type of identifier you have used to annotate the proteins in your
+`MSnSet`. It is also possible to pass patterns to match the species
+e.g. in the code chunk below we pass "Mus musculus", and the
+identifier type for our data (see `featureNames(hl)`) which is
+"Uniprot/Swissprot", for the `Biocpkg("Biomart")` query.
 
@@
 
-Now we have contructed the query parameters we can use the `makeGoSet` function to retrieve and build an auxiliary GO `MSnSet` as described above.
-By default, the cellular component terms are downloaded, without any filtering on evidence codes.
-It is also possible to download terms from the molecular function and biological process GO namespaces, and also apply filtering based on evidence codes as desired, see `?makeGoSet` for more details.
+Now we have contructed the query parameters we can use the `makeGoSet`
+function to retrieve and build an auxiliary GO `MSnSet` as described
+above. By default, the cellular component terms are downloaded,
+without any filtering on evidence codes. It is also possible to
+download terms from the molecular function and biological process GO
+namespaces, and also apply filtering based on evidence codes as
+desired, see `?makeGoSet` for more details.
 
@@
 
-Now we have generated our auxiliary data we can use the *k*-NN implementation of transfer learning available in `Biocpkg("pRoloc")` to integrate this with our primary MS-based quantitative proteomics data using the functions `knntlOptimisation` to estimate the free-parameters for the integration, and `knntlClassification` to do the predictions.
-We have shown that using transfer learning in the context of spatial proteomics results in the assignment of proteins to sub-cellular niches with a higher generalisation accuracy than using standard supervised machine learning with a single source of information [@Breckels:2016].
-
-We begin by performing the optimisation. The `knntlOptimisation` function estimates the optimal weights for the different data sources and the spatial niches described for the data.
+Now we have generated our auxiliary data we can use the *k*-NN
+implementation of transfer learning available in `Biocpkg("pRoloc")`
+to integrate this with our primary MS-based quantitative proteomics
+data using the functions `knntlOptimisation` to estimate the
+free-parameters for the integration, and `knntlClassification` to do
+the predictions. We have shown that using transfer learning in the
+context of spatial proteomics results in the assignment of proteins to
+sub-cellular niches with a higher generalisation accuracy than using
+standard supervised machine learning with a single source of
+information [@Breckels:2016].
+
+The first step, as with any machine learning algorithm, is to
+optimise any free paramaters of the classifier. For the *k*-NN TL
+classifier these are the class weights, one per subcellular niche,
+that control the proportion of primary and auxiliary data to use for
+learning. A weight can take any real value number between 0 and 1. A
+weight of $\theta = 1$ indicates that all weight is given to the primary
+data (and this implicitly implies that a weight of $1 - \theta$ is given to
+the auxiliary data, similarly a weight of $\theta = 0$ implies that all
+weight is given to the auxiliary data (so 0 to the primary source).
+If we conduct a parameter seach and test weights $\theta = {0, 1/3, 2/3, 1}$
+for each class, and if we have say 10 subcellular niches, this
+will result in `4^10` different combinations of parameters to test. The
+parameter optimisation is therefore time consuming and as such
+we recommend you make use of a computing cluster.
+
+<!--In the example below
+we first examine a PCA plot of the data and overlay the remaining
+subcellular markers. We can see that the two ribosomes, two nuclear
+compartments and plasma membrane are quite well-defined and separated
+form the rest of the plot, indicating that they are well-resolved in
+the hyperLOPIT experiments. For this reason, and in the interest of
+parameter optimisation and time we have removed these classes from the
+$k$-NN TL optimisation as we expect their weight to be 1.
+
+- add code here
+
+We begin by performing the optimisation. The `knntlOptimisation`
+function estimates the optimal weights for the different data sources
+and the spatial niches described for the data.
 
@@
 
-getMarkers(hl, fcol = "phenoDisco.Input")
-
 opt <- knntlOptimisation(hl, gocc,
-                         fcol = "phenoDisco.Input",
+                         fcol = ,
                          by = .5, 
@@
 ```
+-->
 

### 2016-08-25 bioc-workflow.Rmd
minor changes

@@
 ```{r readMSnSet2}
-hl <- readMSnSet2(csvfile, ecol = c(8:27), fnames = 1, skip = 1)
+hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
 ```
@@
 
-To list all data distributed with a package, one can use pass the
-package name as input to the `data` function. For instance
+To list all data distributed with a package, one can use use the
+`pRolocdata` function.
 
 ```{r pkgdata, eval=FALSE}
-data(package = "pRolocdata")
-```
-
-will list the `r nrow(data(package = "pRolocdata")$results)` datasets
+pRolocdata()
+```
+
+will list the `r nrow(pRolocdata()$results)` datasets
 that are available in `r Biocexptpkg("pRolocdata")`.

### 2016-08-27 bioc-workflow.Rmd
Close #19: use markers rather than SVM.marker.set, and remove Golgi using fDataToUnknown

@@
 
-```{r env, echo=FALSE}
+```{r env, echo=FALSE, warning=FALSE}
 library("BiocStyle")
@@
 
+As demonstrated in [@hyper] and illustrated in the PCA plot above, the
+Golgi apparatus proteins display a dynamic pattern, noting sets of
+Golgi marker proteins that are distributed amongst other subcellular
+structures, an observation supported by microscopy. As such, we are
+going to reset the annotation of Golgi markers to unknown using the
+`fDataTounknown` function. It is often used to replace empty strings
+("") or missing values in the markers definition to a common
+definition of *unknown localisation*.
+
+```{r}
+hl <- fDataToUnknown(hl, from = "Golgi", to = "unknown")
+getMarkers(hl)
+```
+
 In general, the Gene Ontology (GO) [@Ashburner:2000], and in
@@
 par(mfrow = c(1, 2))
-plot2D(hl[, hl$replicate == 1], fcol = "SVM.marker.set", main = "Replicate 1")
-plot2D(hl[, hl$replicate == 2], fcol = "SVM.marker.set", main = "Replicate 2",
+plot2D(hl[, hl$replicate == 1], main = "Replicate 1")
+plot2D(hl[, hl$replicate == 2], main = "Replicate 2",
        mirrorX = TRUE)
@@
 ## Re-order the colours for the phenoDisco output
-cl <- getMarkerClasses(pdRes, "pd", verbose = FALSE)
+cl <- getMarkerClasses(pdRes, "pd")
 cols <- getStockcol()[seq(cl)]
@@
 In the code chunk below we employ the use of a Support Vector Machine
-(SVM) to learn a classifier on the labelled training data. The
-training data is found in the `featureData` slot in the column called
-`SVM.marker.set`. As previously mentioned one first needs to train the
-classifiers parameters before an algorithm can be used to predict the
-class labels of the proteins with unknown location. One of the most
-common ways to optimise the parameters of a classifier is to partition
-the labelled data in to training and testing subsets. In this
-framework parameters are tested via a grid search using
-cross-validation on the training partition. The best parameters chosen
-from the cross-validation stage are then used to build a classifier to
-predict the class labels of the protein profiles on the test
-partition. Observed and expected classication results can be compared,
-and then used to assess how well a given model works by getting an
-estimate of the classiers ability to achieve a good generalisation
-i.e. that is given an unknown example predict its class label with
-high accuracy. In `pRoloc` algorithmic performance is estimated using
-stratified 80/20 partitioning for the training/testing subsets
-respectively, in conjuction with five-fold cross-validation in order
-to optimise the free parameters via a grid search. This procedure is
-usually repeated 100 times and then the best parameter(s) are selected
-upon investigation of classifier accuracy, here we use the harmonic
-mean of precision and recall; the macro F1 score. In the code chunk
-below we demonstrate how to optimise the free parameters; `sigma` and
-`cost`, of a classical SVM classifier with a Gaussian kernel using the
-function `svmOptimisation`. As the number of labelled instances per
-class varies from organelle to organelle, we can account for class
-imbalance by setting specific class weights when generating the SVM
-model. Below the weights, `w` are set to be inversely proportional to
-the class frequencies.
+(SVM) to learn a classifier on the labelled training data. As
+previously mentioned one first needs to train the classifiers
+parameters before an algorithm can be used to predict the class labels
+of the proteins with unknown location. One of the most common ways to
+optimise the parameters of a classifier is to partition the labelled
+data in to training and testing subsets. In this framework parameters
+are tested via a grid search using cross-validation on the training
+partition. The best parameters chosen from the cross-validation stage
+are then used to build a classifier to predict the class labels of the
+protein profiles on the test partition. Observed and expected
+classication results can be compared, and then used to assess how well
+a given model works by getting an estimate of the classiers ability to
+achieve a good generalisation i.e. that is given an unknown example
+predict its class label with high accuracy. In `pRoloc` algorithmic
+performance is estimated using stratified 80/20 partitioning for the
+training/testing subsets respectively, in conjuction with five-fold
+cross-validation in order to optimise the free parameters via a grid
+search. This procedure is usually repeated 100 times and then the best
+parameter(s) are selected upon investigation of classifier accuracy,
+here we use the harmonic mean of precision and recall; the macro F1
+score. In the code chunk below we demonstrate how to optimise the free
+parameters; `sigma` and `cost`, of a classical SVM classifier with a
+Gaussian kernel using the function `svmOptimisation`. As the number of
+labelled instances per class varies from organelle to organelle, we
+can account for class imbalance by setting specific class weights when
+generating the SVM model. Below the weights, `w` are set to be
+inversely proportional to the class frequencies.
 
@@
 load("Data/fusionparams.rda")
-(w <- table(fData(pdRes)[, "SVM.marker.set"]))
+w <- table(getMarkers(pdRes, verbose = TRUE))
 w <- 1/w[names(w) != "unknown"]
@@
 ```{r optimise, eval=FALSE}
-(w <- table(fData(pdRes)[, "SVM.marker.set"]))
+w <- table(getMarkers(pdRes, verbose = TRUE))
 w <- 1/w[names(w) != "unknown"]
@@
 ## 100 rounds of optimisation with five-fold cross-validation
-params <- svmOptimisation(pdRes, fcol = "SVM.marker.set",
+params <- svmOptimisation(pdRes, fcol = "markers",
                           times = 100, xval = 5,
@@
 ```
+
+As mentioned previously, we reply on the default feature variable
+`"markers"` to define the class labels and hence can ommit it. To use
+another feature variables, one need to explicitly specify its name
+using the `fcol` argument (for example `fcol = "markers2"`).
 
@@
 their most likely sub-cellular location. The algorithm parameters are
-passed to the function, along with the class weights and the `fcol` to
-tell the function where the labelled training data is located, in our
-case, our marker proteins are located in the `featureData` with the
-label `"SVM.marker.set"`.
+passed to the function, along with the class weights. As above, `fcol`
+can be ignored as we use the labels defined in the default `"markers"`
+feature variable.
 
@@
                             class.weights = w,
-                            fcol = "SVM.marker.set")
+                            fcol = "markers")
 ```
@@
 sub-cellular class for stratified 80/20 partitioning and 5-fold
-cross-validation - this allows a minimum of ~10 examples for parameter optimisation on the training partition i.e.
-2 per fold for 5-fold cross-validation, and then 3 for testing the best parameters on the validation set).
-<!--Conversely, with defining markers one must also
+cross-validation - this allows a minimum of ~10 examples for parameter
+optimisation on the training partition i.e. 2 per fold for 5-fold
+cross-validation, and then 3 for testing the best parameters on the
+validation set). <!--Conversely, with defining markers one must also
 be careful to not label too many markers, as this can lead to
@@
 ## First remove the markers
-preds <- unknownMSnSet(svmres, fcol = "SVM.marker.set")
+preds <- unknownMSnSet(svmres)
 
@@
                 scol = "svm.scores",
-                mcol = "SVM.marker.set", 
+                mcol = "markers", 
                 t = .5)
@@
                         scol = "svm.scores", 
-                        mcol = "SVM.marker.set",
+                        mcol = "markers",
                         t = ts)
@@
 	        scol = "svm.scores",
-          mcol = "SVM.marker.set")
+          mcol = "markers")
 ```
@@
 scores and highlighted on the boxplot by blue dots. For more information we refer
-users to the [`r Biocpkg("pRolocGUI")` tutorial vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html).
+users to the `r Biocpkg("pRolocGUI")` tutorial 
+[vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html).
 

### 2016-08-31 bioc-workflow.Rmd
Updated naming of msnsets, added pData

@@
 identify which columns of the spreadsheet contain the
-quantitation data.
-
-We first locate the the spreadsheet of E14TG2a data distributed with
-the `r Biocexptpkg("pRolocdata")` data using `system.file`, that
-locates the `extdata` in the `r Biocexptpkg("pRolocdata")` package on
-the hard drive and `dir`, that displays the full path to the file
-matching the `hyperLOPIT-SIData-ms3-rep12-intersect.csv` pattern,
-which corresponds the the file of interest. In the last line, we print
+quantitation data. In the last line of the code chunk below, we print
 the file name (not the full path, which will vary from computer to
@@
 
-As briefly mentioned above, the quantitation data is stored in the
-`exprs` slot of the `MSnSet` and can be accessed by
-`exprs(hl)`. Below, we examine the quantitative information for
+Below, we examine the quantitative information for
 first 5 proteins. It is also possible to access specific rows and
@@
 ```{r pdata}
-pData(hl)$replicate <- rep(1:2, each = 10)
-pData(hl)$tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
+pData(hl)$Replicate <- rep(1:2, each = 10)
+pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
+pData(hl)
+```
+
+Sample meta-data distributed with
+the `r Biocexptpkg("pRolocdata")` and as above with the quantitation
+data file is located in the `extdata` in the `r Biocexptpkg("pRolocdata")` 
+package on the hard drive. In the code chunk below we again use the `dir` 
+function to locate the filepath to the meta-data `csv` file and then read 
+it into R using `read.csv` and append the meta-data to the `pData` slot. 
+We add information on the gradient fractions used and the associated 
+subcellular fraction densities in % w/v Iodixanol for each replicate.
+
+```{r morepdata}
+expinfo <- dir(extdatadir, full.names = TRUE,
+               pattern = "hyperLOPIT-SIData-fraction-info.csv")
+
+fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
+                     header = FALSE, stringsAsFactors = FALSE)
+
+pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
+pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
 pData(hl)
@@
 (typically 0 and 1) by, for example, diving each intensity by the sum
-of the intensities for that quantitative feature, as illustrated
-below.
+of the intensities for that quantitative feature. This is not necessary 
+in this example as the intensities for each replicate have already 
+been re-scaled to 1 in Proteome Discoverer. However, if one wanted
+to do this they would execute the `normalise` function as demonstrated 
+in the below code chunk.
 
@@
 
-Before combination, the two replicates in the `hl` data that we
-read into R were separately normalised by sum across the 10 channels
-for each replicate respectively. We can verify this by summing each
-rows for each replicate:
+As previously mentioned, before combination, the two replicates in 
+the `hl` data that we read into R were separately normalised 
+by sum (i.e. to 1) across the 10 channels for each replicate 
+respectively. We can verify this by summing each rows for each replicate:
 
@@
 ```
+
 We see that some features do not add up exactly to 1 due to rounding
@@
 values. We start by loading the experimental data package and the
-equivalent replicates using the `load` function.
+equivalent replicates using the `data` function.
 
@@
 fvarLabels(hyperLOPIT2015ms3r2)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r2)[3:5], 2)
+fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5] 
 fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5] 
@@
 ```{r combine}
-hyperLOPIT <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
-hyperLOPIT
+combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
+combined
 ```
@@
 ```{r namap}
-image2(is.na(hyperLOPIT), col = c("black", "white"),
+image2(is.na(combined), col = c("black", "white"),
        main = "Missing values (white cells) after combining replicates")
@@
 ```{r filterNA}
-hyperLOPIT <- filterNA(hyperLOPIT)
-hyperLOPIT
+combined <- filterNA(combined)
+combined
 ```
@@
 library("pRoloc")
-plot2D(hyperLOPIT, fcol = NULL)
+plot2D(hl, fcol = NULL, col = "black")
 ```
@@
 [this](https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32)
-very interesting gist with intriguing tSNE examples. This might be
-better to come after markers.
+very interesting gist with intriguing tSNE examples. <!--This might be
+better to come after markers.-->
 
@@
 par(mfrow = c(1, 2))
-o <- order(hyperLOPIT$Iodixonal.Density)
-plotDist(hyperLOPIT[, o], pcol = "#00000010")
-lines(colMeans(exprs(hyperLOPIT[, o])), col = "red", type = "b")
-boxplot(exprs(hyperLOPIT[, o]), las = 2)
+o <- order(hl$Iodixonal.Density)
+plotDist(hl[, o], pcol = "#00000010")
+lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
+boxplot(exprs(hl[, o]), las = 2)
 ```
@@
 par(mfrow = c(1, 2))
-plot2D(hl, fcol = "markers", main = "pRolocmarkers for mouse")
-addLegend(hl, fcol = "markers", where = "bottomleft", cex = .7)
+plot2D(hl, main = "pRolocmarkers for mouse")
+addLegend(hl, cex = .7)
 plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
@@
 ```{r plotDist}
-hlo <- hyperLOPIT[, order(hyperLOPIT$Iodixonal.Density)]
+hlo <- hl[, order(hl$Iodixonal.Density)]
 plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
@@
 ```{r pdtab, echo=FALSE, eval = FALSE}
-i <- hyperLOPIT$Fraction.No %in% 7:10
-pData(hyperLOPIT)[i, -3]
+i <- c(2, 3, 4, 13, 14)
+pData(hl)[i, -3]
 ```
@@
 par(mfrow = c(1, 2))
-plot2D(hl[, hl$replicate == 1], main = "Replicate 1")
-plot2D(hl[, hl$replicate == 2], main = "Replicate 2",
+plot2D(hl[, hl$Replicate == 1], main = "Replicate 1")
+plot2D(hl[, hl$Replicate == 2], main = "Replicate 2",
        mirrorX = TRUE)
@@
 
-```{r}
-i1 <- exprs(hl[, hl$replicate == 1])
-i2 <- exprs(hl[, hl$replicate == 2])
+```{r pearson}
+i1 <- exprs(hl[, hl$Replicate == 1])
+i2 <- exprs(hl[, hl$Replicate == 2])
 pearcor <- cor(t(i1), t(i2))

### 2016-09-02 bioc-workflow.Rmd
Finished tl section and modified pd section

@@
 ## As per Christoforou et al (2016), 
-pdRes <- phenoDisco(hl, fcol = "phenoDisco.Input", 
+hl <- phenoDisco(hl, fcol = "phenoDisco.Input", 
                  times = 200, 
@@
 load("Data/pd_runs/pd.rda")
+hl <- pdRes
 setStockcol(NULL)
@@
 ```{r showPDres}
-pdRes
-getMarkers(pdRes, fcol = "pd")
+hl
+getMarkers(hl, fcol = "pd")
 ```
@@
 ## Re-order the colours for the phenoDisco output
-cl <- getMarkerClasses(pdRes, "pd")
+cl <- getMarkerClasses(hl, "pd")
 cols <- getStockcol()[seq(cl)]
@@
 par(mfrow = c(1, 2))
-plot2D(pdRes, fcol = "phenodisco.input", 
+plot2D(hl, fcol = "phenodisco.input", 
        main = "phenoDisco input markers", col = getStockcol()[1:6])
-addLegend(pdRes, fcol = "phenodisco.input", 
-          where = "bottomleft", cex = .7)
-plot2D(pdRes, fcol = "pd", main = "phenoDisco output", col = cols)
-addLegend(pdRes, fcol = "pd", where = "bottomleft", cex = .7, col = cols)
+addLegend(hl, fcol = "phenodisco.input", cex = .7)
+plot2D(hl, fcol = "pd", main = "phenoDisco output", col = cols)
+addLegend(hl, fcol = "pd", cex = .7, col = cols)
 ```
@@
 
-```{r writePDres, eval=FALSE}
-fData(pdRes)$pd <- as.character(fData(pdRes)$pd)
-write.exprs(pdRes, fDataCols = "pd", file = "pd-results.csv", sep = ",")
+```{r writehl, eval=FALSE}
+fData(hl)$pd <- as.character(fData(hl)$pd)
+write.exprs(hl, fDataCols = "pd", file = "pd-results.csv", sep = ",")
 ```
@@
 
-```{r visPDres, eval=FALSE}
-pRolocVis(pdRes, fcol = "pd")
-```
+```{r vishl, eval=FALSE}
+pRolocVis(hl, fcol = "pd")
+```
+
+From examining the literature and the Gene Ontology we find that
+phenotype 1 is enriched in proteins that localise to the nucleus
+(non-chromatin associated proteins), phenotype 2 is enriched in
+chromatin associated proteins, phenotype 3 contains many cytosolic
+proteins and phenotypes 4 and 5 contain lysosmal and endosomal
+proteins. From curating these phenotypes we obtain new marker proteins
+that can be used as input for supervised classification.
 
@@
 ```{r loadpdRes, include=FALSE}
-pdRes <- hl 
 load("Data/fusionparams.rda")
-w <- table(getMarkers(pdRes, verbose = TRUE))
+w <- table(getMarkers(hl, verbose = TRUE))
 w <- 1/w[names(w) != "unknown"]
@@
 ```{r optimise, eval=FALSE}
-w <- table(getMarkers(pdRes, verbose = TRUE))
+w <- table(getMarkers(hl, verbose = TRUE))
 w <- 1/w[names(w) != "unknown"]
@@
 ## 100 rounds of optimisation with five-fold cross-validation
-params <- svmOptimisation(pdRes, fcol = "markers",
+params <- svmOptimisation(hl, fcol = "markers",
                           times = 100, xval = 5,
-                          class.weights = w,
-                          verbose = FALSE)
+                          class.weights = w)
 ```
@@
 ```{r visualiseOptHide, echo=FALSE, fig.width = 12}
+library('gridExtra')
 grid.arrange(plot(params), levelPlot(params), ncol = 2)
@@
 ```{r classify, eval=TRUE}
-svmres <- svmClassification(pdRes, params,
+hl <- svmClassification(hl, params,
                         class.weights = w,
@@
 setStockcol(paste0(getStockcol(), 70))
+par(mfrow = c(1, 1))
 ```
@@
 ## set point size of each protein to be inversely proportional to the
-ptsze <- exp(fData(svmres)$svm.scores) - 1
+ptsze <- exp(fData(hl)$svm.scores) - 1
 
 ## plot new predictions
-plot2D(svmres, fcol = "svm", cex = ptsze)
-addLegend(svmres, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
+plot2D(hl, fcol = "svm", cex = ptsze)
+addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
 ```
@@
 the labelled marker proteins. These class/organelle boundaries define
-how non-assigned proteins are classifier and with what
+how non-assigned proteins are classified and with what
 confidence. **TODO** explain how choosing markers
@@
 optimisation on the training partition i.e. 2 per fold for 5-fold
-cross-validation, and then 3 for testing the best parameters on the
-validation set). <!--Conversely, with defining markers one must also
+cross-validation, and then ~3 for testing the best parameters on the
+validation set). <!--Conversely, when defining markers one must also
 be careful to not label too many markers, as this can lead to
@@
 ## First remove the markers
-preds <- unknownMSnSet(svmres)
+preds <- unknownMSnSet(hl)
 
@@
 ```{r orgquants}
-ts <- orgQuants(svmres, 
+ts <- orgQuants(hl, 
                 fcol = "svm", 
@@
 
-preds <- getPredictions(svmres, 
+hl <- getPredictions(hl,
                      fcol = "svm",
@@
                      t = ts)
+
 ```
@@
 
+```{r resetpar}
+dev.off()
+```
 ```{r plotres}
-plot2D(preds, fcol = "svm.pred")
+plot2D(hl, fcol = "svm.pred")
 ```
@@
 library("pRolocGUI")
-pRolocVis(svmres, app = "classify",
+pRolocVis(hl, 
+          app = "classify",
           fcol = "svm",
@@
 
-The first step, as with any machine learning algorithm, is to
-optimise any free paramaters of the classifier. For the *k*-NN TL
-classifier these are the class weights, one per subcellular niche,
-that control the proportion of primary and auxiliary data to use for
-learning. A weight can take any real value number between 0 and 1. A
-weight of $\theta = 1$ indicates that all weight is given to the primary
-data (and this implicitly implies that a weight of $1 - \theta$ is given to
-the auxiliary data, similarly a weight of $\theta = 0$ implies that all
-weight is given to the auxiliary data (so 0 to the primary source).
-If we conduct a parameter seach and test weights $\theta = {0, 1/3, 2/3, 1}$
-for each class, and if we have say 10 subcellular niches, this
-will result in `4^10` different combinations of parameters to test. The
-parameter optimisation is therefore time consuming and as such
-we recommend you make use of a computing cluster.
-
-<!--In the example below
-we first examine a PCA plot of the data and overlay the remaining
-subcellular markers. We can see that the two ribosomes, two nuclear
-compartments and plasma membrane are quite well-defined and separated
-form the rest of the plot, indicating that they are well-resolved in
-the hyperLOPIT experiments. For this reason, and in the interest of
-parameter optimisation and time we have removed these classes from the
-$k$-NN TL optimisation as we expect their weight to be 1.
-
-- add code here
-
-We begin by performing the optimisation. The `knntlOptimisation`
-function estimates the optimal weights for the different data sources
-and the spatial niches described for the data.
-
-- first need to do k-NN optimisation to get best k's for P and A
-- explain weight matrix of thetas, `by` and `length,out`
-- explain how to do this on the cluster, in practise too computationally heavy on a single processor
+The first step, as with any machine learning algorithm, is to optimise
+any free paramaters of the classifier. For the *k*-NN TL classifier
+there are two sets of parameters that need optimising: the first set
+are the $k$'s for the primary and auxiliary data sources required for
+the nearest neighbour calculations for each data source.  The second
+set of parameters (which is noted by a vector of $\theta$ weights)
+that require optimising are the class weights, one per subcellular
+niche, that control the proportion of primary and auxiliary data to
+use for learning. A weight can take any real value number between 0
+and 1. A weight of $\theta = 1$ indicates that all weight is given to
+the primary data (and this implicitly implies that a weight of $1 -
+\theta$ is given to the auxiliary data, similarly a weight of $\theta
+= 0$ implies that all weight is given to the auxiliary data (so 0 is
+given to the primary source). If we conduct a parameter seach and test
+weights $\theta = {0, 1/3, 2/3, 1}$ for each class, and if we have,
+for example 10 subcellular niches, this will result in `4^10`
+different combinations of parameters to test. The parameter
+optimisation is therefore time consuming and as such we recommend you
+make use of a computing cluster (code and submissing scripts are also
+available in the supporting information). The markers in the `hl`
+dataset contain 14 subcellular classes. If we examine these markers
+and classes on the PCA plot above we can see that in particular the
+two ribosomes and two nuclear compartments are highly separated along
+the first two components, this is also evident from the profiles plot
+which gives us a good indication that these subcellular niches are
+well-resolved in the hyperLOPIT dataset. Transfer learning is
+particularly useful for classes that are not as well separated, we
+find that subcellular niches that are well-separated under hyperLOPIT
+and LOPIT obtain a class score of 1 (i.e. use only primary data from
+transfer learning [@Breckels:2016]). Therefore, for the optimisation
+stage of the analyses we can already infer a subcellular class weight
+of 1 for these niches and only optimise over the remaining organelles.
+This can signifciantly cut down optimisation time as by removing these
+4 classes from the optimisation (and not the classification) we only
+have `4^10` class weight combinations to consider instead of `4^14`
+combinations. In the example below we remove these 4 classes from the
+marker set, re-run the `knnOptimisation` for each data source and then
+run the `knntlOptimisation` with the 10 remaining classes. (Note: this
+is not run live as this the `hl` dataset with 10 classes, 707 markers
+and `4^10` combinations of parameters takes ~76 hours to run on the
+University of Cambridge HPC using 256 workers).
+
+In the example code below we first remove the 2 ribosomes and nuclear
+classes and then run the `knnOptimisation` followed by the
+`knntlOptimisation` (specifying snow with a MPI
+backend with 255 workers for parallelisation on the Cambridge HPC).
 
 ```{r tlopt, eval=FALSE}
-
-opt <- knntlOptimisation(hl, gocc,
-                         fcol = ,
-                         by = .5, 
-                         times = 1,
-                         xval = 5, k = c(3,3),
+## create new markers column for tl markers
+fData(hl)$tlmarkers <- fData(hl)$markers
+fData(gocc)$tlmarkers <- fData(gocc)$markers
+
+## Remove 4 classes
+torm  <- c("40S Ribosome", "60S Ribosome",
+           "Nucleus - Chromatin", 
+           "Nucleus - Non-chromatin")
+for (i in seq(torm)) {
+  hl <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
+  gocc <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
+}
+getMarkerClasses(hl, fcol = "tlmarkers")
+getMarkerClasses(hl, fcol = "tlmarkers")
+
+## get best k's
+kp <- knnOptimisation(hl, fcol = "tlmarkers")
+ka <- knnOptimisation(gocc, fcol = "tlmarkers")
+
+## Set appropriate backend and workers for the tl
+par <- SnowParam(255L, type = "MPI")
+
+## Now peform tl optimisation 
+tlopt <- knntlOptimisation(hl, gocc,
+                           fcol = "tlmarkers",
+                           length.out = 4, 
+                           times = 50,
+                           xval = 5, k = c(kp,ka),
                            BPPARAM = par
@@
 ```
--->
+```{r loadtlres, include = FALSE}
+load("Data/tl_runs/tlopt.rda")
+```
+
+The results of the optimisation can be visalised using the `plot`
+method for `"ThetaRegRes"` objects:
+
+```{r plottl}
+plot(tlopt)
+```
+
+Looking at the bubble plot displaying the distribution of best weights
+over the 50 runs we find that for many of the subcellular niches a
+weight of 1 is most popular (i.e. use only primary hyperLOPIT data in
+classification), this is unsuprising as we already know the dataset is
+well resolved for these classes. We see that the most popular weights
+for the proteasome and lysosome tend to be towards 0, indicating that
+these niches are well-resolved in the Gene Ontology. This tells us
+that we would benefit from including auxiliary GO information in our
+classifier for these subcellular compartments. The plasma membrane
+weights are relatively equally spread between using hyperLOPIT and GO
+data. Using the `getParams` function we can return the best weights
+and then use this as input for the classification.
+
+One of the benefits of the algorithm is the ability to manually select
+weights for each class. In the optimisation above, for time
+constraints, we removed the two ribosomal subunits and the two nulcear
+compartments, and therefore in the code chunk below when we extract
+the best parameters, these subcellular niches are not included. To
+include these 4 subcellular niches in the next classification step we
+must include them in the parameters so we infer a weight of 1 for each
+of these niches as we know they are well resolved in hyperLOPIT. We
+then re-order the weights according to `getMarkerClasses` and perform
+the classification using the function `knntlClassification`.
+
+```{r tlclassify}
+## best parameters for the 10 classes
+(bestpar <- getParams(tlopt))
+
+## add weights for classes not included in the optimisation
+otherweights <- rep(1, 4)
+names(otherweights) <- c("40S Ribosome", "60S Ribosome", 
+                         "Nucleus - Chromatin", 
+                         "Nucleus - Non-chromatin")
+(bestpar <- c(bestpar, otherweights))
+
+## re-order classes 
+bestpar <- bestpar[getMarkerClasses(hl)]
+
+## Do the classification
+hl <- knntlClassification(hl, gocc, bestTheta = bestpar, 
+                          k = c(kp, ka))
+
+```
+
+The results from the classification results and associated scores are
+appended to the `fData` slot and named `knntl` and `knntl.scores`
+respectively. Results can be visualised using `plot2D`, scores
+assessed and cutoffs calculated using the `classify` app in
+`pRolocVis`, predictions obtained using `getPredictions` in the same
+way as demonstrated above for the SVM classifier.
 

### 2016-09-06 bioc-workflow.Rmd
add todo re plot3D and marker clustering
remove peason cor, as this is nonsense anyway

@@
 
+**TODO** Possibly mention `plot3D` and `mrkHClust`, although these are
+only available in the development version.
+
 # Replication
@@
 ```
-
-## Peason correlation
-
-```{r pearson}
-i1 <- exprs(hl[, hl$Replicate == 1])
-i2 <- exprs(hl[, hl$Replicate == 2])
-pearcor <- cor(t(i1), t(i2))
-pearcor0 <- cor(t(i1[, seq(1, 10, 2)]), t(i2[, seq(1, 10, 2)]))
-mean(diag(pearcor))
-mean(diag(pearcor0))
-```
-
-- What about classification on different replicates?
 

### 2016-09-07 bioc-workflow.Rmd
Small updates to workflow

--- 2016-09-02 bioc-workflow.Rmd
@@
 author: "Lisa Breckels and Laurent Gatto"
-output: html_document
+output: pdf_document
 ---
@@
 [CRAN](https://cran.r-project.org/) as well as GitHub. Once a package
-has been installed, it needs to be loaded for it functionality become
-available in the R session; this is done with the `library` function.
+has been installed, it needs to be loaded for its functionality to become
+available in the R session; this is done with the `library` function e.g. 
+to load the `r Biocpkg("pRoloc")` one would type `library("pRoloc")` after
+installation.
 
 If you have questions about this workflow in particular, or about
-other Bioconductor packages in general, they a best asked on the
+other Bioconductor packages in general, they are best asked on the
 [Bioconductor support site](https://support.bioconductor.org/)
@@
 [@hyper]. The data was generated using hyperplexed LOPIT (hyperLOPIT),
-a method relying on improved sub-cellular fractionation and
+a state-of-the-art method relying on improved sub-cellular fractionation and
 more accurate quantitation, leading to more reliable classification of
@@
 There are a number of ways to import quantitation data and create an
-`MSnSet` instance and all methods are described in the `r Biocpkg("MSnbase")`
+`MSnSet` instance. All methods are described in the `r Biocpkg("MSnbase")`
 [input/output capabilities vignette](http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf). One
@@
 quantitation data by `ecol` or the feature names by `fnames` is
-deposited to the `fData` slot. As we wish to demonstrate the complete
-analysis of this data we remove the results from the prior analysis
-described in [@hyper] in the code chunk below. We see the `fData`
+deposited to the `fData` slot.
+
+We see the `fData`
 contains 25 columns describing information such as the number of
-peptides, associated markers, machine learning results etc.
-
-For demonstration in the code chunk below keep the 2nd, 8th and 11th
-columns which contain the UniProt entry names and two different marker
-sets to use an input for machine learning analyses (see sections on
-markers and subsequent sections) and rename the first feature variable
-names.
+peptides, associated markers, machine learning results etc. To
+identify the feature variable names we can use the function 
+`fvarLabels`. We see that the first 6 feature variable names 
+contain non-discriminatory label names, so we relabel them to
+help us identify what feature data information is stored in
+the associated columns.
 
@@
 fvarLabels(hl)
-fData(hl) <- fData(hl)[, c(2, 8, 11)]
-fvarLabels(hl)[1] <- "UniProtId"
-head(fData(hl))
+fvarLabels(hl)[1:3] <- c("UniProtAccession", "UniProtID", "Description")
+fvarLabels(hl)[4:6] <- paste0("Peptides.expt", 1:3)
+fData(hl)[1:4, 1:6]
 ```
@@
 
-Sample meta-data distributed with
-the `r Biocexptpkg("pRolocdata")` and as above with the quantitation
-data file is located in the `extdata` in the `r Biocexptpkg("pRolocdata")` 
+The sample meta-data that is distributed with
+the `r Biocexptpkg("pRolocdata")` package and (as above) the quantitation
+data file are located in the `extdata` in the `r Biocexptpkg("pRolocdata")` 
 package on the hard drive. In the code chunk below we again use the `dir` 
 function to locate the filepath to the meta-data `csv` file and then read 
-it into R using `read.csv` and append the meta-data to the `pData` slot. 
-We add information on the gradient fractions used and the associated 
-subcellular fraction densities in % w/v Iodixanol for each replicate.
+it into R using `read.csv`. We then append the meta-data to the `pData` slot. 
+Information about the gradient fractions used and the associated 
+subcellular fraction densities in % w/v Iodixanol for each replicate
+are stored here.
 
@@
 affecting biological variability (i.e. the shape of the quantitative
-profiles) too much, as generally focused on. This normalisation will
+profiles). This normalisation will
 depend on the underlying quantitative technology and the experimental
-design, and will not be addressed in this workflow. In addition, and
-more specific to spatial proteomics data, it is fundamental to scale
+design, and will not be addressed in this workflow. The second aspect, and
+more specific to spatial proteomics data, is scaling
 all the organelle-specific profiles into a same intensity interval
@@
 The spreadsheet that was used to create the `hl` MSnSet
-included two replicates. We also provide individual replicates in the
+included the two replicates within one .csv file. 
+We also provide individual replicates in the
 `r Biocexptpkg("pRolocdata")` package. Below, we show how to combine
 `MSnSet` objects and, subsequently, how to filter and handle missing
-values. We start by loading the experimental data package and the
+values. We start by loading the `r Biocexptpkg("pRolocdata")` package and the
 equivalent replicates using the `data` function.
@@
 
-To list all data distributed with a package, one can use use the
-`pRolocdata` function.
+At the R prompt, typing
 
@@
 of peptides or peptide-spectrum matches (PSMs).
-vfn
+
 ```{r fvarnames}
@@
 fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
+fvarLabels(hyperLOPIT2015ms3r1)
+fvarLabels(hyperLOPIT2015ms3r2)
 ```
@@
 ```
-
-From examining the literature and the Gene Ontology we find that
-phenotype 1 is enriched in proteins that localise to the nucleus
-(non-chromatin associated proteins), phenotype 2 is enriched in
-chromatin associated proteins, phenotype 3 contains many cytosolic
-proteins and phenotypes 4 and 5 contain lysosmal and endosomal
-proteins. From curating these phenotypes we obtain new marker proteins
-that can be used as input for supervised classification.
 
@@
 ## Plot a boxplot of the scores of each organelle
-par(oma= c(10.5, 0, 0, 0))
+par(oma = c(10.5, 0, 0, 0))
 boxplot(svm.scores ~ svm, data = fData(preds), ylab = "SVM scores", las = 2)
@@
 
-```{r tlopt, eval=FALSE}
+```{r loadtlres, include = FALSE}
+load("Data/tl_runs/tlopt.rda")
+data("hyperLOPIT2015goCC")
+gocc <-hyperLOPIT2015goCC
+ka <- kp <- 3
+```
+```{r setuptlopt, eval=TRUE}
 ## create new markers column for tl markers
@@
 getMarkerClasses(hl, fcol = "tlmarkers")
-
+```
+```{r runtlopts, eval=FALSE}
 ## get best k's
@@
   )
-```
-```{r loadtlres, include = FALSE}
-load("Data/tl_runs/tlopt.rda")
 ```

### 2016-09-07 bioc-workflow.Rmd
added data and updated manuscript

@@
 distinct clusters are observed, we assume that there is organellar
-separation present in the data. In the code chunk below we produce a
-PCA plot of the mouse stem cell dataset. One point on the plot
-represents one protein. We can indeed see several distinct protein
-clusters. Although, representing the multi-dimensional data along a
+separation present in the data. Although, representing the 
+multi-dimensional data along a
 limited set of PCs does not give us a hard quantitative measure of
@@
 
-In the code chunk below, we specify `fcol = NULL`, which means not to
+In the code chunk below we produce a
+PCA plot of the mouse stem cell dataset. One point on the plot
+represents one protein. We can indeed see several distinct protein
+clusters. We specify `fcol = NULL`, which means not to
 consider any feature variable to annotate the features (proteins) with
@@
 
-These markers can then be mapped to an `MSnSet`'s `featureNames`. The
+These markers can then be mapped to a `MSnSet`'s `featureNames`. The
 mouse dataset used here has Uniprot IDs stored as the `featureNames`
-(see `head(featureNames(lopit2016))`) and the names of the vector of
+(see `head(featureNames(hl))`) and the names of the vector of
 the mouse markers stored in `r Biocpkg("pRoloc")` (`mmus` markers) 
@@
 makes conversion between identifers relatively straightforward. If
-this is not the case however, conversion can be performed using the 
+this is not the case however, conversion can be performed using 
 `r Biocpkg("biomaRt")`, the Bioconductor
@@
 We now visualise these annotations along the PCA plot using the
-`plot2D` function. We also use the `addLegend` function to map the
+`plot2D` function and then use the `addLegend` function to map the
 marker classes to the pre-defined colours. We also display the data
@@
 ```{r}
-hl <- fDataToUnknown(hl, from = "Golgi", to = "unknown")
+hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
 getMarkers(hl)
@@
 i <- c(2, 3, 4, 13, 14)
-pData(hl)[i, -3]
+pData(hl)[i, ]
 ```
@@
 presence of the nucleus and nucleus sub-compartments. In the code 
-chunk below, we re-run this analysis, highlighting some of the 
+chunk below, we demonstrate how to do this analysis, highlighting some of the 
 optional arguments and parameters available for phenotype extraction 
@@
 These input markers are stored in the `featureData` column of 
-`hl` where `fcol = "phenoDisoc.Input"`. We can use the convenience 
+`hl` where `fcol = "phenoDisco.Input"`. We can use the convenience 
 accessor function `getMarkers` to print out a table of the markers 
@@
 
-In the code chunk below we run the `phenoDisco` function and 
-return a novelty detection result stored in `pdRes`, according 
+In the code chunk below we demonstrate how to run the `phenoDisco` function and 
+return a novelty detection result, according 
 to the specified parameters. The algorithm parameters `times`, 
 `GS` and `p` are passed to the function, along with the `fcol` 
-telling the algorithm where the input training data is contained.
+to tell the algorithm where the input training data is contained.
 
@@
                  times = 200, 
-                 GS = 60)
-
-```
-
-The argument `times` defined the number of times we run unsupervied 
+                 GS = 20)
+```
+
+*Note: We do not evaluate this code chunk in this document as the algorithm is computational intensive and best parallelised over multiple workers.
+This phenoDisco analysis took ~24 hours to complete when parallelised over 40 workers.*
+
+The argument `times` indicated the number of times we run unsupervied 
 Gaussian Mixture Modelling before defining a new phenotype cluster. 
-The recommended minimum and default value is 100. In the below 
+The recommended minimum and default value is 100. In the above 
 code chunk we increase the value to `times = 200` as we have found 
@@
 detection of small complexes they may wish to use a small `GS = 10`, 
-or `=20` etc. If they wish to detect larger more abundant sub-cellular 
-niches a higher `GS` would be preferable. Specifying a small `GS` is be 
-more time consuming than a larger `GS`, and there is a trade off 
+or `= 20` etc. If they wish to detect larger more abundant sub-cellular 
+niches a much higher `GS` would be preferable. Specifying a small `GS` can be 
+more time consuming than using a larger `GS`, and there is a trade off 
 between finding interesting small complexes and those that may not 
@@
 the number of GMM components to test and fit; the default is `G = 1:9` 
-(`mclust`'s default). One should note that the choice of decreasing 
+(`mclust`'s default). One should note that the decreasing the 
 `GS`, and increasing the values of the arguments `times`, `G` 
@@
 
+```{r loadPDres, include=FALSE}
+load("Data/pd_runs/pd.rda")
+hl <- pdRes
+setStockcol(NULL)
+setStockcol(paste0(getStockcol(), 70))
+```
+
 The ouput of running the `phenoDisco` algorithm is a `MSnSet` containing 
 the new data clusters, appended to the `featureData` under the name 
-`pd`. We can see by typing `pdRes` directly into the console the 
+`pd`. We can see by typing `hl@processingData` directly into the console the 
 processing information has been updated to the `MSnSet` recording the 
@@
 
-```{r loadPDres, include=FALSE}
-load("Data/pd_runs/pd.rda")
-hl <- pdRes
-setStockcol(NULL)
-setStockcol(paste0(getStockcol(), 70))
-```
 ```{r showPDres}
@@
 `pRolocGUI`.
+We found that phenotype 1 was enriched in nucleus associated proteins, phenotype 2 in chromatin associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5 in lysosomal and endosomal proteins.
 

### 2016-09-08 main.tex
Project created

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Welcome to Overleaf --- just edit your article on the left,
% and we'll compile it for you on the right. If you give 
% someone the link to this page, they can edit at the same
% time. See the help menu above for more info. Enjoy!
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% For more detailed article preparation guidelines, please see:
% http://f1000research.com/author-guidelines

\documentclass[10pt,a4paper,twocolumn]{article}
\usepackage{f1000_styles}

%% Default: numerical citations
\usepackage[numbers]{natbib}

%% Uncomment this lines for superscript citations instead
% \usepackage[super]{natbib}

%% Uncomment these lines for author-year citations instead
% \usepackage[round]{natbib}
% \let\cite\citep

\begin{document}

\title{\textit{F1000Research} Article Template}
\titlenote{The title should be detailed enough for someone to know whether the article would be of interest to them, but also concise. Please ensure the broadness and claims within the title are appropriate to the content of the article itself.}
\author[1]{Author Name1}
\author[2]{Author Name2}
\affil[1]{Address of author1}
\affil[2]{Address of author2}

\maketitle
\thispagestyle{fancy}

Please list all authors that played a significant role in the research involved in the article. Please provide full affiliation information (including full institutional address, ZIP code and e-mail address) for all authors, and identify who is/are the corresponding author(s).

\begin{abstract}

Abstracts should be up to 300 words and provide a succinct summary of the article. Although the abstract should explain why the article might be interesting, care should be taken not to inappropriately over-emphasise the importance of the work described in the article. Citations should not be used in the abstract, and the use of abbreviations should be minimized.

\end{abstract}

\clearpage

\section*{Introduction}

The format of the main body of the article is flexible: it should be concise and in the format most appropriate to displaying the content of the article.

Some examples of commonly used \LaTeX{}  commands and features are listed below, to help you get started.

\subsection*{Sections}

Use section and subsection commands to organize your document. \LaTeX{} handles all the formatting and numbering automatically. Use ref and label commands for cross-references.

\subsection*{Tables}

Use the table and tabledata commands for basic tables --- see Table~\ref{tab:widgets}, for example.
\begin{table}[h!]
\hrule \vspace{0.1cm}
\caption{\label{tab:widgets}An example of a simple table with caption.}
\centering
\begin{tabledata}{llr} 
\header First name & Last Name & Grade \\ 
\row John & Doe & $7.5$ \\ 
\row Richard & Miles & $2$ \\ 
\end{tabledata}
\end{table}

\subsection*{Figures}
You can upload a figure (JPEG, PNG or PDF) using the files menu. To include it in your document, use the includegraphics command (see the example below in the source code).

Please give figures appropriate filenames eg: figure1.pdf, figure2.png.

Figure legends should briefly describe the key messages of the figure such that the figure can stand alone from the main text. However, all figures should also be discussed in the article text. Each legend should have a concise title of no more than 15 words. The legend itself should be succinct, while still explaining all symbols and abbreviations. Avoid lengthy descriptions of methods.

For any figures reproduced from another publication (as long as appropriate permission has been obtained from the copyright holder —see under the heading 'Submission'), please include a line in the legend to state that: 'This figure has been reproduced with kind permission from [include original publication citation]'.

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{frog.jpg}
\caption{\label{fig:your-figure}Your figure legend goes here; it should be succinct, while still explaining all symbols and abbreviations. }
\end{figure}

\subsection*{Mathematics}

\LaTeX{} is great at typesetting mathematics. Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
$$S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
      = \frac{1}{n}\sum_{i}^{n} X_i$$
denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.

\subsection*{Lists}

You can make lists with automatic numbering \dots

\begin{enumerate}
\item Like this,
\item and like this.
\end{enumerate}
\dots or bullet points \dots
\begin{itemize}
\item Like this,
\item and like this.
\end{itemize}

\section*{Methods}
This section is not essential for Web Tool papers.

Methods should include a brief discussion of allowances made (if any) for controlling bias or unwanted sources of variability, and the limitations of the datasets.

\section*{Results}
This section is not essential for Web Tool papers.

\section*{Discussion}
The discussion should include the implications of the article results in view of prior work in this field.

\section*{Conclusions}
Please state what you think are the main conclusions that can be realistically drawn from the findings in the paper, taking care not to make claims that cannot be supported.

\subsection*{Author contributions}
In order to give appropriate credit to each author of an article, the individual
contributions of each author to the manuscript should be detailed in this section. We
recommend using author initials and then stating briefly how they contributed.

\subsection*{Competing interests}
All financial, personal, or professional competing interests for any of the authors that
could be construed to unduly influence the content of the article must be disclosed and
will be displayed alongside the article.

\subsection*{Grant information}
Please state who funded the work discussed in this article, whether it is your employer,
a grant funder etc. Please do not list funding that you have that is not relevant to this
specific piece of research. For each funder, please state the funder’s name, the grant
number where applicable, and the individual to whom the grant was assigned.
If your work was not funded by any grants, please include the line: ‘The author(s)
declared that no grants were involved in supporting this work.’

\subsection*{Acknowledgements}
This section should acknowledge anyone who contributed to the research or the
article but who does not qualify as an author based on the criteria provided earlier
(e.g. someone or an organisation that provided writing assistance). Please state how
they contributed; authors should obtain permission to acknowledge from all those
mentioned in the Acknowledgements section.

Please do not list grant funding in this section.

{\small\bibliographystyle{unsrtnat}
\bibliography{sample}}

\bigskip
References can be listed in any standard referencing style that uses a numbering system
(i.e. not Harvard referencing style), and should be consistent between references within
a given article.

Reference management systems such as Zotero provide options for exporting bibliographies as Bib\TeX{} files. Bib\TeX{} is a bibliographic tool that is used with \LaTeX{} to help organize the user's references and create a bibliography. This template contains an example of such a file, \texttt{sample.bib}, which can be replaced with your own. Use the \verb|\cite| command  to create in-text citations, like this \cite{Smith:2012qr} and this \cite{Smith:2013jd}.

% See this guide for more information on BibTeX:
% http://libguides.mit.edu/content.php?pid=55482&sid=406343

% For more author guidance please see:
% http://f1000research.com/author-guidelines

% When all authors are happy with the paper, use the 
% ‘Submit to F1000Research' button from the menu above
% to submit directly to the open life science journal F1000Research.

% Please note that this template results in a draft pre-submission PDF document.
% Articles will be professionally typeset when accepted for publication.

% We hope you find the F1000Research Overleaf template useful,
% please let us know if you have any feedback using the help menu above.

\end{document}

### 2016-09-08 bioc-workflow.Rmd
Make pretty for Laurent
added filterBinMSnSet

@@
 elaborate, expensive and time consuming experimental procedures and
-considerable effort is invested in the generation of such
-data. Multiple research groups have described a variety of approaches
-to establish high quality proteome-wide datasets. However, data
-analysis is as critical as data production for reliable and insightful
+considerable effort is invested in the generation of such data.
+Multiple research groups have described a variety of approaches to
+establish high quality proteome-wide datasets. However, data analysis
+is as critical as data production for reliable and insightful
 biological interpretation. Here, we walk the reader through a typical
@@
 
-The main package to analyse protein localisation data is
-`r Biocpkg("pRoloc")`, which offers a set of dedicated functions for the
-analysis of such data. `r Biocpkg("pRoloc")` itself relies on `r Biocpkg("MSnbase")` to
-manipulate and process quantitative proteomics data. Many other
-packages are used by `r Biocpkg("pRoloc")` for clustering, classification and
-visualisation.
+The main package to analyse protein localisation data is `r
+Biocpkg("pRoloc")`, which offers a set of dedicated functions for the
+analysis of such data. `r Biocpkg("pRoloc")` itself relies on `r
+Biocpkg("MSnbase")` to manipulate and process quantitative proteomics
+data. Many other packages are used by `r Biocpkg("pRoloc")` for
+clustering, classification and visualisation.
 
@@
 data file are located in the `extdata` in the `r Biocexptpkg("pRolocdata")` 
-package on the hard drive. In the code chunk below we again use the `dir` 
-function to locate the filepath to the meta-data `csv` file and then read 
-it into R using `read.csv`. We then append the meta-data to the `pData` slot. 
-Information about the gradient fractions used and the associated 
-subcellular fraction densities in % w/v Iodixanol for each replicate
-are stored here.
+package on the hard drive. In the code chunk below we again use the
+`dir` function to locate the filepath to the meta-data `csv` file and
+then read it into R using `read.csv`. We then append the meta-data to
+the `pData` slot.  Information about the gradient fractions used and
+the associated subcellular fraction densities in % w/v Iodixanol for
+each replicate are stored here.
 
@@
 affecting biological variability (i.e. the shape of the quantitative
-profiles). This normalisation will
-depend on the underlying quantitative technology and the experimental
-design, and will not be addressed in this workflow. The second aspect, and
-more specific to spatial proteomics data, is scaling
-all the organelle-specific profiles into a same intensity interval
-(typically 0 and 1) by, for example, diving each intensity by the sum
-of the intensities for that quantitative feature. This is not necessary 
-in this example as the intensities for each replicate have already 
-been re-scaled to 1 in Proteome Discoverer. However, if one wanted
-to do this they would execute the `normalise` function as demonstrated 
-in the below code chunk.
-
-```{r normsum, eval = FALSE}
-hl <- normalise(hl, method = "sum")
-```
+profiles). This normalisation will depend on the underlying
+quantitative technology and the experimental design, and will not be
+addressed in this workflow. The second aspect, and more specific to
+spatial proteomics data, is scaling all the organelle-specific
+profiles into a same intensity interval (typically 0 and 1) by, for
+example, diving each intensity by the sum of the intensities for that
+quantitative feature. This is not necessary in this example as the
+intensities for each replicate have already been re-scaled to 1 in
+Proteome Discoverer. However, if one wanted to do this they would
+execute the `normalise` function as demonstrated in the below code
+chunk.
+
+```{r normsum, eval = FALSE} hl <- normalise(hl, method = "sum") ```
 
@@
 The same `normalise` function (or `normalize`, both spellings are
-supported) can also be applied in the first case described
-above. Different normalisation methods such as mean or median scaling,
+supported) can also be applied in the first case described above.
+Different normalisation methods such as mean or median scaling,
 variance stabilisation or quantile normalisation, to cite a few, can
@@
 
-As previously mentioned, before combination, the two replicates in 
-the `hl` data that we read into R were separately normalised 
-by sum (i.e. to 1) across the 10 channels for each replicate 
-respectively. We can verify this by summing each rows for each replicate:
+As previously mentioned, before combination, the two replicates in the
+`hl` data that we read into R were separately normalised by sum (i.e.
+to 1) across the 10 channels for each replicate respectively. We can
+verify this by summing each rows for each replicate:
 
@@
 
-The spreadsheet that was used to create the `hl` MSnSet
-included the two replicates within one .csv file. 
-We also provide individual replicates in the
-`r Biocexptpkg("pRolocdata")` package. Below, we show how to combine
-`MSnSet` objects and, subsequently, how to filter and handle missing
-values. We start by loading the `r Biocexptpkg("pRolocdata")` package and the
-equivalent replicates using the `data` function.
+The spreadsheet that was used to create the `hl` MSnSet included the
+two replicates within one .csv file.  We also provide individual
+replicates in the `r Biocexptpkg("pRolocdata")` package. Below, we
+show how to combine `MSnSet` objects and, subsequently, how to filter
+and handle missing values. We start by loading the `r
+Biocexptpkg("pRolocdata")` package and the equivalent replicates using
+the `data` function.
 
@@
 separation present in the data. Although, representing the
-multi-dimensional data along a
-limited set of PCs does not give us a hard quantitative measure of
-separation, it is extremely useful summarising complex experimental
-information in one figure, to get an simplified overview of the data.
-
-In the code chunk below we produce a
-PCA plot of the mouse stem cell dataset. One point on the plot
-represents one protein. We can indeed see several distinct protein
-clusters. We specify `fcol = NULL`, which means not to
-consider any feature variable to annotate the features (proteins) with
-colours. We will see later how to use this to annotate the PCA plot
-with prior information about sub-cellular localisation.
+multi-dimensional data along a limited set of PCs does not give us a
+hard quantitative measure of separation, it is extremely useful
+summarising complex experimental information in one figure, to get an
+simplified overview of the data.
+
+In the code chunk below we produce a PCA plot of the mouse stem cell
+dataset. One point on the plot represents one protein. We can indeed
+see several distinct protein clusters. We specify `fcol = NULL`, which
+means not to consider any feature variable to annotate the features
+(proteins) with colours. We will see later how to use this to annotate
+the PCA plot with prior information about sub-cellular localisation.
 
@@
 annotation. The identification of well resolved clusters in the data,
-constitutes an unbiased assessment of the data structure, demonstrating 
-the successful separation of sub-cellular clusters.
+constitutes an unbiased assessment of the data structure,
+demonstrating the successful separation of sub-cellular clusters.
 
@@
 the protein profiles occupancy along the gradient (we also display the
-mean fraction intensities) and a `boxplot` of the column
-intensities. In the two plots displayed below, we re-order the
-fractions to pair corresponding fractions in the two replicates
-(rather than ordering the fractions by replicate).
-
-```{r qcbx, fig.width = 12}
-par(mfrow = c(1, 2))
-o <- order(hl$Iodixonal.Density)
-plotDist(hl[, o], pcol = "#00000010")
+mean fraction intensities) and a `boxplot` of the column intensities.
+In the two plots displayed below, we re-order the fractions to pair
+corresponding fractions in the two replicates (rather than ordering
+the fractions by replicate).
+
+```{r qcbx, fig.width = 12} par(mfrow = c(1, 2)) o <-
+order(hl$Iodixonal.Density) plotDist(hl[, o], pcol = "#00000010")
 lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
-boxplot(exprs(hl[, o]), las = 2)
-```
+boxplot(exprs(hl[, o]), las = 2) ```
 
@@
 
-```{r plot2Drep, fig.width = 12}
-par(mfrow = c(1, 2))
-plot2D(hl[, hl$Replicate == 1], main = "Replicate 1")
-plot2D(hl[, hl$Replicate == 2], main = "Replicate 2",
-       mirrorX = TRUE)
-```
+```{r plot2Drep, fig.width = 12} par(mfrow = c(1, 2)) plot2D(hl[,
+hl$Replicate == 1], main = "Replicate 1") plot2D(hl[, hl$Replicate ==
+2], main = "Replicate 2", mirrorX = TRUE) ```
 
@@
 are also useful for confirming the presence of known or postulated
-clusters in an unbiased fashion. For example, in [@hyper] 
-the `phenoDisco` algorithm was used to confirm the data-specific
-presence of the nucleus and nucleus sub-compartments. In the code 
-chunk below, we demonstrate how to do this analysis, highlighting some of the 
+clusters in an unbiased fashion. For example, in [@hyper] the
+`phenoDisco` algorithm was used to confirm the data-specific presence
+of the nucleus and nucleus sub-compartments. In the code chunk below,
+we demonstrate how to do this analysis, highlighting some of the
 optional arguments and parameters available for phenotype extraction
@@
 
-As the `phenoDisco` algorithm is semi-supervised it uses both 
-labelled (markers) and unlabelled data to explore the data structure 
-and find new sub-cellular data clusters. Thus the first step is to 
-define some input labelled data i.e. markers, that the algorithm 
-will use as input for the supervised learning aspect of the 
-algorithm. As described in [@hyper] we define a set of markers 
-to use as input for the analyses that cover well-known residents 
-from three distinct organelle structures; the mitochondria, plasma 
-membrane and ER, and from three well-known and abundant protein 
-complexes; the proteasome and two ribosomal subunits, 40S and 60S. 
-These input markers are stored in the `featureData` column of 
-`hl` where `fcol = "phenoDisco.Input"`. We can use the convenience 
-accessor function `getMarkers` to print out a table of the markers 
-contained in this marker set. These initial markers were manually 
-curated using information from the UniProt database, the Gene 
-Ontology and the literature.
+As the `phenoDisco` algorithm is semi-supervised it uses both labelled
+(markers) and unlabelled data to explore the data structure and find
+new sub-cellular data clusters. Thus the first step is to define some
+input labelled data i.e. markers, that the algorithm will use as input
+for the supervised learning aspect of the algorithm. As described in
+[@hyper] we define a set of markers to use as input for the analyses
+that cover well-known residents from three distinct organelle
+structures; the mitochondria, plasma membrane and ER, and from three
+well-known and abundant protein complexes; the proteasome and two
+ribosomal subunits, 40S and 60S.  These input markers are stored in
+the `featureData` column of `hl` where `fcol = "phenoDisco.Input"`. We
+can use the convenience accessor function `getMarkers` to print out a
+table of the markers contained in this marker set. These initial
+markers were manually curated using information from the UniProt
+database, the Gene Ontology and the literature.
 
@@
 
-In the code chunk below we demonstrate how to run the `phenoDisco` function and 
-return a novelty detection result, according 
-to the specified parameters. The algorithm parameters `times`, 
-`GS` and `p` are passed to the function, along with the `fcol` 
-to tell the algorithm where the input training data is contained.
+In the code chunk below we demonstrate how to run the `phenoDisco`
+function and return a novelty detection result, according to the
+specified parameters. The algorithm parameters `times`, `GS` and `p`
+are passed to the function, along with the `fcol` to tell the
+algorithm where the input training data is contained.
 
@@
                  times = 200, 
-                 GS = 20)
-```
-
-*Note: We do not evaluate this code chunk in this document as the algorithm is computational intensive and best parallelised over multiple workers.
-This phenoDisco analysis took ~24 hours to complete when parallelised over 40 workers.*
+                 GS = 60)
+```
+
+*Note: We do not evaluate this code chunk in this document as the
+algorithm is computational intensive and best parallelised over
+multiple workers. This phenoDisco analysis took ~24 hours to complete
+when parallelised over 40 workers.*
 
@@
 Gaussian Mixture Modelling before defining a new phenotype cluster.
-The recommended minimum and default value is 100. In the above 
-code chunk we increase the value to `times = 200` as we have found 
-for larger datasets (e.g. 5000+ proteins) a higher times is requried 
-for convergence. `GS` defines the minimum number of proteins allowed 
-per new data cluster and thus heavily influences what type of new 
-clusters are extracted. For example, if a user is interesed in the 
-detection of small complexes they may wish to use a small `GS = 10`, 
-or `= 20` etc. If they wish to detect larger more abundant sub-cellular 
-niches a much higher `GS` would be preferable. Specifying a small `GS` can be 
-more time consuming than using a larger `GS`, and there is a trade off 
-between finding interesting small complexes and those that may not 
-be of interest as we find there is a tendancy to find more noise when 
+The recommended minimum and default value is 100. In the above code
+chunk we increase the value to `times = 200` as we have found for
+larger datasets (e.g. 5000+ proteins) a higher times is requried for
+convergence. `GS` defines the minimum number of proteins allowed per
+new data cluster and thus heavily influences what type of new clusters
+are extracted. For example, if a user is interesed in the detection of
+small complexes they may wish to use a small `GS = 10`, or `= 20` etc.
+If they wish to detect larger more abundant sub-cellular niches a much
+higher `GS` would be preferable. Specifying a small `GS` can be more
+time consuming than using a larger `GS`, and there is a trade off
+between finding interesting small complexes and those that may not be
+of interest as we find there is a tendancy to find more noise when
 using a small `GS` compared to using a higher one.
@@
 the number of GMM components to test and fit; the default is `G = 1:9`
-(`mclust`'s default). One should note that the decreasing the 
-`GS`, and increasing the values of the arguments `times`, `G` 
-(among other function arguments, see `?phenoDisco`) will heavily 
-influence (increase) the total time taken to run the algorithm. 
-`phenoDisco` supports parallelisation and we strongly suggest 
-you make use of a parallel processing to run these analyses (different 
-backends can be set with argument `BPPARAM`).
+(`mclust`'s default). One should note that the decreasing the `GS`,
+and increasing the values of the arguments `times`, `G` (among other
+function arguments, see `?phenoDisco`) will heavily influence
+(increase) the total time taken to run the algorithm.  `phenoDisco`
+supports parallelisation and we strongly suggest you make use of a
+parallel processing to run these analyses (different backends can be
+set with argument `BPPARAM`).
 
@@
 
-The ouput of running the `phenoDisco` algorithm is a `MSnSet` containing 
-the new data clusters, appended to the `featureData` under the name 
-`pd`. We can see by typing `hl@processingData` directly into the console the 
-processing information has been updated to the `MSnSet` recording the 
-parameters that were used to run the analyses. This is handy for keeping 
-track of data analyses. The results can be displayed by using the 
-`getMarkers` function. We see that 5 new phenotype data clusters were 
-found.
+The ouput of running the `phenoDisco` algorithm is a `MSnSet`
+containing the new data clusters, appended to the `featureData` under
+the name `pd`. We can see by typing `hl@processingData` directly into
+the console the processing information has been updated to the
+`MSnSet` recording the parameters that were used to run the analyses.
+This is handy for keeping track of data analyses. The results can be
+displayed by using the `getMarkers` function. We see that 5 new
+phenotype data clusters were found.
 
@@
 their protein profiles by using the `pRolocVis` function in the
-`pRolocGUI`.
-We found that phenotype 1 was enriched in nucleus associated proteins, phenotype 2 in chromatin associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5 in lysosomal and endosomal proteins.
+`pRolocGUI`. We found that phenotype 1 was enriched in nucleus
+associated proteins, phenotype 2 in chromatin associated proteins,
+phenotype 3 in cytosolic and phenotypes 4 and 5 in lysosomal and
+endosomal proteins.
 
@@
 the labelled marker proteins. These class/organelle boundaries define
-how non-assigned proteins are classified and with what
-confidence. **TODO** explain how choosing markers
-positively/negatively affects this, and relation with thresholding.
+how non-assigned proteins are classified and with what confidence.
+**TODO** explain how choosing markers positively/negatively affects
+this, and relation with thresholding.
 
@@
 niches can exhibit different score distributions, as highlighted in
-the boxplot below. We recommend users to set class-specific thresholds.
-In the code chunk below we display a boxplot of the score distributions 
-per organelle.
+the boxplot below. We recommend users to set class-specific
+thresholds.  In the code chunk below we display a boxplot of the score
+distributions per organelle.
 
@@
 
-There are many ways to set thresholds and the choice of method will depend
-on the biological question and experimental design at hand. One viable
-approach in the frame of the above experimetal design would be to manually
-set a FDR, say 5%, per organelle. To do this the user would examine 
-the top scoring predictions for each organelle, and then set a threshold at
-the score at which they achieve 5% of false assignments per organelle. 
-The definintion of a false assignment would depend on the information 
-available, for example, validity or lack of validity for the localisation
-from another experiment as reported in the literature or a reliable database. 
-If such information is not available, one crude method is to set a threshold  
-per organelle by extracting the median or 3rd quantile score per organelle. 
-For example, in the code chunk below, we use the `orgQuants` function
-to extract the median organelle scores and then pass these scores to the 
-`getPrediction` function to extract the new localisations that meet this
-scoring criteria. Any sub-cellular predictions that fall below the specified
-thresholds are labelled as unknown.
+There are many ways to set thresholds and the choice of method will
+depend on the biological question and experimental design at hand. One
+viable approach in the frame of the above experimetal design would be
+to manually set a FDR, say 5%, per organelle. To do this the user
+would examine the top scoring predictions for each organelle, and then
+set a threshold at the score at which they achieve 5% of false
+assignments per organelle.  The definintion of a false assignment
+would depend on the information available, for example, validity or
+lack of validity for the localisation from another experiment as
+reported in the literature or a reliable database.  If such
+information is not available, one crude method is to set a threshold
+per organelle by extracting the median or 3rd quantile score per
+organelle.  For example, in the code chunk below, we use the
+`orgQuants` function to extract the median organelle scores and then
+pass these scores to the `getPrediction` function to extract the new
+localisations that meet this scoring criteria. Any sub-cellular
+predictions that fall below the specified thresholds are labelled as
+unknown.
 
@@
 
-The output of `getPredictons` is the original `MSnSet` dataset with a new feature variable appended to the feature data called `fcol.pred` (i.e.
-in our case `svm.pred`) containing the prediction results.
-The results can also be visualied using `plot2D` function.
+The output of `getPredictons` is the original `MSnSet` dataset with a
+new feature variable appended to the feature data called `fcol.pred`
+(i.e. in our case `svm.pred`) containing the prediction results. The
+results can also be visualied using `plot2D` function.
 
@@
 
-The data is loaded and displayed on a PCA plot and a boxplot is used to
-display the classifier scores by data class. On the left there is
-a sidebar panel with sliders to control the thresholds upon which classifications 
-are made. There are two types of cut-off that the user can choose from: (1)
-"Quantile" and (2) "User-defined". By default, when the application is launched
-quatile scoring is selected and set to 0.5, the median. The 
-class-specific score thresholds that correspond to selecting the desired
-quantile are shown on as red dots on the boxplot. The assignments on the PCA
-plot are also updated according to the selected threshold. The quantile 
-threshold can be set by moving the corresponding quantile slider. If one
-wished to set their own cut-offs the "User-defined" radio button must be selected
-and then the sliders for defining user-specified scores become active and the
-scores and highlighted on the boxplot by blue dots. For more information we refer
-users to the `r Biocpkg("pRolocGUI")` tutorial 
+The data is loaded and displayed on a PCA plot and a boxplot is used
+to display the classifier scores by data class. On the left there is a
+sidebar panel with sliders to control the thresholds upon which
+classifications are made. There are two types of cut-off that the user
+can choose from: (1) "Quantile" and (2) "User-defined". By default,
+when the application is launched quatile scoring is selected and set
+to 0.5, the median. The class-specific score thresholds that
+correspond to selecting the desired quantile are shown on as red dots
+on the boxplot. The assignments on the PCA plot are also updated
+according to the selected threshold. The quantile threshold can be set
+by moving the corresponding quantile slider. If one wished to set
+their own cut-offs the "User-defined" radio button must be selected
+and then the sliders for defining user-specified scores become active
+and the scores and highlighted on the boxplot by blue dots. For more
+information we refer users to the `r Biocpkg("pRolocGUI")` tutorial 
 [vignette](http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html).
@@
                   evidence = NULL)
-gocc
+```
+
+The function `makeGoSet` uses the `r Biocpkg("biomaRt")` package to query the relevent database (e.g.
+Ensembl, Uniprot) for GO terms.
+All GO terms that have been observed for the 5032 proteins in the hyperLOPIT dataset are retieved.
+Users should note that the number of GO terms retreived is also dependent on the database version queried and thus is always subject to change.
+We find it is common to see many GO terms with only one protein assigned to that term, these such terms generally do not bring and information generally for building the classifier and thus we can remove such GO terms by using the function `filterBinMSnSet`.
+
+```{r filtergo, eval=FALSE}
+gocc <- filterBinMSnSet(hl)
 ```
@@
 are the $k$'s for the primary and auxiliary data sources required for
-the nearest neighbour calculations for each data source.  The second
+the nearest neighbour calculations for each data source.
+
+```{r knnoptfortl, eval=FALSE}
+## get best k's
+kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
+kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
+kp <- getParams(kpopt)
+ka <- getParams(kaopt)
+```
+
+The second
 set of parameters (which is noted by a vector of $\theta$ weights)
@@
 ```
+
 ```{r setuptlopt, eval=TRUE}

### 2016-09-08 bioc-workflow.Rmd
added tlres from pRolocdata

@@
 are the $k$'s for the primary and auxiliary data sources required for
-the nearest neighbour calculations for each data source.
-
-```{r knnoptfortl, eval=FALSE}
-## get best k's
-kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
-kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
-kp <- getParams(kpopt)
-ka <- getParams(kaopt)
-```
-
-The second
+the nearest neighbour calculations for each data source. The second
 set of parameters (which is noted by a vector of $\theta$ weights)
@@
 have `4^10` class weight combinations to consider instead of `4^14`
-combinations. In the example below we remove these 4 classes from the
+combinations.
+
+In the example below we remove these 4 classes from the
 marker set, re-run the `knnOptimisation` for each data source and then
@@
 
-In the example code below we first remove the 2 ribosomes and nuclear
-classes and then run the `knnOptimisation` followed by the
-`knntlOptimisation` (specifying snow with a MPI
-backend with 255 workers for parallelisation on the Cambridge HPC).
-
-```{r loadtlres, include = FALSE}
-load("Data/tl_runs/tlopt.rda")
-data("hyperLOPIT2015goCC")
-gocc <-hyperLOPIT2015goCC
-ka <- kp <- 3
-```
-
+To remove the 4 classes and create a new column of markers in the feature data called `tlmarkers` to use for the analysis:
 ```{r setuptlopt, eval=TRUE}
@@
 ```
+
+Optimisation stage 1: run `knnOptimisation` to get the best $k$'s for each data source.
+```{r knnoptfortl, eval=FALSE}
+## get best k's
+kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
+kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
+kp <- getParams(kpopt)
+ka <- getParams(kaopt)
+```
+
+Optimisation stage 2: run `knntlOptimisation` to get the best transfer learning weights for each sub-cellular class.
 ```{r runtlopts, eval=FALSE}
-## get best k's
-kp <- knnOptimisation(hl, fcol = "tlmarkers")
-ka <- knnOptimisation(gocc, fcol = "tlmarkers")
-
-## Set appropriate backend and workers for the tl
+## Set appropriate parallelisation backend and  
+## number of workers for the tl
 par <- SnowParam(255L, type = "MPI")
@@
 
+```{r loadtlres, echo=FALSE}
+tlfile <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-tlopt.rds")
+tlopt <- readRDS(tlfile)
+
+```
 ```{r plottl}

### 2016-09-08 bioc-workflow.Rmd
udpated pd section

@@
 
-In the code chunk below we demonstrate how to run the `phenoDisco`
+In the code chunk below we show how to run the `phenoDisco`
 function and return a novelty detection result, according to the
@@
 
-```{r loadPDres, include=FALSE}
-load("Data/pd_runs/pd.rda")
-hl <- pdRes
+```{r loadPDres, echo=FALSE}
+f0 <- dir(extdatadir, full.names = TRUE,
+          pattern = "bpw-pdres.rds")
+pdres <- readRDS(f0)
+hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
 setStockcol(NULL)

### 2016-09-08 bioc-workflow.Rmd
added svm params

@@
 
-```{r loadpdRes, include=FALSE}
-load("Data/fusionparams.rda")
-w <- table(getMarkers(hl, verbose = TRUE))
-w <- 1/w[names(w) != "unknown"]
-```
-
+```{r loadpdRes, echo=FALSE}
+svmf <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-svmopt.rds") 
+params <- readRDS(svmf)
+```
 ```{r optimise, eval=FALSE}

### 2016-09-08 bioc-workflow.Rmd
all files loaded correctly

@@
 fvarLabels(hl)
-fvarLabels(hl)[1:3] <- c("UniProtAccession", "UniProtID", "Description")
-fvarLabels(hl)[4:6] <- paste0("Peptides.expt", 1:3)
+fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
+fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
 fData(hl)[1:4, 1:6]
@@
 
-```{r}
+```{r removeGA}
 hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
@@
 
-```{r plot2Drep, fig.width = 12} par(mfrow = c(1, 2)) plot2D(hl[,
-hl$Replicate == 1], main = "Replicate 1") plot2D(hl[, hl$Replicate ==
-2], main = "Replicate 2", mirrorX = TRUE) ```
+```{r plot2Drep, fig.width = 12} 
+par(mfrow = c(1, 2)) 
+plot2D(hl[, hl$Replicate == 1], main = "Replicate 1") 
+plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE) 
+```
 
@@
 par(mfrow = c(1, 2))
-plot2D(hl, fcol = "phenodisco.input", 
+plot2D(hl, fcol = "phenoDisco.Input", 
        main = "phenoDisco input markers", col = getStockcol()[1:6])
-addLegend(hl, fcol = "phenodisco.input", cex = .7)
+addLegend(hl, fcol = "phenoDisco.Input", cex = .7)
 plot2D(hl, fcol = "pd", main = "phenoDisco output", col = cols)
@@
 ```
-```{r optimise, eval=FALSE}
+```{r getsvmweight}
 w <- table(getMarkers(hl, verbose = TRUE))
 w <- 1/w[names(w) != "unknown"]
-
+```
+```{r dosvmopt, eval=FALSE}
 ## 100 rounds of optimisation with five-fold cross-validation
@@
 To remove the 4 classes and create a new column of markers in the feature data called `tlmarkers` to use for the analysis:
-```{r setuptlopt, eval=TRUE}
+```{r loadgocc, include=FALSE}
+gofile <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-gocc.rds")
+gocc <- readRDS(gofile)
+```
+```{r setuptlopt}
 ## create new markers column for tl markers
@@
   hl <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
-  gocc <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
+  gocc <- fDataToUnknown(gocc, from = torm[i], fcol = "tlmarkers")
 }
 getMarkerClasses(hl, fcol = "tlmarkers")
-getMarkerClasses(hl, fcol = "tlmarkers")
+getMarkerClasses(gocc, fcol = "tlmarkers")
 ```
@@
 kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
-kp <- getParams(kpopt)
-ka <- getParams(kaopt)
-```
+plot(kpopt)
+plot(kaopt)
+```
+From examining the parameter seach plots we find the best $k$'s for both the primary and auxiliary are 3.
 
@@
                            times = 50,
-                           xval = 5, k = c(kp,ka),
+                           xval = 5, k = c(3, 3),
                            BPPARAM = par
@@
 tlopt <- readRDS(tlfile)
-
+colnames(tlopt@results)[4] <- "Endoplasmic reticulum"
+setStockcol(NULL)
 ```
@@
 hl <- knntlClassification(hl, gocc, bestTheta = bestpar, 
-                          k = c(kp, ka))
-
+                          k = c(3, 3))
 ```

### 2016-09-15 bioc-workflow.Rmd

@@
 
-```{r normsum, eval = FALSE} hl <- normalise(hl, method = "sum") ```
+```{r normsum, eval = FALSE} 
+hl <- normalise(hl, method = "sum") 
+```
 
@@
 
-```{r classify, eval=TRUE}
+```{r classify}
 hl <- svmClassification(hl, params,
@@
 ```
-```{r plotSVM, eval=TRUE}
+```{r plotSVM}
 ## set point size of each protein to be inversely proportional to the

### 2016-09-23 bioc-workflow.Rmd
added section on unsupervised ML

@@
 
+## Unsupervised machine learning
+In `pRoloc` there is functionality for unsupervsied machine learning
+methods. In unsupervised learning, the training data consists of a
+set of input vectors e.g. protein profiles, for which we do not have
+information about the class label e.g. localisation. The main goal in
+unsupervised learning is to uncover groups of similar examples within
+the data, this is termed clustering. Ordinance methods such as
+principal components analysis (PCA) also fall into the category of
+unsupervised learning methods, where the data can be projected from a
+high-dimensional spcae down to two or three dimensions for the purpose
+of visualisation.
+
+As described and demonstrated already above, PCA is a valuable and
+powerful method for data visualisation and quality control. We do not
+do any unsupervised clustering in the frame of the above spatial
+proteomics workflow as although clustering can be useful for grouping
+labelled data into categories e.g. see the `mrkHClust` function, we do
+not find it adaquete for protein localisation prediction and spatial
+proteomics data analysis in general.
+
 # Session information

### 2016-09-23 bioc-workflow.Rmd
more on unsupervised ML

@@
 labelled data into categories e.g. see the `mrkHClust` function, we do
-not find it adaquete for protein localisation prediction and spatial
-proteomics data analysis in general.
+not find it adaquete for spatial proteomics data analysis. We find
+supervised learning more suited to the task of protein localisation
+prediction in which we use high-quality curated marker proteins to
+build a classifier, instead of using an entirely unsupervised approach
+to look for clusters and then look for enrichment of organelles and
+complexes. In the latter we do not make good use of valuable prior
+knowledge, and in our experience unsupervised clustering can be
+extremely difficult due poor estimates of the number of clusters that
+may appear in the data.
 

### 2016-09-29 bioc-workflow.Rmd

@@
 emphasis separation of the denser organelles.  The intersect of
-replicates 1 and 2 was treated as a 20-plex dataset for the analysis
-discussed in the manuscript [@hyper] as it has been shown that
+replicates 1 and 2 was treated as a 20-plex dataset for the analysis.
+As discussed in the manuscript [@hyper] as it has been shown that
 combining replicates across from different gradients can increase
@@
 manuscript (see tab 2 of the `xlsx` supplementary data set 1 in
-[@hyper], which should be exported as a text-based spreadsheet) and
-also as a `csv` in the Bioconductor `r Biocexptpkg("pRolocdata")` data
-package.
+[@hyper], which should be exported as a text-based spreadsheet). It
+is also available as a `csv` in the Bioconductor `r
+Biocexptpkg("pRolocdata")` data package.
 
@@
 
-Below, we examine the quantitative information for
-first 5 proteins. It is also possible to access specific rows and
-columns by naming the proteins and fractions of interest.
+Below, we examine the quantitative information for first 5 proteins.
+It is also possible to access specific rows and columns by naming the
+proteins and TMT tag channels of interest.
 
@@
 
-The feature meta-data is stored in the `fData`
-slot and can be accessed by `fData(hl)`. When using
-`readMSnSet2`, automatically, everything that is not defined as
-quantitation data by `ecol` or the feature names by `fnames` is
-deposited to the `fData` slot.
-
-We see the `fData`
-contains 25 columns describing information such as the number of
-peptides, associated markers, machine learning results etc. To
-identify the feature variable names we can use the function 
-`fvarLabels`. We see that the first 6 feature variable names 
-contain non-discriminatory label names, so we relabel them to
-help us identify what feature data information is stored in
-the associated columns.
+The feature meta-data is stored in the `fData` slot and can be
+accessed by `fData(hl)`. When using `readMSnSet2`, automatically,
+everything that is not defined as quantitation data by `ecol` or the
+feature names by `fnames` is deposited to the `fData` slot.
+
+We see the `fData` contains 25 columns describing information such as
+the number of peptides, associated markers, machine learning results
+etc. To identify the feature variable names we can use the function
+`fvarLabels`. We see that the first 6 feature variable names contain
+non-discriminatory label names, so we relabel them to help us identify
+what feature data information is stored in the associated columns.
 
@@
 
-Note that when using the simple `readMSnSet2` procedure, sample
-meta-data is kept empty. It is advised to annotate the fractions as
+Note that when using the simple `readMSnSet2` procedure, the `pData`
+slot which is used to store information about the samples/channels
+is kept empty. It is advised to annotate the channels as
 well. Below, we annotate the replicate from which the profiles
-originate and the TMT tag (extracted from the sample/fraction names).
+originate and the TMT tag (extracted from the sample/channel names).
 
@@
 
+Throughout this workflow we refer to the different columns that
+are found in the `exprs` (expression data) slot as channels (short for TMT channels).
+In the frame of LOPIT and hyperLOPIT these channels consititue the
+relative abundance of each protein (along the rows) in the channel of
+interest. Each TMT channel originates from fractions collected from
+the density gradient, or a set of pooled fractions or may be a sample
+originating from an alternative preparation e.g. such as from the
+chromatin enrichment performed in Christoforou et al [@hyper].
+Information about which gradient fractions were used for which tag
+should also be stored in the sample meta-data `pData` slot.
+
 The sample meta-data that is distributed with
-the `r Biocexptpkg("pRolocdata")` package and (as above) the quantitation
-data file are located in the `extdata` in the `r Biocexptpkg("pRolocdata")` 
-package on the hard drive. In the code chunk below we again use the
+the `r Biocexptpkg("pRolocdata")` package for Christoforou's
+hyperLOPIT experiment and (as above) the quantitation
+data file, are located in the `extdata` in the `r Biocexptpkg("pRolocdata")` 
+package on the hard drive.
+
+In the code chunk below we again use the
 `dir` function to locate the filepath to the meta-data `csv` file and
@@
 to spatial proteomics data processing. The first one focuses on
-reducing purely technical variation between fractions without
+reducing purely technical variation between channels without
 affecting biological variability (i.e. the shape of the quantitative
@@
 focus subsequent analyses on the relative profiles along the
-sub-cellular fractions.
+sub-cellular channels.
 
@@
 the data. As we want our replicates to be combined along the columns
-(same proteins, different sets of fractions), we need to assure that
+(same proteins, different sets of channels), we need to assure that
 the respective sample names differ so they can be identified from one
@@
 It is also useful to visualise the relative intensities along the
-gradient to identify fractions displaying particularly low yield. This
+gradient to identify channels displaying particularly low yield. This
 can be done using the `plotDist` and `boxplot` functions, that plot
 the protein profiles occupancy along the gradient (we also display the
-mean fraction intensities) and a `boxplot` of the column intensities.
-In the two plots displayed below, we re-order the fractions to pair
-corresponding fractions in the two replicates (rather than ordering
-the fractions by replicate).
+mean channel intensities) and a `boxplot` of the column intensities.
+In the two plots displayed below, we re-order the TMT channles to pair
+corresponding channels in the two replicates (rather than ordering
+the channels by replicate).
 
@@
 few marker clusters is useful to assess how exactly they differ (in
-terms of peak fractions, for example). Below, we plot the profile of
+terms of peak channels, for example). Below, we plot the profile of
 the mitochondrial and peroxisome markers to highlight the differences
 in profiles between these two sets of markers along the 6th and 7th
-fractions, as represented above along the 7th PC on the PCA plot.
+channels, as represented above along the 7th PC on the PCA plot.
 
@@
 
-Direct comparisons of individual fractions in replicated experiments
+Direct comparisons of individual channels in replicated experiments
 does not provide an adequate, goal-driven assessment of different
-experiments. Indeed, due to the nature of the experiment and fraction
-collection, the quantitative channels (loosely referred to as
-fractions) do not correspond to identical actual fractions along the
+experiments. Indeed, due to the nature of the experiment and gradient fraction
+collection, the quantitative channels 
+do not correspond to identical selected fractions along the
 gradient. As can be seen in the table below (taken from the
 `hyperLOPIT`'s `pData`, focusing on channels 7 to 10), different sets
-of fractions are combined to obtain enough material and optimise
+of gradient fractions are pooled to obtain enough material and optimise
 acurate quantitation.
@@
 
-The more relevant comparison unit is not a single fraction, but rather
+The more relevant comparison unit is not a single channel, but rather
 the complete protein occupancy profiles, which are best visualised as
@@
 
+# Writing and exporting data
+A `MSnSet` can be exported from R using the `write.exprs` function.
+This function writes the expression values to a tab (the defualt, or
+other type e.g.e.g. comma, as specified in `write,table`) separated
+file. There argument `fDataCols` can be used to specify which
+`featureData` columns (as column names, column number or `logical`) to
+append to the right of the expression matrix.
+
+In the below code chunk we write the `hl` object to a csv file. The
+`file` argument is used to specify the file path, the `sep` argument
+specifies the field separator string, here we use a comma, finally as
+we want to write all the information in the `featureData` to the file,
+as well as the expression data, we specify `fDataCols = 1:ncol(hl)`
+i.e. write everything in the `featureData` to the file `"hl.csv"`.
+
+```{r writeData, eval = FALSE}
+write.exprs(hl, file = "hl.csv", sep = ",", 
+            fDataCols = 1:ncol(hl))
+```
+
+Individual R objects, such a parameters from the machine learning
+optimisations, can be save using the standard `save` function in R.
+For example, to save and then re-load the parameters from the SVM
+optimisation,
+
+```{r saveData, eval = FALSE}
+## To save the parameters as an R object
+save(params, file = "svmparams.rda")
+
+## To re-load after saving
+load(file = "svmparams.rda")
+```
+
 # Session information

### 2016-09-29 bioc-workflow.tex
Update on Overleaf.

\section{Introduction}\label{introduction}

Quantitative mass spectrometry based spatial proteomics involves
elaborate, expensive and time consuming experimental procedures and
considerable effort is invested in the generation of such data. Multiple
research groups have described a variety of approaches to establish high
quality proteome-wide datasets. However, data analysis is as critical as
data production for reliable and insightful biological interpretation.
Here, we walk the reader through a typical pipeline for the analysis of
such data using several Bioconductor packages for the R statistical
programming environment.

The main package to analyse protein localisation data is
\texttt{r Biocpkg("pRoloc")}, which offers a set of dedicated functions
for the analysis of such data.
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} itself
relies on \texttt{r Biocpkg("MSnbase")} to manipulate and process
quantitative proteomics data. Many other packages are used by
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} for
clustering, classification and visualisation.

In this workflow, we will describe how to prepare the spatial proteomics
data starting from a spreadsheet containing quantitative mass
spectrometry data. We will focus on a recent pluripotent mouse embryonic
stem cells experiment {[}@hyper{]}. Additional annotated and
pre-formatted datasets from various species are readily available in the
\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
package.

Installation of Bioconductor package is documented in details on the
\href{http://bioconductor.org/install/\#install-bioconductor-packages}{Bioconductor
installation help page}. This procedure is also applicable to any
packages, from \href{https://cran.r-project.org/}{CRAN} as well as
GitHub. Once a package has been installed, it needs to be loaded for its
functionality to become available in the R session; this is done with
the \texttt{library} function e.g. to load the
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} one would
type \texttt{library("pRoloc")} after installation.

If you have questions about this workflow in particular, or about other
Bioconductor packages in general, they are best asked on the
\href{https://support.bioconductor.org/}{Bioconductor support site}
following the
\href{http://www.bioconductor.org/help/support/posting-guide/}{posting
guidelines}. Questions can be tagged with specific package names or
keywords. For more general information about mass spectrometry and
proteomics, the readers are invited to read the
\emph{\href{http://bioconductor.org/packages/RforProteomics}{RforProteomics}}
package vignettes and associated papers {[}@Gatto:2014;@Gatto2015{]}.

\section{Reading and processing spatial proteomics
data}\label{reading-and-processing-spatial-proteomics-data}

\subsection{The use-case: predicting sub-cellular localisation in
pluripotent embryonic mouse stem
cells}\label{the-use-case-predicting-sub-cellular-localisation-in-pluripotent-embryonic-mouse-stem-cells}

As a use-case, we analyse a recent high-throughput spatial proteomics
dataset from pluripotent mouse embryonic stem cells (E14TG2a)
{[}@hyper{]}. The data was generated using hyperplexed LOPIT
(hyperLOPIT), a state-of-the-art method relying on improved sub-cellular
fractionation and more accurate quantitation, leading to more reliable
classification of protein localisation across the whole sub-cellular
space. The method uses an elaborate sub-cellular fractionation scheme,
enabled by the use of Tandem Mass Tag (TMT) {[}Thompson:2003{]} 10-plex
and application of the MS data acquisition technique named synchronous
precursor selection MS\textsuperscript{3} (SPS-MS\textsuperscript{3})
{[}@McAlister:2014{]}, for TMT quantification with high accuracy and
precision. Three biological replicates were generated from the E14TG2a
experiment, the first was to target low density fractions and the second
and third were to emphasis separation of the denser organelles. The
intersect of replicates 1 and 2 was treated as a 20-plex dataset for the
analysis. As discussed in the manuscript {[}@hyper{]} as it has been
shown that combining replicates across from different gradients can
increase spatial resolution {[}@Trotter:2010{]}. The combination of
replicates resulted in 5032 proteins common in both experiments.

These, as well as many other data are directly available as properly
structured and annotated computational object from the
\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
experiment package. In this workflow, we will start with a description
of how to generate these ad hoc objects starting from an arbitrary
spreadsheets, as produced by many popular third-party applications.

While we focus here on a LOPIT-type dataset, these analyses are relevant
for any quantitative spatial proteomics data, irrespective of the
fractionation or quantitation (i.e.~labelled or label-free) methods.

\subsection{The infrastructure:
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} and
\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}} in
Bioconductor}\label{the-infrastructure-proloc-and-msnbase-in-bioconductor}

To make use of the full functionality of the
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} software
one needs to import their data into R and prepare them as an
\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for
the efficient manipulation and processing of mass spectrometry and
proteomics data in R. Figure 1 illustrates a simplified view of the
\texttt{MSnSet} structure; there exists 3 key sub-parts (termed slots)
to such a data object: (1) the \texttt{exprs} slot for storing the
quantitation data, (2) the \texttt{fData} slot for storing the feature
meta-data, and finally (3) the \texttt{pData} slot for storing the
sample meta-data.

\begin{figure}[htbp]
\centering
\includegraphics{./Figures/msnset.png}
\caption{Simplified representation of the \texttt{MSnSet} data structure
(reproduced with permission from the
\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
vignette)}
\end{figure}

\subsection{Importing data}\label{importing-data}

There are a number of ways to import quantitation data and create an
\texttt{MSnSet} instance. All methods are described in the
\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf}{input/output
capabilities vignette}. One suggested simple method is to use the
function \texttt{readMSnSet2} in
\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}. The
function takes a single spreadsheet as input and extracts the columns
containing the quantitation data, as identified by the argument
\texttt{ecol}, to create the expression data, while the other columns in
the spreadsheet are appended to the feature meta-data slot. By example,
in the code chunk below we read in the \texttt{csv} spreadsheet
containing the quantitation data from the intersect of replicates 1 and
2 of the mouse map {[}@hyper{]}, using the \texttt{readMSnSet2}
function. The data is as available online with the manuscript (see tab 2
of the \texttt{xlsx} supplementary data set 1 in {[}@hyper{]}, which
should be exported as a text-based spreadsheet). It is also available as
a \texttt{csv} in the Bioconductor \texttt{r Biocexptpkg("pRolocdata")}
data package.

To use the \texttt{readMSnSet2} function, as a minimum one must specify
the file path to the data and which columns of the spreadsheet contain
quantitation data. The \texttt{getEcols} function exists to help users
identify which columns of the spreadsheet contain the quantitation data.
In the last line of the code chunk below, we print the file name (not
the full path, which will vary from computer to computer).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"MSnbase"}\NormalTok{)}
\NormalTok{extdatadir <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"pRolocdata"}\NormalTok{)}
\NormalTok{csvfile <-}\StringTok{ }\KeywordTok{dir}\NormalTok{(extdatadir, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{,}
          \DataTypeTok{pattern =} \StringTok{"hyperLOPIT-SIData-ms3-rep12-intersect.csv"}\NormalTok{)}
\KeywordTok{basename}\NormalTok{(csvfile)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "hyperLOPIT-SIData-ms3-rep12-intersect.csv.gz"
\end{verbatim}

Note that the file is compressed (as indicated by the \texttt{gz}, for
\texttt{gzip}, extension), and will be decompressed on-the-fly when read
into R later on.

The spreadsheet that was deposited by the authors contains two headers,
with the second header containing information about where the
quantitation data is stored.

\begin{figure}[htbp]
\centering
\includegraphics{./Figures/spreadsheet-screenshot.png}
\caption{A screenshot of the data in the spreadsheet.}
\end{figure}

We can display the names of the second header by calling the
\texttt{getEcols} function with the argument \texttt{n = 2} (the default
value is \texttt{n = 1}), to specify that we wish to display the column
names of the second line.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{getEcols}\NormalTok{(csvfile, }\DataTypeTok{split =} \StringTok{","}\NormalTok{, }\DataTypeTok{n =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] ""                                 
##  [2] ""                                 
##  [3] ""                                 
##  [4] "Experiment 1"                     
##  [5] "Experiment 2"                     
##  [6] "Experiment 1"                     
##  [7] "Experiment 2"                     
##  [8] "126"                              
##  [9] "127N"                             
## [10] "127C"                             
## [11] "128N"                             
## [12] "128C"                             
## [13] "129N"                             
## [14] "129C"                             
## [15] "130N"                             
## [16] "130C"                             
## [17] "131"                              
## [18] "126"                              
## [19] "127N"                             
## [20] "127C"                             
## [21] "128N"                             
## [22] "128C"                             
## [23] "129N"                             
## [24] "129C"                             
## [25] "130N"                             
## [26] "130C"                             
## [27] "131"                              
## [28] "phenoDisco Input"                 
## [29] "phenoDisco Output"                
## [30] "Curated phenoDisco Output"        
## [31] "SVM marker set"                   
## [32] "SVM classification"               
## [33] "SVM score"                        
## [34] "SVM classification (top quartile)"
## [35] "Final Localization Assignment"    
## [36] "First localization evidence?"     
## [37] "Curated Organelles"               
## [38] "Cytoskeletal Components"          
## [39] "Trafficking Proteins"             
## [40] "Protein Complexes"                
## [41] "Signaling Cascades"               
## [42] "Oct4 Interactome"                 
## [43] "Nanog Interactome"                
## [44] "Sox2 Interactome"                 
## [45] "Cell Surface Proteins"
\end{verbatim}

It is now easy for one to identify that the quantitation data,
corresponding to the 10 TMT isobaric tags, is located in columns 8 to
27. We now have the two mandatory arguments to \texttt{readMSnSet2},
namely the file name (stored in the \texttt{csvfile} variable) and the
quantitation column indices. In addition to these, it is also possible
to pass the optional argument \texttt{fnames} to indicate which column
to use as the labels by which to identify each protein in the sample.
Here, we use \texttt{fnames = 1} to use the UniProt identifiers
contained in the first (unnamed) column of the spreadsheet. We also need
to specify to skip the first line of the file (for the same reason that
we used \texttt{n = 2} in \texttt{getEcols} above) to read the
\texttt{csv} data and convert it to an \texttt{MSnSet} object, named
\texttt{hl} (for hyperLOPIT).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl <-}\StringTok{ }\KeywordTok{readMSnSet2}\NormalTok{(csvfile, }\DataTypeTok{ecol =} \DecValTok{8}\NormalTok{:}\DecValTok{27}\NormalTok{, }\DataTypeTok{fnames =} \DecValTok{1}\NormalTok{, }\DataTypeTok{skip =} \DecValTok{1}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Below, we display a short summary of the data. The data contains 5032
proteins/features common across the 2 biological replicates for the
respective 2 x 10-plex reporter tags (20 columns/samples), along with
associated feature meta-data such as protein markers, protein
description, number of quantified peptides etc (see below).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData: none
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: X X.1 ... Cell.Surface.Proteins (25 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
##  MSnbase version: 1.99.2
\end{verbatim}

Below, we examine the quantitative information for first 5 proteins. It
is also possible to access specific rows and columns by naming the
proteins and TMT tag channels of interest.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{exprs}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{5}\NormalTok{, ]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##           X126 X127N X127C X128N X128C X129N X129C X130N X130C  X131
## Q9JHU4   0.028 0.034 0.024 0.014 0.026 0.045 0.107 0.341 0.059 0.321
## Q9QXS1-3 0.039 0.134 0.095 0.053 0.084 0.121 0.107 0.128 0.122 0.117
## Q9ERU9   0.021 0.013 0.014 0.009 0.024 0.054 0.116 0.257 0.209 0.284
## P26039   0.120 0.255 0.148 0.091 0.135 0.095 0.041 0.057 0.014 0.043
## Q8BTM8   0.055 0.139 0.078 0.050 0.077 0.098 0.093 0.171 0.079 0.160
##          X126.1 X127N.1 X127C.1 X128N.1 X128C.1 X129N.1 X129C.1 X130N.1
## Q9JHU4    0.037   0.064   0.058   0.059   0.067   0.078   0.140   0.208
## Q9QXS1-3  0.033   0.073   0.074   0.062   0.081   0.142   0.190   0.069
## Q9ERU9    0.026   0.017   0.023   0.029   0.039   0.071   0.105   0.171
## P26039    0.111   0.181   0.141   0.144   0.152   0.119   0.075   0.028
## Q8BTM8    0.062   0.108   0.091   0.086   0.099   0.111   0.117   0.095
##          X130C.1 X131.1
## Q9JHU4     0.141  0.147
## Q9QXS1-3   0.151  0.125
## Q9ERU9     0.304  0.215
## P26039     0.017  0.033
## Q8BTM8     0.144  0.087
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{exprs}\NormalTok{(hl)[}\KeywordTok{c}\NormalTok{(}\StringTok{"Q9ERU9"}\NormalTok{, }\StringTok{"Q9Z2R6"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"X126"}\NormalTok{, }\StringTok{"X131.1"}\NormalTok{)]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##         X126 X131.1
## Q9ERU9 0.021  0.215
## Q9Z2R6 0.563  0.000
\end{verbatim}

The feature meta-data is stored in the \texttt{fData} slot and can be
accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2},
automatically, everything that is not defined as quantitation data by
\texttt{ecol} or the feature names by \texttt{fnames} is deposited to
the \texttt{fData} slot.

We see the \texttt{fData} contains 25 columns describing information
such as the number of peptides, associated markers, machine learning
results etc. To identify the feature variable names we can use the
function \texttt{fvarLabels}. We see that the first 6 feature variable
names contain non-discriminatory label names, so we relabel them to help
us identify what feature data information is stored in the associated
columns.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fvarLabels}\NormalTok{(hl)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] "X"                                
##  [2] "X.1"                              
##  [3] "X.2"                              
##  [4] "Experiment.1"                     
##  [5] "Experiment.2"                     
##  [6] "Experiment.1.1"                   
##  [7] "Experiment.2.1"                   
##  [8] "phenoDisco.Input"                 
##  [9] "phenoDisco.Output"                
## [10] "Curated.phenoDisco.Output"        
## [11] "SVM.marker.set"                   
## [12] "SVM.classification"               
## [13] "SVM.score"                        
## [14] "SVM.classification..top.quartile."
## [15] "Final.Localization.Assignment"    
## [16] "First.localization.evidence."     
## [17] "Curated.Organelles"               
## [18] "Cytoskeletal.Components"          
## [19] "Trafficking.Proteins"             
## [20] "Protein.Complexes"                
## [21] "Signaling.Cascades"               
## [22] "Oct4.Interactome"                 
## [23] "Nanog.Interactome"                
## [24] "Sox2.Interactome"                 
## [25] "Cell.Surface.Proteins"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fvarLabels}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{3}\NormalTok{] <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"uniprot.accession"}\NormalTok{, }\StringTok{"uniprot.id"}\NormalTok{, }\StringTok{"description"}\NormalTok{)}
\KeywordTok{fvarLabels}\NormalTok{(hl)[}\DecValTok{4}\NormalTok{:}\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\StringTok{"peptides.expt"}\NormalTok{, }\DecValTok{1}\NormalTok{:}\DecValTok{3}\NormalTok{)}
\KeywordTok{fData}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{4}\NormalTok{, }\DecValTok{1}\NormalTok{:}\DecValTok{6}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##          uniprot.accession  uniprot.id
## Q9JHU4              Q9JHU4 DYHC1_MOUSE
## Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE
## Q9ERU9              Q9ERU9  RBP2_MOUSE
## P26039              P26039  TLN1_MOUSE
##                                                                                                                                 description
## Q9JHU4                                              Cytoplasmic dynein 1 heavy chain 1 OS=Mus musculus GN=Dync1h1 PE=1 SV=2 - [DYHC1_MOUSE]
## Q9QXS1-3 Isoform PLEC-1 of Plectin OS=Mus musculus GN=Plec - [PLEC_MOUSE]|Isoform PLEC-1A of Plectin OS=Mus musculus GN=Plec - [PLEC_MOUSE]
## Q9ERU9                                                     E3 SUMO-protein ligase RanBP2 OS=Mus musculus GN=Ranbp2 PE=1 SV=2 - [RBP2_MOUSE]
## P26039                                                                             Talin-1 OS=Mus musculus GN=Tln1 PE=1 SV=2 - [TLN1_MOUSE]
##          peptides.expt1 peptides.expt2 peptides.expt3
## Q9JHU4              175            166            322
## Q9QXS1-3            123            150            174
## Q9ERU9              101             90            181
## P26039              101             94            167
\end{verbatim}

Note that when using the simple \texttt{readMSnSet2} procedure, the
\texttt{pData} slot which is used to store information about the
samples/channels is kept empty. It is advised to annotate the channels
as well. Below, we annotate the replicate from which the profiles
originate and the TMT tag (extracted from the sample/channel names).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{pData}\NormalTok{(hl)$Replicate <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{each =} \DecValTok{10}\NormalTok{)}
\KeywordTok{pData}\NormalTok{(hl)$Tag <-}\StringTok{ }\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{.1$"}\NormalTok{, }\StringTok{""}\NormalTok{, }\KeywordTok{sub}\NormalTok{(}\StringTok{"^X"}\NormalTok{, }\StringTok{""}\NormalTok{, }\KeywordTok{sampleNames}\NormalTok{(hl)))}
\KeywordTok{pData}\NormalTok{(hl)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##         Replicate  Tag
## X126            1  126
## X127N           1 127N
## X127C           1 127C
## X128N           1 128N
## X128C           1 128C
## X129N           1 129N
## X129C           1 129C
## X130N           1 130N
## X130C           1 130C
## X131            1  131
## X126.1          2  126
## X127N.1         2 127N
## X127C.1         2 127C
## X128N.1         2 128N
## X128C.1         2 128C
## X129N.1         2 129N
## X129C.1         2 129C
## X130N.1         2 130N
## X130C.1         2 130C
## X131.1          2  131
\end{verbatim}

Throughout this workflow we refer to the different columns that are
found in the \texttt{exprs} (expression data) slot as channels (short
for TMT channels). In the frame of LOPIT and hyperLOPIT these channels
consititue the relative abundance of each protein (along the rows) in
the channel of interest. Each TMT channel originates from fractions
collected from the density gradient, or a set of pooled fractions or may
be a sample originating from an alternative preparation e.g.~such as
from the chromatin enrichment performed in Christoforou et al
{[}@hyper{]}. Information about which gradient fractions were used for
which tag should also be stored in the sample meta-data \texttt{pData}
slot.

The sample meta-data that is distributed with the
\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
package for Christoforou's hyperLOPIT experiment and (as above) the
quantitation data file, are located in the \texttt{extdata} in the
\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
package on the hard drive.

In the code chunk below we again use the \texttt{dir} function to locate
the filepath to the meta-data \texttt{csv} file and then read it into R
using \texttt{read.csv}. We then append the meta-data to the
\texttt{pData} slot. Information about the gradient fractions used and
the associated subcellular fraction densities in \% w/v Iodixanol for
each replicate are stored here.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{expinfo <-}\StringTok{ }\KeywordTok{dir}\NormalTok{(extdatadir, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{,}
               \DataTypeTok{pattern =} \StringTok{"hyperLOPIT-SIData-fraction-info.csv"}\NormalTok{)}

\NormalTok{fracinfo <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(expinfo, }\DataTypeTok{row.names=}\DecValTok{1}\NormalTok{, }\DataTypeTok{skip =} \DecValTok{2}\NormalTok{, }
                     \DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}

\KeywordTok{pData}\NormalTok{(hl)$Gradient.Fraction <-}\StringTok{ }\KeywordTok{c}\NormalTok{(fracinfo[, }\DecValTok{1}\NormalTok{], fracinfo[, }\DecValTok{2}\NormalTok{])}
\KeywordTok{pData}\NormalTok{(hl)$Iodixonal.Density <-}\StringTok{ }\KeywordTok{c}\NormalTok{(fracinfo[, }\DecValTok{4}\NormalTok{], fracinfo[, }\DecValTok{5}\NormalTok{])}
\KeywordTok{pData}\NormalTok{(hl)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##         Replicate  Tag Gradient.Fraction Iodixonal.Density
## X126            1  126           Cytosol               0.0
## X127N           1 127N   1 to 6 (pooled)               6.0
## X127C           1 127C   8 to 9 (pooled)              11.0
## X128N           1 128N 10 to 11 (pooled)              13.3
## X128C           1 128C                12              14.6
## X129N           1 129N                14              17.4
## X129C           1 129C                16              20.1
## X130N           1 130N                18              26.8
## X130C           1 130C         Chromatin                NA
## X131            1  131                19              34.5
## X126.1          2  126           Cytosol               0.0
## X127N.1         2 127N   1 to 6 (pooled)               5.2
## X127C.1         2 127C   7 to 9 (pooled)              10.0
## X128N.1         2 128N 10 to 11 (pooled)              12.5
## X128C.1         2 128C                12              14.0
## X129N.1         2 129N 14 to 15 (pooled)              17.3
## X129C.1         2 129C                17              20.9
## X130N.1         2 130N 18 to 19 (pooled)              24.7
## X130C.1         2 130C         Chromatin                NA
## X131.1          2  131                20              31.9
\end{verbatim}

\subsection{Data processing}\label{data-processing}

\subsubsection{Normalisation}\label{normalisation}

There are two aspects related to data normalisation that are relevant to
spatial proteomics data processing. The first one focuses on reducing
purely technical variation between channels without affecting biological
variability (i.e.~the shape of the quantitative profiles). This
normalisation will depend on the underlying quantitative technology and
the experimental design, and will not be addressed in this workflow. The
second aspect, and more specific to spatial proteomics data, is scaling
all the organelle-specific profiles into a same intensity interval
(typically 0 and 1) by, for example, diving each intensity by the sum of
the intensities for that quantitative feature. This is not necessary in
this example as the intensities for each replicate have already been
re-scaled to 1 in Proteome Discoverer. However, if one wanted to do this
they would execute the \texttt{normalise} function as demonstrated in
the below code chunk.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl <-}\StringTok{ }\KeywordTok{normalise}\NormalTok{(hl, }\DataTypeTok{method =} \StringTok{"sum"}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

This transformation of the data assures to cancel the effect of the
absolute intensities of the quantitative features along the rows, to
focus subsequent analyses on the relative profiles along the
sub-cellular channels.

The same \texttt{normalise} function (or \texttt{normalize}, both
spellings are supported) can also be applied in the first case described
above. Different normalisation methods such as mean or median scaling,
variance stabilisation or quantile normalisation, to cite a few, can be
applied to accomodation different needs.

As previously mentioned, before combination, the two replicates in the
\texttt{hl} data that we read into R were separately normalised by sum
(i.e.~to 1) across the 10 channels for each replicate respectively. We
can verify this by summing each rows for each replicate:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{rowSums}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hl[, hl$replicate ==}\StringTok{ }\DecValTok{1}\NormalTok{])))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##       0       0       0       0       0       0
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{summary}\NormalTok{(}\KeywordTok{rowSums}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hl[, hl$replicate ==}\StringTok{ }\DecValTok{2}\NormalTok{])))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
##       0       0       0       0       0       0
\end{verbatim}

We see that some features do not add up exactly to 1 due to rounding
errors after exporting to intermediate files. These small deviations do
not bear any consequences here.

\subsubsection{Combining acquisitions}\label{combining-acquisitions}

The spreadsheet that was used to create the \texttt{hl} MSnSet included
the two replicates within one .csv file. We also provide individual
replicates in the
\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
package. Below, we show how to combine \texttt{MSnSet} objects and,
subsequently, how to filter and handle missing values. We start by
loading the \texttt{r Biocexptpkg("pRolocdata")} package and the
equivalent replicates using the \texttt{data} function.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocdata"}\NormalTok{)}
\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r1)}
\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r2)}
\end{Highlighting}
\end{Shaded}

At the R prompt, typing

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{pRolocdata}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

will list the 54 datasets that are available in
\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}.

Combining data is performed with the \texttt{combine} function. This
function will inspect the feature and sample names to identify how to
combine the data. As we want our replicates to be combined along the
columns (same proteins, different sets of channels), we need to assure
that the respective sample names differ so they can be identified from
one another. The function \texttt{updateSampleNames} can be used do
this.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{identical}\NormalTok{(}\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r1), }\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r2))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] TRUE
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hyperLOPIT2015ms3r1 <-}\StringTok{ }\KeywordTok{updateSampleNames}\NormalTok{(hyperLOPIT2015ms3r1, }\DecValTok{1}\NormalTok{)}
\NormalTok{hyperLOPIT2015ms3r2 <-}\StringTok{ }\KeywordTok{updateSampleNames}\NormalTok{(hyperLOPIT2015ms3r2, }\DecValTok{2}\NormalTok{)}
\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] "X126.1"  "X127N.1" "X127C.1" "X128N.1" "X128C.1" "X129N.1" "X129C.1"
##  [8] "X130N.1" "X130C.1" "X131.1"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] "X126.2"  "X127N.2" "X127C.2" "X128N.2" "X128C.2" "X129N.2" "X129C.2"
##  [8] "X130N.2" "X130C.2" "X131.2"
\end{verbatim}

In addition, to matching names, the content of the feature metadata for
identical feature annotations must match exactly across the data to be
combined. In particular for these data, we expect the same proteins in
each replicates to be annotated with the same UniProt entry names and
descriptions, but not with the same coverage of number of peptides or
peptide-spectrum matches (PSMs).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "EntryName"          "ProteinDescription" "Peptides"          
## [4] "PSMs"               "ProteinCoverage"    "markers"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "EntryName"          "ProteinDescription" "Peptides"          
## [4] "PSMs"               "ProteinCoverage"    "markers"
\end{verbatim}

Below, we update the replicate specific feature variable names and
remove the shared annotation.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{], }\DecValTok{1}\NormalTok{)}
\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{], }\DecValTok{2}\NormalTok{)}
\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r1) <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{1}\NormalTok{:}\DecValTok{5}\NormalTok{] }
\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r2) <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{]}
\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "EntryName"          "ProteinDescription" "Peptides1"         
## [4] "PSMs1"              "ProteinCoverage1"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "Peptides2"        "PSMs2"            "ProteinCoverage2"
\end{verbatim}

We can now combine the two experiments into a single \texttt{MSnSet}:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{combined <-}\StringTok{ }\KeywordTok{combine}\NormalTok{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
\NormalTok{combined}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 6725 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
##  MSnbase version: 1.21.7
\end{verbatim}

More details above combining data are given in the dedicated
\emph{Combining MSnSet instances} section of the
\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
vignette}.

\subsubsection{Missing data}\label{missing-data}

Missing data are a recurrent issue in mass spectrometry applications,
and should be addressed independently of this workflow
{[}@Webb-Robertson:2015;@Lazar:2016{]}. In {[}@Gatto:2014b{]}, we have
described how a high content in missing values in spatial proteomics
data and their inappropriate handling leads to a reduction of
sub-cellular resolution. Missing data can be imputated using
\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}'s
\texttt{impute} function. The method underlying the imputation method is
then determined by a \texttt{methods} parameter. In our particular case,
missing values are indicative of protein groups that were not profiles
in both replicates.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{image2}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(combined), }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"black"}\NormalTok{, }\StringTok{"white"}\NormalTok{),}
       \DataTypeTok{main =} \StringTok{"Missing values (white cells) after combining replicates"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/namap-1.png}
\caption{plot of chunk namap}
\end{figure}

We prefer to remove proteins that were not assayed in both of our two
replicated experiments. This is done with the \texttt{filterNA} function
that removes features that contain more than a certain proportion
(default is 0) missing values.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{combined <-}\StringTok{ }\KeywordTok{filterNA}\NormalTok{(combined)}
\NormalTok{combined}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
##     total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
## Subset [6725,20][5032,20] Thu Sep 29 10:28:15 2016 
## Removed features with more than 0 NAs: Thu Sep 29 10:28:15 2016 
## Dropped featureData's levels Thu Sep 29 10:28:15 2016 
##  MSnbase version: 1.21.7
\end{verbatim}

When more than 2 data are to be combined and/or too many proteins were
not consistently assayed, leading to too many proteins being filtered
out, we suggest to implement an ensemble of classifiers voting on
protein-sub-cellular niche membership over the output of several
experiments (see section \emph{Supervised machine learning} for the
description of sub-cellular assignments).

\section{Quality Control}\label{quality-control}

Data quality is routinely examined through visualisation to verify that
sub-cellular niches have been separated along the gradient. Based on De
Duve's principle {[}@DeDuve:1981{]} proteins that co-localise exhibit
similar quantitation profiles across the gradient fractions employed.
One approach that has been widely used to visualise and inspect high
throughput mass spectrometry-based proteomics data is principal
components analysis (PCA). PCA is one of many dimensionality reduction
methods, that allow one to effectively summarise multi-dimensional data
in to 2 or 3 dimensions to enable visualisation. Very generally, the
original continuous multi-dimensional data is transformed into a set of
orthogonal components ordered according to the amount of variability
that they describe. The \texttt{plot2D} method in
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} allows one
to plot the principal components (PCs) of a dataset against one another,
by default the first two components are plotted on the x- and y-axis,
respectively (the \texttt{dims} argument can be used to plot other PCs).
If distinct clusters are observed, we assume that there is organellar
separation present in the data. Although, representing the
multi-dimensional data along a limited set of PCs does not give us a
hard quantitative measure of separation, it is extremely useful
summarising complex experimental information in one figure, to get an
simplified overview of the data.

In the code chunk below we produce a PCA plot of the mouse stem cell
dataset. One point on the plot represents one protein. We can indeed see
several distinct protein clusters. We specify \texttt{fcol = NULL},
which means not to consider any feature variable to annotate the
features (proteins) with colours. We will see later how to use this to
annotate the PCA plot with prior information about sub-cellular
localisation.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"pRoloc"}\NormalTok{)}
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/qcplot-1.png}
\caption{plot of chunk qcplot}
\end{figure}

In the first instance we advise one to visualise their data without any
annotation (i.e.~with \texttt{fcol = NULL}), before proceeding with data
annotation. The identification of well resolved clusters in the data,
constitutes an unbiased assessment of the data structure, demonstrating
the successful separation of sub-cellular clusters.

\textbf{TODO} There are other dimensionality reduction methods available
in the \texttt{plot2D} function, which can be parametrised with the
\texttt{method} argument. Advantages, disadvantages? See
\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}{this}
very interesting gist with intriguing tSNE examples.

It is also useful to visualise the relative intensities along the
gradient to identify channels displaying particularly low yield. This
can be done using the \texttt{plotDist} and \texttt{boxplot} functions,
that plot the protein profiles occupancy along the gradient (we also
display the mean channel intensities) and a \texttt{boxplot} of the
column intensities. In the two plots displayed below, we re-order the
TMT channles to pair corresponding channels in the two replicates
(rather than ordering the channels by replicate).

\texttt{\{r qcbx, fig.width = 12\} par(mfrow = c(1, 2)) o \textless{}- order(hl\$Iodixonal.Density) plotDist(hl{[}, o{]}, pcol = "\#00000010") lines(colMeans(exprs(hl{[}, o{]})), col = "red", type = "b") boxplot(exprs(hl{[}, o{]}), las = 2)}

\section{Markers}\label{markers}

In the context of spatial proteomics, a marker protein is defined as a
well-known resident of a specific sub-cellular niche in a species
\emph{and} condition of interest. Applying this to machine learning
(ML), and specifically supervised learning, for the task of protein
localisation prediction, these markers constitute the labelled training
data to use as input to a classification analyses. Defining well-known
residents, and obtaining labelled training data for ML analyses can be
time consuming, but it is important to define markers that are
representative of the multivariate data space and on which a classifier
will be trained and generated.
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} provides a
convenience function, \texttt{addMarkers}, to directly add markers to a
\texttt{MSnSet} object, as demonstrated in the code chunk below. These
marker sets can be accessed using the \texttt{pRolocmarkers()} function.
Marker sets are stored as a simple named vector in R, and originate from
in-house user-defined sets of markers or from previous published studies
{[}@Gatto:2014b{]}, which are continuosly updated and integrated.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## List available marker sets}
\KeywordTok{pRolocmarkers}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## 7 marker lists available:
## Arabidopsis thaliana [atha]:
##  Ids: TAIR, 543 markers
## Drosophila melanogaster [dmel]:
##  Ids: Uniprot, 179 markers
## Gallus gallus [ggal]:
##  Ids: IPI, 102 markers
## Homo sapiens [hsap]:
##  Ids: Uniprot ID, 205 markers
## Mus musculus [mmus]:
##  Ids: Uniprot, 937 markers
## Saccharomyces cerevisiae [scer_sgd]:
##  Ids: SGD, 259 markers
## Saccharomyces cerevisiae [scer_uniprot]:
##  Ids: Uniprot Accession, 259 markers
\end{verbatim}

These markers can then be mapped to a \texttt{MSnSet}'s
\texttt{featureNames}. The mouse dataset used here has Uniprot IDs
stored as the \texttt{featureNames} (see
\texttt{head(featureNames(hl))}) and the names of the vector of the
mouse markers stored in
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}
(\texttt{mmus} markers) are also Uniprot IDs (see \texttt{head(mrk)} in
the code chunk below), so it is straightforward to match names between
the markers and the \texttt{MSnSet} instance using the
\texttt{addMarkers} function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Use mouse markers}
\NormalTok{mrk <-}\StringTok{ }\KeywordTok{pRolocmarkers}\NormalTok{(}\DataTypeTok{species =} \StringTok{"mmus"}\NormalTok{)}
\KeywordTok{head}\NormalTok{(mrk)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##                 P26039                 Q6PB66                 P11276 
##   "Actin cytoskeleton"        "Mitochondrion" "Extracellular matrix" 
##                 Q6PR54                 Q05793                 P19096 
##  "Nucleus - Chromatin" "Extracellular matrix"              "Cytosol"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Add mouse markers}
\NormalTok{hl <-}\StringTok{ }\KeywordTok{addMarkers}\NormalTok{(hl, mrk)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Markers in data: 937 out of 5032
\end{verbatim}

\begin{verbatim}
## organelleMarkers
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      27                      43                      13 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                      43                      95                      12 
##    Extracellular matrix         Golgi apparatus                Lysosome 
##                      10                      27                      33 
##           Mitochondrion     Nucleus - Chromatin Nucleus - Non-chromatin 
##                     383                      64                      85 
##              Peroxisome         Plasma membrane              Proteasome 
##                      17                      51                      34 
##                 unknown 
##                    4095
\end{verbatim}

If the naming between the marker sets and the \texttt{MSnSet} dataset
are different, one will have to convert and match the proteins according
to the appropriate identifier. Sometimes, we find the equivalent entry
name, Uniprot ID or accession number is stored with the data, which
makes conversion between identifers relatively straightforward. If this
is not the case however, conversion can be performed using
\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}}, the
Bioconductor
\href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation
resouces} or any conversion softwares available online.

We now visualise these annotations along the PCA plot using the
\texttt{plot2D} function and then use the \texttt{addLegend} function to
map the marker classes to the pre-defined colours. We also display the
data along the first and seventh PCs using the \texttt{dims} argument.
Note that in this second call to the \texttt{plot2D} function, we have
omitted the \texttt{fcol} argument, as the \texttt{"markers"} feature
variable name is the default value. We choose to display PCs 1 and 7 to
illustrate that while upper principal components explain much less
variability in the data (2.23\% for PC7, as opposed to 48.41\% for PC1),
we see that the mitochondrial and peroxisome clusters can be
differenciated, despite the apparent overlap in the two first PCs.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{main =} \StringTok{"pRolocmarkers for mouse"}\NormalTok{)}
\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{)}
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{dims =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{7}\NormalTok{), }\DataTypeTok{main =} \StringTok{"Marker resolution along PC 1 and 7"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plotmarkers-1.png}
\caption{plot of chunk plotmarkers}
\end{figure}

The colours have been defined so as to enable to differenciate up to 30
classes. If more are provided, different character symbols (circles,
squares, \ldots{} and empty and solid symbols) are used. The colours and
the default plotting characters (solid dots for the markers and empty
circles for the features of unknown localisation) can of course be
changed, as described in the \texttt{setStockcol} manual page.

As demonstrated in {[}@hyper{]} and illustrated in the PCA plot above,
the Golgi apparatus proteins display a dynamic pattern, noting sets of
Golgi marker proteins that are distributed amongst other subcellular
structures, an observation supported by microscopy. As such, we are
going to reset the annotation of Golgi markers to unknown using the
\texttt{fDataTounknown} function. It is often used to replace empty
strings (``'') or missing values in the markers definition to a common
definition of \emph{unknown localisation}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(hl, }\DataTypeTok{from =} \StringTok{"Golgi apparatus"}\NormalTok{, }\DataTypeTok{to =} \StringTok{"unknown"}\NormalTok{)}
\KeywordTok{getMarkers}\NormalTok{(hl)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## organelleMarkers
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      27                      43                      13 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                      43                      95                      12 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      10                      33                     383 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                      64                      85                      17 
##         Plasma membrane              Proteasome                 unknown 
##                      51                      34                    4122
\end{verbatim}

In general, the Gene Ontology (GO) {[}@Ashburner:2000{]}, and in
particular the cellular compartment (CC) namespace are a good starting
point for protein annotation and marker definition. It is important to
note however that automatic retrieval of sub-cellular localisation
information, from
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} or
elsewhere, is only the beginning in defining a marker set for downstream
analyses. Expert curation is vital to check that any annotation added is
in the correct context for the the biological question under
investigation.

Another useful visualisation that relies on marker annotation is the
representation of the protein profiles occupancy along the gradient
using the \texttt{plotDist} function. While the PCA plot enables to
efficiently visualise the complete dataset and assess the relative
separation of different sub-cellular niches, comparing profiles of a few
marker clusters is useful to assess how exactly they differ (in terms of
peak channels, for example). Below, we plot the profile of the
mitochondrial and peroxisome markers to highlight the differences in
profiles between these two sets of markers along the 6th and 7th
channels, as represented above along the 7th PC on the PCA plot.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hlo <-}\StringTok{ }\NormalTok{hl[, }\KeywordTok{order}\NormalTok{(hl$Iodixonal.Density)]}
\KeywordTok{plotDist}\NormalTok{(hlo[}\KeywordTok{fData}\NormalTok{(hlo)$markers ==}\StringTok{ "Mitochondrion"}\NormalTok{, ],}
         \DataTypeTok{pcol =} \StringTok{"purple"}\NormalTok{, }\DataTypeTok{fractions =} \StringTok{"Fraction.No"}\NormalTok{)}
\KeywordTok{title}\NormalTok{(}\DataTypeTok{main =} \StringTok{"Marker occupancy profiles along the gradient"}\NormalTok{)}
\KeywordTok{matlines}\NormalTok{(}\KeywordTok{t}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hlo[}\KeywordTok{fData}\NormalTok{(hlo)$markers ==}\StringTok{ "Peroxisome"}\NormalTok{, ])),}
         \DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"darkblue"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"Mitochondrion"}\NormalTok{, }\StringTok{"Peroxisome"}\NormalTok{),}
       \DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"purple"}\NormalTok{, }\StringTok{"blue"}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plotDist-1.png}
\caption{plot of chunk plotDist}
\end{figure}

\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust},
although these are only available in the development version.

\section{Replication}\label{replication}

With the aim of maximising the sub-cellular resolution and,
consequently, the reliability in protein sub-cellular assignments, we
follow the advice in {[}@Trotter:2010{]} and combine replicated spatial
proteomics experiments as described above. Indeed, Trotter et al. have
shown a significant improvement in protein--organelle association upon
direct combination of single experiments, in particular when these
resolve different subcellular niches.

Direct comparisons of individual channels in replicated experiments does
not provide an adequate, goal-driven assessment of different
experiments. Indeed, due to the nature of the experiment and gradient
fraction collection, the quantitative channels do not correspond to
identical selected fractions along the gradient. As can be seen in the
table below (taken from the \texttt{hyperLOPIT}'s \texttt{pData},
focusing on channels 7 to 10), different sets of gradient fractions are
pooled to obtain enough material and optimise acurate quantitation.

The more relevant comparison unit is not a single channel, but rather
the complete protein occupancy profiles, which are best visualised as
experiment-wide on a PCA plot. As such, we prefer to focus on the
direct, qualitative comparison of individual replicate PCA plots,
assuring that each displays acceptable sub-cellular resolution. Note
that in the code chunk below, we mirror the x-axis to represent the two
figures with the same orientation.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{)) }
\KeywordTok{plot2D}\NormalTok{(hl[, hl$Replicate ==}\StringTok{ }\DecValTok{1}\NormalTok{], }\DataTypeTok{main =} \StringTok{"Replicate 1"}\NormalTok{) }
\KeywordTok{plot2D}\NormalTok{(hl[, hl$Replicate ==}\StringTok{ }\DecValTok{2}\NormalTok{], }\DataTypeTok{main =} \StringTok{"Replicate 2"}\NormalTok{, }\DataTypeTok{mirrorX =} \OtherTok{TRUE}\NormalTok{) }
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot2Drep-1.png}
\caption{plot of chunk plot2Drep}
\end{figure}

\section{Interactive visualisation}\label{interactive-visualisation}

Visualisation and data exploration is an important aspect of data
analyses allowing one to shed light on data structure and patterns of
interest. Using the
\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
package we can interactively visualise, explore and interrogate
quantitative spatial proteomics data. The
\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
package is currently under active development and it relies on the
\texttt{shiny} framework for reactivity and interactivity. The package
currently distributes 3 different GUI's (``main'' (default), ``compare''
or ``compare'') which are wrapped and launched by the \texttt{pRolocVis}
function. In the below code chunk we lauch the main app (note, we do not
need to specify the argument, \texttt{app = "main"} as it is the
default).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocGUI"}\NormalTok{)}
\KeywordTok{pRolocVis}\NormalTok{(hl)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{./Figures/mainapp.png}
\caption{A screen shot of clickable interface and zoomable PCA plot of
the main app in the
\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
package.}
\end{figure}

As diplayed in the screenshot above, the ``main'' application is
designed for exploratory data analysis and is divied into 3 tabs: (1)
PCA, (2) Profiles and (3) Table selection. The default view upon loading
is the PCA tab, which features a clickable interface and zoomable PCA
plot with an interactive data table for displaying the quantitation
information. Particular proteins of interest can be highlighted using
the text search box. There is also an alternate profiles tab for
visualisation of the protein profiles, which can be used to examine the
patterns of proteins of interest. The Table selection tab provides an
interface to control data table column selection.

The ``compare'' application is useful for examining two replicate
experiments, or two experiments from different conditions, treatments
etc. The compare application is called by default if the input object to
\texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets}, but
it can also be specified by calling the argument
\texttt{app = "compare"}. For example, in the code chunk below we first
create a \texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT
data, this is then passed to \texttt{pRolocVis}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r1)}
\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r2)}
\NormalTok{mydata <-}\StringTok{ }\KeywordTok{MSnSetList}\NormalTok{(}\KeywordTok{list}\NormalTok{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
\KeywordTok{pRolocVis}\NormalTok{(mydata, }\DataTypeTok{app =} \StringTok{"compare"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{./Figures/SS_Compare1.jpg}
\caption{The compare application, main panel}
\end{figure}

The comparison app loads the two PCA plots side-by-side. Only common
proteins between the two data sets are displayed. As per the main
application proteins can be searched, identified and highlighted on both
PCA plots and in the dedicated profiles tab. One key feature of the
compare application is the ability to re-map the second dataset onto the
PCA data space of the first (reference) data set (see
\texttt{?pRolocVis} and the argument \texttt{remap = TRUE}). Using the
first dataset as the reference set, PCA is carried out on the first
dataset and the standard deviations of the principal components
(i.e.~the square roots of the eigenvalues of the covariance/correlation
matrix) and the matrix of variable loadings (i.e.~a matrix whose columns
contain the eigenvectors) are stored and then used to calculate the
principal components of the second dataset. Both datasets are scaled and
centered in the usual way. The first dataset appears on the left, and
the second re-mapped data appears on the right. The order of the first
(the reference data for remapping) and second dataset can be changed
through regeneration/re-ordering of the \texttt{MSnSetList} object.

The final application ``classify'', has been designed to view machine
learning classification results according to user-specified thresholds
for the assignment of proteins to its sub-cellular location, this is
discussed later in the subsection \hyperref[thresholding]{thresholding}
in the \hyperref[supervised-machine-learning]{supervised machine
learning section}.

\section{Novelty Detection}\label{novelty-detection}

The extraction of sub-cellular protein clusters can be difficult owing
to the limited number of marker proteins that exist in databases and
elsewhere. Furthermore, given the vast complexity of the cell, automatic
annotation retrieval does not always give a full representation of the
true sub-cellular diversity in the data. For downstream analyses, such
as supervised machine learning, it is desirable to obtain reliable
markers that cover as many sub-cellular niches as possible, as these
markers are directly used in the training phase of the ML
classification. We find that a lack of sub-cellular diversity in the
labelled training data leads to prediction errors, as unlabelled
instances can only be assigned to a class that exists in the training
data {[}@Breckels:2013{]}. In such scenarios novelty detection can be
useful to identify data-specific sub-cellular groupings such as
organelles and protein complexes. The phenotype discovery (phenoDisco)
algorithm {[}@Breckels:2013{]} is one such method and is available in
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}. It is an
iterative semi-supervised learning method that combines the
classification of proteins on existing labelled data with the detection
of new clusters.

In addition to extracting new phenotypes, novelty detection methods are
also useful for confirming the presence of known or postulated clusters
in an unbiased fashion. For example, in {[}@hyper{]} the
\texttt{phenoDisco} algorithm was used to confirm the data-specific
presence of the nucleus and nucleus sub-compartments. In the code chunk
below, we demonstrate how to do this analysis, highlighting some of the
optional arguments and parameters available for phenotype extraction and
give some advice on how to interpret the output.

As the \texttt{phenoDisco} algorithm is semi-supervised it uses both
labelled (markers) and unlabelled data to explore the data structure and
find new sub-cellular data clusters. Thus the first step is to define
some input labelled data i.e.~markers, that the algorithm will use as
input for the supervised learning aspect of the algorithm. As described
in {[}@hyper{]} we define a set of markers to use as input for the
analyses that cover well-known residents from three distinct organelle
structures; the mitochondria, plasma membrane and ER, and from three
well-known and abundant protein complexes; the proteasome and two
ribosomal subunits, 40S and 60S. These input markers are stored in the
\texttt{featureData} column of \texttt{hl} where
\texttt{fcol = "phenoDisco.Input"}. We can use the convenience accessor
function \texttt{getMarkers} to print out a table of the markers
contained in this marker set. These initial markers were manually
curated using information from the UniProt database, the Gene Ontology
and the literature.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                    26 
##                          60S Ribosome 
##                                    43 
## Endoplasmic reticulum/Golgi apparatus 
##                                    76 
##                         Mitochondrion 
##                                   261 
##                       Plasma membrane 
##                                    50 
##                            Proteasome 
##                                    34 
##                               unknown 
##                                  4542
\end{verbatim}

In the code chunk below we show how to run the \texttt{phenoDisco}
function and return a novelty detection result, according to the
specified parameters. The algorithm parameters \texttt{times},
\texttt{GS} and \texttt{p} are passed to the function, along with the
\texttt{fcol} to tell the algorithm where the input training data is
contained.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## As per Christoforou et al (2016), }
\NormalTok{hl <-}\StringTok{ }\KeywordTok{phenoDisco}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }
                 \DataTypeTok{times =} \DecValTok{200}\NormalTok{, }
                 \DataTypeTok{GS =} \DecValTok{60}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Note: We do not evaluate this code chunk in this document as the
algorithm is computational intensive and best parallelised over multiple
workers. This phenoDisco analysis took \textasciitilde{}24 hours to
complete when parallelised over 40 workers.}

The argument \texttt{times} indicated the number of times we run
unsupervied Gaussian Mixture Modelling before defining a new phenotype
cluster. The recommended minimum and default value is 100. In the above
code chunk we increase the value to \texttt{times = 200} as we have
found for larger datasets (e.g.~5000+ proteins) a higher times is
requried for convergence. \texttt{GS} defines the minimum number of
proteins allowed per new data cluster and thus heavily influences what
type of new clusters are extracted. For example, if a user is interesed
in the detection of small complexes they may wish to use a small
\texttt{GS = 10}, or \texttt{= 20} etc. If they wish to detect larger
more abundant sub-cellular niches a much higher \texttt{GS} would be
preferable. Specifying a small \texttt{GS} can be more time consuming
than using a larger \texttt{GS}, and there is a trade off between
finding interesting small complexes and those that may not be of
interest as we find there is a tendancy to find more noise when using a
small \texttt{GS} compared to using a higher one.

One may also consider increasing the search space for new data clusters
by increasing the value of the parameter \texttt{G}. This defines the
number of GMM components to test and fit; the default is
\texttt{G = 1:9} (\texttt{mclust}'s default). One should note that the
decreasing the \texttt{GS}, and increasing the values of the arguments
\texttt{times}, \texttt{G} (among other function arguments, see
\texttt{?phenoDisco}) will heavily influence (increase) the total time
taken to run the algorithm. \texttt{phenoDisco} supports parallelisation
and we strongly suggest you make use of a parallel processing to run
these analyses (different backends can be set with argument
\texttt{BPPARAM}).

The ouput of running the \texttt{phenoDisco} algorithm is a
\texttt{MSnSet} containing the new data clusters, appended to the
\texttt{featureData} under the name \texttt{pd}. We can see by typing
\texttt{hl@processingData} directly into the console the processing
information has been updated to the \texttt{MSnSet} recording the
parameters that were used to run the analyses. This is handy for keeping
track of data analyses. The results can be displayed by using the
\texttt{getMarkers} function. We see that 5 new phenotype data clusters
were found.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## MSnSet (storageMode: lockedEnvironment)
## assayData: 5032 features, 20 samples 
##   element names: exprs 
## protocolData: none
## phenoData
##   sampleNames: X126 X127N ... X131.1 (20 total)
##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
##   varMetadata: labelDescription
## featureData
##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
##   fvarMetadata: labelDescription
## experimentData: use 'experimentData(object)'
## Annotation:  
## - - - Processing information - - -
## Added markers from  'mrk' marker vector. Thu Sep 29 10:28:15 2016 
## Added markers from  'pdres' marker vector. Thu Sep 29 10:28:17 2016 
##  MSnbase version: 1.99.2
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## organelleMarkers
##                          40S Ribosome 
##                                   106 
##                          60S Ribosome 
##                                    95 
## Endoplasmic reticulum/Golgi apparatus 
##                                   393 
##                         Mitochondrion 
##                                   525 
##                           Phenotype 1 
##                                   300 
##                           Phenotype 2 
##                                   253 
##                           Phenotype 3 
##                                   203 
##                           Phenotype 4 
##                                    74 
##                           Phenotype 5 
##                                    91 
##                       Plasma membrane 
##                                   421 
##                            Proteasome 
##                                    92 
##                               unknown 
##                                  2479
\end{verbatim}

We can plot the results using the \texttt{plot2D} function.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Re-order the colours for the phenoDisco output}
\NormalTok{cl <-}\StringTok{ }\KeywordTok{getMarkerClasses}\NormalTok{(hl, }\StringTok{"pd"}\NormalTok{)}
\NormalTok{cols <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)]}
\NormalTok{ind <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"Pheno"}\NormalTok{, cl, }\DataTypeTok{invert =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{cols[ind] <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)][}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(ind)]}
\NormalTok{cols[-ind] <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)][(}\KeywordTok{length}\NormalTok{(ind) +}\StringTok{ }\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(cl)]}

\NormalTok{## Plot the input and output}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }
       \DataTypeTok{main =} \StringTok{"phenoDisco input markers"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{getStockcol}\NormalTok{()[}\DecValTok{1}\NormalTok{:}\DecValTok{6}\NormalTok{])}
\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{)}
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"phenoDisco output"}\NormalTok{, }\DataTypeTok{col =} \NormalTok{cols)}
\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{col =} \NormalTok{cols)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plotPDres-1.png}
\caption{plot of chunk plotPDres}
\end{figure}

The five new phenotype data clusters can be extracted and examined. In
the code chunk below we write the results to a .csv file. We use the
argument \texttt{fDataCols} to specify which columns of the
\texttt{featureData} to write.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{fData}\NormalTok{(hl)$pd <-}\StringTok{ }\KeywordTok{as.character}\NormalTok{(}\KeywordTok{fData}\NormalTok{(hl)$pd)}
\KeywordTok{write.exprs}\NormalTok{(hl, }\DataTypeTok{fDataCols =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"pd-results.csv"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{","}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We can also examine the each phenotype intercatively and visualise their
protein profiles by using the \texttt{pRolocVis} function in the
\texttt{pRolocGUI}. We found that phenotype 1 was enriched in nucleus
associated proteins, phenotype 2 in chromatin associated proteins,
phenotype 3 in cytosolic and phenotypes 4 and 5 in lysosomal and
endosomal proteins.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{pRolocVis}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hyperdef{}{supervised-machine-learning}{\section{Supervised machine
learning}\label{supervised-machine-learning}}

Supervised machine learning, also known as classification, is an
essential tool for the assignment of proteins to distinct sub-cellular
niches. Using a set of labelled training examples i.e.~markers, we can
train a machine learning classifier to learn a mapping between the data
i.e.~the quantitative protein profiles, and a known localisation. The
trained classifier can then be used to predict the localisation of a
protein of unknown localisation, based on its observed protein profile.
To date, this method has been extensively used in spatial quantitative
proteomics to assign thousands of proteins to distinct sub-cellular
niches {[}@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006;
@Tan:2009{]}.

There are several classification algorithms readily available in
\texttt{pRoloc}, which are documented in the dedicated
\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\texttt{pRoloc}
machine learning techniques vignette}. We find the general tendancy to
be that it is not the choice of classifier, but the improper
optimisation of the algorithmic parameters, that limits classification
accuracy. Before employing any classification algorithm and generating a
model on the training data, one must first find the optimal parameters
for the algorithm of choice.

\subsection{Optimisation}\label{optimisation}

In the code chunk below we employ the use of a Support Vector Machine
(SVM) to learn a classifier on the labelled training data. As previously
mentioned one first needs to train the classifiers parameters before an
algorithm can be used to predict the class labels of the proteins with
unknown location. One of the most common ways to optimise the parameters
of a classifier is to partition the labelled data in to training and
testing subsets. In this framework parameters are tested via a grid
search using cross-validation on the training partition. The best
parameters chosen from the cross-validation stage are then used to build
a classifier to predict the class labels of the protein profiles on the
test partition. Observed and expected classication results can be
compared, and then used to assess how well a given model works by
getting an estimate of the classiers ability to achieve a good
generalisation i.e.~that is given an unknown example predict its class
label with high accuracy. In \texttt{pRoloc} algorithmic performance is
estimated using stratified 80/20 partitioning for the training/testing
subsets respectively, in conjuction with five-fold cross-validation in
order to optimise the free parameters via a grid search. This procedure
is usually repeated 100 times and then the best parameter(s) are
selected upon investigation of classifier accuracy, here we use the
harmonic mean of precision and recall; the macro F1 score. In the code
chunk below we demonstrate how to optimise the free parameters;
\texttt{sigma} and \texttt{cost}, of a classical SVM classifier with a
Gaussian kernel using the function \texttt{svmOptimisation}. As the
number of labelled instances per class varies from organelle to
organelle, we can account for class imbalance by setting specific class
weights when generating the SVM model. Below the weights, \texttt{w} are
set to be inversely proportional to the class frequencies.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{w <-}\StringTok{ }\KeywordTok{table}\NormalTok{(}\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## organelleMarkers
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      27                      43                      13 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                      43                      95                      12 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      10                      33                     383 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                      64                      85                      17 
##         Plasma membrane              Proteasome                 unknown 
##                      51                      34                    4122
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{w <-}\StringTok{ }\DecValTok{1}\NormalTok{/w[}\KeywordTok{names}\NormalTok{(w) !=}\StringTok{ "unknown"}\NormalTok{]}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## 100 rounds of optimisation with five-fold cross-validation}
\NormalTok{params <-}\StringTok{ }\KeywordTok{svmOptimisation}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"markers"}\NormalTok{,}
                          \DataTypeTok{times =} \DecValTok{100}\NormalTok{, }\DataTypeTok{xval =} \DecValTok{5}\NormalTok{,}
                          \DataTypeTok{class.weights =} \NormalTok{w)}
\end{Highlighting}
\end{Shaded}

As mentioned previously, we reply on the default feature variable
\texttt{"markers"} to define the class labels and hence can ommit it. To
use another feature variables, one need to explicitly specify its name
using the \texttt{fcol} argument (for example
\texttt{fcol = "markers2"}).

The output \texttt{params} is an object of class \texttt{GenRegRes}; a
dedicated container for the storage of the design and results from a
machine learning optimisation. To assess classifier performance we can
examine the macro F1 scores and the most frequently chosen parameters. A
high macro F1 score indicates that the marker proteins in the test
dataset are consistently and correctly assigned by the the algorithm.
Often more than one parameter or set of parameters gives rise to the
best generalisation accuracy. As such it is always important to
investigate the model parameters and critically assess the best choice.
The best choice may not be as simple as the parameter set that gives
rise to the highest macro F1 score and one must be careful to avoid
overfitting and to choose parameters wisely.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{(best <-}\StringTok{ }\KeywordTok{getParams}\NormalTok{(params))}
\KeywordTok{plot}\NormalTok{(params)}
\KeywordTok{levelPlot}\NormalTok{(params)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/visualiseOptHide-1.png}
\caption{plot of chunk visualiseOptHide}
\end{figure}

By using the function \texttt{getParams} we can extract the best set of
parameters. Currently, \texttt{getParams} retrieves the first best set
is automatically but users are encouraged to critically assess whether
this is the most wise choice by visualising the results using the
methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method
for \texttt{GenRegRes} object shows the respective distributions of the
100 macro F1 scores for the best cost/sigma parameter pairs, and
\texttt{levelPlot} shows the averaged macro F1 scores, for the full
range of parameter values. Once we have selected the best parameters we
can then use them to build a classifier from the labelled marker
proteins.

\subsection{Classification}\label{classification}

We can use the function \texttt{svmClassification} to return a
classification result for all unlabelled instances in the dataset
corresponding to their most likely sub-cellular location. The algorithm
parameters are passed to the function, along with the class weights. As
above, \texttt{fcol} can be ignored as we use the labels defined in the
default \texttt{"markers"} feature variable.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl <-}\StringTok{ }\KeywordTok{svmClassification}\NormalTok{(hl, params,}
                        \DataTypeTok{class.weights =} \NormalTok{w,}
                        \DataTypeTok{fcol =} \StringTok{"markers"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Automatically, the output of the above classification; the organelle
predictions and assignment scores, are stored in the
\texttt{featureData} slot of the \texttt{MSnSet}. In this case, they are
given the labels \texttt{svm} and \texttt{svm.scores} for the
predictions and scores respectively. The resultant predictions can be
visualised using \texttt{plot2D}. In the code chunk below
\texttt{plot2D} is called to generate a PCA plot of the data and
\texttt{fcol} is used to specify where the new assignments are located
e.g. \texttt{fcol =  "svm"}.

Additionally, when calling \texttt{plot2D} we can use the \texttt{cex}
argument to change the size of each point on the plot (where one point
represents one protein) to be inversely proportional to the SVM score.
This gives an initial overview of the high scoring localisations from
the SVM predictions.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## set point size of each protein to be inversely proportional to the}
\NormalTok{ptsze <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(}\KeywordTok{fData}\NormalTok{(hl)$svm.scores) -}\StringTok{ }\DecValTok{1}

\NormalTok{## plot new predictions}
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{ptsze)}
\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }\DataTypeTok{where =} \StringTok{"bottomleft"}\NormalTok{, }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plotSVM-1.png}
\caption{plot of chunk plotSVM}
\end{figure}

The adjustment of the point size intuitively confers important
information that is more difficult to defined formally, but that we will
address in the next section. The classifier (SVM in our case, but this
is also valid of other classifiers) defines boundaries based on the
labelled marker proteins. These class/organelle boundaries define how
non-assigned proteins are classified and with what confidence.
\textbf{TODO} explain how choosing markers positively/negatively affects
this, and relation with thresholding.

\hyperdef{}{thresholding}{\subsection{Thresholding}\label{thresholding}}

It is common when applying a supervised classification algorithm to set
a specific score cutoff on which to define new assignments, below which
classifications are set to unknown/unassigned. This is important as in a
supervised learning setup proteins can only be predicted to be localised
to one of the sub-cellular niches that appear in the labelled training
data. We can not guaranette (and do not expect) that the whole class
diversity to be represented in the labelled training data as (1) finding
markers that represent the whole diversity of the cell is challenging
(especially obtaining dual- and multiply-localised protein markers) and
(2) many sub-cellular niches contain too few proteins to train on (we
recommend a minimum of 13 markers per sub-cellular class for stratified
80/20 partitioning and 5-fold cross-validation - this allows a minimum
of \textasciitilde{}10 examples for parameter optimisation on the
training partition i.e.~2 per fold for 5-fold cross-validation, and then
\textasciitilde{}3 for testing the best parameters on the validation
set).

Deciding on a threshold is not trivial as classifier scores are heavily
dependent upon the classifier used and different sub-cellular niches can
exhibit different score distributions, as highlighted in the boxplot
below. We recommend users to set class-specific thresholds. In the code
chunk below we display a boxplot of the score distributions per
organelle.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## First remove the markers}
\NormalTok{preds <-}\StringTok{ }\KeywordTok{unknownMSnSet}\NormalTok{(hl)}

\NormalTok{## Plot a boxplot of the scores of each organelle}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\FloatTok{10.5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
\KeywordTok{boxplot}\NormalTok{(svm.scores ~}\StringTok{ }\NormalTok{svm, }\DataTypeTok{data =} \KeywordTok{fData}\NormalTok{(preds), }\DataTypeTok{ylab =} \StringTok{"SVM scores"}\NormalTok{, }\DataTypeTok{las =} \DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/thresholds-1.png}
\caption{plot of chunk thresholds}
\end{figure}

There are many ways to set thresholds and the choice of method will
depend on the biological question and experimental design at hand. One
viable approach in the frame of the above experimetal design would be to
manually set a FDR, say 5\%, per organelle. To do this the user would
examine the top scoring predictions for each organelle, and then set a
threshold at the score at which they achieve 5\% of false assignments
per organelle. The definintion of a false assignment would depend on the
information available, for example, validity or lack of validity for the
localisation from another experiment as reported in the literature or a
reliable database. If such information is not available, one crude
method is to set a threshold per organelle by extracting the median or
3rd quantile score per organelle. For example, in the code chunk below,
we use the \texttt{orgQuants} function to extract the median organelle
scores and then pass these scores to the \texttt{getPrediction} function
to extract the new localisations that meet this scoring criteria. Any
sub-cellular predictions that fall below the specified thresholds are
labelled as unknown.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ts <-}\StringTok{ }\KeywordTok{orgQuants}\NormalTok{(hl, }
                \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }
                \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{,}
                \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{, }
                \DataTypeTok{t =} \NormalTok{.}\DecValTok{5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##               0.4356698               0.3028843               0.3825420 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##               0.6825826               0.5995243               0.4245311 
##    Extracellular matrix                Lysosome           Mitochondrion 
##               0.4034908               0.5882168               0.9501241 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##               0.7951459               0.7096694               0.3149301 
##         Plasma membrane              Proteasome 
##               0.7091667               0.4204304
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{hl <-}\StringTok{ }\KeywordTok{getPredictions}\NormalTok{(hl,}
                     \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{,}
                     \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{, }
                     \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{,}
                     \DataTypeTok{t =} \NormalTok{ts)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## ans
##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
##                      86                     169                      86 
##                 Cytosol   Endoplasmic reticulum                Endosome 
##                     295                     479                      91 
##    Extracellular matrix                Lysosome           Mitochondrion 
##                      26                     125                     522 
##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
##                     230                     344                      38 
##         Plasma membrane              Proteasome                 unknown 
##                     325                     159                    2057
\end{verbatim}

The output of \texttt{getPredictons} is the original \texttt{MSnSet}
dataset with a new feature variable appended to the feature data called
\texttt{fcol.pred} (i.e.~in our case \texttt{svm.pred}) containing the
prediction results. The results can also be visualied using
\texttt{plot2D} function.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{dev.off}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## null device 
##           1
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm.pred"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plotres-1.png}
\caption{plot of chunk plotres}
\end{figure}

There is also a dedicated interactive application to help users examine
these distributions in the
\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
package. This app can be launched via the \texttt{pRolocVis} function
and specifying the argument \texttt{app = "classify"} along with the
relevent \texttt{fcol}, \texttt{scol} and \texttt{mcol} which refer to
the columns in the feature data that contain the new assignments,
assignment scores and markers respectively (see also
\texttt{fvarLabels(svmres)}).

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocGUI"}\NormalTok{)}
\KeywordTok{pRolocVis}\NormalTok{(hl, }
          \DataTypeTok{app =} \StringTok{"classify"}\NormalTok{,}
          \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{,}
            \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{,}
          \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{./Figures/classify.png}
\caption{The classify application}
\end{figure}

The data is loaded and displayed on a PCA plot and a boxplot is used to
display the classifier scores by data class. On the left there is a
sidebar panel with sliders to control the thresholds upon which
classifications are made. There are two types of cut-off that the user
can choose from: (1) ``Quantile'' and (2) ``User-defined''. By default,
when the application is launched quatile scoring is selected and set to
0.5, the median. The class-specific score thresholds that correspond to
selecting the desired quantile are shown on as red dots on the boxplot.
The assignments on the PCA plot are also updated according to the
selected threshold. The quantile threshold can be set by moving the
corresponding quantile slider. If one wished to set their own cut-offs
the ``User-defined'' radio button must be selected and then the sliders
for defining user-specified scores become active and the scores and
highlighted on the boxplot by blue dots. For more information we refer
users to the
\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
tutorial
\href{http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html}{vignette}.

\subsection{Transfer learning}\label{transfer-learning}

In addition to high quality MS-based quantitative proteomics data there
exist a number of other sources of information that are freely available
in the public domain that may be useful to assign a protein to its
sub-cellular niche. For example, imaging from immunofluorescence
microscopy, protein annotations and sequences, and protein-protein
interactions among others, represent a rich and vast source of
complementary information. We can integrate this auxiliary information
with our primary MS-based quantitative data using a paradigm known as
transfer learning (TL). The integration of data between different
technologies is one of the biggest challenges in computational biology
to date and the
\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} package
provides functionality to do such analyses. We recently developed two
transfer learning algorithms using a \emph{k}-NN and SVM framework and
applied them to the task of protein localisation prediction
{[}@Breckels:2016{]}. In this section we will begin with explaining the
concept of transfer learning and then show how to apply this in the
frame of spatial proteomics and protein localisation prediction.

In TL one typically has a primary task that they wish to solve, and some
complementary (often heterogeneous) auxiliary information that is
related to the primary learning objective, that can be used to help
solve the primary goal. For example, here our primary task is to assign
proteins to their sub-cellular niche with high generalisation accuracy
from data collected from quantitative MS-based experiments. In the
example below we extract Gene Ontology (GO) information to use as an
auxiliary data source to help solve our task of protein localisation
prediction.

Using the functions \texttt{setAnnotationParams} and \texttt{makeGoSet}
we can contruct an auxiliary \texttt{MSnSet} of GO terms, from the
primary data's features i.e.~the protein accession numbers. All the GO
terms associated to each accession number are retrieved and used to
create a binary matrix where a 1 (0) at position (i; j) indicates that
term j has (not) been used to annotate protein i. The GO terms are
retrieved from an appropriate repository using the
\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}} package.
The specific
\emph{\href{http://bioconductor.org/packages/Biomart}{Biomart}}
repository and query will depend on the species under study and the type
of identifiers. The first step is to construct the annotation parameters
that will enable to perform the query, this is done using
\texttt{setAnnotationParams}. Typing into the R console
\texttt{par \textless{}- setAnnotationParams()} will present two menus,
firstly asking you to identify the species of study, and then what type
of identifier you have used to annotate the proteins in your
\texttt{MSnSet}. It is also possible to pass patterns to match the
species e.g.~in the code chunk below we pass ``Mus musculus'', and the
identifier type for our data (see \texttt{featureNames(hl)}) which is
``Uniprot/Swissprot'', for the \texttt{Biocpkg("Biomart")} query.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{par <-}\StringTok{ }\KeywordTok{setAnnotationParams}\NormalTok{(}\DataTypeTok{inputs =} \KeywordTok{c}\NormalTok{(}\StringTok{"Mus musculus"}\NormalTok{, }
                                      \StringTok{"UniProt/Swissprot"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

Now we have contructed the query parameters we can use the
\texttt{makeGoSet} function to retrieve and build an auxiliary GO
\texttt{MSnSet} as described above. By default, the cellular component
terms are downloaded, without any filtering on evidence codes. It is
also possible to download terms from the molecular function and
biological process GO namespaces, and also apply filtering based on
evidence codes as desired, see \texttt{?makeGoSet} for more details.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gocc <-}\StringTok{ }\KeywordTok{makeGoSet}\NormalTok{(hl, }\DataTypeTok{params =} \NormalTok{par, }
                  \DataTypeTok{namespace =} \StringTok{"cellular_component"}\NormalTok{, }
                  \DataTypeTok{evidence =} \OtherTok{NULL}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The function \texttt{makeGoSet} uses the
\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}} package
to query the relevent database (e.g.~Ensembl, Uniprot) for GO terms. All
GO terms that have been observed for the 5032 proteins in the hyperLOPIT
dataset are retieved. Users should note that the number of GO terms
retreived is also dependent on the database version queried and thus is
always subject to change. We find it is common to see many GO terms with
only one protein assigned to that term, these such terms generally do
not bring and information generally for building the classifier and thus
we can remove such GO terms by using the function
\texttt{filterBinMSnSet}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{gocc <-}\StringTok{ }\KeywordTok{filterBinMSnSet}\NormalTok{(hl)}
\end{Highlighting}
\end{Shaded}

Now we have generated our auxiliary data we can use the \emph{k}-NN
implementation of transfer learning available in
\texttt{Biocpkg("pRoloc")} to integrate this with our primary MS-based
quantitative proteomics data using the functions
\texttt{knntlOptimisation} to estimate the free-parameters for the
integration, and \texttt{knntlClassification} to do the predictions. We
have shown that using transfer learning in the context of spatial
proteomics results in the assignment of proteins to sub-cellular niches
with a higher generalisation accuracy than using standard supervised
machine learning with a single source of information
{[}@Breckels:2016{]}.

The first step, as with any machine learning algorithm, is to optimise
any free paramaters of the classifier. For the \emph{k}-NN TL classifier
there are two sets of parameters that need optimising: the first set are
the $k$'s for the primary and auxiliary data sources required for the
nearest neighbour calculations for each data source. The second set of
parameters (which is noted by a vector of $\theta$ weights) that require
optimising are the class weights, one per subcellular niche, that
control the proportion of primary and auxiliary data to use for
learning. A weight can take any real value number between 0 and 1. A
weight of $\theta = 1$ indicates that all weight is given to the primary
data (and this implicitly implies that a weight of $1 - \theta$ is given
to the auxiliary data, similarly a weight of $\theta = 0$ implies that
all weight is given to the auxiliary data (so 0 is given to the primary
source). If we conduct a parameter seach and test weights
$\theta = {0, 1/3, 2/3, 1}$ for each class, and if we have, for example
10 subcellular niches, this will result in \texttt{4\^{}10} different
combinations of parameters to test. The parameter optimisation is
therefore time consuming and as such we recommend you make use of a
computing cluster (code and submissing scripts are also available in the
supporting information). The markers in the \texttt{hl} dataset contain
14 subcellular classes. If we examine these markers and classes on the
PCA plot above we can see that in particular the two ribosomes and two
nuclear compartments are highly separated along the first two
components, this is also evident from the profiles plot which gives us a
good indication that these subcellular niches are well-resolved in the
hyperLOPIT dataset. Transfer learning is particularly useful for classes
that are not as well separated, we find that subcellular niches that are
well-separated under hyperLOPIT and LOPIT obtain a class score of 1
(i.e.~use only primary data from transfer learning
{[}@Breckels:2016{]}). Therefore, for the optimisation stage of the
analyses we can already infer a subcellular class weight of 1 for these
niches and only optimise over the remaining organelles. This can
signifciantly cut down optimisation time as by removing these 4 classes
from the optimisation (and not the classification) we only have
\texttt{4\^{}10} class weight combinations to consider instead of
\texttt{4\^{}14} combinations.

In the example below we remove these 4 classes from the marker set,
re-run the \texttt{knnOptimisation} for each data source and then run
the \texttt{knntlOptimisation} with the 10 remaining classes. (Note:
this is not run live as this the \texttt{hl} dataset with 10 classes,
707 markers and \texttt{4\^{}10} combinations of parameters takes
\textasciitilde{}76 hours to run on the University of Cambridge HPC
using 256 workers).

To remove the 4 classes and create a new column of markers in the
feature data called \texttt{tlmarkers} to use for the analysis:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## create new markers column for tl markers}
\KeywordTok{fData}\NormalTok{(hl)$tlmarkers <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hl)$markers}
\KeywordTok{fData}\NormalTok{(gocc)$tlmarkers <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(gocc)$markers}

\NormalTok{## Remove 4 classes}
\NormalTok{torm  <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"40S Ribosome"}\NormalTok{, }\StringTok{"60S Ribosome"}\NormalTok{,}
           \StringTok{"Nucleus - Chromatin"}\NormalTok{, }
           \StringTok{"Nucleus - Non-chromatin"}\NormalTok{)}
\NormalTok{for (i in }\KeywordTok{seq}\NormalTok{(torm)) \{}
  \NormalTok{hl <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(hl, }\DataTypeTok{from =} \NormalTok{torm[i], }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
  \NormalTok{gocc <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(gocc, }\DataTypeTok{from =} \NormalTok{torm[i], }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
\NormalTok{\}}
\KeywordTok{getMarkerClasses}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] "Actin cytoskeleton"    "Cytosol"              
##  [3] "Endoplasmic reticulum" "Endosome"             
##  [5] "Extracellular matrix"  "Lysosome"             
##  [7] "Mitochondrion"         "Peroxisome"           
##  [9] "Plasma membrane"       "Proteasome"
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{getMarkerClasses}\NormalTok{(gocc, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##  [1] "Actin cytoskeleton"    "Cytosol"              
##  [3] "Endoplasmic reticulum" "Endosome"             
##  [5] "Extracellular matrix"  "Lysosome"             
##  [7] "Mitochondrion"         "Peroxisome"           
##  [9] "Plasma membrane"       "Proteasome"
\end{verbatim}

Optimisation stage 1: run \texttt{knnOptimisation} to get the best $k$'s
for each data source.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## get best k's}
\NormalTok{kpopt <-}\StringTok{ }\KeywordTok{knnOptimisation}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
\NormalTok{kaopt <-}\StringTok{ }\KeywordTok{knnOptimisation}\NormalTok{(gocc, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(kpopt)}
\KeywordTok{plot}\NormalTok{(kaopt)}
\end{Highlighting}
\end{Shaded}

From examining the parameter seach plots we find the best $k$'s for both
the primary and auxiliary are 3.

Optimisation stage 2: run \texttt{knntlOptimisation} to get the best
transfer learning weights for each sub-cellular class.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## Set appropriate parallelisation backend and  }
\NormalTok{## number of workers for the tl}
\NormalTok{par <-}\StringTok{ }\KeywordTok{SnowParam}\NormalTok{(255L, }\DataTypeTok{type =} \StringTok{"MPI"}\NormalTok{)}

\NormalTok{## Now peform tl optimisation }
\NormalTok{tlopt <-}\StringTok{ }\KeywordTok{knntlOptimisation}\NormalTok{(hl, gocc,}
                           \DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{,}
                           \DataTypeTok{length.out =} \DecValTok{4}\NormalTok{, }
                           \DataTypeTok{times =} \DecValTok{50}\NormalTok{,}
                           \DataTypeTok{xval =} \DecValTok{5}\NormalTok{, }\DataTypeTok{k =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{),}
                           \DataTypeTok{BPPARAM =} \NormalTok{par}
  \NormalTok{)}
\end{Highlighting}
\end{Shaded}

The results of the optimisation can be visalised using the \texttt{plot}
method for \texttt{"ThetaRegRes"} objects:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(tlopt)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plottl-1.png}
\caption{plot of chunk plottl}
\end{figure}

Looking at the bubble plot displaying the distribution of best weights
over the 50 runs we find that for many of the subcellular niches a
weight of 1 is most popular (i.e.~use only primary hyperLOPIT data in
classification), this is unsuprising as we already know the dataset is
well resolved for these classes. We see that the most popular weights
for the proteasome and lysosome tend to be towards 0, indicating that
these niches are well-resolved in the Gene Ontology. This tells us that
we would benefit from including auxiliary GO information in our
classifier for these subcellular compartments. The plasma membrane
weights are relatively equally spread between using hyperLOPIT and GO
data. Using the \texttt{getParams} function we can return the best
weights and then use this as input for the classification.

One of the benefits of the algorithm is the ability to manually select
weights for each class. In the optimisation above, for time constraints,
we removed the two ribosomal subunits and the two nulcear compartments,
and therefore in the code chunk below when we extract the best
parameters, these subcellular niches are not included. To include these
4 subcellular niches in the next classification step we must include
them in the parameters so we infer a weight of 1 for each of these
niches as we know they are well resolved in hyperLOPIT. We then re-order
the weights according to \texttt{getMarkerClasses} and perform the
classification using the function \texttt{knntlClassification}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## best parameters for the 10 classes}
\NormalTok{(bestpar <-}\StringTok{ }\KeywordTok{getParams}\NormalTok{(tlopt))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##    Actin cytoskeleton               Cytosol Endoplasmic reticulum 
##             1.0000000             1.0000000             0.6666667 
##              Endosome  Extracellular matrix              Lysosome 
##             1.0000000             1.0000000             0.3333333 
##         Mitochondrion            Peroxisome       Plasma membrane 
##             1.0000000             1.0000000             0.6666667 
##            Proteasome 
##             0.3333333
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## add weights for classes not included in the optimisation}
\NormalTok{otherweights <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{)}
\KeywordTok{names}\NormalTok{(otherweights) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"40S Ribosome"}\NormalTok{, }\StringTok{"60S Ribosome"}\NormalTok{, }
                         \StringTok{"Nucleus - Chromatin"}\NormalTok{, }
                         \StringTok{"Nucleus - Non-chromatin"}\NormalTok{)}
\NormalTok{(bestpar <-}\StringTok{ }\KeywordTok{c}\NormalTok{(bestpar, otherweights))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
##      Actin cytoskeleton                 Cytosol   Endoplasmic reticulum 
##               1.0000000               1.0000000               0.6666667 
##                Endosome    Extracellular matrix                Lysosome 
##               1.0000000               1.0000000               0.3333333 
##           Mitochondrion              Peroxisome         Plasma membrane 
##               1.0000000               1.0000000               0.6666667 
##              Proteasome            40S Ribosome            60S Ribosome 
##               0.3333333               1.0000000               1.0000000 
##     Nucleus - Chromatin Nucleus - Non-chromatin 
##               1.0000000               1.0000000
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## re-order classes }
\NormalTok{bestpar <-}\StringTok{ }\NormalTok{bestpar[}\KeywordTok{getMarkerClasses}\NormalTok{(hl)]}

\NormalTok{## Do the classification}
\NormalTok{hl <-}\StringTok{ }\KeywordTok{knntlClassification}\NormalTok{(hl, gocc, }\DataTypeTok{bestTheta =} \NormalTok{bestpar, }
                          \DataTypeTok{k =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

The results from the classification results and associated scores are
appended to the \texttt{fData} slot and named \texttt{knntl} and
\texttt{knntl.scores} respectively. Results can be visualised using
\texttt{plot2D}, scores assessed and cutoffs calculated using the
\texttt{classify} app in \texttt{pRolocVis}, predictions obtained using
\texttt{getPredictions} in the same way as demonstrated above for the
SVM classifier.

\subsection{Unsupervised machine
learning}\label{unsupervised-machine-learning}

In \texttt{pRoloc} there is functionality for unsupervsied machine
learning methods. In unsupervised learning, the training data consists
of a set of input vectors e.g.~protein profiles, for which we do not
have information about the class label e.g.~localisation. The main goal
in unsupervised learning is to uncover groups of similar examples within
the data, this is termed clustering. Ordinance methods such as principal
components analysis (PCA) also fall into the category of unsupervised
learning methods, where the data can be projected from a
high-dimensional spcae down to two or three dimensions for the purpose
of visualisation.

As described and demonstrated already above, PCA is a valuable and
powerful method for data visualisation and quality control. We do not do
any unsupervised clustering in the frame of the above spatial proteomics
workflow as although clustering can be useful for grouping labelled data
into categories e.g.~see the \texttt{mrkHClust} function, we do not find
it adaquete for spatial proteomics data analysis. We find supervised
learning more suited to the task of protein localisation prediction in
which we use high-quality curated marker proteins to build a classifier,
instead of using an entirely unsupervised approach to look for clusters
and then look for enrichment of organelles and complexes. In the latter
we do not make good use of valuable prior knowledge, and in our
experience unsupervised clustering can be extremely difficult due poor
estimates of the number of clusters that may appear in the data.

\section{Writing and exporting data}\label{writing-and-exporting-data}

A \texttt{MSnSet} can be exported from R using the \texttt{write.exprs}
function. This function writes the expression values to a tab (the
defualt, or other type e.g.e.g.~comma, as specified in
\texttt{write,table}) separated file. There argument \texttt{fDataCols}
can be used to specify which \texttt{featureData} columns (as column
names, column number or \texttt{logical}) to append to the right of the
expression matrix.

In the below code chunk we write the \texttt{hl} object to a csv file.
The \texttt{file} argument is used to specify the file path, the
\texttt{sep} argument specifies the field separator string, here we use
a comma, finally as we want to write all the information in the
\texttt{featureData} to the file, as well as the expression data, we
specify \texttt{fDataCols = 1:ncol(hl)} i.e.~write everything in the
\texttt{featureData} to the file \texttt{"hl.csv"}.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{write.exprs}\NormalTok{(hl, }\DataTypeTok{file =} \StringTok{"hl.csv"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{","}\NormalTok{, }
            \DataTypeTok{fDataCols =} \DecValTok{1}\NormalTok{:}\KeywordTok{ncol}\NormalTok{(hl))}
\end{Highlighting}
\end{Shaded}

Individual R objects, such a parameters from the machine learning
optimisations, can be save using the standard \texttt{save} function in
R. For example, to save and then re-load the parameters from the SVM
optimisation,

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{## To save the parameters as an R object}
\KeywordTok{save}\NormalTok{(params, }\DataTypeTok{file =} \StringTok{"svmparams.rda"}\NormalTok{)}

\NormalTok{## To re-load after saving}
\KeywordTok{load}\NormalTok{(}\DataTypeTok{file =} \StringTok{"svmparams.rda"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\section{Session information}\label{session-information}

The function \texttt{sessionInfo} provides a summary of all packages and
versions used to generate this document. This enables us to record the
exact state of our session that lead to these exact results. Conversely,
if the script stops working of if it returns different results, we are
in a position to re-generate the original results using the adequate
software versions and retrace changes in the software that lead to
failure and/or different results.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{sessionInfo}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## R version 3.3.1 Patched (2016-08-02 r71022)
## Platform: x86_64-pc-linux-gnu (64-bit)
## Running under: Ubuntu 14.04.5 LTS
## 
## locale:
##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
## 
## attached base packages:
## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
## [8] datasets  base     
## 
## other attached packages:
##  [1] pRolocdata_1.11.9    pRoloc_1.13.15       MLInterfaces_1.53.1 
##  [4] cluster_2.0.4        annotate_1.51.1      XML_3.98-1.4        
##  [7] AnnotationDbi_1.35.4 IRanges_2.7.15       S4Vectors_0.11.17   
## [10] MSnbase_1.99.2       ProtGenerics_1.5.1   BiocParallel_1.7.8  
## [13] mzR_2.7.6            Rcpp_0.12.7          Biobase_2.33.3      
## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.32    
## [19] knitr_1.14          
## 
## loaded via a namespace (and not attached):
##  [1] minqa_1.2.4           colorspace_1.2-6      hwriter_1.3.2        
##  [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
##  [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-16         
## [10] MatrixModels_0.4-1    affyio_1.43.0         flexmix_2.3-13       
## [13] mvtnorm_1.0-5         codetools_0.2-14      splines_3.3.1        
## [16] doParallel_1.0.10     impute_1.47.0         robustbase_0.92-6    
## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-71         
## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-24       
## [25] vsn_3.41.0            sfsmisc_1.1-0         shiny_0.14           
## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
## [31] limma_3.29.21         formatR_1.4           htmltools_0.3.5      
## [34] quantreg_5.29         tools_3.3.1           ggvis_0.4.3          
## [37] gtable_0.2.0          affy_1.51.1           reshape2_1.4.1       
## [40] dplyr_0.5.0           MALDIquant_1.15       trimcluster_0.1-2    
## [43] gdata_2.17.0          preprocessCore_1.35.0 nlme_3.1-128         
## [46] iterators_1.0.8       fpc_2.1-10            stringr_1.1.0        
## [49] lme4_1.1-12           lpSolve_5.6.13        mime_0.5             
## [52] gtools_3.5.0          dendextend_1.3.0      DEoptimR_1.0-6       
## [55] zlibbioc_1.19.0       MASS_7.3-45           scales_0.4.0         
## [58] BiocInstaller_1.23.9  pcaMethods_1.65.0     SparseM_1.72         
## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.29.2       
## [64] rpart_4.1-10          stringi_1.1.1         RSQLite_1.0.0        
## [67] genefilter_1.55.2     randomForest_4.6-12   foreach_1.4.3        
## [70] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
## [73] mzID_1.11.2           evaluate_0.9          lattice_0.20-34      
## [76] htmlwidgets_0.7       gbm_2.1.1             plyr_1.8.4           
## [79] magrittr_1.5          R6_2.1.3              DBI_0.5-1            
## [82] whisker_0.3-2         mgcv_1.8-15           survival_2.39-5      
## [85] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
## [88] msdata_0.12.1         car_2.1-3             mlbench_2.1-1        
## [91] grid_3.3.1            FNN_1.1               threejs_0.2.2        
## [94] digest_0.6.10         diptest_0.75-7        xtable_1.8-2         
## [97] httpuv_1.3.3          munsell_0.4.3
\end{verbatim}

It is always important to include session information details along with
a \href{http://adv-r.had.co.nz/Reproducibility.html}{short reproducible
example} highlighting the problem or
\href{https://support.bioconductor.org/}{question} at hand.

### 2016-09-29 main.tex

@@
 
+\usepackage{hyperref}
+
 \begin{document}
@@
 Please ensure the broadness and claims within the title are appropriate to the content of the article itself.}
-\author[1]{Author Name1}
-\author[2]{Author Name2}
-\affil[1]{Address of author1}
-\affil[2]{Address of author2}
+\author[1,2]{Lisa M. Breckels}
+\author[1,2]{Laurent Gatto}
+\affil[1]{Computational Proteomics Unit, Cambridge Systems Biology Centre, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
+\affil[2]{Cambridge Centre for Proteomics, Department of Biochemistry, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
 
@@
 
-Please list all authors that played a significant role in the research involved in the article.
-Please provide full affiliation information (including full institutional address, ZIP code and e-mail address) for all authors, and identify who is/are the corresponding author(s).
-
 \begin{abstract}
-
 Abstracts should be up to 300 words and provide a succinct summary of the article.
@@
 Citations should not be used in the abstract, and the use of abbreviations should be minimized.
-
 \end{abstract}
@@
 
-\section*{Introduction}
-
-The format of the main body of the article is flexible: it should be concise and in the format most appropriate to displaying the content of the article.
-
-Some examples of commonly used \LaTeX{}  commands and features are listed below, to help you get started.
-
-\subsection*{Sections}
-
-Use section and subsection commands to organize your document. \LaTeX{} handles all the formatting and numbering automatically. Use ref and label commands for cross-references.
-
-\subsection*{Tables}
-
-Use the table and tabledata commands for basic tables --- see Table~\ref{tab:widgets}, for example.
-\begin{table}[h!]
-\hrule \vspace{0.1cm}
-\caption{\label{tab:widgets}An example of a simple table with caption.}
-\centering
-\begin{tabledata}{llr} 
-\header First name & Last Name & Grade \\ 
-\row John & Doe & $7.5$ \\ 
-\row Richard & Miles & $2$ \\ 
-\end{tabledata}
-\end{table}
-
-\subsection*{Figures}
-You can upload a figure (JPEG, PNG or PDF) using the files menu. To include it in your document, use the includegraphics command (see the example below in the source code).
-
-Please give figures appropriate filenames eg: figure1.pdf, figure2.png.
-
-Figure legends should briefly describe the key messages of the figure such that the figure can stand alone from the main text.
-However, all figures should also be discussed in the article text.
-Each legend should have a concise title of no more than 15 words.
-The legend itself should be succinct, while still explaining all symbols and abbreviations.
-Avoid lengthy descriptions of methods.
-
-For any figures reproduced from another publication (as long as appropriate permission has been obtained from the copyright holder —see under the heading 'Submission'), please include a line in the legend to state that: 'This figure has been reproduced with kind permission from [include original publication citation]'.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.4\textwidth]{frog.jpg}
-\caption{\label{fig:your-figure}Your figure legend goes here; it should be succinct, while still explaining all symbols and abbreviations. }
-\end{figure}
-
-\subsection*{Mathematics}
-
-\LaTeX{} is great at typesetting mathematics.
-Let $X_1, X_2, \ldots, X_n$ be a sequence of independent and identically distributed random variables with $\text{E}[X_i] = \mu$ and $\text{Var}[X_i] = \sigma^2 < \infty$, and let
-$$S_n = \frac{X_1 + X_2 + \cdots + X_n}{n}
-      = \frac{1}{n}\sum_{i}^{n} X_i$$
-denote their mean. Then as $n$ approaches infinity, the random variables $\sqrt{n}(S_n - \mu)$ converge in distribution to a normal $\mathcal{N}(0, \sigma^2)$.
-
-\subsection*{Lists}
-
-You can make lists with automatic numbering \dots
-
-\begin{enumerate}
-\item Like this,
-\item and like this.
-\end{enumerate}
-\dots or bullet points \dots
-\begin{itemize}
-\item Like this,
-\item and like this.
-\end{itemize}
-
-\section*{Methods}
-This section is not essential for Web Tool papers.
-
-Methods should include a brief discussion of allowances made (if any) for controlling bias or unwanted sources of variability, and the limitations of the datasets.
-
-\section*{Results}
-This section is not essential for Web Tool papers.
-
-\section*{Discussion}
-The discussion should include the implications of the article results in view of prior work in this field.
-
-\section*{Conclusions}
-Please state what you think are the main conclusions that can be realistically drawn from the findings in the paper, taking care not to make claims that cannot be supported.
-
-\subsection*{Author contributions}
-In order to give appropriate credit to each author of an article, the individual
-contributions of each author to the manuscript should be detailed in this section. We
-recommend using author initials and then stating briefly how they contributed.
-
-\subsection*{Competing interests}
-All financial, personal, or professional competing interests for any of the authors that
-could be construed to unduly influence the content of the article must be disclosed and
-will be displayed alongside the article.
-
-\subsection*{Grant information}
-Please state who funded the work discussed in this article, whether it is your employer,
-a grant funder etc. Please do not list funding that you have that is not relevant to this
-specific piece of research. For each funder, please state the funder’s name, the grant
-number where applicable, and the individual to whom the grant was assigned.
-If your work was not funded by any grants, please include the line: ‘The author(s)
-declared that no grants were involved in supporting this work.’
-
-\subsection*{Acknowledgements}
-This section should acknowledge anyone who contributed to the research or the
-article but who does not qualify as an author based on the criteria provided earlier
-(e.g. someone or an organisation that provided writing assistance). Please state how
-they contributed; authors should obtain permission to acknowledge from all those
-mentioned in the Acknowledgements section.
-
-Please do not list grant funding in this section.
-
-{\small\bibliographystyle{unsrtnat}
-\bibliography{sample}}
-
-\bigskip
-References can be listed in any standard referencing style that uses a numbering system
-(i.e. not Harvard referencing style), and should be consistent between references within
-a given article.
-
-Reference management systems such as Zotero provide options for exporting bibliographies as Bib\TeX{} files.
-Bib\TeX{} is a bibliographic tool that is used with \LaTeX{} to help organize the user's references and create a bibliography.
-This template contains an example of such a file, \texttt{sample.bib}, which can be replaced with your own.
-Use the \verb|\cite| command  to create in-text citations, like this \cite{Smith:2012qr} and this \cite{Smith:2013jd}.
-
-% See this guide for more information on BibTeX:
-% http://libguides.mit.edu/content.php?pid=55482&sid=406343
-
-% For more author guidance please see:
-% http://f1000research.com/author-guidelines
-
-% When all authors are happy with the paper, use the 
-% ‘Submit to F1000Research' button from the menu above
-% to submit directly to the open life science journal F1000Research.
-
-% Please note that this template results in a draft pre-submission PDF document.
-% Articles will be professionally typeset when accepted for publication.
-
-% We hope you find the F1000Research Overleaf template useful,
-% please let us know if you have any feedback using the help menu above.
+\include{bioc-workflow}
 

### 2016-09-29 bioc-workflow.Rmd
fix wrong citation

@@
 readers are invited to read the `r Biocexptpkg("RforProteomics")`
-package vignettes and associated papers [@Gatto:2014;@Gatto2015].
+package vignettes and associated papers [@Gatto:2014;@Gatto:2015].
 

### 2016-09-29 bioc-workflow.tex
fix wrong citation

@@
 spectrometry data. We will focus on a recent pluripotent mouse embryonic
-stem cells experiment {[}@hyper{]}. Additional annotated and
-pre-formatted datasets from various species are readily available in the
+stem cells experiment (A. Christoforou et al. 2016). Additional
+annotated and pre-formatted datasets from various species are readily
+available in the
 \emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
@@
 \emph{\href{http://bioconductor.org/packages/RforProteomics}{RforProteomics}}
-package vignettes and associated papers {[}@Gatto:2014;@Gatto2015{]}.
+package vignettes and associated papers (L. Gatto and Christoforou 2014;
+Gatto et al. 2015).
 
@@
 As a use-case, we analyse a recent high-throughput spatial proteomics
-dataset from pluripotent mouse embryonic stem cells (E14TG2a)
-{[}@hyper{]}. The data was generated using hyperplexed LOPIT
-(hyperLOPIT), a state-of-the-art method relying on improved sub-cellular
-fractionation and more accurate quantitation, leading to more reliable
-classification of protein localisation across the whole sub-cellular
-space. The method uses an elaborate sub-cellular fractionation scheme,
-enabled by the use of Tandem Mass Tag (TMT) {[}Thompson:2003{]} 10-plex
-and application of the MS data acquisition technique named synchronous
-precursor selection MS\textsuperscript{3} (SPS-MS\textsuperscript{3})
-{[}@McAlister:2014{]}, for TMT quantification with high accuracy and
-precision. Three biological replicates were generated from the E14TG2a
-experiment, the first was to target low density fractions and the second
-and third were to emphasis separation of the denser organelles. The
-intersect of replicates 1 and 2 was treated as a 20-plex dataset for the
-analysis. As discussed in the manuscript {[}@hyper{]} as it has been
-shown that combining replicates across from different gradients can
-increase spatial resolution {[}@Trotter:2010{]}. The combination of
+dataset from pluripotent mouse embryonic stem cells (E14TG2a) (A.
+Christoforou et al. 2016). The data was generated using hyperplexed
+LOPIT (hyperLOPIT), a state-of-the-art method relying on improved
+sub-cellular fractionation and more accurate quantitation, leading to
+more reliable classification of protein localisation across the whole
+sub-cellular space. The method uses an elaborate sub-cellular
+fractionation scheme, enabled by the use of Tandem Mass Tag (TMT)
+{[}Thompson:2003{]} 10-plex and application of the MS data acquisition
+technique named synchronous precursor selection MS\textsuperscript{3}
+(SPS-MS\textsuperscript{3}) (McAlister et al. 2014), for TMT
+quantification with high accuracy and precision. Three biological
+replicates were generated from the E14TG2a experiment, the first was to
+target low density fractions and the second and third were to emphasis
+separation of the denser organelles. The intersect of replicates 1 and 2
+was treated as a 20-plex dataset for the analysis. As discussed in the
+manuscript (A. Christoforou et al. 2016) as it has been shown that
+combining replicates across from different gradients can increase
+spatial resolution (M. W. B. Trotter et al. 2010). The combination of
 replicates resulted in 5032 proteins common in both experiments.
@@
 containing the quantitation data from the intersect of replicates 1 and
-2 of the mouse map {[}@hyper{]}, using the \texttt{readMSnSet2}
-function. The data is as available online with the manuscript (see tab 2
-of the \texttt{xlsx} supplementary data set 1 in {[}@hyper{]}, which
-should be exported as a text-based spreadsheet). It is also available as
-a \texttt{csv} in the Bioconductor \texttt{r Biocexptpkg("pRolocdata")}
-data package.
+2 of the mouse map (A. Christoforou et al. 2016), using the
+\texttt{readMSnSet2} function. The data is as available online with the
+manuscript (see tab 2 of the \texttt{xlsx} supplementary data set 1 in
+(A. Christoforou et al. 2016), which should be exported as a text-based
+spreadsheet). It is also available as a \texttt{csv} in the Bioconductor
+\texttt{r Biocexptpkg("pRolocdata")} data package.
 
@@
 be a sample originating from an alternative preparation e.g.~such as
-from the chromatin enrichment performed in Christoforou et al
-{[}@hyper{]}. Information about which gradient fractions were used for
-which tag should also be stored in the sample meta-data \texttt{pData}
-slot.
+from the chromatin enrichment performed in Christoforou et al (A.
+Christoforou et al. 2016). Information about which gradient fractions
+were used for which tag should also be stored in the sample meta-data
+\texttt{pData} slot.
 
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 13:42:11 2016 
 ##  MSnbase version: 1.21.7
@@
 Missing data are a recurrent issue in mass spectrometry applications,
-and should be addressed independently of this workflow
-{[}@Webb-Robertson:2015;@Lazar:2016{]}. In {[}@Gatto:2014b{]}, we have
+and should be addressed independently of this workflow (Webb-Robertson
+et al. 2015; Lazar et al. 2016). In (Gatto et al. 2014), we have
 described how a high content in missing values in spatial proteomics
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
-## Subset [6725,20][5032,20] Thu Sep 29 10:28:15 2016 
-## Removed features with more than 0 NAs: Thu Sep 29 10:28:15 2016 
-## Dropped featureData's levels Thu Sep 29 10:28:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 13:42:11 2016 
+## Subset [6725,20][5032,20] Thu Sep 29 13:42:11 2016 
+## Removed features with more than 0 NAs: Thu Sep 29 13:42:11 2016 
+## Dropped featureData's levels Thu Sep 29 13:42:11 2016 
 ##  MSnbase version: 1.21.7
@@
 sub-cellular niches have been separated along the gradient. Based on De
-Duve's principle {[}@DeDuve:1981{]} proteins that co-localise exhibit
-similar quantitation profiles across the gradient fractions employed.
-One approach that has been widely used to visualise and inspect high
-throughput mass spectrometry-based proteomics data is principal
-components analysis (PCA). PCA is one of many dimensionality reduction
-methods, that allow one to effectively summarise multi-dimensional data
-in to 2 or 3 dimensions to enable visualisation. Very generally, the
-original continuous multi-dimensional data is transformed into a set of
-orthogonal components ordered according to the amount of variability
-that they describe. The \texttt{plot2D} method in
+Duve's principle (De Duve and Beaufay 1981) proteins that co-localise
+exhibit similar quantitation profiles across the gradient fractions
+employed. One approach that has been widely used to visualise and
+inspect high throughput mass spectrometry-based proteomics data is
+principal components analysis (PCA). PCA is one of many dimensionality
+reduction methods, that allow one to effectively summarise
+multi-dimensional data in to 2 or 3 dimensions to enable visualisation.
+Very generally, the original continuous multi-dimensional data is
+transformed into a set of orthogonal components ordered according to the
+amount of variability that they describe. The \texttt{plot2D} method in
 \emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} allows one
@@
 in-house user-defined sets of markers or from previous published studies
-{[}@Gatto:2014b{]}, which are continuosly updated and integrated.
+(Gatto et al. 2014), which are continuosly updated and integrated.
 
@@
 
-As demonstrated in {[}@hyper{]} and illustrated in the PCA plot above,
-the Golgi apparatus proteins display a dynamic pattern, noting sets of
-Golgi marker proteins that are distributed amongst other subcellular
-structures, an observation supported by microscopy. As such, we are
-going to reset the annotation of Golgi markers to unknown using the
-\texttt{fDataTounknown} function. It is often used to replace empty
+As demonstrated in (A. Christoforou et al. 2016) and illustrated in the
+PCA plot above, the Golgi apparatus proteins display a dynamic pattern,
+noting sets of Golgi marker proteins that are distributed amongst other
+subcellular structures, an observation supported by microscopy. As such,
+we are going to reset the annotation of Golgi markers to unknown using
+the \texttt{fDataTounknown} function. It is often used to replace empty
 strings (``'') or missing values in the markers definition to a common
@@
 
-In general, the Gene Ontology (GO) {[}@Ashburner:2000{]}, and in
+In general, the Gene Ontology (GO) (Ashburner et al. 2000), and in
 particular the cellular compartment (CC) namespace are a good starting
@@
 consequently, the reliability in protein sub-cellular assignments, we
-follow the advice in {[}@Trotter:2010{]} and combine replicated spatial
-proteomics experiments as described above. Indeed, Trotter et al. have
-shown a significant improvement in protein--organelle association upon
-direct combination of single experiments, in particular when these
-resolve different subcellular niches.
+follow the advice in (M. W. B. Trotter et al. 2010) and combine
+replicated spatial proteomics experiments as described above. Indeed,
+Trotter et al. have shown a significant improvement in
+protein--organelle association upon direct combination of single
+experiments, in particular when these resolve different subcellular
+niches.
 
@@
 instances can only be assigned to a class that exists in the training
-data {[}@Breckels:2013{]}. In such scenarios novelty detection can be
-useful to identify data-specific sub-cellular groupings such as
+data (L. M. Breckels et al. 2013). In such scenarios novelty detection
+can be useful to identify data-specific sub-cellular groupings such as
 organelles and protein complexes. The phenotype discovery (phenoDisco)
-algorithm {[}@Breckels:2013{]} is one such method and is available in
+algorithm (L. M. Breckels et al. 2013) is one such method and is
+available in
 \emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}. It is an
@@
 also useful for confirming the presence of known or postulated clusters
-in an unbiased fashion. For example, in {[}@hyper{]} the
-\texttt{phenoDisco} algorithm was used to confirm the data-specific
+in an unbiased fashion. For example, in (A. Christoforou et al. 2016)
+the \texttt{phenoDisco} algorithm was used to confirm the data-specific
 presence of the nucleus and nucleus sub-compartments. In the code chunk
@@
 input for the supervised learning aspect of the algorithm. As described
-in {[}@hyper{]} we define a set of markers to use as input for the
-analyses that cover well-known residents from three distinct organelle
-structures; the mitochondria, plasma membrane and ER, and from three
-well-known and abundant protein complexes; the proteasome and two
-ribosomal subunits, 40S and 60S. These input markers are stored in the
-\texttt{featureData} column of \texttt{hl} where
+in (A. Christoforou et al. 2016) we define a set of markers to use as
+input for the analyses that cover well-known residents from three
+distinct organelle structures; the mitochondria, plasma membrane and ER,
+and from three well-known and abundant protein complexes; the proteasome
+and two ribosomal subunits, 40S and 60S. These input markers are stored
+in the \texttt{featureData} column of \texttt{hl} where
 \texttt{fcol = "phenoDisco.Input"}. We can use the convenience accessor
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Sep 29 10:28:15 2016 
-## Added markers from  'pdres' marker vector. Thu Sep 29 10:28:17 2016 
+## Added markers from  'mrk' marker vector. Thu Sep 29 13:42:12 2016 
+## Added markers from  'pdres' marker vector. Thu Sep 29 13:42:13 2016 
 ##  MSnbase version: 1.99.2
@@
 proteomics to assign thousands of proteins to distinct sub-cellular
-niches {[}@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006;
-@Tan:2009{]}.
+niches (A. Christoforou et al. 2016; A. J. Groen et al. 2014; M. W. B.
+Trotter et al. 2010; Hall et al. 2009; Dunkley et al. 2006; Tan et al.
+2009).
 
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4356698               0.3028843               0.3825420 
+##               0.4361037               0.3025581               0.3909670 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6825826               0.5995243               0.4245311 
+##               0.6803283               0.6104122               0.4318449 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4034908               0.5882168               0.9501241 
+##               0.4172007               0.5869100               0.9490415 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951459               0.7096694               0.3149301 
+##               0.7940276               0.7083168               0.3151153 
 ##         Plasma membrane              Proteasome 
-##               0.7091667               0.4204304
+##               0.7224713               0.4169904
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     169                      86 
+##                      86                     169                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     479                      91 
+##                     297                     476                     103 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     125                     522 
+##                      24                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      38 
+##                     229                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     325                     159                    2057
+##                     315                     158                    2057
 \end{verbatim}
@@
 transfer learning algorithms using a \emph{k}-NN and SVM framework and
-applied them to the task of protein localisation prediction
-{[}@Breckels:2016{]}. In this section we will begin with explaining the
-concept of transfer learning and then show how to apply this in the
-frame of spatial proteomics and protein localisation prediction.
+applied them to the task of protein localisation prediction (Breckels et
+al. 2016). In this section we will begin with explaining the concept of
+transfer learning and then show how to apply this in the frame of
+spatial proteomics and protein localisation prediction.
 
@@
 with a higher generalisation accuracy than using standard supervised
-machine learning with a single source of information
-{[}@Breckels:2016{]}.
+machine learning with a single source of information (Breckels et al.
+2016).
 
@@
 well-separated under hyperLOPIT and LOPIT obtain a class score of 1
-(i.e.~use only primary data from transfer learning
-{[}@Breckels:2016{]}). Therefore, for the optimisation stage of the
-analyses we can already infer a subcellular class weight of 1 for these
-niches and only optimise over the remaining organelles. This can
-signifciantly cut down optimisation time as by removing these 4 classes
-from the optimisation (and not the classification) we only have
-\texttt{4\^{}10} class weight combinations to consider instead of
-\texttt{4\^{}14} combinations.
+(i.e.~use only primary data from transfer learning (Breckels et al.
+2016)). Therefore, for the optimisation stage of the analyses we can
+already infer a subcellular class weight of 1 for these niches and only
+optimise over the remaining organelles. This can signifciantly cut down
+optimisation time as by removing these 4 classes from the optimisation
+(and not the classification) we only have \texttt{4\^{}10} class weight
+combinations to consider instead of \texttt{4\^{}14} combinations.
 
@@
 \href{https://support.bioconductor.org/}{question} at hand.
+
+Ashburner, M., C. A. Ball, J. A. Blake, D. Botstein, H. Butler, J. M.
+Cherry, A. P. Davis, et al. 2000. ``Gene Ontology: tool for the
+Unification of Biology. the Gene Ontology Consortium.'' \emph{Nat Genet}
+25 (1): 25--29.
+doi:\href{http://dx.doi.org/10.1038/75556}{10.1038/75556}.
+
+Breckels, L. M., L. Gatto, A. Christoforou, A. J. Groen, K. S. Lilley,
+and M. W. Trotter. 2013. ``The Effect of Organelle Discovery Upon
+Sub-Cellular Protein Localisation.'' \emph{J Proteomics} 88 (Aug):
+129--40.
+doi:\href{http://dx.doi.org/10.1016/j.jprot.2013.02.019}{10.1016/j.jprot.2013.02.019}.
+
+Breckels, Lisa M., Sean B. Holden, David Wojnar, Claire M. Mulvey, Andy
+Christoforou, Arnoud Groen, Matthew W. B. Trotter, Oliver Kohlbacher,
+Kathryn S. Lilley, and Laurent Gatto. 2016. ``Learning from
+Heterogeneous Data Sources: An Application in Spatial Proteomics.''
+Edited by Thomas Dandekar. \emph{PLOS Computational Biology} 12 (5).
+Public Library of Science (PLoS): e1004920.
+doi:\href{http://dx.doi.org/10.1371/journal.pcbi.1004920}{10.1371/journal.pcbi.1004920}.
+\url{http://dx.doi.org/10.1371/journal.pcbi.1004920}.
+
+Christoforou, Andy, Claire M. Mulvey, Lisa M. Breckels, Penelope C.
+Hayward, Ekaterini Geladaki, Tracey Hurrell, Thomas Naake, et al. 2016.
+``A Draft Map of the Mouse Pluripotent Stem Cell Spatial Proteome.''
+\emph{Nat Commun} 7 (January): 9992.
+doi:\href{http://dx.doi.org/10.1038/ncomms9992}{10.1038/ncomms9992}.
+\url{http://dx.doi.org/10.1038/ncomms9992}.
+
+De Duve, C, and H Beaufay. 1981. ``A Short History of Tissue
+Fractionation.'' \emph{J Cell Biol} 91 (3 Pt 2): 293s--s.
+
+Dunkley, T. P. J., S. Hester, I. P. Shadforth, J. Runions, T. Weimar, S.
+L. Hanton, J. L. Griffin, et al. 2006. ``Mapping the Arabidopsis
+Organelle Proteome.'' \emph{Proc Natl Acad Sci USA} 103 (17): 6518--23.
+doi:\href{http://dx.doi.org/10.1073/pnas.0506958103}{10.1073/pnas.0506958103}.
+\url{http://dx.doi.org/10.1073/pnas.0506958103}.
+
+Gatto, L, L M Breckels, T Burger, D J Nightingale, A J Groen, C
+Campbell, N Nikolovski, et al. 2014. ``A Foundation for Reliable Spatial
+Proteomics Data Analysis.'' \emph{Mol Cell Proteomics} 13 (8): 1937--52.
+doi:\href{http://dx.doi.org/10.1074/mcp.M113.036350}{10.1074/mcp.M113.036350}.
+
+Gatto, L., and A. Christoforou. 2014. ``Using R and Bioconductor for
+Proteomics Data Analysis.'' \emph{Biochim Biophys Acta} 1844 (1 Pt A):
+42--51.
+doi:\href{http://dx.doi.org/10.1016/j.bbapap.2013.04.032}{10.1016/j.bbapap.2013.04.032}.
+
+Gatto, Laurent, Lisa M. Breckels, Thomas Naake, and Sebastian Gibb.
+2015. ``Visualization of Proteomics Data Using R and Bioconductor.''
+\emph{PROTEOMICS} 15 (8): 1375--89.
+doi:\href{http://dx.doi.org/10.1002/pmic.201400392}{10.1002/pmic.201400392}.
+\url{http://dx.doi.org/10.1002/pmic.201400392}.
+
+Groen, Arnoud J, Gloria Sancho-Andr{é}s, Lisa M Breckels, Laurent Gatto,
+Fernando Aniento, and Kathryn S Lilley. 2014. ``Identification of
+Trans-Golgi Network Proteins in Arabidopsis Thaliana Root Tissue.''
+\emph{J Proteome Res} 13 (2): 763--76.
+doi:\href{http://dx.doi.org/10.1021/pr4008464}{10.1021/pr4008464}.
+
+Hall, S. L., S. Hester, J. L. Griffin, K. S. Lilley, and A. P. Jackson.
+2009. ``The Organelle Proteome of the DT40 Lymphocyte Cell Line.''
+\emph{Mol Cell Proteomics} 8 (6): 1295--1305.
+doi:\href{http://dx.doi.org/10.1074/mcp.M800394-MCP200}{10.1074/mcp.M800394-MCP200}.
+\url{http://dx.doi.org/10.1074/mcp.M800394-MCP200}.
+
+Lazar, C, L Gatto, M Ferro, C Bruley, and T Burger. 2016. ``Accounting
+for the Multiple Natures of Missing Values in Label-Free Quantitative
+Proteomics Data Sets to Compare Imputation Strategies.'' \emph{J
+Proteome Res} 15 (4): 1116--25.
+doi:\href{http://dx.doi.org/10.1021/acs.jproteome.5b00981}{10.1021/acs.jproteome.5b00981}.
+
+McAlister, G C, D P Nusinow, M P Jedrychowski, M W{ü}hr, E L Huttlin, B
+K Erickson, R Rad, W Haas, and S P Gygi. 2014. ``MultiNotch MS3 Enables
+Accurate, Sensitive, and Multiplexed Detection of Differential
+Expression Across Cancer Cell Line Proteomes.'' \emph{Anal Chem} 86
+(14): 7150--8.
+doi:\href{http://dx.doi.org/10.1021/ac502040v}{10.1021/ac502040v}.
+
+Tan, D. J., H. Dvinge, A. Christoforou, P. Bertone, A. A. Martinez, and
+K. S. Lilley. 2009. ``Mapping Organelle Proteins and Protein Complexes
+in Drosophila Melanogaster.'' \emph{J Proteome Res} 8 (6): 2667--78.
+doi:\href{http://dx.doi.org/10.1021/pr800866n}{10.1021/pr800866n}.
+
+Trotter, M. W. B., P. G. Sadowski, T. P. J. Dunkley, A. J. Groen, and K.
+S. Lilley. 2010. ``Improved Sub-Cellular Resolution via Simultaneous
+Analysis of Organelle Proteomics Data Across Varied Experimental
+Conditions.'' \emph{Proteomics} 10 (23). WILEY-VCH Verlag: 4213--19.
+doi:\href{http://dx.doi.org/10.1002/pmic.201000359}{10.1002/pmic.201000359}.
+\url{http://dx.doi.org/10.1002/pmic.201000359}.
+
+Webb-Robertson, B J, H K Wiberg, M M Matzke, J N Brown, J Wang, J E
+McDermott, R D Smith, et al. 2015. ``Review, Evaluation, and Discussion
+of the Challenges of Missing Value Imputation for Mass
+Spectrometry-Based Label-Free Global Proteomics.'' \emph{J Proteome Res}
+14 (5): 1993--2001.
+doi:\href{http://dx.doi.org/10.1021/pr501138h}{10.1021/pr501138h}.

### 2016-09-29 bioc-workflow.Rmd
move to Rnw, update license to CC-BY

@@
----
-title: "A Bioconductor Workflow for Processing and Analysing Spatial Proteomics Data"
-author: "Lisa Breckels and Laurent Gatto"
-output: pdf_document
----
-
 ```{r env, echo=FALSE, warning=FALSE}
@@
 
-The main package to analyse protein localisation data is `r
-Biocpkg("pRoloc")`, which offers a set of dedicated functions for the
-analysis of such data. `r Biocpkg("pRoloc")` itself relies on `r
-Biocpkg("MSnbase")` to manipulate and process quantitative proteomics
-data. Many other packages are used by `r Biocpkg("pRoloc")` for
-clustering, classification and visualisation.
+The main package to analyse protein localisation data is 
+`r Biocpkg("pRoloc")`, which offers a set of dedicated functions for
+the analysis of such data. `r Biocpkg("pRoloc")` itself relies on 
+`r Biocpkg("MSnbase")` to manipulate and process quantitative
+proteomics data. Many other packages are used by `r Biocpkg("pRoloc")`
+for clustering, classification and visualisation.
 
@@
 
-```{r qcbx, fig.width = 12} par(mfrow = c(1, 2)) o <-
-order(hl$Iodixonal.Density) plotDist(hl[, o], pcol = "#00000010")
+```{r qcbx}
+par(mfrow = c(1, 2))
+o <- order(hl$Iodixonal.Density)
+plotDist(hl[, o], pcol = "#00000010")
 lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
-boxplot(exprs(hl[, o]), las = 2) ```
+boxplot(exprs(hl[, o]), las = 2)
+```
 
@@
 
-```{r plotmarkers, fig.width = 12}
+```{r plotmarkers}
 par(mfrow = c(1, 2))
@@
 
-```{r plot2Drep, fig.width = 12} 
+```{r plot2Drep} 
 par(mfrow = c(1, 2)) 
@@
 
-```{r plotPDres, fig.width=9}
+```{r plotPDres}
 ## Re-order the colours for the phenoDisco output
@@
 ```
-```{r visualiseOptHide, echo=FALSE, fig.width = 12}
+```{r visualiseOptHide, echo=FALSE}
 library('gridExtra')

### 2016-09-29 bioc-workflow.Rnw
move to Rnw, update license to CC-BY

--- 2016-09-29 bioc-workflow.tex
@@
-\section{Introduction}\label{introduction}
+<<env, echo=FALSE, warning=FALSE>>=
+library("BiocStyle")
+suppressPackageStartupMessages(library("knitr"))
+suppressPackageStartupMessages(library("gridExtra"))
+suppressPackageStartupMessages(library("MSnbase"))
+suppressPackageStartupMessages(library("pRoloc"))
+suppressPackageStartupMessages(library("pRolocdata"))
+cache <- TRUE
+@
+
+\section*{Introduction}
 
@@
 elaborate, expensive and time consuming experimental procedures and
-considerable effort is invested in the generation of such data. Multiple
-research groups have described a variety of approaches to establish high
-quality proteome-wide datasets. However, data analysis is as critical as
-data production for reliable and insightful biological interpretation.
-Here, we walk the reader through a typical pipeline for the analysis of
-such data using several Bioconductor packages for the R statistical
-programming environment.
+considerable effort is invested in the generation of such data.
+Multiple research groups have described a variety of approaches to
+establish high quality proteome-wide datasets. However, data analysis
+is as critical as data production for reliable and insightful
+biological interpretation. Here, we walk the reader through a typical
+pipeline for the analysis of such data using several Bioconductor
+packages for the R statistical programming environment.
 
 The main package to analyse protein localisation data is
-\texttt{r Biocpkg("pRoloc")}, which offers a set of dedicated functions
-for the analysis of such data.
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} itself
-relies on \texttt{r Biocpkg("MSnbase")} to manipulate and process
-quantitative proteomics data. Many other packages are used by
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} for
-clustering, classification and visualisation.
-
-In this workflow, we will describe how to prepare the spatial proteomics
-data starting from a spreadsheet containing quantitative mass
-spectrometry data. We will focus on a recent pluripotent mouse embryonic
-stem cells experiment {[}@hyper{]}. Additional annotated and
-pre-formatted datasets from various species are readily available in the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
-package.
+\Biocpkg{pRoloc}, which offers a set of dedicated functions for the
+analysis of such data. \Biocpkg{pRoloc} itself relies on
+\Biocpkg{MSnbase} to manipulate and process quantitative
+proteomics data. Many other packages are used by \Biocpkg{pRoloc}
+for clustering, classification and visualisation.
+
+In this workflow, we will describe how to prepare the spatial
+proteomics data starting from a spreadsheet containing quantitative
+mass spectrometry data. We will focus on a recent pluripotent mouse
+embryonic stem cells experiment \cite{hyper}. Additional annotated and
+pre-formatted datasets from various species are readily available in
+the \Biocexptpkg{pRolocdata} package.
 
 Installation of Bioconductor package is documented in details on the
-\href{http://bioconductor.org/install/\#install-bioconductor-packages}{Bioconductor
+\href{http://bioconductor.org/install/#install-bioconductor-packages}{Bioconductor
   installation help page}. This procedure is also applicable to any
 packages, from \href{https://cran.r-project.org/}{CRAN} as well as
-GitHub. Once a package has been installed, it needs to be loaded for its
-functionality to become available in the R session; this is done with
-the \texttt{library} function e.g. to load the
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} one would
-type \texttt{library("pRoloc")} after installation.
-
-If you have questions about this workflow in particular, or about other
-Bioconductor packages in general, they are best asked on the
+GitHub. Once a package has been installed, it needs to be loaded for
+its functionality to become available in the R session; this is done
+with the \texttt{library} function e.g.  to load the
+\Biocpkg{pRoloc} one would type \texttt{library("pRoloc")}
+after installation.
+
+If you have questions about this workflow in particular, or about
+other Bioconductor packages in general, they are best asked on the
 \href{https://support.bioconductor.org/}{Bioconductor support site}
@@
 proteomics, the readers are invited to read the
-\emph{\href{http://bioconductor.org/packages/RforProteomics}{RforProteomics}}
-package vignettes and associated papers {[}@Gatto:2014;@Gatto2015{]}.
-
-\section{Reading and processing spatial proteomics
-data}\label{reading-and-processing-spatial-proteomics-data}
-
-\subsection{The use-case: predicting sub-cellular localisation in
-pluripotent embryonic mouse stem
-cells}\label{the-use-case-predicting-sub-cellular-localisation-in-pluripotent-embryonic-mouse-stem-cells}
+\Biocexptpkg{"RforProteomics"} package vignettes and associated
+papers \cite{Gatto:2014,Gatto:2015}.
+
+\section*{Reading and processing spatial proteomics data}
+
+\subsection*{The use-case: predicting sub-cellular localisation in pluripotent embryonic mouse stem cells}
 
@@
 dataset from pluripotent mouse embryonic stem cells (E14TG2a)
-{[}@hyper{]}. The data was generated using hyperplexed LOPIT
-(hyperLOPIT), a state-of-the-art method relying on improved sub-cellular
-fractionation and more accurate quantitation, leading to more reliable
-classification of protein localisation across the whole sub-cellular
-space. The method uses an elaborate sub-cellular fractionation scheme,
-enabled by the use of Tandem Mass Tag (TMT) {[}Thompson:2003{]} 10-plex
-and application of the MS data acquisition technique named synchronous
-precursor selection MS\textsuperscript{3} (SPS-MS\textsuperscript{3})
-{[}@McAlister:2014{]}, for TMT quantification with high accuracy and
-precision. Three biological replicates were generated from the E14TG2a
-experiment, the first was to target low density fractions and the second
-and third were to emphasis separation of the denser organelles. The
-intersect of replicates 1 and 2 was treated as a 20-plex dataset for the
-analysis. As discussed in the manuscript {[}@hyper{]} as it has been
-shown that combining replicates across from different gradients can
-increase spatial resolution {[}@Trotter:2010{]}. The combination of
-replicates resulted in 5032 proteins common in both experiments.
+\cite{hyper}. The data was generated using hyperplexed LOPIT
+(hyperLOPIT), a state-of-the-art method relying on improved
+sub-cellular fractionation and more accurate quantitation, leading to
+more reliable classification of protein localisation across the whole
+sub-cellular space. The method uses an elaborate sub-cellular
+fractionation scheme, enabled by the use of Tandem Mass Tag (TMT)
+\cite{Thompson:2003} 10-plex and application of the MS data
+acquisition technique named synchronous precursor selection MS$^3$
+(SPS-MS$^3$) \cite{McAlister:2014}, for TMT quantification with high
+accuracy and precision. Three biological replicates were generated
+from the E14TG2a experiment, the first was to target low density
+fractions and the second and third were to emphasis separation of the
+denser organelles.  The intersect of replicates 1 and 2 was treated as
+a 20-plex dataset for the analysis.  As discussed in the manuscript
+\cite{hyper} as it has been shown that combining replicates across
+from different gradients can increase spatial resolution
+\cite{Trotter:2010}. The combination of replicates resulted in 5032
+proteins common in both experiments.
 
@@
 structured and annotated computational object from the 
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
-experiment package. In this workflow, we will start with a description
-of how to generate these ad hoc objects starting from an arbitrary
-spreadsheets, as produced by many popular third-party applications.
-
-While we focus here on a LOPIT-type dataset, these analyses are relevant
-for any quantitative spatial proteomics data, irrespective of the
-fractionation or quantitation (i.e.~labelled or label-free) methods.
-
-\subsection{The infrastructure:
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} and
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}} in
-Bioconductor}\label{the-infrastructure-proloc-and-msnbase-in-bioconductor}
-
-To make use of the full functionality of the
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} software
-one needs to import their data into R and prepare them as an
-\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for
-the efficient manipulation and processing of mass spectrometry and
-proteomics data in R. Figure 1 illustrates a simplified view of the
-\texttt{MSnSet} structure; there exists 3 key sub-parts (termed slots)
-to such a data object: (1) the \texttt{exprs} slot for storing the
-quantitation data, (2) the \texttt{fData} slot for storing the feature
-meta-data, and finally (3) the \texttt{pData} slot for storing the
-sample meta-data.
-
-\begin{figure}[htbp]
+\Biocexptpkg{pRolocdata} experiment package. In this workflow, we
+will start with a description of how to generate these ad hoc objects
+starting from an arbitrary spreadsheets, as produced by many popular
+third-party applications.
+
+While we focus here on a LOPIT-type dataset, these analyses are
+relevant for any quantitative spatial proteomics data, irrespective of
+the fractionation or quantitation (i.e. labelled or label-free)
+methods.
+
+\subsection*{The infrastructure: \texttt{pRoloc} and \texttt{MSnbase} packages}
+
+To make use of the full functionality of the \Biocpkg{pRoloc}
+software one needs to import their data into R and prepare them as an
+\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for the efficient
+manipulation and processing of mass spectrometry and proteomics data
+in R. Figure 1 illustrates a simplified view of the \texttt{MSnSet}
+structure; there exists 3 key sub-parts (termed slots) to such a data
+object: (1) the \texttt{exprs} slot for storing the quantitation data, (2)
+the \texttt{fData} slot for storing the feature meta-data, and finally (3)
+the \texttt{pData} slot for storing the sample meta-data.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/msnset.png}
-\caption{Simplified representation of the \texttt{MSnSet} data structure
-(reproduced with permission from the
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
+  \includegraphics[width=.45\textwidth]{./Figures/msnset.png}
+  \caption{Simplified representation of the \texttt{MSnSet} data
+    structure (reproduced with permission from the \Biocpkg{MSnbase}
     vignette)}
+  \label{fig:msnset}
 \end{figure}
 
-\subsection{Importing data}\label{importing-data}
+\subsection*{Importing data}
 
@@
 \texttt{MSnSet} instance. All methods are described in the
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
+\Biocpkg{MSnbase}
 \href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf}{input/output
   capabilities vignette}. One suggested simple method is to use the
-function \texttt{readMSnSet2} in
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}. The
-function takes a single spreadsheet as input and extracts the columns
+function \texttt{readMSnSet2} in \Biocpkg{MSnbase}. The function
+takes a single spreadsheet as input and extracts the columns
 containing the quantitation data, as identified by the argument
-\texttt{ecol}, to create the expression data, while the other columns in
-the spreadsheet are appended to the feature meta-data slot. By example,
-in the code chunk below we read in the \texttt{csv} spreadsheet
-containing the quantitation data from the intersect of replicates 1 and
-2 of the mouse map {[}@hyper{]}, using the \texttt{readMSnSet2}
-function. The data is as available online with the manuscript (see tab 2
-of the \texttt{xlsx} supplementary data set 1 in {[}@hyper{]}, which
-should be exported as a text-based spreadsheet). It is also available as
-a \texttt{csv} in the Bioconductor \texttt{r Biocexptpkg("pRolocdata")}
-data package.
-
-To use the \texttt{readMSnSet2} function, as a minimum one must specify
-the file path to the data and which columns of the spreadsheet contain
+\texttt{ecol}, to create the expression data, while the other columns
+in the spreadsheet are appended to the feature meta-data slot.  By
+example, in the code chunk below we read in the \texttt{csv}
+spreadsheet containing the quantitation data from the intersect of
+replicates 1 and 2 of the mouse map \cite{hyper}, using the
+\texttt{readMSnSet2} function. The data is as available online with
+the manuscript (see tab 2 of the \texttt{xlsx} supplementary data set
+1 in \cite{hyper}, which should be exported as a text-based
+spreadsheet). It is also available as a \texttt{csv} in the
+Bioconductor \Biocexptpkg{pRolocdata} data package.
+
+To use the \texttt{readMSnSet2} function, as a minimum one must specify the
+file path to the data and which columns of the spreadsheet contain
 quantitation data. The \texttt{getEcols} function exists to help users
-identify which columns of the spreadsheet contain the quantitation data.
-In the last line of the code chunk below, we print the file name (not
-the full path, which will vary from computer to computer).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"MSnbase"}\NormalTok{)}
-\NormalTok{extdatadir <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"pRolocdata"}\NormalTok{)}
-\NormalTok{csvfile <-}\StringTok{ }\KeywordTok{dir}\NormalTok{(extdatadir, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{,}
-          \DataTypeTok{pattern =} \StringTok{"hyperLOPIT-SIData-ms3-rep12-intersect.csv"}\NormalTok{)}
-\KeywordTok{basename}\NormalTok{(csvfile)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] "hyperLOPIT-SIData-ms3-rep12-intersect.csv.gz"
-\end{verbatim}
+identify which columns of the spreadsheet contain the
+quantitation data. In the last line of the code chunk below, we print
+the file name (not the full path, which will vary from computer to
+computer).
+
+<<getFilename, cache = cache>>=
+library("MSnbase")
+extdatadir <- system.file("extdata", package = "pRolocdata")
+csvfile <- dir(extdatadir, full.names = TRUE,
+          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
+basename(csvfile)
+@
 
 Note that the file is compressed (as indicated by the \texttt{gz}, for
-\texttt{gzip}, extension), and will be decompressed on-the-fly when read
-into R later on.
-
-The spreadsheet that was deposited by the authors contains two headers,
-with the second header containing information about where the
+\texttt{gzip}, extension), and will be decompressed on-the-fly when read into
+R later on.
+
+The spreadsheet that was deposited by the authors contains two
+headers, with the second header containing information about where the
 quantitation data is stored.
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/spreadsheet-screenshot.png}
+  \includegraphics[width=.65\textwidth]{./Figures/spreadsheet-screenshot.png}
   \caption{A screenshot of the data in the spreadsheet.}
+  \label{fig:spreadsheet}
 \end{figure}
@@
 We can display the names of the second header by calling the
-\texttt{getEcols} function with the argument \texttt{n = 2} (the default
-value is \texttt{n = 1}), to specify that we wish to display the column
-names of the second line.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getEcols}\NormalTok{(csvfile, }\DataTypeTok{split =} \StringTok{","}\NormalTok{, }\DataTypeTok{n =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##  [1] ""                                 
-##  [2] ""                                 
-##  [3] ""                                 
-##  [4] "Experiment 1"                     
-##  [5] "Experiment 2"                     
-##  [6] "Experiment 1"                     
-##  [7] "Experiment 2"                     
-##  [8] "126"                              
-##  [9] "127N"                             
-## [10] "127C"                             
-## [11] "128N"                             
-## [12] "128C"                             
-## [13] "129N"                             
-## [14] "129C"                             
-## [15] "130N"                             
-## [16] "130C"                             
-## [17] "131"                              
-## [18] "126"                              
-## [19] "127N"                             
-## [20] "127C"                             
-## [21] "128N"                             
-## [22] "128C"                             
-## [23] "129N"                             
-## [24] "129C"                             
-## [25] "130N"                             
-## [26] "130C"                             
-## [27] "131"                              
-## [28] "phenoDisco Input"                 
-## [29] "phenoDisco Output"                
-## [30] "Curated phenoDisco Output"        
-## [31] "SVM marker set"                   
-## [32] "SVM classification"               
-## [33] "SVM score"                        
-## [34] "SVM classification (top quartile)"
-## [35] "Final Localization Assignment"    
-## [36] "First localization evidence?"     
-## [37] "Curated Organelles"               
-## [38] "Cytoskeletal Components"          
-## [39] "Trafficking Proteins"             
-## [40] "Protein Complexes"                
-## [41] "Signaling Cascades"               
-## [42] "Oct4 Interactome"                 
-## [43] "Nanog Interactome"                
-## [44] "Sox2 Interactome"                 
-## [45] "Cell Surface Proteins"
-\end{verbatim}
+\texttt{getEcols} function with the argument \texttt{n = 2} (the default value is \texttt{n
+= 1}), to specify that we wish to display the column names of the
+second line.
+
+<<getEcols>>=
+getEcols(csvfile, split = ",", n = 2)
+@
 
 It is now easy for one to identify that the quantitation data,
-corresponding to the 10 TMT isobaric tags, is located in columns 8 to
-27. We now have the two mandatory arguments to \texttt{readMSnSet2},
+corresponding to the 10 TMT isobaric tags, is located in columns 8
+to 27. We now have the two mandatory arguments to \texttt{readMSnSet2},
 namely the file name (stored in the \texttt{csvfile} variable) and the
 quantitation column indices. In addition to these, it is also possible
-to pass the optional argument \texttt{fnames} to indicate which column
-to use as the labels by which to identify each protein in the sample.
-Here, we use \texttt{fnames = 1} to use the UniProt identifiers
-contained in the first (unnamed) column of the spreadsheet. We also need
-to specify to skip the first line of the file (for the same reason that
-we used \texttt{n = 2} in \texttt{getEcols} above) to read the
-\texttt{csv} data and convert it to an \texttt{MSnSet} object, named
-\texttt{hl} (for hyperLOPIT).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{readMSnSet2}\NormalTok{(csvfile, }\DataTypeTok{ecol =} \DecValTok{8}\NormalTok{:}\DecValTok{27}\NormalTok{, }\DataTypeTok{fnames =} \DecValTok{1}\NormalTok{, }\DataTypeTok{skip =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Below, we display a short summary of the data. The data contains 5032
-proteins/features common across the 2 biological replicates for the
-respective 2 x 10-plex reporter tags (20 columns/samples), along with
-associated feature meta-data such as protein markers, protein
-description, number of quantified peptides etc (see below).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 5032 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData: none
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
-##   fvarLabels: X X.1 ... Cell.Surface.Proteins (25 total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-## Annotation:  
-## - - - Processing information - - -
-##  MSnbase version: 1.99.2
-\end{verbatim}
-
-Below, we examine the quantitative information for first 5 proteins. It
-is also possible to access specific rows and columns by naming the
+to pass the optional argument \texttt{fnames} to indicate which column to use
+as the labels by which to identify each protein in the sample. Here,
+we use \texttt{fnames = 1} to use the UniProt identifiers contained in the
+first (unnamed) column of the spreadsheet. We also need to specify to
+skip the first line of the file (for the same reason that we used 
+\texttt{n = 2} in \texttt{getEcols} above) to read the \texttt{csv} data and convert it to an
+\texttt{MSnSet} object, named \texttt{hl} (for hyperLOPIT).
+
+<<readMSnSet2>>=
+hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
+@
+
+Below, we display a short summary of the data. The data contains 
+\Sexpr{nrow(hl)} proteins/features common across the 2 biological replicates
+for the respective 2 x 10-plex reporter tags (\Sexpr{ncol(hl)}
+columns/samples), along with associated feature meta-data such as
+protein markers, protein description, number of quantified peptides
+etc (see below).
+
+<<showhl>>=
+hl
+@
+
+Below, we examine the quantitative information for first 5 proteins.
+It is also possible to access specific rows and columns by naming the
 proteins and TMT tag channels of interest.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{exprs}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{5}\NormalTok{, ]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##           X126 X127N X127C X128N X128C X129N X129C X130N X130C  X131
-## Q9JHU4   0.028 0.034 0.024 0.014 0.026 0.045 0.107 0.341 0.059 0.321
-## Q9QXS1-3 0.039 0.134 0.095 0.053 0.084 0.121 0.107 0.128 0.122 0.117
-## Q9ERU9   0.021 0.013 0.014 0.009 0.024 0.054 0.116 0.257 0.209 0.284
-## P26039   0.120 0.255 0.148 0.091 0.135 0.095 0.041 0.057 0.014 0.043
-## Q8BTM8   0.055 0.139 0.078 0.050 0.077 0.098 0.093 0.171 0.079 0.160
-##          X126.1 X127N.1 X127C.1 X128N.1 X128C.1 X129N.1 X129C.1 X130N.1
-## Q9JHU4    0.037   0.064   0.058   0.059   0.067   0.078   0.140   0.208
-## Q9QXS1-3  0.033   0.073   0.074   0.062   0.081   0.142   0.190   0.069
-## Q9ERU9    0.026   0.017   0.023   0.029   0.039   0.071   0.105   0.171
-## P26039    0.111   0.181   0.141   0.144   0.152   0.119   0.075   0.028
-## Q8BTM8    0.062   0.108   0.091   0.086   0.099   0.111   0.117   0.095
-##          X130C.1 X131.1
-## Q9JHU4     0.141  0.147
-## Q9QXS1-3   0.151  0.125
-## Q9ERU9     0.304  0.215
-## P26039     0.017  0.033
-## Q8BTM8     0.144  0.087
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{exprs}\NormalTok{(hl)[}\KeywordTok{c}\NormalTok{(}\StringTok{"Q9ERU9"}\NormalTok{, }\StringTok{"Q9Z2R6"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"X126"}\NormalTok{, }\StringTok{"X131.1"}\NormalTok{)]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##         X126 X131.1
-## Q9ERU9 0.021  0.215
-## Q9Z2R6 0.563  0.000
-\end{verbatim}
+<<assesshl>>=
+exprs(hl)[1:5, ]
+exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
+@
 
 The feature meta-data is stored in the \texttt{fData} slot and can be
-accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2},
-automatically, everything that is not defined as quantitation data by
-\texttt{ecol} or the feature names by \texttt{fnames} is deposited to
-the \texttt{fData} slot.
-
-We see the \texttt{fData} contains 25 columns describing information
-such as the number of peptides, associated markers, machine learning
-results etc. To identify the feature variable names we can use the
-function \texttt{fvarLabels}. We see that the first 6 feature variable
-names contain non-discriminatory label names, so we relabel them to help
-us identify what feature data information is stored in the associated
-columns.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##  [1] "X"                                
-##  [2] "X.1"                              
-##  [3] "X.2"                              
-##  [4] "Experiment.1"                     
-##  [5] "Experiment.2"                     
-##  [6] "Experiment.1.1"                   
-##  [7] "Experiment.2.1"                   
-##  [8] "phenoDisco.Input"                 
-##  [9] "phenoDisco.Output"                
-## [10] "Curated.phenoDisco.Output"        
-## [11] "SVM.marker.set"                   
-## [12] "SVM.classification"               
-## [13] "SVM.score"                        
-## [14] "SVM.classification..top.quartile."
-## [15] "Final.Localization.Assignment"    
-## [16] "First.localization.evidence."     
-## [17] "Curated.Organelles"               
-## [18] "Cytoskeletal.Components"          
-## [19] "Trafficking.Proteins"             
-## [20] "Protein.Complexes"                
-## [21] "Signaling.Cascades"               
-## [22] "Oct4.Interactome"                 
-## [23] "Nanog.Interactome"                
-## [24] "Sox2.Interactome"                 
-## [25] "Cell.Surface.Proteins"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{3}\NormalTok{] <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"uniprot.accession"}\NormalTok{, }\StringTok{"uniprot.id"}\NormalTok{, }\StringTok{"description"}\NormalTok{)}
-\KeywordTok{fvarLabels}\NormalTok{(hl)[}\DecValTok{4}\NormalTok{:}\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\StringTok{"peptides.expt"}\NormalTok{, }\DecValTok{1}\NormalTok{:}\DecValTok{3}\NormalTok{)}
-\KeywordTok{fData}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{4}\NormalTok{, }\DecValTok{1}\NormalTok{:}\DecValTok{6}\NormalTok{]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##          uniprot.accession  uniprot.id
-## Q9JHU4              Q9JHU4 DYHC1_MOUSE
-## Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE
-## Q9ERU9              Q9ERU9  RBP2_MOUSE
-## P26039              P26039  TLN1_MOUSE
-##                                                                                                                                 description
-## Q9JHU4                                              Cytoplasmic dynein 1 heavy chain 1 OS=Mus musculus GN=Dync1h1 PE=1 SV=2 - [DYHC1_MOUSE]
-## Q9QXS1-3 Isoform PLEC-1 of Plectin OS=Mus musculus GN=Plec - [PLEC_MOUSE]|Isoform PLEC-1A of Plectin OS=Mus musculus GN=Plec - [PLEC_MOUSE]
-## Q9ERU9                                                     E3 SUMO-protein ligase RanBP2 OS=Mus musculus GN=Ranbp2 PE=1 SV=2 - [RBP2_MOUSE]
-## P26039                                                                             Talin-1 OS=Mus musculus GN=Tln1 PE=1 SV=2 - [TLN1_MOUSE]
-##          peptides.expt1 peptides.expt2 peptides.expt3
-## Q9JHU4              175            166            322
-## Q9QXS1-3            123            150            174
-## Q9ERU9              101             90            181
-## P26039              101             94            167
-\end{verbatim}
-
-Note that when using the simple \texttt{readMSnSet2} procedure, the
-\texttt{pData} slot which is used to store information about the
-samples/channels is kept empty. It is advised to annotate the channels
-as well. Below, we annotate the replicate from which the profiles
+accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2}, automatically,
+everything that is not defined as quantitation data by \texttt{ecol} or the
+feature names by \texttt{fnames} is deposited to the \texttt{fData} slot.
+
+We see the \texttt{fData} contains 25 columns describing information such as
+the number of peptides, associated markers, machine learning results
+etc. To identify the feature variable names we can use the function
+\texttt{fvarLabels}. We see that the first 6 feature variable names contain
+non-discriminatory label names, so we relabel them to help us identify
+what feature data information is stored in the associated columns.
+
+<<removefData>>=
+fvarLabels(hl)
+fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
+fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
+fData(hl)[1:4, 1:6]
+@
+
+Note that when using the simple \texttt{readMSnSet2} procedure, the \texttt{pData}
+slot which is used to store information about the samples/channels
+is kept empty. It is advised to annotate the channels as
+well. Below, we annotate the replicate from which the profiles
 originate and the TMT tag (extracted from the sample/channel names).
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{pData}\NormalTok{(hl)$Replicate <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{each =} \DecValTok{10}\NormalTok{)}
-\KeywordTok{pData}\NormalTok{(hl)$Tag <-}\StringTok{ }\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{.1$"}\NormalTok{, }\StringTok{""}\NormalTok{, }\KeywordTok{sub}\NormalTok{(}\StringTok{"^X"}\NormalTok{, }\StringTok{""}\NormalTok{, }\KeywordTok{sampleNames}\NormalTok{(hl)))}
-\KeywordTok{pData}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##         Replicate  Tag
-## X126            1  126
-## X127N           1 127N
-## X127C           1 127C
-## X128N           1 128N
-## X128C           1 128C
-## X129N           1 129N
-## X129C           1 129C
-## X130N           1 130N
-## X130C           1 130C
-## X131            1  131
-## X126.1          2  126
-## X127N.1         2 127N
-## X127C.1         2 127C
-## X128N.1         2 128N
-## X128C.1         2 128C
-## X129N.1         2 129N
-## X129C.1         2 129C
-## X130N.1         2 130N
-## X130C.1         2 130C
-## X131.1          2  131
-\end{verbatim}
-
-Throughout this workflow we refer to the different columns that are
-found in the \texttt{exprs} (expression data) slot as channels (short
-for TMT channels). In the frame of LOPIT and hyperLOPIT these channels
-consititue the relative abundance of each protein (along the rows) in
-the channel of interest. Each TMT channel originates from fractions
-collected from the density gradient, or a set of pooled fractions or may
-be a sample originating from an alternative preparation e.g.~such as
-from the chromatin enrichment performed in Christoforou et al
-{[}@hyper{]}. Information about which gradient fractions were used for
-which tag should also be stored in the sample meta-data \texttt{pData}
-slot.
+<<pdata>>=
+pData(hl)$Replicate <- rep(1:2, each = 10)
+pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
+pData(hl)
+@
+
+Throughout this workflow we refer to the different columns that
+are found in the \texttt{exprs} (expression data) slot as channels (short for TMT channels).
+In the frame of LOPIT and hyperLOPIT these channels consititue the
+relative abundance of each protein (along the rows) in the channel of
+interest. Each TMT channel originates from fractions collected from
+the density gradient, or a set of pooled fractions or may be a sample
+originating from an alternative preparation e.g. such as from the
+chromatin enrichment performed in Christoforou et al \cite{hyper}.
+Information about which gradient fractions were used for which tag
+should also be stored in the sample meta-data \texttt{pData} slot.
 
 The sample meta-data that is distributed with the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
-package for Christoforou's hyperLOPIT experiment and (as above) the
-quantitation data file, are located in the \texttt{extdata} in the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
+\Biocexptpkg{pRolocdata} package for Christoforou's
+hyperLOPIT experiment and (as above) the quantitation data file, are
+located in the \texttt{extdata} in the \Biocexptpkg{pRolocdata}
 package on the hard drive.
 
-In the code chunk below we again use the \texttt{dir} function to locate
-the filepath to the meta-data \texttt{csv} file and then read it into R
-using \texttt{read.csv}. We then append the meta-data to the
-\texttt{pData} slot. Information about the gradient fractions used and
-the associated subcellular fraction densities in \% w/v Iodixanol for
+In the code chunk below we again use the
+\texttt{dir} function to locate the filepath to the meta-data \texttt{csv} file and
+then read it into R using \texttt{read.csv}. We then append the meta-data to
+the \texttt{pData} slot.  Information about the gradient fractions used and
+the associated subcellular fraction densities in % w/v Iodixanol for
 each replicate are stored here.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{expinfo <-}\StringTok{ }\KeywordTok{dir}\NormalTok{(extdatadir, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{,}
-               \DataTypeTok{pattern =} \StringTok{"hyperLOPIT-SIData-fraction-info.csv"}\NormalTok{)}
-
-\NormalTok{fracinfo <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(expinfo, }\DataTypeTok{row.names=}\DecValTok{1}\NormalTok{, }\DataTypeTok{skip =} \DecValTok{2}\NormalTok{, }
-                     \DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
-
-\KeywordTok{pData}\NormalTok{(hl)$Gradient.Fraction <-}\StringTok{ }\KeywordTok{c}\NormalTok{(fracinfo[, }\DecValTok{1}\NormalTok{], fracinfo[, }\DecValTok{2}\NormalTok{])}
-\KeywordTok{pData}\NormalTok{(hl)$Iodixonal.Density <-}\StringTok{ }\KeywordTok{c}\NormalTok{(fracinfo[, }\DecValTok{4}\NormalTok{], fracinfo[, }\DecValTok{5}\NormalTok{])}
-\KeywordTok{pData}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##         Replicate  Tag Gradient.Fraction Iodixonal.Density
-## X126            1  126           Cytosol               0.0
-## X127N           1 127N   1 to 6 (pooled)               6.0
-## X127C           1 127C   8 to 9 (pooled)              11.0
-## X128N           1 128N 10 to 11 (pooled)              13.3
-## X128C           1 128C                12              14.6
-## X129N           1 129N                14              17.4
-## X129C           1 129C                16              20.1
-## X130N           1 130N                18              26.8
-## X130C           1 130C         Chromatin                NA
-## X131            1  131                19              34.5
-## X126.1          2  126           Cytosol               0.0
-## X127N.1         2 127N   1 to 6 (pooled)               5.2
-## X127C.1         2 127C   7 to 9 (pooled)              10.0
-## X128N.1         2 128N 10 to 11 (pooled)              12.5
-## X128C.1         2 128C                12              14.0
-## X129N.1         2 129N 14 to 15 (pooled)              17.3
-## X129C.1         2 129C                17              20.9
-## X130N.1         2 130N 18 to 19 (pooled)              24.7
-## X130C.1         2 130C         Chromatin                NA
-## X131.1          2  131                20              31.9
-\end{verbatim}
-
-\subsection{Data processing}\label{data-processing}
-
-\subsubsection{Normalisation}\label{normalisation}
-
-There are two aspects related to data normalisation that are relevant to
-spatial proteomics data processing. The first one focuses on reducing
-purely technical variation between channels without affecting biological
-variability (i.e.~the shape of the quantitative profiles). This
-normalisation will depend on the underlying quantitative technology and
-the experimental design, and will not be addressed in this workflow. The
-second aspect, and more specific to spatial proteomics data, is scaling
-all the organelle-specific profiles into a same intensity interval
-(typically 0 and 1) by, for example, diving each intensity by the sum of
-the intensities for that quantitative feature. This is not necessary in
-this example as the intensities for each replicate have already been
-re-scaled to 1 in Proteome Discoverer. However, if one wanted to do this
-they would execute the \texttt{normalise} function as demonstrated in
-the below code chunk.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{normalise}\NormalTok{(hl, }\DataTypeTok{method =} \StringTok{"sum"}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
+<<morepdata>>=
+expinfo <- dir(extdatadir, full.names = TRUE,
+               pattern = "hyperLOPIT-SIData-fraction-info.csv")
+
+fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
+                     header = FALSE, stringsAsFactors = FALSE)
+
+pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
+pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
+pData(hl)
+@
+
+\subsection*{Data processing}
+
+\subsubsection*{Normalisation}
+
+There are two aspects related to data normalisation that are relevant
+to spatial proteomics data processing. The first one focuses on
+reducing purely technical variation between channels without
+affecting biological variability (i.e. the shape of the quantitative
+profiles). This normalisation will depend on the underlying
+quantitative technology and the experimental design, and will not be
+addressed in this workflow. The second aspect, and more specific to
+spatial proteomics data, is scaling all the organelle-specific
+profiles into a same intensity interval (typically 0 and 1) by, for
+example, diving each intensity by the sum of the intensities for that
+quantitative feature. This is not necessary in this example as the
+intensities for each replicate have already been re-scaled to 1 in
+Proteome Discoverer. However, if one wanted to do this they would
+execute the \texttt{normalise} function as demonstrated in the below code
+chunk.
+
+<<normsum, eval = FALSE>>= 
+hl <- normalise(hl, method = "sum") 
+@
 
@@
 
-The same \texttt{normalise} function (or \texttt{normalize}, both
-spellings are supported) can also be applied in the first case described
-above. Different normalisation methods such as mean or median scaling,
-variance stabilisation or quantile normalisation, to cite a few, can be
-applied to accomodation different needs.
+The same \texttt{normalise} function (or \texttt{normalize}, both spellings are
+supported) can also be applied in the first case described above.
+Different normalisation methods such as mean or median scaling,
+variance stabilisation or quantile normalisation, to cite a few, can
+be applied to accomodation different needs.
 
 As previously mentioned, before combination, the two replicates in the
-\texttt{hl} data that we read into R were separately normalised by sum
-(i.e.~to 1) across the 10 channels for each replicate respectively. We
-can verify this by summing each rows for each replicate:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{summary}\NormalTok{(}\KeywordTok{rowSums}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hl[, hl$replicate ==}\StringTok{ }\DecValTok{1}\NormalTok{])))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-##       0       0       0       0       0       0
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{summary}\NormalTok{(}\KeywordTok{rowSums}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hl[, hl$replicate ==}\StringTok{ }\DecValTok{2}\NormalTok{])))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-##       0       0       0       0       0       0
-\end{verbatim}
+\texttt{hl} data that we read into R were separately normalised by sum (i.e.
+to 1) across the 10 channels for each replicate respectively. We can
+verify this by summing each rows for each replicate:
+
+<<normcheck>>=
+summary(rowSums(exprs(hl[, hl$replicate == 1])))
+summary(rowSums(exprs(hl[, hl$replicate == 2])))
+@
 
 We see that some features do not add up exactly to 1 due to rounding
-errors after exporting to intermediate files. These small deviations do
-not bear any consequences here.
-
-\subsubsection{Combining acquisitions}\label{combining-acquisitions}
-
-The spreadsheet that was used to create the \texttt{hl} MSnSet included
-the two replicates within one .csv file. We also provide individual
-replicates in the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
+errors after exporting to intermediate files. These small deviations
+do not bear any consequences here.
+
+\subsubsection*{Combining acquisitions}
+
+The spreadsheet that was used to create the \texttt{hl} MSnSet
+included the two replicates within one .csv file.  We also provide
+individual replicates in the \Biocexptpkg{pRolocdata}
 package. Below, we show how to combine \texttt{MSnSet} objects and,
 subsequently, how to filter and handle missing values. We start by
-loading the \texttt{r Biocexptpkg("pRolocdata")} package and the
+loading the \Biocexptpkg{pRolocdata} package and the
 equivalent replicates using the \texttt{data} function.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocdata"}\NormalTok{)}
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r1)}
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
+<<reps4combine>>=
+library("pRolocdata")
+data(hyperLOPIT2015ms3r1)
+data(hyperLOPIT2015ms3r2)
+@
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{pRolocdata}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-will list the 54 datasets that are available in
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}.
+<<pkgdata, eval=FALSE>>=
+pRolocdata()
+@
+
+will list the \Sexpr{nrow(pRolocdata()$results)} datasets that are
+available in \Biocexptpkg{pRolocdata}.
+
+%% $
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{identical}\NormalTok{(}\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r1), }\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r2))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] TRUE
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hyperLOPIT2015ms3r1 <-}\StringTok{ }\KeywordTok{updateSampleNames}\NormalTok{(hyperLOPIT2015ms3r1, }\DecValTok{1}\NormalTok{)}
-\NormalTok{hyperLOPIT2015ms3r2 <-}\StringTok{ }\KeywordTok{updateSampleNames}\NormalTok{(hyperLOPIT2015ms3r2, }\DecValTok{2}\NormalTok{)}
-\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r1)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##  [1] "X126.1"  "X127N.1" "X127C.1" "X128N.1" "X128C.1" "X129N.1" "X129C.1"
-##  [8] "X130N.1" "X130C.1" "X131.1"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##  [1] "X126.2"  "X127N.2" "X127C.2" "X128N.2" "X128C.2" "X129N.2" "X129C.2"
-##  [8] "X130N.2" "X130C.2" "X131.2"
-\end{verbatim}
-
-In addition, to matching names, the content of the feature metadata for
-identical feature annotations must match exactly across the data to be
-combined. In particular for these data, we expect the same proteins in
-each replicates to be annotated with the same UniProt entry names and
-descriptions, but not with the same coverage of number of peptides or
-peptide-spectrum matches (PSMs).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] "EntryName"          "ProteinDescription" "Peptides"          
-## [4] "PSMs"               "ProteinCoverage"    "markers"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] "EntryName"          "ProteinDescription" "Peptides"          
-## [4] "PSMs"               "ProteinCoverage"    "markers"
-\end{verbatim}
+<<samplenames>>=
+identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
+hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
+hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
+sampleNames(hyperLOPIT2015ms3r1)
+sampleNames(hyperLOPIT2015ms3r2)
+@
+
+In addition, to matching names, the content of the feature metadata
+for identical feature annotations must match exactly across the data
+to be combined. In particular for these data, we expect the same
+proteins in each replicates to be annotated with the same UniProt
+entry names and descriptions, but not with the same coverage of number
+of peptides or peptide-spectrum matches (PSMs).
+
+<<fvarnames>>=
+fvarLabels(hyperLOPIT2015ms3r1)
+fvarLabels(hyperLOPIT2015ms3r2)
+@
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{], }\DecValTok{1}\NormalTok{)}
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{], }\DecValTok{2}\NormalTok{)}
-\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r1) <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{1}\NormalTok{:}\DecValTok{5}\NormalTok{] }
-\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r2) <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{]}
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] "EntryName"          "ProteinDescription" "Peptides1"         
-## [4] "PSMs1"              "ProteinCoverage1"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## [1] "Peptides2"        "PSMs2"            "ProteinCoverage2"
-\end{verbatim}
+<<fvarnames2>>=
+fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
+fvarLabels(hyperLOPIT2015ms3r2)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r2)[3:5], 2)
+fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5] 
+fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
+fvarLabels(hyperLOPIT2015ms3r1)
+fvarLabels(hyperLOPIT2015ms3r2)
+@
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{combined <-}\StringTok{ }\KeywordTok{combine}\NormalTok{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
-\NormalTok{combined}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 6725 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData
-##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
-##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
-##   varMetadata: labelDescription
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
-##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
-##     total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-## Annotation:  
-## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
-##  MSnbase version: 1.21.7
-\end{verbatim}
+<<combine, cache = cache>>=
+combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
+combined
+@
 
 More details above combining data are given in the dedicated
-\emph{Combining MSnSet instances} section of the
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
+*Combining MSnSet instances* section of the \Biocpkg{MSnbase}
 \href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
@@
 
-\subsubsection{Missing data}\label{missing-data}
+\subsubsection*{Missing data}
 
@@
 and should be addressed independently of this workflow
-{[}@Webb-Robertson:2015;@Lazar:2016{]}. In {[}@Gatto:2014b{]}, we have
+\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
 described how a high content in missing values in spatial proteomics
@@
 sub-cellular resolution. Missing data can be imputated using 
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}'s
-\texttt{impute} function. The method underlying the imputation method is
-then determined by a \texttt{methods} parameter. In our particular case,
-missing values are indicative of protein groups that were not profiles
-in both replicates.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{image2}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(combined), }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"black"}\NormalTok{, }\StringTok{"white"}\NormalTok{),}
-       \DataTypeTok{main =} \StringTok{"Missing values (white cells) after combining replicates"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying the
+imputation method is then determined by a \texttt{methods} parameter. In our
+particular case, missing values are indicative of protein groups that
+were not profiles in both replicates.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/namap-1.png}
-\caption{plot of chunk namap}
+<<namap, out.width="\\textwidth">>=
+image2(is.na(combined), col = c("black", "white"),
+       main = "Missing values (white cells) after combining replicates")
+@  
+  \caption{Heatmap of missing values}
+  \label{fig:namap}
 \end{figure}
@@
 We prefer to remove proteins that were not assayed in both of our two
-replicated experiments. This is done with the \texttt{filterNA} function
-that removes features that contain more than a certain proportion
-(default is 0) missing values.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{combined <-}\StringTok{ }\KeywordTok{filterNA}\NormalTok{(combined)}
-\NormalTok{combined}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 5032 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData
-##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
-##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
-##   varMetadata: labelDescription
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
-##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
-##     total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-## Annotation:  
-## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
-## Subset [6725,20][5032,20] Thu Sep 29 10:28:15 2016 
-## Removed features with more than 0 NAs: Thu Sep 29 10:28:15 2016 
-## Dropped featureData's levels Thu Sep 29 10:28:15 2016 
-##  MSnbase version: 1.21.7
-\end{verbatim}
+replicated experiments. This is done with the \texttt{filterNA} function that
+removes features that contain more than a certain proportion (default
+is 0) missing values.
+
+<<filterNA>>=
+combined <- filterNA(combined)
+combined
+@
 
@@
 protein-sub-cellular niche membership over the output of several
-experiments (see section \emph{Supervised machine learning} for the
+experiments (see section *Supervised machine learning* for the
 description of sub-cellular assignments).
 
-\section{Quality Control}\label{quality-control}
-
-Data quality is routinely examined through visualisation to verify that
-sub-cellular niches have been separated along the gradient. Based on De
-Duve's principle {[}@DeDuve:1981{]} proteins that co-localise exhibit
-similar quantitation profiles across the gradient fractions employed.
-One approach that has been widely used to visualise and inspect high
-throughput mass spectrometry-based proteomics data is principal
-components analysis (PCA). PCA is one of many dimensionality reduction
-methods, that allow one to effectively summarise multi-dimensional data
-in to 2 or 3 dimensions to enable visualisation. Very generally, the
-original continuous multi-dimensional data is transformed into a set of
-orthogonal components ordered according to the amount of variability
-that they describe. The \texttt{plot2D} method in
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} allows one
-to plot the principal components (PCs) of a dataset against one another,
+\section*{Quality Control}
+
+Data quality is routinely examined through visualisation to verify
+that sub-cellular niches have been separated along the gradient. Based
+on De Duve's principle \cite{DeDuve:1981} proteins that co-localise
+exhibit similar quantitation profiles across the gradient fractions
+employed. One approach that has been widely used to visualise and
+inspect high throughput mass spectrometry-based proteomics data is
+principal components analysis (PCA). PCA is one of many dimensionality
+reduction methods, that allow one to effectively summarise
+multi-dimensional data in to 2 or 3 dimensions to enable
+visualisation. Very generally, the original continuous
+multi-dimensional data is transformed into a set of orthogonal
+components ordered according to the amount of variability that they
+describe. The \texttt{plot2D} method in \Biocpkg{pRoloc} allows one to
+plot the principal components (PCs) of a dataset against one another,
 by default the first two components are plotted on the x- and y-axis,
-respectively (the \texttt{dims} argument can be used to plot other PCs).
-If distinct clusters are observed, we assume that there is organellar
+respectively (the \texttt{dims} argument can be used to plot other PCs). If
+distinct clusters are observed, we assume that there is organellar
 separation present in the data. Although, representing the
@@
 In the code chunk below we produce a PCA plot of the mouse stem cell
-dataset. One point on the plot represents one protein. We can indeed see
-several distinct protein clusters. We specify \texttt{fcol = NULL},
-which means not to consider any feature variable to annotate the
-features (proteins) with colours. We will see later how to use this to
-annotate the PCA plot with prior information about sub-cellular
-localisation.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRoloc"}\NormalTok{)}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+dataset. One point on the plot represents one protein. We can indeed
+see several distinct protein clusters. We specify \texttt{fcol = NULL}, which
+means not to consider any feature variable to annotate the features
+(proteins) with colours. We will see later how to use this to annotate
+the PCA plot with prior information about sub-cellular localisation.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/qcplot-1.png}
-\caption{plot of chunk qcplot}
+<<qcplot, message=FALSE, warning=FALSE, out.width="\\textwidth">>=
+library("pRoloc")
+plot2D(hl, fcol = NULL, col = "black")
+@  
+  \caption{PCA plot of the mouse stem cell data \texttt{hl}.}
+  \label{fig:pcahl}
 \end{figure}
 
-In the first instance we advise one to visualise their data without any
-annotation (i.e.~with \texttt{fcol = NULL}), before proceeding with data
+In the first instance we advise one to visualise their data without
+any annotation (i.e. with \texttt{fcol = NULL}), before proceeding with data
 annotation. The identification of well resolved clusters in the data,
-constitutes an unbiased assessment of the data structure, demonstrating
-the successful separation of sub-cellular clusters.
-
-\textbf{TODO} There are other dimensionality reduction methods available
-in the \texttt{plot2D} function, which can be parametrised with the
+constitutes an unbiased assessment of the data structure,
+demonstrating the successful separation of sub-cellular clusters.
+
+**TODO** There are other dimensionality reduction methods available in
+the \texttt{plot2D} function, which can be parametrised with the
 \texttt{method} argument.  Advantages, disadvantages? See
-\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}{this}
-very interesting gist with intriguing tSNE examples.
+\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}
+very interesting gist with intriguing tSNE examples. (This might be
+better to come after markers.)
 
@@
 gradient to identify channels displaying particularly low yield. This
-can be done using the \texttt{plotDist} and \texttt{boxplot} functions,
-that plot the protein profiles occupancy along the gradient (we also
-display the mean channel intensities) and a \texttt{boxplot} of the
-column intensities. In the two plots displayed below, we re-order the
-TMT channles to pair corresponding channels in the two replicates
-(rather than ordering the channels by replicate).
-
-\texttt{\{r qcbx, fig.width = 12\} par(mfrow = c(1, 2)) o \textless{}- order(hl\$Iodixonal.Density) plotDist(hl{[}, o{]}, pcol = "\#00000010") lines(colMeans(exprs(hl{[}, o{]})), col = "red", type = "b") boxplot(exprs(hl{[}, o{]}), las = 2)}
-
-\section{Markers}\label{markers}
+can be done using the \texttt{plotDist} and \texttt{boxplot} functions, that plot
+the protein profiles occupancy along the gradient (we also display the
+mean channel intensities) and a \texttt{boxplot} of the column intensities.
+In the two plots displayed below, we re-order the TMT channles to pair
+corresponding channels in the two replicates (rather than ordering
+the channels by replicate).
+
+\begin{figure}[!ht]
+  \centering
+<<qcbx, out.width="\\textwidth">>=
+par(mfrow = c(1, 2))
+o <- order(hl$Iodixonal.Density)
+plotDist(hl[, o], pcol = "#00000010")
+lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
+boxplot(exprs(hl[, o]), las = 2)
+@  
+  \caption{Protein profiles and distribution of channel intensities.}
+  \label{fig:qcbx}
+\end{figure}
+
+\section*{Markers}
 
@@
 well-known resident of a specific sub-cellular niche in a species
-\emph{and} condition of interest. Applying this to machine learning
-(ML), and specifically supervised learning, for the task of protein
+*and* condition of interest. Applying this to machine learning (ML),
+and specifically supervised learning, for the task of protein
 localisation prediction, these markers constitute the labelled training
@@
 time consuming, but it is important to define markers that are
-representative of the multivariate data space and on which a classifier
-will be trained and generated.
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} provides a
-convenience function, \texttt{addMarkers}, to directly add markers to a
-\texttt{MSnSet} object, as demonstrated in the code chunk below. These
-marker sets can be accessed using the \texttt{pRolocmarkers()} function.
-Marker sets are stored as a simple named vector in R, and originate from
-in-house user-defined sets of markers or from previous published studies
-{[}@Gatto:2014b{]}, which are continuosly updated and integrated.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## List available marker sets}
-\KeywordTok{pRolocmarkers}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## 7 marker lists available:
-## Arabidopsis thaliana [atha]:
-##  Ids: TAIR, 543 markers
-## Drosophila melanogaster [dmel]:
-##  Ids: Uniprot, 179 markers
-## Gallus gallus [ggal]:
-##  Ids: IPI, 102 markers
-## Homo sapiens [hsap]:
-##  Ids: Uniprot ID, 205 markers
-## Mus musculus [mmus]:
-##  Ids: Uniprot, 937 markers
-## Saccharomyces cerevisiae [scer_sgd]:
-##  Ids: SGD, 259 markers
-## Saccharomyces cerevisiae [scer_uniprot]:
-##  Ids: Uniprot Accession, 259 markers
-\end{verbatim}
-
-These markers can then be mapped to a \texttt{MSnSet}'s
-\texttt{featureNames}. The mouse dataset used here has Uniprot IDs
-stored as the \texttt{featureNames} (see
-\texttt{head(featureNames(hl))}) and the names of the vector of the
-mouse markers stored in
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}
-(\texttt{mmus} markers) are also Uniprot IDs (see \texttt{head(mrk)} in
-the code chunk below), so it is straightforward to match names between
-the markers and the \texttt{MSnSet} instance using the
-\texttt{addMarkers} function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Use mouse markers}
-\NormalTok{mrk <-}\StringTok{ }\KeywordTok{pRolocmarkers}\NormalTok{(}\DataTypeTok{species =} \StringTok{"mmus"}\NormalTok{)}
-\KeywordTok{head}\NormalTok{(mrk)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##                 P26039                 Q6PB66                 P11276 
-##   "Actin cytoskeleton"        "Mitochondrion" "Extracellular matrix" 
-##                 Q6PR54                 Q05793                 P19096 
-##  "Nucleus - Chromatin" "Extracellular matrix"              "Cytosol"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Add mouse markers}
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{addMarkers}\NormalTok{(hl, mrk)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## Markers in data: 937 out of 5032
-\end{verbatim}
-
-\begin{verbatim}
-## organelleMarkers
-##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      27                      43                      13 
-##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                      43                      95                      12 
-##    Extracellular matrix         Golgi apparatus                Lysosome 
-##                      10                      27                      33 
-##           Mitochondrion     Nucleus - Chromatin Nucleus - Non-chromatin 
-##                     383                      64                      85 
-##              Peroxisome         Plasma membrane              Proteasome 
-##                      17                      51                      34 
-##                 unknown 
-##                    4095
-\end{verbatim}
+representative of the multivariate data space and on which a
+classifier will be trained and generated. \Biocpkg{pRoloc}
+provides a convenience function, \texttt{addMarkers}, to directly add markers
+to a \texttt{MSnSet} object, as demonstrated in the code chunk below. These
+marker sets can be accessed using the \texttt{pRolocmarkers()}
+function. Marker sets are stored as a simple named vector in R, and
+originate from in-house user-defined sets of markers or from previous
+published studies \cite{Gatto:2014b}, which are continuosly updated and 
+integrated.
+
+<<markers>>=
+## List available marker sets
+pRolocmarkers()
+@
+
+These markers can then be mapped to a \texttt{MSnSet}'s \texttt{featureNames}. The
+mouse dataset used here has Uniprot IDs stored as the \texttt{featureNames}
+(see \texttt{head(featureNames(hl))}) and the names of the vector of
+the mouse markers stored in \Biocpkg{pRoloc} (\texttt{mmus} markers) 
+are also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below), so 
+it is straightforward to match names between the markers and 
+the \texttt{MSnSet} instance using the \texttt{addMarkers} function.
+
+<<addmrkers>>=
+## Use mouse markers
+mrk <- pRolocmarkers(species = "mmus")
+head(mrk)
+
+## Add mouse markers
+hl <- addMarkers(hl, mrk)
+@
 
 If the naming between the marker sets and the \texttt{MSnSet} dataset
-are different, one will have to convert and match the proteins according
-to the appropriate identifier. Sometimes, we find the equivalent entry
-name, Uniprot ID or accession number is stored with the data, which
-makes conversion between identifers relatively straightforward. If this
-is not the case however, conversion can be performed using
-\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}}, the
-Bioconductor
+are different, one will have to convert and match the proteins
+according to the appropriate identifier. Sometimes, we find the
+equivalent entry name, Uniprot ID or accession number is stored with
+the data, which makes conversion between identifers relatively
+straightforward. If this is not the case however, conversion can be
+performed using \Biocpkg{biomaRt}, the Bioconductor
 \href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation
@@
 We now visualise these annotations along the PCA plot using the
-\texttt{plot2D} function and then use the \texttt{addLegend} function to
-map the marker classes to the pre-defined colours. We also display the
-data along the first and seventh PCs using the \texttt{dims} argument.
-Note that in this second call to the \texttt{plot2D} function, we have
-omitted the \texttt{fcol} argument, as the \texttt{"markers"} feature
-variable name is the default value. We choose to display PCs 1 and 7 to
-illustrate that while upper principal components explain much less
-variability in the data (2.23\% for PC7, as opposed to 48.41\% for PC1),
-we see that the mitochondrial and peroxisome clusters can be
-differenciated, despite the apparent overlap in the two first PCs.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{main =} \StringTok{"pRolocmarkers for mouse"}\NormalTok{)}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{)}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{dims =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{7}\NormalTok{), }\DataTypeTok{main =} \StringTok{"Marker resolution along PC 1 and 7"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\texttt{plot2D} function and then use the \texttt{addLegend} function to map the
+marker classes to the pre-defined colours. We also display the data
+along the first and seventh PCs using the \texttt{dims} argument. Note that
+in this second call to the \texttt{plot2D} function, we have omitted the
+\texttt{fcol} argument, as the \texttt{"markers"} feature variable name is the
+default value. We choose to display PCs 1 and 7 to illustrate that
+while upper principal components explain much less variability in the
+data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
+mitochondrial and peroxisome clusters can be differenciated, despite
+the apparent overlap in the two first PCs.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotmarkers-1.png}
-\caption{plot of chunk plotmarkers}
+<<plotmarkers, out.width="\\textwidth">>=
+par(mfrow = c(1, 2))
+plot2D(hl, main = "pRolocmarkers for mouse")
+addLegend(hl, cex = .7)
+plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
+@  
+  \caption{Annotated PCA plots of the \texttt{hl} data.}
+  \label{fig:plotmarkers}
 \end{figure}
 
-The colours have been defined so as to enable to differenciate up to 30
-classes. If more are provided, different character symbols (circles,
-squares, \ldots{} and empty and solid symbols) are used. The colours and
-the default plotting characters (solid dots for the markers and empty
-circles for the features of unknown localisation) can of course be
-changed, as described in the \texttt{setStockcol} manual page.
-
-As demonstrated in {[}@hyper{]} and illustrated in the PCA plot above,
+The colours have been defined so as to enable to differenciate up to
+30 classes. If more are provided, different character symbols
+(circles, squares, ... and empty and solid symbols) are used. The
+colours and the default plotting characters (solid dots for the
+markers and empty circles for the features of unknown localisation)
+can of course be changed, as described in the \texttt{setStockcol} manual
+page.
+
+As demonstrated in \cite{hyper} and illustrated in the PCA plot above,
 the Golgi apparatus proteins display a dynamic pattern, noting sets of
@@
 \texttt{fDataTounknown} function. It is often used to replace empty
-strings (``'') or missing values in the markers definition to a common
-definition of \emph{unknown localisation}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(hl, }\DataTypeTok{from =} \StringTok{"Golgi apparatus"}\NormalTok{, }\DataTypeTok{to =} \StringTok{"unknown"}\NormalTok{)}
-\KeywordTok{getMarkers}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## organelleMarkers
-##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      27                      43                      13 
-##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                      43                      95                      12 
-##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      10                      33                     383 
-##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                      64                      85                      17 
-##         Plasma membrane              Proteasome                 unknown 
-##                      51                      34                    4122
-\end{verbatim}
-
-In general, the Gene Ontology (GO) {[}@Ashburner:2000{]}, and in
+strings ("") or missing values in the markers definition to a common
+definition of *unknown localisation*.
+
+<<removeGA>>=
+hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
+getMarkers(hl)
+@
+
+In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
 particular the cellular compartment (CC) namespace are a good starting
@@
 note however that automatic retrieval of sub-cellular localisation
-information, from
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} or
-elsewhere, is only the beginning in defining a marker set for downstream
-analyses. Expert curation is vital to check that any annotation added is
-in the correct context for the the biological question under
-investigation.
+information, from \Biocpkg{pRoloc} or elsewhere, is only the
+beginning in defining a marker set for downstream analyses. Expert
+curation is vital to check that any annotation added is in the correct
+context for the the biological question under investigation.
 
@@
 efficiently visualise the complete dataset and assess the relative
-separation of different sub-cellular niches, comparing profiles of a few
-marker clusters is useful to assess how exactly they differ (in terms of
-peak channels, for example). Below, we plot the profile of the
-mitochondrial and peroxisome markers to highlight the differences in
-profiles between these two sets of markers along the 6th and 7th
+separation of different sub-cellular niches, comparing profiles of a
+few marker clusters is useful to assess how exactly they differ (in
+terms of peak channels, for example). Below, we plot the profile of
+the mitochondrial and peroxisome markers to highlight the differences
+in profiles between these two sets of markers along the 6th and 7th
 channels, as represented above along the 7th PC on the PCA plot.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hlo <-}\StringTok{ }\NormalTok{hl[, }\KeywordTok{order}\NormalTok{(hl$Iodixonal.Density)]}
-\KeywordTok{plotDist}\NormalTok{(hlo[}\KeywordTok{fData}\NormalTok{(hlo)$markers ==}\StringTok{ "Mitochondrion"}\NormalTok{, ],}
-         \DataTypeTok{pcol =} \StringTok{"purple"}\NormalTok{, }\DataTypeTok{fractions =} \StringTok{"Fraction.No"}\NormalTok{)}
-\KeywordTok{title}\NormalTok{(}\DataTypeTok{main =} \StringTok{"Marker occupancy profiles along the gradient"}\NormalTok{)}
-\KeywordTok{matlines}\NormalTok{(}\KeywordTok{t}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hlo[}\KeywordTok{fData}\NormalTok{(hlo)$markers ==}\StringTok{ "Peroxisome"}\NormalTok{, ])),}
-         \DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"darkblue"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{)}
-\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"Mitochondrion"}\NormalTok{, }\StringTok{"Peroxisome"}\NormalTok{),}
-       \DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"purple"}\NormalTok{, }\StringTok{"blue"}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotDist-1.png}
-\caption{plot of chunk plotDist}
+<<plotDist, out.width="\\textwidth">>=
+hlo <- hl[, order(hl$Iodixonal.Density)]
+plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
+         pcol = "purple", fractions = "Fraction.No")
+title(main = "Marker occupancy profiles along the gradient")
+matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
+         lty = 1, col = "darkblue", type = "l")
+legend("topleft", c("Mitochondrion", "Peroxisome"),
+       lty = 1, col = c("purple", "blue"), bty = "n")
+@  
+  \caption{Mitochondrion and peroxisome protein profiles.}
+  \label{fig:plotDist2}
 \end{figure}
 
-\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust},
-although these are only available in the development version.
-
-\section{Replication}\label{replication}
+\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust}
+(also mentioned later), although these are only available in the
+development version.
+
+\section*{Replication}
 
@@
 consequently, the reliability in protein sub-cellular assignments, we
-follow the advice in {[}@Trotter:2010{]} and combine replicated spatial
-proteomics experiments as described above. Indeed, Trotter et al. have
-shown a significant improvement in protein--organelle association upon
-direct combination of single experiments, in particular when these
-resolve different subcellular niches.
-
-Direct comparisons of individual channels in replicated experiments does
-not provide an adequate, goal-driven assessment of different
-experiments. Indeed, due to the nature of the experiment and gradient
-fraction collection, the quantitative channels do not correspond to
-identical selected fractions along the gradient. As can be seen in the
-table below (taken from the \texttt{hyperLOPIT}'s \texttt{pData},
-focusing on channels 7 to 10), different sets of gradient fractions are
-pooled to obtain enough material and optimise acurate quantitation.
+follow the advice in \cite{Trotter:2010} and combine replicated spatial
+proteomics experiments as described above. Indeed, Trotter et
+al. have shown a significant improvement in protein–organelle
+association upon direct combination of single experiments, in
+particular when these resolve different subcellular niches.
+
+Direct comparisons of individual channels in replicated experiments
+does not provide an adequate, goal-driven assessment of different
+experiments. Indeed, due to the nature of the experiment and gradient fraction
+collection, the quantitative channels 
+do not correspond to identical selected fractions along the
+gradient. As can be seen in the table below (taken from the
+\texttt{hyperLOPIT}'s \texttt{pData}, focusing on channels 7 to 10), different sets
+of gradient fractions are pooled to obtain enough material and optimise
+acurate quantitation.
+
+<<pdtab, echo=FALSE, eval = FALSE>>=
+i <- c(2, 3, 4, 13, 14)
+pData(hl)[i, ]
+@
+
+% On the figure below, we compare the relative intensities of channels 7
+% to 10 on scatter plots, highlighting different marker sets. These
+% differences result from different peak fractions for these niches on
+% the mitochondrial and chromatin occupancy profiles for the two
+% replicates.
+
+% <<repl1, eval = TRUE, echo=FALSE, fig.width = 12>>=
+% sel0 <- fData(hyperLOPIT)$markers != "unknown"
+% cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
+% cls[!sel0] <- getUnknowncol()
+
+% par(mfrow = c(2, 4))
+% for (i in c(7, 8, 10, 9)) {
+%     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
+%          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
+%     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
+%     grid()
+%     abline(0, 1, col = "red")
+% }
+% plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
+% addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
+
+% i <- fData(hyperLOPIT)$markers == "Mitochondrion"
+% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
+%          featureNames(hyperLOPIT)[i], type = "l",
+%          mcol = "orange")
+% title(main = "Mitochondrion")
+% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
+% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
+% legend("topleft", paste("Replicate", 1:2),
+%        col = c("#F781BF", "orange"),
+%        lty = 1, bty = "n")
+
+% i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
+% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
+%          featureNames(hyperLOPIT)[i], type = "l")
+% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
+% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
+% title(main = "Chromatin")
+% legend("topleft", paste("Replicate", 1:2),
+%        col = c("#984EA3", "steelblue"),
+%        lty = 1, bty = "n")
+% @
 
@@
 assuring that each displays acceptable sub-cellular resolution. Note
-that in the code chunk below, we mirror the x-axis to represent the two
-figures with the same orientation.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{)) }
-\KeywordTok{plot2D}\NormalTok{(hl[, hl$Replicate ==}\StringTok{ }\DecValTok{1}\NormalTok{], }\DataTypeTok{main =} \StringTok{"Replicate 1"}\NormalTok{) }
-\KeywordTok{plot2D}\NormalTok{(hl[, hl$Replicate ==}\StringTok{ }\DecValTok{2}\NormalTok{], }\DataTypeTok{main =} \StringTok{"Replicate 2"}\NormalTok{, }\DataTypeTok{mirrorX =} \OtherTok{TRUE}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+that in the code chunk below, we mirror the x-axis to represent the
+two figures with the same orientation.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plot2Drep-1.png}
-\caption{plot of chunk plot2Drep}
+<<plot2Drep, out.width="\\textwidth">>= 
+par(mfrow = c(1, 2)) 
+plot2D(hl[, hl$Replicate == 1], main = "Replicate 1") 
+plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE) 
+@  
+  \caption{PCA plots of replicates 1 and 2}
+  \label{fig:plot2Drep}
 \end{figure}
 
-\section{Interactive visualisation}\label{interactive-visualisation}
+\section*{Interactive visualisation}
 
@@
 analyses allowing one to shed light on data structure and patterns of
-interest. Using the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package we can interactively visualise, explore and interrogate
-quantitative spatial proteomics data. The
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package is currently under active development and it relies on the
-\texttt{shiny} framework for reactivity and interactivity. The package
-currently distributes 3 different GUI's (``main'' (default), ``compare''
-or ``compare'') which are wrapped and launched by the \texttt{pRolocVis}
-function. In the below code chunk we lauch the main app (note, we do not
-need to specify the argument, \texttt{app = "main"} as it is the
-default).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocGUI"}\NormalTok{)}
-\KeywordTok{pRolocVis}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+interest. Using the \Biocpkg{pRolocGUI} package we can
+interactively visualise, explore and interrogate quantitative spatial
+proteomics data. The \Biocpkg{pRolocGUI} package is currently
+under active development and it relies on the \texttt{shiny} framework for
+reactivity and interactivity. The package currently distributes 3
+different GUI's ("main" (default), "compare" or "compare") which are
+wrapped and launched by the \texttt{pRolocVis} function. In the below code
+chunk we lauch the main app (note, we do not need to specify the
+argument, \texttt{app = "main"} as it is the default).
+
+<<mainapp, eval = FALSE, echo = TRUE>>=
+library("pRolocGUI")
+pRolocVis(hl)
+@
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/mainapp.png}
-\caption{A screen shot of clickable interface and zoomable PCA plot of
-the main app in the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package.}
+  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
+  \caption{A screen shot of clickable interface and zoomable PCA plot of the main app in the \Biocpkg{pRolocGUI} package.}
+  \label{fig:app}
 \end{figure}
 
-As diplayed in the screenshot above, the ``main'' application is
+As diplayed in the screenshot above, the "main" application is
 designed for exploratory data analysis and is divied into 3 tabs: (1)
-PCA, (2) Profiles and (3) Table selection. The default view upon loading
-is the PCA tab, which features a clickable interface and zoomable PCA
-plot with an interactive data table for displaying the quantitation
-information. Particular proteins of interest can be highlighted using
-the text search box. There is also an alternate profiles tab for
-visualisation of the protein profiles, which can be used to examine the
-patterns of proteins of interest. The Table selection tab provides an
-interface to control data table column selection.
-
-The ``compare'' application is useful for examining two replicate
+PCA, (2) Profiles and (3) Table selection. The default view upon
+loading is the PCA tab, which features a clickable interface and
+zoomable PCA plot with an interactive data table for displaying the
+quantitation information. Particular proteins of interest can be
+highlighted using the text search box. There is also an alternate
+profiles tab for visualisation of the protein profiles, which can be
+used to examine the patterns of proteins of interest. The Table
+selection tab provides an interface to control data table column
+selection.
+
+The \textit{compare} application is useful for examining two replicate
 experiments, or two experiments from different conditions, treatments
-etc. The compare application is called by default if the input object to
-\texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets}, but
-it can also be specified by calling the argument
-\texttt{app = "compare"}. For example, in the code chunk below we first
-create a \texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT
-data, this is then passed to \texttt{pRolocVis}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r1)}
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r2)}
-\NormalTok{mydata <-}\StringTok{ }\KeywordTok{MSnSetList}\NormalTok{(}\KeywordTok{list}\NormalTok{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
-\KeywordTok{pRolocVis}\NormalTok{(mydata, }\DataTypeTok{app =} \StringTok{"compare"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+etc. The compare application is called by default if the input object
+to \texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets},
+but it can also be specified by calling the argument \texttt{app =
+  "compare"}. For example, in the code chunk below we first create a
+\texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT data, this
+is then passed to \texttt{pRolocVis}.
+
+<<compareapp, eval = FALSE, echo = TRUE>>=
+data(hyperLOPIT2015ms3r1)
+data(hyperLOPIT2015ms3r2)
+mydata <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
+pRolocVis(mydata, app = "compare")
+@
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/SS_Compare1.jpg}
+  \includegraphics[width=\textwidth]{./Figures/SS_Compare1.jpg}
   \caption{The compare application, main panel}
+  \label{fig:compare}
 \end{figure}
@@
 proteins between the two data sets are displayed. As per the main
-application proteins can be searched, identified and highlighted on both
-PCA plots and in the dedicated profiles tab. One key feature of the
-compare application is the ability to re-map the second dataset onto the
-PCA data space of the first (reference) data set (see
-\texttt{?pRolocVis} and the argument \texttt{remap = TRUE}). Using the
-first dataset as the reference set, PCA is carried out on the first
-dataset and the standard deviations of the principal components
-(i.e.~the square roots of the eigenvalues of the covariance/correlation
-matrix) and the matrix of variable loadings (i.e.~a matrix whose columns
-contain the eigenvectors) are stored and then used to calculate the
-principal components of the second dataset. Both datasets are scaled and
+application proteins can be searched, identified and highlighted on
+both PCA plots and in the dedicated profiles tab. One key feature of
+the compare application is the ability to re-map the second dataset
+onto the PCA data space of the first (reference) data set (see
+\texttt{?pRolocVis} and the argument \texttt{remap = TRUE}). Using the first dataset
+as the reference set, PCA is carried out on the first dataset and the
+standard deviations of the principal components (i.e. the square roots
+of the eigenvalues of the covariance/correlation matrix) and the
+matrix of variable loadings (i.e. a matrix whose columns contain the
+eigenvectors) are stored and then used to calculate the principal
+components of the second dataset. Both datasets are scaled and
 centered in the usual way. The first dataset appears on the left, and
@@
 
-The final application ``classify'', has been designed to view machine
+The final application "classify", has been designed to view machine
 learning classification results according to user-specified thresholds
 for the assignment of proteins to its sub-cellular location, this is
-discussed later in the subsection \hyperref[thresholding]{thresholding}
-in the \hyperref[supervised-machine-learning]{supervised machine
-learning section}.
-
-\section{Novelty Detection}\label{novelty-detection}
+discussed later in the subsection \textit{thresholding} in the
+\textit{supervised machine learning section}.
+
+\section*{Novelty Detection}
 
@@
 to the limited number of marker proteins that exist in databases and
-elsewhere. Furthermore, given the vast complexity of the cell, automatic
-annotation retrieval does not always give a full representation of the
-true sub-cellular diversity in the data. For downstream analyses, such
-as supervised machine learning, it is desirable to obtain reliable
-markers that cover as many sub-cellular niches as possible, as these
-markers are directly used in the training phase of the ML
-classification. We find that a lack of sub-cellular diversity in the
-labelled training data leads to prediction errors, as unlabelled
-instances can only be assigned to a class that exists in the training
-data {[}@Breckels:2013{]}. In such scenarios novelty detection can be
-useful to identify data-specific sub-cellular groupings such as
-organelles and protein complexes. The phenotype discovery (phenoDisco)
-algorithm {[}@Breckels:2013{]} is one such method and is available in
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}. It is an
-iterative semi-supervised learning method that combines the
-classification of proteins on existing labelled data with the detection
-of new clusters.
-
-In addition to extracting new phenotypes, novelty detection methods are
-also useful for confirming the presence of known or postulated clusters
-in an unbiased fashion. For example, in {[}@hyper{]} the
-\texttt{phenoDisco} algorithm was used to confirm the data-specific
-presence of the nucleus and nucleus sub-compartments. In the code chunk
-below, we demonstrate how to do this analysis, highlighting some of the
-optional arguments and parameters available for phenotype extraction and
-give some advice on how to interpret the output.
-
-As the \texttt{phenoDisco} algorithm is semi-supervised it uses both
-labelled (markers) and unlabelled data to explore the data structure and
-find new sub-cellular data clusters. Thus the first step is to define
-some input labelled data i.e.~markers, that the algorithm will use as
-input for the supervised learning aspect of the algorithm. As described
-in {[}@hyper{]} we define a set of markers to use as input for the
-analyses that cover well-known residents from three distinct organelle
+elsewhere. Furthermore, given the vast complexity of the cell,
+automatic annotation retrieval does not always give a full
+representation of the true sub-cellular diversity in the data. For
+downstream analyses, such as supervised machine learning, it is
+desirable to obtain reliable markers that cover as many sub-cellular
+niches as possible, as these markers are directly used in the training
+phase of the ML classification. We find that a lack of sub-cellular
+diversity in the labelled training data leads to prediction errors, as
+unlabelled instances can only be assigned to a class that exists in
+the training data \cite{Breckels:2013}. In such scenarios novelty
+detection can be useful to identify data-specific sub-cellular
+groupings such as organelles and protein complexes. The phenotype
+discovery (phenoDisco) algorithm \cite{Breckels:2013} is one such method
+and is available in \Biocpkg{pRoloc}. It is an iterative
+semi-supervised learning method that combines the classification of
+proteins on existing labelled data with the detection of new clusters.
+
+In addition to extracting new phenotypes, novelty detection methods
+are also useful for confirming the presence of known or postulated
+clusters in an unbiased fashion. For example, in \cite{hyper} the
+\texttt{phenoDisco} algorithm was used to confirm the data-specific presence
+of the nucleus and nucleus sub-compartments. In the code chunk below,
+we demonstrate how to do this analysis, highlighting some of the
+optional arguments and parameters available for phenotype extraction
+and give some advice on how to interpret the output.
+
+As the \texttt{phenoDisco} algorithm is semi-supervised it uses both labelled
+(markers) and unlabelled data to explore the data structure and find
+new sub-cellular data clusters. Thus the first step is to define some
+input labelled data i.e. markers, that the algorithm will use as input
+for the supervised learning aspect of the algorithm. As described in
+\cite{hyper} we define a set of markers to use as input for the analyses
+that cover well-known residents from three distinct organelle
 structures; the mitochondria, plasma membrane and ER, and from three
 well-known and abundant protein complexes; the proteasome and two
-ribosomal subunits, 40S and 60S. These input markers are stored in the
-\texttt{featureData} column of \texttt{hl} where
-\texttt{fcol = "phenoDisco.Input"}. We can use the convenience accessor
-function \texttt{getMarkers} to print out a table of the markers
-contained in this marker set. These initial markers were manually
-curated using information from the UniProt database, the Gene Ontology
-and the literature.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## organelleMarkers
-##                          40S Ribosome 
-##                                    26 
-##                          60S Ribosome 
-##                                    43 
-## Endoplasmic reticulum/Golgi apparatus 
-##                                    76 
-##                         Mitochondrion 
-##                                   261 
-##                       Plasma membrane 
-##                                    50 
-##                            Proteasome 
-##                                    34 
-##                               unknown 
-##                                  4542
-\end{verbatim}
+ribosomal subunits, 40S and 60S.  These input markers are stored in
+the \texttt{featureData} column of \texttt{hl} where \texttt{fcol = "phenoDisco.Input"}. We
+can use the convenience accessor function \texttt{getMarkers} to print out a
+table of the markers contained in this marker set. These initial
+markers were manually curated using information from the UniProt
+database, the Gene Ontology and the literature.
+
+<<displayPDmarkers>>=
+getMarkers(hl, fcol = "phenoDisco.Input")
+@
 
@@
 function and return a novelty detection result, according to the
-specified parameters. The algorithm parameters \texttt{times},
-\texttt{GS} and \texttt{p} are passed to the function, along with the
-\texttt{fcol} to tell the algorithm where the input training data is
-contained.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## As per Christoforou et al (2016), }
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{phenoDisco}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }
-                 \DataTypeTok{times =} \DecValTok{200}\NormalTok{, }
-                 \DataTypeTok{GS =} \DecValTok{60}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Note: We do not evaluate this code chunk in this document as the
-algorithm is computational intensive and best parallelised over multiple
-workers. This phenoDisco analysis took \textasciitilde{}24 hours to
-complete when parallelised over 40 workers.}
-
-The argument \texttt{times} indicated the number of times we run
-unsupervied Gaussian Mixture Modelling before defining a new phenotype
-cluster. The recommended minimum and default value is 100. In the above
-code chunk we increase the value to \texttt{times = 200} as we have
-found for larger datasets (e.g.~5000+ proteins) a higher times is
-requried for convergence. \texttt{GS} defines the minimum number of
-proteins allowed per new data cluster and thus heavily influences what
-type of new clusters are extracted. For example, if a user is interesed
-in the detection of small complexes they may wish to use a small
-\texttt{GS = 10}, or \texttt{= 20} etc. If they wish to detect larger
-more abundant sub-cellular niches a much higher \texttt{GS} would be
-preferable. Specifying a small \texttt{GS} can be more time consuming
-than using a larger \texttt{GS}, and there is a trade off between
-finding interesting small complexes and those that may not be of
-interest as we find there is a tendancy to find more noise when using a
-small \texttt{GS} compared to using a higher one.
-
-One may also consider increasing the search space for new data clusters
-by increasing the value of the parameter \texttt{G}. This defines the
-number of GMM components to test and fit; the default is
-\texttt{G = 1:9} (\texttt{mclust}'s default). One should note that the
-decreasing the \texttt{GS}, and increasing the values of the arguments
-\texttt{times}, \texttt{G} (among other function arguments, see
-\texttt{?phenoDisco}) will heavily influence (increase) the total time
-taken to run the algorithm. \texttt{phenoDisco} supports parallelisation
-and we strongly suggest you make use of a parallel processing to run
-these analyses (different backends can be set with argument
-\texttt{BPPARAM}).
-
-The ouput of running the \texttt{phenoDisco} algorithm is a
-\texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. We can see by typing
-\texttt{hl@processingData} directly into the console the processing
-information has been updated to the \texttt{MSnSet} recording the
-parameters that were used to run the analyses. This is handy for keeping
-track of data analyses. The results can be displayed by using the
-\texttt{getMarkers} function. We see that 5 new phenotype data clusters
-were found.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 5032 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData
-##   sampleNames: X126 X127N ... X131.1 (20 total)
-##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
-##   varMetadata: labelDescription
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
-##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-## Annotation:  
-## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Sep 29 10:28:15 2016 
-## Added markers from  'pdres' marker vector. Thu Sep 29 10:28:17 2016 
-##  MSnbase version: 1.99.2
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## organelleMarkers
-##                          40S Ribosome 
-##                                   106 
-##                          60S Ribosome 
-##                                    95 
-## Endoplasmic reticulum/Golgi apparatus 
-##                                   393 
-##                         Mitochondrion 
-##                                   525 
-##                           Phenotype 1 
-##                                   300 
-##                           Phenotype 2 
-##                                   253 
-##                           Phenotype 3 
-##                                   203 
-##                           Phenotype 4 
-##                                    74 
-##                           Phenotype 5 
-##                                    91 
-##                       Plasma membrane 
-##                                   421 
-##                            Proteasome 
-##                                    92 
-##                               unknown 
-##                                  2479
-\end{verbatim}
+specified parameters. The algorithm parameters \texttt{times}, \texttt{GS} and \texttt{p}
+are passed to the function, along with the \texttt{fcol} to tell the
+algorithm where the input training data is contained.
+
+<<runPD, eval = FALSE>>=
+## As per Christoforou et al (2016), 
+hl <- phenoDisco(hl, fcol = "phenoDisco.Input", 
+                 times = 200, 
+                 GS = 60)
+@
+
+Note: We do not evaluate this code chunk in this document as the
+algorithm is computational intensive and best parallelised over
+multiple workers. This phenoDisco analysis took ~24 hours to complete
+when parallelised over 40 workers.
+
+The argument \texttt{times} indicated the number of times we run unsupervied
+Gaussian Mixture Modelling before defining a new phenotype cluster.
+The recommended minimum and default value is 100. In the above code
+chunk we increase the value to \texttt{times = 200} as we have found for
+larger datasets (e.g. 5000+ proteins) a higher times is requried for
+convergence. \texttt{GS} defines the minimum number of proteins allowed per
+new data cluster and thus heavily influences what type of new clusters
+are extracted. For example, if a user is interesed in the detection of
+small complexes they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
+If they wish to detect larger more abundant sub-cellular niches a much
+higher \texttt{GS} would be preferable. Specifying a small \texttt{GS} can be more
+time consuming than using a larger \texttt{GS}, and there is a trade off
+between finding interesting small complexes and those that may not be
+of interest as we find there is a tendancy to find more noise when
+using a small \texttt{GS} compared to using a higher one.
+
+One may also consider increasing the search space for new data
+clusters by increasing the value of the parameter \texttt{G}. This defines
+the number of GMM components to test and fit; the default is \texttt{G = 1:9}
+(\texttt{mclust}'s default). One should note that the decreasing the \texttt{GS},
+and increasing the values of the arguments \texttt{times}, \texttt{G} (among other
+function arguments, see \texttt{?phenoDisco}) will heavily influence
+(increase) the total time taken to run the algorithm.  \texttt{phenoDisco}
+supports parallelisation and we strongly suggest you make use of a
+parallel processing to run these analyses (different backends can be
+set with argument \texttt{BPPARAM}).
+
+<<loadPDres, echo=FALSE>>=
+f0 <- dir(extdatadir, full.names = TRUE,
+          pattern = "bpw-pdres.rds")
+pdres <- readRDS(f0)
+hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
+setStockcol(NULL)
+setStockcol(paste0(getStockcol(), 70))
+@
+
+The ouput of running the \texttt{phenoDisco} algorithm is a \texttt{MSnSet}
+containing the new data clusters, appended to the \texttt{featureData} under
+the name \texttt{pd}. We can see by typing \texttt{hl@processingData} directly into
+the console the processing information has been updated to the
+\texttt{MSnSet} recording the parameters that were used to run the analyses.
+This is handy for keeping track of data analyses. The results can be
+displayed by using the \texttt{getMarkers} function. We see that 5 new
+phenotype data clusters were found.
+
+<<showPDres>>=
+hl
+getMarkers(hl, fcol = "pd")
+@
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Re-order the colours for the phenoDisco output}
-\NormalTok{cl <-}\StringTok{ }\KeywordTok{getMarkerClasses}\NormalTok{(hl, }\StringTok{"pd"}\NormalTok{)}
-\NormalTok{cols <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)]}
-\NormalTok{ind <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"Pheno"}\NormalTok{, cl, }\DataTypeTok{invert =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{cols[ind] <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)][}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(ind)]}
-\NormalTok{cols[-ind] <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)][(}\KeywordTok{length}\NormalTok{(ind) +}\StringTok{ }\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(cl)]}
-
-\NormalTok{## Plot the input and output}
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }
-       \DataTypeTok{main =} \StringTok{"phenoDisco input markers"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{getStockcol}\NormalTok{()[}\DecValTok{1}\NormalTok{:}\DecValTok{6}\NormalTok{])}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{)}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"phenoDisco output"}\NormalTok{, }\DataTypeTok{col =} \NormalTok{cols)}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{col =} \NormalTok{cols)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotPDres-1.png}
-\caption{plot of chunk plotPDres}
+<<plotPDres, out.width="\\textwidth">>=
+## Re-order the colours for the phenoDisco output
+cl <- getMarkerClasses(hl, "pd")
+cols <- getStockcol()[seq(cl)]
+ind <- grep("Pheno", cl, invert = TRUE)
+cols[ind] <- getStockcol()[seq(cl)][1:length(ind)]
+cols[-ind] <- getStockcol()[seq(cl)][(length(ind) + 1):length(cl)]
+
+## Plot the input and output
+par(mfrow = c(1, 2))
+plot2D(hl, fcol = "phenoDisco.Input", 
+       main = "phenoDisco input markers", col = getStockcol()[1:6])
+addLegend(hl, fcol = "phenoDisco.Input", cex = .7)
+plot2D(hl, fcol = "pd", main = "phenoDisco output", col = cols)
+addLegend(hl, fcol = "pd", cex = .7, col = cols)
+@
+  \caption{Results of the novelty detection algorithm.}
+  \label{fig:plotPDres}
 \end{figure}
 
-The five new phenotype data clusters can be extracted and examined. In
-the code chunk below we write the results to a .csv file. We use the
-argument \texttt{fDataCols} to specify which columns of the
+The five new phenotype data clusters can be extracted and examined.
+In the code chunk below we write the results to a .csv file. We use
+the argument \texttt{fDataCols} to specify which columns of the
 \texttt{featureData} to write.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fData}\NormalTok{(hl)$pd <-}\StringTok{ }\KeywordTok{as.character}\NormalTok{(}\KeywordTok{fData}\NormalTok{(hl)$pd)}
-\KeywordTok{write.exprs}\NormalTok{(hl, }\DataTypeTok{fDataCols =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"pd-results.csv"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{","}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-We can also examine the each phenotype intercatively and visualise their
-protein profiles by using the \texttt{pRolocVis} function in the
-\texttt{pRolocGUI}. We found that phenotype 1 was enriched in nucleus
-associated proteins, phenotype 2 in chromatin associated proteins,
-phenotype 3 in cytosolic and phenotypes 4 and 5 in lysosomal and
-endosomal proteins.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{pRolocVis}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hyperdef{}{supervised-machine-learning}{\section{Supervised machine
-learning}\label{supervised-machine-learning}}
+<<writehl, eval=FALSE>>=
+fData(hl)$pd <- as.character(fData(hl)$pd)
+write.exprs(hl, fDataCols = "pd", file = "pd-results.csv", sep = ",")
+@
+
+We can also examine the each phenotype intercatively and visualise
+their protein profiles by using the \texttt{pRolocVis} function in the
+\Biocpkg{pRolocGUI} package. We found that phenotype 1 was
+enriched in nucleus associated proteins, phenotype 2 in chromatin
+associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5
+in lysosomal and endosomal proteins.
+
+<<vishl, eval=FALSE>>=
+pRolocVis(hl, fcol = "pd")
+@
+
+\section*{Supervised machine learning}
 
@@
 essential tool for the assignment of proteins to distinct sub-cellular
-niches. Using a set of labelled training examples i.e.~markers, we can
-train a machine learning classifier to learn a mapping between the data
-i.e.~the quantitative protein profiles, and a known localisation. The
-trained classifier can then be used to predict the localisation of a
-protein of unknown localisation, based on its observed protein profile.
-To date, this method has been extensively used in spatial quantitative
-proteomics to assign thousands of proteins to distinct sub-cellular
-niches {[}@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006;
-@Tan:2009{]}.
+niches. Using a set of labelled training examples i.e. markers, we can
+train a machine learning classifier to learn a mapping between the
+data i.e. the quantitative protein profiles, and a known localisation.
+The trained classifier can then be used to predict the localisation of
+a protein of unknown localisation, based on its observed protein
+profile. To date, this method has been extensively used in spatial
+quantitative proteomics to assign thousands of proteins to distinct
+sub-cellular niches
+\cite{hyper,Groen:2014,Trotter:2010,Hall:2009,Dunkley:2006,Tan:2009}.
 
 There are several classification algorithms readily available in
-\texttt{pRoloc}, which are documented in the dedicated
-\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\texttt{pRoloc}
-machine learning techniques vignette}. We find the general tendancy to
-be that it is not the choice of classifier, but the improper
+\Biocpkg{pRoloc}, which are documented in the dedicated
+\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\Biocpkg{pRoloc}
+  machine learning techniques vignette}.  We find the general tendancy
+to be that it is not the choice of classifier, but the improper
 optimisation of the algorithmic parameters, that limits classification
-accuracy. Before employing any classification algorithm and generating a
-model on the training data, one must first find the optimal parameters
-for the algorithm of choice.
-
-\subsection{Optimisation}\label{optimisation}
+accuracy. Before employing any classification algorithm and generating
+a model on the training data, one must first find the optimal
+parameters for the algorithm of choice.
+
+\subsection*{Optimisation}
 
 In the code chunk below we employ the use of a Support Vector Machine
-(SVM) to learn a classifier on the labelled training data. As previously
-mentioned one first needs to train the classifiers parameters before an
-algorithm can be used to predict the class labels of the proteins with
-unknown location. One of the most common ways to optimise the parameters
-of a classifier is to partition the labelled data in to training and
-testing subsets. In this framework parameters are tested via a grid
-search using cross-validation on the training partition. The best
-parameters chosen from the cross-validation stage are then used to build
-a classifier to predict the class labels of the protein profiles on the
-test partition. Observed and expected classication results can be
-compared, and then used to assess how well a given model works by
-getting an estimate of the classiers ability to achieve a good
-generalisation i.e.~that is given an unknown example predict its class
-label with high accuracy. In \texttt{pRoloc} algorithmic performance is
-estimated using stratified 80/20 partitioning for the training/testing
-subsets respectively, in conjuction with five-fold cross-validation in
-order to optimise the free parameters via a grid search. This procedure
-is usually repeated 100 times and then the best parameter(s) are
-selected upon investigation of classifier accuracy, here we use the
-harmonic mean of precision and recall; the macro F1 score. In the code
-chunk below we demonstrate how to optimise the free parameters;
-\texttt{sigma} and \texttt{cost}, of a classical SVM classifier with a
-Gaussian kernel using the function \texttt{svmOptimisation}. As the
-number of labelled instances per class varies from organelle to
-organelle, we can account for class imbalance by setting specific class
-weights when generating the SVM model. Below the weights, \texttt{w} are
-set to be inversely proportional to the class frequencies.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{w <-}\StringTok{ }\KeywordTok{table}\NormalTok{(}\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## organelleMarkers
-##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      27                      43                      13 
-##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                      43                      95                      12 
-##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      10                      33                     383 
-##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                      64                      85                      17 
-##         Plasma membrane              Proteasome                 unknown 
-##                      51                      34                    4122
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{w <-}\StringTok{ }\DecValTok{1}\NormalTok{/w[}\KeywordTok{names}\NormalTok{(w) !=}\StringTok{ "unknown"}\NormalTok{]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## 100 rounds of optimisation with five-fold cross-validation}
-\NormalTok{params <-}\StringTok{ }\KeywordTok{svmOptimisation}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"markers"}\NormalTok{,}
-                          \DataTypeTok{times =} \DecValTok{100}\NormalTok{, }\DataTypeTok{xval =} \DecValTok{5}\NormalTok{,}
-                          \DataTypeTok{class.weights =} \NormalTok{w)}
-\end{Highlighting}
-\end{Shaded}
+(SVM) to learn a classifier on the labelled training data. As
+previously mentioned one first needs to train the classifiers
+parameters before an algorithm can be used to predict the class labels
+of the proteins with unknown location. One of the most common ways to
+optimise the parameters of a classifier is to partition the labelled
+data in to training and testing subsets. In this framework parameters
+are tested via a grid search using cross-validation on the training
+partition. The best parameters chosen from the cross-validation stage
+are then used to build a classifier to predict the class labels of the
+protein profiles on the test partition. Observed and expected
+classication results can be compared, and then used to assess how well
+a given model works by getting an estimate of the classiers ability to
+achieve a good generalisation i.e. that is given an unknown example
+predict its class label with high accuracy. In
+\Biocpkg{pRoloc}, algorithmic performance is estimated using
+stratified 80/20 partitioning for the training/testing subsets
+respectively, in conjuction with five-fold cross-validation in order
+to optimise the free parameters via a grid search. This procedure is
+usually repeated 100 times and then the best parameter(s) are selected
+upon investigation of classifier accuracy, here we use the harmonic
+mean of precision and recall; the macro F1 score. In the code chunk
+below we demonstrate how to optimise the free parameters; \texttt{sigma} and
+\texttt{cost}, of a classical SVM classifier with a Gaussian kernel using the
+function \texttt{svmOptimisation}. As the number of labelled instances per
+class varies from organelle to organelle, we can account for class
+imbalance by setting specific class weights when generating the SVM
+model. Below the weights, \texttt{w} are set to be inversely proportional to
+the class frequencies.
+
+<<loadpdRes, echo=FALSE>>=
+svmf <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-svmopt.rds") 
+params <- readRDS(svmf)
+@
+
+<<getsvmweight>>=
+w <- table(getMarkers(hl, verbose = TRUE))
+w <- 1/w[names(w) != "unknown"]
+@
+
+<<dosvmopt, eval=FALSE>>=
+## 100 rounds of optimisation with five-fold cross-validation
+params <- svmOptimisation(hl, fcol = "markers",
+                          times = 100, xval = 5,
+                          class.weights = w)
+@
 
 As mentioned previously, we reply on the default feature variable
-\texttt{"markers"} to define the class labels and hence can ommit it. To
-use another feature variables, one need to explicitly specify its name
-using the \texttt{fcol} argument (for example
-\texttt{fcol = "markers2"}).
+\texttt{"markers"} to define the class labels and hence can ommit
+it. To use another feature variables, one need to explicitly specify
+its name using the \texttt{fcol} argument (for example \texttt{fcol =
+  "markers2"}).
 
@@
 machine learning optimisation. To assess classifier performance we can
-examine the macro F1 scores and the most frequently chosen parameters. A
-high macro F1 score indicates that the marker proteins in the test
-dataset are consistently and correctly assigned by the the algorithm.
-Often more than one parameter or set of parameters gives rise to the
-best generalisation accuracy. As such it is always important to
-investigate the model parameters and critically assess the best choice.
-The best choice may not be as simple as the parameter set that gives
-rise to the highest macro F1 score and one must be careful to avoid
-overfitting and to choose parameters wisely.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{(best <-}\StringTok{ }\KeywordTok{getParams}\NormalTok{(params))}
-\KeywordTok{plot}\NormalTok{(params)}
-\KeywordTok{levelPlot}\NormalTok{(params)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+examine the macro F1 scores and the most frequently chosen
+parameters. A high macro F1 score indicates that the marker proteins
+in the test dataset are consistently and correctly assigned by the the
+algorithm. Often more than one parameter or set of parameters gives
+rise to the best generalisation accuracy. As such it is always
+important to investigate the model parameters and critically assess
+the best choice. The best choice may not be as simple as the parameter
+set that gives rise to the highest macro F1 score and one must be
+careful to avoid overfitting and to choose parameters wisely.
+
+<<visualiseOpt, eval=FALSE>>=
+(best <- getParams(params))
+plot(params)
+levelPlot(params)
+@
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/visualiseOptHide-1.png}
-\caption{plot of chunk visualiseOptHide}
+<<visualiseOptHide, echo=FALSE, out.width="\\textwidth">>=
+library('gridExtra')
+grid.arrange(plot(params), levelPlot(params), ncol = 2)
+@  
+  \caption{Assessment of the classification model parameter optimisation.}
+  \label{fig:visualisOptHide}
 \end{figure}
@@
 By using the function \texttt{getParams} we can extract the best set of
-parameters. Currently, \texttt{getParams} retrieves the first best set
-is automatically but users are encouraged to critically assess whether
+parameters. Currently, \texttt{getParams} retrieves the first best set is
+automatically but users are encouraged to critically assess whether
 this is the most wise choice by visualising the results using the
-methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method
-for \texttt{GenRegRes} object shows the respective distributions of the
-100 macro F1 scores for the best cost/sigma parameter pairs, and
-\texttt{levelPlot} shows the averaged macro F1 scores, for the full
-range of parameter values. Once we have selected the best parameters we
-can then use them to build a classifier from the labelled marker
-proteins.
-
-\subsection{Classification}\label{classification}
-
-We can use the function \texttt{svmClassification} to return a
-classification result for all unlabelled instances in the dataset
-corresponding to their most likely sub-cellular location. The algorithm
-parameters are passed to the function, along with the class weights. As
-above, \texttt{fcol} can be ignored as we use the labels defined in the
-default \texttt{"markers"} feature variable.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{svmClassification}\NormalTok{(hl, params,}
-                        \DataTypeTok{class.weights =} \NormalTok{w,}
-                        \DataTypeTok{fcol =} \StringTok{"markers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method for \texttt{GenRegRes}
+object shows the respective distributions of the 100 macro F1 scores
+for the best cost/sigma parameter pairs, and \texttt{levelPlot} shows the
+averaged macro F1 scores, for the full range of parameter values. Once
+we have selected the best parameters we can then use them to build a
+classifier from the labelled marker proteins.
+
+\subsection*{Classification}
+
+We can use the function \texttt{svmClassification} to return a classification
+result for all unlabelled instances in the dataset corresponding to
+their most likely sub-cellular location. The algorithm parameters are
+passed to the function, along with the class weights. As above, \texttt{fcol}
+can be ignored as we use the labels defined in the default \texttt{"markers"}
+feature variable.
+
+<<classify>>=
+hl <- svmClassification(hl, params,
+                        class.weights = w,
+                        fcol = "markers")
+@
 
@@
 predictions and assignment scores, are stored in the
-\texttt{featureData} slot of the \texttt{MSnSet}. In this case, they are
-given the labels \texttt{svm} and \texttt{svm.scores} for the
+\texttt{featureData} slot of the \texttt{MSnSet}. In this case, they
+are given the labels \texttt{svm} and \texttt{svm.scores} for the
 predictions and scores respectively. The resultant predictions can be
@@
 argument to change the size of each point on the plot (where one point
-represents one protein) to be inversely proportional to the SVM score.
-This gives an initial overview of the high scoring localisations from
-the SVM predictions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## set point size of each protein to be inversely proportional to the}
-\NormalTok{ptsze <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(}\KeywordTok{fData}\NormalTok{(hl)$svm.scores) -}\StringTok{ }\DecValTok{1}
-
-\NormalTok{## plot new predictions}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{ptsze)}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }\DataTypeTok{where =} \StringTok{"bottomleft"}\NormalTok{, }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+represents one protein) to be inversely proportional to the SVM
+score. This gives an initial overview of the high scoring
+localisations from the SVM predictions.
+
+<<setColors, include=FALSE>>=
+setStockcol(NULL)
+setStockcol(paste0(getStockcol(), 70))
+par(mfrow = c(1, 1))
+@
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotSVM-1.png}
-\caption{plot of chunk plotSVM}
+<<plotSVM, out.width="\\textwidth">>=
+## set point size of each protein to be inversely proportional to the
+ptsze <- exp(fData(hl)$svm.scores) - 1
+## plot new predictions
+plot2D(hl, fcol = "svm", cex = ptsze)
+addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
+@  
+  \caption{Classification results.}
+  \label{fig:plotSVM}
 \end{figure}
@@
 The adjustment of the point size intuitively confers important
-information that is more difficult to defined formally, but that we will
-address in the next section. The classifier (SVM in our case, but this
-is also valid of other classifiers) defines boundaries based on the
-labelled marker proteins. These class/organelle boundaries define how
-non-assigned proteins are classified and with what confidence.
-\textbf{TODO} explain how choosing markers positively/negatively affects
-this, and relation with thresholding.
-
-\hyperdef{}{thresholding}{\subsection{Thresholding}\label{thresholding}}
-
-It is common when applying a supervised classification algorithm to set
-a specific score cutoff on which to define new assignments, below which
-classifications are set to unknown/unassigned. This is important as in a
-supervised learning setup proteins can only be predicted to be localised
-to one of the sub-cellular niches that appear in the labelled training
-data. We can not guaranette (and do not expect) that the whole class
-diversity to be represented in the labelled training data as (1) finding
-markers that represent the whole diversity of the cell is challenging
-(especially obtaining dual- and multiply-localised protein markers) and
-(2) many sub-cellular niches contain too few proteins to train on (we
-recommend a minimum of 13 markers per sub-cellular class for stratified
-80/20 partitioning and 5-fold cross-validation - this allows a minimum
-of \textasciitilde{}10 examples for parameter optimisation on the
-training partition i.e.~2 per fold for 5-fold cross-validation, and then
-\textasciitilde{}3 for testing the best parameters on the validation
-set).
-
-Deciding on a threshold is not trivial as classifier scores are heavily
-dependent upon the classifier used and different sub-cellular niches can
-exhibit different score distributions, as highlighted in the boxplot
-below. We recommend users to set class-specific thresholds. In the code
-chunk below we display a boxplot of the score distributions per
-organelle.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## First remove the markers}
-\NormalTok{preds <-}\StringTok{ }\KeywordTok{unknownMSnSet}\NormalTok{(hl)}
-
-\NormalTok{## Plot a boxplot of the scores of each organelle}
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\FloatTok{10.5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
-\KeywordTok{boxplot}\NormalTok{(svm.scores ~}\StringTok{ }\NormalTok{svm, }\DataTypeTok{data =} \KeywordTok{fData}\NormalTok{(preds), }\DataTypeTok{ylab =} \StringTok{"SVM scores"}\NormalTok{, }\DataTypeTok{las =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+information that is more difficult to defined formally, but that we
+will address in the next section. The classifier (SVM in our case, but
+this is also valid of other classifiers) defines boundaries based on
+the labelled marker proteins. These class/organelle boundaries define
+how non-assigned proteins are classified and with what confidence.
+
+\textbf{TODO} explain how choosing markers positively/negatively
+affects this, and relation with thresholding.
+
+\subsubsection*{Thresholding}\label{sec:thresholding}
+
+It is common when applying a supervised classification algorithm to
+set a specific score cutoff on which to define new assignments, below
+which classifications are set to unknown/unassigned. This is important
+as in a supervised learning setup proteins can only be predicted to be
+localised to one of the sub-cellular niches that appear in the
+labelled training data. We can not guaranette (and do not expect) that
+the whole class diversity to be represented in the labelled training
+data as (1) finding markers that represent the whole diversity of the
+cell is challenging (especially obtaining dual- and multiply-localised
+protein markers) and (2) many sub-cellular niches contain too few
+proteins to train on (we recommend a minimum of 13 markers per
+sub-cellular class for stratified 80/20 partitioning and 5-fold
+cross-validation - this allows a minimum of ~10 examples for parameter
+optimisation on the training partition i.e. 2 per fold for 5-fold
+cross-validation, and then ~3 for testing the best parameters on the
+validation set). <!--Conversely, when defining markers one must also
+be careful to not label too many markers, as this can lead to
+overfitting and mis-assignment.-->
+
+Deciding on a threshold is not trivial as classifier scores are
+heavily dependent upon the classifier used and different sub-cellular
+niches can exhibit different score distributions, as highlighted in
+the boxplot below. We recommend users to set class-specific
+thresholds.  In the code chunk below we display a boxplot of the score
+distributions per organelle.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/thresholds-1.png}
-\caption{plot of chunk thresholds}
+<<thresholds, out.width="\\textwidth">>=
+## First remove the markers
+preds <- unknownMSnSet(hl)
+## Plot a boxplot of the scores of each organelle
+par(oma = c(10.5, 0, 0, 0))
+boxplot(svm.scores ~ svm, data = fData(preds), 
+        ylab = "SVM scores", las = 2)
+@
+  \caption{Visualistion of class-specific classification score distribution.}
+  \label{fig:threshold}
 \end{figure}
@@
 depend on the biological question and experimental design at hand. One
-viable approach in the frame of the above experimetal design would be to
-manually set a FDR, say 5\%, per organelle. To do this the user would
-examine the top scoring predictions for each organelle, and then set a
-threshold at the score at which they achieve 5\% of false assignments
-per organelle. The definintion of a false assignment would depend on the
-information available, for example, validity or lack of validity for the
-localisation from another experiment as reported in the literature or a
-reliable database. If such information is not available, one crude
-method is to set a threshold per organelle by extracting the median or
-3rd quantile score per organelle. For example, in the code chunk below,
-we use the \texttt{orgQuants} function to extract the median organelle
-scores and then pass these scores to the \texttt{getPrediction} function
-to extract the new localisations that meet this scoring criteria. Any
+viable approach in the frame of the above experimetal design would be
+to manually set a FDR, say 5\%, per organelle. To do this the user
+would examine the top scoring predictions for each organelle, and then
+set a threshold at the score at which they achieve 5\% of false
+assignments per organelle.  The definintion of a false assignment
+would depend on the information available, for example, validity or
+lack of validity for the localisation from another experiment as
+reported in the literature or a reliable database.  If such
+information is not available, one crude method is to set a threshold
+per organelle by extracting the median or 3rd quantile score per
+organelle.  For example, in the code chunk below, we use the
+\texttt{orgQuants} function to extract the median organelle scores and
+then pass these scores to the \texttt{getPrediction} function to
+extract the new localisations that meet this scoring criteria. Any
 sub-cellular predictions that fall below the specified thresholds are
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ts <-}\StringTok{ }\KeywordTok{orgQuants}\NormalTok{(hl, }
-                \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }
-                \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{,}
-                \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{, }
-                \DataTypeTok{t =} \NormalTok{.}\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4356698               0.3028843               0.3825420 
-##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6825826               0.5995243               0.4245311 
-##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4034908               0.5882168               0.9501241 
-##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951459               0.7096694               0.3149301 
-##         Plasma membrane              Proteasome 
-##               0.7091667               0.4204304
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{getPredictions}\NormalTok{(hl,}
-                     \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{,}
-                     \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{, }
-                     \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{,}
-                     \DataTypeTok{t =} \NormalTok{ts)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## ans
-##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     169                      86 
-##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     479                      91 
-##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     125                     522 
-##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      38 
-##         Plasma membrane              Proteasome                 unknown 
-##                     325                     159                    2057
-\end{verbatim}
+<<orgquants>>=
+ts <- orgQuants(hl, 
+                fcol = "svm", 
+                scol = "svm.scores",
+                mcol = "markers", 
+                t = .5)
+
+hl <- getPredictions(hl,
+                     fcol = "svm",
+                     scol = "svm.scores", 
+                     mcol = "markers",
+                     t = ts)                        
+@
 
 The output of \texttt{getPredictons} is the original \texttt{MSnSet}
-dataset with a new feature variable appended to the feature data called
-\texttt{fcol.pred} (i.e.~in our case \texttt{svm.pred}) containing the
-prediction results. The results can also be visualied using
-\texttt{plot2D} function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{dev.off}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## null device 
-##           1
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm.pred"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+dataset with a new feature variable appended to the feature data
+called \texttt{fcol.pred} (i.e. in our case \texttt{svm.pred})
+containing the prediction results. The results can also be visualied
+using \texttt{plot2D} function.
+
+<<resetpar>>=
+dev.off()
+@
+
+\begin{figure}[!ht]
     \centering
-\includegraphics{figure/plotres-1.png}
-\caption{plot of chunk plotres}
+<<plotres, out.width="\\textwidth">>=
+plot2D(hl, fcol = "svm.pred")
+@  
+  \caption{Results of the localisation preductions after thresholding.}
+  \label{fig:plotres}
 \end{figure}
 
-There is also a dedicated interactive application to help users examine
-these distributions in the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package. This app can be launched via the \texttt{pRolocVis} function
-and specifying the argument \texttt{app = "classify"} along with the
-relevent \texttt{fcol}, \texttt{scol} and \texttt{mcol} which refer to
-the columns in the feature data that contain the new assignments,
-assignment scores and markers respectively (see also
-\texttt{fvarLabels(svmres)}).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocGUI"}\NormalTok{)}
-\KeywordTok{pRolocVis}\NormalTok{(hl, }
-          \DataTypeTok{app =} \StringTok{"classify"}\NormalTok{,}
-          \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{,}
-            \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{,}
-          \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+There is also a dedicated interactive application to help users
+examine these distributions in the \Biocpkg{pRolocGUI} package.
+This app can be launched via the \texttt{pRolocVis} function and specifying
+the argument \texttt{app = "classify"} along with the relevent \texttt{fcol}, \texttt{scol}
+and \texttt{mcol} which refer to the columns in the feature data that contain
+the new assignments, assignment scores and markers respectively (see
+also \texttt{fvarLabels(svmres)}).
+
+<<classifyApp, eval=FALSE>>=
+library("pRolocGUI")
+pRolocVis(hl, 
+          app = "classify",
+          fcol = "svm",
+	        scol = "svm.scores",
+          mcol = "markers")
+@
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/classify.png}
+  \includegraphics[width=\textwidth]{./Figures/classify.png}
   \caption{The classify application}
+  \label{fig:classifyapp}
 \end{figure}
 
-The data is loaded and displayed on a PCA plot and a boxplot is used to
-display the classifier scores by data class. On the left there is a
+The data is loaded and displayed on a PCA plot and a boxplot is used
+to display the classifier scores by data class. On the left there is a
 sidebar panel with sliders to control the thresholds upon which
 classifications are made. There are two types of cut-off that the user
-can choose from: (1) ``Quantile'' and (2) ``User-defined''. By default,
-when the application is launched quatile scoring is selected and set to
-0.5, the median. The class-specific score thresholds that correspond to
-selecting the desired quantile are shown on as red dots on the boxplot.
-The assignments on the PCA plot are also updated according to the
-selected threshold. The quantile threshold can be set by moving the
-corresponding quantile slider. If one wished to set their own cut-offs
-the ``User-defined'' radio button must be selected and then the sliders
-for defining user-specified scores become active and the scores and
-highlighted on the boxplot by blue dots. For more information we refer
-users to the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
+can choose from: (1) "Quantile" and (2) "User-defined". By default,
+when the application is launched quatile scoring is selected and set
+to 0.5, the median. The class-specific score thresholds that
+correspond to selecting the desired quantile are shown on as red dots
+on the boxplot. The assignments on the PCA plot are also updated
+according to the selected threshold. The quantile threshold can be set
+by moving the corresponding quantile slider. If one wished to set
+their own cut-offs the "User-defined" radio button must be selected
+and then the sliders for defining user-specified scores become active
+and the scores and highlighted on the boxplot by blue dots. For more
+information we refer users to the \Biocpkg{pRolocGUI}
 tutorial
@@
 
-\subsection{Transfer learning}\label{transfer-learning}
-
-In addition to high quality MS-based quantitative proteomics data there
-exist a number of other sources of information that are freely available
-in the public domain that may be useful to assign a protein to its
-sub-cellular niche. For example, imaging from immunofluorescence
-microscopy, protein annotations and sequences, and protein-protein
-interactions among others, represent a rich and vast source of
-complementary information. We can integrate this auxiliary information
-with our primary MS-based quantitative data using a paradigm known as
-transfer learning (TL). The integration of data between different
-technologies is one of the biggest challenges in computational biology
-to date and the
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} package
+\subsection*{Transfer learning}
+
+In addition to high quality MS-based quantitative proteomics data
+there exist a number of other sources of information that are freely
+available in the public domain that may be useful to assign a protein
+to its sub-cellular niche. For example, imaging from
+immunofluorescence microscopy, protein annotations and sequences, and
+protein-protein interactions among others, represent a rich and vast
+source of complementary information. We can integrate this auxiliary
+information with our primary MS-based quantitative data using a
+paradigm known as transfer learning (TL). The integration of data
+between different technologies is one of the biggest challenges in
+computational biology to date and the \Biocpkg{pRoloc} package
 provides functionality to do such analyses. We recently developed two
-transfer learning algorithms using a \emph{k}-NN and SVM framework and
+transfer learning algorithms using a *k*-NN and SVM framework and
 applied them to the task of protein localisation prediction
-{[}@Breckels:2016{]}. In this section we will begin with explaining the
+\cite{Breckels:2016}. In this section we will begin with explaining the
 concept of transfer learning and then show how to apply this in the
@@
 
-In TL one typically has a primary task that they wish to solve, and some
-complementary (often heterogeneous) auxiliary information that is
+In TL one typically has a primary task that they wish to solve, and
+some complementary (often heterogeneous) auxiliary information that is
 related to the primary learning objective, that can be used to help
-solve the primary goal. For example, here our primary task is to assign
-proteins to their sub-cellular niche with high generalisation accuracy
-from data collected from quantitative MS-based experiments. In the
-example below we extract Gene Ontology (GO) information to use as an
-auxiliary data source to help solve our task of protein localisation
-prediction.
-
-Using the functions \texttt{setAnnotationParams} and \texttt{makeGoSet}
-we can contruct an auxiliary \texttt{MSnSet} of GO terms, from the
-primary data's features i.e.~the protein accession numbers. All the GO
-terms associated to each accession number are retrieved and used to
-create a binary matrix where a 1 (0) at position (i; j) indicates that
-term j has (not) been used to annotate protein i. The GO terms are
-retrieved from an appropriate repository using the
-\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}} package.
-The specific
-\emph{\href{http://bioconductor.org/packages/Biomart}{Biomart}}
-repository and query will depend on the species under study and the type
-of identifiers. The first step is to construct the annotation parameters
+solve the primary goal. For example, here our primary task is to
+assign proteins to their sub-cellular niche with high generalisation
+accuracy from data collected from quantitative MS-based experiments.
+In the example below we extract Gene Ontology (GO) information to use
+as an auxiliary data source to help solve our task of protein
+localisation prediction.
+
+Using the functions \texttt{setAnnotationParams} and
+\texttt{makeGoSet} we can contruct an auxiliary \texttt{MSnSet} of GO
+terms, from the primary data's features i.e. the protein accession
+numbers. All the GO terms associated to each accession number are
+retrieved and used to create a binary matrix where a 1 (0) at position
+(i; j) indicates that term j has (not) been used to annotate protein
+i. The GO terms are retrieved from an appropriate repository using the
+\Biocpkg{biomaRt} package. The specific Biomart repository and query
+will depend on the species under study and the type of
+identifiers. The first step is to construct the annotation parameters
 that will enable to perform the query, this is done using
-\texttt{setAnnotationParams}. Typing into the R console
-\texttt{par \textless{}- setAnnotationParams()} will present two menus,
-firstly asking you to identify the species of study, and then what type
-of identifier you have used to annotate the proteins in your
-\texttt{MSnSet}. It is also possible to pass patterns to match the
-species e.g.~in the code chunk below we pass ``Mus musculus'', and the
-identifier type for our data (see \texttt{featureNames(hl)}) which is
-``Uniprot/Swissprot'', for the \texttt{Biocpkg("Biomart")} query.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{par <-}\StringTok{ }\KeywordTok{setAnnotationParams}\NormalTok{(}\DataTypeTok{inputs =} \KeywordTok{c}\NormalTok{(}\StringTok{"Mus musculus"}\NormalTok{, }
-                                      \StringTok{"UniProt/Swissprot"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
+\texttt{setAnnotationParams}. Typing into the R console \texttt{par <-
+  setAnnotationParams()} will present two menus, firstly asking you to
+identify the species of study, and then what type of identifier you
+have used to annotate the proteins in your \texttt{MSnSet}. It is also
+possible to pass patterns to match the species e.g. in the code chunk
+below we pass "Mus musculus", and the identifier type for our data
+(see \texttt{featureNames(hl)}) which is "Uniprot/Swissprot", for the
+Biomart query.
+
+<<setap, eval=FALSE>>=
+par <- setAnnotationParams(inputs = c("Mus musculus", 
+                                      "UniProt/Swissprot"))
+
+@
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gocc <-}\StringTok{ }\KeywordTok{makeGoSet}\NormalTok{(hl, }\DataTypeTok{params =} \NormalTok{par, }
-                  \DataTypeTok{namespace =} \StringTok{"cellular_component"}\NormalTok{, }
-                  \DataTypeTok{evidence =} \OtherTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-The function \texttt{makeGoSet} uses the
-\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}} package
-to query the relevent database (e.g.~Ensembl, Uniprot) for GO terms. All
-GO terms that have been observed for the 5032 proteins in the hyperLOPIT
-dataset are retieved. Users should note that the number of GO terms
-retreived is also dependent on the database version queried and thus is
-always subject to change. We find it is common to see many GO terms with
-only one protein assigned to that term, these such terms generally do
-not bring and information generally for building the classifier and thus
-we can remove such GO terms by using the function
+<<makego, eval=FALSE>>=
+gocc <- makeGoSet(hl, params = par, 
+                  namespace = "cellular_component",
+                  evidence = NULL)
+@
+
+The function \texttt{makeGoSet} uses the \Biocpkg{biomaRt} package to
+query the relevent database (e.g. Ensembl, Uniprot) for GO terms. All
+GO terms that have been observed for the 5032 proteins in the
+hyperLOPIT dataset are retieved. Users should note that the number of
+GO terms retreived is also dependent on the database version queried
+and thus is always subject to change. We find it is common to see many
+GO terms with only one protein assigned to that term, these such terms
+generally do not bring and information generally for building the
+classifier and thus we can remove such GO terms by using the function
 \texttt{filterBinMSnSet}.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gocc <-}\StringTok{ }\KeywordTok{filterBinMSnSet}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-Now we have generated our auxiliary data we can use the \emph{k}-NN
-implementation of transfer learning available in
-\texttt{Biocpkg("pRoloc")} to integrate this with our primary MS-based
-quantitative proteomics data using the functions
-\texttt{knntlOptimisation} to estimate the free-parameters for the
-integration, and \texttt{knntlClassification} to do the predictions. We
-have shown that using transfer learning in the context of spatial
-proteomics results in the assignment of proteins to sub-cellular niches
-with a higher generalisation accuracy than using standard supervised
-machine learning with a single source of information
-{[}@Breckels:2016{]}.
+<<filtergo, eval=FALSE>>=
+gocc <- filterBinMSnSet(hl)
+@
+
+Now we have generated our auxiliary data we can use the \textit{k}-NN
+implementation of transfer learning available in \Biocpkg{pRoloc} to
+integrate this with our primary MS-based quantitative proteomics data
+using the functions \texttt{knntlOptimisation} to estimate the
+free-parameters for the integration, and \texttt{knntlClassification}
+to do the predictions. We have shown that using transfer learning in
+the context of spatial proteomics results in the assignment of
+proteins to sub-cellular niches with a higher generalisation accuracy
+than using standard supervised machine learning with a single source
+of information \cite{Breckels:2016}.
 
 The first step, as with any machine learning algorithm, is to optimise
-any free paramaters of the classifier. For the \emph{k}-NN TL classifier
-there are two sets of parameters that need optimising: the first set are
-the $k$'s for the primary and auxiliary data sources required for the
-nearest neighbour calculations for each data source. The second set of
-parameters (which is noted by a vector of $\theta$ weights) that require
-optimising are the class weights, one per subcellular niche, that
-control the proportion of primary and auxiliary data to use for
-learning. A weight can take any real value number between 0 and 1. A
-weight of $\theta = 1$ indicates that all weight is given to the primary
-data (and this implicitly implies that a weight of $1 - \theta$ is given
-to the auxiliary data, similarly a weight of $\theta = 0$ implies that
-all weight is given to the auxiliary data (so 0 is given to the primary
-source). If we conduct a parameter seach and test weights
-$\theta = {0, 1/3, 2/3, 1}$ for each class, and if we have, for example
-10 subcellular niches, this will result in \texttt{4\^{}10} different
-combinations of parameters to test. The parameter optimisation is
-therefore time consuming and as such we recommend you make use of a
-computing cluster (code and submissing scripts are also available in the
-supporting information). The markers in the \texttt{hl} dataset contain
-14 subcellular classes. If we examine these markers and classes on the
-PCA plot above we can see that in particular the two ribosomes and two
-nuclear compartments are highly separated along the first two
-components, this is also evident from the profiles plot which gives us a
-good indication that these subcellular niches are well-resolved in the
-hyperLOPIT dataset. Transfer learning is particularly useful for classes
-that are not as well separated, we find that subcellular niches that are
+any free paramaters of the classifier. For the *k*-NN TL classifier
+there are two sets of parameters that need optimising: the first set
+are the $k$'s for the primary and auxiliary data sources required for
+the nearest neighbour calculations for each data source. The second
+set of parameters (which is noted by a vector of $\theta$ weights)
+that require optimising are the class weights, one per subcellular
+niche, that control the proportion of primary and auxiliary data to
+use for learning. A weight can take any real value number between 0
+and 1. A weight of $\theta = 1$ indicates that all weight is given to
+the primary data (and this implicitly implies that a weight of
+$1 - \theta$ is given to the auxiliary data, similarly a weight of
+$\theta = 0$ implies that all weight is given to the auxiliary data
+(so 0 is given to the primary source). If we conduct a parameter seach
+and test weights $\theta = {0, 1/3, 2/3, 1}$ for each class, and if we
+have, for example 10 subcellular niches, this will result in
+\texttt{4$^{10}$} different combinations of parameters to test. The
+parameter optimisation is therefore time consuming and as such we
+recommend you make use of a computing cluster (code and submissing
+scripts are also available in the supporting information). The markers
+in the \texttt{hl} dataset contain 14 subcellular classes. If we
+examine these markers and classes on the PCA plot above we can see
+that in particular the two ribosomes and two nuclear compartments are
+highly separated along the first two components, this is also evident
+from the profiles plot which gives us a good indication that these
+subcellular niches are well-resolved in the hyperLOPIT
+dataset. Transfer learning is particularly useful for classes that are
+not as well separated, we find that subcellular niches that are
 well-separated under hyperLOPIT and LOPIT obtain a class score of 1
-(i.e.~use only primary data from transfer learning
-{[}@Breckels:2016{]}). Therefore, for the optimisation stage of the
-analyses we can already infer a subcellular class weight of 1 for these
-niches and only optimise over the remaining organelles. This can
-signifciantly cut down optimisation time as by removing these 4 classes
-from the optimisation (and not the classification) we only have
-\texttt{4\^{}10} class weight combinations to consider instead of
-\texttt{4\^{}14} combinations.
+(i.e. use only primary data from transfer learning
+\cite{Breckels:2016}). Therefore, for the optimisation stage of the
+analyses we can already infer a subcellular class weight of 1 for
+these niches and only optimise over the remaining organelles.  This
+can signifciantly cut down optimisation time as by removing these 4
+classes from the optimisation (and not the classification) we only
+have \texttt{4$^{10}$} class weight combinations to consider instead
+of \texttt{4$^{14}$} combinations.
 
@@
 this is not run live as this the \texttt{hl} dataset with 10 classes,
-707 markers and \texttt{4\^{}10} combinations of parameters takes
-\textasciitilde{}76 hours to run on the University of Cambridge HPC
-using 256 workers).
+707 markers and 4$^{10}$ combinations of parameters takes ~76 hours to
+run on the University of Cambridge HPC using 256 workers).
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## create new markers column for tl markers}
-\KeywordTok{fData}\NormalTok{(hl)$tlmarkers <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hl)$markers}
-\KeywordTok{fData}\NormalTok{(gocc)$tlmarkers <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(gocc)$markers}
-
-\NormalTok{## Remove 4 classes}
-\NormalTok{torm  <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"40S Ribosome"}\NormalTok{, }\StringTok{"60S Ribosome"}\NormalTok{,}
-           \StringTok{"Nucleus - Chromatin"}\NormalTok{, }
-           \StringTok{"Nucleus - Non-chromatin"}\NormalTok{)}
-\NormalTok{for (i in }\KeywordTok{seq}\NormalTok{(torm)) \{}
-  \NormalTok{hl <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(hl, }\DataTypeTok{from =} \NormalTok{torm[i], }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-  \NormalTok{gocc <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(gocc, }\DataTypeTok{from =} \NormalTok{torm[i], }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\NormalTok{\}}
-\KeywordTok{getMarkerClasses}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##  [1] "Actin cytoskeleton"    "Cytosol"              
-##  [3] "Endoplasmic reticulum" "Endosome"             
-##  [5] "Extracellular matrix"  "Lysosome"             
-##  [7] "Mitochondrion"         "Peroxisome"           
-##  [9] "Plasma membrane"       "Proteasome"
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getMarkerClasses}\NormalTok{(gocc, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##  [1] "Actin cytoskeleton"    "Cytosol"              
-##  [3] "Endoplasmic reticulum" "Endosome"             
-##  [5] "Extracellular matrix"  "Lysosome"             
-##  [7] "Mitochondrion"         "Peroxisome"           
-##  [9] "Plasma membrane"       "Proteasome"
-\end{verbatim}
-
-Optimisation stage 1: run \texttt{knnOptimisation} to get the best $k$'s
-for each data source.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## get best k's}
-\NormalTok{kpopt <-}\StringTok{ }\KeywordTok{knnOptimisation}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\NormalTok{kaopt <-}\StringTok{ }\KeywordTok{knnOptimisation}\NormalTok{(gocc, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\KeywordTok{plot}\NormalTok{(kpopt)}
-\KeywordTok{plot}\NormalTok{(kaopt)}
-\end{Highlighting}
-\end{Shaded}
-
-From examining the parameter seach plots we find the best $k$'s for both
-the primary and auxiliary are 3.
+<<loadgocc, include=FALSE>>=
+gofile <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-gocc.rds")
+gocc <- readRDS(gofile)
+@
+
+<<setuptlopt>>=
+## create new markers column for tl markers
+fData(hl)$tlmarkers <- fData(hl)$markers
+fData(gocc)$tlmarkers <- fData(gocc)$markers
+
+## Remove 4 classes
+torm  <- c("40S Ribosome", "60S Ribosome",
+           "Nucleus - Chromatin", 
+           "Nucleus - Non-chromatin")
+for (i in seq(torm)) {
+  hl <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
+  gocc <- fDataToUnknown(gocc, from = torm[i], fcol = "tlmarkers")
+}
+getMarkerClasses(hl, fcol = "tlmarkers")
+getMarkerClasses(gocc, fcol = "tlmarkers")
+@
+
+Optimisation stage 1: run \texttt{knnOptimisation} to get the best
+$k$'s for each data source.
+
+<<knnoptfortl, eval=FALSE>>=
+## get best k's
+kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
+kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
+plot(kpopt)
+plot(kaopt)
+@
+
+From examining the parameter seach plots we find the best $k$'s for
+both the primary and auxiliary are 3.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Set appropriate parallelisation backend and  }
-\NormalTok{## number of workers for the tl}
-\NormalTok{par <-}\StringTok{ }\KeywordTok{SnowParam}\NormalTok{(255L, }\DataTypeTok{type =} \StringTok{"MPI"}\NormalTok{)}
-
-\NormalTok{## Now peform tl optimisation }
-\NormalTok{tlopt <-}\StringTok{ }\KeywordTok{knntlOptimisation}\NormalTok{(hl, gocc,}
-                           \DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{,}
-                           \DataTypeTok{length.out =} \DecValTok{4}\NormalTok{, }
-                           \DataTypeTok{times =} \DecValTok{50}\NormalTok{,}
-                           \DataTypeTok{xval =} \DecValTok{5}\NormalTok{, }\DataTypeTok{k =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{),}
-                           \DataTypeTok{BPPARAM =} \NormalTok{par}
-  \NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+<<runtlopts, eval=FALSE>>=
+## Set appropriate parallelisation backend and  
+## number of workers for the tl
+par <- SnowParam(255L, type = "MPI")
+
+## Now peform tl optimisation 
+tlopt <- knntlOptimisation(hl, gocc,
+                           fcol = "tlmarkers",
+                           length.out = 4,
+                           times = 50,
+                           xval = 5, k = c(3, 3),
+                           BPPARAM = par)
+@
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot}\NormalTok{(tlopt)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+<<loadtlres, echo=FALSE>>=
+tlfile <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-tlopt.rds")
+tlopt <- readRDS(tlfile)
+colnames(tlopt@results)[4] <- "Endoplasmic reticulum"
+setStockcol(NULL)
+@
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plottl-1.png}
-\caption{plot of chunk plottl}
+<<plottl, out.width="\\textwidth">>=
+plot(tlopt)
+@  
+  \caption{Visualisation of the tTransfer learning parameter optimisation procedure.}
+  \label{fig:plottl}
 \end{figure}
@@
 over the 50 runs we find that for many of the subcellular niches a
-weight of 1 is most popular (i.e.~use only primary hyperLOPIT data in
+weight of 1 is most popular (i.e. use only primary hyperLOPIT data in
 classification), this is unsuprising as we already know the dataset is
@@
 for the proteasome and lysosome tend to be towards 0, indicating that
-these niches are well-resolved in the Gene Ontology. This tells us that
-we would benefit from including auxiliary GO information in our
+these niches are well-resolved in the Gene Ontology. This tells us
+that we would benefit from including auxiliary GO information in our
 classifier for these subcellular compartments. The plasma membrane
@@
 One of the benefits of the algorithm is the ability to manually select
-weights for each class. In the optimisation above, for time constraints,
-we removed the two ribosomal subunits and the two nulcear compartments,
-and therefore in the code chunk below when we extract the best
-parameters, these subcellular niches are not included. To include these
-4 subcellular niches in the next classification step we must include
-them in the parameters so we infer a weight of 1 for each of these
-niches as we know they are well resolved in hyperLOPIT. We then re-order
-the weights according to \texttt{getMarkerClasses} and perform the
-classification using the function \texttt{knntlClassification}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## best parameters for the 10 classes}
-\NormalTok{(bestpar <-}\StringTok{ }\KeywordTok{getParams}\NormalTok{(tlopt))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##    Actin cytoskeleton               Cytosol Endoplasmic reticulum 
-##             1.0000000             1.0000000             0.6666667 
-##              Endosome  Extracellular matrix              Lysosome 
-##             1.0000000             1.0000000             0.3333333 
-##         Mitochondrion            Peroxisome       Plasma membrane 
-##             1.0000000             1.0000000             0.6666667 
-##            Proteasome 
-##             0.3333333
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## add weights for classes not included in the optimisation}
-\NormalTok{otherweights <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{)}
-\KeywordTok{names}\NormalTok{(otherweights) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"40S Ribosome"}\NormalTok{, }\StringTok{"60S Ribosome"}\NormalTok{, }
-                         \StringTok{"Nucleus - Chromatin"}\NormalTok{, }
-                         \StringTok{"Nucleus - Non-chromatin"}\NormalTok{)}
-\NormalTok{(bestpar <-}\StringTok{ }\KeywordTok{c}\NormalTok{(bestpar, otherweights))}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-##      Actin cytoskeleton                 Cytosol   Endoplasmic reticulum 
-##               1.0000000               1.0000000               0.6666667 
-##                Endosome    Extracellular matrix                Lysosome 
-##               1.0000000               1.0000000               0.3333333 
-##           Mitochondrion              Peroxisome         Plasma membrane 
-##               1.0000000               1.0000000               0.6666667 
-##              Proteasome            40S Ribosome            60S Ribosome 
-##               0.3333333               1.0000000               1.0000000 
-##     Nucleus - Chromatin Nucleus - Non-chromatin 
-##               1.0000000               1.0000000
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## re-order classes }
-\NormalTok{bestpar <-}\StringTok{ }\NormalTok{bestpar[}\KeywordTok{getMarkerClasses}\NormalTok{(hl)]}
-
-\NormalTok{## Do the classification}
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{knntlClassification}\NormalTok{(hl, gocc, }\DataTypeTok{bestTheta =} \NormalTok{bestpar, }
-                          \DataTypeTok{k =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
+weights for each class. In the optimisation above, for time
+constraints, we removed the two ribosomal subunits and the two nulcear
+compartments, and therefore in the code chunk below when we extract
+the best parameters, these subcellular niches are not included. To
+include these 4 subcellular niches in the next classification step we
+must include them in the parameters so we infer a weight of 1 for each
+of these niches as we know they are well resolved in hyperLOPIT. We
+then re-order the weights according to \texttt{getMarkerClasses} and
+perform the classification using the function
+\texttt{knntlClassification}.
+
+<<tlclassify, cache = cache>>=
+## best parameters for the 10 classes
+(bestpar <- getParams(tlopt))
+
+## add weights for classes not included in the optimisation
+otherweights <- rep(1, 4)
+names(otherweights) <- c("40S Ribosome", "60S Ribosome", 
+                         "Nucleus - Chromatin", 
+                         "Nucleus - Non-chromatin")
+(bestpar <- c(bestpar, otherweights))
+
+## re-order classes 
+bestpar <- bestpar[getMarkerClasses(hl)]
+
+## Do the classification
+hl <- knntlClassification(hl, gocc, bestTheta = bestpar, 
+                          k = c(3, 3))
+@
 
@@
 \texttt{plot2D}, scores assessed and cutoffs calculated using the
-\texttt{classify} app in \texttt{pRolocVis}, predictions obtained using
-\texttt{getPredictions} in the same way as demonstrated above for the
-SVM classifier.
-
-\subsection{Unsupervised machine
-learning}\label{unsupervised-machine-learning}
+\texttt{classify} app in \texttt{pRolocVis}, predictions obtained
+using \texttt{getPredictions} in the same way as demonstrated above
+for the SVM classifier.
+
+\subsection*{Unsupervised machine learning}
 
@@
 learning methods. In unsupervised learning, the training data consists
-of a set of input vectors e.g.~protein profiles, for which we do not
-have information about the class label e.g.~localisation. The main goal
-in unsupervised learning is to uncover groups of similar examples within
-the data, this is termed clustering. Ordinance methods such as principal
-components analysis (PCA) also fall into the category of unsupervised
-learning methods, where the data can be projected from a
+of a set of input vectors e.g. protein profiles, for which we do not
+have information about the class label e.g. localisation. The main
+goal in unsupervised learning is to uncover groups of similar examples
+within the data, this is termed clustering. Ordinance methods such as
+principal components analysis (PCA) also fall into the category of
+unsupervised learning methods, where the data can be projected from a
 high-dimensional spcae down to two or three dimensions for the purpose
@@
 As described and demonstrated already above, PCA is a valuable and
-powerful method for data visualisation and quality control. We do not do
-any unsupervised clustering in the frame of the above spatial proteomics
-workflow as although clustering can be useful for grouping labelled data
-into categories e.g.~see the \texttt{mrkHClust} function, we do not find
-it adaquete for spatial proteomics data analysis. We find supervised
-learning more suited to the task of protein localisation prediction in
-which we use high-quality curated marker proteins to build a classifier,
-instead of using an entirely unsupervised approach to look for clusters
-and then look for enrichment of organelles and complexes. In the latter
-we do not make good use of valuable prior knowledge, and in our
-experience unsupervised clustering can be extremely difficult due poor
-estimates of the number of clusters that may appear in the data.
-
-\section{Writing and exporting data}\label{writing-and-exporting-data}
-
-A \texttt{MSnSet} can be exported from R using the \texttt{write.exprs}
-function. This function writes the expression values to a tab (the
-defualt, or other type e.g.e.g.~comma, as specified in
-\texttt{write,table}) separated file. There argument \texttt{fDataCols}
-can be used to specify which \texttt{featureData} columns (as column
-names, column number or \texttt{logical}) to append to the right of the
-expression matrix.
-
-In the below code chunk we write the \texttt{hl} object to a csv file.
-The \texttt{file} argument is used to specify the file path, the
-\texttt{sep} argument specifies the field separator string, here we use
-a comma, finally as we want to write all the information in the
+powerful method for data visualisation and quality control. We do not
+do any unsupervised clustering in the frame of the above spatial
+proteomics workflow as although clustering can be useful for grouping
+labelled data into categories e.g. see the \texttt{mrkHClust}
+function, we do not find it adaquete for spatial proteomics data
+analysis. We find supervised learning more suited to the task of
+protein localisation prediction in which we use high-quality curated
+marker proteins to build a classifier, instead of using an entirely
+unsupervised approach to look for clusters and then look for
+enrichment of organelles and complexes. In the latter we do not make
+good use of valuable prior knowledge, and in our experience
+unsupervised clustering can be extremely difficult due poor estimates
+of the number of clusters that may appear in the data.
+
+\section*{Writing and exporting data}
+
+A \texttt{MSnSet} can be exported from R using the
+\texttt{write.exprs} function.  This function writes the expression
+values to a tab (the defualt, or other type e.g.e.g. comma, as
+specified in \texttt{write,table}) separated file. There argument
+\texttt{fDataCols} can be used to specify which \texttt{featureData}
+columns (as column names, column number or \texttt{logical}) to append
+to the right of the expression matrix.
+
+In the below code chunk we write the \texttt{hl} object to a csv
+file. The \texttt{file} argument is used to specify the file path, the
+\texttt{sep} argument specifies the field separator string, here we
+use a comma, finally as we want to write all the information in the
 \texttt{featureData} to the file, as well as the expression data, we
-specify \texttt{fDataCols = 1:ncol(hl)} i.e.~write everything in the
+specify \texttt{fDataCols = 1:ncol(hl)} i.e. write everything in the
 \texttt{featureData} to the file \texttt{"hl.csv"}.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{write.exprs}\NormalTok{(hl, }\DataTypeTok{file =} \StringTok{"hl.csv"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{","}\NormalTok{, }
-            \DataTypeTok{fDataCols =} \DecValTok{1}\NormalTok{:}\KeywordTok{ncol}\NormalTok{(hl))}
-\end{Highlighting}
-\end{Shaded}
+<<writeData, eval = FALSE>>=
+write.exprs(hl, file = "hl.csv", sep = ",", 
+            fDataCols = 1:ncol(hl))
+@
 
 Individual R objects, such a parameters from the machine learning
-optimisations, can be save using the standard \texttt{save} function in
-R. For example, to save and then re-load the parameters from the SVM
-optimisation,
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## To save the parameters as an R object}
-\KeywordTok{save}\NormalTok{(params, }\DataTypeTok{file =} \StringTok{"svmparams.rda"}\NormalTok{)}
-
-\NormalTok{## To re-load after saving}
-\KeywordTok{load}\NormalTok{(}\DataTypeTok{file =} \StringTok{"svmparams.rda"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\section{Session information}\label{session-information}
-
-The function \texttt{sessionInfo} provides a summary of all packages and
-versions used to generate this document. This enables us to record the
-exact state of our session that lead to these exact results. Conversely,
-if the script stops working of if it returns different results, we are
-in a position to re-generate the original results using the adequate
-software versions and retrace changes in the software that lead to
-failure and/or different results.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{sessionInfo}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## R version 3.3.1 Patched (2016-08-02 r71022)
-## Platform: x86_64-pc-linux-gnu (64-bit)
-## Running under: Ubuntu 14.04.5 LTS
-## 
-## locale:
-##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
-##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
-##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
-##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
-##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
-## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
-## 
-## attached base packages:
-## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
-## [8] datasets  base     
-## 
-## other attached packages:
-##  [1] pRolocdata_1.11.9    pRoloc_1.13.15       MLInterfaces_1.53.1 
-##  [4] cluster_2.0.4        annotate_1.51.1      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.35.4 IRanges_2.7.15       S4Vectors_0.11.17   
-## [10] MSnbase_1.99.2       ProtGenerics_1.5.1   BiocParallel_1.7.8  
-## [13] mzR_2.7.6            Rcpp_0.12.7          Biobase_2.33.3      
-## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.32    
-## [19] knitr_1.14          
-## 
-## loaded via a namespace (and not attached):
-##  [1] minqa_1.2.4           colorspace_1.2-6      hwriter_1.3.2        
-##  [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
-##  [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-16         
-## [10] MatrixModels_0.4-1    affyio_1.43.0         flexmix_2.3-13       
-## [13] mvtnorm_1.0-5         codetools_0.2-14      splines_3.3.1        
-## [16] doParallel_1.0.10     impute_1.47.0         robustbase_0.92-6    
-## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-71         
-## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-24       
-## [25] vsn_3.41.0            sfsmisc_1.1-0         shiny_0.14           
-## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
-## [31] limma_3.29.21         formatR_1.4           htmltools_0.3.5      
-## [34] quantreg_5.29         tools_3.3.1           ggvis_0.4.3          
-## [37] gtable_0.2.0          affy_1.51.1           reshape2_1.4.1       
-## [40] dplyr_0.5.0           MALDIquant_1.15       trimcluster_0.1-2    
-## [43] gdata_2.17.0          preprocessCore_1.35.0 nlme_3.1-128         
-## [46] iterators_1.0.8       fpc_2.1-10            stringr_1.1.0        
-## [49] lme4_1.1-12           lpSolve_5.6.13        mime_0.5             
-## [52] gtools_3.5.0          dendextend_1.3.0      DEoptimR_1.0-6       
-## [55] zlibbioc_1.19.0       MASS_7.3-45           scales_0.4.0         
-## [58] BiocInstaller_1.23.9  pcaMethods_1.65.0     SparseM_1.72         
-## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.29.2       
-## [64] rpart_4.1-10          stringi_1.1.1         RSQLite_1.0.0        
-## [67] genefilter_1.55.2     randomForest_4.6-12   foreach_1.4.3        
-## [70] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
-## [73] mzID_1.11.2           evaluate_0.9          lattice_0.20-34      
-## [76] htmlwidgets_0.7       gbm_2.1.1             plyr_1.8.4           
-## [79] magrittr_1.5          R6_2.1.3              DBI_0.5-1            
-## [82] whisker_0.3-2         mgcv_1.8-15           survival_2.39-5      
-## [85] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
-## [88] msdata_0.12.1         car_2.1-3             mlbench_2.1-1        
-## [91] grid_3.3.1            FNN_1.1               threejs_0.2.2        
-## [94] digest_0.6.10         diptest_0.75-7        xtable_1.8-2         
-## [97] httpuv_1.3.3          munsell_0.4.3
-\end{verbatim}
-
-It is always important to include session information details along with
-a \href{http://adv-r.had.co.nz/Reproducibility.html}{short reproducible
-example} highlighting the problem or
+optimisations, can be save using the standard \texttt{save} function
+in R.  For example, to save and then re-load the parameters from the
+SVM optimisation,
+
+<<saveData, eval = FALSE>>=
+## To save the parameters as an R object
+save(params, file = "svmparams.rda")
+
+## To re-load after saving
+load(file = "svmparams.rda")
+@
+
+\section*{Session information}
+
+The function \texttt{sessionInfo} provides a summary of all packages
+and versions used to generate this document. This enables us to record
+the exact state of our session that lead to these exact
+results. Conversely, if the script stops working of if it returns
+different results, we are in a position to re-generate the original
+results using the adequate software versions and retrace changes in
+the software that lead to failure and/or different results.
+
+<<si>>=
+sessionInfo()
+@
+
+It is always important to include session information details along
+with a \href{http://adv-r.had.co.nz/Reproducibility.html}{short
+  reproducible example} highlighting the problem or
 \href{https://support.bioconductor.org/}{question} at hand.

### 2016-09-29 bioc-workflow.Rnw
add final sections

@@
   \centering
-  \includegraphics[width=.45\textwidth]{./Figures/msnset.png}
+  \includegraphics[width=.5\textwidth]{./Figures/msnset.png}
   \caption{Simplified representation of the \texttt{MSnSet} data
@@
   \centering
-<<namap, out.width="\\textwidth">>=
+<<namap, out.width=".65\\textwidth">>=
 image2(is.na(combined), col = c("black", "white"),
@@
   \centering
-<<qcplot, message=FALSE, warning=FALSE, out.width="\\textwidth">>=
+<<qcplot, message=FALSE, warning=FALSE, out.width=".65\\textwidth">>=
 library("pRoloc")
@@
   \centering
-<<qcbx, out.width="\\textwidth">>=
+<<qcbx, out.width="\\textwidth", fig.asp=1/2>>=
 par(mfrow = c(1, 2))
@@
   \centering
-<<plotmarkers, out.width="\\textwidth">>=
+<<plotmarkers, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>=
 par(mfrow = c(1, 2))
@@
   \centering
-<<plotDist, out.width="\\textwidth">>=
+<<plotDist, out.width=".65\\textwidth">>=
 hlo <- hl[, order(hl$Iodixonal.Density)]
@@
   \centering
-<<plot2Drep, out.width="\\textwidth">>= 
+<<plot2Drep, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>= 
 par(mfrow = c(1, 2)) 
@@
   \centering
-<<plotPDres, out.width="\\textwidth">>=
+<<plotPDres, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>=
 ## Re-order the colours for the phenoDisco output
@@
   \centering
-<<visualiseOptHide, echo=FALSE, out.width="\\textwidth">>=
+<<visualiseOptHide, echo=FALSE, out.width=".65\\textwidth">>=
 library('gridExtra')
@@
   \centering
-<<plotSVM, out.width="\\textwidth">>=
+<<plotSVM, out.width=".8\\textwidth">>=
 ## set point size of each protein to be inversely proportional to the
@@
   \centering
-<<thresholds, out.width="\\textwidth">>=
+<<thresholds, out.width=".8\\textwidth">>=
 ## First remove the markers
@@
     \centering
-<<plotres, out.width="\\textwidth">>=
+<<plotres, out.width=".8\\textwidth">>=
 plot2D(hl, fcol = "svm.pred")
@@
   \centering
-<<plottl, out.width="\\textwidth">>=
+<<plottl, out.width=".6\\textwidth">>=
 plot(tlopt)
 @  
-  \caption{Visualisation of the tTransfer learning parameter optimisation procedure.}
+  \caption{Visualisation of the transfer learning parameter optimisation procedure.}
   \label{fig:plottl}
@@
 \href{https://support.bioconductor.org/}{question} at hand.
+
+\bigskip
+
+\subsection*{Author contributions}
+
+\textbf{TODO} In order to give appropriate credit to each author of an
+article, the individual contributions of each author to the manuscript
+should be detailed in this section. We recommend using author initials
+and then stating briefly how they contributed.
+
+\subsection*{Competing interests}
+
+No conflict of interest.
+
+\subsection*{Grant information}
+
+L.M.B is supported by a Wellcome Trust Technology Development Grant
+(grant number 108467/Z/15/Z). L.G. is supported by the BBSRC Strategic
+Longer and Larger grant (Award BB/L002817/1).
+
+\subsection*{Acknowledgements}
+
+The authors would like to thank the Cambridge Computational Biology
+Institute and the Centre for Mathematical Sciences for hosting them
+when writing this article. The authors would like to thank Dr Claire
+M. Mulvey for helpful comments on early versions of this
+manuscript. The authors would also like to thank Dr. Stuart Rankin
+from the High Performance Computing Service for his support. Part of
+this work was performed using the Darwin Supercomputer of the
+University of Cambridge High Performance Computing Service, provided
+by Dell Inc. using Strategic Research Infrastructure Funding from the
+Higher Education Funding Council for England and funding from the
+Science and Technology Facilities Council.

### 2016-09-29 bioc-workflow.tex
add final sections

--- 2016-09-29 bioc-workflow.tex
@@
-\section{Introduction}\label{introduction}
+\section*{Introduction}
 
@@
 structured and annotated computational object from the 
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
-experiment package. In this workflow, we will start with a description
-of how to generate these ad hoc objects starting from an arbitrary
-spreadsheets, as produced by many popular third-party applications.
-
-While we focus here on a LOPIT-type dataset, these analyses are relevant
-for any quantitative spatial proteomics data, irrespective of the
-fractionation or quantitation (i.e.~labelled or label-free) methods.
-
-\subsection{The infrastructure:
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} and
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}} in
-Bioconductor}\label{the-infrastructure-proloc-and-msnbase-in-bioconductor}
-
-To make use of the full functionality of the
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} software
-one needs to import their data into R and prepare them as an
-\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for
-the efficient manipulation and processing of mass spectrometry and
-proteomics data in R. Figure 1 illustrates a simplified view of the
-\texttt{MSnSet} structure; there exists 3 key sub-parts (termed slots)
-to such a data object: (1) the \texttt{exprs} slot for storing the
-quantitation data, (2) the \texttt{fData} slot for storing the feature
-meta-data, and finally (3) the \texttt{pData} slot for storing the
-sample meta-data.
-
-\begin{figure}[htbp]
+\Biocexptpkg{pRolocdata} experiment package. In this workflow, we
+will start with a description of how to generate these ad hoc objects
+starting from an arbitrary spreadsheets, as produced by many popular
+third-party applications.
+
+While we focus here on a LOPIT-type dataset, these analyses are
+relevant for any quantitative spatial proteomics data, irrespective of
+the fractionation or quantitation (i.e. labelled or label-free)
+methods.
+
+\subsection*{The infrastructure: \texttt{pRoloc} and \texttt{MSnbase} packages}
+
+To make use of the full functionality of the \Biocpkg{pRoloc}
+software one needs to import their data into R and prepare them as an
+\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for the efficient
+manipulation and processing of mass spectrometry and proteomics data
+in R. Figure 1 illustrates a simplified view of the \texttt{MSnSet}
+structure; there exists 3 key sub-parts (termed slots) to such a data
+object: (1) the \texttt{exprs} slot for storing the quantitation data, (2)
+the \texttt{fData} slot for storing the feature meta-data, and finally (3)
+the \texttt{pData} slot for storing the sample meta-data.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/msnset.png}
-\caption{Simplified representation of the \texttt{MSnSet} data structure
-(reproduced with permission from the
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
+  \includegraphics[width=.5\textwidth]{./Figures/msnset.png}
+  \caption{Simplified representation of the \texttt{MSnSet} data
+    structure (reproduced with permission from the \Biocpkg{MSnbase}
     vignette)}
+  \label{fig:msnset}
 \end{figure}
 
-\subsection{Importing data}\label{importing-data}
+\subsection*{Importing data}
 
@@
 \texttt{MSnSet} instance. All methods are described in the
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
+\Biocpkg{MSnbase}
 \href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf}{input/output
   capabilities vignette}. One suggested simple method is to use the
-function \texttt{readMSnSet2} in
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}. The
-function takes a single spreadsheet as input and extracts the columns
+function \texttt{readMSnSet2} in \Biocpkg{MSnbase}. The function
+takes a single spreadsheet as input and extracts the columns
 containing the quantitation data, as identified by the argument
-\texttt{ecol}, to create the expression data, while the other columns in
-the spreadsheet are appended to the feature meta-data slot. By example,
-in the code chunk below we read in the \texttt{csv} spreadsheet
-containing the quantitation data from the intersect of replicates 1 and
-2 of the mouse map {[}@hyper{]}, using the \texttt{readMSnSet2}
-function. The data is as available online with the manuscript (see tab 2
-of the \texttt{xlsx} supplementary data set 1 in {[}@hyper{]}, which
-should be exported as a text-based spreadsheet). It is also available as
-a \texttt{csv} in the Bioconductor \texttt{r Biocexptpkg("pRolocdata")}
-data package.
-
-To use the \texttt{readMSnSet2} function, as a minimum one must specify
-the file path to the data and which columns of the spreadsheet contain
+\texttt{ecol}, to create the expression data, while the other columns
+in the spreadsheet are appended to the feature meta-data slot.  By
+example, in the code chunk below we read in the \texttt{csv}
+spreadsheet containing the quantitation data from the intersect of
+replicates 1 and 2 of the mouse map \cite{hyper}, using the
+\texttt{readMSnSet2} function. The data is as available online with
+the manuscript (see tab 2 of the \texttt{xlsx} supplementary data set
+1 in \cite{hyper}, which should be exported as a text-based
+spreadsheet). It is also available as a \texttt{csv} in the
+Bioconductor \Biocexptpkg{pRolocdata} data package.
+
+To use the \texttt{readMSnSet2} function, as a minimum one must specify the
+file path to the data and which columns of the spreadsheet contain
 quantitation data. The \texttt{getEcols} function exists to help users
-identify which columns of the spreadsheet contain the quantitation data.
-In the last line of the code chunk below, we print the file name (not
-the full path, which will vary from computer to computer).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"MSnbase"}\NormalTok{)}
-\NormalTok{extdatadir <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\DataTypeTok{package =} \StringTok{"pRolocdata"}\NormalTok{)}
-\NormalTok{csvfile <-}\StringTok{ }\KeywordTok{dir}\NormalTok{(extdatadir, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{,}
-          \DataTypeTok{pattern =} \StringTok{"hyperLOPIT-SIData-ms3-rep12-intersect.csv"}\NormalTok{)}
-\KeywordTok{basename}\NormalTok{(csvfile)}
-\end{Highlighting}
-\end{Shaded}
-
+identify which columns of the spreadsheet contain the
+quantitation data. In the last line of the code chunk below, we print
+the file name (not the full path, which will vary from computer to
+computer).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"MSnbase"}\hlstd{)}
+\hlstd{extdatadir} \hlkwb{<-} \hlkwd{system.file}\hlstd{(}\hlstr{"extdata"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"pRolocdata"}\hlstd{)}
+\hlstd{csvfile} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+          \hlkwc{pattern} \hlstd{=} \hlstr{"hyperLOPIT-SIData-ms3-rep12-intersect.csv"}\hlstd{)}
+\hlkwd{basename}\hlstd{(csvfile)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 Note that the file is compressed (as indicated by the \texttt{gz}, for
-\texttt{gzip}, extension), and will be decompressed on-the-fly when read
-into R later on.
-
-The spreadsheet that was deposited by the authors contains two headers,
-with the second header containing information about where the
+\texttt{gzip}, extension), and will be decompressed on-the-fly when read into
+R later on.
+
+The spreadsheet that was deposited by the authors contains two
+headers, with the second header containing information about where the
 quantitation data is stored.
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/spreadsheet-screenshot.png}
+  \includegraphics[width=.65\textwidth]{./Figures/spreadsheet-screenshot.png}
   \caption{A screenshot of the data in the spreadsheet.}
+  \label{fig:spreadsheet}
 \end{figure}
@@
 We can display the names of the second header by calling the
-\texttt{getEcols} function with the argument \texttt{n = 2} (the default
-value is \texttt{n = 1}), to specify that we wish to display the column
-names of the second line.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getEcols}\NormalTok{(csvfile, }\DataTypeTok{split =} \StringTok{","}\NormalTok{, }\DataTypeTok{n =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
+\texttt{getEcols} function with the argument \texttt{n = 2} (the default value is \texttt{n
+= 1}), to specify that we wish to display the column names of the
+second line.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{getEcols}\hlstd{(csvfile,} \hlkwc{split} \hlstd{=} \hlstr{","}\hlstd{,} \hlkwc{n} \hlstd{=} \hlnum{2}\hlstd{)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 It is now easy for one to identify that the quantitation data,
-corresponding to the 10 TMT isobaric tags, is located in columns 8 to
-27. We now have the two mandatory arguments to \texttt{readMSnSet2},
+corresponding to the 10 TMT isobaric tags, is located in columns 8
+to 27. We now have the two mandatory arguments to \texttt{readMSnSet2},
 namely the file name (stored in the \texttt{csvfile} variable) and the
 quantitation column indices. In addition to these, it is also possible
-to pass the optional argument \texttt{fnames} to indicate which column
-to use as the labels by which to identify each protein in the sample.
-Here, we use \texttt{fnames = 1} to use the UniProt identifiers
-contained in the first (unnamed) column of the spreadsheet. We also need
-to specify to skip the first line of the file (for the same reason that
-we used \texttt{n = 2} in \texttt{getEcols} above) to read the
-\texttt{csv} data and convert it to an \texttt{MSnSet} object, named
-\texttt{hl} (for hyperLOPIT).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{readMSnSet2}\NormalTok{(csvfile, }\DataTypeTok{ecol =} \DecValTok{8}\NormalTok{:}\DecValTok{27}\NormalTok{, }\DataTypeTok{fnames =} \DecValTok{1}\NormalTok{, }\DataTypeTok{skip =} \DecValTok{1}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-Below, we display a short summary of the data. The data contains 5032
-proteins/features common across the 2 biological replicates for the
-respective 2 x 10-plex reporter tags (20 columns/samples), along with
-associated feature meta-data such as protein markers, protein
-description, number of quantified peptides etc (see below).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl}
-\end{Highlighting}
-\end{Shaded}
-
+to pass the optional argument \texttt{fnames} to indicate which column to use
+as the labels by which to identify each protein in the sample. Here,
+we use \texttt{fnames = 1} to use the UniProt identifiers contained in the
+first (unnamed) column of the spreadsheet. We also need to specify to
+skip the first line of the file (for the same reason that we used 
+\texttt{n = 2} in \texttt{getEcols} above) to read the \texttt{csv} data and convert it to an
+\texttt{MSnSet} object, named \texttt{hl} (for hyperLOPIT).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{readMSnSet2}\hlstd{(csvfile,} \hlkwc{ecol} \hlstd{=} \hlnum{8}\hlopt{:}\hlnum{27}\hlstd{,} \hlkwc{fnames} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{skip} \hlstd{=} \hlnum{1}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Below, we display a short summary of the data. The data contains 
+5032 proteins/features common across the 2 biological replicates
+for the respective 2 x 10-plex reporter tags (20
+columns/samples), along with associated feature meta-data such as
+protein markers, protein description, number of quantified peptides
+etc (see below).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-Below, we examine the quantitative information for first 5 proteins. It
-is also possible to access specific rows and columns by naming the
+\end{kframe}
+\end{knitrout}
+
+Below, we examine the quantitative information for first 5 proteins.
+It is also possible to access specific rows and columns by naming the
 proteins and TMT tag channels of interest.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{exprs}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{5}\NormalTok{, ]}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{exprs}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{5}\hlstd{, ]}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{exprs}\NormalTok{(hl)[}\KeywordTok{c}\NormalTok{(}\StringTok{"Q9ERU9"}\NormalTok{, }\StringTok{"Q9Z2R6"}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\StringTok{"X126"}\NormalTok{, }\StringTok{"X131.1"}\NormalTok{)]}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{exprs}\hlstd{(hl)[}\hlkwd{c}\hlstd{(}\hlstr{"Q9ERU9"}\hlstd{,} \hlstr{"Q9Z2R6"}\hlstd{),} \hlkwd{c}\hlstd{(}\hlstr{"X126"}\hlstd{,} \hlstr{"X131.1"}\hlstd{)]}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 The feature meta-data is stored in the \texttt{fData} slot and can be
-accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2},
-automatically, everything that is not defined as quantitation data by
-\texttt{ecol} or the feature names by \texttt{fnames} is deposited to
-the \texttt{fData} slot.
-
-We see the \texttt{fData} contains 25 columns describing information
-such as the number of peptides, associated markers, machine learning
-results etc. To identify the feature variable names we can use the
-function \texttt{fvarLabels}. We see that the first 6 feature variable
-names contain non-discriminatory label names, so we relabel them to help
-us identify what feature data information is stored in the associated
-columns.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
+accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2}, automatically,
+everything that is not defined as quantitation data by \texttt{ecol} or the
+feature names by \texttt{fnames} is deposited to the \texttt{fData} slot.
+
+We see the \texttt{fData} contains 25 columns describing information such as
+the number of peptides, associated markers, machine learning results
+etc. To identify the feature variable names we can use the function
+\texttt{fvarLabels}. We see that the first 6 feature variable names contain
+non-discriminatory label names, so we relabel them to help us identify
+what feature data information is stored in the associated columns.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hl)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{3}\NormalTok{] <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"uniprot.accession"}\NormalTok{, }\StringTok{"uniprot.id"}\NormalTok{, }\StringTok{"description"}\NormalTok{)}
-\KeywordTok{fvarLabels}\NormalTok{(hl)[}\DecValTok{4}\NormalTok{:}\DecValTok{6}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\StringTok{"peptides.expt"}\NormalTok{, }\DecValTok{1}\NormalTok{:}\DecValTok{3}\NormalTok{)}
-\KeywordTok{fData}\NormalTok{(hl)[}\DecValTok{1}\NormalTok{:}\DecValTok{4}\NormalTok{, }\DecValTok{1}\NormalTok{:}\DecValTok{6}\NormalTok{]}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{3}\hlstd{]} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"uniprot.accession"}\hlstd{,} \hlstr{"uniprot.id"}\hlstd{,} \hlstr{"description"}\hlstd{)}
+\hlkwd{fvarLabels}\hlstd{(hl)[}\hlnum{4}\hlopt{:}\hlnum{6}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlstr{"peptides.expt"}\hlstd{,} \hlnum{1}\hlopt{:}\hlnum{3}\hlstd{)}
+\hlkwd{fData}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{4}\hlstd{,} \hlnum{1}\hlopt{:}\hlnum{6}\hlstd{]}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-Note that when using the simple \texttt{readMSnSet2} procedure, the
-\texttt{pData} slot which is used to store information about the
-samples/channels is kept empty. It is advised to annotate the channels
-as well. Below, we annotate the replicate from which the profiles
+\end{kframe}
+\end{knitrout}
+
+Note that when using the simple \texttt{readMSnSet2} procedure, the \texttt{pData}
+slot which is used to store information about the samples/channels
+is kept empty. It is advised to annotate the channels as
+well. Below, we annotate the replicate from which the profiles
 originate and the TMT tag (extracted from the sample/channel names).
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{pData}\NormalTok{(hl)$Replicate <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{:}\DecValTok{2}\NormalTok{, }\DataTypeTok{each =} \DecValTok{10}\NormalTok{)}
-\KeywordTok{pData}\NormalTok{(hl)$Tag <-}\StringTok{ }\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}\textbackslash{}}\StringTok{.1$"}\NormalTok{, }\StringTok{""}\NormalTok{, }\KeywordTok{sub}\NormalTok{(}\StringTok{"^X"}\NormalTok{, }\StringTok{""}\NormalTok{, }\KeywordTok{sampleNames}\NormalTok{(hl)))}
-\KeywordTok{pData}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Replicate} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{1}\hlopt{:}\hlnum{2}\hlstd{,} \hlkwc{each} \hlstd{=} \hlnum{10}\hlstd{)}
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Tag} \hlkwb{<-} \hlkwd{sub}\hlstd{(}\hlstr{"\textbackslash{}\textbackslash{}.1$"}\hlstd{,} \hlstr{""}\hlstd{,} \hlkwd{sub}\hlstd{(}\hlstr{"^X"}\hlstd{,} \hlstr{""}\hlstd{,} \hlkwd{sampleNames}\hlstd{(hl)))}
+\hlkwd{pData}\hlstd{(hl)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-Throughout this workflow we refer to the different columns that are
-found in the \texttt{exprs} (expression data) slot as channels (short
-for TMT channels). In the frame of LOPIT and hyperLOPIT these channels
-consititue the relative abundance of each protein (along the rows) in
-the channel of interest. Each TMT channel originates from fractions
-collected from the density gradient, or a set of pooled fractions or may
-be a sample originating from an alternative preparation e.g.~such as
-from the chromatin enrichment performed in Christoforou et al
-{[}@hyper{]}. Information about which gradient fractions were used for
-which tag should also be stored in the sample meta-data \texttt{pData}
-slot.
+\end{kframe}
+\end{knitrout}
+
+Throughout this workflow we refer to the different columns that
+are found in the \texttt{exprs} (expression data) slot as channels (short for TMT channels).
+In the frame of LOPIT and hyperLOPIT these channels consititue the
+relative abundance of each protein (along the rows) in the channel of
+interest. Each TMT channel originates from fractions collected from
+the density gradient, or a set of pooled fractions or may be a sample
+originating from an alternative preparation e.g. such as from the
+chromatin enrichment performed in Christoforou et al \cite{hyper}.
+Information about which gradient fractions were used for which tag
+should also be stored in the sample meta-data \texttt{pData} slot.
 
 The sample meta-data that is distributed with the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
-package for Christoforou's hyperLOPIT experiment and (as above) the
-quantitation data file, are located in the \texttt{extdata} in the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
+\Biocexptpkg{pRolocdata} package for Christoforou's
+hyperLOPIT experiment and (as above) the quantitation data file, are
+located in the \texttt{extdata} in the \Biocexptpkg{pRolocdata}
 package on the hard drive.
 
-In the code chunk below we again use the \texttt{dir} function to locate
-the filepath to the meta-data \texttt{csv} file and then read it into R
-using \texttt{read.csv}. We then append the meta-data to the
-\texttt{pData} slot. Information about the gradient fractions used and
-the associated subcellular fraction densities in \% w/v Iodixanol for
+In the code chunk below we again use the
+\texttt{dir} function to locate the filepath to the meta-data \texttt{csv} file and
+then read it into R using \texttt{read.csv}. We then append the meta-data to
+the \texttt{pData} slot.  Information about the gradient fractions used and
+the associated subcellular fraction densities in % w/v Iodixanol for
 each replicate are stored here.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{expinfo <-}\StringTok{ }\KeywordTok{dir}\NormalTok{(extdatadir, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{,}
-               \DataTypeTok{pattern =} \StringTok{"hyperLOPIT-SIData-fraction-info.csv"}\NormalTok{)}
-
-\NormalTok{fracinfo <-}\StringTok{ }\KeywordTok{read.csv}\NormalTok{(expinfo, }\DataTypeTok{row.names=}\DecValTok{1}\NormalTok{, }\DataTypeTok{skip =} \DecValTok{2}\NormalTok{, }
-                     \DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
-
-\KeywordTok{pData}\NormalTok{(hl)$Gradient.Fraction <-}\StringTok{ }\KeywordTok{c}\NormalTok{(fracinfo[, }\DecValTok{1}\NormalTok{], fracinfo[, }\DecValTok{2}\NormalTok{])}
-\KeywordTok{pData}\NormalTok{(hl)$Iodixonal.Density <-}\StringTok{ }\KeywordTok{c}\NormalTok{(fracinfo[, }\DecValTok{4}\NormalTok{], fracinfo[, }\DecValTok{5}\NormalTok{])}
-\KeywordTok{pData}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{expinfo} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+               \hlkwc{pattern} \hlstd{=} \hlstr{"hyperLOPIT-SIData-fraction-info.csv"}\hlstd{)}
+
+\hlstd{fracinfo} \hlkwb{<-} \hlkwd{read.csv}\hlstd{(expinfo,} \hlkwc{row.names}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{skip} \hlstd{=} \hlnum{2}\hlstd{,}
+                     \hlkwc{header} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{stringsAsFactors} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Gradient.Fraction} \hlkwb{<-} \hlkwd{c}\hlstd{(fracinfo[,} \hlnum{1}\hlstd{], fracinfo[,} \hlnum{2}\hlstd{])}
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Iodixonal.Density} \hlkwb{<-} \hlkwd{c}\hlstd{(fracinfo[,} \hlnum{4}\hlstd{], fracinfo[,} \hlnum{5}\hlstd{])}
+\hlkwd{pData}\hlstd{(hl)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\subsection{Data processing}\label{data-processing}
-
-\subsubsection{Normalisation}\label{normalisation}
-
-There are two aspects related to data normalisation that are relevant to
-spatial proteomics data processing. The first one focuses on reducing
-purely technical variation between channels without affecting biological
-variability (i.e.~the shape of the quantitative profiles). This
-normalisation will depend on the underlying quantitative technology and
-the experimental design, and will not be addressed in this workflow. The
-second aspect, and more specific to spatial proteomics data, is scaling
-all the organelle-specific profiles into a same intensity interval
-(typically 0 and 1) by, for example, diving each intensity by the sum of
-the intensities for that quantitative feature. This is not necessary in
-this example as the intensities for each replicate have already been
-re-scaled to 1 in Proteome Discoverer. However, if one wanted to do this
-they would execute the \texttt{normalise} function as demonstrated in
-the below code chunk.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{normalise}\NormalTok{(hl, }\DataTypeTok{method =} \StringTok{"sum"}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
+\end{kframe}
+\end{knitrout}
+
+\subsection*{Data processing}
+
+\subsubsection*{Normalisation}
+
+There are two aspects related to data normalisation that are relevant
+to spatial proteomics data processing. The first one focuses on
+reducing purely technical variation between channels without
+affecting biological variability (i.e. the shape of the quantitative
+profiles). This normalisation will depend on the underlying
+quantitative technology and the experimental design, and will not be
+addressed in this workflow. The second aspect, and more specific to
+spatial proteomics data, is scaling all the organelle-specific
+profiles into a same intensity interval (typically 0 and 1) by, for
+example, diving each intensity by the sum of the intensities for that
+quantitative feature. This is not necessary in this example as the
+intensities for each replicate have already been re-scaled to 1 in
+Proteome Discoverer. However, if one wanted to do this they would
+execute the \texttt{normalise} function as demonstrated in the below code
+chunk.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{normalise}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"sum"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-The same \texttt{normalise} function (or \texttt{normalize}, both
-spellings are supported) can also be applied in the first case described
-above. Different normalisation methods such as mean or median scaling,
-variance stabilisation or quantile normalisation, to cite a few, can be
-applied to accomodation different needs.
+The same \texttt{normalise} function (or \texttt{normalize}, both spellings are
+supported) can also be applied in the first case described above.
+Different normalisation methods such as mean or median scaling,
+variance stabilisation or quantile normalisation, to cite a few, can
+be applied to accomodation different needs.
 
 As previously mentioned, before combination, the two replicates in the
-\texttt{hl} data that we read into R were separately normalised by sum
-(i.e.~to 1) across the 10 channels for each replicate respectively. We
-can verify this by summing each rows for each replicate:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{summary}\NormalTok{(}\KeywordTok{rowSums}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hl[, hl$replicate ==}\StringTok{ }\DecValTok{1}\NormalTok{])))}
-\end{Highlighting}
-\end{Shaded}
-
+\texttt{hl} data that we read into R were separately normalised by sum (i.e.
+to 1) across the 10 channels for each replicate respectively. We can
+verify this by summing each rows for each replicate:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{replicate} \hlopt{==} \hlnum{1}\hlstd{])))}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{summary}\NormalTok{(}\KeywordTok{rowSums}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hl[, hl$replicate ==}\StringTok{ }\DecValTok{2}\NormalTok{])))}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{replicate} \hlopt{==} \hlnum{2}\hlstd{])))}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 We see that some features do not add up exactly to 1 due to rounding
-errors after exporting to intermediate files. These small deviations do
-not bear any consequences here.
-
-\subsubsection{Combining acquisitions}\label{combining-acquisitions}
-
-The spreadsheet that was used to create the \texttt{hl} MSnSet included
-the two replicates within one .csv file. We also provide individual
-replicates in the
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}
+errors after exporting to intermediate files. These small deviations
+do not bear any consequences here.
+
+\subsubsection*{Combining acquisitions}
+
+The spreadsheet that was used to create the \texttt{hl} MSnSet
+included the two replicates within one .csv file.  We also provide
+individual replicates in the \Biocexptpkg{pRolocdata}
 package. Below, we show how to combine \texttt{MSnSet} objects and,
 subsequently, how to filter and handle missing values. We start by
-loading the \texttt{r Biocexptpkg("pRolocdata")} package and the
+loading the \Biocexptpkg{pRolocdata} package and the
 equivalent replicates using the \texttt{data} function.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocdata"}\NormalTok{)}
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r1)}
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRolocdata"}\hlstd{)}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r1)}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{pRolocdata}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
-will list the 54 datasets that are available in
-\emph{\href{http://bioconductor.org/packages/pRolocdata}{pRolocdata}}.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{pRolocdata}\hlstd{()}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+will list the 54 datasets that are
+available in \Biocexptpkg{pRolocdata}.
+
+%% $
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{identical}\NormalTok{(}\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r1), }\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r2))}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{identical}\hlstd{(}\hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r1),} \hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r2))}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hyperLOPIT2015ms3r1 <-}\StringTok{ }\KeywordTok{updateSampleNames}\NormalTok{(hyperLOPIT2015ms3r1, }\DecValTok{1}\NormalTok{)}
-\NormalTok{hyperLOPIT2015ms3r2 <-}\StringTok{ }\KeywordTok{updateSampleNames}\NormalTok{(hyperLOPIT2015ms3r2, }\DecValTok{2}\NormalTok{)}
-\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r1)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlstd{hyperLOPIT2015ms3r1} \hlkwb{<-} \hlkwd{updateSampleNames}\hlstd{(hyperLOPIT2015ms3r1,} \hlnum{1}\hlstd{)}
+\hlstd{hyperLOPIT2015ms3r2} \hlkwb{<-} \hlkwd{updateSampleNames}\hlstd{(hyperLOPIT2015ms3r2,} \hlnum{2}\hlstd{)}
+\hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r1)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{sampleNames}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-In addition, to matching names, the content of the feature metadata for
-identical feature annotations must match exactly across the data to be
-combined. In particular for these data, we expect the same proteins in
-each replicates to be annotated with the same UniProt entry names and
-descriptions, but not with the same coverage of number of peptides or
-peptide-spectrum matches (PSMs).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)}
-\end{Highlighting}
-\end{Shaded}
-
+\end{kframe}
+\end{knitrout}
+
+In addition, to matching names, the content of the feature metadata
+for identical feature annotations must match exactly across the data
+to be combined. In particular for these data, we expect the same
+proteins in each replicates to be annotated with the same UniProt
+entry names and descriptions, but not with the same coverage of number
+of peptides or peptide-spectrum matches (PSMs).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{], }\DecValTok{1}\NormalTok{)}
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{] <-}\StringTok{ }\KeywordTok{paste0}\NormalTok{(}\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{], }\DecValTok{2}\NormalTok{)}
-\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r1) <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r1)[}\DecValTok{1}\NormalTok{:}\DecValTok{5}\NormalTok{] }
-\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r2) <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hyperLOPIT2015ms3r2)[}\DecValTok{3}\NormalTok{:}\DecValTok{5}\NormalTok{]}
-\KeywordTok{fvarLabels}\NormalTok{(hyperLOPIT2015ms3r1)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{],} \hlnum{1}\hlstd{)}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{],} \hlnum{2}\hlstd{)}
+\hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r1)} \hlkwb{<-} \hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{1}\hlopt{:}\hlnum{5}\hlstd{]}
+\hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r2)} \hlkwb{<-} \hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r2)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)}
+\end{alltt}
 \begin{verbatim}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{combined <-}\StringTok{ }\KeywordTok{combine}\NormalTok{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
-\NormalTok{combined}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
+\hlstd{combined}
+\end{alltt}
 \begin{verbatim}
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
 ##  MSnbase version: 1.21.7
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 More details above combining data are given in the dedicated
-\emph{Combining MSnSet instances} section of the
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}
+*Combining MSnSet instances* section of the \Biocpkg{MSnbase}
 \href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
@@
 sub-cellular resolution. Missing data can be imputated using 
-\emph{\href{http://bioconductor.org/packages/MSnbase}{MSnbase}}'s
-\texttt{impute} function. The method underlying the imputation method is
-then determined by a \texttt{methods} parameter. In our particular case,
-missing values are indicative of protein groups that were not profiles
-in both replicates.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{image2}\NormalTok{(}\KeywordTok{is.na}\NormalTok{(combined), }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"black"}\NormalTok{, }\StringTok{"white"}\NormalTok{),}
-       \DataTypeTok{main =} \StringTok{"Missing values (white cells) after combining replicates"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying the
+imputation method is then determined by a \texttt{methods} parameter. In our
+particular case, missing values are indicative of protein groups that
+were not profiles in both replicates.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/namap-1.png}
-\caption{plot of chunk namap}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
+       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/namap-1}
+
+\end{knitrout}
+  \caption{Heatmap of missing values}
+  \label{fig:namap}
 \end{figure}
@@
 We prefer to remove proteins that were not assayed in both of our two
-replicated experiments. This is done with the \texttt{filterNA} function
-that removes features that contain more than a certain proportion
-(default is 0) missing values.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{combined <-}\StringTok{ }\KeywordTok{filterNA}\NormalTok{(combined)}
-\NormalTok{combined}
-\end{Highlighting}
-\end{Shaded}
-
+replicated experiments. This is done with the \texttt{filterNA} function that
+removes features that contain more than a certain proportion (default
+is 0) missing values.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
+\hlstd{combined}
+\end{alltt}
 \begin{verbatim}
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 10:28:15 2016 
-## Subset [6725,20][5032,20] Thu Sep 29 10:28:15 2016 
-## Removed features with more than 0 NAs: Thu Sep 29 10:28:15 2016 
-## Dropped featureData's levels Thu Sep 29 10:28:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
+## Subset [6725,20][5032,20] Thu Sep 29 21:02:54 2016 
+## Removed features with more than 0 NAs: Thu Sep 29 21:02:54 2016 
+## Dropped featureData's levels Thu Sep 29 21:02:54 2016 
 ##  MSnbase version: 1.21.7
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 In the code chunk below we produce a PCA plot of the mouse stem cell
-dataset. One point on the plot represents one protein. We can indeed see
-several distinct protein clusters. We specify \texttt{fcol = NULL},
-which means not to consider any feature variable to annotate the
-features (proteins) with colours. We will see later how to use this to
-annotate the PCA plot with prior information about sub-cellular
-localisation.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRoloc"}\NormalTok{)}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \OtherTok{NULL}\NormalTok{, }\DataTypeTok{col =} \StringTok{"black"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+dataset. One point on the plot represents one protein. We can indeed
+see several distinct protein clusters. We specify \texttt{fcol = NULL}, which
+means not to consider any feature variable to annotate the features
+(proteins) with colours. We will see later how to use this to annotate
+the PCA plot with prior information about sub-cellular localisation.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/qcplot-1.png}
-\caption{plot of chunk qcplot}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRoloc"}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlkwa{NULL}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"black"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/qcplot-1}
+
+\end{knitrout}
+  \caption{PCA plot of the mouse stem cell data \texttt{hl}.}
+  \label{fig:pcahl}
 \end{figure}
 
-In the first instance we advise one to visualise their data without any
-annotation (i.e.~with \texttt{fcol = NULL}), before proceeding with data
+In the first instance we advise one to visualise their data without
+any annotation (i.e. with \texttt{fcol = NULL}), before proceeding with data
 annotation. The identification of well resolved clusters in the data,
-constitutes an unbiased assessment of the data structure, demonstrating
-the successful separation of sub-cellular clusters.
-
-\textbf{TODO} There are other dimensionality reduction methods available
-in the \texttt{plot2D} function, which can be parametrised with the
+constitutes an unbiased assessment of the data structure,
+demonstrating the successful separation of sub-cellular clusters.
+
+**TODO** There are other dimensionality reduction methods available in
+the \texttt{plot2D} function, which can be parametrised with the
 \texttt{method} argument.  Advantages, disadvantages? See
-\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}{this}
-very interesting gist with intriguing tSNE examples.
+\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}
+very interesting gist with intriguing tSNE examples. (This might be
+better to come after markers.)
 
@@
 gradient to identify channels displaying particularly low yield. This
-can be done using the \texttt{plotDist} and \texttt{boxplot} functions,
-that plot the protein profiles occupancy along the gradient (we also
-display the mean channel intensities) and a \texttt{boxplot} of the
-column intensities. In the two plots displayed below, we re-order the
-TMT channles to pair corresponding channels in the two replicates
-(rather than ordering the channels by replicate).
-
-\texttt{\{r qcbx, fig.width = 12\} par(mfrow = c(1, 2)) o \textless{}- order(hl\$Iodixonal.Density) plotDist(hl{[}, o{]}, pcol = "\#00000010") lines(colMeans(exprs(hl{[}, o{]})), col = "red", type = "b") boxplot(exprs(hl{[}, o{]}), las = 2)}
-
-\section{Markers}\label{markers}
+can be done using the \texttt{plotDist} and \texttt{boxplot} functions, that plot
+the protein profiles occupancy along the gradient (we also display the
+mean channel intensities) and a \texttt{boxplot} of the column intensities.
+In the two plots displayed below, we re-order the TMT channles to pair
+corresponding channels in the two replicates (rather than ordering
+the channels by replicate).
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlstd{o} \hlkwb{<-} \hlkwd{order}\hlstd{(hl}\hlopt{$}\hlstd{Iodixonal.Density)}
+\hlkwd{plotDist}\hlstd{(hl[, o],} \hlkwc{pcol} \hlstd{=} \hlstr{"#00000010"}\hlstd{)}
+\hlkwd{lines}\hlstd{(}\hlkwd{colMeans}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o])),} \hlkwc{col} \hlstd{=} \hlstr{"red"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"b"}\hlstd{)}
+\hlkwd{boxplot}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o]),} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/qcbx-1}
+
+\end{knitrout}
+  \caption{Protein profiles and distribution of channel intensities.}
+  \label{fig:qcbx}
+\end{figure}
+
+\section*{Markers}
 
@@
 time consuming, but it is important to define markers that are
-representative of the multivariate data space and on which a classifier
-will be trained and generated.
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}} provides a
-convenience function, \texttt{addMarkers}, to directly add markers to a
-\texttt{MSnSet} object, as demonstrated in the code chunk below. These
-marker sets can be accessed using the \texttt{pRolocmarkers()} function.
-Marker sets are stored as a simple named vector in R, and originate from
-in-house user-defined sets of markers or from previous published studies
-{[}@Gatto:2014b{]}, which are continuosly updated and integrated.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## List available marker sets}
-\KeywordTok{pRolocmarkers}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
+representative of the multivariate data space and on which a
+classifier will be trained and generated. \Biocpkg{pRoloc}
+provides a convenience function, \texttt{addMarkers}, to directly add markers
+to a \texttt{MSnSet} object, as demonstrated in the code chunk below. These
+marker sets can be accessed using the \texttt{pRolocmarkers()}
+function. Marker sets are stored as a simple named vector in R, and
+originate from in-house user-defined sets of markers or from previous
+published studies \cite{Gatto:2014b}, which are continuosly updated and 
+integrated.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## List available marker sets}
+\hlkwd{pRolocmarkers}\hlstd{()}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-These markers can then be mapped to a \texttt{MSnSet}'s
-\texttt{featureNames}. The mouse dataset used here has Uniprot IDs
-stored as the \texttt{featureNames} (see
-\texttt{head(featureNames(hl))}) and the names of the vector of the
-mouse markers stored in
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}
-(\texttt{mmus} markers) are also Uniprot IDs (see \texttt{head(mrk)} in
-the code chunk below), so it is straightforward to match names between
-the markers and the \texttt{MSnSet} instance using the
-\texttt{addMarkers} function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Use mouse markers}
-\NormalTok{mrk <-}\StringTok{ }\KeywordTok{pRolocmarkers}\NormalTok{(}\DataTypeTok{species =} \StringTok{"mmus"}\NormalTok{)}
-\KeywordTok{head}\NormalTok{(mrk)}
-\end{Highlighting}
-\end{Shaded}
-
+\end{kframe}
+\end{knitrout}
+
+These markers can then be mapped to a \texttt{MSnSet}'s \texttt{featureNames}. The
+mouse dataset used here has Uniprot IDs stored as the \texttt{featureNames}
+(see \texttt{head(featureNames(hl))}) and the names of the vector of
+the mouse markers stored in \Biocpkg{pRoloc} (\texttt{mmus} markers) 
+are also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below), so 
+it is straightforward to match names between the markers and 
+the \texttt{MSnSet} instance using the \texttt{addMarkers} function.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Use mouse markers}
+\hlstd{mrk} \hlkwb{<-} \hlkwd{pRolocmarkers}\hlstd{(}\hlkwc{species} \hlstd{=} \hlstr{"mmus"}\hlstd{)}
+\hlkwd{head}\hlstd{(mrk)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Add mouse markers}
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{addMarkers}\NormalTok{(hl, mrk)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-## Markers in data: 937 out of 5032
-\end{verbatim}
-
-\begin{verbatim}
+\begin{alltt}
+\hlcom{## Add mouse markers}
+\hlstd{hl} \hlkwb{<-} \hlkwd{addMarkers}\hlstd{(hl, mrk)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Markers in data: 937 out of 5032}}\begin{verbatim}
 ## organelleMarkers
@@
 \end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 If the naming between the marker sets and the \texttt{MSnSet} dataset
-are different, one will have to convert and match the proteins according
-to the appropriate identifier. Sometimes, we find the equivalent entry
-name, Uniprot ID or accession number is stored with the data, which
-makes conversion between identifers relatively straightforward. If this
-is not the case however, conversion can be performed using
-\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}}, the
-Bioconductor
+are different, one will have to convert and match the proteins
+according to the appropriate identifier. Sometimes, we find the
+equivalent entry name, Uniprot ID or accession number is stored with
+the data, which makes conversion between identifers relatively
+straightforward. If this is not the case however, conversion can be
+performed using \Biocpkg{biomaRt}, the Bioconductor
 \href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation
@@
 We now visualise these annotations along the PCA plot using the
-\texttt{plot2D} function and then use the \texttt{addLegend} function to
-map the marker classes to the pre-defined colours. We also display the
-data along the first and seventh PCs using the \texttt{dims} argument.
-Note that in this second call to the \texttt{plot2D} function, we have
-omitted the \texttt{fcol} argument, as the \texttt{"markers"} feature
-variable name is the default value. We choose to display PCs 1 and 7 to
-illustrate that while upper principal components explain much less
-variability in the data (2.23\% for PC7, as opposed to 48.41\% for PC1),
-we see that the mitochondrial and peroxisome clusters can be
-differenciated, despite the apparent overlap in the two first PCs.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{main =} \StringTok{"pRolocmarkers for mouse"}\NormalTok{)}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{)}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{dims =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{7}\NormalTok{), }\DataTypeTok{main =} \StringTok{"Marker resolution along PC 1 and 7"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\texttt{plot2D} function and then use the \texttt{addLegend} function to map the
+marker classes to the pre-defined colours. We also display the data
+along the first and seventh PCs using the \texttt{dims} argument. Note that
+in this second call to the \texttt{plot2D} function, we have omitted the
+\texttt{fcol} argument, as the \texttt{"markers"} feature variable name is the
+default value. We choose to display PCs 1 and 7 to illustrate that
+while upper principal components explain much less variability in the
+data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
+mitochondrial and peroxisome clusters can be differenciated, despite
+the apparent overlap in the two first PCs.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotmarkers-1.png}
-\caption{plot of chunk plotmarkers}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{main} \hlstd{=} \hlstr{"pRolocmarkers for mouse"}\hlstd{)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.7}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{7}\hlstd{),} \hlkwc{main} \hlstd{=} \hlstr{"Marker resolution along PC 1 and 7"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plotmarkers-1}
+
+\end{knitrout}
+  \caption{Annotated PCA plots of the \texttt{hl} data.}
+  \label{fig:plotmarkers}
 \end{figure}
 
-The colours have been defined so as to enable to differenciate up to 30
-classes. If more are provided, different character symbols (circles,
-squares, \ldots{} and empty and solid symbols) are used. The colours and
-the default plotting characters (solid dots for the markers and empty
-circles for the features of unknown localisation) can of course be
-changed, as described in the \texttt{setStockcol} manual page.
-
-As demonstrated in {[}@hyper{]} and illustrated in the PCA plot above,
+The colours have been defined so as to enable to differenciate up to
+30 classes. If more are provided, different character symbols
+(circles, squares, ... and empty and solid symbols) are used. The
+colours and the default plotting characters (solid dots for the
+markers and empty circles for the features of unknown localisation)
+can of course be changed, as described in the \texttt{setStockcol} manual
+page.
+
+As demonstrated in \cite{hyper} and illustrated in the PCA plot above,
 the Golgi apparatus proteins display a dynamic pattern, noting sets of
@@
 \texttt{fDataTounknown} function. It is often used to replace empty
-strings (``'') or missing values in the markers definition to a common
-definition of \emph{unknown localisation}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(hl, }\DataTypeTok{from =} \StringTok{"Golgi apparatus"}\NormalTok{, }\DataTypeTok{to =} \StringTok{"unknown"}\NormalTok{)}
-\KeywordTok{getMarkers}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
+strings ("") or missing values in the markers definition to a common
+definition of *unknown localisation*.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{fDataToUnknown}\hlstd{(hl,} \hlkwc{from} \hlstd{=} \hlstr{"Golgi apparatus"}\hlstd{,} \hlkwc{to} \hlstd{=} \hlstr{"unknown"}\hlstd{)}
+\hlkwd{getMarkers}\hlstd{(hl)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-In general, the Gene Ontology (GO) {[}@Ashburner:2000{]}, and in
+\end{kframe}
+\end{knitrout}
+
+In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
 particular the cellular compartment (CC) namespace are a good starting
@@
 efficiently visualise the complete dataset and assess the relative
-separation of different sub-cellular niches, comparing profiles of a few
-marker clusters is useful to assess how exactly they differ (in terms of
-peak channels, for example). Below, we plot the profile of the
-mitochondrial and peroxisome markers to highlight the differences in
-profiles between these two sets of markers along the 6th and 7th
+separation of different sub-cellular niches, comparing profiles of a
+few marker clusters is useful to assess how exactly they differ (in
+terms of peak channels, for example). Below, we plot the profile of
+the mitochondrial and peroxisome markers to highlight the differences
+in profiles between these two sets of markers along the 6th and 7th
 channels, as represented above along the 7th PC on the PCA plot.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hlo <-}\StringTok{ }\NormalTok{hl[, }\KeywordTok{order}\NormalTok{(hl$Iodixonal.Density)]}
-\KeywordTok{plotDist}\NormalTok{(hlo[}\KeywordTok{fData}\NormalTok{(hlo)$markers ==}\StringTok{ "Mitochondrion"}\NormalTok{, ],}
-         \DataTypeTok{pcol =} \StringTok{"purple"}\NormalTok{, }\DataTypeTok{fractions =} \StringTok{"Fraction.No"}\NormalTok{)}
-\KeywordTok{title}\NormalTok{(}\DataTypeTok{main =} \StringTok{"Marker occupancy profiles along the gradient"}\NormalTok{)}
-\KeywordTok{matlines}\NormalTok{(}\KeywordTok{t}\NormalTok{(}\KeywordTok{exprs}\NormalTok{(hlo[}\KeywordTok{fData}\NormalTok{(hlo)$markers ==}\StringTok{ "Peroxisome"}\NormalTok{, ])),}
-         \DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \StringTok{"darkblue"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"l"}\NormalTok{)}
-\KeywordTok{legend}\NormalTok{(}\StringTok{"topleft"}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"Mitochondrion"}\NormalTok{, }\StringTok{"Peroxisome"}\NormalTok{),}
-       \DataTypeTok{lty =} \DecValTok{1}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"purple"}\NormalTok{, }\StringTok{"blue"}\NormalTok{), }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotDist-1.png}
-\caption{plot of chunk plotDist}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hlo} \hlkwb{<-} \hlstd{hl[,} \hlkwd{order}\hlstd{(hl}\hlopt{$}\hlstd{Iodixonal.Density)]}
+\hlkwd{plotDist}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Mitochondrion"}\hlstd{, ],}
+         \hlkwc{pcol} \hlstd{=} \hlstr{"purple"}\hlstd{,} \hlkwc{fractions} \hlstd{=} \hlstr{"Fraction.No"}\hlstd{)}
+\hlkwd{title}\hlstd{(}\hlkwc{main} \hlstd{=} \hlstr{"Marker occupancy profiles along the gradient"}\hlstd{)}
+\hlkwd{matlines}\hlstd{(}\hlkwd{t}\hlstd{(}\hlkwd{exprs}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Peroxisome"}\hlstd{, ])),}
+         \hlkwc{lty} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"darkblue"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"l"}\hlstd{)}
+\hlkwd{legend}\hlstd{(}\hlstr{"topleft"}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"Mitochondrion"}\hlstd{,} \hlstr{"Peroxisome"}\hlstd{),}
+       \hlkwc{lty} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"purple"}\hlstd{,} \hlstr{"blue"}\hlstd{),} \hlkwc{bty} \hlstd{=} \hlstr{"n"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/plotDist-1}
+
+\end{knitrout}
+  \caption{Mitochondrion and peroxisome protein profiles.}
+  \label{fig:plotDist2}
 \end{figure}
 
-\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust},
-although these are only available in the development version.
-
-\section{Replication}\label{replication}
+\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust}
+(also mentioned later), although these are only available in the
+development version.
+
+\section*{Replication}
 
@@
 consequently, the reliability in protein sub-cellular assignments, we
-follow the advice in {[}@Trotter:2010{]} and combine replicated spatial
-proteomics experiments as described above. Indeed, Trotter et al. have
-shown a significant improvement in protein--organelle association upon
-direct combination of single experiments, in particular when these
-resolve different subcellular niches.
-
-Direct comparisons of individual channels in replicated experiments does
-not provide an adequate, goal-driven assessment of different
-experiments. Indeed, due to the nature of the experiment and gradient
-fraction collection, the quantitative channels do not correspond to
-identical selected fractions along the gradient. As can be seen in the
-table below (taken from the \texttt{hyperLOPIT}'s \texttt{pData},
-focusing on channels 7 to 10), different sets of gradient fractions are
-pooled to obtain enough material and optimise acurate quantitation.
+follow the advice in \cite{Trotter:2010} and combine replicated spatial
+proteomics experiments as described above. Indeed, Trotter et
+al. have shown a significant improvement in protein–organelle
+association upon direct combination of single experiments, in
+particular when these resolve different subcellular niches.
+
+Direct comparisons of individual channels in replicated experiments
+does not provide an adequate, goal-driven assessment of different
+experiments. Indeed, due to the nature of the experiment and gradient fraction
+collection, the quantitative channels 
+do not correspond to identical selected fractions along the
+gradient. As can be seen in the table below (taken from the
+\texttt{hyperLOPIT}'s \texttt{pData}, focusing on channels 7 to 10), different sets
+of gradient fractions are pooled to obtain enough material and optimise
+acurate quantitation.
+
+% On the figure below, we compare the relative intensities of channels 7
+% to 10 on scatter plots, highlighting different marker sets. These
+% differences result from different peak fractions for these niches on
+% the mitochondrial and chromatin occupancy profiles for the two
+% replicates.
+
+% <<repl1, eval = TRUE, echo=FALSE, fig.width = 12>>=
+% sel0 <- fData(hyperLOPIT)$markers != "unknown"
+% cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
+% cls[!sel0] <- getUnknowncol()
+
+% par(mfrow = c(2, 4))
+% for (i in c(7, 8, 10, 9)) {
+%     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
+%          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
+%     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
+%     grid()
+%     abline(0, 1, col = "red")
+% }
+% plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
+% addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
+
+% i <- fData(hyperLOPIT)$markers == "Mitochondrion"
+% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
+%          featureNames(hyperLOPIT)[i], type = "l",
+%          mcol = "orange")
+% title(main = "Mitochondrion")
+% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
+% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
+% legend("topleft", paste("Replicate", 1:2),
+%        col = c("#F781BF", "orange"),
+%        lty = 1, bty = "n")
+
+% i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
+% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
+%          featureNames(hyperLOPIT)[i], type = "l")
+% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
+% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
+% title(main = "Chromatin")
+% legend("topleft", paste("Replicate", 1:2),
+%        col = c("#984EA3", "steelblue"),
+%        lty = 1, bty = "n")
+% @
 
@@
 assuring that each displays acceptable sub-cellular resolution. Note
-that in the code chunk below, we mirror the x-axis to represent the two
-figures with the same orientation.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{)) }
-\KeywordTok{plot2D}\NormalTok{(hl[, hl$Replicate ==}\StringTok{ }\DecValTok{1}\NormalTok{], }\DataTypeTok{main =} \StringTok{"Replicate 1"}\NormalTok{) }
-\KeywordTok{plot2D}\NormalTok{(hl[, hl$Replicate ==}\StringTok{ }\DecValTok{2}\NormalTok{], }\DataTypeTok{main =} \StringTok{"Replicate 2"}\NormalTok{, }\DataTypeTok{mirrorX =} \OtherTok{TRUE}\NormalTok{) }
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+that in the code chunk below, we mirror the x-axis to represent the
+two figures with the same orientation.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plot2Drep-1.png}
-\caption{plot of chunk plot2Drep}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{1}\hlstd{],} \hlkwc{main} \hlstd{=} \hlstr{"Replicate 1"}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{2}\hlstd{],} \hlkwc{main} \hlstd{=} \hlstr{"Replicate 2"}\hlstd{,} \hlkwc{mirrorX} \hlstd{=} \hlnum{TRUE}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plot2Drep-1}
+
+\end{knitrout}
+  \caption{PCA plots of replicates 1 and 2}
+  \label{fig:plot2Drep}
 \end{figure}
 
-\section{Interactive visualisation}\label{interactive-visualisation}
+\section*{Interactive visualisation}
 
@@
 analyses allowing one to shed light on data structure and patterns of
-interest. Using the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package we can interactively visualise, explore and interrogate
-quantitative spatial proteomics data. The
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package is currently under active development and it relies on the
-\texttt{shiny} framework for reactivity and interactivity. The package
-currently distributes 3 different GUI's (``main'' (default), ``compare''
-or ``compare'') which are wrapped and launched by the \texttt{pRolocVis}
-function. In the below code chunk we lauch the main app (note, we do not
-need to specify the argument, \texttt{app = "main"} as it is the
-default).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocGUI"}\NormalTok{)}
-\KeywordTok{pRolocVis}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+interest. Using the \Biocpkg{pRolocGUI} package we can
+interactively visualise, explore and interrogate quantitative spatial
+proteomics data. The \Biocpkg{pRolocGUI} package is currently
+under active development and it relies on the \texttt{shiny} framework for
+reactivity and interactivity. The package currently distributes 3
+different GUI's ("main" (default), "compare" or "compare") which are
+wrapped and launched by the \texttt{pRolocVis} function. In the below code
+chunk we lauch the main app (note, we do not need to specify the
+argument, \texttt{app = "main"} as it is the default).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRolocGUI"}\hlstd{)}
+\hlkwd{pRolocVis}\hlstd{(hl)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/mainapp.png}
-\caption{A screen shot of clickable interface and zoomable PCA plot of
-the main app in the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package.}
+  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
+  \caption{A screen shot of clickable interface and zoomable PCA plot of the main app in the \Biocpkg{pRolocGUI} package.}
+  \label{fig:app}
 \end{figure}
 
-As diplayed in the screenshot above, the ``main'' application is
+As diplayed in the screenshot above, the "main" application is
 designed for exploratory data analysis and is divied into 3 tabs: (1)
-PCA, (2) Profiles and (3) Table selection. The default view upon loading
-is the PCA tab, which features a clickable interface and zoomable PCA
-plot with an interactive data table for displaying the quantitation
-information. Particular proteins of interest can be highlighted using
-the text search box. There is also an alternate profiles tab for
-visualisation of the protein profiles, which can be used to examine the
-patterns of proteins of interest. The Table selection tab provides an
-interface to control data table column selection.
-
-The ``compare'' application is useful for examining two replicate
+PCA, (2) Profiles and (3) Table selection. The default view upon
+loading is the PCA tab, which features a clickable interface and
+zoomable PCA plot with an interactive data table for displaying the
+quantitation information. Particular proteins of interest can be
+highlighted using the text search box. There is also an alternate
+profiles tab for visualisation of the protein profiles, which can be
+used to examine the patterns of proteins of interest. The Table
+selection tab provides an interface to control data table column
+selection.
+
+The \textit{compare} application is useful for examining two replicate
 experiments, or two experiments from different conditions, treatments
-etc. The compare application is called by default if the input object to
-\texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets}, but
-it can also be specified by calling the argument
-\texttt{app = "compare"}. For example, in the code chunk below we first
-create a \texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT
-data, this is then passed to \texttt{pRolocVis}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r1)}
-\KeywordTok{data}\NormalTok{(hyperLOPIT2015ms3r2)}
-\NormalTok{mydata <-}\StringTok{ }\KeywordTok{MSnSetList}\NormalTok{(}\KeywordTok{list}\NormalTok{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
-\KeywordTok{pRolocVis}\NormalTok{(mydata, }\DataTypeTok{app =} \StringTok{"compare"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+etc. The compare application is called by default if the input object
+to \texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets},
+but it can also be specified by calling the argument \texttt{app =
+  "compare"}. For example, in the code chunk below we first create a
+\texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT data, this
+is then passed to \texttt{pRolocVis}.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r1)}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r2)}
+\hlstd{mydata} \hlkwb{<-} \hlkwd{MSnSetList}\hlstd{(}\hlkwd{list}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
+\hlkwd{pRolocVis}\hlstd{(mydata,} \hlkwc{app} \hlstd{=} \hlstr{"compare"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/SS_Compare1.jpg}
+  \includegraphics[width=\textwidth]{./Figures/SS_Compare1.jpg}
   \caption{The compare application, main panel}
+  \label{fig:compare}
 \end{figure}
@@
 to the limited number of marker proteins that exist in databases and
-elsewhere. Furthermore, given the vast complexity of the cell, automatic
-annotation retrieval does not always give a full representation of the
-true sub-cellular diversity in the data. For downstream analyses, such
-as supervised machine learning, it is desirable to obtain reliable
-markers that cover as many sub-cellular niches as possible, as these
-markers are directly used in the training phase of the ML
-classification. We find that a lack of sub-cellular diversity in the
-labelled training data leads to prediction errors, as unlabelled
-instances can only be assigned to a class that exists in the training
-data {[}@Breckels:2013{]}. In such scenarios novelty detection can be
-useful to identify data-specific sub-cellular groupings such as
-organelles and protein complexes. The phenotype discovery (phenoDisco)
-algorithm {[}@Breckels:2013{]} is one such method and is available in
-\emph{\href{http://bioconductor.org/packages/pRoloc}{pRoloc}}. It is an
-iterative semi-supervised learning method that combines the
-classification of proteins on existing labelled data with the detection
-of new clusters.
-
-In addition to extracting new phenotypes, novelty detection methods are
-also useful for confirming the presence of known or postulated clusters
-in an unbiased fashion. For example, in {[}@hyper{]} the
-\texttt{phenoDisco} algorithm was used to confirm the data-specific
-presence of the nucleus and nucleus sub-compartments. In the code chunk
-below, we demonstrate how to do this analysis, highlighting some of the
-optional arguments and parameters available for phenotype extraction and
-give some advice on how to interpret the output.
-
-As the \texttt{phenoDisco} algorithm is semi-supervised it uses both
-labelled (markers) and unlabelled data to explore the data structure and
-find new sub-cellular data clusters. Thus the first step is to define
-some input labelled data i.e.~markers, that the algorithm will use as
-input for the supervised learning aspect of the algorithm. As described
-in {[}@hyper{]} we define a set of markers to use as input for the
-analyses that cover well-known residents from three distinct organelle
+elsewhere. Furthermore, given the vast complexity of the cell,
+automatic annotation retrieval does not always give a full
+representation of the true sub-cellular diversity in the data. For
+downstream analyses, such as supervised machine learning, it is
+desirable to obtain reliable markers that cover as many sub-cellular
+niches as possible, as these markers are directly used in the training
+phase of the ML classification. We find that a lack of sub-cellular
+diversity in the labelled training data leads to prediction errors, as
+unlabelled instances can only be assigned to a class that exists in
+the training data \cite{Breckels:2013}. In such scenarios novelty
+detection can be useful to identify data-specific sub-cellular
+groupings such as organelles and protein complexes. The phenotype
+discovery (phenoDisco) algorithm \cite{Breckels:2013} is one such method
+and is available in \Biocpkg{pRoloc}. It is an iterative
+semi-supervised learning method that combines the classification of
+proteins on existing labelled data with the detection of new clusters.
+
+In addition to extracting new phenotypes, novelty detection methods
+are also useful for confirming the presence of known or postulated
+clusters in an unbiased fashion. For example, in \cite{hyper} the
+\texttt{phenoDisco} algorithm was used to confirm the data-specific presence
+of the nucleus and nucleus sub-compartments. In the code chunk below,
+we demonstrate how to do this analysis, highlighting some of the
+optional arguments and parameters available for phenotype extraction
+and give some advice on how to interpret the output.
+
+As the \texttt{phenoDisco} algorithm is semi-supervised it uses both labelled
+(markers) and unlabelled data to explore the data structure and find
+new sub-cellular data clusters. Thus the first step is to define some
+input labelled data i.e. markers, that the algorithm will use as input
+for the supervised learning aspect of the algorithm. As described in
+\cite{hyper} we define a set of markers to use as input for the analyses
+that cover well-known residents from three distinct organelle
 structures; the mitochondria, plasma membrane and ER, and from three
 well-known and abundant protein complexes; the proteasome and two
-ribosomal subunits, 40S and 60S. These input markers are stored in the
-\texttt{featureData} column of \texttt{hl} where
-\texttt{fcol = "phenoDisco.Input"}. We can use the convenience accessor
-function \texttt{getMarkers} to print out a table of the markers
-contained in this marker set. These initial markers were manually
-curated using information from the UniProt database, the Gene Ontology
-and the literature.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
+ribosomal subunits, 40S and 60S.  These input markers are stored in
+the \texttt{featureData} column of \texttt{hl} where \texttt{fcol = "phenoDisco.Input"}. We
+can use the convenience accessor function \texttt{getMarkers} to print out a
+table of the markers contained in this marker set. These initial
+markers were manually curated using information from the UniProt
+database, the Gene Ontology and the literature.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{)}
+\end{alltt}
 \begin{verbatim}
@@
 function and return a novelty detection result, according to the
-specified parameters. The algorithm parameters \texttt{times},
-\texttt{GS} and \texttt{p} are passed to the function, along with the
-\texttt{fcol} to tell the algorithm where the input training data is
-contained.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## As per Christoforou et al (2016), }
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{phenoDisco}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }
-                 \DataTypeTok{times =} \DecValTok{200}\NormalTok{, }
-                 \DataTypeTok{GS =} \DecValTok{60}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\emph{Note: We do not evaluate this code chunk in this document as the
-algorithm is computational intensive and best parallelised over multiple
-workers. This phenoDisco analysis took \textasciitilde{}24 hours to
-complete when parallelised over 40 workers.}
-
-The argument \texttt{times} indicated the number of times we run
-unsupervied Gaussian Mixture Modelling before defining a new phenotype
-cluster. The recommended minimum and default value is 100. In the above
-code chunk we increase the value to \texttt{times = 200} as we have
-found for larger datasets (e.g.~5000+ proteins) a higher times is
-requried for convergence. \texttt{GS} defines the minimum number of
-proteins allowed per new data cluster and thus heavily influences what
-type of new clusters are extracted. For example, if a user is interesed
-in the detection of small complexes they may wish to use a small
-\texttt{GS = 10}, or \texttt{= 20} etc. If they wish to detect larger
-more abundant sub-cellular niches a much higher \texttt{GS} would be
-preferable. Specifying a small \texttt{GS} can be more time consuming
-than using a larger \texttt{GS}, and there is a trade off between
-finding interesting small complexes and those that may not be of
-interest as we find there is a tendancy to find more noise when using a
-small \texttt{GS} compared to using a higher one.
-
-One may also consider increasing the search space for new data clusters
-by increasing the value of the parameter \texttt{G}. This defines the
-number of GMM components to test and fit; the default is
-\texttt{G = 1:9} (\texttt{mclust}'s default). One should note that the
-decreasing the \texttt{GS}, and increasing the values of the arguments
-\texttt{times}, \texttt{G} (among other function arguments, see
-\texttt{?phenoDisco}) will heavily influence (increase) the total time
-taken to run the algorithm. \texttt{phenoDisco} supports parallelisation
-and we strongly suggest you make use of a parallel processing to run
-these analyses (different backends can be set with argument
-\texttt{BPPARAM}).
-
-The ouput of running the \texttt{phenoDisco} algorithm is a
-\texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. We can see by typing
-\texttt{hl@processingData} directly into the console the processing
-information has been updated to the \texttt{MSnSet} recording the
-parameters that were used to run the analyses. This is handy for keeping
-track of data analyses. The results can be displayed by using the
-\texttt{getMarkers} function. We see that 5 new phenotype data clusters
-were found.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl}
-\end{Highlighting}
-\end{Shaded}
-
+specified parameters. The algorithm parameters \texttt{times}, \texttt{GS} and \texttt{p}
+are passed to the function, along with the \texttt{fcol} to tell the
+algorithm where the input training data is contained.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## As per Christoforou et al (2016), }
+\hlstd{hl} \hlkwb{<-} \hlkwd{phenoDisco}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,}
+                 \hlkwc{times} \hlstd{=} \hlnum{200}\hlstd{,}
+                 \hlkwc{GS} \hlstd{=} \hlnum{60}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Note: We do not evaluate this code chunk in this document as the
+algorithm is computational intensive and best parallelised over
+multiple workers. This phenoDisco analysis took ~24 hours to complete
+when parallelised over 40 workers.
+
+The argument \texttt{times} indicated the number of times we run unsupervied
+Gaussian Mixture Modelling before defining a new phenotype cluster.
+The recommended minimum and default value is 100. In the above code
+chunk we increase the value to \texttt{times = 200} as we have found for
+larger datasets (e.g. 5000+ proteins) a higher times is requried for
+convergence. \texttt{GS} defines the minimum number of proteins allowed per
+new data cluster and thus heavily influences what type of new clusters
+are extracted. For example, if a user is interesed in the detection of
+small complexes they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
+If they wish to detect larger more abundant sub-cellular niches a much
+higher \texttt{GS} would be preferable. Specifying a small \texttt{GS} can be more
+time consuming than using a larger \texttt{GS}, and there is a trade off
+between finding interesting small complexes and those that may not be
+of interest as we find there is a tendancy to find more noise when
+using a small \texttt{GS} compared to using a higher one.
+
+One may also consider increasing the search space for new data
+clusters by increasing the value of the parameter \texttt{G}. This defines
+the number of GMM components to test and fit; the default is \texttt{G = 1:9}
+(\texttt{mclust}'s default). One should note that the decreasing the \texttt{GS},
+and increasing the values of the arguments \texttt{times}, \texttt{G} (among other
+function arguments, see \texttt{?phenoDisco}) will heavily influence
+(increase) the total time taken to run the algorithm.  \texttt{phenoDisco}
+supports parallelisation and we strongly suggest you make use of a
+parallel processing to run these analyses (different backends can be
+set with argument \texttt{BPPARAM}).
+
+The ouput of running the \texttt{phenoDisco} algorithm is a \texttt{MSnSet}
+containing the new data clusters, appended to the \texttt{featureData} under
+the name \texttt{pd}. We can see by typing \texttt{hl@processingData} directly into
+the console the processing information has been updated to the
+\texttt{MSnSet} recording the parameters that were used to run the analyses.
+This is handy for keeping track of data analyses. The results can be
+displayed by using the \texttt{getMarkers} function. We see that 5 new
+phenotype data clusters were found.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl}
+\end{alltt}
 \begin{verbatim}
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Sep 29 10:28:15 2016 
-## Added markers from  'pdres' marker vector. Thu Sep 29 10:28:17 2016 
+## Added markers from  'mrk' marker vector. Thu Sep 29 21:02:55 2016 
+## Added markers from  'pdres' marker vector. Thu Sep 29 21:02:56 2016 
 ##  MSnbase version: 1.99.2
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
+\end{alltt}
 \begin{verbatim}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Re-order the colours for the phenoDisco output}
-\NormalTok{cl <-}\StringTok{ }\KeywordTok{getMarkerClasses}\NormalTok{(hl, }\StringTok{"pd"}\NormalTok{)}
-\NormalTok{cols <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)]}
-\NormalTok{ind <-}\StringTok{ }\KeywordTok{grep}\NormalTok{(}\StringTok{"Pheno"}\NormalTok{, cl, }\DataTypeTok{invert =} \OtherTok{TRUE}\NormalTok{)}
-\NormalTok{cols[ind] <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)][}\DecValTok{1}\NormalTok{:}\KeywordTok{length}\NormalTok{(ind)]}
-\NormalTok{cols[-ind] <-}\StringTok{ }\KeywordTok{getStockcol}\NormalTok{()[}\KeywordTok{seq}\NormalTok{(cl)][(}\KeywordTok{length}\NormalTok{(ind) +}\StringTok{ }\DecValTok{1}\NormalTok{):}\KeywordTok{length}\NormalTok{(cl)]}
-
-\NormalTok{## Plot the input and output}
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{2}\NormalTok{))}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }
-       \DataTypeTok{main =} \StringTok{"phenoDisco input markers"}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{getStockcol}\NormalTok{()[}\DecValTok{1}\NormalTok{:}\DecValTok{6}\NormalTok{])}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"phenoDisco.Input"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{)}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{main =} \StringTok{"phenoDisco output"}\NormalTok{, }\DataTypeTok{col =} \NormalTok{cols)}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{col =} \NormalTok{cols)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotPDres-1.png}
-\caption{plot of chunk plotPDres}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Re-order the colours for the phenoDisco output}
+\hlstd{cl} \hlkwb{<-} \hlkwd{getMarkerClasses}\hlstd{(hl,} \hlstr{"pd"}\hlstd{)}
+\hlstd{cols} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)]}
+\hlstd{ind} \hlkwb{<-} \hlkwd{grep}\hlstd{(}\hlstr{"Pheno"}\hlstd{, cl,} \hlkwc{invert} \hlstd{=} \hlnum{TRUE}\hlstd{)}
+\hlstd{cols[ind]} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)][}\hlnum{1}\hlopt{:}\hlkwd{length}\hlstd{(ind)]}
+\hlstd{cols[}\hlopt{-}\hlstd{ind]} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)][(}\hlkwd{length}\hlstd{(ind)} \hlopt{+} \hlnum{1}\hlstd{)}\hlopt{:}\hlkwd{length}\hlstd{(cl)]}
+
+\hlcom{## Plot the input and output}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,}
+       \hlkwc{main} \hlstd{=} \hlstr{"phenoDisco input markers"}\hlstd{,} \hlkwc{col} \hlstd{=} \hlkwd{getStockcol}\hlstd{()[}\hlnum{1}\hlopt{:}\hlnum{6}\hlstd{])}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{.7}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{main} \hlstd{=} \hlstr{"phenoDisco output"}\hlstd{,} \hlkwc{col} \hlstd{= cols)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{.7}\hlstd{,} \hlkwc{col} \hlstd{= cols)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plotPDres-1}
+
+\end{knitrout}
+  \caption{Results of the novelty detection algorithm.}
+  \label{fig:plotPDres}
 \end{figure}
 
-The five new phenotype data clusters can be extracted and examined. In
-the code chunk below we write the results to a .csv file. We use the
-argument \texttt{fDataCols} to specify which columns of the
+The five new phenotype data clusters can be extracted and examined.
+In the code chunk below we write the results to a .csv file. We use
+the argument \texttt{fDataCols} to specify which columns of the
 \texttt{featureData} to write.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{fData}\NormalTok{(hl)$pd <-}\StringTok{ }\KeywordTok{as.character}\NormalTok{(}\KeywordTok{fData}\NormalTok{(hl)$pd)}
-\KeywordTok{write.exprs}\NormalTok{(hl, }\DataTypeTok{fDataCols =} \StringTok{"pd"}\NormalTok{, }\DataTypeTok{file =} \StringTok{"pd-results.csv"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{","}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-We can also examine the each phenotype intercatively and visualise their
-protein profiles by using the \texttt{pRolocVis} function in the
-\texttt{pRolocGUI}. We found that phenotype 1 was enriched in nucleus
-associated proteins, phenotype 2 in chromatin associated proteins,
-phenotype 3 in cytosolic and phenotypes 4 and 5 in lysosomal and
-endosomal proteins.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{pRolocVis}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"pd"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\hyperdef{}{supervised-machine-learning}{\section{Supervised machine
-learning}\label{supervised-machine-learning}}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{pd} \hlkwb{<-} \hlkwd{as.character}\hlstd{(}\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{pd)}
+\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{fDataCols} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{file} \hlstd{=} \hlstr{"pd-results.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+We can also examine the each phenotype intercatively and visualise
+their protein profiles by using the \texttt{pRolocVis} function in the
+\Biocpkg{pRolocGUI} package. We found that phenotype 1 was
+enriched in nucleus associated proteins, phenotype 2 in chromatin
+associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5
+in lysosomal and endosomal proteins.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{pRolocVis}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\section*{Supervised machine learning}
 
@@
 essential tool for the assignment of proteins to distinct sub-cellular
-niches. Using a set of labelled training examples i.e.~markers, we can
-train a machine learning classifier to learn a mapping between the data
-i.e.~the quantitative protein profiles, and a known localisation. The
-trained classifier can then be used to predict the localisation of a
-protein of unknown localisation, based on its observed protein profile.
-To date, this method has been extensively used in spatial quantitative
-proteomics to assign thousands of proteins to distinct sub-cellular
-niches {[}@hyper; @Groen:2014; @Trotter:2010; @Hall:2009; @Dunkley:2006;
-@Tan:2009{]}.
+niches. Using a set of labelled training examples i.e. markers, we can
+train a machine learning classifier to learn a mapping between the
+data i.e. the quantitative protein profiles, and a known localisation.
+The trained classifier can then be used to predict the localisation of
+a protein of unknown localisation, based on its observed protein
+profile. To date, this method has been extensively used in spatial
+quantitative proteomics to assign thousands of proteins to distinct
+sub-cellular niches
+\cite{hyper,Groen:2014,Trotter:2010,Hall:2009,Dunkley:2006,Tan:2009}.
 
 There are several classification algorithms readily available in
-\texttt{pRoloc}, which are documented in the dedicated
-\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\texttt{pRoloc}
-machine learning techniques vignette}. We find the general tendancy to
-be that it is not the choice of classifier, but the improper
+\Biocpkg{pRoloc}, which are documented in the dedicated
+\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\Biocpkg{pRoloc}
+  machine learning techniques vignette}.  We find the general tendancy
+to be that it is not the choice of classifier, but the improper
 optimisation of the algorithmic parameters, that limits classification
-accuracy. Before employing any classification algorithm and generating a
-model on the training data, one must first find the optimal parameters
-for the algorithm of choice.
-
-\subsection{Optimisation}\label{optimisation}
+accuracy. Before employing any classification algorithm and generating
+a model on the training data, one must first find the optimal
+parameters for the algorithm of choice.
+
+\subsection*{Optimisation}
 
 In the code chunk below we employ the use of a Support Vector Machine
-(SVM) to learn a classifier on the labelled training data. As previously
-mentioned one first needs to train the classifiers parameters before an
-algorithm can be used to predict the class labels of the proteins with
-unknown location. One of the most common ways to optimise the parameters
-of a classifier is to partition the labelled data in to training and
-testing subsets. In this framework parameters are tested via a grid
-search using cross-validation on the training partition. The best
-parameters chosen from the cross-validation stage are then used to build
-a classifier to predict the class labels of the protein profiles on the
-test partition. Observed and expected classication results can be
-compared, and then used to assess how well a given model works by
-getting an estimate of the classiers ability to achieve a good
-generalisation i.e.~that is given an unknown example predict its class
-label with high accuracy. In \texttt{pRoloc} algorithmic performance is
-estimated using stratified 80/20 partitioning for the training/testing
-subsets respectively, in conjuction with five-fold cross-validation in
-order to optimise the free parameters via a grid search. This procedure
-is usually repeated 100 times and then the best parameter(s) are
-selected upon investigation of classifier accuracy, here we use the
-harmonic mean of precision and recall; the macro F1 score. In the code
-chunk below we demonstrate how to optimise the free parameters;
-\texttt{sigma} and \texttt{cost}, of a classical SVM classifier with a
-Gaussian kernel using the function \texttt{svmOptimisation}. As the
-number of labelled instances per class varies from organelle to
-organelle, we can account for class imbalance by setting specific class
-weights when generating the SVM model. Below the weights, \texttt{w} are
-set to be inversely proportional to the class frequencies.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{w <-}\StringTok{ }\KeywordTok{table}\NormalTok{(}\KeywordTok{getMarkers}\NormalTok{(hl, }\DataTypeTok{verbose =} \OtherTok{TRUE}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
-
+(SVM) to learn a classifier on the labelled training data. As
+previously mentioned one first needs to train the classifiers
+parameters before an algorithm can be used to predict the class labels
+of the proteins with unknown location. One of the most common ways to
+optimise the parameters of a classifier is to partition the labelled
+data in to training and testing subsets. In this framework parameters
+are tested via a grid search using cross-validation on the training
+partition. The best parameters chosen from the cross-validation stage
+are then used to build a classifier to predict the class labels of the
+protein profiles on the test partition. Observed and expected
+classication results can be compared, and then used to assess how well
+a given model works by getting an estimate of the classiers ability to
+achieve a good generalisation i.e. that is given an unknown example
+predict its class label with high accuracy. In
+\Biocpkg{pRoloc}, algorithmic performance is estimated using
+stratified 80/20 partitioning for the training/testing subsets
+respectively, in conjuction with five-fold cross-validation in order
+to optimise the free parameters via a grid search. This procedure is
+usually repeated 100 times and then the best parameter(s) are selected
+upon investigation of classifier accuracy, here we use the harmonic
+mean of precision and recall; the macro F1 score. In the code chunk
+below we demonstrate how to optimise the free parameters; \texttt{sigma} and
+\texttt{cost}, of a classical SVM classifier with a Gaussian kernel using the
+function \texttt{svmOptimisation}. As the number of labelled instances per
+class varies from organelle to organelle, we can account for class
+imbalance by setting specific class weights when generating the SVM
+model. Below the weights, \texttt{w} are set to be inversely proportional to
+the class frequencies.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{w} \hlkwb{<-} \hlkwd{table}\hlstd{(}\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{verbose} \hlstd{=} \hlnum{TRUE}\hlstd{))}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{w <-}\StringTok{ }\DecValTok{1}\NormalTok{/w[}\KeywordTok{names}\NormalTok{(w) !=}\StringTok{ "unknown"}\NormalTok{]}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## 100 rounds of optimisation with five-fold cross-validation}
-\NormalTok{params <-}\StringTok{ }\KeywordTok{svmOptimisation}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"markers"}\NormalTok{,}
-                          \DataTypeTok{times =} \DecValTok{100}\NormalTok{, }\DataTypeTok{xval =} \DecValTok{5}\NormalTok{,}
-                          \DataTypeTok{class.weights =} \NormalTok{w)}
-\end{Highlighting}
-\end{Shaded}
+\begin{alltt}
+\hlstd{w} \hlkwb{<-} \hlnum{1}\hlopt{/}\hlstd{w[}\hlkwd{names}\hlstd{(w)} \hlopt{!=} \hlstr{"unknown"}\hlstd{]}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## 100 rounds of optimisation with five-fold cross-validation}
+\hlstd{params} \hlkwb{<-} \hlkwd{svmOptimisation}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{,}
+                          \hlkwc{times} \hlstd{=} \hlnum{100}\hlstd{,} \hlkwc{xval} \hlstd{=} \hlnum{5}\hlstd{,}
+                          \hlkwc{class.weights} \hlstd{= w)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
 As mentioned previously, we reply on the default feature variable
-\texttt{"markers"} to define the class labels and hence can ommit it. To
-use another feature variables, one need to explicitly specify its name
-using the \texttt{fcol} argument (for example
-\texttt{fcol = "markers2"}).
+\texttt{"markers"} to define the class labels and hence can ommit
+it. To use another feature variables, one need to explicitly specify
+its name using the \texttt{fcol} argument (for example \texttt{fcol =
+  "markers2"}).
 
@@
 machine learning optimisation. To assess classifier performance we can
-examine the macro F1 scores and the most frequently chosen parameters. A
-high macro F1 score indicates that the marker proteins in the test
-dataset are consistently and correctly assigned by the the algorithm.
-Often more than one parameter or set of parameters gives rise to the
-best generalisation accuracy. As such it is always important to
-investigate the model parameters and critically assess the best choice.
-The best choice may not be as simple as the parameter set that gives
-rise to the highest macro F1 score and one must be careful to avoid
-overfitting and to choose parameters wisely.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{(best <-}\StringTok{ }\KeywordTok{getParams}\NormalTok{(params))}
-\KeywordTok{plot}\NormalTok{(params)}
-\KeywordTok{levelPlot}\NormalTok{(params)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+examine the macro F1 scores and the most frequently chosen
+parameters. A high macro F1 score indicates that the marker proteins
+in the test dataset are consistently and correctly assigned by the the
+algorithm. Often more than one parameter or set of parameters gives
+rise to the best generalisation accuracy. As such it is always
+important to investigate the model parameters and critically assess
+the best choice. The best choice may not be as simple as the parameter
+set that gives rise to the highest macro F1 score and one must be
+careful to avoid overfitting and to choose parameters wisely.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{(best} \hlkwb{<-} \hlkwd{getParams}\hlstd{(params))}
+\hlkwd{plot}\hlstd{(params)}
+\hlkwd{levelPlot}\hlstd{(params)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/visualiseOptHide-1.png}
-\caption{plot of chunk visualiseOptHide}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/visualiseOptHide-1}
+
+\end{knitrout}
+  \caption{Assessment of the classification model parameter optimisation.}
+  \label{fig:visualisOptHide}
 \end{figure}
@@
 By using the function \texttt{getParams} we can extract the best set of
-parameters. Currently, \texttt{getParams} retrieves the first best set
-is automatically but users are encouraged to critically assess whether
+parameters. Currently, \texttt{getParams} retrieves the first best set is
+automatically but users are encouraged to critically assess whether
 this is the most wise choice by visualising the results using the
-methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method
-for \texttt{GenRegRes} object shows the respective distributions of the
-100 macro F1 scores for the best cost/sigma parameter pairs, and
-\texttt{levelPlot} shows the averaged macro F1 scores, for the full
-range of parameter values. Once we have selected the best parameters we
-can then use them to build a classifier from the labelled marker
-proteins.
-
-\subsection{Classification}\label{classification}
-
-We can use the function \texttt{svmClassification} to return a
-classification result for all unlabelled instances in the dataset
-corresponding to their most likely sub-cellular location. The algorithm
-parameters are passed to the function, along with the class weights. As
-above, \texttt{fcol} can be ignored as we use the labels defined in the
-default \texttt{"markers"} feature variable.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{svmClassification}\NormalTok{(hl, params,}
-                        \DataTypeTok{class.weights =} \NormalTok{w,}
-                        \DataTypeTok{fcol =} \StringTok{"markers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method for \texttt{GenRegRes}
+object shows the respective distributions of the 100 macro F1 scores
+for the best cost/sigma parameter pairs, and \texttt{levelPlot} shows the
+averaged macro F1 scores, for the full range of parameter values. Once
+we have selected the best parameters we can then use them to build a
+classifier from the labelled marker proteins.
+
+\subsection*{Classification}
+
+We can use the function \texttt{svmClassification} to return a classification
+result for all unlabelled instances in the dataset corresponding to
+their most likely sub-cellular location. The algorithm parameters are
+passed to the function, along with the class weights. As above, \texttt{fcol}
+can be ignored as we use the labels defined in the default \texttt{"markers"}
+feature variable.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl, params,}
+                        \hlkwc{class.weights} \hlstd{= w,}
+                        \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 argument to change the size of each point on the plot (where one point
-represents one protein) to be inversely proportional to the SVM score.
-This gives an initial overview of the high scoring localisations from
-the SVM predictions.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## set point size of each protein to be inversely proportional to the}
-\NormalTok{ptsze <-}\StringTok{ }\KeywordTok{exp}\NormalTok{(}\KeywordTok{fData}\NormalTok{(hl)$svm.scores) -}\StringTok{ }\DecValTok{1}
-
-\NormalTok{## plot new predictions}
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{ptsze)}
-\KeywordTok{addLegend}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }\DataTypeTok{where =} \StringTok{"bottomleft"}\NormalTok{, }\DataTypeTok{bty =} \StringTok{"n"}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+represents one protein) to be inversely proportional to the SVM
+score. This gives an initial overview of the high scoring
+localisations from the SVM predictions.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plotSVM-1.png}
-\caption{plot of chunk plotSVM}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## set point size of each protein to be inversely proportional to the}
+\hlstd{ptsze} \hlkwb{<-} \hlkwd{exp}\hlstd{(}\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{svm.scores)} \hlopt{-} \hlnum{1}
+\hlcom{## plot new predictions}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{cex} \hlstd{= ptsze)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{where} \hlstd{=} \hlstr{"bottomleft"}\hlstd{,} \hlkwc{bty} \hlstd{=} \hlstr{"n"}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{.5}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/plotSVM-1}
+
+\end{knitrout}
+  \caption{Classification results.}
+  \label{fig:plotSVM}
 \end{figure}
@@
 The adjustment of the point size intuitively confers important
-information that is more difficult to defined formally, but that we will
-address in the next section. The classifier (SVM in our case, but this
-is also valid of other classifiers) defines boundaries based on the
-labelled marker proteins. These class/organelle boundaries define how
-non-assigned proteins are classified and with what confidence.
-\textbf{TODO} explain how choosing markers positively/negatively affects
-this, and relation with thresholding.
-
-\hyperdef{}{thresholding}{\subsection{Thresholding}\label{thresholding}}
-
-It is common when applying a supervised classification algorithm to set
-a specific score cutoff on which to define new assignments, below which
-classifications are set to unknown/unassigned. This is important as in a
-supervised learning setup proteins can only be predicted to be localised
-to one of the sub-cellular niches that appear in the labelled training
-data. We can not guaranette (and do not expect) that the whole class
-diversity to be represented in the labelled training data as (1) finding
-markers that represent the whole diversity of the cell is challenging
-(especially obtaining dual- and multiply-localised protein markers) and
-(2) many sub-cellular niches contain too few proteins to train on (we
-recommend a minimum of 13 markers per sub-cellular class for stratified
-80/20 partitioning and 5-fold cross-validation - this allows a minimum
-of \textasciitilde{}10 examples for parameter optimisation on the
-training partition i.e.~2 per fold for 5-fold cross-validation, and then
-\textasciitilde{}3 for testing the best parameters on the validation
-set).
-
-Deciding on a threshold is not trivial as classifier scores are heavily
-dependent upon the classifier used and different sub-cellular niches can
-exhibit different score distributions, as highlighted in the boxplot
-below. We recommend users to set class-specific thresholds. In the code
-chunk below we display a boxplot of the score distributions per
-organelle.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## First remove the markers}
-\NormalTok{preds <-}\StringTok{ }\KeywordTok{unknownMSnSet}\NormalTok{(hl)}
-
-\NormalTok{## Plot a boxplot of the scores of each organelle}
-\KeywordTok{par}\NormalTok{(}\DataTypeTok{oma =} \KeywordTok{c}\NormalTok{(}\FloatTok{10.5}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{))}
-\KeywordTok{boxplot}\NormalTok{(svm.scores ~}\StringTok{ }\NormalTok{svm, }\DataTypeTok{data =} \KeywordTok{fData}\NormalTok{(preds), }\DataTypeTok{ylab =} \StringTok{"SVM scores"}\NormalTok{, }\DataTypeTok{las =} \DecValTok{2}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+information that is more difficult to defined formally, but that we
+will address in the next section. The classifier (SVM in our case, but
+this is also valid of other classifiers) defines boundaries based on
+the labelled marker proteins. These class/organelle boundaries define
+how non-assigned proteins are classified and with what confidence.
+
+\textbf{TODO} explain how choosing markers positively/negatively
+affects this, and relation with thresholding.
+
+\subsubsection*{Thresholding}\label{sec:thresholding}
+
+It is common when applying a supervised classification algorithm to
+set a specific score cutoff on which to define new assignments, below
+which classifications are set to unknown/unassigned. This is important
+as in a supervised learning setup proteins can only be predicted to be
+localised to one of the sub-cellular niches that appear in the
+labelled training data. We can not guaranette (and do not expect) that
+the whole class diversity to be represented in the labelled training
+data as (1) finding markers that represent the whole diversity of the
+cell is challenging (especially obtaining dual- and multiply-localised
+protein markers) and (2) many sub-cellular niches contain too few
+proteins to train on (we recommend a minimum of 13 markers per
+sub-cellular class for stratified 80/20 partitioning and 5-fold
+cross-validation - this allows a minimum of ~10 examples for parameter
+optimisation on the training partition i.e. 2 per fold for 5-fold
+cross-validation, and then ~3 for testing the best parameters on the
+validation set). <!--Conversely, when defining markers one must also
+be careful to not label too many markers, as this can lead to
+overfitting and mis-assignment.-->
+
+Deciding on a threshold is not trivial as classifier scores are
+heavily dependent upon the classifier used and different sub-cellular
+niches can exhibit different score distributions, as highlighted in
+the boxplot below. We recommend users to set class-specific
+thresholds.  In the code chunk below we display a boxplot of the score
+distributions per organelle.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/thresholds-1.png}
-\caption{plot of chunk thresholds}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## First remove the markers}
+\hlstd{preds} \hlkwb{<-} \hlkwd{unknownMSnSet}\hlstd{(hl)}
+\hlcom{## Plot a boxplot of the scores of each organelle}
+\hlkwd{par}\hlstd{(}\hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{10.5}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))}
+\hlkwd{boxplot}\hlstd{(svm.scores} \hlopt{~} \hlstd{svm,} \hlkwc{data} \hlstd{=} \hlkwd{fData}\hlstd{(preds),}
+        \hlkwc{ylab} \hlstd{=} \hlstr{"SVM scores"}\hlstd{,} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/thresholds-1}
+
+\end{knitrout}
+  \caption{Visualistion of class-specific classification score distribution.}
+  \label{fig:threshold}
 \end{figure}
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ts <-}\StringTok{ }\KeywordTok{orgQuants}\NormalTok{(hl, }
-                \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{, }
-                \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{,}
-                \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{, }
-                \DataTypeTok{t =} \NormalTok{.}\DecValTok{5}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,}
+                \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,}
+                \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,}
+                \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,}
+                \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
+\end{alltt}
 \begin{verbatim}
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4356698               0.3028843               0.3825420 
+##               0.4383808               0.3028615               0.3856431 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6825826               0.5995243               0.4245311 
+##               0.6859801               0.5995991               0.4350839 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4034908               0.5882168               0.9501241 
+##               0.4307964               0.5935764               0.9500433 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951459               0.7096694               0.3149301 
+##               0.7953738               0.7087561               0.3155507 
 ##         Plasma membrane              Proteasome 
-##               0.7091667               0.4204304
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{getPredictions}\NormalTok{(hl,}
-                     \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{,}
-                     \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{, }
-                     \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{,}
-                     \DataTypeTok{t =} \NormalTok{ts)}
-\end{Highlighting}
-\end{Shaded}
-
+##               0.7171990               0.4185577
+\end{verbatim}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,}
+                     \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,}
+                     \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,}
+                     \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,}
+                     \hlkwc{t} \hlstd{= ts)}
+\end{alltt}
 \begin{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     169                      86 
+##                      87                     169                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     479                      91 
+##                     298                     475                      98 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     125                     522 
+##                      31                     123                     521 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      38 
+##                     229                     344                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     325                     159                    2057
-\end{verbatim}
+##                     319                     157                    2055
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
 The output of \texttt{getPredictons} is the original \texttt{MSnSet}
-dataset with a new feature variable appended to the feature data called
-\texttt{fcol.pred} (i.e.~in our case \texttt{svm.pred}) containing the
-prediction results. The results can also be visualied using
-\texttt{plot2D} function.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{dev.off}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
+dataset with a new feature variable appended to the feature data
+called \texttt{fcol.pred} (i.e. in our case \texttt{svm.pred})
+containing the prediction results. The results can also be visualied
+using \texttt{plot2D} function.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{dev.off}\hlstd{()}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot2D}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"svm.pred"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\end{kframe}
+\end{knitrout}
+
+\begin{figure}[!ht]
     \centering
-\includegraphics{figure/plotres-1.png}
-\caption{plot of chunk plotres}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm.pred"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/plotres-1}
+
+\end{knitrout}
+  \caption{Results of the localisation preductions after thresholding.}
+  \label{fig:plotres}
 \end{figure}
 
-There is also a dedicated interactive application to help users examine
-these distributions in the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
-package. This app can be launched via the \texttt{pRolocVis} function
-and specifying the argument \texttt{app = "classify"} along with the
-relevent \texttt{fcol}, \texttt{scol} and \texttt{mcol} which refer to
-the columns in the feature data that contain the new assignments,
-assignment scores and markers respectively (see also
-\texttt{fvarLabels(svmres)}).
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{library}\NormalTok{(}\StringTok{"pRolocGUI"}\NormalTok{)}
-\KeywordTok{pRolocVis}\NormalTok{(hl, }
-          \DataTypeTok{app =} \StringTok{"classify"}\NormalTok{,}
-          \DataTypeTok{fcol =} \StringTok{"svm"}\NormalTok{,}
-            \DataTypeTok{scol =} \StringTok{"svm.scores"}\NormalTok{,}
-          \DataTypeTok{mcol =} \StringTok{"markers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+There is also a dedicated interactive application to help users
+examine these distributions in the \Biocpkg{pRolocGUI} package.
+This app can be launched via the \texttt{pRolocVis} function and specifying
+the argument \texttt{app = "classify"} along with the relevent \texttt{fcol}, \texttt{scol}
+and \texttt{mcol} which refer to the columns in the feature data that contain
+the new assignments, assignment scores and markers respectively (see
+also \texttt{fvarLabels(svmres)}).
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRolocGUI"}\hlstd{)}
+\hlkwd{pRolocVis}\hlstd{(hl,}
+          \hlkwc{app} \hlstd{=} \hlstr{"classify"}\hlstd{,}
+          \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,}
+                \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,}
+          \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{figure}[!ht]
   \centering
-\includegraphics{./Figures/classify.png}
+  \includegraphics[width=\textwidth]{./Figures/classify.png}
   \caption{The classify application}
+  \label{fig:classifyapp}
 \end{figure}
 
-The data is loaded and displayed on a PCA plot and a boxplot is used to
-display the classifier scores by data class. On the left there is a
+The data is loaded and displayed on a PCA plot and a boxplot is used
+to display the classifier scores by data class. On the left there is a
 sidebar panel with sliders to control the thresholds upon which
 classifications are made. There are two types of cut-off that the user
-can choose from: (1) ``Quantile'' and (2) ``User-defined''. By default,
-when the application is launched quatile scoring is selected and set to
-0.5, the median. The class-specific score thresholds that correspond to
-selecting the desired quantile are shown on as red dots on the boxplot.
-The assignments on the PCA plot are also updated according to the
-selected threshold. The quantile threshold can be set by moving the
-corresponding quantile slider. If one wished to set their own cut-offs
-the ``User-defined'' radio button must be selected and then the sliders
-for defining user-specified scores become active and the scores and
-highlighted on the boxplot by blue dots. For more information we refer
-users to the
-\emph{\href{http://bioconductor.org/packages/pRolocGUI}{pRolocGUI}}
+can choose from: (1) "Quantile" and (2) "User-defined". By default,
+when the application is launched quatile scoring is selected and set
+to 0.5, the median. The class-specific score thresholds that
+correspond to selecting the desired quantile are shown on as red dots
+on the boxplot. The assignments on the PCA plot are also updated
+according to the selected threshold. The quantile threshold can be set
+by moving the corresponding quantile slider. If one wished to set
+their own cut-offs the "User-defined" radio button must be selected
+and then the sliders for defining user-specified scores become active
+and the scores and highlighted on the boxplot by blue dots. For more
+information we refer users to the \Biocpkg{pRolocGUI}
 tutorial
@@
 
-In TL one typically has a primary task that they wish to solve, and some
-complementary (often heterogeneous) auxiliary information that is
+In TL one typically has a primary task that they wish to solve, and
+some complementary (often heterogeneous) auxiliary information that is
 related to the primary learning objective, that can be used to help
-solve the primary goal. For example, here our primary task is to assign
-proteins to their sub-cellular niche with high generalisation accuracy
-from data collected from quantitative MS-based experiments. In the
-example below we extract Gene Ontology (GO) information to use as an
-auxiliary data source to help solve our task of protein localisation
-prediction.
-
-Using the functions \texttt{setAnnotationParams} and \texttt{makeGoSet}
-we can contruct an auxiliary \texttt{MSnSet} of GO terms, from the
-primary data's features i.e.~the protein accession numbers. All the GO
-terms associated to each accession number are retrieved and used to
-create a binary matrix where a 1 (0) at position (i; j) indicates that
-term j has (not) been used to annotate protein i. The GO terms are
-retrieved from an appropriate repository using the
-\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}} package.
-The specific
-\emph{\href{http://bioconductor.org/packages/Biomart}{Biomart}}
-repository and query will depend on the species under study and the type
-of identifiers. The first step is to construct the annotation parameters
+solve the primary goal. For example, here our primary task is to
+assign proteins to their sub-cellular niche with high generalisation
+accuracy from data collected from quantitative MS-based experiments.
+In the example below we extract Gene Ontology (GO) information to use
+as an auxiliary data source to help solve our task of protein
+localisation prediction.
+
+Using the functions \texttt{setAnnotationParams} and
+\texttt{makeGoSet} we can contruct an auxiliary \texttt{MSnSet} of GO
+terms, from the primary data's features i.e. the protein accession
+numbers. All the GO terms associated to each accession number are
+retrieved and used to create a binary matrix where a 1 (0) at position
+(i; j) indicates that term j has (not) been used to annotate protein
+i. The GO terms are retrieved from an appropriate repository using the
+\Biocpkg{biomaRt} package. The specific Biomart repository and query
+will depend on the species under study and the type of
+identifiers. The first step is to construct the annotation parameters
 that will enable to perform the query, this is done using
-\texttt{setAnnotationParams}. Typing into the R console
-\texttt{par \textless{}- setAnnotationParams()} will present two menus,
-firstly asking you to identify the species of study, and then what type
-of identifier you have used to annotate the proteins in your
-\texttt{MSnSet}. It is also possible to pass patterns to match the
-species e.g.~in the code chunk below we pass ``Mus musculus'', and the
-identifier type for our data (see \texttt{featureNames(hl)}) which is
-``Uniprot/Swissprot'', for the \texttt{Biocpkg("Biomart")} query.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{par <-}\StringTok{ }\KeywordTok{setAnnotationParams}\NormalTok{(}\DataTypeTok{inputs =} \KeywordTok{c}\NormalTok{(}\StringTok{"Mus musculus"}\NormalTok{, }
-                                      \StringTok{"UniProt/Swissprot"}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
+\texttt{setAnnotationParams}. Typing into the R console \texttt{par <-
+  setAnnotationParams()} will present two menus, firstly asking you to
+identify the species of study, and then what type of identifier you
+have used to annotate the proteins in your \texttt{MSnSet}. It is also
+possible to pass patterns to match the species e.g. in the code chunk
+below we pass "Mus musculus", and the identifier type for our data
+(see \texttt{featureNames(hl)}) which is "Uniprot/Swissprot", for the
+Biomart query.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{par} \hlkwb{<-} \hlkwd{setAnnotationParams}\hlstd{(}\hlkwc{inputs} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"Mus musculus"}\hlstd{,}
+                                      \hlstr{"UniProt/Swissprot"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gocc <-}\StringTok{ }\KeywordTok{makeGoSet}\NormalTok{(hl, }\DataTypeTok{params =} \NormalTok{par, }
-                  \DataTypeTok{namespace =} \StringTok{"cellular_component"}\NormalTok{, }
-                  \DataTypeTok{evidence =} \OtherTok{NULL}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-The function \texttt{makeGoSet} uses the
-\emph{\href{http://bioconductor.org/packages/biomaRt}{biomaRt}} package
-to query the relevent database (e.g.~Ensembl, Uniprot) for GO terms. All
-GO terms that have been observed for the 5032 proteins in the hyperLOPIT
-dataset are retieved. Users should note that the number of GO terms
-retreived is also dependent on the database version queried and thus is
-always subject to change. We find it is common to see many GO terms with
-only one protein assigned to that term, these such terms generally do
-not bring and information generally for building the classifier and thus
-we can remove such GO terms by using the function
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{gocc} \hlkwb{<-} \hlkwd{makeGoSet}\hlstd{(hl,} \hlkwc{params} \hlstd{= par,}
+                  \hlkwc{namespace} \hlstd{=} \hlstr{"cellular_component"}\hlstd{,}
+                  \hlkwc{evidence} \hlstd{=} \hlkwa{NULL}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+The function \texttt{makeGoSet} uses the \Biocpkg{biomaRt} package to
+query the relevent database (e.g. Ensembl, Uniprot) for GO terms. All
+GO terms that have been observed for the 5032 proteins in the
+hyperLOPIT dataset are retieved. Users should note that the number of
+GO terms retreived is also dependent on the database version queried
+and thus is always subject to change. We find it is common to see many
+GO terms with only one protein assigned to that term, these such terms
+generally do not bring and information generally for building the
+classifier and thus we can remove such GO terms by using the function
 \texttt{filterBinMSnSet}.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{gocc <-}\StringTok{ }\KeywordTok{filterBinMSnSet}\NormalTok{(hl)}
-\end{Highlighting}
-\end{Shaded}
-
-Now we have generated our auxiliary data we can use the \emph{k}-NN
-implementation of transfer learning available in
-\texttt{Biocpkg("pRoloc")} to integrate this with our primary MS-based
-quantitative proteomics data using the functions
-\texttt{knntlOptimisation} to estimate the free-parameters for the
-integration, and \texttt{knntlClassification} to do the predictions. We
-have shown that using transfer learning in the context of spatial
-proteomics results in the assignment of proteins to sub-cellular niches
-with a higher generalisation accuracy than using standard supervised
-machine learning with a single source of information
-{[}@Breckels:2016{]}.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{gocc} \hlkwb{<-} \hlkwd{filterBinMSnSet}\hlstd{(hl)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Now we have generated our auxiliary data we can use the \textit{k}-NN
+implementation of transfer learning available in \Biocpkg{pRoloc} to
+integrate this with our primary MS-based quantitative proteomics data
+using the functions \texttt{knntlOptimisation} to estimate the
+free-parameters for the integration, and \texttt{knntlClassification}
+to do the predictions. We have shown that using transfer learning in
+the context of spatial proteomics results in the assignment of
+proteins to sub-cellular niches with a higher generalisation accuracy
+than using standard supervised machine learning with a single source
+of information \cite{Breckels:2016}.
 
 The first step, as with any machine learning algorithm, is to optimise
-any free paramaters of the classifier. For the \emph{k}-NN TL classifier
-there are two sets of parameters that need optimising: the first set are
-the $k$'s for the primary and auxiliary data sources required for the
-nearest neighbour calculations for each data source. The second set of
-parameters (which is noted by a vector of $\theta$ weights) that require
-optimising are the class weights, one per subcellular niche, that
-control the proportion of primary and auxiliary data to use for
-learning. A weight can take any real value number between 0 and 1. A
-weight of $\theta = 1$ indicates that all weight is given to the primary
-data (and this implicitly implies that a weight of $1 - \theta$ is given
-to the auxiliary data, similarly a weight of $\theta = 0$ implies that
-all weight is given to the auxiliary data (so 0 is given to the primary
-source). If we conduct a parameter seach and test weights
-$\theta = {0, 1/3, 2/3, 1}$ for each class, and if we have, for example
-10 subcellular niches, this will result in \texttt{4\^{}10} different
-combinations of parameters to test. The parameter optimisation is
-therefore time consuming and as such we recommend you make use of a
-computing cluster (code and submissing scripts are also available in the
-supporting information). The markers in the \texttt{hl} dataset contain
-14 subcellular classes. If we examine these markers and classes on the
-PCA plot above we can see that in particular the two ribosomes and two
-nuclear compartments are highly separated along the first two
-components, this is also evident from the profiles plot which gives us a
-good indication that these subcellular niches are well-resolved in the
-hyperLOPIT dataset. Transfer learning is particularly useful for classes
-that are not as well separated, we find that subcellular niches that are
+any free paramaters of the classifier. For the *k*-NN TL classifier
+there are two sets of parameters that need optimising: the first set
+are the $k$'s for the primary and auxiliary data sources required for
+the nearest neighbour calculations for each data source. The second
+set of parameters (which is noted by a vector of $\theta$ weights)
+that require optimising are the class weights, one per subcellular
+niche, that control the proportion of primary and auxiliary data to
+use for learning. A weight can take any real value number between 0
+and 1. A weight of $\theta = 1$ indicates that all weight is given to
+the primary data (and this implicitly implies that a weight of
+$1 - \theta$ is given to the auxiliary data, similarly a weight of
+$\theta = 0$ implies that all weight is given to the auxiliary data
+(so 0 is given to the primary source). If we conduct a parameter seach
+and test weights $\theta = {0, 1/3, 2/3, 1}$ for each class, and if we
+have, for example 10 subcellular niches, this will result in
+\texttt{4$^{10}$} different combinations of parameters to test. The
+parameter optimisation is therefore time consuming and as such we
+recommend you make use of a computing cluster (code and submissing
+scripts are also available in the supporting information). The markers
+in the \texttt{hl} dataset contain 14 subcellular classes. If we
+examine these markers and classes on the PCA plot above we can see
+that in particular the two ribosomes and two nuclear compartments are
+highly separated along the first two components, this is also evident
+from the profiles plot which gives us a good indication that these
+subcellular niches are well-resolved in the hyperLOPIT
+dataset. Transfer learning is particularly useful for classes that are
+not as well separated, we find that subcellular niches that are
 well-separated under hyperLOPIT and LOPIT obtain a class score of 1
-(i.e.~use only primary data from transfer learning
-{[}@Breckels:2016{]}). Therefore, for the optimisation stage of the
-analyses we can already infer a subcellular class weight of 1 for these
-niches and only optimise over the remaining organelles. This can
-signifciantly cut down optimisation time as by removing these 4 classes
-from the optimisation (and not the classification) we only have
-\texttt{4\^{}10} class weight combinations to consider instead of
-\texttt{4\^{}14} combinations.
+(i.e. use only primary data from transfer learning
+\cite{Breckels:2016}). Therefore, for the optimisation stage of the
+analyses we can already infer a subcellular class weight of 1 for
+these niches and only optimise over the remaining organelles.  This
+can signifciantly cut down optimisation time as by removing these 4
+classes from the optimisation (and not the classification) we only
+have \texttt{4$^{10}$} class weight combinations to consider instead
+of \texttt{4$^{14}$} combinations.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## create new markers column for tl markers}
-\KeywordTok{fData}\NormalTok{(hl)$tlmarkers <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(hl)$markers}
-\KeywordTok{fData}\NormalTok{(gocc)$tlmarkers <-}\StringTok{ }\KeywordTok{fData}\NormalTok{(gocc)$markers}
-
-\NormalTok{## Remove 4 classes}
-\NormalTok{torm  <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"40S Ribosome"}\NormalTok{, }\StringTok{"60S Ribosome"}\NormalTok{,}
-           \StringTok{"Nucleus - Chromatin"}\NormalTok{, }
-           \StringTok{"Nucleus - Non-chromatin"}\NormalTok{)}
-\NormalTok{for (i in }\KeywordTok{seq}\NormalTok{(torm)) \{}
-  \NormalTok{hl <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(hl, }\DataTypeTok{from =} \NormalTok{torm[i], }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-  \NormalTok{gocc <-}\StringTok{ }\KeywordTok{fDataToUnknown}\NormalTok{(gocc, }\DataTypeTok{from =} \NormalTok{torm[i], }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\NormalTok{\}}
-\KeywordTok{getMarkerClasses}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## create new markers column for tl markers}
+\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{tlmarkers} \hlkwb{<-} \hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{markers}
+\hlkwd{fData}\hlstd{(gocc)}\hlopt{$}\hlstd{tlmarkers} \hlkwb{<-} \hlkwd{fData}\hlstd{(gocc)}\hlopt{$}\hlstd{markers}
+
+\hlcom{## Remove 4 classes}
+\hlstd{torm}  \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"40S Ribosome"}\hlstd{,} \hlstr{"60S Ribosome"}\hlstd{,}
+           \hlstr{"Nucleus - Chromatin"}\hlstd{,}
+           \hlstr{"Nucleus - Non-chromatin"}\hlstd{)}
+\hlkwa{for} \hlstd{(i} \hlkwa{in} \hlkwd{seq}\hlstd{(torm)) \{}
+  \hlstd{hl} \hlkwb{<-} \hlkwd{fDataToUnknown}\hlstd{(hl,} \hlkwc{from} \hlstd{= torm[i],} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+  \hlstd{gocc} \hlkwb{<-} \hlkwd{fDataToUnknown}\hlstd{(gocc,} \hlkwc{from} \hlstd{= torm[i],} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\hlstd{\}}
+\hlkwd{getMarkerClasses}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{getMarkerClasses}\NormalTok{(gocc, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlkwd{getMarkerClasses}\hlstd{(gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-Optimisation stage 1: run \texttt{knnOptimisation} to get the best $k$'s
-for each data source.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## get best k's}
-\NormalTok{kpopt <-}\StringTok{ }\KeywordTok{knnOptimisation}\NormalTok{(hl, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\NormalTok{kaopt <-}\StringTok{ }\KeywordTok{knnOptimisation}\NormalTok{(gocc, }\DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{)}
-\KeywordTok{plot}\NormalTok{(kpopt)}
-\KeywordTok{plot}\NormalTok{(kaopt)}
-\end{Highlighting}
-\end{Shaded}
-
-From examining the parameter seach plots we find the best $k$'s for both
-the primary and auxiliary are 3.
+\end{kframe}
+\end{knitrout}
+
+Optimisation stage 1: run \texttt{knnOptimisation} to get the best
+$k$'s for each data source.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## get best k's}
+\hlstd{kpopt} \hlkwb{<-} \hlkwd{knnOptimisation}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\hlstd{kaopt} \hlkwb{<-} \hlkwd{knnOptimisation}\hlstd{(gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\hlkwd{plot}\hlstd{(kpopt)}
+\hlkwd{plot}\hlstd{(kaopt)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+From examining the parameter seach plots we find the best $k$'s for
+both the primary and auxiliary are 3.
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## Set appropriate parallelisation backend and  }
-\NormalTok{## number of workers for the tl}
-\NormalTok{par <-}\StringTok{ }\KeywordTok{SnowParam}\NormalTok{(255L, }\DataTypeTok{type =} \StringTok{"MPI"}\NormalTok{)}
-
-\NormalTok{## Now peform tl optimisation }
-\NormalTok{tlopt <-}\StringTok{ }\KeywordTok{knntlOptimisation}\NormalTok{(hl, gocc,}
-                           \DataTypeTok{fcol =} \StringTok{"tlmarkers"}\NormalTok{,}
-                           \DataTypeTok{length.out =} \DecValTok{4}\NormalTok{, }
-                           \DataTypeTok{times =} \DecValTok{50}\NormalTok{,}
-                           \DataTypeTok{xval =} \DecValTok{5}\NormalTok{, }\DataTypeTok{k =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{),}
-                           \DataTypeTok{BPPARAM =} \NormalTok{par}
-  \NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Set appropriate parallelisation backend and  }
+\hlcom{## number of workers for the tl}
+\hlstd{par} \hlkwb{<-} \hlkwd{SnowParam}\hlstd{(}\hlnum{255L}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"MPI"}\hlstd{)}
+
+\hlcom{## Now peform tl optimisation }
+\hlstd{tlopt} \hlkwb{<-} \hlkwd{knntlOptimisation}\hlstd{(hl, gocc,}
+                           \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{,}
+                           \hlkwc{length.out} \hlstd{=} \hlnum{4}\hlstd{,}
+                           \hlkwc{times} \hlstd{=} \hlnum{50}\hlstd{,}
+                           \hlkwc{xval} \hlstd{=} \hlnum{5}\hlstd{,} \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{),}
+                           \hlkwc{BPPARAM} \hlstd{= par)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{plot}\NormalTok{(tlopt)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-\includegraphics{figure/plottl-1.png}
-\caption{plot of chunk plottl}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot}\hlstd{(tlopt)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.6\textwidth]{figure/plottl-1}
+
+\end{knitrout}
+  \caption{Visualisation of the transfer learning parameter optimisation procedure.}
+  \label{fig:plottl}
 \end{figure}
@@
 One of the benefits of the algorithm is the ability to manually select
-weights for each class. In the optimisation above, for time constraints,
-we removed the two ribosomal subunits and the two nulcear compartments,
-and therefore in the code chunk below when we extract the best
-parameters, these subcellular niches are not included. To include these
-4 subcellular niches in the next classification step we must include
-them in the parameters so we infer a weight of 1 for each of these
-niches as we know they are well resolved in hyperLOPIT. We then re-order
-the weights according to \texttt{getMarkerClasses} and perform the
-classification using the function \texttt{knntlClassification}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## best parameters for the 10 classes}
-\NormalTok{(bestpar <-}\StringTok{ }\KeywordTok{getParams}\NormalTok{(tlopt))}
-\end{Highlighting}
-\end{Shaded}
-
+weights for each class. In the optimisation above, for time
+constraints, we removed the two ribosomal subunits and the two nulcear
+compartments, and therefore in the code chunk below when we extract
+the best parameters, these subcellular niches are not included. To
+include these 4 subcellular niches in the next classification step we
+must include them in the parameters so we infer a weight of 1 for each
+of these niches as we know they are well resolved in hyperLOPIT. We
+then re-order the weights according to \texttt{getMarkerClasses} and
+perform the classification using the function
+\texttt{knntlClassification}.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## best parameters for the 10 classes}
+\hlstd{(bestpar} \hlkwb{<-} \hlkwd{getParams}\hlstd{(tlopt))}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## add weights for classes not included in the optimisation}
-\NormalTok{otherweights <-}\StringTok{ }\KeywordTok{rep}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{4}\NormalTok{)}
-\KeywordTok{names}\NormalTok{(otherweights) <-}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"40S Ribosome"}\NormalTok{, }\StringTok{"60S Ribosome"}\NormalTok{, }
-                         \StringTok{"Nucleus - Chromatin"}\NormalTok{, }
-                         \StringTok{"Nucleus - Non-chromatin"}\NormalTok{)}
-\NormalTok{(bestpar <-}\StringTok{ }\KeywordTok{c}\NormalTok{(bestpar, otherweights))}
-\end{Highlighting}
-\end{Shaded}
-
+\begin{alltt}
+\hlcom{## add weights for classes not included in the optimisation}
+\hlstd{otherweights} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{4}\hlstd{)}
+\hlkwd{names}\hlstd{(otherweights)} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"40S Ribosome"}\hlstd{,} \hlstr{"60S Ribosome"}\hlstd{,}
+                         \hlstr{"Nucleus - Chromatin"}\hlstd{,}
+                         \hlstr{"Nucleus - Non-chromatin"}\hlstd{)}
+\hlstd{(bestpar} \hlkwb{<-} \hlkwd{c}\hlstd{(bestpar, otherweights))}
+\end{alltt}
 \begin{verbatim}
@@
 \end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## re-order classes }
-\NormalTok{bestpar <-}\StringTok{ }\NormalTok{bestpar[}\KeywordTok{getMarkerClasses}\NormalTok{(hl)]}
-
-\NormalTok{## Do the classification}
-\NormalTok{hl <-}\StringTok{ }\KeywordTok{knntlClassification}\NormalTok{(hl, gocc, }\DataTypeTok{bestTheta =} \NormalTok{bestpar, }
-                          \DataTypeTok{k =} \KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\DecValTok{3}\NormalTok{))}
-\end{Highlighting}
-\end{Shaded}
+\begin{alltt}
+\hlcom{## re-order classes }
+\hlstd{bestpar} \hlkwb{<-} \hlstd{bestpar[}\hlkwd{getMarkerClasses}\hlstd{(hl)]}
+
+\hlcom{## Do the classification}
+\hlstd{hl} \hlkwb{<-} \hlkwd{knntlClassification}\hlstd{(hl, gocc,} \hlkwc{bestTheta} \hlstd{= bestpar,}
+                          \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 As described and demonstrated already above, PCA is a valuable and
-powerful method for data visualisation and quality control. We do not do
-any unsupervised clustering in the frame of the above spatial proteomics
-workflow as although clustering can be useful for grouping labelled data
-into categories e.g.~see the \texttt{mrkHClust} function, we do not find
-it adaquete for spatial proteomics data analysis. We find supervised
-learning more suited to the task of protein localisation prediction in
-which we use high-quality curated marker proteins to build a classifier,
-instead of using an entirely unsupervised approach to look for clusters
-and then look for enrichment of organelles and complexes. In the latter
-we do not make good use of valuable prior knowledge, and in our
-experience unsupervised clustering can be extremely difficult due poor
-estimates of the number of clusters that may appear in the data.
-
-\section{Writing and exporting data}\label{writing-and-exporting-data}
-
-A \texttt{MSnSet} can be exported from R using the \texttt{write.exprs}
-function. This function writes the expression values to a tab (the
-defualt, or other type e.g.e.g.~comma, as specified in
-\texttt{write,table}) separated file. There argument \texttt{fDataCols}
-can be used to specify which \texttt{featureData} columns (as column
-names, column number or \texttt{logical}) to append to the right of the
-expression matrix.
-
-In the below code chunk we write the \texttt{hl} object to a csv file.
-The \texttt{file} argument is used to specify the file path, the
-\texttt{sep} argument specifies the field separator string, here we use
-a comma, finally as we want to write all the information in the
+powerful method for data visualisation and quality control. We do not
+do any unsupervised clustering in the frame of the above spatial
+proteomics workflow as although clustering can be useful for grouping
+labelled data into categories e.g. see the \texttt{mrkHClust}
+function, we do not find it adaquete for spatial proteomics data
+analysis. We find supervised learning more suited to the task of
+protein localisation prediction in which we use high-quality curated
+marker proteins to build a classifier, instead of using an entirely
+unsupervised approach to look for clusters and then look for
+enrichment of organelles and complexes. In the latter we do not make
+good use of valuable prior knowledge, and in our experience
+unsupervised clustering can be extremely difficult due poor estimates
+of the number of clusters that may appear in the data.
+
+\section*{Writing and exporting data}
+
+A \texttt{MSnSet} can be exported from R using the
+\texttt{write.exprs} function.  This function writes the expression
+values to a tab (the defualt, or other type e.g.e.g. comma, as
+specified in \texttt{write,table}) separated file. There argument
+\texttt{fDataCols} can be used to specify which \texttt{featureData}
+columns (as column names, column number or \texttt{logical}) to append
+to the right of the expression matrix.
+
+In the below code chunk we write the \texttt{hl} object to a csv
+file. The \texttt{file} argument is used to specify the file path, the
+\texttt{sep} argument specifies the field separator string, here we
+use a comma, finally as we want to write all the information in the
 \texttt{featureData} to the file, as well as the expression data, we
-specify \texttt{fDataCols = 1:ncol(hl)} i.e.~write everything in the
+specify \texttt{fDataCols = 1:ncol(hl)} i.e. write everything in the
 \texttt{featureData} to the file \texttt{"hl.csv"}.
 
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{write.exprs}\NormalTok{(hl, }\DataTypeTok{file =} \StringTok{"hl.csv"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{","}\NormalTok{, }
-            \DataTypeTok{fDataCols =} \DecValTok{1}\NormalTok{:}\KeywordTok{ncol}\NormalTok{(hl))}
-\end{Highlighting}
-\end{Shaded}
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{file} \hlstd{=} \hlstr{"hl.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{,}
+            \hlkwc{fDataCols} \hlstd{=} \hlnum{1}\hlopt{:}\hlkwd{ncol}\hlstd{(hl))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
 Individual R objects, such a parameters from the machine learning
-optimisations, can be save using the standard \texttt{save} function in
-R. For example, to save and then re-load the parameters from the SVM
-optimisation,
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{## To save the parameters as an R object}
-\KeywordTok{save}\NormalTok{(params, }\DataTypeTok{file =} \StringTok{"svmparams.rda"}\NormalTok{)}
-
-\NormalTok{## To re-load after saving}
-\KeywordTok{load}\NormalTok{(}\DataTypeTok{file =} \StringTok{"svmparams.rda"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\section{Session information}\label{session-information}
-
-The function \texttt{sessionInfo} provides a summary of all packages and
-versions used to generate this document. This enables us to record the
-exact state of our session that lead to these exact results. Conversely,
-if the script stops working of if it returns different results, we are
-in a position to re-generate the original results using the adequate
-software versions and retrace changes in the software that lead to
-failure and/or different results.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\KeywordTok{sessionInfo}\NormalTok{()}
-\end{Highlighting}
-\end{Shaded}
-
+optimisations, can be save using the standard \texttt{save} function
+in R.  For example, to save and then re-load the parameters from the
+SVM optimisation,
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## To save the parameters as an R object}
+\hlkwd{save}\hlstd{(params,} \hlkwc{file} \hlstd{=} \hlstr{"svmparams.rda"}\hlstd{)}
+
+\hlcom{## To re-load after saving}
+\hlkwd{load}\hlstd{(}\hlkwc{file} \hlstd{=} \hlstr{"svmparams.rda"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\section*{Session information}
+
+The function \texttt{sessionInfo} provides a summary of all packages
+and versions used to generate this document. This enables us to record
+the exact state of our session that lead to these exact
+results. Conversely, if the script stops working of if it returns
+different results, we are in a position to re-generate the original
+results using the adequate software versions and retrace changes in
+the software that lead to failure and/or different results.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{sessionInfo}\hlstd{()}
+\end{alltt}
 \begin{verbatim}
@@
 ## [64] rpart_4.1-10          stringi_1.1.1         RSQLite_1.0.0        
-## [67] genefilter_1.55.2     randomForest_4.6-12   foreach_1.4.3        
-## [70] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
-## [73] mzID_1.11.2           evaluate_0.9          lattice_0.20-34      
-## [76] htmlwidgets_0.7       gbm_2.1.1             plyr_1.8.4           
-## [79] magrittr_1.5          R6_2.1.3              DBI_0.5-1            
-## [82] whisker_0.3-2         mgcv_1.8-15           survival_2.39-5      
-## [85] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
-## [88] msdata_0.12.1         car_2.1-3             mlbench_2.1-1        
-## [91] grid_3.3.1            FNN_1.1               threejs_0.2.2        
-## [94] digest_0.6.10         diptest_0.75-7        xtable_1.8-2         
-## [97] httpuv_1.3.3          munsell_0.4.3
-\end{verbatim}
-
-It is always important to include session information details along with
-a \href{http://adv-r.had.co.nz/Reproducibility.html}{short reproducible
-example} highlighting the problem or
+## [67] highr_0.6             genefilter_1.55.2     randomForest_4.6-12  
+## [70] foreach_1.4.3         e1071_1.6-7           prabclus_2.2-6       
+## [73] bitops_1.0-6          mzID_1.11.2           evaluate_0.9         
+## [76] lattice_0.20-34       htmlwidgets_0.7       gbm_2.1.1            
+## [79] plyr_1.8.4            magrittr_1.5          R6_2.1.3             
+## [82] DBI_0.5-1             whisker_0.3-2         mgcv_1.8-15          
+## [85] survival_2.39-5       RCurl_1.95-4.8        nnet_7.3-12          
+## [88] tibble_1.2            msdata_0.12.1         car_2.1-3            
+## [91] mlbench_2.1-1         grid_3.3.1            FNN_1.1              
+## [94] threejs_0.2.2         digest_0.6.10         diptest_0.75-7       
+## [97] xtable_1.8-2          httpuv_1.3.3          munsell_0.4.3
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+It is always important to include session information details along
+with a \href{http://adv-r.had.co.nz/Reproducibility.html}{short
+  reproducible example} highlighting the problem or
 \href{https://support.bioconductor.org/}{question} at hand.
+
+\bigskip
+
+\subsection*{Author contributions}
+
+\textbf{TODO} In order to give appropriate credit to each author of an
+article, the individual contributions of each author to the manuscript
+should be detailed in this section. We recommend using author initials
+and then stating briefly how they contributed.
+
+\subsection*{Competing interests}
+
+No conflict of interest.
+
+\subsection*{Grant information}
+
+L.M.B is supported by a Wellcome Trust Technology Development Grant
+(grant number 108467/Z/15/Z). L.G. is supported by the BBSRC Strategic
+Longer and Larger grant (Award BB/L002817/1).
+
+\subsection*{Acknowledgements}
+
+The authors would like to thank the Cambridge Computational Biology
+Institute and the Centre for Mathematical Sciences for hosting them
+when writing this article. The authors would like to thank Dr Claire
+M. Mulvey for helpful comments on early versions of this
+manuscript. The authors would also like to thank Dr. Stuart Rankin
+from the High Performance Computing Service for his support. Part of
+this work was performed using the Darwin Supercomputer of the
+University of Cambridge High Performance Computing Service, provided
+by Dell Inc. using Strategic Research Infrastructure Funding from the
+Higher Education Funding Council for England and funding from the
+Science and Technology Facilities Council.

### 2016-10-01 bioc-workflow.tex
contributions, competing interests

@@
 \includegraphics[width=.8\textwidth]{figure/plotSVM-1}
+
+%% $
 
@@
 
-\textbf{TODO} In order to give appropriate credit to each author of an
-article, the individual contributions of each author to the manuscript
-should be detailed in this section. We recommend using author initials
-and then stating briefly how they contributed.
+All authors developed the software presented in this workflow and
+wrote and approved the manuscript.
 
@@
 
-No conflict of interest.
+The authors declare that they have no competing interest.
 
@@
 
-L.M.B is supported by a Wellcome Trust Technology Development Grant
-(grant number 108467/Z/15/Z). L.G. is supported by the BBSRC Strategic
+LMB is supported by a Wellcome Trust Technology Development Grant
+(grant number 108467/Z/15/Z). LG is supported by the BBSRC Strategic
 Longer and Larger grant (Award BB/L002817/1).

### 2016-10-02 bioc-workflow.Rnw
last changes to rnw (not tex) and remove/ammend todos

@@
 
-**TODO** There are other dimensionality reduction methods available in
-the \texttt{plot2D} function, which can be parametrised with the
-\texttt{method} argument.  Advantages, disadvantages? See
-\href{https://gist.github.com/mikelove/74bbf5c41010ae1dc94281cface90d32}
-very interesting gist with intriguing tSNE examples. (This might be
-better to come after markers.)
-
 It is also useful to visualise the relative intensities along the
@@
 
-\textbf{TODO} Possibly mention \texttt{plot3D} and \texttt{mrkHClust}
-(also mentioned later), although these are only available in the
-development version.
+Finally, in addition to \texttt{plot2D}, the \texttt{plot3D} function
+allows to interactively explore a 3-dimensional plot of the data.
 

### 2016-10-05 bioc-workflow.tex
re-compile

@@
 ## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
-## Subset [6725,20][5032,20] Thu Sep 29 21:02:54 2016 
-## Removed features with more than 0 NAs: Thu Sep 29 21:02:54 2016 
-## Dropped featureData's levels Thu Sep 29 21:02:54 2016 
+## Subset [6725,20][5032,20] Wed Oct  5 09:13:01 2016 
+## Removed features with more than 0 NAs: Wed Oct  5 09:13:01 2016 
+## Dropped featureData's levels Wed Oct  5 09:13:01 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Sep 29 21:02:55 2016 
-## Added markers from  'pdres' marker vector. Thu Sep 29 21:02:56 2016 
+## Added markers from  'mrk' marker vector. Wed Oct  5 09:13:02 2016 
+## Added markers from  'pdres' marker vector. Wed Oct  5 09:13:03 2016 
 ##  MSnbase version: 1.99.2
@@
 \includegraphics[width=.8\textwidth]{figure/plotSVM-1}
-
-%% $
 
@@
 ## [85] survival_2.39-5       RCurl_1.95-4.8        nnet_7.3-12          
-## [88] tibble_1.2            msdata_0.12.1         car_2.1-3            
+## [88] tibble_1.2            msdata_0.12.2         car_2.1-3            
 ## [91] mlbench_2.1-1         grid_3.3.1            FNN_1.1              

### 2016-10-08 bioc-workflow.Rnw
re-compile
some changes in intro

@@
 mass spectrometry data. We will focus on a recent pluripotent mouse
-embryonic stem cells experiment \cite{hyper}. Additional annotated and
-pre-formatted datasets from various species are readily available in
-the \Biocexptpkg{pRolocdata} package.
+embryonic stem cells experiment \cite{hyper}. These data, as well as
+additional annotated and pre-formatted datasets from various species
+are readily available in the \Biocexptpkg{pRolocdata} package.
 
@@
 a 20-plex dataset for the analysis.  As discussed in the manuscript
-\cite{hyper} as it has been shown that combining replicates across
-from different gradients can increase spatial resolution
+\cite{hyper}, it has been shown that combining replicates across from
+different gradients can increase spatial resolution
 \cite{Trotter:2010}. The combination of replicates resulted in 5032
@@
 @  
-  \caption{Heatmap of missing values}
+\caption{Heatmap of missing values. Note that the features are
+  re-ordered to highlight cluster of proteins with similar numbers of
+  missing values.}
   \label{fig:namap}
@@
 
-\subsubsection*{Thresholding}\label{sec:thresholding}
+\subsection*{Thresholding}\label{sec:thresholding}
 
@@
 
-\subsection*{Transfer learning}
+\section*{Transfer learning}
 
@@
 
-\subsection*{Unsupervised machine learning}
+\section*{Unsupervised machine learning}
 

### 2016-10-08 bioc-workflow.tex
some changes in intro

@@
 \end{knitrout}
-  \caption{Heatmap of missing values}
+\caption{Heatmap of missing values. Note that the features are
+  re-ordered to highlight cluster of proteins with similar numbers of
+  missing values.}
   \label{fig:namap}
@@
 ## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
-## Subset [6725,20][5032,20] Wed Oct  5 09:13:01 2016 
-## Removed features with more than 0 NAs: Wed Oct  5 09:13:01 2016 
-## Dropped featureData's levels Wed Oct  5 09:13:01 2016 
+## Subset [6725,20][5032,20] Fri Oct  7 10:16:26 2016 
+## Removed features with more than 0 NAs: Fri Oct  7 10:16:26 2016 
+## Dropped featureData's levels Fri Oct  7 10:16:26 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Oct  5 09:13:02 2016 
-## Added markers from  'pdres' marker vector. Wed Oct  5 09:13:03 2016 
+## Added markers from  'mrk' marker vector. Fri Oct  7 10:16:28 2016 
+## Added markers from  'pdres' marker vector. Fri Oct  7 10:16:28 2016 
 ##  MSnbase version: 1.99.2
@@
 ##  [4] cluster_2.0.4        annotate_1.51.1      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.35.4 IRanges_2.7.15       S4Vectors_0.11.17   
+##  [7] AnnotationDbi_1.35.4 IRanges_2.7.17       S4Vectors_0.11.18   
 ## [10] MSnbase_1.99.2       ProtGenerics_1.5.1   BiocParallel_1.7.8  
-## [13] mzR_2.7.6            Rcpp_0.12.7          Biobase_2.33.3      
-## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.32    
+## [13] mzR_2.7.9            Rcpp_0.12.7          Biobase_2.33.4      
+## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.33    
 ## [19] knitr_1.14          
@@
 ## [10] MatrixModels_0.4-1    affyio_1.43.0         flexmix_2.3-13       
-## [13] mvtnorm_1.0-5         codetools_0.2-14      splines_3.3.1        
+## [13] mvtnorm_1.0-5         codetools_0.2-15      splines_3.3.1        
 ## [16] doParallel_1.0.10     impute_1.47.0         robustbase_0.92-6    
 ## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-71         
-## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-24       
-## [25] vsn_3.41.0            sfsmisc_1.1-0         shiny_0.14           
+## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-25       
+## [25] vsn_3.41.0            sfsmisc_1.1-0         shiny_0.14.1         
 ## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
@@
 ## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.29.2       
-## [64] rpart_4.1-10          stringi_1.1.1         RSQLite_1.0.0        
+## [64] rpart_4.1-10          stringi_1.1.2         RSQLite_1.0.0        
 ## [67] highr_0.6             genefilter_1.55.2     randomForest_4.6-12  
@@
 ## [76] lattice_0.20-34       htmlwidgets_0.7       gbm_2.1.1            
-## [79] plyr_1.8.4            magrittr_1.5          R6_2.1.3             
+## [79] plyr_1.8.4            magrittr_1.5          R6_2.2.0             
 ## [82] DBI_0.5-1             whisker_0.3-2         mgcv_1.8-15          
 ## [85] survival_2.39-5       RCurl_1.95-4.8        nnet_7.3-12          
-## [88] tibble_1.2            msdata_0.12.2         car_2.1-3            
+## [88] tibble_1.2            msdata_0.12.3         car_2.1-3            
 ## [91] mlbench_2.1-1         grid_3.3.1            FNN_1.1              

### 2016-11-01 bioc-workflow.Rnw
updating text, fixing last errors

@@
 proteomics, the readers are invited to read the
-\Biocexptpkg{"RforProteomics"} package vignettes and associated
+\Biocexptpkg{RforProteomics} package vignettes and associated
 papers \cite{Gatto:2014,Gatto:2015}.
@@
 
-To make use of the full functionality of the \Biocpkg{pRoloc}
-software one needs to import their data into R and prepare them as an
-\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for the efficient
-manipulation and processing of mass spectrometry and proteomics data
-in R. Figure 1 illustrates a simplified view of the \texttt{MSnSet}
-structure; there exists 3 key sub-parts (termed slots) to such a data
-object: (1) the \texttt{exprs} slot for storing the quantitation data, (2)
-the \texttt{fData} slot for storing the feature meta-data, and finally (3)
-the \texttt{pData} slot for storing the sample meta-data.
+To make use of the full functionality of the \Biocpkg{pRoloc} software
+one needs to import their data into R and prepare them as an
+\texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for
+the efficient manipulation and processing of mass spectrometry and
+proteomics data in R. Figure 1 illustrates a simplified view of the
+\texttt{MSnSet} structure; there exists 3 key sub-parts (termed slots)
+to such a data object: (1) the \texttt{exprs} (short for
+\textit{expression} data) slot for storing the quantitation data, (2)
+the \texttt{fData} slot (short for \textit{feature}-metadata) for
+storing the feature meta-data, and finally (3) the \texttt{pData} slot
+(short for \textit{pheno}-metadata, i.e. sample phenotypic data) for
+storing the sample meta-data.
 
@@
   capabilities vignette}. One suggested simple method is to use the
-function \texttt{readMSnSet2} in \Biocpkg{MSnbase}. The function
-takes a single spreadsheet as input and extracts the columns
-containing the quantitation data, as identified by the argument
-\texttt{ecol}, to create the expression data, while the other columns
-in the spreadsheet are appended to the feature meta-data slot.  By
-example, in the code chunk below we read in the \texttt{csv}
-spreadsheet containing the quantitation data from the intersect of
-replicates 1 and 2 of the mouse map \cite{hyper}, using the
-\texttt{readMSnSet2} function. The data is as available online with
-the manuscript (see tab 2 of the \texttt{xlsx} supplementary data set
-1 in \cite{hyper}, which should be exported as a text-based
-spreadsheet). It is also available as a \texttt{csv} in the
-Bioconductor \Biocexptpkg{pRolocdata} data package.
+function \texttt{readMSnSet2}. The function takes a single spreadsheet
+filename as input and extracts the columns containing the quantitation
+data, as identified by the argument \texttt{ecol}, to create the
+expression data, while the other columns in the spreadsheet are
+appended to the feature meta-data slot.  By example, in the code chunk
+below we read in the \texttt{csv} spreadsheet containing the
+quantitation data from the intersect of replicates 1 and 2 of the
+mouse map \cite{hyper}, using the \texttt{readMSnSet2} function. The
+data is as available online with the manuscript (see tab 2 of the
+\texttt{xlsx} supplementary data set 1 in \cite{hyper}, which should
+be exported as a text-based spreadsheet). It is also available as a
+\texttt{csv} in the Bioconductor \Biocexptpkg{pRolocdata} data
+package, which we make use of below.
 
@@
 Note that the file is compressed (as indicated by the \texttt{gz}, for
-\texttt{gzip}, extension), and will be decompressed on-the-fly when read into
-R later on.
+\texttt{gzip}, extension), and will be decompressed on-the-fly when
+read into R.
 
@@
 <<normcheck>>=
-summary(rowSums(exprs(hl[, hl$replicate == 1])))
-summary(rowSums(exprs(hl[, hl$replicate == 2])))
+summary(rowSums(exprs(hl[, hl$Replicate == 1])))
+summary(rowSums(exprs(hl[, hl$Replicate == 2])))
 @

### 2016-11-01 bioc-workflow.tex
updating text, fixing last errors

@@
 ## - - - Processing information - - -
-##  MSnbase version: 1.99.2
+##  MSnbase version: 1.99.7
 \end{verbatim}
@@
 \begin{alltt}
-\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{replicate} \hlopt{==} \hlnum{1}\hlstd{])))}
+\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{1}\hlstd{])))}
 \end{alltt}
@@
 ##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-##       0       0       0       0       0       0
-\end{verbatim}
-\begin{alltt}
-\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{replicate} \hlopt{==} \hlnum{2}\hlstd{])))}
+##   0.997   0.999   1.000   1.000   1.001   1.003
+\end{verbatim}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{2}\hlstd{])))}
 \end{alltt}
@@
 ##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
-##       0       0       0       0       0       0
+##   0.997   0.999   1.000   1.000   1.001   1.003
 \end{verbatim}
@@
 ## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
-## Subset [6725,20][5032,20] Fri Oct  7 10:16:26 2016 
-## Removed features with more than 0 NAs: Fri Oct  7 10:16:26 2016 
-## Dropped featureData's levels Fri Oct  7 10:16:26 2016 
+## Subset [6725,20][5032,20] Tue Nov  1 11:26:17 2016 
+## Removed features with more than 0 NAs: Tue Nov  1 11:26:17 2016 
+## Dropped featureData's levels Tue Nov  1 11:26:17 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Oct  7 10:16:28 2016 
-## Added markers from  'pdres' marker vector. Fri Oct  7 10:16:28 2016 
-##  MSnbase version: 1.99.2
+## Added markers from  'mrk' marker vector. Tue Nov  1 11:26:18 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  1 11:26:19 2016 
+##  MSnbase version: 1.99.7
 \end{verbatim}
@@
 ## other attached packages:
-##  [1] pRolocdata_1.11.9    pRoloc_1.13.15       MLInterfaces_1.53.1 
-##  [4] cluster_2.0.4        annotate_1.51.1      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.35.4 IRanges_2.7.17       S4Vectors_0.11.18   
-## [10] MSnbase_1.99.2       ProtGenerics_1.5.1   BiocParallel_1.7.8  
-## [13] mzR_2.7.9            Rcpp_0.12.7          Biobase_2.33.4      
+##  [1] pRolocdata_1.11.9    pRoloc_1.13.17       MLInterfaces_1.53.1 
+##  [4] cluster_2.0.5        annotate_1.51.1      XML_3.98-1.4        
+##  [7] AnnotationDbi_1.35.4 IRanges_2.7.17       S4Vectors_0.11.19   
+## [10] MSnbase_1.99.7       ProtGenerics_1.5.1   BiocParallel_1.7.9  
+## [13] mzR_2.7.12           Rcpp_0.12.7          Biobase_2.33.4      
 ## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.33    
@@
 ## loaded via a namespace (and not attached):
-##  [1] minqa_1.2.4           colorspace_1.2-6      hwriter_1.3.2        
+##  [1] minqa_1.2.4           colorspace_1.2-7      hwriter_1.3.2        
 ##  [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
@@
 ## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-25       
-## [25] vsn_3.41.0            sfsmisc_1.1-0         shiny_0.14.1         
+## [25] vsn_3.41.5            sfsmisc_1.1-0         shiny_0.14.1         
 ## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
@@
 ## [55] zlibbioc_1.19.0       MASS_7.3-45           scales_0.4.0         
-## [58] BiocInstaller_1.23.9  pcaMethods_1.65.0     SparseM_1.72         
+## [58] BiocInstaller_1.24.0  pcaMethods_1.65.0     SparseM_1.72         
 ## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.29.2       
@@
 ## [70] foreach_1.4.3         e1071_1.6-7           prabclus_2.2-6       
-## [73] bitops_1.0-6          mzID_1.11.2           evaluate_0.9         
+## [73] bitops_1.0-6          mzID_1.11.2           evaluate_0.10        
 ## [76] lattice_0.20-34       htmlwidgets_0.7       gbm_2.1.1            
@@
 ## [85] survival_2.39-5       RCurl_1.95-4.8        nnet_7.3-12          
-## [88] tibble_1.2            msdata_0.12.3         car_2.1-3            
+## [88] tibble_1.2            msdata_0.14.0         car_2.1-3            
 ## [91] mlbench_2.1-1         grid_3.3.1            FNN_1.1              

### 2016-11-01 bioc-workflow.Rnw
small changes in text

@@
 More details above combining data are given in the dedicated
-*Combining MSnSet instances* section of the \Biocpkg{MSnbase}
+\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
 \href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
@@
 sub-cellular resolution. Missing data can be imputated using
-\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying the
-imputation method is then determined by a \texttt{methods} parameter. In our
-particular case, missing values are indicative of protein groups that
-were not profiles in both replicates.
+\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
+the imputation method is then determined by a \texttt{methods}
+parameter. In our particular case, missing values are indicative of
+protein groups that were not acquired in both replicates (Figures
+\ref{fig:namap}).
 
@@
 
-We prefer to remove proteins that were not assayed in both of our two
-replicated experiments. This is done with the \texttt{filterNA} function that
+We prefer to remove proteins that were not assayed in both replicated
+experiments. This is done with the \texttt{filterNA} function that
 removes features that contain more than a certain proportion (default
@@
 
-When more than 2 data are to be combined and/or too many proteins were
+When more than 2 data are to be combined and too many proteins were
 not consistently assayed, leading to too many proteins being filtered
@@
 protein-sub-cellular niche membership over the output of several
-experiments (see section *Supervised machine learning* for the
+experiments (see section \textit{Supervised machine learning} for the
 description of sub-cellular assignments).

### 2016-11-01 bioc-workflow.tex
small changes in text

@@
 ## - - - Processing information - - -
-##  MSnbase version: 1.99.7
+##  MSnbase version: 1.21.8
 \end{verbatim}
@@
 
-will list the 54 datasets that are
+will list the 41 datasets that are
 available in \Biocexptpkg{pRolocdata}.
@@
 ## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
-## Subset [6725,20][5032,20] Tue Nov  1 11:26:17 2016 
-## Removed features with more than 0 NAs: Tue Nov  1 11:26:17 2016 
-## Dropped featureData's levels Tue Nov  1 11:26:17 2016 
+## Subset [6725,20][5032,20] Tue Nov  1 14:26:12 2016 
+## Removed features with more than 0 NAs: Tue Nov  1 14:26:12 2016 
+## Dropped featureData's levels Tue Nov  1 14:26:12 2016 
 ##  MSnbase version: 1.21.7
@@
 
-When more than 2 data are to be combined and/or too many proteins were
-not consistently assayed, leading to too many proteins being filtered
-out, we suggest to implement an ensemble of classifiers voting on
-protein-sub-cellular niche membership over the output of several
-experiments (see section *Supervised machine learning* for the
-description of sub-cellular assignments).
+When more than 2 data are to be combined and, as a result, too many
+proteins were not consistently assayed, leading to too many proteins
+being filtered out, we suggest to implement an ensemble of classifiers
+voting on protein-sub-cellular niche membership over the output of
+several experiments (see section \textit{Supervised machine learning}
+for the description of sub-cellular assignments).
 
@@
 \end{kframe}
-\includegraphics[width=\textwidth]{figure/plotmarkers-1}
-
 \end{knitrout}
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  1 11:26:18 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  1 11:26:19 2016 
-##  MSnbase version: 1.99.7
+## Added markers from  'mrk' marker vector. Tue Nov  1 14:26:14 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  1 14:26:15 2016 
+##  MSnbase version: 1.21.8
 \end{verbatim}
@@
 \hlstd{cl} \hlkwb{<-} \hlkwd{getMarkerClasses}\hlstd{(hl,} \hlstr{"pd"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+##  [1] "40S Ribosome"                         
+##  [2] "60S Ribosome"                         
+##  [3] "Endoplasmic reticulum/Golgi apparatus"
+##  [4] "Mitochondrion"                        
+##  [5] "Phenotype 1"                          
+##  [6] "Phenotype 2"                          
+##  [7] "Phenotype 3"                          
+##  [8] "Phenotype 4"                          
+##  [9] "Phenotype 5"                          
+## [10] "Plasma membrane"                      
+## [11] "Proteasome"
+\end{verbatim}
+\begin{alltt}
 \hlstd{cols} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)]}
@@
 \end{kframe}
-\includegraphics[width=\textwidth]{figure/plotPDres-1}
-
 \end{knitrout}
@@
 \begin{verbatim}
-## null device 
-##           1
+## pdf 
+##   3
 \end{verbatim}
@@
 \end{alltt}
-\end{kframe}
-\includegraphics[width=.6\textwidth]{figure/plottl-1}
-
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in .local(x, y, ...): could not find function "{}aes"{}}}\end{kframe}
 \end{knitrout}
@@
 \begin{verbatim}
-## R version 3.3.1 Patched (2016-08-02 r71022)
+## R Under development (unstable) (2016-05-11 r70599)
 ## Platform: x86_64-pc-linux-gnu (64-bit)
@@
 ## other attached packages:
-##  [1] pRolocdata_1.11.9    pRoloc_1.13.17       MLInterfaces_1.53.1 
-##  [4] cluster_2.0.5        annotate_1.51.1      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.35.4 IRanges_2.7.17       S4Vectors_0.11.19   
-## [10] MSnbase_1.99.7       ProtGenerics_1.5.1   BiocParallel_1.7.9  
-## [13] mzR_2.7.12           Rcpp_0.12.7          Biobase_2.33.4      
-## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.33    
-## [19] knitr_1.14          
+##  [1] pRolocdata_1.11.0    pRoloc_1.13.5        MLInterfaces_1.53.0 
+##  [4] cluster_2.0.4        annotate_1.51.0      XML_3.98-1.4        
+##  [7] AnnotationDbi_1.35.3 IRanges_2.7.11       S4Vectors_0.11.7    
+## [10] MSnbase_1.21.8       ProtGenerics_1.5.1   BiocParallel_1.7.4  
+## [13] mzR_2.7.3            Rcpp_0.12.5          Biobase_2.33.0      
+## [16] BiocGenerics_0.19.1  gridExtra_2.2.1      BiocStyle_2.1.7     
+## [19] knitr_1.13          
 ## 
 ## loaded via a namespace (and not attached):
-##  [1] minqa_1.2.4           colorspace_1.2-7      hwriter_1.3.2        
+##  [1] minqa_1.2.4           colorspace_1.2-6      hwriter_1.3.2        
 ##  [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
-##  [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-16         
+##  [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-15         
 ## [10] MatrixModels_0.4-1    affyio_1.43.0         flexmix_2.3-13       
-## [13] mvtnorm_1.0-5         codetools_0.2-15      splines_3.3.1        
+## [13] mvtnorm_1.0-5         codetools_0.2-14      splines_3.4.0        
 ## [16] doParallel_1.0.10     impute_1.47.0         robustbase_0.92-6    
-## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-71         
-## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-25       
-## [25] vsn_3.41.5            sfsmisc_1.1-0         shiny_0.14.1         
-## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
-## [31] limma_3.29.21         formatR_1.4           htmltools_0.3.5      
-## [34] quantreg_5.29         tools_3.3.1           ggvis_0.4.3          
-## [37] gtable_0.2.0          affy_1.51.1           reshape2_1.4.1       
-## [40] dplyr_0.5.0           MALDIquant_1.15       trimcluster_0.1-2    
+## [19] jsonlite_0.9.22       nloptr_1.0.4          caret_6.0-70         
+## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-24       
+## [25] vsn_3.41.0            sfsmisc_1.1-0         shiny_0.13.2         
+## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-6         
+## [31] limma_3.29.12         formatR_1.4           htmltools_0.3.5      
+## [34] quantreg_5.26         tools_3.4.0           ggvis_0.4.2          
+## [37] gtable_0.2.0          affy_1.51.0           reshape2_1.4.1       
+## [40] dplyr_0.4.3           MALDIquant_1.15       trimcluster_0.1-2    
 ## [43] gdata_2.17.0          preprocessCore_1.35.0 nlme_3.1-128         
-## [46] iterators_1.0.8       fpc_2.1-10            stringr_1.1.0        
-## [49] lme4_1.1-12           lpSolve_5.6.13        mime_0.5             
-## [52] gtools_3.5.0          dendextend_1.3.0      DEoptimR_1.0-6       
-## [55] zlibbioc_1.19.0       MASS_7.3-45           scales_0.4.0         
-## [58] BiocInstaller_1.24.0  pcaMethods_1.65.0     SparseM_1.72         
-## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.29.2       
-## [64] rpart_4.1-10          stringi_1.1.2         RSQLite_1.0.0        
-## [67] highr_0.6             genefilter_1.55.2     randomForest_4.6-12  
-## [70] foreach_1.4.3         e1071_1.6-7           prabclus_2.2-6       
-## [73] bitops_1.0-6          mzID_1.11.2           evaluate_0.10        
-## [76] lattice_0.20-34       htmlwidgets_0.7       gbm_2.1.1            
-## [79] plyr_1.8.4            magrittr_1.5          R6_2.2.0             
-## [82] DBI_0.5-1             whisker_0.3-2         mgcv_1.8-15          
-## [85] survival_2.39-5       RCurl_1.95-4.8        nnet_7.3-12          
-## [88] tibble_1.2            msdata_0.14.0         car_2.1-3            
-## [91] mlbench_2.1-1         grid_3.3.1            FNN_1.1              
-## [94] threejs_0.2.2         digest_0.6.10         diptest_0.75-7       
-## [97] xtable_1.8-2          httpuv_1.3.3          munsell_0.4.3
+## [46] iterators_1.0.8       fpc_2.1-10            stringr_1.0.0        
+## [49] lme4_1.1-12           lpSolve_5.6.13        mime_0.4             
+## [52] gtools_3.5.0          DEoptimR_1.0-4        zlibbioc_1.19.0      
+## [55] MASS_7.3-45           scales_0.4.0          BiocInstaller_1.23.4 
+## [58] pcaMethods_1.65.0     SparseM_1.7           RColorBrewer_1.1-2   
+## [61] ggplot2_2.1.0         biomaRt_2.29.2        rpart_4.1-10         
+## [64] stringi_1.1.1         RSQLite_1.0.0         highr_0.6            
+## [67] genefilter_1.55.2     randomForest_4.6-12   foreach_1.4.3        
+## [70] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
+## [73] rgl_0.95.1441         mzID_1.11.2           evaluate_0.9         
+## [76] lattice_0.20-33       htmlwidgets_0.6       gbm_2.1.1            
+## [79] plyr_1.8.4            magrittr_1.5          R6_2.1.2             
+## [82] DBI_0.4-1             mgcv_1.8-12           survival_2.39-4      
+## [85] RCurl_1.95-4.8        nnet_7.3-12           msdata_0.12.0        
+## [88] car_2.1-2             mlbench_2.1-1         grid_3.4.0           
+## [91] FNN_1.1               threejs_0.2.2         digest_0.6.9         
+## [94] diptest_0.75-7        xtable_1.8-2          httpuv_1.3.3         
+## [97] munsell_0.4.3
 \end{verbatim}

### 2016-11-01 bioc-workflow.Rnw
more updates, up to replication section

@@
 suppressPackageStartupMessages(library("pRolocdata"))
-cache <- TRUE
+cache <- FALSE
 @
@@
 In the code chunk below we produce a PCA plot of the mouse stem cell
-dataset. One point on the plot represents one protein. We can indeed
-see several distinct protein clusters. We specify \texttt{fcol = NULL}, which
-means not to consider any feature variable to annotate the features
-(proteins) with colours. We will see later how to use this to annotate
-the PCA plot with prior information about sub-cellular localisation.
+dataset (Figure \ref{fig:pcahl}). One point on the plot represents one
+protein. We can indeed see several distinct protein clusters. We
+specify \texttt{fcol = NULL}, which means not to consider any feature
+variable to annotate the features (proteins) with colours. We will see
+later how to use this to annotate the PCA plot with prior information
+about sub-cellular localisation.
 
@@
 gradient to identify channels displaying particularly low yield. This
-can be done using the \texttt{plotDist} and \texttt{boxplot} functions, that plot
-the protein profiles occupancy along the gradient (we also display the
-mean channel intensities) and a \texttt{boxplot} of the column intensities.
-In the two plots displayed below, we re-order the TMT channles to pair
-corresponding channels in the two replicates (rather than ordering
-the channels by replicate).
+can be done using the \texttt{plotDist} and \texttt{boxplot}
+functions, that plot the protein profiles occupancy along the gradient
+(we also display the mean channel intensities) and a \texttt{boxplot}
+of the column intensities.  In the two plots displayed on figure
+\ref{fig:qcbx}, we re-order the TMT channles to pair corresponding
+channels in the two replicates (rather than ordering the channels by
+replicate).
 
@@
 well-known resident of a specific sub-cellular niche in a species
-*and* condition of interest. Applying this to machine learning (ML),
-and specifically supervised learning, for the task of protein
-localisation prediction, these markers constitute the labelled training
-data to use as input to a classification analyses. Defining well-known
-residents, and obtaining labelled training data for ML analyses can be
-time consuming, but it is important to define markers that are
-representative of the multivariate data space and on which a
-classifier will be trained and generated. \Biocpkg{pRoloc}
-provides a convenience function, \texttt{addMarkers}, to directly add markers
-to a \texttt{MSnSet} object, as demonstrated in the code chunk below. These
-marker sets can be accessed using the \texttt{pRolocmarkers()}
-function. Marker sets are stored as a simple named vector in R, and
-originate from in-house user-defined sets of markers or from previous
-published studies \cite{Gatto:2014b}, which are continuosly updated and 
-integrated.
+\textit{and} condition of interest. Applying this to machine learning
+(ML), and specifically supervised learning, for the task of protein
+localisation prediction, these markers constitute the labelled
+training data to use as input to a classification analyses. Defining
+well-known residents, and obtaining labelled training data for ML
+analyses can be time consuming, but it is important to define markers
+that are representative of the multivariate data space and on which a
+classifier will be trained and generated. \Biocpkg{pRoloc} provides a
+convenience function, \texttt{addMarkers}, to directly add markers to
+a \texttt{MSnSet} object, as demonstrated in the code chunk
+below. These marker sets can be accessed using the
+\texttt{pRolocmarkers()} function. Marker sets are stored as a simple
+named vector in R, and originate from in-house user-defined sets of
+markers or from previous published studies \cite{Gatto:2014b}, which
+are continuosly updated and integrated.
 
@@
 
-<<addmrkers>>=
+<<addmarkers>>=
 ## Use mouse markers
@@
 We now visualise these annotations along the PCA plot using the
-\texttt{plot2D} function and then use the \texttt{addLegend} function to map the
-marker classes to the pre-defined colours. We also display the data
-along the first and seventh PCs using the \texttt{dims} argument. Note that
-in this second call to the \texttt{plot2D} function, we have omitted the
-\texttt{fcol} argument, as the \texttt{"markers"} feature variable name is the
-default value. We choose to display PCs 1 and 7 to illustrate that
-while upper principal components explain much less variability in the
-data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
-mitochondrial and peroxisome clusters can be differenciated, despite
-the apparent overlap in the two first PCs.
+\texttt{plot2D} function and then use the \texttt{addLegend} function
+to map the marker classes to the pre-defined colours. We also display
+the data along the first and seventh PCs using the \texttt{dims}
+argument. Note that in these calls to the \texttt{plot2D} function, we
+have omitted the \texttt{fcol} argument and use of the default
+\texttt{"markers"} feature variable to annotated the plot. We choose
+to display PCs 1 and 7 to illustrate that while upper principal
+components explain much less variability in the data (2.23\% for PC7,
+as opposed to 48.41\% for PC1), we see that the mitochondrial (purple)
+and peroxisome (dark blue) clusters can be differenciated, despite the
+apparent overlap in the two first PCs.
 
@@
   \centering
-<<plotmarkers, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>=
+<<plotmarkers, out.width="\\textwidth", fig.width=12, fig.asp=1/2>>=
 par(mfrow = c(1, 2))
 plot2D(hl, main = "pRolocmarkers for mouse")
-addLegend(hl, cex = .7)
+addLegend(hl, cex = .6)
 plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
 @  
-  \caption{Annotated PCA plots of the \texttt{hl} data.}
+  \caption{Annotated PCA plots of the \texttt{hl} dataset.}
   \label{fig:plotmarkers}
@@
 As demonstrated in \cite{hyper} and illustrated in the PCA plot above,
-the Golgi apparatus proteins display a dynamic pattern, noting sets of
-Golgi marker proteins that are distributed amongst other subcellular
-structures, an observation supported by microscopy. As such, we are
-going to reset the annotation of Golgi markers to unknown using the
-\texttt{fDataTounknown} function. It is often used to replace empty
-strings ("") or missing values in the markers definition to a common
-definition of *unknown localisation*.
+the Golgi apparatus proteins (brown) display a dynamic pattern, noting
+sets of Golgi marker proteins that are distributed amongst other
+subcellular structures, an observation supported by microscopy. As
+such, we are going to reset the annotation of Golgi markers to unknown
+using the \texttt{fDataTounknown} function. It is often used to
+replace empty strings ("") or missing values in the markers definition
+to a common definition of \textit{unknown} localisation.
 
@@
 curation is vital to check that any annotation added is in the correct
-context for the the biological question under investigation.
+context for the biological question under investigation.
 
@@
 few marker clusters is useful to assess how exactly they differ (in
-terms of peak channels, for example). Below, we plot the profile of
-the mitochondrial and peroxisome markers to highlight the differences
-in profiles between these two sets of markers along the 6th and 7th
-channels, as represented above along the 7th PC on the PCA plot.
+terms of peak channels, for example). On figure \ref{fig:plotDist2},
+we plot the profile of the mitochondrial and peroxisome markers to
+highlight the differences in profiles between these two sets of
+markers along the 6th and 7th channels, as represented above along the
+7th PC on the PCA plot on figure \ref{plotmarkers}.
 

### 2016-11-01 bioc-workflow.tex
more updates, up to replication section

--- 2016-11-01 bioc-workflow.tex
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  1 15:41:54 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Sep 29 20:32:05 2016 
-## Subset [6725,20][5032,20] Tue Nov  1 11:26:17 2016 
-## Removed features with more than 0 NAs: Tue Nov  1 11:26:17 2016 
-## Dropped featureData's levels Tue Nov  1 11:26:17 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  1 15:41:54 2016 
+## Subset [6725,20][5032,20] Tue Nov  1 15:41:54 2016 
+## Removed features with more than 0 NAs: Tue Nov  1 15:41:54 2016 
+## Dropped featureData's levels Tue Nov  1 15:41:54 2016 
 ##  MSnbase version: 1.21.7
@@
 \hlkwd{plot2D}\hlstd{(hl,} \hlkwc{main} \hlstd{=} \hlstr{"pRolocmarkers for mouse"}\hlstd{)}
-\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.7}\hlstd{)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
 \hlkwd{plot2D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{7}\hlstd{),} \hlkwc{main} \hlstd{=} \hlstr{"Marker resolution along PC 1 and 7"}\hlstd{)}
@@
 \end{knitrout}
-  \caption{Annotated PCA plots of the \texttt{hl} data.}
+  \caption{Annotated PCA plots of the \texttt{hl} dataset.}
   \label{fig:plotmarkers}
@@
 few marker clusters is useful to assess how exactly they differ (in
-terms of peak channels, for example). Below, we plot the profile of
-the mitochondrial and peroxisome markers to highlight the differences
-in profiles between these two sets of markers along the 6th and 7th
-channels, as represented above along the 7th PC on the PCA plot.
+terms of peak channels, for example). On figure \ref{fig:plotDist2},
+we plot the profile of the mitochondrial and peroxisome markers to
+highlight the differences in profiles between these two sets of
+markers along the 6th and 7th channels, as represented above along the
+7th PC on the PCA plot.
 
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  1 11:26:18 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  1 11:26:19 2016 
+## Added markers from  'mrk' marker vector. Tue Nov  1 15:41:56 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  1 15:41:56 2016 
 ##  MSnbase version: 1.99.7
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4383808               0.3028615               0.3856431 
+##               0.4346677               0.3059684               0.3941820 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6859801               0.5995991               0.4350839 
+##               0.6893290               0.5923417               0.4231235 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4307964               0.5935764               0.9500433 
+##               0.4140720               0.5899672               0.9502352 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7953738               0.7087561               0.3155507 
+##               0.7958140               0.7091452               0.3132634 
 ##         Plasma membrane              Proteasome 
-##               0.7171990               0.4185577
+##               0.7135761               0.4179717
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      87                     169                      88 
+##                      85                     170                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     298                     475                      98 
+##                     295                     481                      96 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      31                     123                     521 
+##                      26                     123                     521 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     344                      38 
+##                     229                     345                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     319                     157                    2055
+##                     319                     157                    2058
 \end{verbatim}

### 2016-11-02 bioc-workflow.Rnw
replication section

@@
 suppressPackageStartupMessages(library("knitr"))
+suppressPackageStartupMessages(library("xtable"))
 suppressPackageStartupMessages(library("gridExtra"))
@@
 markers along the 6th and 7th channels, as represented above along the
-7th PC on the PCA plot on figure \ref{plotmarkers}.
+7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
 
@@
 does not provide an adequate, goal-driven assessment of different
-experiments. Indeed, due to the nature of the experiment and gradient fraction
-collection, the quantitative channels 
-do not correspond to identical selected fractions along the
-gradient. As can be seen in the table below (taken from the
-\texttt{hyperLOPIT}'s \texttt{pData}, focusing on channels 7 to 10), different sets
-of gradient fractions are pooled to obtain enough material and optimise
-acurate quantitation.
-
-<<pdtab, echo=FALSE, eval = FALSE>>=
+experiments. Indeed, due to the nature of the experiment and gradient
+fraction collection, the quantitative channels do not correspond to
+identical selected fractions along the gradient. As can be seen in the
+table below (taken from the \texttt{hl}'s \texttt{pData}, focusing on
+channels 7 to 10), different sets of gradient fractions are pooled to
+obtain enough material and optimise acurate quantitation.
+
+<<pdtab, echo=FALSE, eval = TRUE, results = 'asis'>>=
 i <- c(2, 3, 4, 13, 14)
-pData(hl)[i, ]
-@
-
-% On the figure below, we compare the relative intensities of channels 7
-% to 10 on scatter plots, highlighting different marker sets. These
-% differences result from different peak fractions for these niches on
-% the mitochondrial and chromatin occupancy profiles for the two
-% replicates.
-
-% <<repl1, eval = TRUE, echo=FALSE, fig.width = 12>>=
-% sel0 <- fData(hyperLOPIT)$markers != "unknown"
-% cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
-% cls[!sel0] <- getUnknowncol()
-
-% par(mfrow = c(2, 4))
-% for (i in c(7, 8, 10, 9)) {
-%     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
-%          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
-%     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
-%     grid()
-%     abline(0, 1, col = "red")
-% }
-% plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
-% addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
-
-% i <- fData(hyperLOPIT)$markers == "Mitochondrion"
-% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
-%          featureNames(hyperLOPIT)[i], type = "l",
-%          mcol = "orange")
-% title(main = "Mitochondrion")
-% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
-% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
-% legend("topleft", paste("Replicate", 1:2),
-%        col = c("#F781BF", "orange"),
-%        lty = 1, bty = "n")
-
-% i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
-% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
-%          featureNames(hyperLOPIT)[i], type = "l")
-% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
-% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
-% title(main = "Chromatin")
-% legend("topleft", paste("Replicate", 1:2),
-%        col = c("#984EA3", "steelblue"),
-%        lty = 1, bty = "n")
-% @
+xtable(pData(hl)[i, ], 
+       caption = "Differences in gradient fraction pooling.", 
+       label = "tab:pdtab")
+@
 

### 2016-11-02 bioc-workflow.tex
replication section

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  1 15:41:54 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:41:44 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  1 15:41:54 2016 
-## Subset [6725,20][5032,20] Tue Nov  1 15:41:54 2016 
-## Removed features with more than 0 NAs: Tue Nov  1 15:41:54 2016 
-## Dropped featureData's levels Tue Nov  1 15:41:54 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:41:44 2016 
+## Subset [6725,20][5032,20] Wed Nov  2 14:41:44 2016 
+## Removed features with more than 0 NAs: Wed Nov  2 14:41:44 2016 
+## Dropped featureData's levels Wed Nov  2 14:41:44 2016 
 ##  MSnbase version: 1.21.7
@@
 markers along the 6th and 7th channels, as represented above along the
-7th PC on the PCA plot.
+7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
 
@@
 does not provide an adequate, goal-driven assessment of different
-experiments. Indeed, due to the nature of the experiment and gradient fraction
-collection, the quantitative channels 
-do not correspond to identical selected fractions along the
-gradient. As can be seen in the table below (taken from the
-\texttt{hyperLOPIT}'s \texttt{pData}, focusing on channels 7 to 10), different sets
-of gradient fractions are pooled to obtain enough material and optimise
-acurate quantitation.
+experiments. Indeed, due to the nature of the experiment and gradient
+fraction collection, the quantitative channels do not correspond to
+identical selected fractions along the gradient. As can be seen in the
+table below (taken from the \texttt{hl}'s \texttt{pData}, focusing on
+channels 7 to 10), different sets of gradient fractions are pooled to
+obtain enough material and optimise acurate quantitation.
+
+% latex table generated in R 3.3.2 by xtable 1.8-2 package
+% Wed Nov  2 14:41:46 2016
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrllr}
+  \hline
+ & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
+  \hline
+X127N &   1 & 127N & 1 to 6 (pooled) & 6.00 \\ 
+  X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
+  X128N &   1 & 128N & 10 to 11 (pooled) & 13.30 \\ 
+  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
+  X128N.1 &   2 & 128N & 10 to 11 (pooled) & 12.50 \\ 
+   \hline
+\end{tabular}
+\caption{Differences in gradient fraction pooling.} 
+\label{tab:pdtab}
+\end{table}
 
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  1 15:41:56 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  1 15:41:56 2016 
+## Added markers from  'mrk' marker vector. Wed Nov  2 14:41:46 2016 
+## Added markers from  'pdres' marker vector. Wed Nov  2 14:41:46 2016 
 ##  MSnbase version: 1.99.7
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4346677               0.3059684               0.3941820 
+##               0.4348738               0.3025063               0.3885913 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6893290               0.5923417               0.4231235 
+##               0.6911470               0.6332177               0.4388085 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4140720               0.5899672               0.9502352 
+##               0.4206513               0.5959339               0.9495535 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7958140               0.7091452               0.3132634 
+##               0.7946359               0.7066474               0.3160929 
 ##         Plasma membrane              Proteasome 
-##               0.7135761               0.4179717
+##               0.7169909               0.4203854
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      89 
+##                      85                     172                      87 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     481                      96 
+##                     301                     465                      96 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     123                     521 
+##                      26                     123                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     345                      38 
+##                     230                     343                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     319                     157                    2058
+##                     330                     156                    2056
 \end{verbatim}
@@
 \begin{verbatim}
-## R version 3.3.1 Patched (2016-08-02 r71022)
+## R version 3.3.2 Patched (2016-11-01 r71616)
 ## Platform: x86_64-pc-linux-gnu (64-bit)
@@
 ## other attached packages:
-##  [1] pRolocdata_1.11.9    pRoloc_1.13.17       MLInterfaces_1.53.1 
+##  [1] pRolocdata_1.11.9    pRoloc_1.15.1        MLInterfaces_1.53.1 
 ##  [4] cluster_2.0.5        annotate_1.51.1      XML_3.98-1.4        
@@
 ## [13] mzR_2.7.12           Rcpp_0.12.7          Biobase_2.33.4      
-## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      BiocStyle_2.1.33    
-## [19] knitr_1.14          
+## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      xtable_1.8-2        
+## [19] BiocStyle_2.1.33     knitr_1.14          
 ## 
@@
 ## [10] MatrixModels_0.4-1    affyio_1.43.0         flexmix_2.3-13       
-## [13] mvtnorm_1.0-5         codetools_0.2-15      splines_3.3.1        
+## [13] mvtnorm_1.0-5         codetools_0.2-15      splines_3.3.2        
 ## [16] doParallel_1.0.10     impute_1.47.0         robustbase_0.92-6    
-## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-71         
+## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-72         
 ## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-25       
@@
 ## [31] limma_3.29.21         formatR_1.4           htmltools_0.3.5      
-## [34] quantreg_5.29         tools_3.3.1           ggvis_0.4.3          
+## [34] quantreg_5.29         tools_3.3.2           ggvis_0.4.3          
 ## [37] gtable_0.2.0          affy_1.51.1           reshape2_1.4.1       
@@
 ## [88] tibble_1.2            msdata_0.14.0         car_2.1-3            
-## [91] mlbench_2.1-1         grid_3.3.1            FNN_1.1              
-## [94] threejs_0.2.2         digest_0.6.10         diptest_0.75-7       
-## [97] xtable_1.8-2          httpuv_1.3.3          munsell_0.4.3
+## [91] mlbench_2.1-1         grid_3.3.2            FNN_1.1              
+## [94] ModelMetrics_1.1.0    threejs_0.2.2         digest_0.6.10        
+## [97] diptest_0.75-7        httpuv_1.3.3          munsell_0.4.3
 \end{verbatim}

### 2016-11-02 bioc-workflow.Rnw
interactive vis section

@@
   \centering
-<<plotDist, out.width=".65\\textwidth">>=
+<<plotDist, out.width=".6\\textwidth", fig.asp=1/3>>=
 hlo <- hl[, order(hl$Iodixonal.Density)]
@@
 
+\clearpage
+
 \section*{Interactive visualisation}
@@
 analyses allowing one to shed light on data structure and patterns of
-interest. Using the \Biocpkg{pRolocGUI} package we can
-interactively visualise, explore and interrogate quantitative spatial
-proteomics data. The \Biocpkg{pRolocGUI} package is currently
-under active development and it relies on the \texttt{shiny} framework for
+interest. Using the \Biocpkg{pRolocGUI} package we can interactively
+visualise, explore and interrogate quantitative spatial proteomics
+data. The \Biocpkg{pRolocGUI} package is currently under active
+development and it relies on the \texttt{shiny} framework for
 reactivity and interactivity. The package currently distributes 3
-different GUI's ("main" (default), "compare" or "compare") which are
-wrapped and launched by the \texttt{pRolocVis} function. In the below code
-chunk we lauch the main app (note, we do not need to specify the
-argument, \texttt{app = "main"} as it is the default).
+different GUI's (\textit{main} (default), \textit{compare} or
+\textit{compare}) which are wrapped and launched by the
+\texttt{pRolocVis} function. In the below code chunk we lauch the main
+app (note, we do not need to specify the argument, \texttt{app =
+  "main"} as it is the default).
 
@@
 
-As diplayed in the screenshot above, the "main" application is
+As diplayed in the screenshot above, the \textit{main} application is
 designed for exploratory data analysis and is divied into 3 tabs: (1)
@@
 data(hyperLOPIT2015ms3r2)
-mydata <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
-pRolocVis(mydata, app = "compare")
+hllst <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
+pRolocVis(hllst, app = "compare")
 @
@@
 proteins between the two data sets are displayed. As per the main
-application proteins can be searched, identified and highlighted on
+application, proteins can be searched, identified and highlighted on
 both PCA plots and in the dedicated profiles tab. One key feature of
@@
 
-The final application "classify", has been designed to view machine
-learning classification results according to user-specified thresholds
-for the assignment of proteins to its sub-cellular location, this is
-discussed later in the subsection \textit{thresholding} in the
-\textit{supervised machine learning section}.
+The final application \textit{classify}, has been designed to view
+machine learning classification results according to user-specified
+thresholds for the assignment of proteins to its sub-cellular
+location, as discussed later in the subsection \textit{thresholding}
+in the \textit{supervised machine learning section}.
 

### 2016-11-02 bioc-workflow.tex
interactive vis section

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:41:44 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:56:28 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:41:44 2016 
-## Subset [6725,20][5032,20] Wed Nov  2 14:41:44 2016 
-## Removed features with more than 0 NAs: Wed Nov  2 14:41:44 2016 
-## Dropped featureData's levels Wed Nov  2 14:41:44 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:56:28 2016 
+## Subset [6725,20][5032,20] Wed Nov  2 14:56:29 2016 
+## Removed features with more than 0 NAs: Wed Nov  2 14:56:29 2016 
+## Dropped featureData's levels Wed Nov  2 14:56:29 2016 
 ##  MSnbase version: 1.21.7
@@
 \end{kframe}
-\includegraphics[width=.65\textwidth]{figure/plotDist-1}
+\includegraphics[width=.6\textwidth]{figure/plotDist-1}
 
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Wed Nov  2 14:41:46 2016
+% Wed Nov  2 14:56:30 2016
 \begin{table}[ht]
@@
 
-% On the figure below, we compare the relative intensities of channels 7
-% to 10 on scatter plots, highlighting different marker sets. These
-% differences result from different peak fractions for these niches on
-% the mitochondrial and chromatin occupancy profiles for the two
-% replicates.
-
-% <<repl1, eval = TRUE, echo=FALSE, fig.width = 12>>=
-% sel0 <- fData(hyperLOPIT)$markers != "unknown"
-% cls <- paste0(getStockcol(), 80)[factor(fData(hyperLOPIT)$markers)]
-% cls[!sel0] <- getUnknowncol()
-
-% par(mfrow = c(2, 4))
-% for (i in c(7, 8, 10, 9)) {
-%     plot(exprs(hyperLOPIT[, c(i, i+10)]), col = cls,
-%          main = paste("Channel", pData(hyperLOPIT)[i, "Fraction.No"]))
-%     points(exprs(hyperLOPIT[sel0, c(i, i+10)]), col = cls[sel0], pch = 19)
-%     grid()
-%     abline(0, 1, col = "red")
-% }
-% plot(1:100, type = "n", xaxt = "n", yaxt = "n", xlab = "", bty = "n", ylab = "")
-% addLegend(hyperLOPIT, fcol = "markers", where = "center", cex = .7)
-
-% i <- fData(hyperLOPIT)$markers == "Mitochondrion"
-% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
-%          featureNames(hyperLOPIT)[i], type = "l",
-%          mcol = "orange")
-% title(main = "Mitochondrion")
-% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
-% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#F781BF30", lty = 1)
-% legend("topleft", paste("Replicate", 1:2),
-%        col = c("#F781BF", "orange"),
-%        lty = 1, bty = "n")
-
-% i <- fData(hyperLOPIT)$markers == "Nucleus - Chromatin"
-% plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2],
-%          featureNames(hyperLOPIT)[i], type = "l")
-% ## plotDist(hyperLOPIT[i, hyperLOPIT$Replicate == 2])
-% matlines(t(exprs(hyperLOPIT[i, hyperLOPIT$Replicate == 1])), col = "#984EA330", lty = 1)
-% title(main = "Chromatin")
-% legend("topleft", paste("Replicate", 1:2),
-%        col = c("#984EA3", "steelblue"),
-%        lty = 1, bty = "n")
-% @
-
 The more relevant comparison unit is not a single channel, but rather
@@
 \hlkwd{data}\hlstd{(hyperLOPIT2015ms3r2)}
-\hlstd{mydata} \hlkwb{<-} \hlkwd{MSnSetList}\hlstd{(}\hlkwd{list}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
-\hlkwd{pRolocVis}\hlstd{(mydata,} \hlkwc{app} \hlstd{=} \hlstr{"compare"}\hlstd{)}
+\hlstd{hllst} \hlkwb{<-} \hlkwd{MSnSetList}\hlstd{(}\hlkwd{list}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
+\hlkwd{pRolocVis}\hlstd{(hllst,} \hlkwc{app} \hlstd{=} \hlstr{"compare"}\hlstd{)}
 \end{alltt}
@@
 
-The final application "classify", has been designed to view machine
-learning classification results according to user-specified thresholds
-for the assignment of proteins to its sub-cellular location, this is
-discussed later in the subsection \textit{thresholding} in the
-\textit{supervised machine learning section}.
+The final application \textit{classify}, has been designed to view
+machine learning classification results according to user-specified
+thresholds for the assignment of proteins to its sub-cellular
+location, this is discussed later in the subsection
+\textit{thresholding} in the \textit{supervised machine learning
+  section}.
 
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov  2 14:41:46 2016 
-## Added markers from  'pdres' marker vector. Wed Nov  2 14:41:46 2016 
+## Added markers from  'mrk' marker vector. Wed Nov  2 14:56:30 2016 
+## Added markers from  'pdres' marker vector. Wed Nov  2 14:56:31 2016 
 ##  MSnbase version: 1.99.7
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4348738               0.3025063               0.3885913 
+##               0.4332906               0.3032423               0.3840370 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6911470               0.6332177               0.4388085 
+##               0.6878996               0.5997368               0.4245347 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4206513               0.5959339               0.9495535 
+##               0.4291978               0.5892503               0.9494459 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946359               0.7066474               0.3160929 
+##               0.7950790               0.7079212               0.3151288 
 ##         Plasma membrane              Proteasome 
-##               0.7169909               0.4203854
+##               0.7086233               0.4158484
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      87 
+##                      85                     170                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     301                     465                      96 
+##                     297                     476                      93 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     123                     523 
+##                      27                     125                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      39 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     330                     156                    2056
+##                     321                     157                    2059
 \end{verbatim}

### 2016-11-02 bioc-workflow.Rnw
phenodisco section

@@
 unlabelled instances can only be assigned to a class that exists in
-the training data \cite{Breckels:2013}. In such scenarios novelty
+the training data \cite{Breckels:2013}. In such scenarios, novelty
 detection can be useful to identify data-specific sub-cellular
@@
 ## As per Christoforou et al (2016), 
-hl <- phenoDisco(hl, fcol = "phenoDisco.Input", 
-                 times = 200, 
-                 GS = 60)
+hl <- phenoDisco(hl, fcol = "phenoDisco.Input", times = 200, GS = 60)
 @
@@
 
-The argument \texttt{times} indicated the number of times we run unsupervied
-Gaussian Mixture Modelling before defining a new phenotype cluster.
-The recommended minimum and default value is 100. In the above code
-chunk we increase the value to \texttt{times = 200} as we have found for
-larger datasets (e.g. 5000+ proteins) a higher times is requried for
-convergence. \texttt{GS} defines the minimum number of proteins allowed per
-new data cluster and thus heavily influences what type of new clusters
-are extracted. For example, if a user is interesed in the detection of
-small complexes they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
-If they wish to detect larger more abundant sub-cellular niches a much
-higher \texttt{GS} would be preferable. Specifying a small \texttt{GS} can be more
-time consuming than using a larger \texttt{GS}, and there is a trade off
+The argument \texttt{times} indicated the number of times we run
+unsupervied Gaussian Mixture Modelling before defining a new phenotype
+cluster.  The recommended minimum and default value is 100. In the
+above code chunk we increase the value to \texttt{times = 200} as we
+have found for larger datasets (e.g. 5000+ proteins) a higher times is
+requried for convergence. \texttt{GS} defines the minimum number of
+proteins allowed per new data cluster and thus heavily influences what
+type of new clusters are extracted. For example, if a user is
+interesed in the detection of small complexes they may wish to use a
+small \texttt{GS = 10}, or \texttt{= 20} etc.  If they wish to detect
+larger, more abundant sub-cellular niches a much higher \texttt{GS}
+would be preferable. Specifying a small \texttt{GS} can be more time
+consuming than using a larger \texttt{GS}, and there is a trade off
 between finding interesting small complexes and those that may not be
@@
 One may also consider increasing the search space for new data
-clusters by increasing the value of the parameter \texttt{G}. This defines
-the number of GMM components to test and fit; the default is \texttt{G = 1:9}
-(\texttt{mclust}'s default). One should note that the decreasing the \texttt{GS},
-and increasing the values of the arguments \texttt{times}, \texttt{G} (among other
-function arguments, see \texttt{?phenoDisco}) will heavily influence
-(increase) the total time taken to run the algorithm.  \texttt{phenoDisco}
-supports parallelisation and we strongly suggest you make use of a
-parallel processing to run these analyses (different backends can be
-set with argument \texttt{BPPARAM}).
+clusters by increasing the value of the parameter \texttt{G}. This
+defines the number of GMM components to test and fit; the default is
+\texttt{G = 1:9} (the default value in the
+\href{https://cran.r-project.org/web/packages/mclust/index.html}{\textit{mclust}}
+package \cite{mclust}). One should note that the decreasing the
+\texttt{GS}, and increasing the values of the arguments
+\texttt{times}, \texttt{G} (among other function arguments, see
+\texttt{?phenoDisco}) will heavily influence (increase) the total time
+taken to run the algorithm.  \texttt{phenoDisco} supports
+parallelisation and we strongly suggest you make use of a parallel
+processing to run these analyses.
 
@@
 
-The ouput of running the \texttt{phenoDisco} algorithm is a \texttt{MSnSet}
-containing the new data clusters, appended to the \texttt{featureData} under
-the name \texttt{pd}. We can see by typing \texttt{hl@processingData} directly into
-the console the processing information has been updated to the
-\texttt{MSnSet} recording the parameters that were used to run the analyses.
-This is handy for keeping track of data analyses. The results can be
-displayed by using the \texttt{getMarkers} function. We see that 5 new
-phenotype data clusters were found.
+The ouput of running the \texttt{phenoDisco} algorithm is an
+\texttt{MSnSet} containing the new data clusters, appended to the
+\texttt{featureData} under the name \texttt{pd}. We can see by typing
+\texttt{processingData(hl)} directly into the console the processing
+information has been updated to the \texttt{MSnSet} recording the
+parameters that were used to run the analyses.  This is handy for
+keeping track of data analyses. The results can be displayed by using
+the \texttt{getMarkers} function. We see that 5 new phenotype data
+clusters were found.
 
@@
 @
-
-We can plot the results using the \texttt{plot2D} function.
 
@@
 
-The five new phenotype data clusters can be extracted and examined.
-In the code chunk below we write the results to a .csv file. We use
-the argument \texttt{fDataCols} to specify which columns of the
-\texttt{featureData} to write.
+We can plot the results using the \texttt{plot2D} function (Figure
+\ref{fig:plotPDres}). The five new phenotype data clusters can be
+extracted and examined.  In the code chunk below we write the results
+to a .csv file. We use the argument \texttt{fDataCols} to specify
+which columns of the \texttt{featureData} to write.
 
@@
 
-<<vishl, eval=FALSE>>=
-pRolocVis(hl, fcol = "pd")
-@
+% <<vishl, eval=FALSE>>=
+% pRolocVis(hl, fcol = "pd")
+% @
 

### 2016-11-02 bioc-workflow.tex
phenodisco section

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:56:28 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 15:23:51 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 14:56:28 2016 
-## Subset [6725,20][5032,20] Wed Nov  2 14:56:29 2016 
-## Removed features with more than 0 NAs: Wed Nov  2 14:56:29 2016 
-## Dropped featureData's levels Wed Nov  2 14:56:29 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 15:23:51 2016 
+## Subset [6725,20][5032,20] Wed Nov  2 15:23:52 2016 
+## Removed features with more than 0 NAs: Wed Nov  2 15:23:52 2016 
+## Dropped featureData's levels Wed Nov  2 15:23:52 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Wed Nov  2 14:56:30 2016
+% Wed Nov  2 15:23:54 2016
 \begin{table}[ht]
@@
 thresholds for the assignment of proteins to its sub-cellular
-location, this is discussed later in the subsection
-\textit{thresholding} in the \textit{supervised machine learning
-  section}.
+location, as discussed later in the subsection \textit{thresholding}
+in the \textit{supervised machine learning section}.
 
@@
 \hlcom{## As per Christoforou et al (2016), }
-\hlstd{hl} \hlkwb{<-} \hlkwd{phenoDisco}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,}
-                 \hlkwc{times} \hlstd{=} \hlnum{200}\hlstd{,}
-                 \hlkwc{GS} \hlstd{=} \hlnum{60}\hlstd{)}
+\hlstd{hl} \hlkwb{<-} \hlkwd{phenoDisco}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,} \hlkwc{times} \hlstd{=} \hlnum{200}\hlstd{,} \hlkwc{GS} \hlstd{=} \hlnum{60}\hlstd{)}
 \end{alltt}
@@
 One may also consider increasing the search space for new data
-clusters by increasing the value of the parameter \texttt{G}. This defines
-the number of GMM components to test and fit; the default is \texttt{G = 1:9}
-(\texttt{mclust}'s default). One should note that the decreasing the \texttt{GS},
-and increasing the values of the arguments \texttt{times}, \texttt{G} (among other
-function arguments, see \texttt{?phenoDisco}) will heavily influence
-(increase) the total time taken to run the algorithm.  \texttt{phenoDisco}
-supports parallelisation and we strongly suggest you make use of a
-parallel processing to run these analyses (different backends can be
-set with argument \texttt{BPPARAM}).
-
-The ouput of running the \texttt{phenoDisco} algorithm is a \texttt{MSnSet}
-containing the new data clusters, appended to the \texttt{featureData} under
-the name \texttt{pd}. We can see by typing \texttt{hl@processingData} directly into
-the console the processing information has been updated to the
-\texttt{MSnSet} recording the parameters that were used to run the analyses.
-This is handy for keeping track of data analyses. The results can be
-displayed by using the \texttt{getMarkers} function. We see that 5 new
-phenotype data clusters were found.
+clusters by increasing the value of the parameter \texttt{G}. This
+defines the number of GMM components to test and fit; the default is
+\texttt{G = 1:9} (the default value in the
+\href{https://cran.r-project.org/web/packages/mclust/index.html}{\textit{mclust}}
+package \cite{mclust}). One should note that the decreasing the
+\texttt{GS}, and increasing the values of the arguments
+\texttt{times}, \texttt{G} (among other function arguments, see
+\texttt{?phenoDisco}) will heavily influence (increase) the total time
+taken to run the algorithm.  \texttt{phenoDisco} supports
+parallelisation and we strongly suggest you make use of a parallel
+processing to run these analyses.
+
+The ouput of running the \texttt{phenoDisco} algorithm is an
+\texttt{MSnSet} containing the new data clusters, appended to the
+\texttt{featureData} under the name \texttt{pd}. We can see by typing
+\texttt{processingData(hl)} directly into the console the processing
+information has been updated to the \texttt{MSnSet} recording the
+parameters that were used to run the analyses.  This is handy for
+keeping track of data analyses. The results can be displayed by using
+the \texttt{getMarkers} function. We see that 5 new phenotype data
+clusters were found.
 
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov  2 14:56:30 2016 
-## Added markers from  'pdres' marker vector. Wed Nov  2 14:56:31 2016 
+## Added markers from  'mrk' marker vector. Wed Nov  2 15:23:53 2016 
+## Added markers from  'pdres' marker vector. Wed Nov  2 15:23:54 2016 
 ##  MSnbase version: 1.99.7
@@
 
-We can plot the results using the \texttt{plot2D} function.
-
 \begin{figure}[!ht]
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{pRolocVis}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
+% <<vishl, eval=FALSE>>=
+% pRolocVis(hl, fcol = "pd")
+% @
 
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4332906               0.3032423               0.3840370 
+##               0.4360688               0.3016508               0.3807145 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6878996               0.5997368               0.4245347 
+##               0.6916080               0.6060887               0.4329083 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4291978               0.5892503               0.9494459 
+##               0.4140261               0.5960367               0.9497141 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950790               0.7079212               0.3151288 
+##               0.7955514               0.7064597               0.3145894 
 ##         Plasma membrane              Proteasome 
-##               0.7086233               0.4158484
+##               0.7030837               0.4154482
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      88 
+##                      85                     171                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     297                     476                      93 
+##                     293                     475                     102 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     125                     523 
+##                      26                     123                     524 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
@@
 ##         Plasma membrane              Proteasome                 unknown 
-##                     321                     157                    2059
+##                     320                     160                    2058
 \end{verbatim}

### 2016-11-02 bioc-workflow.Rnw
optimisation section

@@
 
-In the code chunk below we employ the use of a Support Vector Machine
-(SVM) to learn a classifier on the labelled training data. As
-previously mentioned one first needs to train the classifiers
-parameters before an algorithm can be used to predict the class labels
-of the proteins with unknown location. One of the most common ways to
-optimise the parameters of a classifier is to partition the labelled
-data in to training and testing subsets. In this framework parameters
-are tested via a grid search using cross-validation on the training
+In the code chunk below we use a Support Vector Machine (SVM) to learn
+a classifier on the labelled training data. As previously mentioned,
+one first needs to train the classifiers parameters before an
+algorithm can be used to predict the class labels of the proteins with
+unknown location. One of the most common ways to optimise the
+parameters of a classifier is to partition the labelled data in to
+training and testing subsets. In this framework parameters are tested
+via a grid search using cross-validation on the training
 partition. The best parameters chosen from the cross-validation stage
@@
 achieve a good generalisation i.e. that is given an unknown example
-predict its class label with high accuracy. In
-\Biocpkg{pRoloc}, algorithmic performance is estimated using
-stratified 80/20 partitioning for the training/testing subsets
-respectively, in conjuction with five-fold cross-validation in order
-to optimise the free parameters via a grid search. This procedure is
-usually repeated 100 times and then the best parameter(s) are selected
-upon investigation of classifier accuracy, here we use the harmonic
-mean of precision and recall; the macro F1 score. In the code chunk
-below we demonstrate how to optimise the free parameters; \texttt{sigma} and
-\texttt{cost}, of a classical SVM classifier with a Gaussian kernel using the
-function \texttt{svmOptimisation}. As the number of labelled instances per
-class varies from organelle to organelle, we can account for class
-imbalance by setting specific class weights when generating the SVM
-model. Below the weights, \texttt{w} are set to be inversely proportional to
-the class frequencies.
+predict its class label with high accuracy. In \Biocpkg{pRoloc},
+algorithmic performance is estimated using stratified 80/20
+partitioning for the training/testing subsets respectively, in
+conjuction with five-fold cross-validation in order to optimise the
+free parameters via a grid search. This procedure is usually repeated
+100 times and then the best parameter(s) are selected upon
+investigation of classifier accuracy. Here, we use the macro F1 score,
+i.e. the harmonic mean of precision and recall. In the code chunk below we
+demonstrate how to optimise the free parameters, \texttt{sigma} and
+\texttt{cost}, of a classical SVM classifier with a Gaussian kernel
+using the function \texttt{svmOptimisation}. As the number of labelled
+instances per class varies from organelle to organelle, we can account
+for class imbalance by setting specific class weights when generating
+the SVM model. Below the weights, \texttt{w} are set to be inversely
+proportional to the class frequencies.
 
@@
 
-As mentioned previously, we reply on the default feature variable
+As mentioned previously, we rely on the default feature variable
 \texttt{"markers"} to define the class labels and hence can ommit
@@
 important to investigate the model parameters and critically assess
-the best choice. The best choice may not be as simple as the parameter
-set that gives rise to the highest macro F1 score and one must be
-careful to avoid overfitting and to choose parameters wisely.
+the best choice. The \texttt{f1Count} function counts the number of
+parameter occurences above a certain F1 value. The best choice may not
+be as simple as the parameter set that gives rise to the highest macro
+F1 score and one must be careful to avoid overfitting and to choose
+parameters wisely.
+
+<<f1count>>=
+f1Count(params, 0.6)
+@
+
+The parameter optimistion results can also be visualised as a boxplot
+or heatmap, as shown in figure \ref{fig:visualisOptHide}.  The
+\texttt{plot} method for \texttt{GenRegRes} object shows the
+respective distributions of the 100 macro F1 scores for the best
+cost/sigma parameter pairs, and \texttt{levelPlot} shows the averaged
+macro F1 scores, for the full range of parameter values.
 
 <<visualiseOpt, eval=FALSE>>=
-(best <- getParams(params))
 plot(params)
@@
 
-By using the function \texttt{getParams} we can extract the best set of
-parameters. Currently, \texttt{getParams} retrieves the first best set is
-automatically but users are encouraged to critically assess whether
-this is the most wise choice by visualising the results using the
-methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method for \texttt{GenRegRes}
-object shows the respective distributions of the 100 macro F1 scores
-for the best cost/sigma parameter pairs, and \texttt{levelPlot} shows the
-averaged macro F1 scores, for the full range of parameter values. Once
-we have selected the best parameters we can then use them to build a
-classifier from the labelled marker proteins.
+By using the function \texttt{getParams} we can extract the best set
+of parameters. Currently, \texttt{getParams} retrieves the first best
+set is automatically but users are encouraged to critically assess
+whether this is the most wise choice.
+
+<<getbestpars>>=
+(best <- getParams(params))
+@
+
+Once we have selected the best parameters we can then use them to
+build a classifier from the labelled marker proteins.
 

### 2016-11-02 bioc-workflow.tex
optimisation section

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 15:23:51 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 16:12:22 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 15:23:51 2016 
-## Subset [6725,20][5032,20] Wed Nov  2 15:23:52 2016 
-## Removed features with more than 0 NAs: Wed Nov  2 15:23:52 2016 
-## Dropped featureData's levels Wed Nov  2 15:23:52 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 16:12:22 2016 
+## Subset [6725,20][5032,20] Wed Nov  2 16:12:23 2016 
+## Removed features with more than 0 NAs: Wed Nov  2 16:12:23 2016 
+## Dropped featureData's levels Wed Nov  2 16:12:23 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Wed Nov  2 15:23:54 2016
+% Wed Nov  2 16:12:25 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov  2 15:23:53 2016 
-## Added markers from  'pdres' marker vector. Wed Nov  2 15:23:54 2016 
+## Added markers from  'mrk' marker vector. Wed Nov  2 16:12:24 2016 
+## Added markers from  'pdres' marker vector. Wed Nov  2 16:12:25 2016 
 ##  MSnbase version: 1.99.7
@@
 important to investigate the model parameters and critically assess
-the best choice. The best choice may not be as simple as the parameter
-set that gives rise to the highest macro F1 score and one must be
-careful to avoid overfitting and to choose parameters wisely.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{(best} \hlkwb{<-} \hlkwd{getParams}\hlstd{(params))}
+the best choice. The \texttt{f1Count} function counts the number of
+parameter occurences above a certain F1 value. The best choice may not
+be as simple as the parameter set that gives rise to the highest macro
+F1 score and one must be careful to avoid overfitting and to choose
+parameters wisely.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{f1Count}\hlstd{(params,} \hlnum{0.6}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+##     4  8 16
+## 0.1 1 10 89
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+The parameter optimistion results can also be visualised as a boxplot
+or heatmap, as shown in figure \ref{fig:visualisOptHide}.  The
+\texttt{plot} method for \texttt{GenRegRes} object shows the
+respective distributions of the 100 macro F1 scores for the best
+cost/sigma parameter pairs, and \texttt{levelPlot} shows the averaged
+macro F1 scores, for the full range of parameter values.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
 \hlkwd{plot}\hlstd{(params)}
@@
 
-By using the function \texttt{getParams} we can extract the best set of
-parameters. Currently, \texttt{getParams} retrieves the first best set is
-automatically but users are encouraged to critically assess whether
-this is the most wise choice by visualising the results using the
-methods \texttt{plot} and \texttt{levelPlot}. The \texttt{plot} method for \texttt{GenRegRes}
-object shows the respective distributions of the 100 macro F1 scores
-for the best cost/sigma parameter pairs, and \texttt{levelPlot} shows the
-averaged macro F1 scores, for the full range of parameter values. Once
-we have selected the best parameters we can then use them to build a
-classifier from the labelled marker proteins.
+By using the function \texttt{getParams} we can extract the best set
+of parameters. Currently, \texttt{getParams} retrieves the first best
+set is automatically but users are encouraged to critically assess
+whether this is the most wise choice.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{(best} \hlkwb{<-} \hlkwd{getParams}\hlstd{(params))}
+\end{alltt}
+\begin{verbatim}
+## sigma  cost 
+##   0.1  16.0
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+Once we have selected the best parameters we can then use them to
+build a classifier from the labelled marker proteins.
 
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4360688               0.3016508               0.3807145 
+##               0.4391696               0.3042111               0.3869018 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6916080               0.6060887               0.4329083 
+##               0.6964096               0.6130827               0.4300618 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4140261               0.5960367               0.9497141 
+##               0.4186188               0.6092759               0.9497908 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7955514               0.7064597               0.3145894 
+##               0.7950972               0.7103401               0.3158138 
 ##         Plasma membrane              Proteasome 
-##               0.7030837               0.4154482
+##               0.7221735               0.4144763
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      84 
+##                      86                     170                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     293                     475                     102 
+##                     296                     477                     105 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     123                     524 
+##                      24                     120                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     230                     342                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     320                     160                    2058
+##                     315                     158                    2059
 \end{verbatim}

### 2016-11-03 bioc-workflow.Rnw
Update discussion on markers
- Describe conservative vs loose definition of markers, their effect,
  and reference figure reflecting classification scores.
- Move algorithmic argument for 13 markers in optimisation section.

@@
 
-These markers can then be mapped to a \texttt{MSnSet}'s \texttt{featureNames}. The
-mouse dataset used here has Uniprot IDs stored as the \texttt{featureNames}
-(see \texttt{head(featureNames(hl))}) and the names of the vector of
-the mouse markers stored in \Biocpkg{pRoloc} (\texttt{mmus} markers) 
-are also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below), so 
-it is straightforward to match names between the markers and 
-the \texttt{MSnSet} instance using the \texttt{addMarkers} function.
+These markers can then be mapped to a \texttt{MSnSet}'s
+\texttt{featureNames}. The mouse dataset used here has Uniprot IDs
+stored as the \texttt{featureNames} (see
+\texttt{head(featureNames(hl))}) and the names of the vector of the
+mouse markers stored in \Biocpkg{pRoloc} (\texttt{mmus} markers) are
+also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below), so
+it is straightforward to match names between the markers and the
+\texttt{MSnSet} instance using the \texttt{addMarkers} function.
 
@@
 @
+
+We recommend at least 13 markers per sub-cellular class (see the
+\textit{Optimisation} section for details about the algorithmic
+motivation of this number). Markers should be chosen to confidently
+represent to distribution of genuine residents of a sub-cellular
+niche. We generally recommend such a conservative approach in defining
+markers to avoid false assignments when assigning sub-cellular
+localisation of proteins of unknown localisation. A more relaxed
+definition of markers, i.e. one that broadly or over-confidently
+defines markers, risks to erroneously assign proteins to a single
+location, when, in reality, they reside in more that a single location
+(including the assumed unique location). One can not expect to
+identify exact boundaries between sub-cellular classes through marker
+annotation alone; the definition of these boundaries is better handled
+algorithmically, i.e. after application of the supervised learning
+algorithm, using the prediction scores (as described in the
+\textit{Classification} section, in particular Figure
+\ref{fig:plotSVM}).
 
@@
   \centering
-<<plotDist, out.width=".6\\textwidth", fig.asp=1/3>>=
+<<plotDist, out.width=".6\\textwidth", fig.asp=3/5>>=
 hlo <- hl[, order(hl$Iodixonal.Density)]
@@
 100 times and then the best parameter(s) are selected upon
-investigation of classifier accuracy. Here, we use the macro F1 score,
-i.e. the harmonic mean of precision and recall. In the code chunk below we
+investigation of classifier accuracy. We recommend a minimum of 13
+markers per sub-cellular class for stratified 80/20 partitioning and
+5-fold cross-validation; this allows a minimum of 10 examples for
+parameter optimisation on the training partition i.e. 2 per fold for
+5-fold cross-validation, and then 3 for testing the best parameters on
+the validation set.
+
+Classifier accuracy is estimated using the macro F1 score, i.e. the
+harmonic mean of precision and recall. In the code chunk below we
 demonstrate how to optimise the free parameters, \texttt{sigma} and
@@
 <<classify>>=
-hl <- svmClassification(hl, params,
-                        class.weights = w,
-                        fcol = "markers")
-@
-
-Automatically, the output of the above classification; the organelle
+hl <- svmClassification(hl, params, class.weights = w, fcol = "markers")
+@
+
+Automatically, the output of the above classification, the organelle
 predictions and assignment scores, are stored in the
@@
 Additionally, when calling \texttt{plot2D} we can use the \texttt{cex}
-argument to change the size of each point on the plot (where one point
-represents one protein) to be inversely proportional to the SVM
-score. This gives an initial overview of the high scoring
-localisations from the SVM predictions.
+argument to change the size of each point on the plot to be inversely
+proportional to the SVM score. This gives an initial overview of the
+high scoring localisations from the SVM predictions.
 
@@
 @  
-  \caption{Classification results.}
+\caption{Classification results. Colours indicate class membership and
+  point size are representative or the classification confidence.}
   \label{fig:plotSVM}
@@
 
-\textbf{TODO} explain how choosing markers positively/negatively
-affects this, and relation with thresholding.
-
 \subsection*{Thresholding}\label{sec:thresholding}
@@
 set a specific score cutoff on which to define new assignments, below
-which classifications are set to unknown/unassigned. This is important
-as in a supervised learning setup proteins can only be predicted to be
+which classifications are kept unknown/unassigned. This is important
+as in a supervised learning setup, proteins can only be predicted to be
 localised to one of the sub-cellular niches that appear in the
 labelled training data. We can not guaranette (and do not expect) that
-the whole class diversity to be represented in the labelled training
+the whole sub-cellular diversity to be represented in the labelled training
 data as (1) finding markers that represent the whole diversity of the
@@
 protein markers) and (2) many sub-cellular niches contain too few
-proteins to train on (we recommend a minimum of 13 markers per
-sub-cellular class for stratified 80/20 partitioning and 5-fold
-cross-validation - this allows a minimum of ~10 examples for parameter
-optimisation on the training partition i.e. 2 per fold for 5-fold
-cross-validation, and then ~3 for testing the best parameters on the
-validation set). <!--Conversely, when defining markers one must also
-be careful to not label too many markers, as this can lead to
-overfitting and mis-assignment.-->
+proteins to train on (see XXX)
 
@@
 <<orgquants>>=
-ts <- orgQuants(hl, 
-                fcol = "svm", 
-                scol = "svm.scores",
-                mcol = "markers", 
-                t = .5)
-
-hl <- getPredictions(hl,
-                     fcol = "svm",
-                     scol = "svm.scores", 
-                     mcol = "markers",
-                     t = ts)                        
+ts <- orgQuants(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = .5)
+hl <- getPredictions(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = ts)
 @
@@
 
-<<resetpar>>=
-dev.off()
+<<resetpar, echo=FALSE>>=
+invisible(dev.off())
 @

### 2016-11-03 bioc-workflow.tex
Update discussion on markers
- Describe conservative vs loose definition of markers, their effect,
  and reference figure reflecting classification scores.
- Move algorithmic argument for 13 markers in optimisation section.

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 16:12:22 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 10:06:16 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov  2 16:12:22 2016 
-## Subset [6725,20][5032,20] Wed Nov  2 16:12:23 2016 
-## Removed features with more than 0 NAs: Wed Nov  2 16:12:23 2016 
-## Dropped featureData's levels Wed Nov  2 16:12:23 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 10:06:16 2016 
+## Subset [6725,20][5032,20] Thu Nov  3 10:06:16 2016 
+## Removed features with more than 0 NAs: Thu Nov  3 10:06:16 2016 
+## Dropped featureData's levels Thu Nov  3 10:06:16 2016 
 ##  MSnbase version: 1.21.7
@@
 
+We recommend at least 13 markers per sub-cellular class (see the
+\textit{Optimisation} section for details about the algorithmic
+motivation of this number). Markers should be chosen to confidently
+represent to distribution of genuine residents of a sub-cellular
+niche. We generally recommend such a conservative approach in defining
+markers to avoid false assignments when assigning sub-cellular
+localisation of proteins of unknown localisation. A more relaxed
+definition of markers, i.e. one that broadly or over-confidently
+defines markers, risks to erroneously assign proteins to a single
+location, when, in reality, they reside in more that a single location
+(including the assumed unique location). One can not expect to
+identify exact boundaries between sub-cellular classes through marker
+annotation alone; the definition of these boundaries is better handled
+algorithmically, i.e. after application of the supervised learning
+algorithm, using the prediction scores (as described in the
+\textit{Classification} section, in particular Figure
+\ref{fig:plotSVM}).
+
 If the naming between the marker sets and the \texttt{MSnSet} dataset
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Wed Nov  2 16:12:25 2016
+% Thu Nov  3 10:06:18 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov  2 16:12:24 2016 
-## Added markers from  'pdres' marker vector. Wed Nov  2 16:12:25 2016 
+## Added markers from  'mrk' marker vector. Thu Nov  3 10:06:18 2016 
+## Added markers from  'pdres' marker vector. Thu Nov  3 10:06:18 2016 
 ##  MSnbase version: 1.99.7
@@
 \begin{alltt}
-\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl, params,}
-                        \hlkwc{class.weights} \hlstd{= w,}
-                        \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-Automatically, the output of the above classification; the organelle
+\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl, params,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Automatically, the output of the above classification, the organelle
 predictions and assignment scores, are stored in the
@@
 \end{knitrout}
-  \caption{Classification results.}
+\caption{Classification results. Colours indicate class membership and
+  point size are representative or the classification confidence.}
   \label{fig:plotSVM}
@@
 \begin{alltt}
-\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,}
-                \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,}
-                \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,}
-                \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,}
-                \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
+\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
 \end{alltt}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4391696               0.3042111               0.3869018 
+##               0.4341324               0.3058579               0.3830383 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6964096               0.6130827               0.4300618 
+##               0.7027971               0.6112125               0.4319591 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4186188               0.6092759               0.9497908 
+##               0.4219879               0.5982771               0.9503209 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950972               0.7103401               0.3158138 
+##               0.7950258               0.7086536               0.3168774 
 ##         Plasma membrane              Proteasome 
-##               0.7221735               0.4144763
-\end{verbatim}
-\begin{alltt}
-\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,}
-                     \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,}
-                     \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,}
-                     \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,}
-                     \hlkwc{t} \hlstd{= ts)}
+##               0.7163298               0.4144068
+\end{verbatim}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
 \end{alltt}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     170                      89 
+##                      85                     172                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     477                     105 
+##                     297                     475                      98 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      24                     120                     522 
+##                      27                     123                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
@@
 ##         Plasma membrane              Proteasome                 unknown 
-##                     315                     158                    2059
+##                     323                     157                    2058
 \end{verbatim}
@@
 using \texttt{plot2D} function.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{dev.off}\hlstd{()}
-\end{alltt}
-\begin{verbatim}
-## null device 
-##           1
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
 

### 2016-11-03 bioc-workflow.Rnw
more markers

@@
 which classifications are kept unknown/unassigned. This is important
-as in a supervised learning setup, proteins can only be predicted to be
-localised to one of the sub-cellular niches that appear in the
-labelled training data. We can not guaranette (and do not expect) that
-the whole sub-cellular diversity to be represented in the labelled training
-data as (1) finding markers that represent the whole diversity of the
-cell is challenging (especially obtaining dual- and multiply-localised
-protein markers) and (2) many sub-cellular niches contain too few
-proteins to train on (see XXX)
+as in a supervised learning setup, proteins can only be predicted to
+be localised to one of the sub-cellular niches that appear in the
+labelled training data. We can not guarantee (and do not expect) that
+the whole sub-cellular diversity to be represented in the labelled
+training data as (1) finding markers that represent the whole
+diversity of the cell is challenging (especially obtaining dual- and
+multiply-localised protein markers) and (2) many sub-cellular niches
+contain too few proteins to train on (see above for a motivation of a
+minimum of 13 markers).
 

### 2016-11-03 bioc-workflow.Rnw
thresholding section

@@
 thresholds.  In the code chunk below we display a boxplot of the score
-distributions per organelle.
+distributions per organelle (Figure \ref{fig:threshold}).
 
@@
 containing the prediction results. The results can also be visualied
-using \texttt{plot2D} function.
+using \texttt{plot2D} function (Figure \ref{fig:plotres}).
 
@@
     \centering
-<<plotres, out.width=".8\\textwidth">>=
+<<plotres, out.width=".7\\textwidth">>=
 plot2D(hl, fcol = "svm.pred")
@@
 There is also a dedicated interactive application to help users
-examine these distributions in the \Biocpkg{pRolocGUI} package.
-This app can be launched via the \texttt{pRolocVis} function and specifying
-the argument \texttt{app = "classify"} along with the relevent \texttt{fcol}, \texttt{scol}
-and \texttt{mcol} which refer to the columns in the feature data that contain
-the new assignments, assignment scores and markers respectively (see
-also \texttt{fvarLabels(svmres)}).
+examine these distributions in the \Biocpkg{pRolocGUI} package (Figure
+\ref{fig:classifyapp}).  This app can be launched via the
+\texttt{pRolocVis} function and specifying the argument \texttt{app =
+  "classify"} along with the relevent \texttt{fcol}, \texttt{scol} and
+\texttt{mcol} which refer to the columns in the feature data that
+contain the new assignments, assignment scores and markers
+respectively (see also \texttt{fvarLabels(svmres)}).
 
@@
 library("pRolocGUI")
-pRolocVis(hl, 
-          app = "classify",
-          fcol = "svm",
-	        scol = "svm.scores",
-          mcol = "markers")
+pRolocVis(hl, app = "classify", fcol = "svm", scol = "svm.scores", mcol = "markers")
 @
@@
   \centering
-  \includegraphics[width=\textwidth]{./Figures/classify.png}
-  \caption{The classify application}
+  \includegraphics[width=0.7\textwidth]{./Figures/classify.png}
+  \caption{The classify application enable the interactive exploration
+    of classification score thresholding.}
   \label{fig:classifyapp}
@@
 The data is loaded and displayed on a PCA plot and a boxplot is used
-to display the classifier scores by data class. On the left there is a
-sidebar panel with sliders to control the thresholds upon which
+to display the classifier scores by data class. On the left, there is
+a sidebar panel with sliders to control the thresholds upon which
 classifications are made. There are two types of cut-off that the user
-can choose from: (1) "Quantile" and (2) "User-defined". By default,
-when the application is launched quatile scoring is selected and set
-to 0.5, the median. The class-specific score thresholds that
-correspond to selecting the desired quantile are shown on as red dots
-on the boxplot. The assignments on the PCA plot are also updated
-according to the selected threshold. The quantile threshold can be set
-by moving the corresponding quantile slider. If one wished to set
-their own cut-offs the "User-defined" radio button must be selected
-and then the sliders for defining user-specified scores become active
-and the scores and highlighted on the boxplot by blue dots. For more
-information we refer users to the \Biocpkg{pRolocGUI}
-tutorial
+can choose from: (1) \textit{Quantile} and (2)
+\textit{User-defined}. By default, when the application is launched
+quatile scoring is selected and set to 0.5, the median. The
+class-specific score thresholds that correspond to selecting the
+desired quantile are shown on as red dots on the boxplot. The
+assignments on the PCA plot are also updated according to the selected
+threshold. The quantile threshold can be set by moving the
+corresponding quantile slider. If one wished to set their own cut-offs
+the \textit{User-defined} radio button must be selected and then the
+sliders for defining user-specified scores become active and the
+scores and highlighted on the boxplot by blue dots. For more
+information we refer users to the \Biocpkg{pRolocGUI} tutorial
 \href{http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html}{vignette}.
+
+\clearpage
 

### 2016-11-03 bioc-workflow.tex
thresholding section

@@
 ## - - - Processing information - - -
-##  MSnbase version: 1.99.7
+##  MSnbase version: 2.0.0
 \end{verbatim}
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 10:06:16 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:25:04 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 10:06:16 2016 
-## Subset [6725,20][5032,20] Thu Nov  3 10:06:16 2016 
-## Removed features with more than 0 NAs: Thu Nov  3 10:06:16 2016 
-## Dropped featureData's levels Thu Nov  3 10:06:16 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:25:04 2016 
+## Subset [6725,20][5032,20] Thu Nov  3 17:25:05 2016 
+## Removed features with more than 0 NAs: Thu Nov  3 17:25:05 2016 
+## Dropped featureData's levels Thu Nov  3 17:25:05 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Thu Nov  3 10:06:18 2016
+% Thu Nov  3 17:25:06 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Nov  3 10:06:18 2016 
-## Added markers from  'pdres' marker vector. Thu Nov  3 10:06:18 2016 
-##  MSnbase version: 1.99.7
+## Added markers from  'mrk' marker vector. Thu Nov  3 17:25:06 2016 
+## Added markers from  'pdres' marker vector. Thu Nov  3 17:25:07 2016 
+##  MSnbase version: 2.0.0
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4341324               0.3058579               0.3830383 
+##               0.4362653               0.3034648               0.3816205 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.7027971               0.6112125               0.4319591 
+##               0.6867997               0.6097258               0.4288042 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4219879               0.5982771               0.9503209 
+##               0.4156074               0.5909675               0.9497999 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950258               0.7086536               0.3168774 
+##               0.7967810               0.7107021               0.3135143 
 ##         Plasma membrane              Proteasome 
-##               0.7163298               0.4144068
+##               0.7125828               0.4223579
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      84 
+##                      85                     170                      86 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     297                     475                      98 
+##                     298                     476                      97 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     522 
+##                      27                     124                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     342                      39 
+##                     229                     347                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     323                     157                    2058
+##                     320                     155                    2057
 \end{verbatim}
@@
 \end{kframe}
-\includegraphics[width=.8\textwidth]{figure/plotres-1}
+\includegraphics[width=.7\textwidth]{figure/plotres-1}
 
@@
 \hlkwd{library}\hlstd{(}\hlstr{"pRolocGUI"}\hlstd{)}
-\hlkwd{pRolocVis}\hlstd{(hl,}
-          \hlkwc{app} \hlstd{=} \hlstr{"classify"}\hlstd{,}
-          \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,}
-                \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,}
-          \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\hlkwd{pRolocVis}\hlstd{(hl,} \hlkwc{app} \hlstd{=} \hlstr{"classify"}\hlstd{,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
 \end{alltt}
@@
 ## other attached packages:
-##  [1] pRolocdata_1.11.9    pRoloc_1.15.1        MLInterfaces_1.53.1 
-##  [4] cluster_2.0.5        annotate_1.51.1      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.35.4 IRanges_2.7.17       S4Vectors_0.11.19   
-## [10] MSnbase_1.99.7       ProtGenerics_1.5.1   BiocParallel_1.7.9  
-## [13] mzR_2.7.12           Rcpp_0.12.7          Biobase_2.33.4      
-## [16] BiocGenerics_0.19.2  gridExtra_2.2.1      xtable_1.8-2        
-## [19] BiocStyle_2.1.33     knitr_1.14          
+##  [1] pRolocdata_1.12.0    pRoloc_1.15.1        MLInterfaces_1.54.0 
+##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.4        
+##  [7] AnnotationDbi_1.36.0 IRanges_2.8.0        S4Vectors_0.12.0    
+## [10] MSnbase_2.0.0        ProtGenerics_1.6.0   BiocParallel_1.8.1  
+## [13] mzR_2.8.0            Rcpp_0.12.7          Biobase_2.34.0      
+## [16] BiocGenerics_0.20.0  gridExtra_2.2.1      xtable_1.8-2        
+## [19] BiocStyle_2.2.0      knitr_1.14          
 ## 
@@
 ##  [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-16         
-## [10] MatrixModels_0.4-1    affyio_1.43.0         flexmix_2.3-13       
+## [10] MatrixModels_0.4-1    affyio_1.44.0         flexmix_2.3-13       
 ## [13] mvtnorm_1.0-5         codetools_0.2-15      splines_3.3.2        
-## [16] doParallel_1.0.10     impute_1.47.0         robustbase_0.92-6    
+## [16] doParallel_1.0.10     impute_1.48.0         robustbase_0.92-6    
 ## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-72         
 ## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-25       
-## [25] vsn_3.41.5            sfsmisc_1.1-0         shiny_0.14.1         
+## [25] vsn_3.42.3            sfsmisc_1.1-0         shiny_0.14.2         
 ## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
-## [31] limma_3.29.21         formatR_1.4           htmltools_0.3.5      
+## [31] limma_3.30.2          formatR_1.4           htmltools_0.3.5      
 ## [34] quantreg_5.29         tools_3.3.2           ggvis_0.4.3          
-## [37] gtable_0.2.0          affy_1.51.1           reshape2_1.4.1       
+## [37] gtable_0.2.0          affy_1.52.0           reshape2_1.4.2       
 ## [40] dplyr_0.5.0           MALDIquant_1.15       trimcluster_0.1-2    
-## [43] gdata_2.17.0          preprocessCore_1.35.0 nlme_3.1-128         
+## [43] gdata_2.17.0          preprocessCore_1.36.0 nlme_3.1-128         
 ## [46] iterators_1.0.8       fpc_2.1-10            stringr_1.1.0        
@@
 ## [52] gtools_3.5.0          dendextend_1.3.0      DEoptimR_1.0-6       
-## [55] zlibbioc_1.19.0       MASS_7.3-45           scales_0.4.0         
-## [58] BiocInstaller_1.24.0  pcaMethods_1.65.0     SparseM_1.72         
-## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.29.2       
+## [55] zlibbioc_1.20.0       MASS_7.3-45           scales_0.4.0         
+## [58] BiocInstaller_1.24.0  pcaMethods_1.66.0     SparseM_1.72         
+## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.30.0       
 ## [64] rpart_4.1-10          stringi_1.1.2         RSQLite_1.0.0        
-## [67] highr_0.6             genefilter_1.55.2     randomForest_4.6-12  
+## [67] highr_0.6             genefilter_1.56.0     randomForest_4.6-12  
 ## [70] foreach_1.4.3         e1071_1.6-7           prabclus_2.2-6       
-## [73] bitops_1.0-6          mzID_1.11.2           evaluate_0.10        
+## [73] bitops_1.0-6          mzID_1.12.0           evaluate_0.10        
 ## [76] lattice_0.20-34       htmlwidgets_0.7       gbm_2.1.1            
@@
 ## [82] DBI_0.5-1             whisker_0.3-2         mgcv_1.8-15          
-## [85] survival_2.39-5       RCurl_1.95-4.8        nnet_7.3-12          
+## [85] survival_2.40-1       RCurl_1.95-4.8        nnet_7.3-12          
 ## [88] tibble_1.2            msdata_0.14.0         car_2.1-3            

### 2016-11-03 bioc-workflow.Rnw
tl optimisation

@@
 
-In addition to high quality MS-based quantitative proteomics data
+In addition to high quality MS-based quantitative proteomics data,
 there exist a number of other sources of information that are freely
@@
 provides functionality to do such analyses. We recently developed two
-transfer learning algorithms using a *k*-NN and SVM framework and
+transfer learning algorithms using a $k$-NN and SVM framework and
 applied them to the task of protein localisation prediction
@@
 
-In TL one typically has a primary task that they wish to solve, and
+In TL one typically has a primary task that one wishes to solve, and
 some complementary (often heterogeneous) auxiliary information that is
@@
 retrieved and used to create a binary matrix where a 1 (0) at position
-(i; j) indicates that term j has (not) been used to annotate protein
-i. The GO terms are retrieved from an appropriate repository using the
-\Biocpkg{biomaRt} package. The specific Biomart repository and query
-will depend on the species under study and the type of
+$(i, j)$ indicates that term $j$ has (not) been used to annotate
+protein $i$. The GO terms are retrieved from an appropriate repository
+using the \Biocpkg{biomaRt} package. The specific Biomart repository
+and query will depend on the species under study and the type of
 identifiers. The first step is to construct the annotation parameters
-that will enable to perform the query, this is done using
+that will enable to perform the query, which is done using
 \texttt{setAnnotationParams}. Typing into the R console \texttt{par <-
@@
 possible to pass patterns to match the species e.g. in the code chunk
-below we pass "Mus musculus", and the identifier type for our data
-(see \texttt{featureNames(hl)}) which is "Uniprot/Swissprot", for the
-Biomart query.
+below we pass \texttt{"Mus musculus"}, and the identifier type for our
+data (see \texttt{featureNames(hl)}) which is
+\texttt{"Uniprot/Swissprot"}, for the Biomart query.
 
 <<setap, eval=FALSE>>=
-par <- setAnnotationParams(inputs = c("Mus musculus", 
-                                      "UniProt/Swissprot"))
+par <- setAnnotationParams(inputs = c("Mus musculus", "UniProt/Swissprot"))
 
@@
 <<makego, eval=FALSE>>=
-gocc <- makeGoSet(hl, params = par, 
-                  namespace = "cellular_component",
-                  evidence = NULL)
+gocc <- makeGoSet(hl, params = par, namespace = "cellular_component")
 @
@@
 GO terms retreived is also dependent on the database version queried
-and thus is always subject to change. We find it is common to see many
-GO terms with only one protein assigned to that term, these such terms
-generally do not bring and information generally for building the
-classifier and thus we can remove such GO terms by using the function
-\texttt{filterBinMSnSet}.
+and thus is always subject to change. We find it is common to see GO
+terms with only one protein assigned to that term. Such terms do not
+bring any information for building the classifier and are thus removes
+using the \texttt{filterBinMSnSet} function.
 
@@
 
-Now we have generated our auxiliary data we can use the \textit{k}-NN
-implementation of transfer learning available in \Biocpkg{pRoloc} to
-integrate this with our primary MS-based quantitative proteomics data
-using the functions \texttt{knntlOptimisation} to estimate the
-free-parameters for the integration, and \texttt{knntlClassification}
-to do the predictions. We have shown that using transfer learning in
-the context of spatial proteomics results in the assignment of
-proteins to sub-cellular niches with a higher generalisation accuracy
-than using standard supervised machine learning with a single source
-of information \cite{Breckels:2016}.
+Now that we have generated our auxiliary data, we can use the
+\textit{k}-NN implementation of transfer learning available in
+\Biocpkg{pRoloc} to integrate this with our primary MS-based
+quantitative proteomics data using the functions
+\texttt{knntlOptimisation} to estimate the free-parameters for the
+integration, and \texttt{knntlClassification} to do the
+predictions. We have shown that using transfer learning results in the
+assignment of proteins to sub-cellular niches with a higher
+generalisation accuracy than using standard supervised machine
+learning with a single source of information \cite{Breckels:2016}.
 
 The first step, as with any machine learning algorithm, is to optimise
-any free paramaters of the classifier. For the *k*-NN TL classifier
+any free paramaters of the classifier. For the $k$-NN TL classifier
 there are two sets of parameters that need optimising: the first set
@@
 the nearest neighbour calculations for each data source. The second
-set of parameters (which is noted by a vector of $\theta$ weights)
-that require optimising are the class weights, one per subcellular
-niche, that control the proportion of primary and auxiliary data to
-use for learning. A weight can take any real value number between 0
-and 1. A weight of $\theta = 1$ indicates that all weight is given to
-the primary data (and this implicitly implies that a weight of
-$1 - \theta$ is given to the auxiliary data, similarly a weight of
-$\theta = 0$ implies that all weight is given to the auxiliary data
-(so 0 is given to the primary source). If we conduct a parameter seach
-and test weights $\theta = {0, 1/3, 2/3, 1}$ for each class, and if we
-have, for example 10 subcellular niches, this will result in
-\texttt{4$^{10}$} different combinations of parameters to test. The
-parameter optimisation is therefore time consuming and as such we
-recommend you make use of a computing cluster (code and submissing
-scripts are also available in the supporting information). The markers
-in the \texttt{hl} dataset contain 14 subcellular classes. If we
-examine these markers and classes on the PCA plot above we can see
-that in particular the two ribosomes and two nuclear compartments are
-highly separated along the first two components, this is also evident
-from the profiles plot which gives us a good indication that these
-subcellular niches are well-resolved in the hyperLOPIT
-dataset. Transfer learning is particularly useful for classes that are
-not as well separated, we find that subcellular niches that are
-well-separated under hyperLOPIT and LOPIT obtain a class score of 1
-(i.e. use only primary data from transfer learning
+set of parameters (noted by a vector of $\theta$ weights) that require
+optimising are the class weights, one per subcellular niche, that
+control the proportion of primary and auxiliary data to use for
+learning. A weight can take any real value number between 0 and 1. A
+weight of $\theta = 1$ indicates that all weight is given to the
+primary data (and this implicitly implies that a weight of
+$1 - \theta = 0$ is given to the auxiliary data), and similarly a
+weight of $\theta = 0$ implies that all weight is given to the
+auxiliary data (so 0 is given to the primary source). If we conduct a
+parameter seach and test weights $\theta = {0, 1/3, 2/3, 1}$ for each
+class, and if we have, for example 10 subcellular niches, this will
+result in \texttt{4$^{10}$} different combinations of parameters to
+test. The parameter optimisation is therefore time consuming and as
+such we recommend making use of a computing cluster (code and
+submissing scripts are also available in the supporting
+information). The markers in the \texttt{hl} dataset contain 14
+subcellular classes. If we examine these markers and classes on the
+PCA plot above we can see that in particular the two ribosomes and two
+nuclear compartments are highly separated along the first two
+components, this is also evident from the profiles plot which gives us
+a good indication that these subcellular niches are well-resolved in
+the hyperLOPIT dataset. Transfer learning is particularly useful for
+classes that are not as well separated. We find that subcellular
+niches that are well-separated under hyperLOPIT and LOPIT obtain a
+class score of 1 (i.e. use only primary data from transfer learning
 \cite{Breckels:2016}). Therefore, for the optimisation stage of the
@@
 this is not run live as this the \texttt{hl} dataset with 10 classes,
-707 markers and 4$^{10}$ combinations of parameters takes ~76 hours to
-run on the University of Cambridge HPC using 256 workers).
+707 markers and 4$^{10}$ combinations of parameters takes around 76
+hours to run on the University of Cambridge HPC using 256 workers).
 
@@
 
-Optimisation stage 1: run \texttt{knnOptimisation} to get the best
-$k$'s for each data source.
+\paragraph{TL optimisation stage 1} Run \texttt{knnOptimisation} to get
+the best $k$'s for each data source.
 
@@
 kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
-plot(kpopt)
-plot(kaopt)
-@
-
-From examining the parameter seach plots we find the best $k$'s for
-both the primary and auxiliary are 3.
-
-Optimisation stage 2: run \texttt{knntlOptimisation} to get the best
-transfer learning weights for each sub-cellular class.
+@
+
+From examining the parameter seach plots as described in section
+\textit{Optimisation}, we find the best $k$'s for both the primary and
+auxiliary are 3.
+
+\paragraph{TL optimisation stage 2} Run \texttt{knntlOptimisation} to
+get the best transfer learning weights for each sub-cellular class.
 
@@
 ## Now peform tl optimisation 
-tlopt <- knntlOptimisation(hl, gocc,
-                           fcol = "tlmarkers",
-                           length.out = 4,
-                           times = 50,
+tlopt <- knntlOptimisation(hl, gocc, fcol = "tlmarkers",
+                           length.out = 4, times = 50,
                            xval = 5, k = c(3, 3),
@@
 
-The results of the optimisation can be visalised using the \texttt{plot}
-method for \texttt{"ThetaRegRes"} objects:
+The results of the optimisation can be visalised using the
+\texttt{plot} method for \texttt{tlopt} optimisation result:
 
@@
 
-Looking at the bubble plot displaying the distribution of best weights
-over the 50 runs we find that for many of the subcellular niches a
-weight of 1 is most popular (i.e. use only primary hyperLOPIT data in
-classification), this is unsuprising as we already know the dataset is
-well resolved for these classes. We see that the most popular weights
-for the proteasome and lysosome tend to be towards 0, indicating that
-these niches are well-resolved in the Gene Ontology. This tells us
-that we would benefit from including auxiliary GO information in our
-classifier for these subcellular compartments. The plasma membrane
-weights are relatively equally spread between using hyperLOPIT and GO
-data. Using the \texttt{getParams} function we can return the best
-weights and then use this as input for the classification.
+\paragraph{TL classification} Looking at the bubble plot displaying
+the distribution of best weights over the 50 runs we find that for
+many of the subcellular niches a weight of 1 is most popular (i.e. use
+only primary hyperLOPIT data in classification), this is unsuprising
+as we already know the dataset is well resolved for these classes. We
+see that the most popular weights for the proteasome and lysosome tend
+to be towards 0, indicating that these niches are well-resolved in the
+Gene Ontology. This tells us that we would benefit from including
+auxiliary GO information in our classifier for these subcellular
+compartments. The plasma membrane weights are relatively equally
+spread between using hyperLOPIT and GO data. Using the
+\texttt{getParams} function we can return the best weights and then
+use this as input for the classification.
 
@@
 ## Do the classification
-hl <- knntlClassification(hl, gocc, bestTheta = bestpar, 
-                          k = c(3, 3))
+hl <- knntlClassification(hl, gocc, bestTheta = bestpar, k = c(3, 3))
 @
@@
 values to a tab (the defualt, or other type e.g.e.g. comma, as
-specified in \texttt{write,table}) separated file. There argument
+specified in \texttt{write.table}) separated file. There argument
 \texttt{fDataCols} can be used to specify which \texttt{featureData}

### 2016-11-03 bioc-workflow.tex
tl optimisation

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:25:04 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:54:23 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:25:04 2016 
-## Subset [6725,20][5032,20] Thu Nov  3 17:25:05 2016 
-## Removed features with more than 0 NAs: Thu Nov  3 17:25:05 2016 
-## Dropped featureData's levels Thu Nov  3 17:25:05 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:54:23 2016 
+## Subset [6725,20][5032,20] Thu Nov  3 17:54:24 2016 
+## Removed features with more than 0 NAs: Thu Nov  3 17:54:24 2016 
+## Dropped featureData's levels Thu Nov  3 17:54:24 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Thu Nov  3 17:25:06 2016
+% Thu Nov  3 17:54:25 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Nov  3 17:25:06 2016 
-## Added markers from  'pdres' marker vector. Thu Nov  3 17:25:07 2016 
+## Added markers from  'mrk' marker vector. Thu Nov  3 17:54:25 2016 
+## Added markers from  'pdres' marker vector. Thu Nov  3 17:54:26 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4362653               0.3034648               0.3816205 
+##               0.4358215               0.3027190               0.3838820 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6867997               0.6097258               0.4288042 
+##               0.6873138               0.5919056               0.4206391 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4156074               0.5909675               0.9497999 
+##               0.4163956               0.5946202               0.9493401 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7967810               0.7107021               0.3135143 
+##               0.7952464               0.7053250               0.3164791 
 ##         Plasma membrane              Proteasome 
-##               0.7125828               0.4223579
+##               0.7128863               0.4133411
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      86 
+##                      85                     171                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     298                     476                      97 
+##                     298                     470                     102 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     522 
+##                      27                     125                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     347                      39 
+##                     229                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     320                     155                    2057
+##                     315                     158                    2058
 \end{verbatim}
@@
 possible to pass patterns to match the species e.g. in the code chunk
-below we pass "Mus musculus", and the identifier type for our data
-(see \texttt{featureNames(hl)}) which is "Uniprot/Swissprot", for the
-Biomart query.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{par} \hlkwb{<-} \hlkwd{setAnnotationParams}\hlstd{(}\hlkwc{inputs} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"Mus musculus"}\hlstd{,}
-                                      \hlstr{"UniProt/Swissprot"}\hlstd{))}
+below we pass \texttt{"Mus musculus"}, and the identifier type for our
+data (see \texttt{featureNames(hl)}) which is
+\texttt{"Uniprot/Swissprot"}, for the Biomart query.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{par} \hlkwb{<-} \hlkwd{setAnnotationParams}\hlstd{(}\hlkwc{inputs} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"Mus musculus"}\hlstd{,} \hlstr{"UniProt/Swissprot"}\hlstd{))}
 \end{alltt}
@@
 \begin{alltt}
-\hlstd{gocc} \hlkwb{<-} \hlkwd{makeGoSet}\hlstd{(hl,} \hlkwc{params} \hlstd{= par,}
-                  \hlkwc{namespace} \hlstd{=} \hlstr{"cellular_component"}\hlstd{,}
-                  \hlkwc{evidence} \hlstd{=} \hlkwa{NULL}\hlstd{)}
+\hlstd{gocc} \hlkwb{<-} \hlkwd{makeGoSet}\hlstd{(hl,} \hlkwc{params} \hlstd{= par,} \hlkwc{namespace} \hlstd{=} \hlstr{"cellular_component"}\hlstd{)}
 \end{alltt}
@@
 \hlstd{kaopt} \hlkwb{<-} \hlkwd{knnOptimisation}\hlstd{(gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
-\hlkwd{plot}\hlstd{(kpopt)}
-\hlkwd{plot}\hlstd{(kaopt)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-From examining the parameter seach plots we find the best $k$'s for
-both the primary and auxiliary are 3.
-
-Optimisation stage 2: run \texttt{knntlOptimisation} to get the best
-transfer learning weights for each sub-cellular class.
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+From examining the parameter seach plots as described in section
+\textit{Optimisation}, we find the best $k$'s for both the primary and
+auxiliary are 3.
+
+\paragraph{TL optimisation stage 2} Run \texttt{knntlOptimisation} to
+get the best transfer learning weights for each sub-cellular class.
 
@@
 \hlcom{## Now peform tl optimisation }
-\hlstd{tlopt} \hlkwb{<-} \hlkwd{knntlOptimisation}\hlstd{(hl, gocc,}
-                           \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{,}
-                           \hlkwc{length.out} \hlstd{=} \hlnum{4}\hlstd{,}
-                           \hlkwc{times} \hlstd{=} \hlnum{50}\hlstd{,}
+\hlstd{tlopt} \hlkwb{<-} \hlkwd{knntlOptimisation}\hlstd{(hl, gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{,}
+                           \hlkwc{length.out} \hlstd{=} \hlnum{4}\hlstd{,} \hlkwc{times} \hlstd{=} \hlnum{50}\hlstd{,}
                            \hlkwc{xval} \hlstd{=} \hlnum{5}\hlstd{,} \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{),}
@@
 \hlcom{## Do the classification}
-\hlstd{hl} \hlkwb{<-} \hlkwd{knntlClassification}\hlstd{(hl, gocc,} \hlkwc{bestTheta} \hlstd{= bestpar,}
-                          \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{))}
+\hlstd{hl} \hlkwb{<-} \hlkwd{knntlClassification}\hlstd{(hl, gocc,} \hlkwc{bestTheta} \hlstd{= bestpar,} \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{))}
 \end{alltt}

### 2016-11-04 bioc-workflow.Rnw
mention processingData early on

@@
 \end{figure}
+
+Another slot of interest is \texttt{processingData}, that logs the
+processing \texttt{MSnSet} objects undergo. The processing log can be
+accessed with the \texttt{processingData} function and is displayed
+under \textit{Processing information} in the textual object summary
+when an \texttt{MSnSet}'s name it typed in the R console.
 
@@
 \texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. We can see by typing
-\texttt{processingData(hl)} directly into the console the processing
-information has been updated to the \texttt{MSnSet} recording the
-parameters that were used to run the analyses.  This is handy for
-keeping track of data analyses. The results can be displayed by using
-the \texttt{getMarkers} function. We see that 5 new phenotype data
-clusters were found.
+\texttt{featureData} under the name \texttt{pd}. The results can be
+displayed by using the \texttt{getMarkers} function. We see that 5 new
+phenotype data clusters were found.
 

### 2016-11-04 bioc-workflow.Rnw
udpates, related to issues

@@
   \centering
-<<qcplot, message=FALSE, warning=FALSE, out.width=".65\\textwidth">>=
+<<qcplot, fig.show = "hold", message=FALSE, warning=FALSE, out.width=".45\\textwidth">>=
 library("pRoloc")
 plot2D(hl, fcol = NULL, col = "black")
-@  
-  \caption{PCA plot of the mouse stem cell data \texttt{hl}.}
+plot2D(hl, method = "hexbin")
+@  
+\caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
+  represent a single protein, and cluster of proteins represent
+  proteins residing in the same sub-cellular niche. The figure on the
+  right bins proteins and represent the bins density to highlight the
+  presence of protein clusters. }
   \label{fig:pcahl}
@@
 using the \texttt{fDataTounknown} function. It is often used to
-replace empty strings ("") or missing values in the markers definition
+replace empty strings (\texttt{""}) or missing values in the markers definition
 to a common definition of \textit{unknown} localisation.
@@
 @  
-  \caption{Visualisation of the transfer learning parameter optimisation procedure.}
+\caption{Visualisation of the transfer learning parameter optimisation
+  procedure. Each row displays the frequency of observed weights
+  (along the columns) for a specific sub-cellular class, with large
+  dots representing higher observation frequencies.}
   \label{fig:plottl}
@@
 
+The source of this document, including the code necessary to reproduce
+the analyses and figures is available in the public manuscript
+repository \cite{ghrepo}.
+
 \bigskip

### 2016-11-04 bioc-workflow.tex
udpates, related to issues

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:54:23 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 11:36:17 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Nov  3 17:54:23 2016 
-## Subset [6725,20][5032,20] Thu Nov  3 17:54:24 2016 
-## Removed features with more than 0 NAs: Thu Nov  3 17:54:24 2016 
-## Dropped featureData's levels Thu Nov  3 17:54:24 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 11:36:17 2016 
+## Subset [6725,20][5032,20] Fri Nov  4 11:36:18 2016 
+## Removed features with more than 0 NAs: Fri Nov  4 11:36:18 2016 
+## Dropped featureData's levels Fri Nov  4 11:36:18 2016 
 ##  MSnbase version: 1.21.7
@@
 \hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlkwa{NULL}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"black"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=.65\textwidth]{figure/qcplot-1}
-
-\end{knitrout}
-  \caption{PCA plot of the mouse stem cell data \texttt{hl}.}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"hexbin"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.45\textwidth]{figure/qcplot-1} 
+\includegraphics[width=.45\textwidth]{figure/qcplot-2}
+
+\end{knitrout}
+\caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
+  represent a single protein, and cluster of proteins represent
+  proteins residing in the same sub-cellular niche. The figure on the
+  right bins proteins and represent the bins density to highlight the
+  presence of protein clusters. }
   \label{fig:pcahl}
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Thu Nov  3 17:54:25 2016
+% Fri Nov  4 11:36:20 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Nov  3 17:54:25 2016 
-## Added markers from  'pdres' marker vector. Thu Nov  3 17:54:26 2016 
+## Added markers from  'mrk' marker vector. Fri Nov  4 11:36:19 2016 
+## Added markers from  'pdres' marker vector. Fri Nov  4 11:36:20 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4358215               0.3027190               0.3838820 
+##               0.4344085               0.3070891               0.3823223 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6873138               0.5919056               0.4206391 
+##               0.6928348               0.6017963               0.4249301 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4163956               0.5946202               0.9493401 
+##               0.4226011               0.5911111               0.9498226 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7952464               0.7053250               0.3164791 
+##               0.7957737               0.7096282               0.3152061 
 ##         Plasma membrane              Proteasome 
-##               0.7128863               0.4133411
+##               0.7053008               0.4133848
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      88 
+##                      84                     170                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     298                     470                     102 
+##                     301                     479                      94 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     125                     523 
+##                      27                     123                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     344                      39 
+##                     230                     341                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     315                     158                    2058
+##                     326                     154                    2058
 \end{verbatim}
@@
 \end{knitrout}
-  \caption{Visualisation of the transfer learning parameter optimisation procedure.}
+\caption{Visualisation of the transfer learning parameter optimisation
+  procedure. Each row displays the frequency of observed weights
+  (along the columns) for a specific sub-cellular class, with large
+  dots representing higher observation frequencies.}
   \label{fig:plottl}
@@
 ## other attached packages:
-##  [1] pRolocdata_1.12.0    pRoloc_1.15.1        MLInterfaces_1.54.0 
+##  [1] pRolocdata_1.12.0    pRoloc_1.15.3        MLInterfaces_1.54.0 
 ##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.4        
@@
 ##   [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-16         
-## [10] MatrixModels_0.4-1    affyio_1.44.0         flexmix_2.3-13       
-## [13] mvtnorm_1.0-5         codetools_0.2-15      splines_3.3.2        
-## [16] doParallel_1.0.10     impute_1.48.0         robustbase_0.92-6    
-## [19] jsonlite_1.1          nloptr_1.0.4          caret_6.0-72         
-## [22] pbkrtest_0.4-6        rda_1.0.2-2           kernlab_0.9-25       
-## [25] vsn_3.42.3            sfsmisc_1.1-0         shiny_0.14.2         
-## [28] sampling_2.7          assertthat_0.1        Matrix_1.2-7.1       
-## [31] limma_3.30.2          formatR_1.4           htmltools_0.3.5      
-## [34] quantreg_5.29         tools_3.3.2           ggvis_0.4.3          
-## [37] gtable_0.2.0          affy_1.52.0           reshape2_1.4.2       
-## [40] dplyr_0.5.0           MALDIquant_1.15       trimcluster_0.1-2    
-## [43] gdata_2.17.0          preprocessCore_1.36.0 nlme_3.1-128         
-## [46] iterators_1.0.8       fpc_2.1-10            stringr_1.1.0        
-## [49] lme4_1.1-12           lpSolve_5.6.13        mime_0.5             
-## [52] gtools_3.5.0          dendextend_1.3.0      DEoptimR_1.0-6       
-## [55] zlibbioc_1.20.0       MASS_7.3-45           scales_0.4.0         
-## [58] BiocInstaller_1.24.0  pcaMethods_1.66.0     SparseM_1.72         
-## [61] RColorBrewer_1.1-2    ggplot2_2.1.0         biomaRt_2.30.0       
-## [64] rpart_4.1-10          stringi_1.1.2         RSQLite_1.0.0        
-## [67] highr_0.6             genefilter_1.56.0     randomForest_4.6-12  
-## [70] foreach_1.4.3         e1071_1.6-7           prabclus_2.2-6       
-## [73] bitops_1.0-6          mzID_1.12.0           evaluate_0.10        
-## [76] lattice_0.20-34       htmlwidgets_0.7       gbm_2.1.1            
-## [79] plyr_1.8.4            magrittr_1.5          R6_2.2.0             
-## [82] DBI_0.5-1             whisker_0.3-2         mgcv_1.8-15          
-## [85] survival_2.40-1       RCurl_1.95-4.8        nnet_7.3-12          
-## [88] tibble_1.2            msdata_0.14.0         car_2.1-3            
-## [91] mlbench_2.1-1         grid_3.3.2            FNN_1.1              
-## [94] ModelMetrics_1.1.0    threejs_0.2.2         digest_0.6.10        
-## [97] diptest_0.75-7        httpuv_1.3.3          munsell_0.4.3
+##  [10] hexbin_1.27.1         MatrixModels_0.4-1    affyio_1.44.0        
+##  [13] flexmix_2.3-13        mvtnorm_1.0-5         codetools_0.2-15     
+##  [16] splines_3.3.2         doParallel_1.0.10     impute_1.48.0        
+##  [19] robustbase_0.92-6     jsonlite_1.1          nloptr_1.0.4         
+##  [22] caret_6.0-72          pbkrtest_0.4-6        rda_1.0.2-2          
+##  [25] kernlab_0.9-25        vsn_3.42.3            sfsmisc_1.1-0        
+##  [28] shiny_0.14.2          sampling_2.7          assertthat_0.1       
+##  [31] Matrix_1.2-7.1        limma_3.30.2          formatR_1.4          
+##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.2          
+##  [37] ggvis_0.4.3           gtable_0.2.0          affy_1.52.0          
+##  [40] reshape2_1.4.2        dplyr_0.5.0           MALDIquant_1.15      
+##  [43] trimcluster_0.1-2     gdata_2.17.0          preprocessCore_1.36.0
+##  [46] nlme_3.1-128          iterators_1.0.8       fpc_2.1-10           
+##  [49] stringr_1.1.0         lme4_1.1-12           lpSolve_5.6.13       
+##  [52] mime_0.5              gtools_3.5.0          dendextend_1.3.0     
+##  [55] DEoptimR_1.0-6        zlibbioc_1.20.0       MASS_7.3-45          
+##  [58] scales_0.4.0          BiocInstaller_1.24.0  pcaMethods_1.66.0    
+##  [61] SparseM_1.72          RColorBrewer_1.1-2    ggplot2_2.1.0        
+##  [64] biomaRt_2.30.0        rpart_4.1-10          stringi_1.1.2        
+##  [67] RSQLite_1.0.0         highr_0.6             genefilter_1.56.0    
+##  [70] randomForest_4.6-12   foreach_1.4.3         e1071_1.6-7          
+##  [73] prabclus_2.2-6        bitops_1.0-6          mzID_1.12.0          
+##  [76] evaluate_0.10         lattice_0.20-34       htmlwidgets_0.7      
+##  [79] gbm_2.1.1             plyr_1.8.4            magrittr_1.5         
+##  [82] R6_2.2.0              DBI_0.5-1             whisker_0.3-2        
+##  [85] mgcv_1.8-15           survival_2.40-1       RCurl_1.95-4.8       
+##  [88] nnet_7.3-12           tibble_1.2            msdata_0.14.0        
+##  [91] car_2.1-3             mlbench_2.1-1         grid_3.3.2           
+##  [94] FNN_1.1               ModelMetrics_1.1.0    threejs_0.2.2        
+##  [97] digest_0.6.10         diptest_0.75-7        httpuv_1.3.3         
+## [100] munsell_0.4.3
 \end{verbatim}
@@
 \href{https://support.bioconductor.org/}{question} at hand.
+
+The source of this document, including the code necessary to reproduce
+the analyses and figures is available in the public manuscript
+repository \cite{ghrepo}.
 

### 2016-11-04 bioc-workflow.Rnw
link to pRolocVis gif

@@
   \includegraphics[width=\textwidth]{./Figures/mainapp.png}
-  \caption{A screen shot of clickable interface and zoomable PCA plot of the main app in the \Biocpkg{pRolocGUI} package.}
+  \caption{A screen shot of clickable interface and zoomable PCA plot
+    of the main app in the \Biocpkg{pRolocGUI} package. }
   \label{fig:app}
@@
 selection tab provides an interface to control data table column
-selection.
+selection. A short
+animation\footnote{\url{https://github.com/lmsimp/bioc-pRoloc-hyperLOPIT-workflow/blob/master/Figures/pRolocVis\_pca.gif}}
+illustrating the interface is available in the manuscript repository~\cite{ghrepo}.
 

### 2016-11-04 bioc-workflow.tex
link to pRolocVis gif

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 11:36:17 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 12:03:40 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 11:36:17 2016 
-## Subset [6725,20][5032,20] Fri Nov  4 11:36:18 2016 
-## Removed features with more than 0 NAs: Fri Nov  4 11:36:18 2016 
-## Dropped featureData's levels Fri Nov  4 11:36:18 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 12:03:40 2016 
+## Subset [6725,20][5032,20] Fri Nov  4 12:03:41 2016 
+## Removed features with more than 0 NAs: Fri Nov  4 12:03:41 2016 
+## Dropped featureData's levels Fri Nov  4 12:03:41 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Nov  4 11:36:20 2016
+% Fri Nov  4 12:03:43 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Nov  4 11:36:19 2016 
-## Added markers from  'pdres' marker vector. Fri Nov  4 11:36:20 2016 
+## Added markers from  'mrk' marker vector. Fri Nov  4 12:03:43 2016 
+## Added markers from  'pdres' marker vector. Fri Nov  4 12:03:43 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4344085               0.3070891               0.3823223 
+##               0.4369956               0.3026572               0.3845674 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6928348               0.6017963               0.4249301 
+##               0.6735594               0.6449643               0.4375399 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4226011               0.5911111               0.9498226 
+##               0.4123260               0.6045583               0.9494543 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7957737               0.7096282               0.3152061 
+##               0.7957342               0.7090410               0.3143893 
 ##         Plasma membrane              Proteasome 
-##               0.7053008               0.4133848
+##               0.7176337               0.4177169
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     170                      84 
+##                      83                     174                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     301                     479                      94 
+##                     300                     460                     106 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     523 
+##                      27                     121                     524 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     341                      38 
+##                     229                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     326                     154                    2058
+##                     328                     157                    2058
 \end{verbatim}

### 2016-11-04 bioc-workflow.Rnw
unsupervised learning section

@@
 learning with a single source of information \cite{Breckels:2016}.
+
+\subsection*{TL optimisation}
 
@@
 
-\paragraph{TL classification} Looking at the bubble plot displaying
-the distribution of best weights over the 50 runs we find that for
-many of the subcellular niches a weight of 1 is most popular (i.e. use
-only primary hyperLOPIT data in classification), this is unsuprising
-as we already know the dataset is well resolved for these classes. We
-see that the most popular weights for the proteasome and lysosome tend
-to be towards 0, indicating that these niches are well-resolved in the
-Gene Ontology. This tells us that we would benefit from including
-auxiliary GO information in our classifier for these subcellular
-compartments. The plasma membrane weights are relatively equally
-spread between using hyperLOPIT and GO data. Using the
-\texttt{getParams} function we can return the best weights and then
-use this as input for the classification.
+\subsection*{TL classification}
+
+Looking at the bubble plot displaying the distribution of best weights
+over the 50 runs we find that for many of the subcellular niches a
+weight of 1 is most popular (i.e. use only primary hyperLOPIT data in
+classification), this is unsuprising as we already know the dataset is
+well resolved for these classes. We see that the most popular weights
+for the proteasome and lysosome tend to be towards 0, indicating that
+these niches are well-resolved in the Gene Ontology. This tells us
+that we would benefit from including auxiliary GO information in our
+classifier for these subcellular compartments. The plasma membrane
+weights are relatively equally spread between using hyperLOPIT and GO
+data. Using the \texttt{getParams} function we can return the best
+weights and then use this as input for the classification.
 
@@
 include these 4 subcellular niches in the next classification step we
-must include them in the parameters so we infer a weight of 1 for each
-of these niches as we know they are well resolved in hyperLOPIT. We
+must include them in the parameters. We define a weight of 1 for each
+of these niches, as we know they are well resolved in hyperLOPIT. We
 then re-order the weights according to \texttt{getMarkerClasses} and
@@
 learning methods. In unsupervised learning, the training data consists
-of a set of input vectors e.g. protein profiles, for which we do not
-have information about the class label e.g. localisation. The main
-goal in unsupervised learning is to uncover groups of similar examples
-within the data, this is termed clustering. Ordinance methods such as
-principal components analysis (PCA) also fall into the category of
-unsupervised learning methods, where the data can be projected from a
-high-dimensional spcae down to two or three dimensions for the purpose
-of visualisation.
+of a set of input vectors e.g. protein profiles, ignoring the
+information about the class label e.g. localisation, other than for
+annotation purposes. The main goal in unsupervised learning is to
+uncover groups of similar features within the data, termed
+clustering. Ordination methods such as principal components analysis
+(PCA) also fall into the category of unsupervised learning methods,
+where the data can be projected from a high-dimensional spcae down to
+two or three dimensions for the purpose of visualisation.
 
 As described and demonstrated already above, PCA is a valuable and
-powerful method for data visualisation and quality control. We do not
-do any unsupervised clustering in the frame of the above spatial
-proteomics workflow as although clustering can be useful for grouping
-labelled data into categories e.g. see the \texttt{mrkHClust}
-function, we do not find it adaquete for spatial proteomics data
-analysis. We find supervised learning more suited to the task of
+powerful method for data visualisation and quality control. Another
+application uses hierarchical clustering to summarise the relation
+between marker proteins using he \texttt{mrkHClust} function, where
+the euclidean distance bewteen average class-specific profiles is used
+to produce a dendrogramme describing a simple relationship between the
+sub-cellular classes (Figure~\ref{fig:mrkHclust}).
+
+\begin{figure}[!ht]
+  \centering
+<<hclust, out.width=".6\\textwidth", fig.asp = 1.5, out.extra='angle=-90'>>=
+par(mar = c(15, 4, 1, 0))
+mrkHClust(hl)
+@  
+  \caption{Hierarchical clustering of the average marker profiles summarising the relation between organelles profiles.}
+  \label{fig:mrkHclust}
+\end{figure}
+
+We generally find supervised learning more suited to the task of
 protein localisation prediction in which we use high-quality curated

### 2016-11-04 bioc-workflow.tex
unsupervised learning section

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 12:03:40 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:47:41 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 12:03:40 2016 
-## Subset [6725,20][5032,20] Fri Nov  4 12:03:41 2016 
-## Removed features with more than 0 NAs: Fri Nov  4 12:03:41 2016 
-## Dropped featureData's levels Fri Nov  4 12:03:41 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:47:41 2016 
+## Subset [6725,20][5032,20] Fri Nov  4 13:47:41 2016 
+## Removed features with more than 0 NAs: Fri Nov  4 13:47:41 2016 
+## Dropped featureData's levels Fri Nov  4 13:47:41 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Nov  4 12:03:43 2016
+% Fri Nov  4 13:47:44 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Nov  4 12:03:43 2016 
-## Added markers from  'pdres' marker vector. Fri Nov  4 12:03:43 2016 
+## Added markers from  'mrk' marker vector. Fri Nov  4 13:47:43 2016 
+## Added markers from  'pdres' marker vector. Fri Nov  4 13:47:44 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4369956               0.3026572               0.3845674 
+##               0.4336259               0.3016425               0.3822712 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6735594               0.6449643               0.4375399 
+##               0.6765106               0.6083550               0.4291844 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4123260               0.6045583               0.9494543 
+##               0.4223252               0.5894097               0.9493407 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7957342               0.7090410               0.3143893 
+##               0.7950836               0.7082548               0.3133443 
 ##         Plasma membrane              Proteasome 
-##               0.7176337               0.4177169
+##               0.7083989               0.4215825
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      83                     174                      84 
+##                      86                     168                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     300                     460                     106 
+##                     302                     477                      98 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     121                     524 
+##                      27                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     343                      38 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     328                     157                    2058
+##                     322                     153                    2057
 \end{verbatim}
@@
 learning methods. In unsupervised learning, the training data consists
-of a set of input vectors e.g. protein profiles, for which we do not
-have information about the class label e.g. localisation. The main
-goal in unsupervised learning is to uncover groups of similar examples
-within the data, this is termed clustering. Ordinance methods such as
-principal components analysis (PCA) also fall into the category of
-unsupervised learning methods, where the data can be projected from a
-high-dimensional spcae down to two or three dimensions for the purpose
-of visualisation.
+of a set of input vectors e.g. protein profiles, ignoring the
+information about the class label e.g. localisation, other than for
+annotation purposes. The main goal in unsupervised learning is to
+uncover groups of similar features within the data, termed
+clustering. Ordination methods such as principal components analysis
+(PCA) also fall into the category of unsupervised learning methods,
+where the data can be projected from a high-dimensional spcae down to
+two or three dimensions for the purpose of visualisation.
 
 As described and demonstrated already above, PCA is a valuable and
-powerful method for data visualisation and quality control. We do not
-do any unsupervised clustering in the frame of the above spatial
-proteomics workflow as although clustering can be useful for grouping
-labelled data into categories e.g. see the \texttt{mrkHClust}
-function, we do not find it adaquete for spatial proteomics data
-analysis. We find supervised learning more suited to the task of
-protein localisation prediction in which we use high-quality curated
-marker proteins to build a classifier, instead of using an entirely
+powerful method for data visualisation and quality control. Another
+application uses hierarchical clustering to summarise the relation
+between marker proteins using he \texttt{mrkHClust} function, where
+the euclidean distance bewteen average class-specific profiles is used
+to produce a dendrogramme describing a simple relationship between the
+sub-cellular classes.
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{15}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{0}\hlstd{))}
+\hlkwd{mrkHClust}\hlstd{(hl)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.6\textwidth,angle=-90]{figure/hclust-1}
+
+\end{knitrout}
+  \caption{Hierarchical clustering of the average marker profiles summarising the relation between organelles profiles.}
+  \label{fig:mrkHclust}
+\end{figure}
+
+We find supervised learning more suited to the task of protein
+localisation prediction in which we use high-quality curated marker
+proteins to build a classifier, instead of using an entirely
 unsupervised approach to look for clusters and then look for

### 2016-11-04 bioc-workflow.Rnw
a bit more to unsupervised learning section

@@
 to produce a dendrogramme describing a simple relationship between the
-sub-cellular classes (Figure~\ref{fig:mrkHclust}).
+sub-cellular classes (Figure~\ref{fig:mrkHclust}). The
+\texttt{mrkHClust} uses the same defaults as all other function, using
+the \texttt{markers} feature variable to define marker proteins. In
+the code chunk, we adapt the figure margins to fully display the class
+names.
 
@@
   \centering
-<<hclust, out.width=".6\\textwidth", fig.asp = 1.5, out.extra='angle=-90'>>=
+<<hclust, out.width=".5\\textwidth", fig.asp = 1.5, out.extra='angle=-90'>>=
 par(mar = c(15, 4, 1, 0))

### 2016-11-04 bioc-workflow.tex
a bit more to unsupervised learning section

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:47:41 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:53:42 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:47:41 2016 
-## Subset [6725,20][5032,20] Fri Nov  4 13:47:41 2016 
-## Removed features with more than 0 NAs: Fri Nov  4 13:47:41 2016 
-## Dropped featureData's levels Fri Nov  4 13:47:41 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:53:42 2016 
+## Subset [6725,20][5032,20] Fri Nov  4 13:53:43 2016 
+## Removed features with more than 0 NAs: Fri Nov  4 13:53:43 2016 
+## Dropped featureData's levels Fri Nov  4 13:53:43 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Nov  4 13:47:44 2016
+% Fri Nov  4 13:53:45 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Nov  4 13:47:43 2016 
-## Added markers from  'pdres' marker vector. Fri Nov  4 13:47:44 2016 
+## Added markers from  'mrk' marker vector. Fri Nov  4 13:53:45 2016 
+## Added markers from  'pdres' marker vector. Fri Nov  4 13:53:45 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4336259               0.3016425               0.3822712 
+##               0.4344841               0.3054605               0.3801168 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6765106               0.6083550               0.4291844 
+##               0.6958931               0.5977237               0.4161283 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4223252               0.5894097               0.9493407 
+##               0.4300573               0.5962921               0.9500664 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950836               0.7082548               0.3133443 
+##               0.7946555               0.7105355               0.3145718 
 ##         Plasma membrane              Proteasome 
-##               0.7083989               0.4215825
+##               0.7112584               0.4155859
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     168                      84 
+##                      84                     172                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     302                     477                      98 
+##                     301                     475                      99 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     523 
+##                      29                     125                     521 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     230                     342                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     153                    2057
+##                     319                     155                    2057
 \end{verbatim}
@@
 to produce a dendrogramme describing a simple relationship between the
-sub-cellular classes.
+sub-cellular classes (Figure~\ref{fig:mrkHclust}). The
+\texttt{mrkHClust} uses the same defaults as all other function, using
+the \texttt{markers} feature variable to define marker proteins. In
+the code chunk, we adapt the figure margins to fully display the class
+names.
 
@@
 \end{kframe}
-\includegraphics[width=.6\textwidth,angle=-90]{figure/hclust-1}
+\includegraphics[width=.5\textwidth,angle=-90]{figure/hclust-1}
 
@@
 
-We find supervised learning more suited to the task of protein
-localisation prediction in which we use high-quality curated marker
-proteins to build a classifier, instead of using an entirely
+We generally find supervised learning more suited to the task of
+protein localisation prediction in which we use high-quality curated
+marker proteins to build a classifier, instead of using an entirely
 unsupervised approach to look for clusters and then look for

### 2016-11-04 bioc-workflow.Rnw
finished ammending manuscript

@@
 
+Feature metadata typically contains general annotation about the
+proteins (accession numbers, description, \ldots), information related
+to the identification search (confidence scores, number of peptides,
+\ldots) as well as annotation about know sub-cellular location (see in
+particular the \textit{Markers} section) and results from data
+analysis. The sample metadata would, for example, record what
+labelling ions were used for the respective fraction (when labelled
+quantitation is used), replicate number, fraction number along the
+gradient and pooling information.
+
 Another slot of interest is \texttt{processingData}, that logs the
@@
 the imputation method is then determined by a \texttt{methods}
-parameter. In our particular case, missing values are indicative of
-protein groups that were not acquired in both replicates (Figures
+parameter. To impute missing values using nearest neighbour
+imputation, one would
+
+<<impute, eval=FALSE>>=
+hl <- impute(hl, method = "knn")
+@
+
+In our particular case, missing values are indicative of protein
+groups that were not acquired in both replicates (Figures
 \ref{fig:namap}).
@@
   resouces} or any conversion softwares available online.
+
+It is also possible to use one's own marker list with the
+\texttt{addMarkers} function. The user needs to create a named vector
+of marker localisation, where the names match some (not necessarily
+all) of the \texttt{MSnSet}'s feature names.
 
@@
 be as simple as the parameter set that gives rise to the highest macro
-F1 score and one must be careful to avoid overfitting and to choose
-parameters wisely.
+F1 score. One must be careful to avoid overfitting, and choose
+parameters that frequently provide high classification
+accuracy. Below, we see that only a sigma of 0.1 produces macro F1
+scores above 0.6, but that a cost of 16 is much more frequently chosen
+than lower values.
 
@@
 cost/sigma parameter pairs, and \texttt{levelPlot} shows the averaged
-macro F1 scores, for the full range of parameter values.
+macro F1 scores, for the full range of parameter values. These figures
+also indicate that values of 0.1 and 16 for sigma and cost
+consistently deliver better classification scores.
 
@@
 set is automatically but users are encouraged to critically assess
-whether this is the most wise choice.
+whether this is the most wise choice (which it is, as demonstrated
+above).
 
@@
 containing the prediction results. The results can also be visualied
-using \texttt{plot2D} function (Figure \ref{fig:plotres}).
+using \texttt{plot2D} function (Figure \ref{fig:plotres}) and
+extracted by retrieving that specific column from the feature metadata
+using, for example, \texttt{fData(hl)\$svm.pred}.
 
@@
 
-A \texttt{MSnSet} can be exported from R using the
+An \texttt{MSnSet} can be exported from R using the
 \texttt{write.exprs} function.  This function writes the expression
-values to a tab (the defualt, or other type e.g.e.g. comma, as
-specified in \texttt{write.table}) separated file. There argument
-\texttt{fDataCols} can be used to specify which \texttt{featureData}
-columns (as column names, column number or \texttt{logical}) to append
-to the right of the expression matrix.
+values to a text-based spreadsheet. The \texttt{fcol} argument can be
+used to specify which \texttt{featureData} columns (as column names,
+column number or \texttt{logical}) to append to the right of the
+expression matrix.
 
@@
 \texttt{featureData} to the file, as well as the expression data, we
-specify \texttt{fDataCols = 1:ncol(hl)} i.e. write everything in the
-\texttt{featureData} to the file \texttt{"hl.csv"}.
+specify \texttt{fvarLabels(hl)}, that returns all feature variable
+names, and write the resulting data to the file \texttt{"hl.csv"}.
 
 <<writeData, eval = FALSE>>=
-write.exprs(hl, file = "hl.csv", sep = ",", 
-            fDataCols = 1:ncol(hl))
-@
-
-Individual R objects, such a parameters from the machine learning
-optimisations, can be save using the standard \texttt{save} function
-in R.  For example, to save and then re-load the parameters from the
-SVM optimisation,
+write.exprs(hl, file = "hl.csv", sep = ",", fcol = fvarLabels(hl))
+@
+
+Exporting to a spreadsheet however looses a lot of important
+information, such as the processing data, and the sample metadata in
+the \textit{phenoData} slot. Other objects, such a parameters from the
+machine learning optimisation, cannot be represented as tabular
+data. To directly serialise R objects to disk, on can use the standard
+\texttt{save} function, and later reload the object using
+\texttt{save}. For example, to save and then re-load the parameters
+from the SVM optimisation,
 
@@
 and versions used to generate this document. This enables us to record
-the exact state of our session that lead to these exact
-results. Conversely, if the script stops working of if it returns
-different results, we are in a position to re-generate the original
-results using the adequate software versions and retrace changes in
-the software that lead to failure and/or different results.
+the exact state of our session that lead to these results. Conversely,
+if the script stops working or if it returns different results, we are
+in a position to re-generate the original results using the adequate
+software versions and retrace changes in the software that lead to
+failure and/or different results.
 
@@
 The source of this document, including the code necessary to reproduce
-the analyses and figures is available in the public manuscript
-repository \cite{ghrepo}.
+the analyses and figures is available in a public manuscript
+repository on GitHub~\cite{ghrepo}.
 
@@
 
-All authors developed the software presented in this workflow and
+All authors developed the software presented in this workflow, and
 wrote and approved the manuscript.
@@
 
-The authors would like to thank the Cambridge Computational Biology
-Institute and the Centre for Mathematical Sciences for hosting them
-when writing this article. The authors would like to thank Dr Claire
-M. Mulvey for helpful comments on early versions of this
-manuscript. The authors would also like to thank Dr. Stuart Rankin
-from the High Performance Computing Service for his support. Part of
-this work was performed using the Darwin Supercomputer of the
-University of Cambridge High Performance Computing Service, provided
-by Dell Inc. using Strategic Research Infrastructure Funding from the
-Higher Education Funding Council for England and funding from the
-Science and Technology Facilities Council.
+The authors would like to thank Dr Claire M. Mulvey for helpful
+comments on early versions of this manuscript. The authors would also
+like to thank Dr. Stuart Rankin from the High Performance Computing
+Service for his support. Part of this work was performed using the
+Darwin Supercomputer of the University of Cambridge High Performance
+Computing Service, provided by Dell Inc. using Strategic Research
+Infrastructure Funding from the Higher Education Funding Council for
+England and funding from the Science and Technology Facilities
+Council.

### 2016-11-04 bioc-workflow.tex
finished ammending manuscript

@@
 \end{figure}
+
+Feature metadata typically contains general annotation about the
+proteins (accession numbers, description, \ldots), information related
+to the identification search (confidence scores, number of peptides,
+\ldots) as well as annotation about know sub-cellular location (see in
+particular the \textit{Markers} section) and results from data
+analysis. The sample metadata would, for example, record what
+labelling ions were used for the respective fraction (when labelled
+quantitation is used), replicate number, fraction number along the
+gradient and pooling information.
 
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:53:42 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 14:46:45 2016 
 ##  MSnbase version: 1.21.7
@@
 the imputation method is then determined by a \texttt{methods}
-parameter. In our particular case, missing values are indicative of
-protein groups that were not acquired in both replicates (Figures
+parameter. To impute missing values using nearest neighbour
+imputation, one would
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+In our particular case, missing values are indicative of protein
+groups that were not acquired in both replicates (Figures
 \ref{fig:namap}).
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 13:53:42 2016 
-## Subset [6725,20][5032,20] Fri Nov  4 13:53:43 2016 
-## Removed features with more than 0 NAs: Fri Nov  4 13:53:43 2016 
-## Dropped featureData's levels Fri Nov  4 13:53:43 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 14:46:45 2016 
+## Subset [6725,20][5032,20] Fri Nov  4 14:46:45 2016 
+## Removed features with more than 0 NAs: Fri Nov  4 14:46:45 2016 
+## Dropped featureData's levels Fri Nov  4 14:46:45 2016 
 ##  MSnbase version: 1.21.7
@@
 
+It is also possible to use one's own marker list with the
+\texttt{addMarkers} function. The user needs to create a named vector
+of marker localisation, where the names match some (not necessarily
+all) of the \texttt{MSnSet}'s feature names.
+
 We now visualise these annotations along the PCA plot using the
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Nov  4 13:53:45 2016
+% Fri Nov  4 14:46:48 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Nov  4 13:53:45 2016 
-## Added markers from  'pdres' marker vector. Fri Nov  4 13:53:45 2016 
+## Added markers from  'mrk' marker vector. Fri Nov  4 14:46:47 2016 
+## Added markers from  'pdres' marker vector. Fri Nov  4 14:46:48 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4344841               0.3054605               0.3801168 
+##               0.4344078               0.3022144               0.3779558 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6958931               0.5977237               0.4161283 
+##               0.6798372               0.5931386               0.4103045 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4300573               0.5962921               0.9500664 
+##               0.4226277               0.6020915               0.9489050 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946555               0.7105355               0.3145718 
+##               0.7946252               0.7055619               0.3165817 
 ##         Plasma membrane              Proteasome 
-##               0.7112584               0.4155859
+##               0.7079566               0.4183278
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     172                      84 
+##                      85                     172                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     301                     475                      99 
+##                     295                     473                     101 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      29                     125                     521 
+##                      27                     126                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     342                      39 
+##                     230                     343                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     319                     155                    2057
+##                     320                     159                    2056
 \end{verbatim}
@@
 \texttt{featureData} to the file, as well as the expression data, we
-specify \texttt{fDataCols = 1:ncol(hl)} i.e. write everything in the
-\texttt{featureData} to the file \texttt{"hl.csv"}.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{file} \hlstd{=} \hlstr{"hl.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{,}
-            \hlkwc{fDataCols} \hlstd{=} \hlnum{1}\hlopt{:}\hlkwd{ncol}\hlstd{(hl))}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-Individual R objects, such a parameters from the machine learning
-optimisations, can be save using the standard \texttt{save} function
-in R.  For example, to save and then re-load the parameters from the
-SVM optimisation,
+specify \texttt{fvarLabels(hl)}, that returns all feature variable
+names, and write the resulting data to the file \texttt{"hl.csv"}.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{file} \hlstd{=} \hlstr{"hl.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{,} \hlkwc{fcol} \hlstd{=} \hlkwd{fvarLabels}\hlstd{(hl))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+Exporting to a spreadsheet however looses a lot of important
+information, such as the processing data, and the sample metadata in
+the \textit{phenoData} slot. Other objects, such a parameters from the
+machine learning optimisation, cannot be represented as tabular
+data. To directly serialise R objects to disk, on can use the standard
+\texttt{save} function, and later reload the object using
+\texttt{save}. For example, to save and then re-load the parameters
+from the SVM optimisation,
 
@@
 ## other attached packages:
-##  [1] pRolocdata_1.12.0    pRoloc_1.15.3        MLInterfaces_1.54.0 
+##  [1] pRolocdata_1.12.0    pRoloc_1.14.3        MLInterfaces_1.54.0 
 ##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.4        

### 2016-11-08 bioc-workflow.Rmd
Spelling/grammar typos

--- 2016-09-29 bioc-workflow.Rmd
@@
 
-Installation of Bioconductor package is documented in details on the
+Installation of Bioconductor packages are documented in detail on the
 [Bioconductor installation help page](http://bioconductor.org/install/#install-bioconductor-packages). This
@@
 replicates 1 and 2 was treated as a 20-plex dataset for the analysis.
-As discussed in the manuscript [@hyper] as it has been shown that
+As discussed in the manuscript [@hyper] it has been shown that
 combining replicates across from different gradients can increase
@@
 variance stabilisation or quantile normalisation, to cite a few, can
-be applied to accomodation different needs.
+be applied to accomodate different needs.
 

### 2016-11-08 abstract.tex

Spatial proteomics is the systematic study of protein sub-cellular
localisation. In this workflow, we describe the analysis of a typical
quantitative mass spectrometry-based spatial proteomics experiment
using the \Biocpkg{MSnbase} and \Biocpkg{pRoloc} Bioconductor package
suite. To walk the user through the computational pipeline, we use a
recently published experiment predicting protein sub-cellular
localisation in pluripotent embryonic mouse stem cells. We describe
the software infrastructure at hand, importing and processing data,
quality control, sub-cellular marker definition and visualisation. We
then demonstrate the application and interpretation of statistical
learning methods, including novelty detection using semi-supervised
learning, classification, clustering and transfer learning and
conclude the pipeline with data export. The workflow is aimed at
beginners who are familiar with proteomics in general and spatial
proteomics in particular.

### 2016-11-08 bioc-workflow.Rnw
More typos fixed

--- 2016-11-04 bioc-workflow.Rnw
@@
 
-Installation of Bioconductor package is documented in details on the
+Installation of Bioconductor package is documented in detail on the
 \href{http://bioconductor.org/install/#install-bioconductor-packages}{Bioconductor
@@
 parameter. In our particular case, missing values are indicative of
-protein groups that were not acquired in both replicates (Figures
+protein groups that were not acquired in both replicates (Figure
 \ref{fig:namap}).
@@
 experiments. This is done with the \texttt{filterNA} function that
-removes features that contain more than a certain proportion (default
+removes features (proteins) that contain more than a certain proportion (default
 is 0) missing values.
@@
 
-When more than 2 data are to be combined and too many proteins were
+When more than 2 datasets are to be combined and too many proteins were
 not consistently assayed, leading to too many proteins being filtered
@@
 principal components analysis (PCA). PCA is one of many dimensionality
-reduction methods, that allow one to effectively summarise
+reduction methods, that allows one to effectively summarise
 multi-dimensional data in to 2 or 3 dimensions to enable
@@
 hard quantitative measure of separation, it is extremely useful
-summarising complex experimental information in one figure, to get an
+summarising complex experimental information in one figure, to get a
 simplified overview of the data.
@@
 variable to annotate the features (proteins) with colours. We will see
-later how to use this to annotate the PCA plot with prior information
+later how to use this argument to annotate the PCA plot with prior information
 about sub-cellular localisation.
@@
 motivation of this number). Markers should be chosen to confidently
-represent to distribution of genuine residents of a sub-cellular
-niche. We generally recommend such a conservative approach in defining
+represent the distribution of genuine residents of a sub-cellular
+niche. We generally recommend a conservative approach in defining
 markers to avoid false assignments when assigning sub-cellular
@@
 defines markers, risks to erroneously assign proteins to a single
-location, when, in reality, they reside in more that a single location
+location, when, in reality, they reside in multiple locations
 (including the assumed unique location). One can not expect to
@@
 
-If the naming between the marker sets and the \texttt{MSnSet} dataset
-are different, one will have to convert and match the proteins
-according to the appropriate identifier. Sometimes, we find the
-equivalent entry name, Uniprot ID or accession number is stored with
-the data, which makes conversion between identifers relatively
-straightforward. If this is not the case however, conversion can be
-performed using \Biocpkg{biomaRt}, the Bioconductor
+If the protein naming between the marker sets and the \texttt{MSnSet}
+dataset are different e.g. the markers are labelled by Uniprot
+accession numbers and the dataset entries are labelled by Uniprot entry
+names, one will have to convert and match the proteins according to
+the appropriate identifier. Sometimes, we find the equivalent entry
+name, Uniprot ID or accession number is stored with the data, which
+makes conversion between identifers relatively straightforward. If
+this is not the case however, conversion can be performed using
+\Biocpkg{biomaRt}, the Bioconductor
 \href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation

### 2016-11-08 bioc-workflow.Rnw

--- 2016-11-04 bioc-workflow.Rnw
@@
 In our particular case, missing values are indicative of protein
-groups that were not acquired in both replicates (Figures
+groups that were not acquired in both replicates (Figure
 \ref{fig:namap}).
@@
 
-If the naming between the marker sets and the \texttt{MSnSet} dataset
-are different, one will have to convert and match the proteins
-according to the appropriate identifier. Sometimes, we find the
-equivalent entry name, Uniprot ID or accession number is stored with
-the data, which makes conversion between identifers relatively
-straightforward. If this is not the case however, conversion can be
-performed using \Biocpkg{biomaRt}, the Bioconductor
+If the protein naming between the marker sets and the \texttt{MSnSet}
+dataset are different e.g. the markers are labelled by Uniprot
+accession numbers and the dataset entries are labelled by Uniprot
+entry names, one will have to convert and match the proteins according
+to the appropriate identifier. Sometimes, we find the equivalent entry
+name, Uniprot ID or accession number is stored with the data, which
+makes conversion between identifers relatively straightforward. If
+this is not the case however, conversion can be performed using
+\Biocpkg{biomaRt}, the Bioconductor
 \href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation

### 2016-11-08 bioc-workflow.tex
make all

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 14:46:45 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 14:13:46 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Nov  4 14:46:45 2016 
-## Subset [6725,20][5032,20] Fri Nov  4 14:46:45 2016 
-## Removed features with more than 0 NAs: Fri Nov  4 14:46:45 2016 
-## Dropped featureData's levels Fri Nov  4 14:46:45 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 14:13:46 2016 
+## Subset [6725,20][5032,20] Tue Nov  8 14:13:47 2016 
+## Removed features with more than 0 NAs: Tue Nov  8 14:13:47 2016 
+## Dropped featureData's levels Tue Nov  8 14:13:47 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Nov  4 14:46:48 2016
+% Tue Nov  8 14:13:50 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Nov  4 14:46:47 2016 
-## Added markers from  'pdres' marker vector. Fri Nov  4 14:46:48 2016 
+## Added markers from  'mrk' marker vector. Tue Nov  8 14:13:49 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  8 14:13:50 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4344078               0.3022144               0.3779558 
+##               0.4359409               0.3031417               0.3896552 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6798372               0.5931386               0.4103045 
+##               0.6869014               0.6046754               0.4336399 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4226277               0.6020915               0.9489050 
+##               0.4165647               0.5910830               0.9498244 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946252               0.7055619               0.3165817 
+##               0.7955788               0.7079722               0.3121358 
 ##         Plasma membrane              Proteasome 
-##               0.7079566               0.4183278
+##               0.7187306               0.4168615
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      84 
+##                      86                     171                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     473                     101 
+##                     294                     477                      99 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     126                     522 
+##                      27                     123                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      39 
+##                     230                     342                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     320                     159                    2056
+##                     316                     160                    2058
 \end{verbatim}

### 2016-11-08 bioc-workflow.Rnw
update author list and ack
Lisa's corrections to manuscript

@@
 
-It is also possible to use one's own marker list with the
+It is also possible for users to use their own marker list with the
 \texttt{addMarkers} function. The user needs to create a named vector
-of marker localisation, where the names match some (not necessarily
-all) of the \texttt{MSnSet}'s feature names.
+of marker localisation, or a create a csv file with two columns (one
+for the protein names, one for the corresponding sub-cellular marker
+annotation) and pass the vector or filename respectively to the
+function. As previously mentioned, the protein names of these markers
+must match some (but not necessarily all) of the \texttt{MSnSet}'s
+feature names. See `?addMarkers` for more details.
 
@@
 have omitted the \texttt{fcol} argument and use of the default
-\texttt{"markers"} feature variable to annotated the plot. We choose
+\texttt{"markers"} feature variable to annotate the plot. We choose
 to display PCs 1 and 7 to illustrate that while upper principal
@@
 and peroxisome (dark blue) clusters can be differenciated, despite the
-apparent overlap in the two first PCs.
+apparent overlap in the visualisation of the two first PCs
+(Figure \ref{fig:plotmarkers}).
 
@@
 
-The colours have been defined so as to enable to differenciate up to
-30 classes. If more are provided, different character symbols
-(circles, squares, ... and empty and solid symbols) are used. The
-colours and the default plotting characters (solid dots for the
-markers and empty circles for the features of unknown localisation)
-can of course be changed, as described in the \texttt{setStockcol} manual
-page.
+The defualt colours for plotting have been defined so as to enable to
+differenciate up to 30 classes. If more are provided, different
+character symbols (circles, squares, ... and empty and solid symbols)
+are used. The colours and the default plotting characters (solid dots
+for the markers and empty circles for the features of unknown
+localisation) can of course be changed, as described in the
+\texttt{setStockcol} manual page.
 
 As demonstrated in \cite{hyper} and illustrated in the PCA plot above,
-the Golgi apparatus proteins (brown) display a dynamic pattern, noting
-sets of Golgi marker proteins that are distributed amongst other
-subcellular structures, an observation supported by microscopy. As
-such, we are going to reset the annotation of Golgi markers to unknown
-using the \texttt{fDataTounknown} function. It is often used to
-replace empty strings (\texttt{""}) or missing values in the markers definition
-to a common definition of \textit{unknown} localisation.
+the Golgi apparatus proteins (dark brown) display a dynamic pattern,
+noting sets of Golgi marker proteins that are distributed amongst
+other subcellular structures, an observation supported by microscopy.
+As such, we are going to reset the annotation of Golgi markers to
+unknown using the \texttt{fDataTounknown} function. It is often used
+to replace empty strings (\texttt{""}) or missing values in the
+markers definition to a common definition of \textit{unknown}
+localisation.
 
@@
 highlight the differences in profiles between these two sets of
-markers along the 6th and 7th channels, as represented above along the
-7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
+markers in the channels labelled with tag 129C, as represented above
+along the 7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
 
@@
 plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
-         pcol = "purple", fractions = "Fraction.No")
+         pcol = "purple", fractions = "Tag")
 title(main = "Marker occupancy profiles along the gradient")
@@
 fraction collection, the quantitative channels do not correspond to
-identical selected fractions along the gradient. As can be seen in the
-table below (taken from the \texttt{hl}'s \texttt{pData}, focusing on
-channels 7 to 10), different sets of gradient fractions are pooled to
-obtain enough material and optimise acurate quantitation.
+identical selected fractions along the gradient. For example, in the
+table below (taken from \texttt{hl}'s \texttt{pData}) TMT channels
+127C (among others) in both replicates originate from different sets
+of gradient fractions (gradient fractions 7 - 9 and 8 - 9 for each
+replicate, respectively). Different sets of gradient fractions are
+often pooled to obtain enough material and optimise acurate
+quantitation.
 
 <<pdtab, echo=FALSE, eval = TRUE, results = 'asis'>>=
-i <- c(2, 3, 4, 13, 14)
+i <- c(3, 13)
 xtable(pData(hl)[i, ], 
@@
 The more relevant comparison unit is not a single channel, but rather
-the complete protein occupancy profiles, which are best visualised as
+the complete protein occupancy profiles, which are best visualised 
 experiment-wide on a PCA plot. As such, we prefer to focus on the
@@
 
-As diplayed in the screenshot above, the \textit{main} application is
-designed for exploratory data analysis and is divied into 3 tabs: (1)
-PCA, (2) Profiles and (3) Table selection. The default view upon
-loading is the PCA tab, which features a clickable interface and
-zoomable PCA plot with an interactive data table for displaying the
-quantitation information. Particular proteins of interest can be
-highlighted using the text search box. There is also an alternate
-profiles tab for visualisation of the protein profiles, which can be
-used to examine the patterns of proteins of interest. The Table
-selection tab provides an interface to control data table column
-selection. A short
+As diplayed in the screenshot in figure \ref{fig:app}, the
+\textit{main} application is designed for exploratory data analysis
+and is divied into 3 tabs: (1) PCA, (2) Profiles and (3) Table
+selection. The default view upon loading is the PCA tab, which
+features a clickable interface and zoomable PCA plot with an
+interactive data table for displaying the quantitation information.
+Particular proteins of interest can be highlighted using the text
+search box. There is also an alternate profiles tab for visualisation
+of the protein profiles, which can be used to examine the patterns of
+proteins of interest. The Table selection tab provides an interface to
+control data table column selection. A short
 animation\footnote{\url{https://github.com/lmsimp/bioc-pRoloc-hyperLOPIT-workflow/blob/master/Figures/pRolocVis\_pca.gif}}
@@
 Note: We do not evaluate this code chunk in this document as the
-algorithm is computational intensive and best parallelised over
+algorithm is computationally intensive and best parallelised over
 multiple workers. This phenoDisco analysis took ~24 hours to complete
@@
 
-The argument \texttt{times} indicated the number of times we run
+The argument \texttt{times} indicates the number of times we run
 unsupervied Gaussian Mixture Modelling before defining a new phenotype
@@
 above code chunk we increase the value to \texttt{times = 200} as we
-have found for larger datasets (e.g. 5000+ proteins) a higher times is
-requried for convergence. \texttt{GS} defines the minimum number of
-proteins allowed per new data cluster and thus heavily influences what
-type of new clusters are extracted. For example, if a user is
-interesed in the detection of small complexes they may wish to use a
-small \texttt{GS = 10}, or \texttt{= 20} etc.  If they wish to detect
-larger, more abundant sub-cellular niches a much higher \texttt{GS}
-would be preferable. Specifying a small \texttt{GS} can be more time
-consuming than using a larger \texttt{GS}, and there is a trade off
-between finding interesting small complexes and those that may not be
-of interest as we find there is a tendancy to find more noise when
-using a small \texttt{GS} compared to using a higher one.
+have found for larger datasets (e.g. 5000+ proteins) a higher
+\texttt{times} is requried for convergence. \texttt{GS} defines the
+minimum number of proteins allowed per new data cluster and thus
+heavily influences what type of new clusters are extracted. For
+example, if a user is interesed in the detection of small complexes
+they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
+If they wish to detect larger, more abundant sub-cellular niches a
+much higher \texttt{GS} would be preferable. Specifying a small
+\texttt{GS} can be more time consuming than using a larger
+\texttt{GS}, and there is a trade off between finding interesting
+small complexes and those that may not be of interest as we find there
+is a tendancy to find more noise when using a small \texttt{GS}
+compared to using a higher one.
 
@@
 extracted and examined.  In the code chunk below we write the results
-to a .csv file. We use the argument \texttt{fDataCols} to specify
-which columns of the \texttt{featureData} to write.
+to a .csv file using the \texttt{write.exprs} function. We use the
+argument \texttt{fDataCols} to specify which columns of the
+\texttt{featureData} to write.
 
@@
 
-We can also examine the each phenotype intercatively and visualise
+We can also examine each phenotype interactively and visualise
 their protein profiles by using the \texttt{pRolocVis} function in the
@@
 a protein of unknown localisation, based on its observed protein
-profile. To date, this method has been extensively used in spatial
+profiles. To date, this method has been extensively used in spatial
 quantitative proteomics to assign thousands of proteins to distinct
@@
 unknown location. One of the most common ways to optimise the
-parameters of a classifier is to partition the labelled data in to
+parameters of a classifier is to partition the labelled data into
 training and testing subsets. In this framework parameters are tested
@@
 As mentioned previously, we rely on the default feature variable
-\texttt{"markers"} to define the class labels and hence can ommit
-it. To use another feature variables, one need to explicitly specify
-its name using the \texttt{fcol} argument (for example \texttt{fcol =
-  "markers2"}).
+\texttt{"markers"} to define the class labels and hence do not need to
+specify it in the above code chunk.  To use another feature variables,
+one need to explicitly specify its name using the \texttt{fcol}
+argument (for example \texttt{fcol = "markers2"}).
 
@@
 of parameters. Currently, \texttt{getParams} retrieves the first best
-set is automatically but users are encouraged to critically assess
+set automatically, but users are encouraged to critically assess
 whether this is the most wise choice (which it is, as demonstrated
@@
 
-We can use the function \texttt{svmClassification} to return a classification
-result for all unlabelled instances in the dataset corresponding to
-their most likely sub-cellular location. The algorithm parameters are
-passed to the function, along with the class weights. As above, \texttt{fcol}
-can be ignored as we use the labels defined in the default \texttt{"markers"}
-feature variable.
+We can use the function \texttt{svmClassification} to return a
+classification result for all unlabelled instances in the dataset
+corresponding to their most likely sub-cellular location. The
+algorithm parameters are passed to the function, along with the class
+weights. As above, the \texttt{fcol} argument does not need to be
+specified as we use the labels defined in the default
+\texttt{"markers"} feature variable.
 
@@
 The adjustment of the point size intuitively confers important
-information that is more difficult to defined formally, but that we
-will address in the next section. The classifier (SVM in our case, but
+information that is more difficult to define formally (we
+will address in the next section). The classifier (SVM in our case, but
 this is also valid of other classifiers) defines boundaries based on
@@
 \texttt{orgQuants} function to extract the median organelle scores and
-then pass these scores to the \texttt{getPrediction} function to
+then pass these scores to the \texttt{getPredictions} function to
 extract the new localisations that meet this scoring criteria. Any
@@
 (PCA) also fall into the category of unsupervised learning methods,
-where the data can be projected from a high-dimensional spcae down to
+where the data can be projected from a high-dimensional space down to
 two or three dimensions for the purpose of visualisation.

### 2016-11-08 bioc-workflow.tex

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 14:13:46 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 16:12:13 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 14:13:46 2016 
-## Subset [6725,20][5032,20] Tue Nov  8 14:13:47 2016 
-## Removed features with more than 0 NAs: Tue Nov  8 14:13:47 2016 
-## Dropped featureData's levels Tue Nov  8 14:13:47 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 16:12:13 2016 
+## Subset [6725,20][5032,20] Tue Nov  8 16:12:14 2016 
+## Removed features with more than 0 NAs: Tue Nov  8 16:12:14 2016 
+## Dropped featureData's levels Tue Nov  8 16:12:14 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Tue Nov  8 14:13:50 2016
+% Tue Nov  8 16:12:16 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  8 14:13:49 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  8 14:13:50 2016 
+## Added markers from  'mrk' marker vector. Tue Nov  8 16:12:16 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  8 16:12:16 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4359409               0.3031417               0.3896552 
+##               0.4365906               0.3054928               0.3796591 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6869014               0.6046754               0.4336399 
+##               0.6956146               0.5986152               0.4308889 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4165647               0.5910830               0.9498244 
+##               0.4270518               0.5894672               0.9501479 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7955788               0.7079722               0.3121358 
+##               0.7954126               0.7078652               0.3176106 
 ##         Plasma membrane              Proteasome 
-##               0.7187306               0.4168615
+##               0.7149800               0.4148348
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     171                      89 
+##                      84                     173                      86 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     294                     477                      99 
+##                     296                     476                     100 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     522 
+##                      27                     124                     521 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     342                      38 
+##                     229                     343                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     316                     160                    2058
+##                     319                     158                    2057
 \end{verbatim}
@@
 
-The authors would like to thank Dr Claire M. Mulvey for helpful
-comments on early versions of this manuscript. The authors would also
-like to thank Dr. Stuart Rankin from the High Performance Computing
-Service for his support. Part of this work was performed using the
-Darwin Supercomputer of the University of Cambridge High Performance
-Computing Service, provided by Dell Inc. using Strategic Research
-Infrastructure Funding from the Higher Education Funding Council for
-England and funding from the Science and Technology Facilities
-Council.
+The authors would like to thank Dr. Stuart Rankin from the High
+Performance Computing Service for his support. Part of this work was
+performed using the Darwin Supercomputer of the University of
+Cambridge High Performance Computing Service, provided by Dell
+Inc. using Strategic Research Infrastructure Funding from the Higher
+Education Funding Council for England and funding from the Science and
+Technology Facilities Council.

### 2016-11-08 bioc-workflow.tex
recompile with Lisa's latest ammendments

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 16:12:13 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:46:40 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 16:12:13 2016 
-## Subset [6725,20][5032,20] Tue Nov  8 16:12:14 2016 
-## Removed features with more than 0 NAs: Tue Nov  8 16:12:14 2016 
-## Dropped featureData's levels Tue Nov  8 16:12:14 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:46:40 2016 
+## Subset [6725,20][5032,20] Tue Nov  8 19:46:41 2016 
+## Removed features with more than 0 NAs: Tue Nov  8 19:46:41 2016 
+## Dropped featureData's levels Tue Nov  8 19:46:41 2016 
 ##  MSnbase version: 1.21.7
@@
 \hlkwd{plotDist}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Mitochondrion"}\hlstd{, ],}
-         \hlkwc{pcol} \hlstd{=} \hlstr{"purple"}\hlstd{,} \hlkwc{fractions} \hlstd{=} \hlstr{"Fraction.No"}\hlstd{)}
+         \hlkwc{pcol} \hlstd{=} \hlstr{"purple"}\hlstd{,} \hlkwc{fractions} \hlstd{=} \hlstr{"Tag"}\hlstd{)}
 \hlkwd{title}\hlstd{(}\hlkwc{main} \hlstd{=} \hlstr{"Marker occupancy profiles along the gradient"}\hlstd{)}
@@
 fraction collection, the quantitative channels do not correspond to
-identical selected fractions along the gradient. As can be seen in the
-table below (taken from the \texttt{hl}'s \texttt{pData}, focusing on
-channels 7 to 10), different sets of gradient fractions are pooled to
-obtain enough material and optimise acurate quantitation.
+identical selected fractions along the gradient. For example, in the
+table below (taken from \texttt{hl}'s \texttt{pData}) TMT channels
+127C (among others) in both replicates originate from different sets
+of gradient fractions (gradient fractions 7 - 9 and 8 - 9 for each
+replicate, respectively). Different sets of gradient fractions are
+often pooled to obtain enough material and optimise acurate
+quantitation.
 
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Tue Nov  8 16:12:16 2016
+% Tue Nov  8 19:46:43 2016
 \begin{table}[ht]
@@
   \hline
-X127N &   1 & 127N & 1 to 6 (pooled) & 6.00 \\ 
 X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
-  X128N &   1 & 128N & 10 to 11 (pooled) & 13.30 \\ 
   X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
-  X128N.1 &   2 & 128N & 10 to 11 (pooled) & 12.50 \\ 
    \hline
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  8 16:12:16 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  8 16:12:16 2016 
+## Added markers from  'mrk' marker vector. Tue Nov  8 19:46:42 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  8 19:46:43 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4365906               0.3054928               0.3796591 
+##               0.4325847               0.3029421               0.3817418 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6956146               0.5986152               0.4308889 
+##               0.6933699               0.6027103               0.4145111 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4270518               0.5894672               0.9501479 
+##               0.4276585               0.5845256               0.9490812 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7954126               0.7078652               0.3176106 
+##               0.7947588               0.7079465               0.3161762 
 ##         Plasma membrane              Proteasome 
-##               0.7149800               0.4148348
+##               0.7101150               0.4125186
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     173                      86 
+##                      85                     172                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     476                     100 
+##                     297                     475                      98 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     521 
+##                      28                     125                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     343                      39 
+##                     228                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     319                     158                    2057
+##                     321                     158                    2057
 \end{verbatim}

### 2016-11-08 bioc-workflow.Rnw
Fixed a and an MSnSet

@@
 convenience function, \texttt{addMarkers}, to directly add markers to
-a \texttt{MSnSet} object, as demonstrated in the code chunk
+an \texttt{MSnSet} object, as demonstrated in the code chunk
 below. These marker sets can be accessed using the
@@
 etc. The compare application is called by default if the input object
-to \texttt{pRolocVis} is a \texttt{MSnSetList} of 2 \texttt{MSnSets},
+to \texttt{pRolocVis} is an \texttt{MSnSetList} of 2 \texttt{MSnSets},
 but it can also be specified by calling the argument \texttt{app =
-  "compare"}. For example, in the code chunk below we first create a
+  "compare"}. For example, in the code chunk below we first create an
 \texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT data, this

### 2016-11-08 bioc-workflow.Rnw
use a*n* MSnSet - closes issue #31

@@
 
-The spreadsheet that was used to create the \texttt{hl} MSnSet
-included the two replicates within one .csv file.  We also provide
-individual replicates in the \Biocexptpkg{pRolocdata}
+The spreadsheet that was used to create the \texttt{hl}
+\texttt{MSnSet} included the two replicates within one .csv file.  We
+also provide individual replicates in the \Biocexptpkg{pRolocdata}
 package. Below, we show how to combine \texttt{MSnSet} objects and,
 subsequently, how to filter and handle missing values. We start by
-loading the \Biocexptpkg{pRolocdata} package and the
-equivalent replicates using the \texttt{data} function.
+loading the \Biocexptpkg{pRolocdata} package and the equivalent
+replicates using the \texttt{data} function.
 
@@
 
-These markers can then be mapped to a \texttt{MSnSet}'s
+These markers can then be mapped to an \texttt{MSnSet}'s
 \texttt{featureNames}. The mouse dataset used here has Uniprot IDs

### 2016-11-08 bioc-workflow.tex
use a*n* MSnSet - closes issue #31

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:46:40 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:55:51 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:46:40 2016 
-## Subset [6725,20][5032,20] Tue Nov  8 19:46:41 2016 
-## Removed features with more than 0 NAs: Tue Nov  8 19:46:41 2016 
-## Dropped featureData's levels Tue Nov  8 19:46:41 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:55:51 2016 
+## Subset [6725,20][5032,20] Tue Nov  8 19:55:51 2016 
+## Removed features with more than 0 NAs: Tue Nov  8 19:55:51 2016 
+## Dropped featureData's levels Tue Nov  8 19:55:51 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Tue Nov  8 19:46:43 2016
+% Tue Nov  8 19:55:53 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  8 19:46:42 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  8 19:46:43 2016 
+## Added markers from  'mrk' marker vector. Tue Nov  8 19:55:53 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  8 19:55:54 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4325847               0.3029421               0.3817418 
+##               0.4357194               0.3041525               0.3811012 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6933699               0.6027103               0.4145111 
+##               0.6836987               0.6053745               0.4209542 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4276585               0.5845256               0.9490812 
+##               0.4212970               0.5861141               0.9494434 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7947588               0.7079465               0.3161762 
+##               0.7966054               0.7108833               0.3132191 
 ##         Plasma membrane              Proteasome 
-##               0.7101150               0.4125186
+##               0.7131673               0.4134914
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      84 
+##                      85                     171                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     297                     475                      98 
+##                     292                     476                      99 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      28                     125                     523 
+##                      25                     124                     524 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     228                     343                      38 
+##                     229                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     321                     158                    2057
+##                     322                     162                    2058
 \end{verbatim}

### 2016-11-08 bioc-workflow.tex
a couple of additional a*n* MSnSet

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:55:51 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:57:45 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:55:51 2016 
-## Subset [6725,20][5032,20] Tue Nov  8 19:55:51 2016 
-## Removed features with more than 0 NAs: Tue Nov  8 19:55:51 2016 
-## Dropped featureData's levels Tue Nov  8 19:55:51 2016 
+## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:57:45 2016 
+## Subset [6725,20][5032,20] Tue Nov  8 19:57:46 2016 
+## Removed features with more than 0 NAs: Tue Nov  8 19:57:46 2016 
+## Dropped featureData's levels Tue Nov  8 19:57:46 2016 
 ##  MSnbase version: 1.21.7
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Tue Nov  8 19:55:53 2016
+% Tue Nov  8 19:57:48 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  8 19:55:53 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  8 19:55:54 2016 
+## Added markers from  'mrk' marker vector. Tue Nov  8 19:57:47 2016 
+## Added markers from  'pdres' marker vector. Tue Nov  8 19:57:48 2016 
 ##  MSnbase version: 2.0.0
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4357194               0.3041525               0.3811012 
+##               0.4361397               0.3054709               0.3865287 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6836987               0.6053745               0.4209542 
+##               0.6876809               0.6130306               0.4310852 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4212970               0.5861141               0.9494434 
+##               0.4143871               0.6198125               0.9504927 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7966054               0.7108833               0.3132191 
+##               0.7985420               0.7115356               0.3171008 
 ##         Plasma membrane              Proteasome 
-##               0.7131673               0.4134914
+##               0.7140531               0.4219360
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      84 
+##                      85                     172                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     292                     476                      99 
+##                     296                     475                     107 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      25                     124                     524 
+##                      27                     118                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     343                      38 
+##                     229                     347                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     162                    2058
+##                     315                     155                    2056
 \end{verbatim}

### 2016-11-09 code/submission.sh
TL code for cluster

@@ -1,0 +1,123 @@
+#!/bin/bash
+#!
+#! Example SLURM job script for Darwin (Sandy Bridge, ConnectX3)
+#!
+
+#!#############################################################
+#!#### Modify the options in this section as appropriate ######
+#!#############################################################
+
+#! sbatch directives begin here ###############################
+#! Name of the job:
+#SBATCH -J transfer_learning
+#! Which project should be charged:
+#SBATCH -A GATTO
+#! How many whole nodes should be allocated?
+#SBATCH --nodes=16
+#! How many (MPI) tasks will there be in total? (<= nodes*16)
+#SBATCH --ntasks=256
+#! How much wallclock time will be required?
+#SBATCH --time=12:00:00
+#! What types of email messages do you wish to receive?
+#SBATCH --mail-type=ALL
+#! Uncomment this to prevent the job from being requeued (e.g. if
+#! interrupted by node failure or system downtime):
+##SBATCH --no-requeue
+
+#! Do not change:
+#SBATCH -p sandybridge
+
+#! sbatch directives end here (put any additional directives above this line)
+
+#! Notes:
+#! Charging is determined by node number*walltime. Allocation is in entire nodes.
+#! The --ntasks value refers to the number of tasks to be launched by SLURM only. This
+#! usually equates to the number of MPI tasks launched. Reduce this from nodes*16 if
+#! demanded by memory requirements, or if OMP_NUM_THREADS>1.
+
+#! Number of nodes and tasks per node allocated by SLURM (do not change):
+numnodes=$SLURM_JOB_NUM_NODES
+numtasks=$SLURM_NTASKS
+mpi_tasks_per_node=$(echo "$SLURM_TASKS_PER_NODE" | sed -e  's/^\([0-9][0-9]*\).*$/\1/')
+#! ############################################################
+#! Modify the settings below to specify the application's environment, location 
+#! and launch method:
+
+#! Optionally modify the environment seen by the application
+#! (note that SLURM reproduces the environment at submission irrespective of ~/.bashrc):
+. /etc/profile.d/modules.sh                # Leave this line (enables the module command)
+module purge                               # Removes all modules still loaded
+module load default-impi                   # REQUIRED - loads the basic environment
+
[…]

### 2016-11-09 code/tl.R

@@ -1,0 +1,50 @@
+## Tell R where to look for my packages 
+
+print(.libPaths())
+lp = "/scratch/gatto/R_local"
+
+## Load libraries in my home directory
+library("MLInterfaces", lib.loc = lp)
+library("pRoloc", lib.loc = lp)
+library("pRolocdata", lib.loc = lp)
+library("BiocParallel", lib.loc = lp)
+library("Rmpi")
+stopifnot(packageVersion("MLInterfaces") >= "1.53.1")
+
+
+## Load LOPIT and GOCC data from pRolocdata package
+data("hyperLOPIT2015")
+data("hyperLOPIT2015goCC")
+
+
+## keep only 7 overlapping classes for example run for workflow
+## rename the following classes as "unknown"
+torm  <- c("40S Ribosome", "60S Ribosome",
+           "Nucleus - Chromatin", "Nucleus - Non-chromatin")
+for (i in seq(torm)) {
+  hyperLOPIT2015 <- fDataToUnknown(hyperLOPIT2015,
+                                   from = torm[i])
+  hyperLOPIT2015goCC <- fDataToUnknown(hyperLOPIT2015goCC,
+                                       from = torm[i])
+}
+
+
+## Set params for parallelisation
+par <- SnowParam(255L, type = "MPI",
+                 stop.on.error = FALSE,
+                 log = TRUE,
+                 threshold = "DEBUG")
+
+
+## run opt
+res <- knntlOptimisation(hyperLOPIT2015, hyperLOPIT2015goCC, 
+                            fcol = "markers", 
+                            length.out = 4, 
+                            times = 50,
+                            xval = 5, k = c(3,3), 
+                            BPPARAM = par)
+
+save(res, file = "tl-hyperlopit.rda")
+
+# Tell all slaves to close down, and exit the program
+mpi.quit()

### 2016-11-09 bioc-workflow.Rnw
Added sentence to TL section

@@
 information with our primary MS-based quantitative data using a
-paradigm known as transfer learning (TL). The integration of data
-between different technologies is one of the biggest challenges in
-computational biology to date and the \Biocpkg{pRoloc} package
-provides functionality to do such analyses. We recently developed two
-transfer learning algorithms using a $k$-NN and SVM framework and
-applied them to the task of protein localisation prediction
-\cite{Breckels:2016}. In this section we will begin with explaining the
-concept of transfer learning and then show how to apply this in the
-frame of spatial proteomics and protein localisation prediction.
+paradigm known as transfer learning (TL). Transfer learning may be
+particuarly useful when sub-cellular niches are not equally
+well-resolved in a particular experiment, or if we have overlapping
+clusters in the data.
+
+The integration of data between different technologies is one of the
+biggest challenges in computational biology to date and the
+\Biocpkg{pRoloc} package provides functionality to do such analyses.
+We recently developed two transfer learning algorithms using a $k$-NN
+and SVM framework and applied them to the task of protein localisation
+prediction \cite{Breckels:2016}. In this section we will begin with
+explaining the concept of transfer learning and then show how to apply
+this in the frame of spatial proteomics and protein localisation
+prediction.
 
@@
 accuracy from data collected from quantitative MS-based experiments.
+We have already seen that straightforward supervised ML works well for
+these types of experiments, however, 
+Transfer learning is particularly useful for
+classes that are not as well separated.
+
 In the example below we extract Gene Ontology (GO) information to use

### 2016-11-10 hpc/tl.R
don't need to check MLInterfaces' version

@@ -9,8 +9,6 @@ library("pRoloc", lib.loc = lp)
 library("pRolocdata", lib.loc = lp)
 library("BiocParallel", lib.loc = lp)
 library("Rmpi")
-stopifnot(packageVersion("MLInterfaces") >= "1.53.1")
-
 
 ## Load LOPIT and GOCC data from pRolocdata package
 data("hyperLOPIT2015")

### 2016-11-30 bioc-workflow.Rnw
few updates

--- 2016-11-08 bioc-workflow.Rnw
@@
 These, as well as many other data are directly available as properly
-structured and annotated computational object from the 
+structured and annotated computational objects from the 
 \Biocexptpkg{pRolocdata} experiment package. In this workflow, we
@@
   \centering
-  \includegraphics[width=.65\textwidth]{./Figures/spreadsheet-screenshot.png}
+  \includegraphics[width=.85\textwidth]{./Figures/spreadsheet-screenshot.png}
   \caption{A screenshot of the data in the spreadsheet.}
@@
 components ordered according to the amount of variability that they
-describe. The \texttt{plot2D} method in \Biocpkg{pRoloc} allows one to
+describe. The \texttt{plot2D} and \texttt{plot3D} methods in \Biocpkg{pRoloc} allows one to
 plot the principal components (PCs) of a dataset against one another,
-by default the first two components are plotted on the x- and y-axis,
+by default the first two components are plotted on the x- and y-axis for the \texttt{plot2D} function, and first three components are loaded for the \texttt{plot3D} function,
 respectively (the \texttt{dims} argument can be used to plot other PCs). If
@@
 
-In the code chunk below we produce a PCA plot of the mouse stem cell
+In the code chunk below we produce a 2-dimensional PCA plot of the mouse stem cell
 dataset (Figure \ref{fig:pcahl}). One point on the plot represents one
@@
 must match some (but not necessarily all) of the \texttt{MSnSet}'s
-feature names. See `?addMarkers` for more details.
+feature names. See \texttt{?addMarkers} for more details.
 
@@
 two figures with the same orientation.
+The interactive "compare" application part of the \Biocpkg{pRolocGUI} package is also useful for examining replicate experiments (see the next section interactive visualisation for details).
 
@@
 \textit{compare}) which are wrapped and launched by the
-\texttt{pRolocVis} function. In the below code chunk we lauch the main
+\texttt{pRolocVis} function.
+
+\subsubsection*{The main application}
+In the below code chunk we lauch the main
 app (note, we do not need to specify the argument, \texttt{app =
@@
 
+\subsubsection*{The compare application}
 The \textit{compare} application is useful for examining two replicate
@@
 
+\subsubsection*{The classify application}
 The final application \textit{classify}, has been designed to view
@@
 applied them to the task of protein localisation prediction
-\cite{Breckels:2016}. In this section we will begin with explaining the
-concept of transfer learning and then show how to apply this in the
-frame of spatial proteomics and protein localisation prediction.
+\cite{Breckels:2016}. In this section we will begin with explaining
+the concept of transfer learning and then show how to apply this in
+the frame of spatial proteomics and protein localisation prediction.
 
@@
 accuracy from data collected from quantitative MS-based experiments.
+We have already seen that straightforward supervised ML works well for
+these types of experiments, however, Transfer learning is particularly
+useful for classes that are not as well separated.
+
 In the example below we extract Gene Ontology (GO) information to use

### 2016-11-30 bioc-workflow.Rnw
Added section on foi

@@
 
+\subsection{*Adding user-defined marker lists}
 It is also possible for users to use their own marker list with the
@@
 
-We now visualise these annotations along the PCA plot using the
+In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
+particular the cellular compartment (CC) namespace are a good starting
+point for protein annotation and marker definition. It is important to
+note however that automatic retrieval of sub-cellular localisation
+information, from \Biocpkg{pRoloc} or elsewhere, is only the
+beginning in defining a marker set for downstream analyses. Expert
+curation is vital to check that any annotation added is in the correct
+context for the biological question under investigation.
+
+\subsection{*Visualising markers}
+Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
 \texttt{plot2D} function and then use the \texttt{addLegend} function
@@
 
-In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
-particular the cellular compartment (CC) namespace are a good starting
-point for protein annotation and marker definition. It is important to
-note however that automatic retrieval of sub-cellular localisation
-information, from \Biocpkg{pRoloc} or elsewhere, is only the
-beginning in defining a marker set for downstream analyses. Expert
-curation is vital to check that any annotation added is in the correct
-context for the biological question under investigation.
-
 Another useful visualisation that relies on marker annotation is the
@@
 allows to interactively explore a 3-dimensional plot of the data.
+
+\subsection*{Features of interest}
+In addition to adding annotation using the \texttt{addMarkers} function one can store specific sets of proteins by using the \textit{Features of interest} infrastructure from the \Biocpkg{MSnbase} package.
+If users have a specific subset of proteins they wish to highlight in their data they would first create a \texttt{FeaturesOfInterest} object.
+For example, if we wanted to highlight a proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some of the proteins known to form part of the 13S condensin complex, we would create a first create a \texttt{FeaturesOfInterest} object then use the function \texttt{highlightOnPlot} to highlight their location on a PCA plot.
+
+\begin{figure}[!ht]
+  \centering
+<<foi, fig.width=10, out.width=".8\\textwidth">>= 
+prots <- c("Q8CG48", "Q8CG47", "Q8K2Z4", "Q8C156")
+foi13s <- FeaturesOfInterest(description = "13S consensin proteins",
+                             fnames = prots, 
+                             object = hl)
+foi13s
+
+plot2D(hl)
+addLegend(hl, cex = .6)
+highlightOnPlot(hl, foi13s)
+@  
+  \caption{Highlighting protein features of interest}
+  \label{fig:foi}
+\end{figure}
+
+Users can also create several sets of \texttt{FeaturesOfInterest} object and store them in a \texttt{FoICollection}.
+
+It is also worthly of note that it is possible to search for a specific protein of interest by \texttt{featureNames} or using any identifying information found in the \texttt{fData} columns by using the search box on the \texttt{pRolocVis} application part of the \Biocpkg{pRolocGUI} package (see section on interactive visualisation).
+This can be handy for quickly searching and highlighting proteins on the fly, the disavanatge here is that proteins can only be searched for a one-by-one basis.
 

### 2016-11-30 bioc-workflow.tex
Added section on foi

--- 2016-11-08 bioc-workflow.tex
@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.0.0
+##  MSnbase version: 2.0.1
 \end{verbatim}
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:55:51 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 14:38:29 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Nov  8 19:55:51 2016 
-## Subset [6725,20][5032,20] Tue Nov  8 19:55:51 2016 
-## Removed features with more than 0 NAs: Tue Nov  8 19:55:51 2016 
-## Dropped featureData's levels Tue Nov  8 19:55:51 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 14:38:29 2016 
+## Subset [6725,20][5032,20] Wed Nov 30 14:38:29 2016 
+## Removed features with more than 0 NAs: Wed Nov 30 14:38:29 2016 
+## Dropped featureData's levels Wed Nov 30 14:38:29 2016 
 ##  MSnbase version: 1.21.7
@@
 must match some (but not necessarily all) of the \texttt{MSnSet}'s
-feature names. See `?addMarkers` for more details.
-
-We now visualise these annotations along the PCA plot using the
+feature names. See \texttt{?addMarkers} for more details.
+
+In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
+particular the cellular compartment (CC) namespace are a good starting
+point for protein annotation and marker definition. It is important to
+note however that automatic retrieval of sub-cellular localisation
+information, from \Biocpkg{pRoloc} or elsewhere, is only the
+beginning in defining a marker set for downstream analyses. Expert
+curation is vital to check that any annotation added is in the correct
+context for the biological question under investigation.
+
+\subsection{*Visualising markers}
+Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
 \texttt{plot2D} function and then use the \texttt{addLegend} function
@@
 
+\subsection*{Features of interest}
+In addition to adding annotation using the \texttt{addMarkers} function one can store specific sets of proteins by using the \textit{Features of interest} infrastructure from the \Biocpkg{MSnbase} package.
+If users have a specific subset of proteins they wish to highlight in their data they would first create a \texttt{FeaturesOfInterest} object.
+For example, if we wanted to highlight a proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some of the proteins known to form part of the 13S condensin complex, we would create a first create a \texttt{FeaturesOfInterest} object then use the function \texttt{highlightOnPlot} to highlight their location on a PCA plot.
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{prots} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Q8CG48"}\hlstd{,} \hlstr{"Q8CG47"}\hlstd{,} \hlstr{"Q8K2Z4"}\hlstd{,} \hlstr{"Q8C156"}\hlstd{)}
+\hlstd{foi13s} \hlkwb{<-} \hlkwd{FeaturesOfInterest}\hlstd{(}\hlkwc{description} \hlstd{=} \hlstr{"13S consensin proteins"}\hlstd{,}
+                             \hlkwc{fnames} \hlstd{= prots,}
+                             \hlkwc{object} \hlstd{= hl)}
+\hlstd{foi13s}
+\end{alltt}
+\begin{verbatim}
+## Traceable object of class "FeaturesOfInterest"
+##  Created on Wed Nov 30 14:38:32 2016 
+##  Description:
+##   13S consensin proteins
+##  4 features of interest:
+##    Q8CG48, Q8CG47, Q8K2Z4, Q8C156
+\end{verbatim}
+\begin{alltt}
+\hlkwd{plot2D}\hlstd{(hl)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
+\hlkwd{highlightOnPlot}\hlstd{(hl, foi13s)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/foi-1}
+
+\end{knitrout}
+  \caption{Highlighting protein features of interest}
+  \label{fig:foi}
+\end{figure}
+
+Users can also create several sets of \texttt{FeaturesOfInterest} object and store them in a \texttt{FoICollection}.
+
+It is also worthly of note that it is possible to search for a specific protein of interest by \texttt{featureNames} or using any identifying information found in the \texttt{fData} columns by using the search box on the \texttt{pRolocVis} application part of the \Biocpkg{pRolocGUI} package (see section on interactive visualisation).
+This can be handy for quickly searching and highlighting proteins on the fly, the disavanatge here is that proteins can only be searched for a one-by-one basis.
+
 \section*{Replication}
@@
 
-% latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Tue Nov  8 19:55:53 2016
+% latex table generated in R 3.3.0 by xtable 1.8-2 package
+% Wed Nov 30 14:38:32 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Nov  8 19:55:53 2016 
-## Added markers from  'pdres' marker vector. Tue Nov  8 19:55:54 2016 
-##  MSnbase version: 2.0.0
+## Added markers from  'mrk' marker vector. Wed Nov 30 14:38:31 2016 
+## Added markers from  'pdres' marker vector. Wed Nov 30 14:38:32 2016 
+##  MSnbase version: 2.0.1
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4357194               0.3041525               0.3811012 
+##               0.4346815               0.3050109               0.3898494 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6836987               0.6053745               0.4209542 
+##               0.6915344               0.6018374               0.4277405 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4212970               0.5861141               0.9494434 
+##               0.4215606               0.6043865               0.9499822 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7966054               0.7108833               0.3132191 
+##               0.7959813               0.7080413               0.3162206 
 ##         Plasma membrane              Proteasome 
-##               0.7131673               0.4134914
+##               0.7248143               0.4135359
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      84 
+##                      84                     171                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     292                     476                      99 
+##                     296                     476                     109 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      25                     124                     524 
+##                      24                     119                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     343                      38 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     162                    2058
+##                     315                     158                    2058
 \end{verbatim}
@@
 \begin{verbatim}
-## R version 3.3.2 Patched (2016-11-01 r71616)
-## Platform: x86_64-pc-linux-gnu (64-bit)
-## Running under: Ubuntu 14.04.5 LTS
+## R Under development (unstable) (2015-11-11 r69621)
+## Platform: x86_64-apple-darwin13.4.0 (64-bit)
+## Running under: OS X 10.11.5 (El Capitan)
 ## 
 ## locale:
-##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
-##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
-##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
-##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
-##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
-## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
+## [1] en_GB.UTF-8/en_GB.UTF-8/en_GB.UTF-8/C/en_GB.UTF-8/en_GB.UTF-8
 ## 
@@
 ## other attached packages:
-##  [1] pRolocdata_1.12.0    pRoloc_1.14.3        MLInterfaces_1.54.0 
-##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.36.0 IRanges_2.8.0        S4Vectors_0.12.0    
-## [10] MSnbase_2.0.0        ProtGenerics_1.6.0   BiocParallel_1.8.1  
-## [13] mzR_2.8.0            Rcpp_0.12.7          Biobase_2.34.0      
+##  [1] pRolocdata_1.12.0    pRoloc_1.14.4        MLInterfaces_1.54.0 
+##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.5        
+##  [7] AnnotationDbi_1.36.0 IRanges_2.5.40       S4Vectors_0.9.45    
+## [10] MSnbase_2.0.1        ProtGenerics_1.6.0   BiocParallel_1.8.1  
+## [13] mzR_2.8.0            Rcpp_0.12.8          Biobase_2.34.0      
 ## [16] BiocGenerics_0.20.0  gridExtra_2.2.1      xtable_1.8-2        
-## [19] BiocStyle_2.2.0      knitr_1.14          
+## [19] BiocStyle_2.2.1      knitr_1.15.1        
 ## 
 ## loaded via a namespace (and not attached):
-##   [1] minqa_1.2.4           colorspace_1.2-7      hwriter_1.3.2        
+##   [1] minqa_1.2.4           colorspace_1.3-1      hwriter_1.3.2        
 ##   [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
@@
 ##  [13] flexmix_2.3-13        mvtnorm_1.0-5         codetools_0.2-15     
-##  [16] splines_3.3.2         doParallel_1.0.10     impute_1.48.0        
+##  [16] splines_3.3.0         doParallel_1.0.10     impute_1.48.0        
 ##  [19] robustbase_0.92-6     jsonlite_1.1          nloptr_1.0.4         
-##  [22] caret_6.0-72          pbkrtest_0.4-6        rda_1.0.2-2          
+##  [22] caret_6.0-73          pbkrtest_0.4-6        rda_1.0.2-2          
 ##  [25] kernlab_0.9-25        vsn_3.42.3            sfsmisc_1.1-0        
 ##  [28] shiny_0.14.2          sampling_2.7          assertthat_0.1       
-##  [31] Matrix_1.2-7.1        limma_3.30.2          formatR_1.4          
-##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.2          
+##  [31] Matrix_1.2-7.1        lazyeval_0.2.0        limma_3.30.5         
+##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.0          
 ##  [37] ggvis_0.4.3           gtable_0.2.0          affy_1.52.0          
@@
 ##  [52] mime_0.5              gtools_3.5.0          dendextend_1.3.0     
-##  [55] DEoptimR_1.0-6        zlibbioc_1.20.0       MASS_7.3-45          
-##  [58] scales_0.4.0          BiocInstaller_1.24.0  pcaMethods_1.66.0    
-##  [61] SparseM_1.72          RColorBrewer_1.1-2    ggplot2_2.1.0        
-##  [64] biomaRt_2.30.0        rpart_4.1-10          stringi_1.1.2        
-##  [67] RSQLite_1.0.0         highr_0.6             genefilter_1.56.0    
-##  [70] randomForest_4.6-12   foreach_1.4.3         e1071_1.6-7          
-##  [73] prabclus_2.2-6        bitops_1.0-6          mzID_1.12.0          
-##  [76] evaluate_0.10         lattice_0.20-34       htmlwidgets_0.7      
-##  [79] gbm_2.1.1             plyr_1.8.4            magrittr_1.5         
-##  [82] R6_2.2.0              DBI_0.5-1             whisker_0.3-2        
-##  [85] mgcv_1.8-15           survival_2.40-1       RCurl_1.95-4.8       
-##  [88] nnet_7.3-12           tibble_1.2            msdata_0.14.0        
-##  [91] car_2.1-3             mlbench_2.1-1         grid_3.3.2           
+##  [55] DEoptimR_1.0-8        zlibbioc_1.20.0       MASS_7.3-45          
+##  [58] scales_0.4.1          BiocInstaller_1.24.0  pcaMethods_1.66.0    
+##  [61] SparseM_1.74          RColorBrewer_1.1-2    memoise_1.0.0        
+##  [64] ggplot2_2.2.0         biomaRt_2.30.0        rpart_4.1-10         
+##  [67] stringi_1.1.2         RSQLite_1.1           highr_0.6            
+##  [70] genefilter_1.56.0     randomForest_4.6-12   foreach_1.4.3        
+##  [73] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
+##  [76] mzID_1.12.0           evaluate_0.10         lattice_0.20-34      
+##  [79] htmlwidgets_0.8       gbm_2.1.1             plyr_1.8.4           
+##  [82] magrittr_1.5          R6_2.2.0              DBI_0.5-1            
+##  [85] whisker_0.3-2         mgcv_1.8-16           survival_2.40-1      
+##  [88] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
+##  [91] car_2.1-3             mlbench_2.1-1         grid_3.3.0           
 ##  [94] FNN_1.1               ModelMetrics_1.1.0    threejs_0.2.2        

### 2016-11-30 bioc-workflow.Rnw
rephrasong and latex syntax

--- 2016-11-30 bioc-workflow.Rnw
@@
 
+\subsection*{Adding user-defined marker lists}
 It is also possible for users to use their own marker list with the
@@
 
-We now visualise these annotations along the PCA plot using the
+In general, the Gene Ontology (GO) \cite{Ashburner:2000}, and in
+particular the cellular compartment (CC) namespace are a good starting
+point for protein annotation and marker definition. It is important to
+note however that automatic retrieval of sub-cellular localisation
+information, from \Biocpkg{pRoloc} or elsewhere, is only the
+beginning in defining a marker set for downstream analyses. Expert
+curation is vital to check that any annotation added is in the correct
+context for the biological question under investigation.
+
+\subsection*{Visualising markers}
+Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
 \texttt{plot2D} function and then use the \texttt{addLegend} function
@@
 allows to interactively explore a 3-dimensional plot of the data.
+
+\subsection*{Features of interest}
+
+In addition to adding annotation using the \texttt{addMarkers}
+function, one can store specific sets of proteins by using the
+\textit{Features of interest} infrastructure from the
+\Biocpkg{MSnbase} package. If users have specific subsets of proteins
+they wish to highlight in their data (possibly across multiple
+experiments) they would first create a \texttt{FeaturesOfInterest}
+object. For example, if we wanted to highlight a proteins with the
+accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some
+of the proteins known to form part of the 13S condensin complex, we
+would create a first create a \texttt{FeaturesOfInterest} object, and
+subsequently highlight their location on a PCA plot with the
+\texttt{highlightOnPlot} function.
+
+\begin{figure}[!ht]
+  \centering
+<<foi, fig.width=10, out.width=".8\\textwidth">>= 
+prots <- c("Q8CG48", "Q8CG47", "Q8K2Z4", "Q8C156")
+foi13s <- FeaturesOfInterest(description = "13S consensin proteins",
+                             fnames = prots, 
+                             object = hl)
+foi13s
+
+plot2D(hl)
+addLegend(hl, cex = .6)
+highlightOnPlot(hl, foi13s)
+@  
+  \caption{Highlighting protein features of interest.}
+  \label{fig:foi}
+\end{figure}
+
+Users can also create several sets of \texttt{FeaturesOfInterest}
+object and store them in a \texttt{FoICollection}.
+
+It is also worthly of note that it is possible to search for a
+specific protein of interest by \texttt{featureNames} or using any
+identifying information found in the \texttt{fData} columns by using
+the search box on the \texttt{pRolocVis} application part of the
+\Biocpkg{pRolocGUI} package (see section on interactive
+visualisation). This can be handy for quickly searching and
+highlighting proteins on the fly, the disavanatge here is that
+proteins can only be searched for a one-by-one basis.
 

### 2016-11-30 bioc-workflow.tex
rephrasong and latex syntax

@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.0.1
+##  MSnbase version: 2.1.4
 \end{verbatim}
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 14:38:29 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:02:25 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 14:38:29 2016 
-## Subset [6725,20][5032,20] Wed Nov 30 14:38:29 2016 
-## Removed features with more than 0 NAs: Wed Nov 30 14:38:29 2016 
-## Dropped featureData's levels Wed Nov 30 14:38:29 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:02:25 2016 
+## Subset [6725,20][5032,20] Wed Nov 30 15:02:26 2016 
+## Removed features with more than 0 NAs: Wed Nov 30 15:02:26 2016 
+## Dropped featureData's levels Wed Nov 30 15:02:26 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Homo sapiens [hsap]:
-##  Ids: Uniprot ID, 205 markers
+##  Ids: Uniprot, 872 markers
 ## Mus musculus [mmus]:
@@
 
-\subsection{*Adding user-defined marker lists}
+\subsection*{Adding user-defined marker lists}
 It is also possible for users to use their own marker list with the
@@
 
-\subsection{*Visualising markers}
+\subsection*{Visualising markers}
 Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
@@
          \hlkwc{pcol} \hlstd{=} \hlstr{"purple"}\hlstd{,} \hlkwc{fractions} \hlstd{=} \hlstr{"Tag"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[.data.frame`(pData(object), , .frac): object '.frac' not found}}\begin{alltt}
 \hlkwd{title}\hlstd{(}\hlkwc{main} \hlstd{=} \hlstr{"Marker occupancy profiles along the gradient"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in title(main = "{}Marker occupancy profiles along the gradient"{}): plot.new has not been called yet}}\begin{alltt}
 \hlkwd{matlines}\hlstd{(}\hlkwd{t}\hlstd{(}\hlkwd{exprs}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Peroxisome"}\hlstd{, ])),}
          \hlkwc{lty} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"darkblue"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"l"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot.xy(xy.coords(x, y), type = type, ...): plot.new has not been called yet}}\begin{alltt}
 \hlkwd{legend}\hlstd{(}\hlstr{"topleft"}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"Mitochondrion"}\hlstd{,} \hlstr{"Peroxisome"}\hlstd{),}
@@
 \end{alltt}
-\end{kframe}
-\includegraphics[width=.6\textwidth]{figure/plotDist-1}
-
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strwidth(legend, units = "{}user"{}, cex = cex, font = text.font): plot.new has not been called yet}}\end{kframe}
 \end{knitrout}
@@
 \subsection*{Features of interest}
-In addition to adding annotation using the \texttt{addMarkers} function one can store specific sets of proteins by using the \textit{Features of interest} infrastructure from the \Biocpkg{MSnbase} package.
-If users have a specific subset of proteins they wish to highlight in their data they would first create a \texttt{FeaturesOfInterest} object.
-For example, if we wanted to highlight a proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some of the proteins known to form part of the 13S condensin complex, we would create a first create a \texttt{FeaturesOfInterest} object then use the function \texttt{highlightOnPlot} to highlight their location on a PCA plot.
+
+In addition to adding annotation using the \texttt{addMarkers}
+function, one can store specific sets of proteins by using the
+\textit{Features of interest} infrastructure from the
+\Biocpkg{MSnbase} package. If users have specific subsets of proteins
+they wish to highlight in their data (possibly across multiple
+experiments) they would first create a \texttt{FeaturesOfInterest}
+object. For example, if we wanted to highlight a proteins with the
+accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some
+of the proteins known to form part of the 13S condensin complex, we
+would create a first create a \texttt{FeaturesOfInterest} object, and
+subsequently highlight their location on a PCA plot with the
+\texttt{highlightOnPlot} function.
 
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Wed Nov 30 14:38:32 2016 
+##  Created on Wed Nov 30 15:02:28 2016 
 ##  Description:
@@
 \end{knitrout}
-  \caption{Highlighting protein features of interest}
+  \caption{Highlighting protein features of interest.}
   \label{fig:foi}
@@
 
-Users can also create several sets of \texttt{FeaturesOfInterest} object and store them in a \texttt{FoICollection}.
-
-It is also worthly of note that it is possible to search for a specific protein of interest by \texttt{featureNames} or using any identifying information found in the \texttt{fData} columns by using the search box on the \texttt{pRolocVis} application part of the \Biocpkg{pRolocGUI} package (see section on interactive visualisation).
-This can be handy for quickly searching and highlighting proteins on the fly, the disavanatge here is that proteins can only be searched for a one-by-one basis.
+Users can also create several sets of \texttt{FeaturesOfInterest}
+object and store them in a \texttt{FoICollection}.
+
+It is also worthly of note that it is possible to search for a
+specific protein of interest by \texttt{featureNames} or using any
+identifying information found in the \texttt{fData} columns by using
+the search box on the \texttt{pRolocVis} application part of the
+\Biocpkg{pRolocGUI} package (see section on interactive
+visualisation). This can be handy for quickly searching and
+highlighting proteins on the fly, the disavanatge here is that
+proteins can only be searched for a one-by-one basis.
 
@@
 
-% latex table generated in R 3.3.0 by xtable 1.8-2 package
-% Wed Nov 30 14:38:32 2016
+% latex table generated in R 3.3.2 by xtable 1.8-2 package
+% Wed Nov 30 15:02:28 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov 30 14:38:31 2016 
-## Added markers from  'pdres' marker vector. Wed Nov 30 14:38:32 2016 
-##  MSnbase version: 2.0.1
+## Added markers from  'mrk' marker vector. Wed Nov 30 15:02:27 2016 
+## Added markers from  'pdres' marker vector. Wed Nov 30 15:02:28 2016 
+##  MSnbase version: 2.1.4
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4346815               0.3050109               0.3898494 
+##               0.4348219               0.3034081               0.3800417 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6915344               0.6018374               0.4277405 
+##               0.6838374               0.5953475               0.4285470 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4215606               0.6043865               0.9499822 
+##               0.4261188               0.5933032               0.9494300 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7959813               0.7080413               0.3162206 
+##               0.7951832               0.7086165               0.3130830 
 ##         Plasma membrane              Proteasome 
-##               0.7248143               0.4135359
+##               0.7004839               0.4193368
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     171                      88 
+##                      84                     173                      83 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     476                     109 
+##                     294                     473                     101 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      24                     119                     523 
+##                      28                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     230                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     315                     158                    2058
+##                     321                     158                    2057
 \end{verbatim}
@@
 \begin{verbatim}
-## R Under development (unstable) (2015-11-11 r69621)
-## Platform: x86_64-apple-darwin13.4.0 (64-bit)
-## Running under: OS X 10.11.5 (El Capitan)
+## R version 3.3.2 Patched (2016-11-01 r71616)
+## Platform: x86_64-pc-linux-gnu (64-bit)
+## Running under: Ubuntu 14.04.5 LTS
 ## 
 ## locale:
-## [1] en_GB.UTF-8/en_GB.UTF-8/en_GB.UTF-8/C/en_GB.UTF-8/en_GB.UTF-8
+##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
+##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
+##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
+##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
+##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
+## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
 ## 
@@
 ## other attached packages:
-##  [1] pRolocdata_1.12.0    pRoloc_1.14.4        MLInterfaces_1.54.0 
-##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.5        
-##  [7] AnnotationDbi_1.36.0 IRanges_2.5.40       S4Vectors_0.9.45    
-## [10] MSnbase_2.0.1        ProtGenerics_1.6.0   BiocParallel_1.8.1  
-## [13] mzR_2.8.0            Rcpp_0.12.8          Biobase_2.34.0      
+##  [1] pRolocdata_1.12.0    pRoloc_1.15.5        MLInterfaces_1.54.0 
+##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.4        
+##  [7] AnnotationDbi_1.36.0 IRanges_2.8.0        S4Vectors_0.12.0    
+## [10] MSnbase_2.1.4        ProtGenerics_1.6.0   BiocParallel_1.8.1  
+## [13] mzR_2.8.0            Rcpp_0.12.7          Biobase_2.34.0      
 ## [16] BiocGenerics_0.20.0  gridExtra_2.2.1      xtable_1.8-2        
-## [19] BiocStyle_2.2.1      knitr_1.15.1        
+## [19] BiocStyle_2.2.0      knitr_1.14          
 ## 
 ## loaded via a namespace (and not attached):
-##   [1] minqa_1.2.4           colorspace_1.3-1      hwriter_1.3.2        
+##   [1] minqa_1.2.4           colorspace_1.2-7      hwriter_1.3.2        
 ##   [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
@@
 ##  [13] flexmix_2.3-13        mvtnorm_1.0-5         codetools_0.2-15     
-##  [16] splines_3.3.0         doParallel_1.0.10     impute_1.48.0        
+##  [16] splines_3.3.2         doParallel_1.0.10     impute_1.48.0        
 ##  [19] robustbase_0.92-6     jsonlite_1.1          nloptr_1.0.4         
-##  [22] caret_6.0-73          pbkrtest_0.4-6        rda_1.0.2-2          
+##  [22] caret_6.0-72          pbkrtest_0.4-6        rda_1.0.2-2          
 ##  [25] kernlab_0.9-25        vsn_3.42.3            sfsmisc_1.1-0        
 ##  [28] shiny_0.14.2          sampling_2.7          assertthat_0.1       
-##  [31] Matrix_1.2-7.1        lazyeval_0.2.0        limma_3.30.5         
-##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.0          
+##  [31] Matrix_1.2-7.1        limma_3.30.2          formatR_1.4          
+##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.2          
 ##  [37] ggvis_0.4.3           gtable_0.2.0          affy_1.52.0          
@@
 ##  [52] mime_0.5              gtools_3.5.0          dendextend_1.3.0     
-##  [55] DEoptimR_1.0-8        zlibbioc_1.20.0       MASS_7.3-45          
-##  [58] scales_0.4.1          BiocInstaller_1.24.0  pcaMethods_1.66.0    
-##  [61] SparseM_1.74          RColorBrewer_1.1-2    memoise_1.0.0        
-##  [64] ggplot2_2.2.0         biomaRt_2.30.0        rpart_4.1-10         
-##  [67] stringi_1.1.2         RSQLite_1.1           highr_0.6            
-##  [70] genefilter_1.56.0     randomForest_4.6-12   foreach_1.4.3        
-##  [73] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
-##  [76] mzID_1.12.0           evaluate_0.10         lattice_0.20-34      
-##  [79] htmlwidgets_0.8       gbm_2.1.1             plyr_1.8.4           
-##  [82] magrittr_1.5          R6_2.2.0              DBI_0.5-1            
-##  [85] whisker_0.3-2         mgcv_1.8-16           survival_2.40-1      
-##  [88] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
-##  [91] car_2.1-3             mlbench_2.1-1         grid_3.3.0           
+##  [55] DEoptimR_1.0-6        zlibbioc_1.20.0       MASS_7.3-45          
+##  [58] scales_0.4.0          BiocInstaller_1.24.0  pcaMethods_1.66.0    
+##  [61] SparseM_1.72          RColorBrewer_1.1-2    ggplot2_2.1.0        
+##  [64] biomaRt_2.30.0        rpart_4.1-10          stringi_1.1.2        
+##  [67] RSQLite_1.0.0         highr_0.6             genefilter_1.56.0    
+##  [70] randomForest_4.6-12   foreach_1.4.3         e1071_1.6-7          
+##  [73] prabclus_2.2-6        bitops_1.0-6          mzID_1.12.0          
+##  [76] evaluate_0.10         lattice_0.20-34       htmlwidgets_0.7      
+##  [79] gbm_2.1.1             plyr_1.8.4            magrittr_1.5         
+##  [82] R6_2.2.0              DBI_0.5-1             whisker_0.3-2        
+##  [85] mgcv_1.8-15           survival_2.40-1       RCurl_1.95-4.8       
+##  [88] nnet_7.3-12           tibble_1.2            msdata_0.14.0        
+##  [91] car_2.1-3             mlbench_2.1-1         grid_3.3.2           
 ##  [94] FNN_1.1               ModelMetrics_1.1.0    threejs_0.2.2        

### 2016-11-30 bioc-workflow.tex
update author contribs

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:02:25 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:14:03 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:02:25 2016 
-## Subset [6725,20][5032,20] Wed Nov 30 15:02:26 2016 
-## Removed features with more than 0 NAs: Wed Nov 30 15:02:26 2016 
-## Dropped featureData's levels Wed Nov 30 15:02:26 2016 
+## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:14:03 2016 
+## Subset [6725,20][5032,20] Wed Nov 30 15:14:04 2016 
+## Removed features with more than 0 NAs: Wed Nov 30 15:14:04 2016 
+## Dropped featureData's levels Wed Nov 30 15:14:04 2016 
 ##  MSnbase version: 1.21.7
@@
          \hlkwc{pcol} \hlstd{=} \hlstr{"purple"}\hlstd{,} \hlkwc{fractions} \hlstd{=} \hlstr{"Tag"}\hlstd{)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[.data.frame`(pData(object), , .frac): object '.frac' not found}}\begin{alltt}
 \hlkwd{title}\hlstd{(}\hlkwc{main} \hlstd{=} \hlstr{"Marker occupancy profiles along the gradient"}\hlstd{)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in title(main = "{}Marker occupancy profiles along the gradient"{}): plot.new has not been called yet}}\begin{alltt}
 \hlkwd{matlines}\hlstd{(}\hlkwd{t}\hlstd{(}\hlkwd{exprs}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Peroxisome"}\hlstd{, ])),}
          \hlkwc{lty} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"darkblue"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"l"}\hlstd{)}
-\end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot.xy(xy.coords(x, y), type = type, ...): plot.new has not been called yet}}\begin{alltt}
 \hlkwd{legend}\hlstd{(}\hlstr{"topleft"}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"Mitochondrion"}\hlstd{,} \hlstr{"Peroxisome"}\hlstd{),}
@@
 \end{alltt}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in strwidth(legend, units = "{}user"{}, cex = cex, font = text.font): plot.new has not been called yet}}\end{kframe}
+\end{kframe}
+\includegraphics[width=.6\textwidth]{figure/plotDist-1}
+
 \end{knitrout}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Wed Nov 30 15:02:28 2016 
+##  Created on Wed Nov 30 15:14:06 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Wed Nov 30 15:02:28 2016
+% Wed Nov 30 15:14:06 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov 30 15:02:27 2016 
-## Added markers from  'pdres' marker vector. Wed Nov 30 15:02:28 2016 
+## Added markers from  'mrk' marker vector. Wed Nov 30 15:14:05 2016 
+## Added markers from  'pdres' marker vector. Wed Nov 30 15:14:06 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4348219               0.3034081               0.3800417 
+##               0.4366549               0.3041349               0.3799370 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6838374               0.5953475               0.4285470 
+##               0.6880299               0.6002876               0.4310921 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4261188               0.5933032               0.9494300 
+##               0.4165667               0.5997487               0.9494945 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951832               0.7086165               0.3130830 
+##               0.7934108               0.7117905               0.3160580 
 ##         Plasma membrane              Proteasome 
-##               0.7004839               0.4193368
+##               0.7110269               0.4120044
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     173                      83 
+##                      85                     170                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     294                     473                     101 
+##                     296                     480                      96 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      28                     124                     523 
+##                      27                     125                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      39 
+##                     230                     340                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     321                     158                    2057
+##                     322                     158                    2058
 \end{verbatim}
@@
 
-All authors developed the software presented in this workflow, and
-wrote and approved the manuscript.
+LMB and LG developed the software presented in this workflow. All
+authors wrote and approved the manuscript.
 

### 2016-12-01 bioc-workflow.Rnw
update author contribs
Merging lisa's changes

@@
 proteomics data starting from a spreadsheet containing quantitative
-mass spectrometry data. We will focus on a recent pluripotent mouse
+mass spectrometry data, through to some essential data processing
+steps, and finish with different applications of machine learning
+(Figure \ref{fig:overview}). We focus on a recent pluripotent mouse
 embryonic stem cells experiment \cite{hyper}. These data, as well as
@@
 papers \cite{Gatto:2014,Gatto:2015}.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=.5\textwidth]{./Figures/overview.png}
+  \caption{Schematic overview of the pRoloc pipeline from data import, through to data processing, machine learning and data export.}
+  \label{fig:overview}
+\end{figure}
 
@@
 the efficient manipulation and processing of mass spectrometry and
-proteomics data in R. Figure 1 illustrates a simplified view of the
+proteomics data in R. Figure \ref{fig:msnset} illustrates a simplified view of the
 \texttt{MSnSet} structure; there exists 3 key sub-parts (termed slots)
@@
 plot the principal components (PCs) of a dataset against one another,
-by default the first two components are plotted on the x- and y-axis for the \texttt{plot2D} function, and first three components are loaded for the \texttt{plot3D} function,
-respectively (the \texttt{dims} argument can be used to plot other PCs). If
+by default the first two components are plotted on the x- and y-axis
+for the \texttt{plot2D} function, and first three components are
+loaded for the \texttt{plot3D} function, respectively (the
+\texttt{dims} argument can be used to plot other PCs). If
 distinct clusters are observed, we assume that there is organellar
@@
 
-\subsection*{Adding user-defined marker lists}
+\subsection{*Adding user-defined marker lists}
 It is also possible for users to use their own marker list with the
@@
 
-\subsection*{Visualising markers}
-Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
-\texttt{plot2D} function and then use the \texttt{addLegend} function
-to map the marker classes to the pre-defined colours. We also display
-the data along the first and seventh PCs using the \texttt{dims}
-argument. Note that in these calls to the \texttt{plot2D} function, we
-have omitted the \texttt{fcol} argument and use of the default
-\texttt{"markers"} feature variable to annotate the plot. We choose
-to display PCs 1 and 7 to illustrate that while upper principal
-components explain much less variability in the data (2.23\% for PC7,
-as opposed to 48.41\% for PC1), we see that the mitochondrial (purple)
-and peroxisome (dark blue) clusters can be differenciated, despite the
-apparent overlap in the visualisation of the two first PCs
-(Figure \ref{fig:plotmarkers}).
+\subsection{*Visualising markers}
+Having added a the mouse markers to our \texttt{fData} from the
+\texttt{pRolocmarkers} we can now visualise these annotations along
+the PCA plot using the \texttt{plot2D} function and then use the
+\texttt{addLegend} function to map the marker classes to the
+pre-defined colours. We also display the data along the first and
+seventh PCs using the \texttt{dims} argument. Note that in these calls
+to the \texttt{plot2D} function, we have omitted the \texttt{fcol}
+argument and use of the default \texttt{"markers"} feature variable to
+annotate the plot. We choose to display PCs 1 and 7 to illustrate that
+while upper principal components explain much less variability in the
+data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
+mitochondrial (purple) and peroxisome (dark blue) clusters can be
+differenciated, despite the apparent overlap in the visualisation of
+the two first PCs (Figure \ref{fig:plotmarkers}).
 
@@
 \subsection*{Features of interest}
-
 In addition to adding annotation using the \texttt{addMarkers}
-function, one can store specific sets of proteins by using the
+function one can store specific sets of proteins by using the
 \textit{Features of interest} infrastructure from the
-\Biocpkg{MSnbase} package. If users have specific subsets of proteins
-they wish to highlight in their data (possibly across multiple
-experiments) they would first create a \texttt{FeaturesOfInterest}
-object. For example, if we wanted to highlight a proteins with the
-accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some
-of the proteins known to form part of the 13S condensin complex, we
-would create a first create a \texttt{FeaturesOfInterest} object, and
-subsequently highlight their location on a PCA plot with the
-\texttt{highlightOnPlot} function.
+\Biocpkg{MSnbase} package. If users have a specific subset of proteins
+they wish to highlight in their data they would first create a
+\texttt{FeaturesOfInterest} object. For example, if we wanted to
+highlight a proteins with the accession numbers Q8CG48, Q8CG47,
+Q8K2Z4, and Q8C156, which are some of the proteins known to form part
+of the 13S condensin complex, we would create a first create a
+\texttt{FeaturesOfInterest} object then use the function
+\texttt{highlightOnPlot} to highlight their location on a PCA plot.
 
@@
 @  
-  \caption{Highlighting protein features of interest.}
+  \caption{Highlighting protein features of interest}
   \label{fig:foi}
@@
 
-Users can also create several sets of \texttt{FeaturesOfInterest}
-object and store them in a \texttt{FoICollection}.
+Users can also create several sets of \texttt{FeaturesOfInterest} object and store them in a \texttt{FoICollection}.
 
@@
 that in the code chunk below, we mirror the x-axis to represent the
-two figures with the same orientation.
-The interactive "compare" application part of the \Biocpkg{pRolocGUI} package is also useful for examining replicate experiments (see the next section interactive visualisation for details).
+two figures with the same orientation. The interactive "compare"
+application part of the \Biocpkg{pRolocGUI} package is also useful for
+examining replicate experiments (see the next section interactive
+visualisation for details).
 

### 2016-12-01 bioc-workflow.Rnw

@@
 
-\subsection{*Visualising markers}
+\subsection*{Visualising markers}
 Having added a the mouse markers to our \texttt{fData} from the
@@
 In addition to adding annotation using the \texttt{addMarkers}
-function one can store specific sets of proteins by using the
+function, one can store specific sets of proteins by using the
 \textit{Features of interest} infrastructure from the
-\Biocpkg{MSnbase} package. If users have a specific subset of proteins
-they wish to highlight in their data they would first create a
-\texttt{FeaturesOfInterest} object. For example, if we wanted to
-highlight a proteins with the accession numbers Q8CG48, Q8CG47,
-Q8K2Z4, and Q8C156, which are some of the proteins known to form part
-of the 13S condensin complex, we would create a first create a
-\texttt{FeaturesOfInterest} object then use the function
-\texttt{highlightOnPlot} to highlight their location on a PCA plot.
+\Biocpkg{MSnbase} package. If users have specific subsets of proteins
+they wish to highlight in their data (possibly across multiple
+experiments) they would first create a \texttt{FeaturesOfInterest}
+object. For example, if we wanted to highlight a proteins with the
+accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some
+of the proteins known to form part of the 13S condensin complex, we
+would create a first create a \texttt{FeaturesOfInterest} object, and
+subsequently highlight their location on a PCA plot with the
+\texttt{highlightOnPlot} function.
 
@@
 @  
-  \caption{Highlighting protein features of interest}
+  \caption{Highlighting protein features of interest.}
   \label{fig:foi}
@@
 
-Users can also create several sets of \texttt{FeaturesOfInterest} object and store them in a \texttt{FoICollection}.
-
-It is also worthly of note that it is possible to search for a
+Users can also create several sets of \texttt{FeaturesOfInterest}
+object and store them in a \texttt{FoICollection}.
+
+It is also worthy of note that it is possible to search for a
 specific protein of interest by \texttt{featureNames} or using any

### 2016-12-01 bioc-workflow.tex
merge Lisa's latest, with overview fig

@@
 papers \cite{Gatto:2014,Gatto:2015}.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=.5\textwidth]{./Figures/overview.pdf}
+  \caption{Schematic overview of the pRoloc pipeline from data import, through to data processing, machine learning and data export.}
+  \label{fig:overview}
+\end{figure}
 
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:14:03 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:39:06 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Wed Nov 30 15:14:03 2016 
-## Subset [6725,20][5032,20] Wed Nov 30 15:14:04 2016 
-## Removed features with more than 0 NAs: Wed Nov 30 15:14:04 2016 
-## Dropped featureData's levels Wed Nov 30 15:14:04 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:39:06 2016 
+## Subset [6725,20][5032,20] Thu Dec  1 10:39:07 2016 
+## Removed features with more than 0 NAs: Thu Dec  1 10:39:07 2016 
+## Dropped featureData's levels Thu Dec  1 10:39:07 2016 
 ##  MSnbase version: 1.21.7
@@
 \subsection*{Visualising markers}
-Having added a the mouse markers to our \texttt{fData} from the \texttt{pRolocmarkers} we can now visualise these annotations along the PCA plot using the
-\texttt{plot2D} function and then use the \texttt{addLegend} function
-to map the marker classes to the pre-defined colours. We also display
-the data along the first and seventh PCs using the \texttt{dims}
-argument. Note that in these calls to the \texttt{plot2D} function, we
-have omitted the \texttt{fcol} argument and use of the default
-\texttt{"markers"} feature variable to annotate the plot. We choose
-to display PCs 1 and 7 to illustrate that while upper principal
-components explain much less variability in the data (2.23\% for PC7,
-as opposed to 48.41\% for PC1), we see that the mitochondrial (purple)
-and peroxisome (dark blue) clusters can be differenciated, despite the
-apparent overlap in the visualisation of the two first PCs
-(Figure \ref{fig:plotmarkers}).
+Having added a the mouse markers to our \texttt{fData} from the
+\texttt{pRolocmarkers} we can now visualise these annotations along
+the PCA plot using the \texttt{plot2D} function and then use the
+\texttt{addLegend} function to map the marker classes to the
+pre-defined colours. We also display the data along the first and
+seventh PCs using the \texttt{dims} argument. Note that in these calls
+to the \texttt{plot2D} function, we have omitted the \texttt{fcol}
+argument and use of the default \texttt{"markers"} feature variable to
+annotate the plot. We choose to display PCs 1 and 7 to illustrate that
+while upper principal components explain much less variability in the
+data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
+mitochondrial (purple) and peroxisome (dark blue) clusters can be
+differenciated, despite the apparent overlap in the visualisation of
+the two first PCs (Figure \ref{fig:plotmarkers}).
 
@@
 \subsection*{Features of interest}
-
 In addition to adding annotation using the \texttt{addMarkers}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Wed Nov 30 15:14:06 2016 
+##  Created on Thu Dec  1 10:39:09 2016 
 ##  Description:
@@
 
-It is also worthly of note that it is possible to search for a
+It is also worthy of note that it is possible to search for a
 specific protein of interest by \texttt{featureNames} or using any
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Wed Nov 30 15:14:06 2016
+% Thu Dec  1 10:39:09 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Nov 30 15:14:05 2016 
-## Added markers from  'pdres' marker vector. Wed Nov 30 15:14:06 2016 
+## Added markers from  'mrk' marker vector. Thu Dec  1 10:39:08 2016 
+## Added markers from  'pdres' marker vector. Thu Dec  1 10:39:09 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4366549               0.3041349               0.3799370 
+##               0.4350790               0.3036502               0.3798695 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6880299               0.6002876               0.4310921 
+##               0.6854737               0.6132301               0.4278794 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4165667               0.5997487               0.9494945 
+##               0.4176288               0.6005527               0.9494418 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7934108               0.7117905               0.3160580 
+##               0.7949941               0.7094819               0.3150650 
 ##         Plasma membrane              Proteasome 
-##               0.7110269               0.4120044
+##               0.7117859               0.4212111
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      84 
+##                      85                     171                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     480                      96 
+##                     299                     467                     103 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     125                     523 
+##                      27                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     340                      38 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     158                    2058
+##                     323                     156                    2059
 \end{verbatim}

### 2016-12-01 bioc-workflow.tex
update caption plot2D/3D

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:39:06 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:49:34 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:39:06 2016 
-## Subset [6725,20][5032,20] Thu Dec  1 10:39:07 2016 
-## Removed features with more than 0 NAs: Thu Dec  1 10:39:07 2016 
-## Dropped featureData's levels Thu Dec  1 10:39:07 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:49:34 2016 
+## Subset [6725,20][5032,20] Thu Dec  1 10:49:35 2016 
+## Removed features with more than 0 NAs: Thu Dec  1 10:49:35 2016 
+## Dropped featureData's levels Thu Dec  1 10:49:35 2016 
 ##  MSnbase version: 1.21.7
@@
 \end{knitrout}
-  \caption{Annotated PCA plots of the \texttt{hl} dataset.}
+\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
+  with \texttt{plt2D} and \texttt{plot3D} (right).}
   \label{fig:plotmarkers}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Thu Dec  1 10:39:09 2016 
+##  Created on Thu Dec  1 10:49:37 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Thu Dec  1 10:39:09 2016
+% Thu Dec  1 10:49:38 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Dec  1 10:39:08 2016 
-## Added markers from  'pdres' marker vector. Thu Dec  1 10:39:09 2016 
+## Added markers from  'mrk' marker vector. Thu Dec  1 10:49:37 2016 
+## Added markers from  'pdres' marker vector. Thu Dec  1 10:49:38 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4350790               0.3036502               0.3798695 
+##               0.4343505               0.3047508               0.3888373 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6854737               0.6132301               0.4278794 
+##               0.6858101               0.5897977               0.4283659 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4176288               0.6005527               0.9494418 
+##               0.4175965               0.5981079               0.9494104 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7949941               0.7094819               0.3150650 
+##               0.7967111               0.7085718               0.3154399 
 ##         Plasma membrane              Proteasome 
-##               0.7117859               0.4212111
+##               0.7107861               0.4112064
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      84 
+##                      85                     171                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     299                     467                     103 
+##                     296                     473                      95 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     523 
+##                      26                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     229                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     323                     156                    2059
+##                     323                     158                    2058
 \end{verbatim}

### 2016-12-02 abstract.tex
updates based on Claire's comments - up to p10

@@
 the software infrastructure at hand, importing and processing data,
-quality control, sub-cellular marker definition and visualisation. We
-then demonstrate the application and interpretation of statistical
-learning methods, including novelty detection using semi-supervised
-learning, classification, clustering and transfer learning and
-conclude the pipeline with data export. The workflow is aimed at
-beginners who are familiar with proteomics in general and spatial
-proteomics in particular.
+quality control, sub-cellular marker definition, visualisation and
+interactive exploration. We then demonstrate the application and
+interpretation of statistical learning methods, including novelty
+detection using semi-supervised learning, classification, clustering
+and transfer learning and conclude the pipeline with data export. The
+workflow is aimed at beginners who are familiar with proteomics in
+general and spatial proteomics in particular.

### 2016-12-02 bioc-workflow.Rnw
merge Lisa's latest, with overview fig
update caption plot2D/3D
updates based on Claire's comments - up to p10

@@
 Multiple research groups have described a variety of approaches to
-establish high quality proteome-wide datasets. However, data analysis
-is as critical as data production for reliable and insightful
-biological interpretation. Here, we walk the reader through a typical
-pipeline for the analysis of such data using several Bioconductor
-packages for the R statistical programming environment.
+establish high quality proteome-wide datasets (see for example
+\cite{Gatto2010} for a review, and \cite{hyper,Itzhak:2016} for recent
+examples). However, data analysis is as critical as data production
+for reliable and insightful biological interpretation. Here, we walk
+the reader through a typical pipeline for the analysis of such data
+using several Bioconductor packages for the R statistical programming
+environment.
 
@@
 analysis of such data. \Biocpkg{pRoloc} itself relies on
-\Biocpkg{MSnbase} to manipulate and process quantitative
-proteomics data. Many other packages are used by \Biocpkg{pRoloc}
-for clustering, classification and visualisation.
+\Biocpkg{MSnbase} to manipulate and process quantitative proteomics
+data. Many other packages are used by \Biocpkg{pRoloc} for clustering,
+classification and visualisation. Support for interactive
+visualisation is offered by the \texttt{pRolocGUI} package.
 
@@
 \href{http://bioconductor.org/install/#install-bioconductor-packages}{Bioconductor
-  installation help page}. This procedure is also applicable to any
-packages, from \href{https://cran.r-project.org/}{CRAN} as well as
-GitHub. Once a package has been installed, it needs to be loaded for
-its functionality to become available in the R session; this is done
-with the \texttt{library} function e.g.  to load the
-\Biocpkg{pRoloc} one would type \texttt{library("pRoloc")}
-after installation.
+  installation help page}. Below, we show how to install the four main
+packages used in this workflow:
+
+<<install, eval=FALSE>>=
+source("https://bioconductor.org/biocLite.R")
+biocLite(c("MSnase", "pRoloc", "pRolocdata", "pRolocGUI"))
+@
+
+This procedure is also applicable to any packages, from
+\href{https://cran.r-project.org/}{CRAN} as well as GitHub. Once a
+package has been installed, it needs to be loaded for its
+functionality to become available in the R session; this is done with
+the \texttt{library} function e.g.  to load the \Biocpkg{pRoloc} one
+would type \texttt{library("pRoloc")} after installation.
 
@@
   \centering
-  \includegraphics[width=.5\textwidth]{./Figures/overview.png}
+  \includegraphics[width=.5\textwidth]{./Figures/overview.pdf}
   \caption{Schematic overview of the pRoloc pipeline from data import, through to data processing, machine learning and data export.}
@@
 a 20-plex dataset for the analysis.  As discussed in the manuscript
-\cite{hyper}, it has been shown that combining replicates across from
+\cite{hyper}, it has been shown that combining replicates from
 different gradients can increase spatial resolution
@@
 These, as well as many other data are directly available as properly
-structured and annotated computational objects from the 
-\Biocexptpkg{pRolocdata} experiment package. In this workflow, we
-will start with a description of how to generate these ad hoc objects
-starting from an arbitrary spreadsheets, as produced by many popular
-third-party applications.
+structured and annotated datasets from the \Biocexptpkg{pRolocdata}
+experiment package. In this workflow, we will start with a description
+of how to generate these ad hoc data objects starting from an
+arbitrary spreadsheet, as produced by many popular third-party
+applications.
 
@@
 function \texttt{readMSnSet2}. The function takes a single spreadsheet
-filename as input and extracts the columns containing the quantitation
-data, as identified by the argument \texttt{ecol}, to create the
-expression data, while the other columns in the spreadsheet are
-appended to the feature meta-data slot.  By example, in the code chunk
-below we read in the \texttt{csv} spreadsheet containing the
+file name as input and extracts the columns containing the
+quantitation data, as identified by the argument \texttt{ecol}, to
+create the expression data, while the other columns in the spreadsheet
+are appended to the feature meta-data slot.  By example, in the code
+chunk below we read in the \texttt{csv} spreadsheet containing the
 quantitation data from the intersect of replicates 1 and 2 of the
@@
 
-To use the \texttt{readMSnSet2} function, as a minimum one must specify the
-file path to the data and which columns of the spreadsheet contain
-quantitation data. The \texttt{getEcols} function exists to help users
-identify which columns of the spreadsheet contain the
-quantitation data. In the last line of the code chunk below, we print
-the file name (not the full path, which will vary from computer to
-computer).
+To use the \texttt{readMSnSet2} function, as a minimum one must
+specify the file path to the data and which columns of the spreadsheet
+contain quantitation data. In the code chunk below, we start by
+identifying the file that we want to use. The \texttt{system.file}
+function is used to return the path to the \texttt{extdata} directory
+from the \texttt{pRolocdata} package, which is where our file of
+interest resides. We then use the \texttt{dir} function to list the
+content of that directory and store the path that matches the file
+name of interest in the \texttt{csvfile}. Note that these two lines
+are only needed here to locate a file in a package; in a more general
+use case, the user would defined the \texttt{csvfile} variable
+containing the file name of interest directly.
+
+A common pitfall here is to provide only the file name, rather than
+full path to the file (which is what is shown below with
+\texttt{basename}; we don't print the full path, as it will vary from
+computer to computer). Note that only specifying the name of the file
+is sufficient when it exists in the working directory (i.e. the
+directory in which R is running, which can be queried and changed with
+the \texttt{getwd} and \texttt{setwd} functions respectively).
 
@@
 
-The spreadsheet that was deposited by the authors contains two
-headers, with the second header containing information about where the
-quantitation data is stored.
+Next, we need to identify which columns in the spreadsheet contain the
+quantitation data. This can be done using the \texttt{getEcols}
+function inside R. The spreadsheet that was deposited by the authors
+contains two headers, with the second header containing information
+about where the quantitation data is stored.
 
@@
 We can display the names of the second header by calling the
-\texttt{getEcols} function with the argument \texttt{n = 2} (the default value is \texttt{n
-= 1}), to specify that we wish to display the column names of the
-second line.
+\texttt{getEcols} function with the argument \texttt{n = 2} (the
+default value is \texttt{n = 1}), to specify that we wish to display
+the column names of the second line. We also specify the name of the
+spreadsheet file (defined as \texttt{csvfile} above) and the
+separator that splits cells.
 
@@
 
-Below, we examine the quantitative information for first 5 proteins.
-It is also possible to access specific rows and columns by naming the
-proteins and TMT tag channels of interest.
+Below, we examine the quantitative information along the whole
+gradient for first 5 proteins.  It is also possible to access specific
+rows and columns by naming the proteins and TMT tag channels of
+interest.
 
@@
 fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
-fData(hl)[1:4, 1:6]
-@
-
-Note that when using the simple \texttt{readMSnSet2} procedure, the \texttt{pData}
-slot which is used to store information about the samples/channels
-is kept empty. It is advised to annotate the channels as
-well. Below, we annotate the replicate from which the profiles
-originate and the TMT tag (extracted from the sample/channel names).
+## feature vars 1, 2, and 4 to 6
+fData(hl)[1:4, c(1:2, 4:6)]
+@
+
+Note that when using the simple \texttt{readMSnSet2} procedure, the
+\texttt{pData} slot which is used to store information about the
+samples/channels is kept empty. As illustrated below, one can use the
+\texttt{\$} operator to access (or create) individual columns in the
+metadata slot. It is advised to annotate the channels as well. Below,
+we annotate the replicate from which the profiles originate and the
+TMT tag (extracted from the sample/channel names). To do so, we use
+the sample names that were assigned automatically using the
+quantiation column names and remove leading \texttt{X} and trailing
+\texttt{.1} using the \texttt{sub} function.
 
@@
 are found in the \texttt{exprs} (expression data) slot as channels (short for TMT channels).
-In the frame of LOPIT and hyperLOPIT these channels consititue the
+In the frame of LOPIT and hyperLOPIT these channels constitute the
 relative abundance of each protein (along the rows) in the channel of
@@
 to spatial proteomics data processing. The first one focuses on
-reducing purely technical variation between channels without
-affecting biological variability (i.e. the shape of the quantitative
+reducing purely technical variation between channels without affecting
+biological variability (i.e. the shape of the quantitative
 profiles). This normalisation will depend on the underlying
@@
 profiles into a same intensity interval (typically 0 and 1) by, for
-example, diving each intensity by the sum of the intensities for that
-quantitative feature. This is not necessary in this example as the
-intensities for each replicate have already been re-scaled to 1 in
-Proteome Discoverer. However, if one wanted to do this they would
-execute the \texttt{normalise} function as demonstrated in the below code
-chunk.
+example, dividing each intensity by the sum of the intensities for
+that quantitative feature. This is not necessary in this example as
+the intensities for each replicate have already been re-scaled to 1 in
+Proteome Discoverer v1.4, Thermo Fisher. However, if one wanted to do
+this they would execute the \texttt{normalise} function as
+demonstrated in the below code chunk.
 
@@
 
-The same \texttt{normalise} function (or \texttt{normalize}, both spellings are
-supported) can also be applied in the first case described above.
-Different normalisation methods such as mean or median scaling,
-variance stabilisation or quantile normalisation, to cite a few, can
-be applied to accomodate different needs.
+The same \texttt{normalise} function (or \texttt{normalize}, both
+spellings are supported) can also be applied in the first case
+described above.  Different normalisation methods, such as mean or
+median scaling, variance stabilisation or quantile normalisation, to
+cite a few, can be applied to accomodate different needs (see
+\texttt{?normalise} for available options).
 
@@
 
-In addition, to matching names, the content of the feature metadata
-for identical feature annotations must match exactly across the data
-to be combined. In particular for these data, we expect the same
-proteins in each replicates to be annotated with the same UniProt
-entry names and descriptions, but not with the same coverage of number
-of peptides or peptide-spectrum matches (PSMs).
+In addition to matching names, the content of the feature metadata for
+identical feature annotations must match exactly across the data to be
+combined. In particular for these data, we expect the same proteins in
+each replicate to be annotated with the same UniProt entry names and
+descriptions, but not with the same coverage of number of peptides or
+peptide-spectrum matches (PSMs).
 
@@
 Below, we update the replicate specific feature variable names and
-remove the shared annotation.
+remove the shared annotation. In the first line, we update only the
+feature variable names 3 to 5 (by appending a \texttt{1}) and in the
+second line, we apply the \texttt{updateFvarLabels} function to update
+all feature variable names (by appending a \texttt{2}).
 
@@
 fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
-fvarLabels(hyperLOPIT2015ms3r2)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r2)[3:5], 2)
+hyperLOPIT2015ms3r2 <- updateFvarLabels(hyperLOPIT2015ms3r2, "2", sep = "")
 fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5]
@@
 
-More details above combining data are given in the dedicated
+More details about combining data are given in the dedicated
 \textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
@@
 experiments. This is done with the \texttt{filterNA} function that
-removes features (proteins) that contain more than a certain proportion (default
-is 0) missing values.
+removes features (proteins) that contain more than a certain
+proportion (default is 0) missing values. The \textit{Processing
+  information} section summarises how combining and filtering missing
+values (subsetting) changed the dimensions of the data.
 
@@
 for the protein names, one for the corresponding sub-cellular marker
-annotation) and pass the vector or filename respectively to the
+annotation) and pass the vector or file name respectively to the
 function. As previously mentioned, the protein names of these markers
@@
 @  
-  \caption{Annotated PCA plots of the \texttt{hl} dataset.}
+\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
+  with \texttt{plt2D} and \texttt{plot3D} (right).}
   \label{fig:plotmarkers}
@@
 
+We also recommend that users regularly update the packages as well as
+the R itself. This can be done with the \texttt{biocLite} function.
+
+<<update, eval=FALSE>>=
+library("BiocInstaller")
+biocLite()
+@
+
 It is always important to include session information details along

### 2016-12-02 bioc-workflow.tex
updates based on Claire's comments - up to p10

@@
 \href{http://bioconductor.org/install/#install-bioconductor-packages}{Bioconductor
-  installation help page}. This procedure is also applicable to any
-packages, from \href{https://cran.r-project.org/}{CRAN} as well as
-GitHub. Once a package has been installed, it needs to be loaded for
-its functionality to become available in the R session; this is done
-with the \texttt{library} function e.g.  to load the
-\Biocpkg{pRoloc} one would type \texttt{library("pRoloc")}
-after installation.
+  installation help page}. Below, we show how to install the four main
+packages used in this workflow:
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{source}\hlstd{(}\hlstr{"https://bioconductor.org/biocLite.R"}\hlstd{)}
+\hlkwd{biocLite}\hlstd{(}\hlkwd{c}\hlstd{(}\hlstr{"MSnase"}\hlstd{,} \hlstr{"pRoloc"}\hlstd{,} \hlstr{"pRolocdata"}\hlstd{,} \hlstr{"pRolocGUI"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+This procedure is also applicable to any packages, from
+\href{https://cran.r-project.org/}{CRAN} as well as GitHub. Once a
+package has been installed, it needs to be loaded for its
+functionality to become available in the R session; this is done with
+the \texttt{library} function e.g.  to load the \Biocpkg{pRoloc} one
+would type \texttt{library("pRoloc")} after installation.
 
@@
 \hlkwd{fvarLabels}\hlstd{(hl)[}\hlnum{4}\hlopt{:}\hlnum{6}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlstr{"peptides.expt"}\hlstd{,} \hlnum{1}\hlopt{:}\hlnum{3}\hlstd{)}
-\hlkwd{fData}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{4}\hlstd{,} \hlnum{1}\hlopt{:}\hlnum{6}\hlstd{]}
-\end{alltt}
-\begin{verbatim}
-##          uniprot.accession  uniprot.id
-## Q9JHU4              Q9JHU4 DYHC1_MOUSE
-## Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE
-## Q9ERU9              Q9ERU9  RBP2_MOUSE
-## P26039              P26039  TLN1_MOUSE
-##                                                                                                                                 description
-## Q9JHU4                                              Cytoplasmic dynein 1 heavy chain 1 OS=Mus musculus GN=Dync1h1 PE=1 SV=2 - [DYHC1_MOUSE]
-## Q9QXS1-3 Isoform PLEC-1 of Plectin OS=Mus musculus GN=Plec - [PLEC_MOUSE]|Isoform PLEC-1A of Plectin OS=Mus musculus GN=Plec - [PLEC_MOUSE]
-## Q9ERU9                                                     E3 SUMO-protein ligase RanBP2 OS=Mus musculus GN=Ranbp2 PE=1 SV=2 - [RBP2_MOUSE]
-## P26039                                                                             Talin-1 OS=Mus musculus GN=Tln1 PE=1 SV=2 - [TLN1_MOUSE]
-##          peptides.expt1 peptides.expt2 peptides.expt3
-## Q9JHU4              175            166            322
-## Q9QXS1-3            123            150            174
-## Q9ERU9              101             90            181
-## P26039              101             94            167
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-Note that when using the simple \texttt{readMSnSet2} procedure, the \texttt{pData}
-slot which is used to store information about the samples/channels
-is kept empty. It is advised to annotate the channels as
-well. Below, we annotate the replicate from which the profiles
-originate and the TMT tag (extracted from the sample/channel names).
+\hlcom{## feature vars 1, 2, and 4 to 6}
+\hlkwd{fData}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{4}\hlstd{,} \hlkwd{c}\hlstd{(}\hlnum{1}\hlopt{:}\hlnum{2}\hlstd{,} \hlnum{4}\hlopt{:}\hlnum{6}\hlstd{)]}
+\end{alltt}
+\begin{verbatim}
+##          uniprot.accession  uniprot.id peptides.expt1 peptides.expt2
+## Q9JHU4              Q9JHU4 DYHC1_MOUSE            175            166
+## Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE            123            150
+## Q9ERU9              Q9ERU9  RBP2_MOUSE            101             90
+## P26039              P26039  TLN1_MOUSE            101             94
+##          peptides.expt3
+## Q9JHU4              322
+## Q9QXS1-3            174
+## Q9ERU9              181
+## P26039              167
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+Note that when using the simple \texttt{readMSnSet2} procedure, the
+\texttt{pData} slot which is used to store information about the
+samples/channels is kept empty. As illustrated below, one can use the
+\texttt{\$} operator to access (or create) individual columns in the
+metadata slot. It is advised to annotate the channels as well. Below,
+we annotate the replicate from which the profiles originate and the
+TMT tag (extracted from the sample/channel names). To do so, we use
+the sample names that were assigned automatically using the
+quantiation column names and remove leading \texttt{X} and trailing
+\texttt{.1} using the \texttt{sub} function.
 
@@
 \hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{],} \hlnum{1}\hlstd{)}
-\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{],} \hlnum{2}\hlstd{)}
+\hlstd{hyperLOPIT2015ms3r2} \hlkwb{<-} \hlkwd{updateFvarLabels}\hlstd{(hyperLOPIT2015ms3r2,} \hlstr{"2"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
 \hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r1)} \hlkwb{<-} \hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{1}\hlopt{:}\hlnum{5}\hlstd{]}
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:49:34 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:06:03 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Dec  1 10:49:34 2016 
-## Subset [6725,20][5032,20] Thu Dec  1 10:49:35 2016 
-## Removed features with more than 0 NAs: Thu Dec  1 10:49:35 2016 
-## Dropped featureData's levels Thu Dec  1 10:49:35 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:06:03 2016 
+## Subset [6725,20][5032,20] Fri Dec  2 17:06:04 2016 
+## Removed features with more than 0 NAs: Fri Dec  2 17:06:04 2016 
+## Dropped featureData's levels Fri Dec  2 17:06:04 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Thu Dec  1 10:49:37 2016 
+##  Created on Fri Dec  2 17:06:06 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Thu Dec  1 10:49:38 2016
+% Fri Dec  2 17:06:07 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Dec  1 10:49:37 2016 
-## Added markers from  'pdres' marker vector. Thu Dec  1 10:49:38 2016 
+## Added markers from  'mrk' marker vector. Fri Dec  2 17:06:05 2016 
+## Added markers from  'pdres' marker vector. Fri Dec  2 17:06:07 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4343505               0.3047508               0.3888373 
+##               0.4350585               0.3020114               0.3827713 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6858101               0.5897977               0.4283659 
+##               0.6916987               0.5981016               0.4325049 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4175965               0.5981079               0.9494104 
+##               0.4337431               0.5912701               0.9497307 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7967111               0.7085718               0.3154399 
+##               0.7942905               0.7137491               0.3148182 
 ##         Plasma membrane              Proteasome 
-##               0.7107861               0.4112064
+##               0.7173760               0.4123604
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      88 
+##                      86                     169                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     473                      95 
+##                     297                     477                      98 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     124                     523 
+##                      28                     123                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     344                      39 
+##                     230                     342                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     323                     158                    2058
+##                     318                     158                    2057
 \end{verbatim}
@@
 ## [16] BiocGenerics_0.20.0  gridExtra_2.2.1      xtable_1.8-2        
-## [19] BiocStyle_2.2.0      knitr_1.14          
+## [19] BiocStyle_2.2.1      knitr_1.14          
 ## 
 ## loaded via a namespace (and not attached):
-##   [1] minqa_1.2.4           colorspace_1.2-7      hwriter_1.3.2        
+##   [1] minqa_1.2.4           colorspace_1.3-1      hwriter_1.3.2        
 ##   [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
@@
 
+We also recommend that users regularly update the packages as well as
+the R itself. This can be done with the \texttt{biocLite} function.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"BiocInstaller"}\hlstd{)}
+\hlkwd{biocLite}\hlstd{()}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
 It is always important to include session information details along

### 2016-12-02 bioc-workflow.Rnw
updates based on Claire's comments - up to p13

@@
 of the column intensities.  In the two plots displayed on figure
-\ref{fig:qcbx}, we re-order the TMT channles to pair corresponding
+\ref{fig:qcbx}, we re-order the TMT channels to pair corresponding
 channels in the two replicates (rather than ordering the channels by
@@
 <<qcbx, out.width="\\textwidth", fig.asp=1/2>>=
-par(mfrow = c(1, 2))
+par(mfrow = c(1, 2)) ## creates a two-panel figure
 o <- order(hl$Iodixonal.Density)
@@
 mitochondrial (purple) and peroxisome (dark blue) clusters can be
-differenciated, despite the apparent overlap in the visualisation of
+differentiated, despite the apparent overlap in the visualisation of
 the two first PCs (Figure \ref{fig:plotmarkers}).
@@
 
-The defualt colours for plotting have been defined so as to enable to
-differenciate up to 30 classes. If more are provided, different
+The default colours for plotting have been defined so as to enable the
+differentiation of up to 30 classes. If more are provided, different
 character symbols (circles, squares, ... and empty and solid symbols)
@@
 
-As demonstrated in \cite{hyper} and illustrated in the PCA plot above,
-the Golgi apparatus proteins (dark brown) display a dynamic pattern,
-noting sets of Golgi marker proteins that are distributed amongst
-other subcellular structures, an observation supported by microscopy.
-As such, we are going to reset the annotation of Golgi markers to
-unknown using the \texttt{fDataTounknown} function. It is often used
-to replace empty strings (\texttt{""}) or missing values in the
-markers definition to a common definition of \textit{unknown}
-localisation.
+As demonstrated in \cite{hyper} and illustrated in the PCA plot
+(figure \ref{fig:plotmarkers}), the Golgi apparatus proteins (dark
+brown) display a dynamic pattern, noting sets of Golgi marker proteins
+that are distributed amongst other subcellular structures, an
+observation supported by microscopy.  As such, we are going to reset
+the annotation of Golgi markers to unknown using the
+\texttt{fDataToUnknown} function. It is often used to replace empty
+strings (\texttt{""}) or missing values in the markers definition to a
+common definition of \textit{unknown} localisation.
 

### 2016-12-02 bioc-workflow.tex
updates based on Claire's comments - up to p13

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:06:03 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:19:06 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:06:03 2016 
-## Subset [6725,20][5032,20] Fri Dec  2 17:06:04 2016 
-## Removed features with more than 0 NAs: Fri Dec  2 17:06:04 2016 
-## Dropped featureData's levels Fri Dec  2 17:06:04 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:19:06 2016 
+## Subset [6725,20][5032,20] Fri Dec  2 17:19:07 2016 
+## Removed features with more than 0 NAs: Fri Dec  2 17:19:07 2016 
+## Dropped featureData's levels Fri Dec  2 17:19:07 2016 
 ##  MSnbase version: 1.21.7
@@
 \begin{alltt}
-\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))} \hlcom{## creates a two-panel figure}
 \hlstd{o} \hlkwb{<-} \hlkwd{order}\hlstd{(hl}\hlopt{$}\hlstd{Iodixonal.Density)}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Fri Dec  2 17:06:06 2016 
+##  Created on Fri Dec  2 17:19:09 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Dec  2 17:06:07 2016
+% Fri Dec  2 17:19:09 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Dec  2 17:06:05 2016 
-## Added markers from  'pdres' marker vector. Fri Dec  2 17:06:07 2016 
+## Added markers from  'mrk' marker vector. Fri Dec  2 17:19:08 2016 
+## Added markers from  'pdres' marker vector. Fri Dec  2 17:19:09 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4350585               0.3020114               0.3827713 
+##               0.4368179               0.3037653               0.3856993 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6916987               0.5981016               0.4325049 
+##               0.6916475               0.5951696               0.4377555 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4337431               0.5912701               0.9497307 
+##               0.4317514               0.5919750               0.9527663 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7942905               0.7137491               0.3148182 
+##               0.7947663               0.7096575               0.3269003 
 ##         Plasma membrane              Proteasome 
-##               0.7173760               0.4123604
+##               0.7210996               0.4155492
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     169                      88 
+##                      85                     172                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     297                     477                      98 
+##                     296                     477                     100 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      28                     123                     522 
+##                      27                     123                     513 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     342                      39 
+##                     231                     344                      42 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     318                     158                    2057
+##                     318                     157                    2059
 \end{verbatim}
@@
 ##  [1] pRolocdata_1.12.0    pRoloc_1.15.5        MLInterfaces_1.54.0 
-##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.4        
-##  [7] AnnotationDbi_1.36.0 IRanges_2.8.0        S4Vectors_0.12.0    
+##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.5        
+##  [7] AnnotationDbi_1.36.0 IRanges_2.8.1        S4Vectors_0.12.1    
 ## [10] MSnbase_2.1.4        ProtGenerics_1.6.0   BiocParallel_1.8.1  
-## [13] mzR_2.8.0            Rcpp_0.12.7          Biobase_2.34.0      
+## [13] mzR_2.8.0            Rcpp_0.12.8          Biobase_2.34.0      
 ## [16] BiocGenerics_0.20.0  gridExtra_2.2.1      xtable_1.8-2        
-## [19] BiocStyle_2.2.1      knitr_1.14          
+## [19] BiocStyle_2.2.1      knitr_1.15.1        
 ## 
@@
 ##  [19] robustbase_0.92-6     jsonlite_1.1          nloptr_1.0.4         
-##  [22] caret_6.0-72          pbkrtest_0.4-6        rda_1.0.2-2          
+##  [22] caret_6.0-73          pbkrtest_0.4-6        rda_1.0.2-2          
 ##  [25] kernlab_0.9-25        vsn_3.42.3            sfsmisc_1.1-0        
 ##  [28] shiny_0.14.2          sampling_2.7          assertthat_0.1       
-##  [31] Matrix_1.2-7.1        limma_3.30.2          formatR_1.4          
+##  [31] Matrix_1.2-7.1        lazyeval_0.2.0        limma_3.30.6         
 ##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.2          
@@
 ##  [52] mime_0.5              gtools_3.5.0          dendextend_1.3.0     
-##  [55] DEoptimR_1.0-6        zlibbioc_1.20.0       MASS_7.3-45          
-##  [58] scales_0.4.0          BiocInstaller_1.24.0  pcaMethods_1.66.0    
-##  [61] SparseM_1.72          RColorBrewer_1.1-2    ggplot2_2.1.0        
-##  [64] biomaRt_2.30.0        rpart_4.1-10          stringi_1.1.2        
-##  [67] RSQLite_1.0.0         highr_0.6             genefilter_1.56.0    
-##  [70] randomForest_4.6-12   foreach_1.4.3         e1071_1.6-7          
-##  [73] prabclus_2.2-6        bitops_1.0-6          mzID_1.12.0          
-##  [76] evaluate_0.10         lattice_0.20-34       htmlwidgets_0.7      
-##  [79] gbm_2.1.1             plyr_1.8.4            magrittr_1.5         
-##  [82] R6_2.2.0              DBI_0.5-1             whisker_0.3-2        
-##  [85] mgcv_1.8-15           survival_2.40-1       RCurl_1.95-4.8       
-##  [88] nnet_7.3-12           tibble_1.2            msdata_0.14.0        
-##  [91] car_2.1-3             mlbench_2.1-1         grid_3.3.2           
-##  [94] FNN_1.1               ModelMetrics_1.1.0    threejs_0.2.2        
-##  [97] digest_0.6.10         diptest_0.75-7        httpuv_1.3.3         
-## [100] munsell_0.4.3
+##  [55] DEoptimR_1.0-8        zlibbioc_1.20.0       MASS_7.3-45          
+##  [58] scales_0.4.1          BiocInstaller_1.24.0  pcaMethods_1.66.0    
+##  [61] SparseM_1.74          RColorBrewer_1.1-2    memoise_1.0.0        
+##  [64] ggplot2_2.2.0         biomaRt_2.30.0        rpart_4.1-10         
+##  [67] stringi_1.1.2         RSQLite_1.1           highr_0.6            
+##  [70] genefilter_1.56.0     randomForest_4.6-12   foreach_1.4.3        
+##  [73] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
+##  [76] mzID_1.12.0           evaluate_0.10         lattice_0.20-34      
+##  [79] htmlwidgets_0.8       gbm_2.1.1             plyr_1.8.4           
+##  [82] magrittr_1.5          R6_2.2.0              DBI_0.5-1            
+##  [85] whisker_0.3-2         mgcv_1.8-16           survival_2.40-1      
+##  [88] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
+##  [91] msdata_0.14.0         car_2.1-3             mlbench_2.1-1        
+##  [94] grid_3.3.2            FNN_1.1               ModelMetrics_1.1.0   
+##  [97] threejs_0.2.2         digest_0.6.10         diptest_0.75-7       
+## [100] httpuv_1.3.3          munsell_0.4.3
 \end{verbatim}

### 2016-12-02 bioc-workflow.Rnw
updates based on Claire's comments

@@
 As demonstrated in \cite{hyper} and illustrated in the PCA plot
-(figure \ref{fig:plotmarkers}), the Golgi apparatus proteins (dark
+(Figure \ref{fig:plotmarkers}), the Golgi apparatus proteins (dark
 brown) display a dynamic pattern, noting sets of Golgi marker proteins
@@
 \end{figure}
+
+In addition, the reproducibility can be assessed by performing
+independent classification analyses on each replicate (see the section
+on \textit{Supervised machine learning} below) and comparing the the
+results. Even when the gradient conditions different (for unexpected
+technical or voluntary reasons, to maximise resolution when combining
+experiments \cite{Trotter:2010}), one expects agreement in the most
+confident organelle assignments.
 

### 2016-12-02 bioc-workflow.Rnw
updates based on Claire's comments

@@
 establish high quality proteome-wide datasets (see for example
-\cite{Gatto2010} for a review, and \cite{hyper,Itzhak:2016} for recent
+\cite{Gatto:2010} for a review, and \cite{hyper,Itzhak:2016} for recent
 examples). However, data analysis is as critical as data production
@@
 classification and visualisation. Support for interactive
-visualisation is offered by the \texttt{pRolocGUI} package.
+visualisation is offered by the \Biocpkg{pRolocGUI} package.
 
@@
 
-\subsection*{The infrastructure: \texttt{pRoloc} and \texttt{MSnbase} packages}
+\subsection*{The infrastructure: \Biocpkg{pRoloc} and \Biocpkg{MSnbase} packages}
 
@@
 function is used to return the path to the \texttt{extdata} directory
-from the \texttt{pRolocdata} package, which is where our file of
+from the \Biocexptpkg{pRolocdata} package, which is where our file of
 interest resides. We then use the \texttt{dir} function to list the
@@
 different GUI's (\textit{main} (default), \textit{compare} or
-\textit{compare}) which are wrapped and launched by the
+\textit{classify}) which are wrapped and launched by the
 \texttt{pRolocVis} function.
@@
 \textit{main} application is designed for exploratory data analysis
-and is divied into 3 tabs: (1) PCA, (2) Profiles and (3) Table
+and is divided into 3 tabs: (1) PCA, (2) Profiles and (3) Table
 selection. The default view upon loading is the PCA tab, which
@@
 function and return a novelty detection result, according to the
-specified parameters. The algorithm parameters \texttt{times}, \texttt{GS} and \texttt{p}
-are passed to the function, along with the \texttt{fcol} to tell the
-algorithm where the input training data is contained.
+specified parameters. The algorithm parameters \texttt{times} and
+\texttt{GS} are passed to the function, along with the \texttt{fcol}
+to tell the algorithm where the input training data is contained.
 
@@
 heavily influences what type of new clusters are extracted. For
-example, if a user is interesed in the detection of small complexes
+example, if a user is interested in the detection of small complexes
 they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
@@
 classication results can be compared, and then used to assess how well
-a given model works by getting an estimate of the classiers ability to
+a given model works by getting an estimate of the classifiers ability to
 achieve a good generalisation i.e. that is given an unknown example
@@
 specify it in the above code chunk.  To use another feature variables,
-one need to explicitly specify its name using the \texttt{fcol}
+one needs to explicitly specify its name using the \texttt{fcol}
 argument (for example \texttt{fcol = "markers2"}).
@@
 
+In the code chunk above, we pass the whole \texttt{params} parameter
+results and, internally, the first pair that return the highest F1
+score are returned (using the \texttt{getParams} function above). It
+is advised to always check that these are actually good parameters
+and, if necessary, set them explicitly, as shown below.
+
+<<classify2, eval = FALSE>>=
+hl <- svmClassification(hl, cost = 16, sigma = 0.1, class.weights = w, fcol = "markers")
+@
+
 Automatically, the output of the above classification, the organelle
@@
 \caption{Classification results. Colours indicate class membership and
-  point size are representative or the classification confidence.}
+  point size are representative of the classification confidence.}
   \label{fig:plotSVM}
@@
 ## Plot a boxplot of the scores of each organelle
-par(oma = c(10.5, 0, 0, 0))
+par(oma = c(10.5, 0, 0, 0)) ## sets outer margins
 boxplot(svm.scores ~ svm, data = fData(preds), 
@@
 <<orgquants>>=
-ts <- orgQuants(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = .5)
+(ts <- orgQuants(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = .5))
 hl <- getPredictions(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = ts)
+@
+
+The organelle threshold (\texttt{ts} above) can also be set manually
+using an interactive app (see below) or by using a named vector of
+thresholds, as shown in the putative example below for 4 organelles.
+
+<<orgts>>=
+(ts <- setNames(c(0.612, 0.701, 0.81, 0.92), c("PM", "Mito", "Golgi", "ER")))
 @
@@
 \textit{User-defined}. By default, when the application is launched
-quatile scoring is selected and set to 0.5, the median. The
+quantile scoring is selected and set to 0.5, the median. The
 class-specific score thresholds that correspond to selecting the
-desired quantile are shown on as red dots on the boxplot. The
-assignments on the PCA plot are also updated according to the selected
+desired quantile are shown as red dots on the boxplot. The assignments
+on the PCA plot are also updated according to the selected
 threshold. The quantile threshold can be set by moving the
@@
 GO terms that have been observed for the 5032 proteins in the
-hyperLOPIT dataset are retieved. Users should note that the number of
-GO terms retreived is also dependent on the database version queried
-and thus is always subject to change. We find it is common to see GO
-terms with only one protein assigned to that term. Such terms do not
-bring any information for building the classifier and are thus removes
-using the \texttt{filterBinMSnSet} function.
+hyperLOPIT dataset are retrieved. Users should note that the number of
+GO terms used is also dependent on the database version queried and
+thus is always subject to change. We find it is common to see GO terms
+with only one protein assigned to that term. Such terms do not bring
+any information for building the classifier and are thus removed using
+the \texttt{filterBinMSnSet} function.
 
@@
 these niches and only optimise over the remaining organelles.  This
-can signifciantly cut down optimisation time as by removing these 4
+can significantly cut down optimisation time as by removing these 4
 classes from the optimisation (and not the classification) we only
@@
 the \texttt{knntlOptimisation} with the 10 remaining classes. (Note:
-this is not run live as this the \texttt{hl} dataset with 10 classes,
+this is not run live as the \texttt{hl} dataset with 10 classes,
 707 markers and 4$^{10}$ combinations of parameters takes around 76
@@
 weights for each class. In the optimisation above, for time
-constraints, we removed the two ribosomal subunits and the two nulcear
+constraints, we removed the two ribosomal subunits and the two nuclear
 compartments, and therefore in the code chunk below when we extract
@@
 
+In \Biocpkg{pRoloc}'s
+\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-transfer-learning.pdf}{transfer
+  learning vignette}, we demonstrate how to use imaging data from the
+Human Protein Atlas \cite{Uhlen:2010} via the \Biocpkg{hpar} package
+\cite{hpar} as an auxiliary data source.
+
 \section*{Unsupervised machine learning}
 
-In \texttt{pRoloc} there is functionality for unsupervsied machine
+In \Biocpkg{pRoloc} there is functionality for unsupervsied machine
 learning methods. In unsupervised learning, the training data consists
@@
 between marker proteins using he \texttt{mrkHClust} function, where
-the euclidean distance bewteen average class-specific profiles is used
+the euclidean distance between average class-specific profiles is used
 to produce a dendrogramme describing a simple relationship between the
@@
 <<hclust, out.width=".5\\textwidth", fig.asp = 1.5, out.extra='angle=-90'>>=
-par(mar = c(15, 4, 1, 0))
+par(mar = c(15, 4, 1, 0)) ## set figure margin
 mrkHClust(hl)
@@
 good use of valuable prior knowledge, and in our experience
-unsupervised clustering can be extremely difficult due poor estimates
-of the number of clusters that may appear in the data.
+unsupervised clustering can be extremely difficult due to poor
+estimates of the number of clusters that may appear in the data.
 
@@
 
-Exporting to a spreadsheet however looses a lot of important
+Exporting to a spreadsheet however loses a lot of important
 information, such as the processing data, and the sample metadata in
-the \textit{phenoData} slot. Other objects, such a parameters from the
+the \textit{phenoData} slot. Other objects, such as parameters from the
 machine learning optimisation, cannot be represented as tabular
@@
 
-LMB is supported by a Wellcome Trust Technology Development Grant
-(grant number 108467/Z/15/Z). LG is supported by the BBSRC Strategic
-Longer and Larger grant (Award BB/L002817/1).
+LMB and CMM are supported by a Wellcome Trust Technology Development
+Grant (grant number 108441/Z/15/Z). LG is supported by the BBSRC
+Strategic Longer and Larger grant (Award BB/L002817/1).
 

### 2016-12-02 bioc-workflow.tex
updates based on Claire's comments

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:19:06 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 19:09:16 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 17:19:06 2016 
-## Subset [6725,20][5032,20] Fri Dec  2 17:19:07 2016 
-## Removed features with more than 0 NAs: Fri Dec  2 17:19:07 2016 
-## Dropped featureData's levels Fri Dec  2 17:19:07 2016 
+## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 19:09:16 2016 
+## Subset [6725,20][5032,20] Fri Dec  2 19:09:16 2016 
+## Removed features with more than 0 NAs: Fri Dec  2 19:09:16 2016 
+## Dropped featureData's levels Fri Dec  2 19:09:16 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Fri Dec  2 17:19:09 2016 
+##  Created on Fri Dec  2 19:09:18 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Dec  2 17:19:09 2016
+% Fri Dec  2 19:09:19 2016
 \begin{table}[ht]
@@
 
+In addition, the reproducibility can be assessed by performing
+independent classification analyses on each replicate (see the section
+on \textit{Supervised machine learning} below) and comparing the the
+results. Even when the gradient conditions different (for unexpected
+technical or voluntary reasons, to maximise resolution when combining
+experiments \cite{Trotter:2010}), one expects agreement in the most
+confident organelle assignments.
+
 \clearpage
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Dec  2 17:19:08 2016 
-## Added markers from  'pdres' marker vector. Fri Dec  2 17:19:09 2016 
+## Added markers from  'mrk' marker vector. Fri Dec  2 19:09:18 2016 
+## Added markers from  'pdres' marker vector. Fri Dec  2 19:09:19 2016 
 ##  MSnbase version: 2.1.4
@@
 
+In the code chunk above, we pass the whole \texttt{params} parameter
+results and, internally, the first pair that return the highest F1
+score are returned (using the \texttt{getParams} function above). It
+is advised to always check that these are actually good parameters
+and, if necessary, set them explicitly, as shown below.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl,} \hlkwc{cost} \hlstd{=} \hlnum{16}\hlstd{,} \hlkwc{sigma} \hlstd{=} \hlnum{0.1}\hlstd{,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
 Automatically, the output of the above classification, the organelle
@@
 \hlcom{## Plot a boxplot of the scores of each organelle}
-\hlkwd{par}\hlstd{(}\hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{10.5}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))}
+\hlkwd{par}\hlstd{(}\hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{10.5}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))} \hlcom{## sets outer margins}
 \hlkwd{boxplot}\hlstd{(svm.scores} \hlopt{~} \hlstd{svm,} \hlkwc{data} \hlstd{=} \hlkwd{fData}\hlstd{(preds),}
@@
 \begin{alltt}
-\hlstd{ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{)}
+\hlstd{(ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{))}
 \end{alltt}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4368179               0.3037653               0.3856993 
+##               0.4359719               0.3027894               0.3807832 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6916475               0.5951696               0.4377555 
+##               0.6994864               0.6018155               0.4232624 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4317514               0.5919750               0.9527663 
+##               0.4141776               0.5871408               0.9494709 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7947663               0.7096575               0.3269003 
+##               0.7959998               0.7097865               0.3121881 
 ##         Plasma membrane              Proteasome 
-##               0.7210996               0.4155492
+##               0.7104260               0.4136511
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##               0.4359719               0.3027894               0.3807832 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##               0.6994864               0.6018155               0.4232624 
+##    Extracellular matrix                Lysosome           Mitochondrion 
+##               0.4141776               0.5871408               0.9494709 
+##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+##               0.7959998               0.7097865               0.3121881 
+##         Plasma membrane              Proteasome 
+##               0.7104260               0.4136511
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      88 
+##                      86                     169                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     477                     100 
+##                     296                     477                      93 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     513 
+##                      27                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     231                     344                      42 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     318                     157                    2059
+##                     326                     158                    2058
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+The organelle threshold (\texttt{ts} above) can also be set manually
+using an interactive app (see below) or by using a named vector of
+thresholds, as shown in the putative example below for 4 organelles.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{(ts} \hlkwb{<-} \hlkwd{setNames}\hlstd{(}\hlkwd{c}\hlstd{(}\hlnum{0.612}\hlstd{,} \hlnum{0.701}\hlstd{,} \hlnum{0.81}\hlstd{,} \hlnum{0.92}\hlstd{),} \hlkwd{c}\hlstd{(}\hlstr{"PM"}\hlstd{,} \hlstr{"Mito"}\hlstd{,} \hlstr{"Golgi"}\hlstd{,} \hlstr{"ER"}\hlstd{)))}
+\end{alltt}
+\begin{verbatim}
+##    PM  Mito Golgi    ER 
+## 0.612 0.701 0.810 0.920
 \end{verbatim}
@@
 \begin{alltt}
-\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{15}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{0}\hlstd{))}
+\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{15}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{0}\hlstd{))} \hlcom{## set figure margin}
 \hlkwd{mrkHClust}\hlstd{(hl)}

### 2016-12-03 bioc-workflow.Rnw
updates based on Kathryn's comments - up to page 17

@@
 Quantitative mass spectrometry based spatial proteomics involves
-elaborate, expensive and time consuming experimental procedures and
+elaborate, expensive and time consuming experimental protocols and
 considerable effort is invested in the generation of such data.
@@
 establish high quality proteome-wide datasets (see for example
-\cite{Gatto:2010} for a review, and \cite{hyper,Itzhak:2016} for recent
-examples). However, data analysis is as critical as data production
-for reliable and insightful biological interpretation. Here, we walk
-the reader through a typical pipeline for the analysis of such data
-using several Bioconductor packages for the R statistical programming
-environment.
+\cite{Gatto:2010} for a review, and \cite{hyper,Itzhak:2016} for
+recent examples). However, data analysis is as critical as data
+production for reliable and insightful biological
+interpretation. Here, we walk the reader through a typical pipeline
+for the analysis of such data using several Bioconductor packages for
+the R statistical programming environment.
 
@@
 denser organelles.  The intersect of replicates 1 and 2 was treated as
-a 20-plex dataset for the analysis.  As discussed in the manuscript
+a 20-plex dataset for the analysis.  As discussed in the publication
 \cite{hyper}, it has been shown that combining replicates from
@@
 \cite{Trotter:2010}. The combination of replicates resulted in 5032
-proteins common in both experiments.
+proteins common to both experiments.
 
@@
 particular the \textit{Markers} section) and results from data
-analysis. The sample metadata would, for example, record what
-labelling ions were used for the respective fraction (when labelled
+analysis. The sample metadata would, for example, record what stable
+isotope labels were used for the respective fraction (when labelled
 quantitation is used), replicate number, fraction number along the
@@
 under \textit{Processing information} in the textual object summary
-when an \texttt{MSnSet}'s name it typed in the R console.
+when an \texttt{MSnSet}'s name is typed in the R console.
 
@@
 quantitation data. This can be done using the \texttt{getEcols}
-function inside R. The spreadsheet that was deposited by the authors
-contains two headers, with the second header containing information
-about where the quantitation data is stored.
+function inside R. The spreadsheet deposited by the authors contains
+two headers, with the second header containing information about where
+the quantitation data is stored.
 
@@
 for the respective 2 x 10-plex reporter tags (\Sexpr{ncol(hl)}
-columns/samples), along with associated feature meta-data such as
+columns or samples), along with associated feature meta-data such as
 protein markers, protein description, number of quantified peptides
@@
 the intensities for each replicate have already been re-scaled to 1 in
-Proteome Discoverer v1.4, Thermo Fisher. However, if one wanted to do
-this they would execute the \texttt{normalise} function as
+Proteome Discoverer v1.4 Thermo Fisher. However, if the data require
+normalisation, the user can execute the \texttt{normalise} function as
 demonstrated in the below code chunk.
@@
 
-This transformation of the data assures to cancel the effect of the
-absolute intensities of the quantitative features along the rows, to
-focus subsequent analyses on the relative profiles along the
+This transformation of the data assures cancellation of the effect of
+the absolute intensities of the quantitative features along the rows,
+and focus subsequent analyses on the relative profiles along the
 sub-cellular channels.
@@
 the imputation method is then determined by a \texttt{methods}
-parameter. To impute missing values using nearest neighbour
-imputation, one would
+parameter (see \texttt{?impute} for available options). To impute
+missing values using nearest neighbour imputation, one would
 
@@
 
-When more than 2 datasets are to be combined and too many proteins were
-not consistently assayed, leading to too many proteins being filtered
-out, we suggest to implement an ensemble of classifiers voting on
-protein-sub-cellular niche membership over the output of several
-experiments (see section \textit{Supervised machine learning} for the
-description of sub-cellular assignments).
+When more than 2 datasets are to be combined and too many proteins
+have not been consistently assayed, leading to too many proteins being
+filtered out, we suggest to implement an ensemble of classifiers
+voting on protein-sub-cellular niche membership over the output of
+several experiments (see section \textit{Supervised machine learning}
+for the description of sub-cellular assignments).
 
@@
 that sub-cellular niches have been separated along the gradient. Based
-on De Duve's principle \cite{DeDuve:1981} proteins that co-localise
-exhibit similar quantitation profiles across the gradient fractions
-employed. One approach that has been widely used to visualise and
-inspect high throughput mass spectrometry-based proteomics data is
-principal components analysis (PCA). PCA is one of many dimensionality
-reduction methods, that allows one to effectively summarise
-multi-dimensional data in to 2 or 3 dimensions to enable
+on De Duve's principle \cite{DeDuve:1981} proteins that co-localise in
+a cell, exhibit similar quantitation profiles across the gradient
+fractions employed. One approach that has been widely used to
+visualise and inspect high throughput mass spectrometry-based
+proteomics data is principal components analysis (PCA). PCA is one of
+many dimensionality reduction methods, that allows one to effectively
+summarise multi-dimensional data in to 2 or 3 dimensions to enable
 visualisation. Very generally, the original continuous
@@
 components ordered according to the amount of variability that they
-describe. The \texttt{plot2D} and \texttt{plot3D} methods in \Biocpkg{pRoloc} allows one to
-plot the principal components (PCs) of a dataset against one another,
-by default the first two components are plotted on the x- and y-axis
-for the \texttt{plot2D} function, and first three components are
-loaded for the \texttt{plot3D} function, respectively (the
-\texttt{dims} argument can be used to plot other PCs). If
-distinct clusters are observed, we assume that there is organellar
-separation present in the data. Although, representing the
+describe. The \texttt{plot2D} and \texttt{plot3D} methods in
+\Biocpkg{pRoloc} allows one to plot the principal components (PCs) of
+a dataset against one another, by default the first two components are
+plotted on the x- and y-axis for the \texttt{plot2D} function, and
+first three components are loaded for the \texttt{plot3D} function,
+respectively (the \texttt{dims} argument can be used to plot other
+PCs). If distinct clusters are observed, we assume that there is
+organellar separation present in the data. Although, representing the
 multi-dimensional data along a limited set of PCs does not give us a
@@
 
-In the code chunk below we produce a 2-dimensional PCA plot of the mouse stem cell
-dataset (Figure \ref{fig:pcahl}). One point on the plot represents one
-protein. We can indeed see several distinct protein clusters. We
-specify \texttt{fcol = NULL}, which means not to consider any feature
-variable to annotate the features (proteins) with colours. We will see
-later how to use this argument to annotate the PCA plot with prior information
-about sub-cellular localisation.
+In the code chunk below we produce a 2-dimensional PCA plot of the
+mouse stem cell dataset (Figure \ref{fig:pcahl}). Each point on the
+plot represents one protein. We can indeed see several distinct
+protein clusters. We specify \texttt{fcol = NULL} to ignore feature
+metadata columns and not annotate any feature (protein) with a
+colour. We will see later how to use this argument to annotate the PCA
+plot with prior information about sub-cellular localisation.
 
@@
 \caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
-  represent a single protein, and cluster of proteins represent
+  represents a single protein, and cluster of proteins represent
   proteins residing in the same sub-cellular niche. The figure on the
@@
 lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
-boxplot(exprs(hl[, o]), las = 2)
-@  
-  \caption{Protein profiles and distribution of channel intensities.}
+boxplot(exprs(hl[, o]), las = 2, cex.axis = .7)
+@  
+\caption{Protein profiles and distribution of channel intensities. The
+  red dots represent the mean relative intensity for each channel. }
   \label{fig:qcbx}
@@
 mouse markers stored in \Biocpkg{pRoloc} (\texttt{mmus} markers) are
-also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below), so
-it is straightforward to match names between the markers and the
-\texttt{MSnSet} instance using the \texttt{addMarkers} function.
+also Uniprot IDs (see \texttt{head(mrk)} in the code chunk below, that
+displays the 6 first markers), so it is straightforward to match names
+between the markers and the \texttt{MSnSet} instance using the
+\texttt{addMarkers} function.
 
@@
 definition of markers, i.e. one that broadly or over-confidently
-defines markers, risks to erroneously assign proteins to a single
-location, when, in reality, they reside in multiple locations
+defines markers, risks the erroneous assignment of proteins to a
+single location, when, in reality, they reside in multiple locations
 (including the assumed unique location). One can not expect to
@@
 to the appropriate identifier. Sometimes, we find the equivalent entry
-name, Uniprot ID or accession number is stored with the data, which
-makes conversion between identifers relatively straightforward. If
-this is not the case however, conversion can be performed using
-\Biocpkg{biomaRt}, the Bioconductor
+name, Uniprot ID or accession number is stored in the feature
+metadata, which makes conversion between identifers relatively
+straightforward. If this is not the case however, conversion can be
+performed using \Biocpkg{biomaRt}, the Bioconductor
 \href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation
@@
 \subsection*{Visualising markers}
+
 Having added a the mouse markers to our \texttt{fData} from the
-\texttt{pRolocmarkers} we can now visualise these annotations along
+\texttt{pRolocmarkers}, we can now visualise these annotations along
 the PCA plot using the \texttt{plot2D} function and then use the
@@
 \caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
-  with \texttt{plt2D} and \texttt{plot3D} (right).}
+  with \texttt{plot2D}.}
   \label{fig:plotmarkers}
+\end{figure}
+
+The data can also be visualised along three PCs using the
+\texttt{plot3D} function (Figure \ref{fig:plotmarkers3d}). When
+produced interactively, the plot can be rotated and zoomined using the
+mouse.
+
+\begin{figure}[htb]
+  \centering
+<<plotmarkers3d, eval=FALSE>>=
+plot3D(hl, dims = c(1, 2, 7))
+@
+  \includegraphics[width=.55\textwidth]{./Figures/plot3d.pdf}
+  \caption{Using the \texttt{plot3D} function to visualise the
+    \texttt{hl} dataset along PCs 1, 2 and 7. }
+  \label{fig:plotmarkers3d}
 \end{figure}
@@
 representation of the protein profiles occupancy along the gradient
-using the \texttt{plotDist} function. While the PCA plot enables to
-efficiently visualise the complete dataset and assess the relative
-separation of different sub-cellular niches, comparing profiles of a
-few marker clusters is useful to assess how exactly they differ (in
-terms of peak channels, for example). On figure \ref{fig:plotDist2},
-we plot the profile of the mitochondrial and peroxisome markers to
-highlight the differences in profiles between these two sets of
-markers in the channels labelled with tag 129C, as represented above
-along the 7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
+using the \texttt{plotDist} function. While the PCA plot enables
+efficient visualisation of the complete dataset and assessment the
+relative separation of different sub-cellular niches, comparing
+profiles of a few marker clusters is useful to assess how exactly they
+differ (in terms of peak channels, for example). On figure
+\ref{fig:plotDist2}, we plot the profile of the mitochondrial and
+peroxisome markers to highlight the differences in profiles between
+these two sets of markers in the channels labelled with tag 129C, as
+represented above along the 7th PC on the PCA plot on figure
+\ref{fig:plotmarkers}.
 
@@
 and is divided into 3 tabs: (1) PCA, (2) Profiles and (3) Table
-selection. The default view upon loading is the PCA tab, which
-features a clickable interface and zoomable PCA plot with an
+selection. The default view upon loading is the \textit{PCA} tab,
+which features a clickable interface and zoomable PCA plot with an
 interactive data table for displaying the quantitation information.
 Particular proteins of interest can be highlighted using the text
-search box. There is also an alternate profiles tab for visualisation
-of the protein profiles, which can be used to examine the patterns of
-proteins of interest. The Table selection tab provides an interface to
-control data table column selection. A short
+search box. There is also a \textit{Profiles} tab for visualisation of
+the protein profiles, which can be used to examine the patterns of
+proteins of interest. The \textit{Table selection} tab provides an
+interface to control data table column selection. A short
 animation\footnote{\url{https://github.com/lmsimp/bioc-pRoloc-hyperLOPIT-workflow/blob/master/Figures/pRolocVis\_pca.gif}}
-illustrating the interface is available in the manuscript repository~\cite{ghrepo}.
+illustrating the interface is available in the manuscript
+repository~\cite{ghrepo}.
 

### 2016-12-03 bioc-workflow.tex
updates based on Kathryn's comments - up to page 17

@@
 for the respective 2 x 10-plex reporter tags (20
-columns/samples), along with associated feature meta-data such as
+columns or samples), along with associated feature meta-data such as
 protein markers, protein description, number of quantified peptides
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 19:09:16 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:13:15 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Fri Dec  2 19:09:16 2016 
-## Subset [6725,20][5032,20] Fri Dec  2 19:09:16 2016 
-## Removed features with more than 0 NAs: Fri Dec  2 19:09:16 2016 
-## Dropped featureData's levels Fri Dec  2 19:09:16 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:13:15 2016 
+## Subset [6725,20][5032,20] Sat Dec  3 19:13:15 2016 
+## Removed features with more than 0 NAs: Sat Dec  3 19:13:15 2016 
+## Dropped featureData's levels Sat Dec  3 19:13:15 2016 
 ##  MSnbase version: 1.21.7
@@
 \hlkwd{lines}\hlstd{(}\hlkwd{colMeans}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o])),} \hlkwc{col} \hlstd{=} \hlstr{"red"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"b"}\hlstd{)}
-\hlkwd{boxplot}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o]),} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{)}
+\hlkwd{boxplot}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o]),} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{cex.axis} \hlstd{=} \hlnum{.7}\hlstd{)}
 \end{alltt}
@@
 \end{knitrout}
-  \caption{Protein profiles and distribution of channel intensities.}
+\caption{Protein profiles and distribution of channel intensities. The
+  red dots represent the mean relative intensity for each channel. }
   \label{fig:qcbx}
@@
 \caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
-  with \texttt{plt2D} and \texttt{plot3D} (right).}
+  with \texttt{plot2D}.}
   \label{fig:plotmarkers}
+\end{figure}
+
+The data can also be visualised along three PCs using the
+\texttt{plot3D} function (Figure \ref{fig:plotmarkers3d}). When
+produced interactively, the plot can be rotated and zoomined using the
+mouse.
+
+\begin{figure}[htb]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot3D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{,} \hlnum{7}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+  \includegraphics[width=.55\textwidth]{./Figures/plot3d.pdf}
+  \caption{Using the \texttt{plot3D} function to visualise the
+    \texttt{hl} dataset along PCs 1, 2 and 7. }
+  \label{fig:plotmarkers3d}
 \end{figure}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Fri Dec  2 19:09:18 2016 
+##  Created on Sat Dec  3 19:13:17 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Fri Dec  2 19:09:19 2016
+% Sat Dec  3 19:13:18 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Fri Dec  2 19:09:18 2016 
-## Added markers from  'pdres' marker vector. Fri Dec  2 19:09:19 2016 
+## Added markers from  'mrk' marker vector. Sat Dec  3 19:13:17 2016 
+## Added markers from  'pdres' marker vector. Sat Dec  3 19:13:18 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4359719               0.3027894               0.3807832 
+##               0.4365435               0.3040122               0.3854846 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6994864               0.6018155               0.4232624 
+##               0.6882175               0.5970631               0.4277083 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4141776               0.5871408               0.9494709 
+##               0.4184674               0.5929501               0.9507120 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7959998               0.7097865               0.3121881 
+##               0.7951351               0.7077463               0.3169610 
 ##         Plasma membrane              Proteasome 
-##               0.7104260               0.4136511
+##               0.7256612               0.4192384
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4359719               0.3027894               0.3807832 
+##               0.4365435               0.3040122               0.3854846 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6994864               0.6018155               0.4232624 
+##               0.6882175               0.5970631               0.4277083 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4141776               0.5871408               0.9494709 
+##               0.4184674               0.5929501               0.9507120 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7959998               0.7097865               0.3121881 
+##               0.7951351               0.7077463               0.3169610 
 ##         Plasma membrane              Proteasome 
-##               0.7104260               0.4136511
+##               0.7256612               0.4192384
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     169                      84 
+##                      84                     171                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     477                      93 
+##                     295                     478                      99 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     523 
+##                      27                     124                     520 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     229                     345                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     326                     158                    2058
+##                     315                     158                    2060
 \end{verbatim}

### 2016-12-03 bioc-workflow.Rnw
last updates based on Kathryn's comments

@@
 function and return a novelty detection result, according to the
-specified parameters. The algorithm parameters \texttt{times} and
-\texttt{GS} are passed to the function, along with the \texttt{fcol}
-to tell the algorithm where the input training data is contained.
+specified parameters. The algorithm parameters \texttt{times} (number
+of iterations) and \texttt{GS} (minimum number of proteins required to
+form a new phenotype) are passed to the function, along with the
+\texttt{fcol} to tell the algorithm where the input training data is
+contained.
 
@@
 a classifier on the labelled training data. As previously mentioned,
-one first needs to train the classifiers parameters before an
+one first needs to train the classifier's parameters before an
 algorithm can be used to predict the class labels of the proteins with
@@
 classication results can be compared, and then used to assess how well
-a given model works by getting an estimate of the classifiers ability to
-achieve a good generalisation i.e. that is given an unknown example
+a given model works by getting an estimate of the classifiers ability
+to achieve a good generalisation i.e. that is given an unknown example
 predict its class label with high accuracy. In \Biocpkg{pRoloc},
@@
 harmonic mean of precision and recall. In the code chunk below we
-demonstrate how to optimise the free parameters, \texttt{sigma} and
-\texttt{cost}, of a classical SVM classifier with a Gaussian kernel
-using the function \texttt{svmOptimisation}. As the number of labelled
-instances per class varies from organelle to organelle, we can account
-for class imbalance by setting specific class weights when generating
-the SVM model. Below the weights, \texttt{w} are set to be inversely
-proportional to the class frequencies.
+demonstrate how to optimise the free parameters, \texttt{sigma} (the
+inverse kernel width for the radial basis kernel) and \texttt{cost}
+(the cost of constraints violation), of a classical SVM classifier
+with a Gaussian kernel using the function \texttt{svmOptimisation}. As
+the number of labelled instances per class varies from organelle to
+organelle, we can account for class imbalance by setting specific
+class weights when generating the SVM model. Below the weights,
+\texttt{w} are set to be inversely proportional to the class
+frequencies.
 
@@
 threshold. The quantile threshold can be set by moving the
-corresponding quantile slider. If one wished to set their own cut-offs
-the \textit{User-defined} radio button must be selected and then the
-sliders for defining user-specified scores become active and the
-scores and highlighted on the boxplot by blue dots. For more
+corresponding quantile slider. If the users wishes to set their own
+cut-offs, the \textit{User-defined} radio button must be selected and
+then the sliders for defining user-specified scores become active and
+the scores and highlighted on the boxplot by blue dots. For more
 information we refer users to the \Biocpkg{pRolocGUI} tutorial
@@
 have used to annotate the proteins in your \texttt{MSnSet}. It is also
-possible to pass patterns to match the species e.g. in the code chunk
-below we pass \texttt{"Mus musculus"}, and the identifier type for our
-data (see \texttt{featureNames(hl)}) which is
+possible to pass arbitrary text to match the species e.g. in the code
+chunk below we pass \texttt{"Mus musculus"}, and the identifier type
+for our data (see \texttt{featureNames(hl)}) which is
 \texttt{"Uniprot/Swissprot"}, for the Biomart query.
@@
 LMB and CMM are supported by a Wellcome Trust Technology Development
-Grant (grant number 108441/Z/15/Z). LG is supported by the BBSRC
-Strategic Longer and Larger grant (Award BB/L002817/1).
+Grant (grant number 108441/Z/15/Z). KSL is a Wellcome Trust Joint
+Investigator (110170/Z/15/Z). LG is supported by the BBSRC Strategic
+Longer and Larger grant (Award BB/L002817/1).
 

### 2016-12-03 bioc-workflow.tex
last updates based on Kathryn's comments

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:13:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:46:20 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:13:15 2016 
-## Subset [6725,20][5032,20] Sat Dec  3 19:13:15 2016 
-## Removed features with more than 0 NAs: Sat Dec  3 19:13:15 2016 
-## Dropped featureData's levels Sat Dec  3 19:13:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:46:20 2016 
+## Subset [6725,20][5032,20] Sat Dec  3 19:46:21 2016 
+## Removed features with more than 0 NAs: Sat Dec  3 19:46:21 2016 
+## Dropped featureData's levels Sat Dec  3 19:46:21 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Sat Dec  3 19:13:17 2016 
+##  Created on Sat Dec  3 19:46:22 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Sat Dec  3 19:13:18 2016
+% Sat Dec  3 19:46:23 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Sat Dec  3 19:13:17 2016 
-## Added markers from  'pdres' marker vector. Sat Dec  3 19:13:18 2016 
+## Added markers from  'mrk' marker vector. Sat Dec  3 19:46:22 2016 
+## Added markers from  'pdres' marker vector. Sat Dec  3 19:46:23 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4365435               0.3040122               0.3854846 
+##               0.4332618               0.3058476               0.3871539 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6882175               0.5970631               0.4277083 
+##               0.6745670               0.6093870               0.4335865 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4184674               0.5929501               0.9507120 
+##               0.4193893               0.6121402               0.9498369 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951351               0.7077463               0.3169610 
+##               0.7946622               0.7092443               0.3158764 
 ##         Plasma membrane              Proteasome 
-##               0.7256612               0.4192384
+##               0.7096980               0.4169878
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4365435               0.3040122               0.3854846 
+##               0.4332618               0.3058476               0.3871539 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6882175               0.5970631               0.4277083 
+##               0.6745670               0.6093870               0.4335865 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4184674               0.5929501               0.9507120 
+##               0.4193893               0.6121402               0.9498369 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951351               0.7077463               0.3169610 
+##               0.7946622               0.7092443               0.3158764 
 ##         Plasma membrane              Proteasome 
-##               0.7256612               0.4192384
+##               0.7096980               0.4169878
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     171                      88 
+##                      85                     172                      85 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     478                      99 
+##                     298                     475                     107 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     520 
+##                      25                     119                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     345                      39 
+##                     230                     342                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     315                     158                    2060
+##                     320                     157                    2056
 \end{verbatim}

### 2016-12-03 bioc-workflow.tex
add Jean Beltran reference

@@
 establish high quality proteome-wide datasets (see for example
-\cite{Gatto:2010} for a review, and \cite{hyper,Itzhak:2016} for
-recent examples). However, data analysis is as critical as data
-production for reliable and insightful biological
-interpretation. Here, we walk the reader through a typical pipeline
-for the analysis of such data using several Bioconductor packages for
-the R statistical programming environment.
+\cite{Gatto:2010} for a review, and
+\cite{hyper,Itzhak:2016,Jean_Beltran:2016} for recent
+examples). However, data analysis is as critical as data production
+for reliable and insightful biological interpretation. Here, we walk
+the reader through a typical pipeline for the analysis of such data
+using several Bioconductor packages for the R statistical programming
+environment.
 
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:46:20 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 20:38:06 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 19:46:20 2016 
-## Subset [6725,20][5032,20] Sat Dec  3 19:46:21 2016 
-## Removed features with more than 0 NAs: Sat Dec  3 19:46:21 2016 
-## Dropped featureData's levels Sat Dec  3 19:46:21 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 20:38:06 2016 
+## Subset [6725,20][5032,20] Sat Dec  3 20:38:06 2016 
+## Removed features with more than 0 NAs: Sat Dec  3 20:38:06 2016 
+## Dropped featureData's levels Sat Dec  3 20:38:06 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Sat Dec  3 19:46:22 2016 
+##  Created on Sat Dec  3 20:38:08 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Sat Dec  3 19:46:23 2016
+% Sat Dec  3 20:38:08 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Sat Dec  3 19:46:22 2016 
-## Added markers from  'pdres' marker vector. Sat Dec  3 19:46:23 2016 
+## Added markers from  'mrk' marker vector. Sat Dec  3 20:38:08 2016 
+## Added markers from  'pdres' marker vector. Sat Dec  3 20:38:09 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4332618               0.3058476               0.3871539 
+##               0.4347094               0.3030356               0.3854031 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6745670               0.6093870               0.4335865 
+##               0.6849680               0.5994838               0.4265235 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4193893               0.6121402               0.9498369 
+##               0.4183489               0.5905225               0.9493973 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946622               0.7092443               0.3158764 
+##               0.7954761               0.7092834               0.3116761 
 ##         Plasma membrane              Proteasome 
-##               0.7096980               0.4169878
+##               0.7196883               0.4190620
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4332618               0.3058476               0.3871539 
+##               0.4347094               0.3030356               0.3854031 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6745670               0.6093870               0.4335865 
+##               0.6849680               0.5994838               0.4265235 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4193893               0.6121402               0.9498369 
+##               0.4183489               0.5905225               0.9493973 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946622               0.7092443               0.3158764 
+##               0.7954761               0.7092834               0.3116761 
 ##         Plasma membrane              Proteasome 
-##               0.7096980               0.4169878
+##               0.7196883               0.4190620
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      85 
+##                      85                     170                      87 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     298                     475                     107 
+##                     300                     472                     102 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      25                     119                     523 
+##                      27                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     342                      38 
+##                     229                     345                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     320                     157                    2056
+##                     317                     155                    2058
 \end{verbatim}

### 2016-12-04 bioc-workflow.Rnw
add Jean Beltran reference
update annotation orientation and size - addresses #39

@@
 <<qcbx, out.width="\\textwidth", fig.asp=1/2>>=
-par(mfrow = c(1, 2)) ## creates a two-panel figure
+par(mfrow = c(1, 2), ## creates a two-panel figure
+    las = 2,         ## axis labels orientation
+    cex.axis = .7)   ## axis label size
 o <- order(hl$Iodixonal.Density)
-plotDist(hl[, o], pcol = "#00000010")
+plotDist(hl[, o], pcol = "#00000010", xlab = "")
 lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
-boxplot(exprs(hl[, o]), las = 2, cex.axis = .7)
+boxplot(exprs(hl[, o]))
 @  
@@
 associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5
-in lysosomal and endosomal proteins.
+in both lysosomal and endosomal proteins.
 

### 2016-12-04 bioc-workflow.tex
update annotation orientation and size - addresses #39

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 20:38:06 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sun Dec  4 09:22:15 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sat Dec  3 20:38:06 2016 
-## Subset [6725,20][5032,20] Sat Dec  3 20:38:06 2016 
-## Removed features with more than 0 NAs: Sat Dec  3 20:38:06 2016 
-## Dropped featureData's levels Sat Dec  3 20:38:06 2016 
+## Combined [6725,20] and [6268,10] MSnSets Sun Dec  4 09:22:15 2016 
+## Subset [6725,20][5032,20] Sun Dec  4 09:22:16 2016 
+## Removed features with more than 0 NAs: Sun Dec  4 09:22:16 2016 
+## Dropped featureData's levels Sun Dec  4 09:22:16 2016 
 ##  MSnbase version: 1.21.7
@@
 \begin{alltt}
-\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))} \hlcom{## creates a two-panel figure}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{),} \hlcom{## creates a two-panel figure}
+    \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{,}         \hlcom{## axis labels orientation}
+    \hlkwc{cex.axis} \hlstd{=} \hlnum{.7}\hlstd{)}   \hlcom{## axis label size}
 \hlstd{o} \hlkwb{<-} \hlkwd{order}\hlstd{(hl}\hlopt{$}\hlstd{Iodixonal.Density)}
-\hlkwd{plotDist}\hlstd{(hl[, o],} \hlkwc{pcol} \hlstd{=} \hlstr{"#00000010"}\hlstd{)}
+\hlkwd{plotDist}\hlstd{(hl[, o],} \hlkwc{pcol} \hlstd{=} \hlstr{"#00000010"}\hlstd{,} \hlkwc{xlab} \hlstd{=} \hlstr{""}\hlstd{)}
 \hlkwd{lines}\hlstd{(}\hlkwd{colMeans}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o])),} \hlkwc{col} \hlstd{=} \hlstr{"red"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"b"}\hlstd{)}
-\hlkwd{boxplot}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o]),} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{,} \hlkwc{cex.axis} \hlstd{=} \hlnum{.7}\hlstd{)}
+\hlkwd{boxplot}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o]))}
 \end{alltt}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Sat Dec  3 20:38:08 2016 
+##  Created on Sun Dec  4 09:22:17 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Sat Dec  3 20:38:08 2016
+% Sun Dec  4 09:22:18 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Sat Dec  3 20:38:08 2016 
-## Added markers from  'pdres' marker vector. Sat Dec  3 20:38:09 2016 
+## Added markers from  'mrk' marker vector. Sun Dec  4 09:22:17 2016 
+## Added markers from  'pdres' marker vector. Sun Dec  4 09:22:18 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4347094               0.3030356               0.3854031 
+##               0.4351539               0.3024251               0.3833827 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6849680               0.5994838               0.4265235 
+##               0.6782466               0.6004558               0.4345984 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4183489               0.5905225               0.9493973 
+##               0.4251852               0.5903493               0.9501766 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7954761               0.7092834               0.3116761 
+##               0.7953310               0.7096086               0.3138186 
 ##         Plasma membrane              Proteasome 
-##               0.7196883               0.4190620
+##               0.7165863               0.4200117
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4347094               0.3030356               0.3854031 
+##               0.4351539               0.3024251               0.3833827 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6849680               0.5994838               0.4265235 
+##               0.6782466               0.6004558               0.4345984 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4183489               0.5905225               0.9493973 
+##               0.4251852               0.5903493               0.9501766 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7954761               0.7092834               0.3116761 
+##               0.7953310               0.7096086               0.3138186 
 ##         Plasma membrane              Proteasome 
-##               0.7196883               0.4190620
+##               0.7165863               0.4200117
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      87 
+##                      85                     170                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     300                     472                     102 
+##                     297                     476                     100 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     523 
+##                      27                     123                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     345                      38 
+##                     230                     343                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     317                     155                    2058
+##                     322                     157                    2057
 \end{verbatim}

### 2016-12-05 bioc-workflow.tex
remake pdf, good to go

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sun Dec  4 09:22:15 2016 
+## Combined [6725,20] and [6268,10] MSnSets Mon Dec  5 09:40:22 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Sun Dec  4 09:22:15 2016 
-## Subset [6725,20][5032,20] Sun Dec  4 09:22:16 2016 
-## Removed features with more than 0 NAs: Sun Dec  4 09:22:16 2016 
-## Dropped featureData's levels Sun Dec  4 09:22:16 2016 
+## Combined [6725,20] and [6268,10] MSnSets Mon Dec  5 09:40:22 2016 
+## Subset [6725,20][5032,20] Mon Dec  5 09:40:22 2016 
+## Removed features with more than 0 NAs: Mon Dec  5 09:40:22 2016 
+## Dropped featureData's levels Mon Dec  5 09:40:22 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Sun Dec  4 09:22:17 2016 
+##  Created on Mon Dec  5 09:40:24 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Sun Dec  4 09:22:18 2016
+% Mon Dec  5 09:40:24 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Sun Dec  4 09:22:17 2016 
-## Added markers from  'pdres' marker vector. Sun Dec  4 09:22:18 2016 
+## Added markers from  'mrk' marker vector. Mon Dec  5 09:40:24 2016 
+## Added markers from  'pdres' marker vector. Mon Dec  5 09:40:25 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4351539               0.3024251               0.3833827 
+##               0.4334288               0.3056337               0.3985514 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6782466               0.6004558               0.4345984 
+##               0.6981462               0.6008776               0.4186913 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4251852               0.5903493               0.9501766 
+##               0.4206833               0.5867712               0.9495542 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7953310               0.7096086               0.3138186 
+##               0.7955894               0.7084349               0.3137019 
 ##         Plasma membrane              Proteasome 
-##               0.7165863               0.4200117
+##               0.7297792               0.4140046
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4351539               0.3024251               0.3833827 
+##               0.4334288               0.3056337               0.3985514 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6782466               0.6004558               0.4345984 
+##               0.6981462               0.6008776               0.4186913 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4251852               0.5903493               0.9501766 
+##               0.4206833               0.5867712               0.9495542 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7953310               0.7096086               0.3138186 
+##               0.7955894               0.7084349               0.3137019 
 ##         Plasma membrane              Proteasome 
-##               0.7165863               0.4200117
+##               0.7297792               0.4140046
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      84 
+##                      85                     170                      92 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     297                     476                     100 
+##                     296                     478                      96 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     522 
+##                      27                     125                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
@@
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     157                    2057
+##                     313                     158                    2057
 \end{verbatim}

### 2016-12-08 bioc-workflow.Rnw
remake pdf, good to go
add bioc ref

@@
 the reader through a typical pipeline for the analysis of such data
-using several Bioconductor packages for the R statistical programming
+using several Bioconductor~\cite{Huber:2015} packages for the R statistical programming
 environment.

### 2016-12-08 bioc-workflow.tex
add bioc ref

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Mon Dec  5 09:40:22 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Dec  8 12:09:01 2016 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Mon Dec  5 09:40:22 2016 
-## Subset [6725,20][5032,20] Mon Dec  5 09:40:22 2016 
-## Removed features with more than 0 NAs: Mon Dec  5 09:40:22 2016 
-## Dropped featureData's levels Mon Dec  5 09:40:22 2016 
+## Combined [6725,20] and [6268,10] MSnSets Thu Dec  8 12:09:01 2016 
+## Subset [6725,20][5032,20] Thu Dec  8 12:09:01 2016 
+## Removed features with more than 0 NAs: Thu Dec  8 12:09:01 2016 
+## Dropped featureData's levels Thu Dec  8 12:09:01 2016 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Mon Dec  5 09:40:24 2016 
+##  Created on Thu Dec  8 12:09:03 2016 
 ##  Description:
@@
 % latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Mon Dec  5 09:40:24 2016
+% Thu Dec  8 12:09:03 2016
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Mon Dec  5 09:40:24 2016 
-## Added markers from  'pdres' marker vector. Mon Dec  5 09:40:25 2016 
+## Added markers from  'mrk' marker vector. Thu Dec  8 12:09:03 2016 
+## Added markers from  'pdres' marker vector. Thu Dec  8 12:09:04 2016 
 ##  MSnbase version: 2.1.4
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4334288               0.3056337               0.3985514 
+##               0.4322999               0.3055155               0.3838354 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6981462               0.6008776               0.4186913 
+##               0.6883460               0.6037025               0.4275289 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4206833               0.5867712               0.9495542 
+##               0.4276601               0.5906588               0.9501307 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7955894               0.7084349               0.3137019 
+##               0.7947781               0.7105373               0.3137337 
 ##         Plasma membrane              Proteasome 
-##               0.7297792               0.4140046
+##               0.7191668               0.4144056
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4334288               0.3056337               0.3985514 
+##               0.4322999               0.3055155               0.3838354 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6981462               0.6008776               0.4186913 
+##               0.6883460               0.6037025               0.4275289 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4206833               0.5867712               0.9495542 
+##               0.4276601               0.5906588               0.9501307 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7955894               0.7084349               0.3137019 
+##               0.7947781               0.7105373               0.3137337 
 ##         Plasma membrane              Proteasome 
-##               0.7297792               0.4140046
+##               0.7191668               0.4144056
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      92 
+##                      84                     171                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     478                      96 
+##                     296                     478                     100 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     125                     523 
+##                      27                     123                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      39 
+##                     229                     342                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     313                     158                    2057
+##                     316                     159                    2058
 \end{verbatim}

### 2018-03-26 bioc-workflow_march2017.Rnw
Daniel's comment 1

--- 2016-12-08 bioc-workflow.Rnw
@@
 the reader through a typical pipeline for the analysis of such data
-using several Bioconductor~\cite{Huber:2015} packages for the R statistical programming
-environment.
+using several Bioconductor~\cite{Huber:2015} packages for the R
+statistical programming environment.
 
@@
 \texttt{addLegend} function to map the marker classes to the
-pre-defined colours. We also display the data along the first and
-seventh PCs using the \texttt{dims} argument. Note that in these calls
-to the \texttt{plot2D} function, we have omitted the \texttt{fcol}
-argument and use of the default \texttt{"markers"} feature variable to
-annotate the plot. We choose to display PCs 1 and 7 to illustrate that
-while upper principal components explain much less variability in the
-data (2.23\% for PC7, as opposed to 48.41\% for PC1), we see that the
-mitochondrial (purple) and peroxisome (dark blue) clusters can be
-differentiated, despite the apparent overlap in the visualisation of
-the two first PCs (Figure \ref{fig:plotmarkers}).
+pre-defined colours. \hl{As mentioned previously, PCA transforms the
+original high dimensional data into a set of linearly uncorrelated
+principal components such that the first accounts for as much
+variability in the data as possible and each succeeding component in
+turn has the highest variance possible under the constraint that it
+be orthogonal to the preceding components. We saw in the previous
+section how this is useful for quality control and checking
+organelle seperation.  Adding marker definiton allows one to quickly
+see if known residents appear in defined clusters. One must be
+careful though as different organelles may be resolved in different
+dimensions.} For example, we can display the data along the first and
+seventh PCs using the \texttt{dims} argument. Note that in these
+calls to the \texttt{plot2D} function, we have omitted the
+\texttt{fcol} argument and use of the default \texttt{"markers"}
+feature variable to annotate the plot. We choose to display PCs 1
+and 7 to illustrate that while upper principal components explain
+much less variability in the data (2.23\% for PC7, as opposed to
+48.41\% for PC1), we see that the mitochondrial (purple) and
+peroxisome (dark blue) clusters can be differentiated, despite the
+apparent overlap in the visualisation of the two first PCs (Figure
+\ref{fig:plotmarkers}).  This is further highlighted if we plot the
+profiles of these two clusters using the \texttt{plotDist} function
+(Fig \ref{fig:plotDist2}).
 
@@
   \label{fig:plotmarkers}
+\end{figure}
+
+\begin{figure}[!ht]
+  \centering
+<<plotDist, out.width=".6\\textwidth", fig.asp=3/5>>=
+hlo <- hl[, order(hl$Iodixonal.Density)]
+plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
+         pcol = "purple", fractions = "Tag")
+title(main = "Marker occupancy profiles along the gradient")
+matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
+         lty = 1, col = "darkblue", type = "l")
+legend("topleft", c("Mitochondrion", "Peroxisome"),
+       lty = 1, col = c("purple", "blue"), bty = "n")
+@  
+  \caption{Mitochondrion and peroxisome protein profiles.}
+  \label{fig:plotDist2}
 \end{figure}
@@
 \ref{fig:plotmarkers}.
-
-\begin{figure}[!ht]
-  \centering
-<<plotDist, out.width=".6\\textwidth", fig.asp=3/5>>=
-hlo <- hl[, order(hl$Iodixonal.Density)]
-plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
-         pcol = "purple", fractions = "Tag")
-title(main = "Marker occupancy profiles along the gradient")
-matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
-         lty = 1, col = "darkblue", type = "l")
-legend("topleft", c("Mitochondrion", "Peroxisome"),
-       lty = 1, col = c("purple", "blue"), bty = "n")
-@  
-  \caption{Mitochondrion and peroxisome protein profiles.}
-  \label{fig:plotDist2}
-\end{figure}
 

### 2018-03-26 replies-to-reviewers.tex
Started reply to reviewers

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fixltx2e}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{marvosym}
\usepackage{wasysym}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{xcolor}
\setlength{\parindent}{0pt}
\geometry{top = 0.5in, bottom = 0.5 in, left = 0.5 in, right = 0.5 in}
\tolerance=1000
\setcounter{secnumdepth}{0}
\author{Lisa Breckels and Laurent Gatto}
\date{\today}

\title{A Bioconductor workflow for processing and analysing spatial proteomics data \\ Reply to reviewers}

\begin{document}

\maketitle

\section*{Reviewer 1 - Daniel J. Stekhoven}

We thank reviewer 1 for their comments. Please find our responses to
these inset below.

\begin{quote} \textcolor{gray}{
Next to reducing the dimensions of data for visualisation, PCA also offers a way to understand how the variability is distributed across the multidimensional data by providing linear combinations of the variables which then constitute the actual PCs. On that note it would be nice to mention this in Visualising markers section on page 16, where PC7 explains not much variability but due to the correct weighing of the variables we do get a separation between mitochondrial and peroxisome. This then can be further motivated with Figure 9 - where we probably can see that the weights for the fractions where the two localisations differ are larger than otherwise.}
\end{quote}

We have added a paragraph reminding users what PCA does in the Visualising Markers section to motivate the choice of looking at PC's 1 and 7. Figure 9 (now Figure 8) and corresponding code has been moved to this section and explanation of the plotDist function to lead on from the reference to marker profiles and separation.

\begin{quote}  \textcolor{gray}{
}
\end{quote}

We thank the reviewer for their constructive and positive comments.

\section*{Reviewer 2: Leonard J. Foster}

\begin{quote}\textcolor{gray}{ }
\end{quote}

\end{document}

### 2018-03-26 bioc-workflow_march2017.Rnw
Started reply to reviewers

@@
 apparent overlap in the visualisation of the two first PCs (Figure
-\ref{fig:plotmarkers}).  This is further highlighted if we plot the
+\ref{fig:plotmarkers}).  \hl{This is further highlighted if we plot the
 profiles of these two clusters using the \texttt{plotDist} function
-(Fig \ref{fig:plotDist2}).
+(Fig \ref{fig:plotDist2}).} \hl{The \texttt{plotDist} function is
+another useful visualisation that relies on marker annotation. It
+allows one to represent the protein profiles occupancy along the
+gradient \sout{using the \texttt{plotDist} function}. While the PCA
+plot enables efficient visualisation of the complete dataset and
+assessment the relative separation of different sub-cellular niches,
+comparing profiles of a few marker clusters is useful to assess how
+exactly they differ (in terms of peak channels, for example). On
+figure \ref{fig:plotDist2}, we plot the profile of the mitochondrial
+and peroxisome markers to highlight the differences in profiles
+between these two sets of markers in the channels labelled with tag
+129C, as represented above along the 7th PC on the PCA plot on
+figure \ref{fig:plotmarkers}.}
 
@@
 @
-
-Another useful visualisation that relies on marker annotation is the
-representation of the protein profiles occupancy along the gradient
-using the \texttt{plotDist} function. While the PCA plot enables
-efficient visualisation of the complete dataset and assessment the
-relative separation of different sub-cellular niches, comparing
-profiles of a few marker clusters is useful to assess how exactly they
-differ (in terms of peak channels, for example). On figure
-\ref{fig:plotDist2}, we plot the profile of the mitochondrial and
-peroxisome markers to highlight the differences in profiles between
-these two sets of markers in the channels labelled with tag 129C, as
-represented above along the 7th PC on the PCA plot on figure
-\ref{fig:plotmarkers}.
 

### 2018-03-26 replies-to-reviewers.tex
Added reply to Daniels points on parallelisation

@@
 \usepackage[T1]{fontenc}
-\usepackage{fixltx2e}
-\usepackage{graphicx}
-\usepackage{longtable}
-\usepackage{float}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
-\usepackage{marvosym}
-\usepackage{wasysym}
-\usepackage{amssymb}
-\usepackage{hyperref}
+\usepackage{listings}
 \usepackage{geometry}
@@
 \tolerance=1000
-\setcounter{secnumdepth}{0}
 \author{Lisa Breckels and Laurent Gatto}
@@
 
-\begin{quote} \textcolor{gray}{
-Next to reducing the dimensions of data for visualisation, PCA also offers a way to understand how the variability is distributed across the multidimensional data by providing linear combinations of the variables which then constitute the actual PCs.
-On that note it would be nice to mention this in Visualising markers section on page 16, where PC7 explains not much variability but due to the correct weighing of the variables we do get a separation between mitochondrial and peroxisome.
-This then can be further motivated with Figure 9 - where we probably can see that the weights for the fractions where the two localisations differ are larger than otherwise.}
-\end{quote}
+\begin{quote} \textcolor{gray}{ Next to reducing the dimensions of
+data for visualisation, PCA also offers a way to understand how the
+variability is distributed across the multidimensional data by
+providing linear combinations of the variables which then constitute
+the actual PCs. On that note it would be nice to mention this in
+Visualising markers section on page 16, where PC7 explains not much
+variability but due to the correct weighing of the variables we do
+get a separation between mitochondrial and peroxisome. This then can
+be further motivated with Figure 9 - where we probably can see that
+the weights for the fractions where the two localisations differ are
+larger than otherwise.} \end{quote}
 
-We have added a paragraph reminding users what PCA does in the Visualising Markers section to motivate the choice of looking at PC's 1 and 7.
-Figure 9 (now Figure 8) and corresponding code has been moved to this section and explanation of the plotDist function to lead on from the reference to marker profiles and separation.
+We have added a paragraph reminding users what PCA does in the
+Visualising Markers section to motivate the choice of looking at
+PC's 1 and 7. Figure 9 (now Figure 8), the corresponding code and an
+explanation of the plotDist function has been moved to this section
+to lead on from the reference to marker profiles and separation.
 
-\begin{quote}  \textcolor{gray}{
-}
-\end{quote}
+\begin{quote}  \textcolor{gray}{ I was unable to reproduce Figure 13
+comparing the two MSnSets. While I was able to look at each set
+separately using pRolocVis(hllst@x[[I]]), where i is 1 or 2, I only
+got an error using the code from the manuscript. When using
+?remap=FALSE? it actually works, but since this makes barely sense
+it is of no use - but just as a hint at debugging it.  } \end{quote}
+
+This should be fixed in the latest version of \texttt{pRolocGUI}.
+
+\begin{quote}  \textcolor{gray}{ You really need to make the results
+from the phenoDisco classification available too. It is super
+disappointing that one cannot continue reproducing the code from
+page 23 on, because it takes 24 hours to compute it using 40 cores?
+} \end{quote}
+
+The results are available and stored in `pRolocdata` for users. This
+is what is called in the manuscript under the hood:
+
+\begin{lstlisting}[language=R] extdatadir <- system.file("extdata",
+package = "pRolocdata") csvfile <- dir(extdatadir, full.names =
+TRUE, pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv") f0 <-
+dir(extdatadir, full.names = TRUE, pattern = "bpw-pdres.rds") pdres
+<- readRDS(f0) hl <- addMarkers(hl, pdres, mcol = "pd", verbose =
+FALSE) \end{lstlisting}
+
+Please note that we already say in the workflow: "Note: We do not
+evaluate this code chunk in this document as the algorithm is
+computationally intensive and best parallelised over multiple
+workers. This phenoDisco analysis took 24 hours to complete when
+parallelised over 40 workers." \newline \\ However, we have moved
+this statement to the end of the section for clarity. We state the
+output of these analyses is readily available for readers and stored
+in the \texttt{pd} column of the \texttt{MSnSet} object called
+\texttt{hl}. Readers can examine these results and get an idea of
+the type of output one can expect from this a \texttt{phenoDisco}
+analysis.  \newline \\ Specifically we have added the following to
+the end of the section to make this clear.  \newline \\ "Please
+note, in this document we can not evaluate the call to
+\texttt{phenoDisco} in the code chunk above as the algorithm is
+computationally intensive and best parallelised over multiple
+workers. This phenoDisco analysis took ~24 hours to complete when
+parallelised over 40 workers.  The ouput of running the
+\texttt{phenoDisco} algorithm is an \texttt{MSnSet} containing the
+new data clusters, appended to the \texttt{featureData} under the
+name \texttt{pd}. We have made the results readily available for
+users to interrogate and get an idea of the type of output one may
+gain from using this function. The results can be displayed by using
+the \texttt{getMarkers} function. We see that 5 new phenotype data
+clusters were found."
+
+\begin{quote}  \textcolor{gray}{ The above comment is of course also
+true for the KNN TL Optimisation on page 33 - this needs to be
+downloadable, since not everyone has access to Cambridge?s HPC and
+probably even less have 76 hours to spare.} \end{quote}
+
+The same as for the phenoDisco analysis and svm, the TL results are
+stored as a RDS in pRolocdata and loaded in the backgroun
 
@@
 
-\begin{quote}\textcolor{gray}{ }
-\end{quote}
+\begin{quote}\textcolor{gray}{ } \end{quote}
 

### 2018-03-26 bioc-workflow_march2017.Rnw
Added reply to Daniels points on parallelisation

@@
 
-Note: We do not evaluate this code chunk in this document as the
+\sout{Note: We do not evaluate this code chunk in this document as the
 algorithm is computationally intensive and best parallelised over
 multiple workers. This phenoDisco analysis took ~24 hours to complete
-when parallelised over 40 workers.
+when parallelised over 40 workers.}
 
@@
 \texttt{times}, \texttt{G} (among other function arguments, see
-\texttt{?phenoDisco}) will heavily influence (increase) the total time
-taken to run the algorithm.  \texttt{phenoDisco} supports
+\texttt{?phenoDisco}) will heavily influence (increase) the total
+time taken to run the algorithm.  \texttt{phenoDisco} supports
 parallelisation and we strongly suggest you make use of a parallel
 processing to run these analyses.
+
+\hl{Please note, in this document we can not evaluate the call to 
+`phenoDisco` in the code chunk above as the algorithm is computationally 
+intensive and best parallelised over multiple workers. This phenoDisco 
+analysis took ~24 hours to complete when parallelised over 40 workers. 
+The ouput of running the \texttt{phenoDisco} algorithm is an
+\texttt{MSnSet} containing the new data clusters, appended to the
+\texttt{featureData} under the name \texttt{pd}. We have made the results
+readily available for users to interrogate and get an idea of the type
+of output one may gain from using this function.} The results can be
+displayed by using the \texttt{getMarkers} function. We see that 5 new
+phenotype data clusters were found.
 
@@
 @
-
-The ouput of running the \texttt{phenoDisco} algorithm is an
-\texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. The results can be
-displayed by using the \texttt{getMarkers} function. We see that 5 new
-phenotype data clusters were found.
 

### 2018-04-03 main.tex
add f1000 files

@@
 
-\documentclass[10pt,a4paper,twocolumn]{article}
+%% \documentclass[10pt,a4paper,twocolumn]{article}
+\documentclass[10pt,a4paper]{article}
 \usepackage{f1000_styles}
@@
 
-%% Uncomment this lines for superscript citations instead
+%% Uncomment this lines for superscript citations insteadl
 % \usepackage[super]{natbib}
@@
 \usepackage{hyperref}
+\usepackage{framed}
+\usepackage{graphicx}
+\usepackage{xcolor}
+\usepackage{tikz}
+\usepackage{knitr}
+\usepackage{alltt}
+
+\definecolor{fgcolor}{rgb}{.5,.6,.7}
+
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}
+
+\newcommand{\Rpackage}[1]{\textsl{#1}}
+\newcommand\Biocpkg[1]{%
+  {\href{http://bioconductor.org/packages/#1}%
+    {\Rpackage{#1}}}}
+\newcommand\Biocannopkg[1]{\Biocpkg{#1}}
+\newcommand\Biocexptpkg[1]{\Biocpkg{#1}}
 
@@
 
-\title{\textit{F1000Research} Article Template}
-\titlenote{The title should be detailed enough for someone to know whether the article would be of interest to them, but also concise.
-Please ensure the broadness and claims within the title are appropriate to the content of the article itself.}
+\title{A Bioconductor workflow for processing and analysing spatial proteomics data}
+
+% \titlenote{The title should be detailed enough for someone to know
+% whether the article would be of interest to them, but also
+% concise. Please ensure the broadness and claims within the title are
+% appropriate to the content of the article itself.}
+
 \author[1,2]{Lisa M. Breckels}
+\author[2]{Claire M. Mulvey}
+\author[2]{Kathryn S. Lilley}
 \author[1,2]{Laurent Gatto}
@@
 \begin{abstract}
-Abstracts should be up to 300 words and provide a succinct summary of the article.
-Although the abstract should explain why the article might be interesting, care should be taken not to inappropriately over-emphasise the importance of the work described in the article.
-Citations should not be used in the abstract, and the use of abbreviations should be minimized.
+\input{abstract}
 \end{abstract}
 
-\clearpage
+\input{bioc-workflow}
 
-\include{bioc-workflow}
+\bigskip
+
+{\small\bibliographystyle{unsrtnat}
+\bibliography{bibliography}}
 

### 2018-04-03 replies-to-reviewers.tex
started to address comments

@@
 
-\begin{lstlisting}[language=R] extdatadir <- system.file("extdata",
-package = "pRolocdata") csvfile <- dir(extdatadir, full.names =
-TRUE, pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv") f0 <-
-dir(extdatadir, full.names = TRUE, pattern = "bpw-pdres.rds") pdres
-<- readRDS(f0) hl <- addMarkers(hl, pdres, mcol = "pd", verbose =
-FALSE) \end{lstlisting}
+\begin{lstlisting}[language=R] 
+extdatadir <- system.file("extdata", package = "pRolocdata") 
+csvfile <- dir(extdatadir, full.names = TRUE, pattern = 
+                            "hyperLOPIT-SIData-ms3-rep12-intersect.csv") 
+f0 <- dir(extdatadir, full.names = TRUE, pattern = "bpw-pdres.rds") 
+pdres <- readRDS(f0) 
+hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE) 
+\end{lstlisting}
 

### 2018-04-03 bioc-workflow.Rnw
started to address comments

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Welcome to Overleaf --- just edit your article on the left,
+% and we'll compile it for you on the right. If you give 
+% someone the link to this page, they can edit at the same
+% time. See the help menu above for more info. Enjoy!
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% For more detailed article preparation guidelines, please see:
+% http://f1000research.com/author-guidelines
+
+%% \documentclass[10pt,a4paper,twocolumn]{article}
+\documentclass[10pt,a4paper]{article}
+\usepackage{f1000_styles}
+
+%% Default: numerical citations
+\usepackage[numbers]{natbib}
+
+%% Uncomment this lines for superscript citations insteadl
+% \usepackage[super]{natbib}
+
+%% Uncomment these lines for author-year citations instead
+% \usepackage[round]{natbib}
+% \let\cite\citep
+
+\usepackage{hyperref}
+\usepackage{framed}
+\usepackage{graphicx}
+\usepackage{xcolor}
+\usepackage{tikz}
+\usepackage{knitr}
+\usepackage{alltt}
+
+\definecolor{fgcolor}{rgb}{.5,.6,.7}
+
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}
+
+\newcommand{\Rpackage}[1]{\textsl{#1}}
+\newcommand\Biocpkg[1]{%
+  {\href{http://bioconductor.org/packages/#1}%
+    {\Rpackage{#1}}}}
+\newcommand\Biocannopkg[1]{\Biocpkg{#1}}
+\newcommand\Biocexptpkg[1]{\Biocpkg{#1}}
+
+\begin{document}
+\SweaveOpts{concordance=TRUE}
+
+\title{A Bioconductor workflow for processing and analysing spatial proteomics data}
+
+% \titlenote{The title should be detailed enough for someone to know
+% whether the article would be of interest to them, but also
+% concise. Please ensure the broadness and claims within the title are
+% appropriate to the content of the article itself.}
+
+\author[1,2]{Lisa M. Breckels}
+\author[2]{Claire M. Mulvey}
+\author[2]{Kathryn S. Lilley}
+\author[1,2]{Laurent Gatto}
+\affil[1]{Computational Proteomics Unit, Cambridge Systems Biology Centre, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
+\affil[2]{Cambridge Centre for Proteomics, Department of Biochemistry, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
+
+\maketitle
+\thispagestyle{fancy}
+
+\begin{abstract}
+\input{abstract}
+\end{abstract}
+
 <<env, echo=FALSE, warning=FALSE>>=
@@
 Technology Facilities Council.
+
+\bigskip
+
+{\small\bibliographystyle{unsrtnat}
+\bibliography{bibliography}}
+
+\end{document}

### 2018-04-03 bioc-workflow.tex
started to address comments

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Welcome to Overleaf --- just edit your article on the left,
+% and we'll compile it for you on the right. If you give 
+% someone the link to this page, they can edit at the same
+% time. See the help menu above for more info. Enjoy!
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% For more detailed article preparation guidelines, please see:
+% http://f1000research.com/author-guidelines
+
+%% \documentclass[10pt,a4paper,twocolumn]{article}
+\documentclass[10pt,a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
+\usepackage{f1000_styles}
+
+%% Default: numerical citations
+\usepackage[numbers]{natbib}
+
+%% Uncomment this lines for superscript citations insteadl
+% \usepackage[super]{natbib}
+
+%% Uncomment these lines for author-year citations instead
+% \usepackage[round]{natbib}
+% \let\cite\citep
+
+\usepackage{hyperref}
+\usepackage{framed}
+\usepackage{graphicx}
+\usepackage{xcolor}
+\usepackage{tikz}
+\usepackage{knitr}
+\usepackage{alltt}
+
+\definecolor{fgcolor}{rgb}{.5,.6,.7}
+
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}
+
+\newcommand{\Rpackage}[1]{\textsl{#1}}
+\newcommand\Biocpkg[1]{%
+  {\href{http://bioconductor.org/packages/#1}%
+    {\Rpackage{#1}}}}
+\newcommand\Biocannopkg[1]{\Biocpkg{#1}}
+\newcommand\Biocexptpkg[1]{\Biocpkg{#1}}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\begin{document}
+\SweaveOpts{concordance=TRUE}
+
+\title{A Bioconductor workflow for processing and analysing spatial proteomics data}
+
+% \titlenote{The title should be detailed enough for someone to know
+% whether the article would be of interest to them, but also
+% concise. Please ensure the broadness and claims within the title are
+% appropriate to the content of the article itself.}
+
+\author[1,2]{Lisa M. Breckels}
+\author[2]{Claire M. Mulvey}
+\author[2]{Kathryn S. Lilley}
+\author[1,2]{Laurent Gatto}
+\affil[1]{Computational Proteomics Unit, Cambridge Systems Biology Centre, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
+\affil[2]{Cambridge Centre for Proteomics, Department of Biochemistry, University of Cambridge, Tennis Court Road, Cambridge, CB2 1GA, UK}
+
+\maketitle
+\thispagestyle{fancy}
+
+\begin{abstract}
+\input{abstract}
+\end{abstract}
+
 \section*{Introduction}
@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.1.4
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 
-will list the 54 datasets that are
+will list the 75 datasets that are
 available in \Biocexptpkg{pRolocdata}.
@@
 \hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
-\hlstd{combined}
-\end{alltt}
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 6725 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData
-##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
-##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
-##   varMetadata: labelDescription
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
-##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
-##     total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-## Annotation:  
-## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Dec  8 12:09:01 2016 
-##  MSnbase version: 1.21.7
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-More details about combining data are given in the dedicated
-\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
-\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
-  vignette}.
-
-\subsubsection*{Missing data}
-
-Missing data are a recurrent issue in mass spectrometry applications,
-and should be addressed independently of this workflow
-\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
-described how a high content in missing values in spatial proteomics
-data and their inappropriate handling leads to a reduction of
-sub-cellular resolution. Missing data can be imputated using
-\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
-the imputation method is then determined by a \texttt{methods}
-parameter (see \texttt{?impute} for available options). To impute
-missing values using nearest neighbour imputation, one would
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-In our particular case, missing values are indicative of protein
-groups that were not acquired in both replicates (Figure
-\ref{fig:namap}).
-
-\begin{figure}[!ht]
-  \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
-       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=.65\textwidth]{figure/namap-1}
-
-\end{knitrout}
-\caption{Heatmap of missing values. Note that the features are
-  re-ordered to highlight cluster of proteins with similar numbers of
-  missing values.}
-  \label{fig:namap}
-\end{figure}
-
-We prefer to remove proteins that were not assayed in both replicated
-experiments. This is done with the \texttt{filterNA} function that
-removes features (proteins) that contain more than a certain
-proportion (default is 0) missing values. The \textit{Processing
-  information} section summarises how combining and filtering missing
-values (subsetting) changed the dimensions of the data.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
+\end{alltt}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning: 'combine' is deprecated.\\\#\# Use 'gtable\_combine' instead.\\\#\# See help("{}Deprecated"{})}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[[<-.data.frame`(`*tmp*`, i, value = structure(list(t = numeric(0), : replacement has 0 rows, data has 10}}\begin{alltt}
 \hlstd{combined}
@@
 ## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Thu Dec  8 12:09:01 2016 
-## Subset [6725,20][5032,20] Thu Dec  8 12:09:01 2016 
-## Removed features with more than 0 NAs: Thu Dec  8 12:09:01 2016 
-## Dropped featureData's levels Thu Dec  8 12:09:01 2016 
-##  MSnbase version: 1.21.7
+## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
+## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
+## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
+## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 12:00:57 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 12:00:57 2018 
+## Dropped featureData's levels Tue Apr  3 12:00:57 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 12:56:07 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 12:56:07 2018 
+## Dropped featureData's levels Tue Apr  3 12:56:07 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 12:56:42 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 12:56:42 2018 
+## Dropped featureData's levels Tue Apr  3 12:56:42 2018 
+##  MSnbase version: 2.5.9
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+More details about combining data are given in the dedicated
+\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
+\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
+  vignette}.
+
+\subsubsection*{Missing data}
+
+Missing data are a recurrent issue in mass spectrometry applications,
+and should be addressed independently of this workflow
+\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
+described how a high content in missing values in spatial proteomics
+data and their inappropriate handling leads to a reduction of
+sub-cellular resolution. Missing data can be imputated using
+\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
+the imputation method is then determined by a \texttt{methods}
+parameter (see \texttt{?impute} for available options). To impute
+missing values using nearest neighbour imputation, one would
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+In our particular case, missing values are indicative of protein
+groups that were not acquired in both replicates (Figure
+\ref{fig:namap}).
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
+       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/namap-1}
+
+\end{knitrout}
+\caption{Heatmap of missing values. Note that the features are
+  re-ordered to highlight cluster of proteins with similar numbers of
+  missing values.}
+  \label{fig:namap}
+\end{figure}
+
+We prefer to remove proteins that were not assayed in both replicated
+experiments. This is done with the \texttt{filterNA} function that
+removes features (proteins) that contain more than a certain
+proportion (default is 0) missing values. The \textit{Processing
+  information} section summarises how combining and filtering missing
+values (subsetting) changed the dimensions of the data.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
+\hlstd{combined}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 5032 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData
+##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+##   varMetadata: labelDescription
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+##     total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
+## Annotation:  
+## - - - Processing information - - -
+## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
+## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
+## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
+## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 12:00:57 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 12:00:57 2018 
+## Dropped featureData's levels Tue Apr  3 12:00:57 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 12:56:07 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 12:56:07 2018 
+## Dropped featureData's levels Tue Apr  3 12:56:07 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 12:56:42 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 12:56:42 2018 
+## Dropped featureData's levels Tue Apr  3 12:56:42 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 13:37:03 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 13:37:03 2018 
+## Dropped featureData's levels Tue Apr  3 13:37:03 2018 
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Thu Dec  8 12:09:03 2016 
+##  Created on Tue Apr  3 13:37:05 2018 
 ##  Description:
@@
 
-% latex table generated in R 3.3.2 by xtable 1.8-2 package
-% Thu Dec  8 12:09:03 2016
+% latex table generated in R 3.5.0 by xtable 1.8-2 package
+% Tue Apr  3 13:37:05 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Thu Dec  8 12:09:03 2016 
-## Added markers from  'pdres' marker vector. Thu Dec  8 12:09:04 2016 
-##  MSnbase version: 2.1.4
+## Added markers from  'mrk' marker vector. Tue Apr  3 13:37:04 2018 
+## Added markers from  'pdres' marker vector. Tue Apr  3 13:37:06 2018 
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 \end{alltt}
-\end{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in seq\_len(trpr): first element used of 'length.out' argument}}\end{kframe}
 \end{knitrout}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4322999               0.3055155               0.3838354 
+##               0.4401665               0.3022831               0.3793515 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6883460               0.6037025               0.4275289 
+##               0.6907964               0.6076611               0.4275322 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4276601               0.5906588               0.9501307 
+##               0.4235684               0.5973413               0.9495950 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7947781               0.7105373               0.3137337 
+##               0.7960356               0.7115733               0.3153275 
 ##         Plasma membrane              Proteasome 
-##               0.7191668               0.4144056
+##               0.7118966               0.4169883
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4322999               0.3055155               0.3838354 
+##               0.4401665               0.3022831               0.3793515 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6883460               0.6037025               0.4275289 
+##               0.6907964               0.6076611               0.4275322 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4276601               0.5906588               0.9501307 
+##               0.4235684               0.5973413               0.9495950 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7947781               0.7105373               0.3137337 
+##               0.7960356               0.7115733               0.3153275 
 ##         Plasma membrane              Proteasome 
-##               0.7191668               0.4144056
+##               0.7118966               0.4169883
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     171                      88 
+##                      85                     170                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     478                     100 
+##                     296                     473                     100 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     522 
+##                      28                     123                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     342                      39 
+##                     230                     344                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     316                     159                    2058
+##                     321                     159                    2058
 \end{verbatim}
@@
 \begin{verbatim}
-## R version 3.3.2 Patched (2016-11-01 r71616)
-## Platform: x86_64-pc-linux-gnu (64-bit)
-## Running under: Ubuntu 14.04.5 LTS
+## R Under development (unstable) (2018-03-21 r74432)
+## Platform: x86_64-apple-darwin15.6.0 (64-bit)
+## Running under: macOS High Sierra 10.13.3
+## 
+## Matrix products: default
+## BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
+## LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib
 ## 
 ## locale:
-##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
-##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
-##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
-##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
-##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
-## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
+## [1] en_GB.UTF-8/en_GB.UTF-8/en_GB.UTF-8/C/en_GB.UTF-8/en_GB.UTF-8
 ## 
 ## attached base packages:
-## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
-## [8] datasets  base     
+## [1] stats4    parallel  stats     graphics  grDevices utils     datasets 
+## [8] methods   base     
 ## 
 ## other attached packages:
-##  [1] pRolocdata_1.12.0    pRoloc_1.15.5        MLInterfaces_1.54.0 
-##  [4] cluster_2.0.5        annotate_1.52.0      XML_3.98-1.5        
-##  [7] AnnotationDbi_1.36.0 IRanges_2.8.1        S4Vectors_0.12.1    
-## [10] MSnbase_2.1.4        ProtGenerics_1.6.0   BiocParallel_1.8.1  
-## [13] mzR_2.8.0            Rcpp_0.12.8          Biobase_2.34.0      
-## [16] BiocGenerics_0.20.0  gridExtra_2.2.1      xtable_1.8-2        
-## [19] BiocStyle_2.2.1      knitr_1.15.1        
+##  [1] pRolocdata_1.17.2    gridExtra_2.3        xtable_1.8-2        
+##  [4] BiocStyle_2.7.8      knitr_1.20           rgl_0.99.9          
+##  [7] pRolocGUI_1.13.1     pRoloc_1.19.3        MLInterfaces_1.59.1 
+## [10] cluster_2.0.6        annotate_1.57.2      XML_3.98-1.10       
+## [13] AnnotationDbi_1.41.4 IRanges_2.13.28      S4Vectors_0.17.37   
+## [16] MSnbase_2.5.9        ProtGenerics_1.11.0  BiocParallel_1.13.1 
+## [19] mzR_2.13.5           Rcpp_0.12.16         Biobase_2.39.2      
+## [22] BiocGenerics_0.25.3 
 ## 
 ## loaded via a namespace (and not attached):
-##   [1] minqa_1.2.4           colorspace_1.3-1      hwriter_1.3.2        
-##   [4] class_7.3-14          modeltools_0.2-21     mclust_5.2           
-##   [7] pls_2.5-0             base64enc_0.1-3       proxy_0.4-16         
-##  [10] hexbin_1.27.1         MatrixModels_0.4-1    affyio_1.44.0        
-##  [13] flexmix_2.3-13        mvtnorm_1.0-5         codetools_0.2-15     
-##  [16] splines_3.3.2         doParallel_1.0.10     impute_1.48.0        
-##  [19] robustbase_0.92-6     jsonlite_1.1          nloptr_1.0.4         
-##  [22] caret_6.0-73          pbkrtest_0.4-6        rda_1.0.2-2          
-##  [25] kernlab_0.9-25        vsn_3.42.3            sfsmisc_1.1-0        
-##  [28] shiny_0.14.2          sampling_2.7          assertthat_0.1       
-##  [31] Matrix_1.2-7.1        lazyeval_0.2.0        limma_3.30.6         
-##  [34] htmltools_0.3.5       quantreg_5.29         tools_3.3.2          
-##  [37] ggvis_0.4.3           gtable_0.2.0          affy_1.52.0          
-##  [40] reshape2_1.4.2        dplyr_0.5.0           MALDIquant_1.15      
-##  [43] trimcluster_0.1-2     gdata_2.17.0          preprocessCore_1.36.0
-##  [46] nlme_3.1-128          iterators_1.0.8       fpc_2.1-10           
-##  [49] stringr_1.1.0         lme4_1.1-12           lpSolve_5.6.13       
-##  [52] mime_0.5              gtools_3.5.0          dendextend_1.3.0     
-##  [55] DEoptimR_1.0-8        zlibbioc_1.20.0       MASS_7.3-45          
-##  [58] scales_0.4.1          BiocInstaller_1.24.0  pcaMethods_1.66.0    
-##  [61] SparseM_1.74          RColorBrewer_1.1-2    memoise_1.0.0        
-##  [64] ggplot2_2.2.0         biomaRt_2.30.0        rpart_4.1-10         
-##  [67] stringi_1.1.2         RSQLite_1.1           highr_0.6            
-##  [70] genefilter_1.56.0     randomForest_4.6-12   foreach_1.4.3        
-##  [73] e1071_1.6-7           prabclus_2.2-6        bitops_1.0-6         
-##  [76] mzID_1.12.0           evaluate_0.10         lattice_0.20-34      
-##  [79] htmlwidgets_0.8       gbm_2.1.1             plyr_1.8.4           
-##  [82] magrittr_1.5          R6_2.2.0              DBI_0.5-1            
-##  [85] whisker_0.3-2         mgcv_1.8-16           survival_2.40-1      
-##  [88] RCurl_1.95-4.8        nnet_7.3-12           tibble_1.2           
-##  [91] msdata_0.14.0         car_2.1-3             mlbench_2.1-1        
-##  [94] grid_3.3.2            FNN_1.1               ModelMetrics_1.1.0   
-##  [97] threejs_0.2.2         digest_0.6.10         diptest_0.75-7       
-## [100] httpuv_1.3.3          munsell_0.4.3
+##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
+##   [4] lazyeval_0.2.1        splines_3.5.0         ggvis_0.4.3          
+##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
+##  [10] foreach_1.4.4         BiocInstaller_1.29.4  htmltools_0.3.6      
+##  [13] viridis_0.5.0         gdata_2.18.0          magrittr_1.5         
+##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
+##  [19] limma_3.35.13         recipes_0.1.2         gower_0.1.2          
+##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
+##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.0           
+##  [28] dplyr_0.7.4           tcltk_3.5.0           jsonlite_1.5         
+##  [31] RCurl_1.95-4.10       hexbin_1.27.2         genefilter_1.61.1    
+##  [34] bindr_0.1.1           impute_1.53.0         survival_2.41-3      
+##  [37] iterators_1.0.9       glue_1.2.0            DRR_0.0.3            
+##  [40] gtable_0.2.0          ipred_0.9-6           zlibbioc_1.25.0      
+##  [43] kernlab_0.9-25        ddalpha_1.3.1.1       prabclus_2.2-6       
+##  [46] DEoptimR_1.0-8        scales_0.5.0          vsn_3.47.1           
+##  [49] mvtnorm_1.0-7         DBI_0.8               viridisLite_0.3.0    
+##  [52] progress_1.1.2        foreign_0.8-70        bit_1.1-12           
+##  [55] proxy_0.4-21          mclust_5.4            preprocessCore_1.41.0
+##  [58] DT_0.4                lava_1.6              prodlim_1.6.1        
+##  [61] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
+##  [64] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
+##  [67] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
+##  [70] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
+##  [73] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
+##  [76] munsell_0.4.3         mlbench_2.1-1         tools_3.5.0          
+##  [79] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
+##  [82] evaluate_0.10.1       stringr_1.3.0         yaml_2.1.18          
+##  [85] mzID_1.17.0           ModelMetrics_1.1.0    bit64_0.9-7          
+##  [88] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
+##  [91] dendextend_1.7.0      bindrcpp_0.2          nlme_3.1-136         
+##  [94] whisker_0.3-2         mime_0.5              RcppRoll_0.2.2       
+##  [97] biomaRt_2.35.12       compiler_3.5.0        e1071_1.6-8          
+## [100] affyio_1.49.1         tibble_1.4.2          stringi_1.1.7        
+## [103] highr_0.6             lattice_0.20-35       trimcluster_0.1-2    
+## [106] Matrix_1.2-12         psych_1.7.8           gbm_2.1.3            
+## [109] pillar_1.2.1          MALDIquant_1.17       bitops_1.0-6         
+## [112] httpuv_1.3.6.2        R6_2.2.2              pcaMethods_1.71.0    
+## [115] affy_1.57.0           hwriter_1.3.2         codetools_0.2-15     
+## [118] MASS_7.3-49           gtools_3.5.0          assertthat_0.2.0     
+## [121] CVST_0.2-1            rprojroot_1.3-2       withr_2.1.2          
+## [124] mnormt_1.5-5          diptest_0.75-7        grid_3.5.0           
+## [127] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
+## [130] class_7.3-14          rmarkdown_1.9         shiny_1.0.5          
+## [133] lubridate_1.7.3       base64enc_0.1-3
 \end{verbatim}

### 2018-04-03 bioc-workflow_march2017.Rnw
started to address comments

@@
 reducing purely technical variation between channels without affecting
-biological variability (i.e. the shape of the quantitative
+biological variability (i.e. the shape of the quantitatives
 profiles). This normalisation will depend on the underlying
@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into a same intensity interval (typically 0 and 1) by, for
+profiles into \sout{a}\hl{the} same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
 data and their inappropriate handling leads to a reduction of
-sub-cellular resolution. Missing data can be imputated using
+sub-cellular resolution. \sout{Missing data can be imputated} \hl{We can impute missing data} using
 \Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
@@
 \caption{Heatmap of missing values. Note that the features are
-  re-ordered to highlight cluster of proteins with similar numbers of
+  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
   missing values.}
@@
 experiments) they would first create a \texttt{FeaturesOfInterest}
-object. For example, if we wanted to highlight a proteins with the
-accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, which are some
-of the proteins known to form part of the 13S condensin complex, we
-would create a first create a \texttt{FeaturesOfInterest} object, and
+object \hl{and then use the \textit{highlightOnPlot} function to 
+visualise these}. For example, if we wanted to highlight \sout{a} 
+proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, 
+which are some of the proteins known to form part of the 13S condensin complex, 
+we \hl{would call the following code:}
+\sout{would create a first create a \texttt{FeaturesOfInterest} object, and
 subsequently highlight their location on a PCA plot with the
-\texttt{highlightOnPlot} function.
+\texttt{highlightOnPlot} function.}
 
@@
 readily available for users to interrogate and get an idea of the type
-of output one may gain from using this function.} The results can be
+of output one may expect from using this function.} The results can be
 displayed by using the \texttt{getMarkers} function. We see that 5 new
@@
 
-% <<vishl, eval=FALSE>>=
-% pRolocVis(hl, fcol = "pd")
-% @
+<<vishl, eval=FALSE>>=
+pRolocVis(hl, fcol = "pd")
+@
 
@@
 par <- setAnnotationParams(inputs = c("Mus musculus", "UniProt/Swissprot"))
-
 @

### 2018-04-03 bioc-workflow_march2017.tex
started to address comments

--- 2018-04-03 bioc-workflow_march2017.Rnw
@@
-<<env, echo=FALSE, warning=FALSE>>=
-library("BiocStyle")
-suppressPackageStartupMessages(library("knitr"))
-suppressPackageStartupMessages(library("xtable"))
-suppressPackageStartupMessages(library("gridExtra"))
-suppressPackageStartupMessages(library("MSnbase"))
-suppressPackageStartupMessages(library("pRoloc"))
-suppressPackageStartupMessages(library("pRolocdata"))
-cache <- FALSE
-@
-
 \section*{Introduction}
@@
 
-<<install, eval=FALSE>>=
-source("https://bioconductor.org/biocLite.R")
-biocLite(c("MSnase", "pRoloc", "pRolocdata", "pRolocGUI"))
-@
+\begin{Schunk}
+\begin{Sinput}
+> source("https://bioconductor.org/biocLite.R")
+> biocLite(c("MSnase", "pRoloc", "pRolocdata", "pRolocGUI"))
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<getFilename, cache = cache>>=
-library("MSnbase")
-extdatadir <- system.file("extdata", package = "pRolocdata")
-csvfile <- dir(extdatadir, full.names = TRUE,
-          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
-basename(csvfile)
-@
+\begin{Schunk}
+\begin{Sinput}
+> library("MSnbase")
+> extdatadir <- system.file("extdata", package = "pRolocdata")
+> csvfile <- dir(extdatadir, full.names = TRUE,
++           pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
+> basename(csvfile)
+\end{Sinput}
+\begin{Soutput}
+[1] "hyperLOPIT-SIData-ms3-rep12-intersect.csv.gz"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<getEcols>>=
-getEcols(csvfile, split = ",", n = 2)
-@
+\begin{Schunk}
+\begin{Sinput}
+> getEcols(csvfile, split = ",", n = 2)
+\end{Sinput}
+\begin{Soutput}
+ [1] ""                                  ""                                 
+ [3] ""                                  "Experiment 1"                     
+ [5] "Experiment 2"                      "Experiment 1"                     
+ [7] "Experiment 2"                      "126"                              
+ [9] "127N"                              "127C"                             
+[11] "128N"                              "128C"                             
+[13] "129N"                              "129C"                             
+[15] "130N"                              "130C"                             
+[17] "131"                               "126"                              
+[19] "127N"                              "127C"                             
+[21] "128N"                              "128C"                             
+[23] "129N"                              "129C"                             
+[25] "130N"                              "130C"                             
+[27] "131"                               "phenoDisco Input"                 
+[29] "phenoDisco Output"                 "Curated phenoDisco Output"        
+[31] "SVM marker set"                    "SVM classification"               
+[33] "SVM score"                         "SVM classification (top quartile)"
+[35] "Final Localization Assignment"     "First localization evidence?"     
+[37] "Curated Organelles"                "Cytoskeletal Components"          
+[39] "Trafficking Proteins"              "Protein Complexes"                
+[41] "Signaling Cascades"                "Oct4 Interactome"                 
+[43] "Nanog Interactome"                 "Sox2 Interactome"                 
+[45] "Cell Surface Proteins"            
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<readMSnSet2>>=
-hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
-@
+\begin{Schunk}
+\begin{Sinput}
+> hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
+\end{Sinput}
+\end{Schunk}
 
 Below, we display a short summary of the data. The data contains 
-\Sexpr{nrow(hl)} proteins/features common across the 2 biological replicates
-for the respective 2 x 10-plex reporter tags (\Sexpr{ncol(hl)}
+5032 proteins/features common across the 2 biological replicates
+for the respective 2 x 10-plex reporter tags (20
 columns or samples), along with associated feature meta-data such as
@@
 
-<<showhl>>=
-hl
-@
+\begin{Schunk}
+\begin{Sinput}
+> hl
+\end{Sinput}
+\begin{Soutput}
+MSnSet (storageMode: lockedEnvironment)
+assayData: 5032 features, 20 samples 
+  element names: exprs 
+protocolData: none
+phenoData: none
+featureData
+  featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+  fvarLabels: X X.1 ... Cell.Surface.Proteins (25 total)
+  fvarMetadata: labelDescription
+experimentData: use 'experimentData(object)'
+Annotation:  
+- - - Processing information - - -
+ MSnbase version: 2.5.9 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<assesshl>>=
-exprs(hl)[1:5, ]
-exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
-@
+\begin{Schunk}
+\begin{Sinput}
+> exprs(hl)[1:5, ]
+\end{Sinput}
+\begin{Soutput}
+          X126 X127N X127C X128N X128C X129N X129C X130N X130C  X131 X126.1
+Q9JHU4   0.028 0.034 0.024 0.014 0.026 0.045 0.107 0.341 0.059 0.321  0.037
+Q9QXS1-3 0.039 0.134 0.095 0.053 0.084 0.121 0.107 0.128 0.122 0.117  0.033
+Q9ERU9   0.021 0.013 0.014 0.009 0.024 0.054 0.116 0.257 0.209 0.284  0.026
+P26039   0.120 0.255 0.148 0.091 0.135 0.095 0.041 0.057 0.014 0.043  0.111
+Q8BTM8   0.055 0.139 0.078 0.050 0.077 0.098 0.093 0.171 0.079 0.160  0.062
+         X127N.1 X127C.1 X128N.1 X128C.1 X129N.1 X129C.1 X130N.1 X130C.1 X131.1
+Q9JHU4     0.064   0.058   0.059   0.067   0.078   0.140   0.208   0.141  0.147
+Q9QXS1-3   0.073   0.074   0.062   0.081   0.142   0.190   0.069   0.151  0.125
+Q9ERU9     0.017   0.023   0.029   0.039   0.071   0.105   0.171   0.304  0.215
+P26039     0.181   0.141   0.144   0.152   0.119   0.075   0.028   0.017  0.033
+Q8BTM8     0.108   0.091   0.086   0.099   0.111   0.117   0.095   0.144  0.087
+\end{Soutput}
+\begin{Sinput}
+> exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
+\end{Sinput}
+\begin{Soutput}
+        X126 X131.1
+Q9ERU9 0.021  0.215
+Q9Z2R6 0.563  0.000
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<removefData>>=
-fvarLabels(hl)
-fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
-fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
-## feature vars 1, 2, and 4 to 6
-fData(hl)[1:4, c(1:2, 4:6)]
-@
+\begin{Schunk}
+\begin{Sinput}
+> fvarLabels(hl)
+\end{Sinput}
+\begin{Soutput}
+ [1] "X"                                 "X.1"                              
+ [3] "X.2"                               "Experiment.1"                     
+ [5] "Experiment.2"                      "Experiment.1.1"                   
+ [7] "Experiment.2.1"                    "phenoDisco.Input"                 
+ [9] "phenoDisco.Output"                 "Curated.phenoDisco.Output"        
+[11] "SVM.marker.set"                    "SVM.classification"               
+[13] "SVM.score"                         "SVM.classification..top.quartile."
+[15] "Final.Localization.Assignment"     "First.localization.evidence."     
+[17] "Curated.Organelles"                "Cytoskeletal.Components"          
+[19] "Trafficking.Proteins"              "Protein.Complexes"                
+[21] "Signaling.Cascades"                "Oct4.Interactome"                 
+[23] "Nanog.Interactome"                 "Sox2.Interactome"                 
+[25] "Cell.Surface.Proteins"            
+\end{Soutput}
+\begin{Sinput}
+> fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
+> fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
+> ## feature vars 1, 2, and 4 to 6
+> fData(hl)[1:4, c(1:2, 4:6)]
+\end{Sinput}
+\begin{Soutput}
+         uniprot.accession  uniprot.id peptides.expt1 peptides.expt2
+Q9JHU4              Q9JHU4 DYHC1_MOUSE            175            166
+Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE            123            150
+Q9ERU9              Q9ERU9  RBP2_MOUSE            101             90
+P26039              P26039  TLN1_MOUSE            101             94
+         peptides.expt3
+Q9JHU4              322
+Q9QXS1-3            174
+Q9ERU9              181
+P26039              167
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<pdata>>=
-pData(hl)$Replicate <- rep(1:2, each = 10)
-pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
-pData(hl)
-@
+\begin{Schunk}
+\begin{Sinput}
+> pData(hl)$Replicate <- rep(1:2, each = 10)
+> pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
+> pData(hl)
+\end{Sinput}
+\begin{Soutput}
+        Replicate  Tag
+X126            1  126
+X127N           1 127N
+X127C           1 127C
+X128N           1 128N
+X128C           1 128C
+X129N           1 129N
+X129C           1 129C
+X130N           1 130N
+X130C           1 130C
+X131            1  131
+X126.1          2  126
+X127N.1         2 127N
+X127C.1         2 127C
+X128N.1         2 128N
+X128C.1         2 128C
+X129N.1         2 129N
+X129C.1         2 129C
+X130N.1         2 130N
+X130C.1         2 130C
+X131.1          2  131
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<morepdata>>=
-expinfo <- dir(extdatadir, full.names = TRUE,
-               pattern = "hyperLOPIT-SIData-fraction-info.csv")
-
-fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
-                     header = FALSE, stringsAsFactors = FALSE)
-
-pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
-pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
-pData(hl)
-@
+\begin{Schunk}
+\begin{Sinput}
+> expinfo <- dir(extdatadir, full.names = TRUE,
++                pattern = "hyperLOPIT-SIData-fraction-info.csv")
+> fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
++                      header = FALSE, stringsAsFactors = FALSE)
+> pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
+> pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
+> pData(hl)
+\end{Sinput}
+\begin{Soutput}
+        Replicate  Tag Gradient.Fraction Iodixonal.Density
+X126            1  126           Cytosol               0.0
+X127N           1 127N   1 to 6 (pooled)               6.0
+X127C           1 127C   8 to 9 (pooled)              11.0
+X128N           1 128N 10 to 11 (pooled)              13.3
+X128C           1 128C                12              14.6
+X129N           1 129N                14              17.4
+X129C           1 129C                16              20.1
+X130N           1 130N                18              26.8
+X130C           1 130C         Chromatin                NA
+X131            1  131                19              34.5
+X126.1          2  126           Cytosol               0.0
+X127N.1         2 127N   1 to 6 (pooled)               5.2
+X127C.1         2 127C   7 to 9 (pooled)              10.0
+X128N.1         2 128N 10 to 11 (pooled)              12.5
+X128C.1         2 128C                12              14.0
+X129N.1         2 129N 14 to 15 (pooled)              17.3
+X129C.1         2 129C                17              20.9
+X130N.1         2 130N 18 to 19 (pooled)              24.7
+X130C.1         2 130C         Chromatin                NA
+X131.1          2  131                20              31.9
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<normsum, eval = FALSE>>= 
-hl <- normalise(hl, method = "sum") 
-@
+\begin{Schunk}
+\begin{Sinput}
+> hl <- normalise(hl, method = "sum") 
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<normcheck>>=
-summary(rowSums(exprs(hl[, hl$Replicate == 1])))
-summary(rowSums(exprs(hl[, hl$Replicate == 2])))
-@
+\begin{Schunk}
+\begin{Sinput}
+> summary(rowSums(exprs(hl[, hl$Replicate == 1])))
+\end{Sinput}
+\begin{Soutput}
+   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+  0.997   0.999   1.000   1.000   1.001   1.003 
+\end{Soutput}
+\begin{Sinput}
+> summary(rowSums(exprs(hl[, hl$Replicate == 2])))
+\end{Sinput}
+\begin{Soutput}
+   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+  0.997   0.999   1.000   1.000   1.001   1.003 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<reps4combine>>=
-library("pRolocdata")
-data(hyperLOPIT2015ms3r1)
-data(hyperLOPIT2015ms3r2)
-@
+\begin{Schunk}
+\begin{Sinput}
+> library("pRolocdata")
+> data(hyperLOPIT2015ms3r1)
+> data(hyperLOPIT2015ms3r2)
+\end{Sinput}
+\end{Schunk}
 
@@
 
-<<pkgdata, eval=FALSE>>=
-pRolocdata()
-@
-
-will list the \Sexpr{nrow(pRolocdata()$results)} datasets that are
+\begin{Schunk}
+\begin{Sinput}
+> pRolocdata()
+\end{Sinput}
+\end{Schunk}
+
+will list the 75 datasets that are
 available in \Biocexptpkg{pRolocdata}.
@@
 
-<<samplenames>>=
-identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
-hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
-hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
-sampleNames(hyperLOPIT2015ms3r1)
-sampleNames(hyperLOPIT2015ms3r2)
-@
+\begin{Schunk}
+\begin{Sinput}
+> identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
+\end{Sinput}
+\begin{Soutput}
+[1] TRUE
+\end{Soutput}
+\begin{Sinput}
+> hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
+> hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
+> sampleNames(hyperLOPIT2015ms3r1)
+\end{Sinput}
+\begin{Soutput}
+ [1] "X126.1"  "X127N.1" "X127C.1" "X128N.1" "X128C.1" "X129N.1" "X129C.1"
+ [8] "X130N.1" "X130C.1" "X131.1" 
+\end{Soutput}
+\begin{Sinput}
+> sampleNames(hyperLOPIT2015ms3r2)
+\end{Sinput}
+\begin{Soutput}
+ [1] "X126.2"  "X127N.2" "X127C.2" "X128N.2" "X128C.2" "X129N.2" "X129C.2"
+ [8] "X130N.2" "X130C.2" "X131.2" 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<fvarnames>>=
-fvarLabels(hyperLOPIT2015ms3r1)
-fvarLabels(hyperLOPIT2015ms3r2)
-@
+\begin{Schunk}
+\begin{Sinput}
+> fvarLabels(hyperLOPIT2015ms3r1)
+\end{Sinput}
+\begin{Soutput}
+[1] "EntryName"          "ProteinDescription" "Peptides"          
+[4] "PSMs"               "ProteinCoverage"    "markers"           
+\end{Soutput}
+\begin{Sinput}
+> fvarLabels(hyperLOPIT2015ms3r2)
+\end{Sinput}
+\begin{Soutput}
+[1] "EntryName"          "ProteinDescription" "Peptides"          
+[4] "PSMs"               "ProteinCoverage"    "markers"           
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<fvarnames2>>=
-fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
-hyperLOPIT2015ms3r2 <- updateFvarLabels(hyperLOPIT2015ms3r2, "2", sep = "")
-fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5]
-fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
-fvarLabels(hyperLOPIT2015ms3r1)
-fvarLabels(hyperLOPIT2015ms3r2)
-@
+\begin{Schunk}
+\begin{Sinput}
+> fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
+> hyperLOPIT2015ms3r2 <- updateFvarLabels(hyperLOPIT2015ms3r2, "2", sep = "")
+> fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5]
+> fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
+> fvarLabels(hyperLOPIT2015ms3r1)
+\end{Sinput}
+\begin{Soutput}
+[1] "EntryName"          "ProteinDescription" "Peptides1"         
+[4] "PSMs1"              "ProteinCoverage1"  
+\end{Soutput}
+\begin{Sinput}
+> fvarLabels(hyperLOPIT2015ms3r2)
+\end{Sinput}
+\begin{Soutput}
+[1] "Peptides2"        "PSMs2"            "ProteinCoverage2"
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<combine, cache = cache>>=
-combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
-combined
-@
+\begin{Schunk}
+\begin{Sinput}
+> combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
+> combined
+\end{Sinput}
+\begin{Soutput}
+MSnSet (storageMode: lockedEnvironment)
+assayData: 6725 features, 20 samples 
+  element names: exprs 
+protocolData: none
+phenoData
+  sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+  varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+  varMetadata: labelDescription
+featureData
+  featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
+  fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+    total)
+  fvarMetadata: labelDescription
+experimentData: use 'experimentData(object)'
+  pubMedIds: 26754106 
+Annotation:  
+- - - Processing information - - -
+Combined [5489,10] and [6268,10] MSnSets Tue Apr  3 12:53:18 2018 
+ MSnbase version: 2.5.9 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<impute, eval=FALSE>>=
-hl <- impute(hl, method = "knn")
-@
+\begin{Schunk}
+\begin{Sinput}
+> hl <- impute(hl, method = "knn")
+\end{Sinput}
+\end{Schunk}
 
@@
   \centering
-<<namap, out.width=".65\\textwidth">>=
-image2(is.na(combined), col = c("black", "white"),
-       main = "Missing values (white cells) after combining replicates")
-@  
+\begin{Schunk}
+\begin{Sinput}
+> image2(is.na(combined), col = c("black", "white"),
++        main = "Missing values (white cells) after combining replicates")
+\end{Sinput}
+\end{Schunk}
 \caption{Heatmap of missing values. Note that the features are
@@
 
-<<filterNA>>=
-combined <- filterNA(combined)
-combined
-@
+\begin{Schunk}
+\begin{Sinput}
+> combined <- filterNA(combined)
+> combined
+\end{Sinput}
+\begin{Soutput}
+MSnSet (storageMode: lockedEnvironment)
+assayData: 5032 features, 20 samples 
+  element names: exprs 
+protocolData: none
+phenoData
+  sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+  varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+  varMetadata: labelDescription
+featureData
+  featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+  fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+    total)
+  fvarMetadata: labelDescription
+experimentData: use 'experimentData(object)'
+  pubMedIds: 26754106 
+Annotation:  
+- - - Processing information - - -
+Combined [5489,10] and [6268,10] MSnSets Tue Apr  3 12:53:18 2018 
+Subset [6725,20][5032,20] Tue Apr  3 12:53:18 2018 
+Removed features with more than 0 NAs: Tue Apr  3 12:53:18 2018 
+Dropped featureData's levels Tue Apr  3 12:53:18 2018 
+ MSnbase version: 2.5.9 
+\end{Soutput}
+\end{Schunk}
 
@@
   \centering
-<<qcplot, fig.show = "hold", message=FALSE, warning=FALSE, out.width=".45\\textwidth">>=
-library("pRoloc")
-plot2D(hl, fcol = NULL, col = "black")
-plot2D(hl, method = "hexbin")
-@  
+\begin{Schunk}
+\begin{Sinput}
+> library("pRoloc")
+> plot2D(hl, fcol = NULL, col = "black")
+> plot2D(hl, method = "hexbin")
+\end{Sinput}
+\end{Schunk}
 \caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
@@
   \centering
-<<qcbx, out.width="\\textwidth", fig.asp=1/2>>=
-par(mfrow = c(1, 2), ## creates a two-panel figure
-    las = 2,         ## axis labels orientation
-    cex.axis = .7)   ## axis label size
-o <- order(hl$Iodixonal.Density)
-plotDist(hl[, o], pcol = "#00000010", xlab = "")
-lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
-boxplot(exprs(hl[, o]))
-@  
+\begin{Schunk}
+\begin{Sinput}
+> par(mfrow = c(1, 2), ## creates a two-panel figure
++     las = 2,         ## axis labels orientation
++     cex.axis = .7)   ## axis label size
+> o <- order(hl$Iodixonal.Density)
+> plotDist(hl[, o], pcol = "#00000010", xlab = "")
+> lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
+> boxplot(exprs(hl[, o]))
+\end{Sinput}
+\end{Schunk}
 \caption{Protein profiles and distribution of channel intensities. The
@@
 
-<<markers>>=
-## List available marker sets
-pRolocmarkers()
-@
+\begin{Schunk}
+\begin{Sinput}
+> ## List available marker sets
+> pRolocmarkers()
+\end{Sinput}
+\begin{Soutput}
+7 marker lists available:
+Arabidopsis thaliana [atha]:
+ Ids: TAIR, 543 markers
+Drosophila melanogaster [dmel]:
+ Ids: Uniprot, 179 markers
+Gallus gallus [ggal]:
+ Ids: IPI, 102 markers
+Homo sapiens [hsap]:
+ Ids: Uniprot, 872 markers
+Mus musculus [mmus]:
+ Ids: Uniprot, 937 markers
+Saccharomyces cerevisiae [scer_sgd]:
+ Ids: SGD, 259 markers
+Saccharomyces cerevisiae [scer_uniprot]:
+ Ids: Uniprot Accession, 259 markers
+\end{Soutput}
+\end{Schunk}
 
@@
 
-<<addmarkers>>=
-## Use mouse markers
-mrk <- pRolocmarkers(species = "mmus")
-head(mrk)
-
-## Add mouse markers
-hl <- addMarkers(hl, mrk)
-@
+\begin{Schunk}
+\begin{Sinput}
+> ## Use mouse markers
+> mrk <- pRolocmarkers(species = "mmus")
+> head(mrk)
+\end{Sinput}
+\begin{Soutput}
+                P26039                 Q6PB66                 P11276 
+  "Actin cytoskeleton"        "Mitochondrion" "Extracellular matrix" 
+                Q6PR54                 Q05793                 P19096 
+ "Nucleus - Chromatin" "Extracellular matrix"              "Cytosol" 
+\end{Soutput}
+\begin{Sinput}
+> ## Add mouse markers
+> hl <- addMarkers(hl, mrk)
+\end{Sinput}
+\begin{Soutput}
+organelleMarkers
+           40S Ribosome            60S Ribosome      Actin cytoskeleton 
+                     27                      43                      13 
+                Cytosol   Endoplasmic reticulum                Endosome 
+                     43                      95                      12 
+   Extracellular matrix         Golgi apparatus                Lysosome 
+                     10                      27                      33 
+          Mitochondrion     Nucleus - Chromatin Nucleus - Non-chromatin 
+                    383                      64                      85 
+             Peroxisome         Plasma membrane              Proteasome 
+                     17                      51                      34 
+                unknown 
+                   4095 
+\end{Soutput}
+\end{Schunk}
 
@@
   \centering
-<<plotmarkers, out.width="\\textwidth", fig.width=12, fig.asp=1/2>>=
-par(mfrow = c(1, 2))
-plot2D(hl, main = "pRolocmarkers for mouse")
-addLegend(hl, cex = .6)
-plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
-@  
+\begin{Schunk}
+\begin{Sinput}
+> par(mfrow = c(1, 2))
+> plot2D(hl, main = "pRolocmarkers for mouse")
+> addLegend(hl, cex = .6)
+> plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
+\end{Sinput}
+\end{Schunk}
 \caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
@@
   \centering
-<<plotDist, out.width=".6\\textwidth", fig.asp=3/5>>=
-hlo <- hl[, order(hl$Iodixonal.Density)]
-plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
-         pcol = "purple", fractions = "Tag")
-title(main = "Marker occupancy profiles along the gradient")
-matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
-         lty = 1, col = "darkblue", type = "l")
-legend("topleft", c("Mitochondrion", "Peroxisome"),
-       lty = 1, col = c("purple", "blue"), bty = "n")
-@  
+\begin{Schunk}
+\begin{Sinput}
+> hlo <- hl[, order(hl$Iodixonal.Density)]
+> plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
++          pcol = "purple", fractions = "Tag")
+> title(main = "Marker occupancy profiles along the gradient")
+> matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
++          lty = 1, col = "darkblue", type = "l")
+> legend("topleft", c("Mitochondrion", "Peroxisome"),
++        lty = 1, col = c("purple", "blue"), bty = "n")
+\end{Sinput}
+\end{Schunk}
   \caption{Mitochondrion and peroxisome protein profiles.}
@@
   \centering
-<<plotmarkers3d, eval=FALSE>>=
-plot3D(hl, dims = c(1, 2, 7))
-@
+\begin{Schunk}
+\begin{Sinput}
+> plot3D(hl, dims = c(1, 2, 7))
+\end{Sinput}
+\end{Schunk}
   \includegraphics[width=.55\textwidth]{./Figures/plot3d.pdf}
@@
 
-<<removeGA>>=
-hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
-getMarkers(hl)
-@
+\begin{Schunk}
+\begin{Sinput}
+> hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
+> getMarkers(hl)
+\end{Sinput}
+\begin{Soutput}
+organelleMarkers
+           40S Ribosome            60S Ribosome      Actin cytoskeleton 
+                     27                      43                      13 
+                Cytosol   Endoplasmic reticulum                Endosome 
+                     43                      95                      12 
+   Extracellular matrix                Lysosome           Mitochondrion 
+                     10                      33                     383 
+    Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+                     64                      85                      17 
+        Plasma membrane              Proteasome                 unknown 
+                     51                      34                    4122 
+\end{Soutput}
+\end{Schunk}
 
@@
 
-\begin{figure}[!ht]
-  \centering
-<<foi, fig.width=10, out.width=".8\\textwidth">>=
-prots <- c("Q8CG48", "Q8CG47", "Q8K2Z4", "Q8C156")
-foi13s <- FeaturesOfInterest(description = "13S consensin proteins",
-                             fnames = prots, 
-                             object = hl)
-foi13s
-
-plot2D(hl)
-addLegend(hl, cex = .6)
-highlightOnPlot(hl, foi13s)
-@
-  \caption{Highlighting protein features of interest.}
-  \label{fig:foi}
-\end{figure}
-
 Users can also create several sets of \texttt{FeaturesOfInterest}
@@
 quantitation.
-
-<<pdtab, echo=FALSE, eval = TRUE, results = 'asis'>>=
-i <- c(3, 13)
-xtable(pData(hl)[i, ], 
-       caption = "Differences in gradient fraction pooling.", 
-       label = "tab:pdtab")
-@
-
-The more relevant comparison unit is not a single channel, but rather
-the complete protein occupancy profiles, which are best visualised 
-experiment-wide on a PCA plot. As such, we prefer to focus on the
-direct, qualitative comparison of individual replicate PCA plots,
-assuring that each displays acceptable sub-cellular resolution. Note
-that in the code chunk below, we mirror the x-axis to represent the
-two figures with the same orientation. The interactive "compare"
-application part of the \Biocpkg{pRolocGUI} package is also useful for
-examining replicate experiments (see the next section interactive
-visualisation for details).
-
-\begin{figure}[!ht]
-  \centering
-<<plot2Drep, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>= 
-par(mfrow = c(1, 2)) 
-plot2D(hl[, hl$Replicate == 1], main = "Replicate 1")
-plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE)
-@  
-  \caption{PCA plots of replicates 1 and 2}
-  \label{fig:plot2Drep}
-\end{figure}
-
-In addition, the reproducibility can be assessed by performing
-independent classification analyses on each replicate (see the section
-on \textit{Supervised machine learning} below) and comparing the the
-results. Even when the gradient conditions different (for unexpected
-technical or voluntary reasons, to maximise resolution when combining
-experiments \cite{Trotter:2010}), one expects agreement in the most
-confident organelle assignments.
-
-\clearpage
-
-\section*{Interactive visualisation}
-
-Visualisation and data exploration is an important aspect of data
-analyses allowing one to shed light on data structure and patterns of
-interest. Using the \Biocpkg{pRolocGUI} package we can interactively
-visualise, explore and interrogate quantitative spatial proteomics
-data. The \Biocpkg{pRolocGUI} package is currently under active
-development and it relies on the \texttt{shiny} framework for
-reactivity and interactivity. The package currently distributes 3
-different GUI's (\textit{main} (default), \textit{compare} or
-\textit{classify}) which are wrapped and launched by the
-\texttt{pRolocVis} function.
-
-\subsubsection*{The main application}
-In the below code chunk we lauch the main
-app (note, we do not need to specify the argument, \texttt{app =
-  "main"} as it is the default).
-
-<<mainapp, eval = FALSE, echo = TRUE>>=
-library("pRolocGUI")
-pRolocVis(hl)
-@
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{./Figures/mainapp.png}
-  \caption{A screen shot of clickable interface and zoomable PCA plot
-    of the main app in the \Biocpkg{pRolocGUI} package. }
-  \label{fig:app}
-\end{figure}
-
-As diplayed in the screenshot in figure \ref{fig:app}, the
-\textit{main} application is designed for exploratory data analysis
-and is divided into 3 tabs: (1) PCA, (2) Profiles and (3) Table
-selection. The default view upon loading is the \textit{PCA} tab,
-which features a clickable interface and zoomable PCA plot with an
-interactive data table for displaying the quantitation information.
-Particular proteins of interest can be highlighted using the text
-search box. There is also a \textit{Profiles} tab for visualisation of
-the protein profiles, which can be used to examine the patterns of
-proteins of interest. The \textit{Table selection} tab provides an
-interface to control data table column selection. A short
-animation\footnote{\url{https://github.com/lmsimp/bioc-pRoloc-hyperLOPIT-workflow/blob/master/Figures/pRolocVis\_pca.gif}}
-illustrating the interface is available in the manuscript
-repository~\cite{ghrepo}.
-
-\subsubsection*{The compare application}
-The \textit{compare} application is useful for examining two replicate
-experiments, or two experiments from different conditions, treatments
-etc. The compare application is called by default if the input object
-to \texttt{pRolocVis} is an \texttt{MSnSetList} of 2 \texttt{MSnSets},
-but it can also be specified by calling the argument \texttt{app =
-  "compare"}. For example, in the code chunk below we first create an
-\texttt{MSnSetList} of replicates 1 and 2 of the hyperLOPIT data, this
-is then passed to \texttt{pRolocVis}.
-
-<<compareapp, eval = FALSE, echo = TRUE>>=
-data(hyperLOPIT2015ms3r1)
-data(hyperLOPIT2015ms3r2)
-hllst <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
-pRolocVis(hllst, app = "compare")
-@
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=\textwidth]{./Figures/SS_Compare1.jpg}
-  \caption{The compare application, main panel}
-  \label{fig:compare}
-\end{figure}
-
-The comparison app loads the two PCA plots side-by-side. Only common
-proteins between the two data sets are displayed. As per the main
-application, proteins can be searched, identified and highlighted on
-both PCA plots and in the dedicated profiles tab. One key feature of
-the compare application is the ability to re-map the second dataset
-onto the PCA data space of the first (reference) data set (see
-\texttt{?pRolocVis} and the argument \texttt{remap = TRUE}). Using the first dataset
-as the reference set, PCA is carried out on the first dataset and the
-standard deviations of the principal components (i.e. the square roots
-of the eigenvalues of the covariance/correlation matrix) and the
-matrix of variable loadings (i.e. a matrix whose columns contain the
-eigenvectors) are stored and then used to calculate the principal
-components of the second dataset. Both datasets are scaled and
-centered in the usual way. The first dataset appears on the left, and
-the second re-mapped data appears on the right. The order of the first
-(the reference data for remapping) and second dataset can be changed
-through regeneration/re-ordering of the \texttt{MSnSetList} object.
-
-\subsubsection*{The classify application}
-The final application \textit{classify}, has been designed to view
-machine learning classification results according to user-specified
-thresholds for the assignment of proteins to its sub-cellular
-location, as discussed later in the subsection \textit{thresholding}
-in the \textit{supervised machine learning section}.
-
-\section*{Novelty Detection}
-
-The extraction of sub-cellular protein clusters can be difficult owing
-to the limited number of marker proteins that exist in databases and
-elsewhere. Furthermore, given the vast complexity of the cell,
-automatic annotation retrieval does not always give a full
-representation of the true sub-cellular diversity in the data. For
-downstream analyses, such as supervised machine learning, it is
-desirable to obtain reliable markers that cover as many sub-cellular
-niches as possible, as these markers are directly used in the training
-phase of the ML classification. We find that a lack of sub-cellular
-diversity in the labelled training data leads to prediction errors, as
-unlabelled instances can only be assigned to a class that exists in
-the training data \cite{Breckels:2013}. In such scenarios, novelty
-detection can be useful to identify data-specific sub-cellular
-groupings such as organelles and protein complexes. The phenotype
-discovery (phenoDisco) algorithm \cite{Breckels:2013} is one such method
-and is available in \Biocpkg{pRoloc}. It is an iterative
-semi-supervised learning method that combines the classification of
-proteins on existing labelled data with the detection of new clusters.
-
-In addition to extracting new phenotypes, novelty detection methods
-are also useful for confirming the presence of known or postulated
-clusters in an unbiased fashion. For example, in \cite{hyper} the
-\texttt{phenoDisco} algorithm was used to confirm the data-specific presence
-of the nucleus and nucleus sub-compartments. In the code chunk below,
-we demonstrate how to do this analysis, highlighting some of the
-optional arguments and parameters available for phenotype extraction
-and give some advice on how to interpret the output.
-
-As the \texttt{phenoDisco} algorithm is semi-supervised it uses both labelled
-(markers) and unlabelled data to explore the data structure and find
-new sub-cellular data clusters. Thus the first step is to define some
-input labelled data i.e. markers, that the algorithm will use as input
-for the supervised learning aspect of the algorithm. As described in
-\cite{hyper} we define a set of markers to use as input for the analyses
-that cover well-known residents from three distinct organelle
-structures; the mitochondria, plasma membrane and ER, and from three
-well-known and abundant protein complexes; the proteasome and two
-ribosomal subunits, 40S and 60S.  These input markers are stored in
-the \texttt{featureData} column of \texttt{hl} where \texttt{fcol = "phenoDisco.Input"}. We
-can use the convenience accessor function \texttt{getMarkers} to print out a
-table of the markers contained in this marker set. These initial
-markers were manually curated using information from the UniProt
-database, the Gene Ontology and the literature.
-
-<<displayPDmarkers>>=
-getMarkers(hl, fcol = "phenoDisco.Input")
-@
-
-In the code chunk below we show how to run the \texttt{phenoDisco}
-function and return a novelty detection result, according to the
-specified parameters. The algorithm parameters \texttt{times} (number
-of iterations) and \texttt{GS} (minimum number of proteins required to
-form a new phenotype) are passed to the function, along with the
-\texttt{fcol} to tell the algorithm where the input training data is
-contained.
-
-<<runPD, eval = FALSE>>=
-## As per Christoforou et al (2016), 
-hl <- phenoDisco(hl, fcol = "phenoDisco.Input", times = 200, GS = 60)
-@
-
-\sout{Note: We do not evaluate this code chunk in this document as the
-algorithm is computationally intensive and best parallelised over
-multiple workers. This phenoDisco analysis took ~24 hours to complete
-when parallelised over 40 workers.}
-
-The argument \texttt{times} indicates the number of times we run
-unsupervied Gaussian Mixture Modelling before defining a new phenotype
-cluster.  The recommended minimum and default value is 100. In the
-above code chunk we increase the value to \texttt{times = 200} as we
-have found for larger datasets (e.g. 5000+ proteins) a higher
-\texttt{times} is requried for convergence. \texttt{GS} defines the
-minimum number of proteins allowed per new data cluster and thus
-heavily influences what type of new clusters are extracted. For
-example, if a user is interested in the detection of small complexes
-they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
-If they wish to detect larger, more abundant sub-cellular niches a
-much higher \texttt{GS} would be preferable. Specifying a small
-\texttt{GS} can be more time consuming than using a larger
-\texttt{GS}, and there is a trade off between finding interesting
-small complexes and those that may not be of interest as we find there
-is a tendancy to find more noise when using a small \texttt{GS}
-compared to using a higher one.
-
-One may also consider increasing the search space for new data
-clusters by increasing the value of the parameter \texttt{G}. This
-defines the number of GMM components to test and fit; the default is
-\texttt{G = 1:9} (the default value in the
-\href{https://cran.r-project.org/web/packages/mclust/index.html}{\textit{mclust}}
-package \cite{mclust}). One should note that the decreasing the
-\texttt{GS}, and increasing the values of the arguments
-\texttt{times}, \texttt{G} (among other function arguments, see
-\texttt{?phenoDisco}) will heavily influence (increase) the total
-time taken to run the algorithm.  \texttt{phenoDisco} supports
-parallelisation and we strongly suggest you make use of a parallel
-processing to run these analyses.
-
-\hl{Please note, in this document we can not evaluate the call to 
-`phenoDisco` in the code chunk above as the algorithm is computationally 
-intensive and best parallelised over multiple workers. This phenoDisco 
-analysis took ~24 hours to complete when parallelised over 40 workers. 
-The ouput of running the \texttt{phenoDisco} algorithm is an
-\texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. We have made the results
-readily available for users to interrogate and get an idea of the type
-of output one may expect from using this function.} The results can be
-displayed by using the \texttt{getMarkers} function. We see that 5 new
-phenotype data clusters were found.
-
-<<loadPDres, echo=FALSE>>=
-f0 <- dir(extdatadir, full.names = TRUE,
-          pattern = "bpw-pdres.rds")
-pdres <- readRDS(f0)
-hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
-setStockcol(NULL)
-setStockcol(paste0(getStockcol(), 70))
-@
-
-<<showPDres>>=
-hl
-getMarkers(hl, fcol = "pd")
-@
-
-\begin{figure}[!ht]
-  \centering
-<<plotPDres, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>=
-## Re-order the colours for the phenoDisco output
-cl <- getMarkerClasses(hl, "pd")
-cols <- getStockcol()[seq(cl)]
-ind <- grep("Pheno", cl, invert = TRUE)
-cols[ind] <- getStockcol()[seq(cl)][1:length(ind)]
-cols[-ind] <- getStockcol()[seq(cl)][(length(ind) + 1):length(cl)]
-
-## Plot the input and output
-par(mfrow = c(1, 2))
-plot2D(hl, fcol = "phenoDisco.Input", 
-       main = "phenoDisco input markers", col = getStockcol()[1:6])
-addLegend(hl, fcol = "phenoDisco.Input", cex = .7)
-plot2D(hl, fcol = "pd", main = "phenoDisco output", col = cols)
-addLegend(hl, fcol = "pd", cex = .7, col = cols)
-@
-  \caption{Results of the novelty detection algorithm.}
-  \label{fig:plotPDres}
-\end{figure}
-
-We can plot the results using the \texttt{plot2D} function (Figure
-\ref{fig:plotPDres}). The five new phenotype data clusters can be
-extracted and examined.  In the code chunk below we write the results
-to a .csv file using the \texttt{write.exprs} function. We use the
-argument \texttt{fDataCols} to specify which columns of the
-\texttt{featureData} to write.
-
-<<writehl, eval=FALSE>>=
-fData(hl)$pd <- as.character(fData(hl)$pd)
-write.exprs(hl, fDataCols = "pd", file = "pd-results.csv", sep = ",")
-@
-
-We can also examine each phenotype interactively and visualise
-their protein profiles by using the \texttt{pRolocVis} function in the
-\Biocpkg{pRolocGUI} package. We found that phenotype 1 was
-enriched in nucleus associated proteins, phenotype 2 in chromatin
-associated proteins, phenotype 3 in cytosolic and phenotypes 4 and 5
-in both lysosomal and endosomal proteins.
-
-<<vishl, eval=FALSE>>=
-pRolocVis(hl, fcol = "pd")
-@
-
-\section*{Supervised machine learning}
-
-Supervised machine learning, also known as classification, is an
-essential tool for the assignment of proteins to distinct sub-cellular
-niches. Using a set of labelled training examples i.e. markers, we can
-train a machine learning classifier to learn a mapping between the
-data i.e. the quantitative protein profiles, and a known localisation.
-The trained classifier can then be used to predict the localisation of
-a protein of unknown localisation, based on its observed protein
-profiles. To date, this method has been extensively used in spatial
-quantitative proteomics to assign thousands of proteins to distinct
-sub-cellular niches
-\cite{hyper,Groen:2014,Trotter:2010,Hall:2009,Dunkley:2006,Tan:2009}.
-
-There are several classification algorithms readily available in
-\Biocpkg{pRoloc}, which are documented in the dedicated
-\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\Biocpkg{pRoloc}
-  machine learning techniques vignette}.  We find the general tendancy
-to be that it is not the choice of classifier, but the improper
-optimisation of the algorithmic parameters, that limits classification
-accuracy. Before employing any classification algorithm and generating
-a model on the training data, one must first find the optimal
-parameters for the algorithm of choice.
-
-\subsection*{Optimisation}
-
-In the code chunk below we use a Support Vector Machine (SVM) to learn
-a classifier on the labelled training data. As previously mentioned,
-one first needs to train the classifier's parameters before an
-algorithm can be used to predict the class labels of the proteins with
-unknown location. One of the most common ways to optimise the
-parameters of a classifier is to partition the labelled data into
-training and testing subsets. In this framework parameters are tested
-via a grid search using cross-validation on the training
-partition. The best parameters chosen from the cross-validation stage
-are then used to build a classifier to predict the class labels of the
-protein profiles on the test partition. Observed and expected
-classication results can be compared, and then used to assess how well
-a given model works by getting an estimate of the classifiers ability
-to achieve a good generalisation i.e. that is given an unknown example
-predict its class label with high accuracy. In \Biocpkg{pRoloc},
-algorithmic performance is estimated using stratified 80/20
-partitioning for the training/testing subsets respectively, in
-conjuction with five-fold cross-validation in order to optimise the
-free parameters via a grid search. This procedure is usually repeated
-100 times and then the best parameter(s) are selected upon
-investigation of classifier accuracy. We recommend a minimum of 13
-markers per sub-cellular class for stratified 80/20 partitioning and
-5-fold cross-validation; this allows a minimum of 10 examples for
-parameter optimisation on the training partition i.e. 2 per fold for
-5-fold cross-validation, and then 3 for testing the best parameters on
-the validation set.
-
-Classifier accuracy is estimated using the macro F1 score, i.e. the
-harmonic mean of precision and recall. In the code chunk below we
-demonstrate how to optimise the free parameters, \texttt{sigma} (the
-inverse kernel width for the radial basis kernel) and \texttt{cost}
-(the cost of constraints violation), of a classical SVM classifier
-with a Gaussian kernel using the function \texttt{svmOptimisation}. As
-the number of labelled instances per class varies from organelle to
-organelle, we can account for class imbalance by setting specific
-class weights when generating the SVM model. Below the weights,
-\texttt{w} are set to be inversely proportional to the class
-frequencies.
-
-<<loadpdRes, echo=FALSE>>=
-svmf <- dir(extdatadir, full.names = TRUE,
-              pattern = "bpw-svmopt.rds") 
-params <- readRDS(svmf)
-@
-
-<<getsvmweight>>=
-w <- table(getMarkers(hl, verbose = TRUE))
-w <- 1/w[names(w) != "unknown"]
-@
-
-<<dosvmopt, eval=FALSE>>=
-## 100 rounds of optimisation with five-fold cross-validation
-params <- svmOptimisation(hl, fcol = "markers",
-                          times = 100, xval = 5,
-                          class.weights = w)
-@
-
-As mentioned previously, we rely on the default feature variable
-\texttt{"markers"} to define the class labels and hence do not need to
-specify it in the above code chunk.  To use another feature variables,
-one needs to explicitly specify its name using the \texttt{fcol}
-argument (for example \texttt{fcol = "markers2"}).
-
-The output \texttt{params} is an object of class \texttt{GenRegRes}; a
-dedicated container for the storage of the design and results from a
-machine learning optimisation. To assess classifier performance we can
-examine the macro F1 scores and the most frequently chosen
-parameters. A high macro F1 score indicates that the marker proteins
-in the test dataset are consistently and correctly assigned by the the
-algorithm. Often more than one parameter or set of parameters gives
-rise to the best generalisation accuracy. As such it is always
-important to investigate the model parameters and critically assess
-the best choice. The \texttt{f1Count} function counts the number of
-parameter occurences above a certain F1 value. The best choice may not
-be as simple as the parameter set that gives rise to the highest macro
-F1 score. One must be careful to avoid overfitting, and choose
-parameters that frequently provide high classification
-accuracy. Below, we see that only a sigma of 0.1 produces macro F1
-scores above 0.6, but that a cost of 16 is much more frequently chosen
-than lower values.
-
-<<f1count>>=
-f1Count(params, 0.6)
-@
-
-The parameter optimistion results can also be visualised as a boxplot
-or heatmap, as shown in figure \ref{fig:visualisOptHide}.  The
-\texttt{plot} method for \texttt{GenRegRes} object shows the
-respective distributions of the 100 macro F1 scores for the best
-cost/sigma parameter pairs, and \texttt{levelPlot} shows the averaged
-macro F1 scores, for the full range of parameter values. These figures
-also indicate that values of 0.1 and 16 for sigma and cost
-consistently deliver better classification scores.
-
-<<visualiseOpt, eval=FALSE>>=
-plot(params)
-levelPlot(params)
-@
-
-\begin{figure}[!ht]
-  \centering
-<<visualiseOptHide, echo=FALSE, out.width=".65\\textwidth">>=
-library('gridExtra')
-grid.arrange(plot(params), levelPlot(params), ncol = 2)
-@  
-  \caption{Assessment of the classification model parameter optimisation.}
-  \label{fig:visualisOptHide}
-\end{figure}
-
-By using the function \texttt{getParams} we can extract the best set
-of parameters. Currently, \texttt{getParams} retrieves the first best
-set automatically, but users are encouraged to critically assess
-whether this is the most wise choice (which it is, as demonstrated
-above).
-
-<<getbestpars>>=
-(best <- getParams(params))
-@
-
-Once we have selected the best parameters we can then use them to
-build a classifier from the labelled marker proteins.
-
-\subsection*{Classification}
-
-We can use the function \texttt{svmClassification} to return a
-classification result for all unlabelled instances in the dataset
-corresponding to their most likely sub-cellular location. The
-algorithm parameters are passed to the function, along with the class
-weights. As above, the \texttt{fcol} argument does not need to be
-specified as we use the labels defined in the default
-\texttt{"markers"} feature variable.
-
-<<classify>>=
-hl <- svmClassification(hl, params, class.weights = w, fcol = "markers")
-@
-
-In the code chunk above, we pass the whole \texttt{params} parameter
-results and, internally, the first pair that return the highest F1
-score are returned (using the \texttt{getParams} function above). It
-is advised to always check that these are actually good parameters
-and, if necessary, set them explicitly, as shown below.
-
-<<classify2, eval = FALSE>>=
-hl <- svmClassification(hl, cost = 16, sigma = 0.1, class.weights = w, fcol = "markers")
-@
-
-Automatically, the output of the above classification, the organelle
-predictions and assignment scores, are stored in the
-\texttt{featureData} slot of the \texttt{MSnSet}. In this case, they
-are given the labels \texttt{svm} and \texttt{svm.scores} for the
-predictions and scores respectively. The resultant predictions can be
-visualised using \texttt{plot2D}. In the code chunk below
-\texttt{plot2D} is called to generate a PCA plot of the data and
-\texttt{fcol} is used to specify where the new assignments are located
-e.g. \texttt{fcol = "svm"}.
-
-Additionally, when calling \texttt{plot2D} we can use the \texttt{cex}
-argument to change the size of each point on the plot to be inversely
-proportional to the SVM score. This gives an initial overview of the
-high scoring localisations from the SVM predictions.
-
-<<setColors, include=FALSE>>=
-setStockcol(NULL)
-setStockcol(paste0(getStockcol(), 70))
-par(mfrow = c(1, 1))
-@
-
-\begin{figure}[!ht]
-  \centering
-<<plotSVM, out.width=".8\\textwidth">>=
-## set point size of each protein to be inversely proportional to the
-ptsze <- exp(fData(hl)$svm.scores) - 1
-## plot new predictions
-plot2D(hl, fcol = "svm", cex = ptsze)
-addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
-@  
-\caption{Classification results. Colours indicate class membership and
-  point size are representative of the classification confidence.}
-  \label{fig:plotSVM}
-\end{figure}
-
-The adjustment of the point size intuitively confers important
-information that is more difficult to define formally (we
-will address in the next section). The classifier (SVM in our case, but
-this is also valid of other classifiers) defines boundaries based on
-the labelled marker proteins. These class/organelle boundaries define
-how non-assigned proteins are classified and with what confidence.
-
-\subsection*{Thresholding}\label{sec:thresholding}
-
-It is common when applying a supervised classification algorithm to
-set a specific score cutoff on which to define new assignments, below
-which classifications are kept unknown/unassigned. This is important
-as in a supervised learning setup, proteins can only be predicted to
-be localised to one of the sub-cellular niches that appear in the
-labelled training data. We can not guarantee (and do not expect) that
-the whole sub-cellular diversity to be represented in the labelled
-training data as (1) finding markers that represent the whole
-diversity of the cell is challenging (especially obtaining dual- and
-multiply-localised protein markers) and (2) many sub-cellular niches
-contain too few proteins to train on (see above for a motivation of a
-minimum of 13 markers).
-
-Deciding on a threshold is not trivial as classifier scores are
-heavily dependent upon the classifier used and different sub-cellular
-niches can exhibit different score distributions, as highlighted in
-the boxplot below. We recommend users to set class-specific
-thresholds.  In the code chunk below we display a boxplot of the score
-distributions per organelle (Figure \ref{fig:threshold}).
-
-\begin{figure}[!ht]
-  \centering
-<<thresholds, out.width=".8\\textwidth">>=
-## First remove the markers
-preds <- unknownMSnSet(hl)
-## Plot a boxplot of the scores of each organelle
-par(oma = c(10.5, 0, 0, 0)) ## sets outer margins
-boxplot(svm.scores ~ svm, data = fData(preds), 
-        ylab = "SVM scores", las = 2)
-@
-  \caption{Visualistion of class-specific classification score distribution.}
-  \label{fig:threshold}
-\end{figure}
-
-There are many ways to set thresholds and the choice of method will
-depend on the biological question and experimental design at hand. One
-viable approach in the frame of the above experimetal design would be
-to manually set a FDR, say 5\%, per organelle. To do this the user
-would examine the top scoring predictions for each organelle, and then
-set a threshold at the score at which they achieve 5\% of false
-assignments per organelle.  The definintion of a false assignment
-would depend on the information available, for example, validity or
-lack of validity for the localisation from another experiment as
-reported in the literature or a reliable database.  If such
-information is not available, one crude method is to set a threshold
-per organelle by extracting the median or 3rd quantile score per
-organelle.  For example, in the code chunk below, we use the
-\texttt{orgQuants} function to extract the median organelle scores and
-then pass these scores to the \texttt{getPredictions} function to
-extract the new localisations that meet this scoring criteria. Any
-sub-cellular predictions that fall below the specified thresholds are
-labelled as unknown.
-
-<<orgquants>>=
-(ts <- orgQuants(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = .5))
-hl <- getPredictions(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = ts)
-@
-
-The organelle threshold (\texttt{ts} above) can also be set manually
-using an interactive app (see below) or by using a named vector of
-thresholds, as shown in the putative example below for 4 organelles.
-
-<<orgts>>=
-(ts <- setNames(c(0.612, 0.701, 0.81, 0.92), c("PM", "Mito", "Golgi", "ER")))
-@
-
-The output of \texttt{getPredictons} is the original \texttt{MSnSet}
-dataset with a new feature variable appended to the feature data
-called \texttt{fcol.pred} (i.e. in our case \texttt{svm.pred})
-containing the prediction results. The results can also be visualied
-using \texttt{plot2D} function (Figure \ref{fig:plotres}) and
-extracted by retrieving that specific column from the feature metadata
-using, for example, \texttt{fData(hl)\$svm.pred}.
-
-<<resetpar, echo=FALSE>>=
-invisible(dev.off())
-@
-
-\begin{figure}[!ht]
-    \centering
-<<plotres, out.width=".7\\textwidth">>=
-plot2D(hl, fcol = "svm.pred")
-@  
-  \caption{Results of the localisation preductions after thresholding.}
-  \label{fig:plotres}
-\end{figure}
-
-There is also a dedicated interactive application to help users
-examine these distributions in the \Biocpkg{pRolocGUI} package (Figure
-\ref{fig:classifyapp}).  This app can be launched via the
-\texttt{pRolocVis} function and specifying the argument \texttt{app =
-  "classify"} along with the relevent \texttt{fcol}, \texttt{scol} and
-\texttt{mcol} which refer to the columns in the feature data that
-contain the new assignments, assignment scores and markers
-respectively (see also \texttt{fvarLabels(svmres)}).
-
-<<classifyApp, eval=FALSE>>=
-library("pRolocGUI")
-pRolocVis(hl, app = "classify", fcol = "svm", scol = "svm.scores", mcol = "markers")
-@
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=0.7\textwidth]{./Figures/classify.png}
-  \caption{The classify application enable the interactive exploration
-    of classification score thresholding.}
-  \label{fig:classifyapp}
-\end{figure}
-
-The data is loaded and displayed on a PCA plot and a boxplot is used
-to display the classifier scores by data class. On the left, there is
-a sidebar panel with sliders to control the thresholds upon which
-classifications are made. There are two types of cut-off that the user
-can choose from: (1) \textit{Quantile} and (2)
-\textit{User-defined}. By default, when the application is launched
-quantile scoring is selected and set to 0.5, the median. The
-class-specific score thresholds that correspond to selecting the
-desired quantile are shown as red dots on the boxplot. The assignments
-on the PCA plot are also updated according to the selected
-threshold. The quantile threshold can be set by moving the
-corresponding quantile slider. If the users wishes to set their own
-cut-offs, the \textit{User-defined} radio button must be selected and
-then the sliders for defining user-specified scores become active and
-the scores and highlighted on the boxplot by blue dots. For more
-information we refer users to the \Biocpkg{pRolocGUI} tutorial
-\href{http://bioconductor.org/packages/release/bioc/vignettes/pRolocGUI/inst/doc/pRolocGUI.html}{vignette}.
-
-\clearpage
-
-\section*{Transfer learning}
-
-In addition to high quality MS-based quantitative proteomics data,
-there exist a number of other sources of information that are freely
-available in the public domain that may be useful to assign a protein
-to its sub-cellular niche. For example, imaging from
-immunofluorescence microscopy, protein annotations and sequences, and
-protein-protein interactions among others, represent a rich and vast
-source of complementary information. We can integrate this auxiliary
-information with our primary MS-based quantitative data using a
-paradigm known as transfer learning (TL). The integration of data
-between different technologies is one of the biggest challenges in
-computational biology to date and the \Biocpkg{pRoloc} package
-provides functionality to do such analyses. We recently developed two
-transfer learning algorithms using a $k$-NN and SVM framework and
-applied them to the task of protein localisation prediction
-\cite{Breckels:2016}. In this section we will begin with explaining
-the concept of transfer learning and then show how to apply this in
-the frame of spatial proteomics and protein localisation prediction.
-
-In TL one typically has a primary task that one wishes to solve, and
-some complementary (often heterogeneous) auxiliary information that is
-related to the primary learning objective, that can be used to help
-solve the primary goal. For example, here our primary task is to
-assign proteins to their sub-cellular niche with high generalisation
-accuracy from data collected from quantitative MS-based experiments.
-We have already seen that straightforward supervised ML works well for
-these types of experiments, however, Transfer learning is particularly
-useful for classes that are not as well separated.
-
-In the example below we extract Gene Ontology (GO) information to use
-as an auxiliary data source to help solve our task of protein
-localisation prediction.
-
-Using the functions \texttt{setAnnotationParams} and
-\texttt{makeGoSet} we can contruct an auxiliary \texttt{MSnSet} of GO
-terms, from the primary data's features i.e. the protein accession
-numbers. All the GO terms associated to each accession number are
-retrieved and used to create a binary matrix where a 1 (0) at position
-$(i, j)$ indicates that term $j$ has (not) been used to annotate
-protein $i$. The GO terms are retrieved from an appropriate repository
-using the \Biocpkg{biomaRt} package. The specific Biomart repository
-and query will depend on the species under study and the type of
-identifiers. The first step is to construct the annotation parameters
-that will enable to perform the query, which is done using
-\texttt{setAnnotationParams}. Typing into the R console \texttt{par <-
-  setAnnotationParams()} will present two menus, firstly asking you to
-identify the species of study, and then what type of identifier you
-have used to annotate the proteins in your \texttt{MSnSet}. It is also
-possible to pass arbitrary text to match the species e.g. in the code
-chunk below we pass \texttt{"Mus musculus"}, and the identifier type
-for our data (see \texttt{featureNames(hl)}) which is
-\texttt{"Uniprot/Swissprot"}, for the Biomart query.
-
-<<setap, eval=FALSE>>=
-par <- setAnnotationParams(inputs = c("Mus musculus", "UniProt/Swissprot"))
-@
-
-Now we have contructed the query parameters we can use the
-\texttt{makeGoSet} function to retrieve and build an auxiliary GO
-\texttt{MSnSet} as described above. By default, the cellular component
-terms are downloaded, without any filtering on evidence codes. It is
-also possible to download terms from the molecular function and
-biological process GO namespaces, and also apply filtering based on
-evidence codes as desired, see \texttt{?makeGoSet} for more details.
-
-<<makego, eval=FALSE>>=
-gocc <- makeGoSet(hl, params = par, namespace = "cellular_component")
-@
-
-The function \texttt{makeGoSet} uses the \Biocpkg{biomaRt} package to
-query the relevent database (e.g. Ensembl, Uniprot) for GO terms. All
-GO terms that have been observed for the 5032 proteins in the
-hyperLOPIT dataset are retrieved. Users should note that the number of
-GO terms used is also dependent on the database version queried and
-thus is always subject to change. We find it is common to see GO terms
-with only one protein assigned to that term. Such terms do not bring
-any information for building the classifier and are thus removed using
-the \texttt{filterBinMSnSet} function.
-
-<<filtergo, eval=FALSE>>=
-gocc <- filterBinMSnSet(hl)
-@
-
-Now that we have generated our auxiliary data, we can use the
-\textit{k}-NN implementation of transfer learning available in
-\Biocpkg{pRoloc} to integrate this with our primary MS-based
-quantitative proteomics data using the functions
-\texttt{knntlOptimisation} to estimate the free-parameters for the
-integration, and \texttt{knntlClassification} to do the
-predictions. We have shown that using transfer learning results in the
-assignment of proteins to sub-cellular niches with a higher
-generalisation accuracy than using standard supervised machine
-learning with a single source of information \cite{Breckels:2016}.
-
-\subsection*{TL optimisation}
-
-The first step, as with any machine learning algorithm, is to optimise
-any free paramaters of the classifier. For the $k$-NN TL classifier
-there are two sets of parameters that need optimising: the first set
-are the $k$'s for the primary and auxiliary data sources required for
-the nearest neighbour calculations for each data source. The second
-set of parameters (noted by a vector of $\theta$ weights) that require
-optimising are the class weights, one per subcellular niche, that
-control the proportion of primary and auxiliary data to use for
-learning. A weight can take any real value number between 0 and 1. A
-weight of $\theta = 1$ indicates that all weight is given to the
-primary data (and this implicitly implies that a weight of
-$1 - \theta = 0$ is given to the auxiliary data), and similarly a
-weight of $\theta = 0$ implies that all weight is given to the
-auxiliary data (so 0 is given to the primary source). If we conduct a
-parameter seach and test weights $\theta = {0, 1/3, 2/3, 1}$ for each
-class, and if we have, for example 10 subcellular niches, this will
-result in \texttt{4$^{10}$} different combinations of parameters to
-test. The parameter optimisation is therefore time consuming and as
-such we recommend making use of a computing cluster (code and
-submissing scripts are also available in the supporting
-information). The markers in the \texttt{hl} dataset contain 14
-subcellular classes. If we examine these markers and classes on the
-PCA plot above we can see that in particular the two ribosomes and two
-nuclear compartments are highly separated along the first two
-components, this is also evident from the profiles plot which gives us
-a good indication that these subcellular niches are well-resolved in
-the hyperLOPIT dataset. Transfer learning is particularly useful for
-classes that are not as well separated. We find that subcellular
-niches that are well-separated under hyperLOPIT and LOPIT obtain a
-class score of 1 (i.e. use only primary data from transfer learning
-\cite{Breckels:2016}). Therefore, for the optimisation stage of the
-analyses we can already infer a subcellular class weight of 1 for
-these niches and only optimise over the remaining organelles.  This
-can significantly cut down optimisation time as by removing these 4
-classes from the optimisation (and not the classification) we only
-have \texttt{4$^{10}$} class weight combinations to consider instead
-of \texttt{4$^{14}$} combinations.
-
-In the example below we remove these 4 classes from the marker set,
-re-run the \texttt{knnOptimisation} for each data source and then run
-the \texttt{knntlOptimisation} with the 10 remaining classes. (Note:
-this is not run live as the \texttt{hl} dataset with 10 classes,
-707 markers and 4$^{10}$ combinations of parameters takes around 76
-hours to run on the University of Cambridge HPC using 256 workers).
-
-To remove the 4 classes and create a new column of markers in the
-feature data called \texttt{tlmarkers} to use for the analysis:
-
-<<loadgocc, include=FALSE>>=
-gofile <- dir(extdatadir, full.names = TRUE,
-              pattern = "bpw-gocc.rds")
-gocc <- readRDS(gofile)
-@
-
-<<setuptlopt>>=
-## create new markers column for tl markers
-fData(hl)$tlmarkers <- fData(hl)$markers
-fData(gocc)$tlmarkers <- fData(gocc)$markers
-
-## Remove 4 classes
-torm  <- c("40S Ribosome", "60S Ribosome",
-           "Nucleus - Chromatin", 
-           "Nucleus - Non-chromatin")
-for (i in seq(torm)) {
-  hl <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
-  gocc <- fDataToUnknown(gocc, from = torm[i], fcol = "tlmarkers")
-}
-getMarkerClasses(hl, fcol = "tlmarkers")
-getMarkerClasses(gocc, fcol = "tlmarkers")
-@
-
-\paragraph{TL optimisation stage 1} Run \texttt{knnOptimisation} to get
-the best $k$'s for each data source.
-
-<<knnoptfortl, eval=FALSE>>=
-## get best k's
-kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
-kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
-@
-
-From examining the parameter seach plots as described in section
-\textit{Optimisation}, we find the best $k$'s for both the primary and
-auxiliary are 3.
-
-\paragraph{TL optimisation stage 2} Run \texttt{knntlOptimisation} to
-get the best transfer learning weights for each sub-cellular class.
-
-<<runtlopts, eval=FALSE>>=
-## Set appropriate parallelisation backend and  
-## number of workers for the tl
-par <- SnowParam(255L, type = "MPI")
-
-## Now peform tl optimisation 
-tlopt <- knntlOptimisation(hl, gocc, fcol = "tlmarkers",
-                           length.out = 4, times = 50,
-                           xval = 5, k = c(3, 3),
-                           BPPARAM = par)
-@
-
-The results of the optimisation can be visalised using the
-\texttt{plot} method for \texttt{tlopt} optimisation result:
-
-<<loadtlres, echo=FALSE>>=
-tlfile <- dir(extdatadir, full.names = TRUE,
-              pattern = "bpw-tlopt.rds")
-tlopt <- readRDS(tlfile)
-colnames(tlopt@results)[4] <- "Endoplasmic reticulum"
-setStockcol(NULL)
-@
-
-\begin{figure}[!ht]
-  \centering
-<<plottl, out.width=".6\\textwidth">>=
-plot(tlopt)
-@  
-\caption{Visualisation of the transfer learning parameter optimisation
-  procedure. Each row displays the frequency of observed weights
-  (along the columns) for a specific sub-cellular class, with large
-  dots representing higher observation frequencies.}
-  \label{fig:plottl}
-\end{figure}
-
-\subsection*{TL classification}
-
-Looking at the bubble plot displaying the distribution of best weights
-over the 50 runs we find that for many of the subcellular niches a
-weight of 1 is most popular (i.e. use only primary hyperLOPIT data in
-classification), this is unsuprising as we already know the dataset is
-well resolved for these classes. We see that the most popular weights
-for the proteasome and lysosome tend to be towards 0, indicating that
-these niches are well-resolved in the Gene Ontology. This tells us
-that we would benefit from including auxiliary GO information in our
-classifier for these subcellular compartments. The plasma membrane
-weights are relatively equally spread between using hyperLOPIT and GO
-data. Using the \texttt{getParams} function we can return the best
-weights and then use this as input for the classification.
-
-One of the benefits of the algorithm is the ability to manually select
-weights for each class. In the optimisation above, for time
-constraints, we removed the two ribosomal subunits and the two nuclear
-compartments, and therefore in the code chunk below when we extract
-the best parameters, these subcellular niches are not included. To
-include these 4 subcellular niches in the next classification step we
-must include them in the parameters. We define a weight of 1 for each
-of these niches, as we know they are well resolved in hyperLOPIT. We
-then re-order the weights according to \texttt{getMarkerClasses} and
-perform the classification using the function
-\texttt{knntlClassification}.
-
-<<tlclassify, cache = cache>>=
-## best parameters for the 10 classes
-(bestpar <- getParams(tlopt))
-
-## add weights for classes not included in the optimisation
-otherweights <- rep(1, 4)
-names(otherweights) <- c("40S Ribosome", "60S Ribosome", 
-                         "Nucleus - Chromatin", 
-                         "Nucleus - Non-chromatin")
-(bestpar <- c(bestpar, otherweights))
-
-## re-order classes 
-bestpar <- bestpar[getMarkerClasses(hl)]
-
-## Do the classification
-hl <- knntlClassification(hl, gocc, bestTheta = bestpar, k = c(3, 3))
-@
-
-The results from the classification results and associated scores are
-appended to the \texttt{fData} slot and named \texttt{knntl} and
-\texttt{knntl.scores} respectively. Results can be visualised using
-\texttt{plot2D}, scores assessed and cutoffs calculated using the
-\texttt{classify} app in \texttt{pRolocVis}, predictions obtained
-using \texttt{getPredictions} in the same way as demonstrated above
-for the SVM classifier.
-
-In \Biocpkg{pRoloc}'s
-\href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-transfer-learning.pdf}{transfer
-  learning vignette}, we demonstrate how to use imaging data from the
-Human Protein Atlas \cite{Uhlen:2010} via the \Biocpkg{hpar} package
-\cite{hpar} as an auxiliary data source.
-
-\section*{Unsupervised machine learning}
-
-In \Biocpkg{pRoloc} there is functionality for unsupervsied machine
-learning methods. In unsupervised learning, the training data consists
-of a set of input vectors e.g. protein profiles, ignoring the
-information about the class label e.g. localisation, other than for
-annotation purposes. The main goal in unsupervised learning is to
-uncover groups of similar features within the data, termed
-clustering. Ordination methods such as principal components analysis
-(PCA) also fall into the category of unsupervised learning methods,
-where the data can be projected from a high-dimensional space down to
-two or three dimensions for the purpose of visualisation.
-
-As described and demonstrated already above, PCA is a valuable and
-powerful method for data visualisation and quality control. Another
-application uses hierarchical clustering to summarise the relation
-between marker proteins using he \texttt{mrkHClust} function, where
-the euclidean distance between average class-specific profiles is used
-to produce a dendrogramme describing a simple relationship between the
-sub-cellular classes (Figure~\ref{fig:mrkHclust}). The
-\texttt{mrkHClust} uses the same defaults as all other function, using
-the \texttt{markers} feature variable to define marker proteins. In
-the code chunk, we adapt the figure margins to fully display the class
-names.
-
-\begin{figure}[!ht]
-  \centering
-<<hclust, out.width=".5\\textwidth", fig.asp = 1.5, out.extra='angle=-90'>>=
-par(mar = c(15, 4, 1, 0)) ## set figure margin
-mrkHClust(hl)
-@  
-  \caption{Hierarchical clustering of the average marker profiles summarising the relation between organelles profiles.}
-  \label{fig:mrkHclust}
-\end{figure}
-
-We generally find supervised learning more suited to the task of
-protein localisation prediction in which we use high-quality curated
-marker proteins to build a classifier, instead of using an entirely
-unsupervised approach to look for clusters and then look for
-enrichment of organelles and complexes. In the latter we do not make
-good use of valuable prior knowledge, and in our experience
-unsupervised clustering can be extremely difficult due to poor
-estimates of the number of clusters that may appear in the data.
-
-\section*{Writing and exporting data}
-
-An \texttt{MSnSet} can be exported from R using the
-\texttt{write.exprs} function.  This function writes the expression
-values to a text-based spreadsheet. The \texttt{fcol} argument can be
-used to specify which \texttt{featureData} columns (as column names,
-column number or \texttt{logical}) to append to the right of the
-expression matrix.
-
-In the below code chunk we write the \texttt{hl} object to a csv
-file. The \texttt{file} argument is used to specify the file path, the
-\texttt{sep} argument specifies the field separator string, here we
-use a comma, finally as we want to write all the information in the
-\texttt{featureData} to the file, as well as the expression data, we
-specify \texttt{fvarLabels(hl)}, that returns all feature variable
-names, and write the resulting data to the file \texttt{"hl.csv"}.
-
-<<writeData, eval = FALSE>>=
-write.exprs(hl, file = "hl.csv", sep = ",", fcol = fvarLabels(hl))
-@
-
-Exporting to a spreadsheet however loses a lot of important
-information, such as the processing data, and the sample metadata in
-the \textit{phenoData} slot. Other objects, such as parameters from the
-machine learning optimisation, cannot be represented as tabular
-data. To directly serialise R objects to disk, on can use the standard
-\texttt{save} function, and later reload the object using
-\texttt{save}. For example, to save and then re-load the parameters
-from the SVM optimisation,
-
-<<saveData, eval = FALSE>>=
-## To save the parameters as an R object
-save(params, file = "svmparams.rda")
-
-## To re-load after saving
-load(file = "svmparams.rda")
-@
-
-\section*{Session information}
-
-The function \texttt{sessionInfo} provides a summary of all packages
-and versions used to generate this document. This enables us to record
-the exact state of our session that lead to these results. Conversely,
-if the script stops working or if it returns different results, we are
-in a position to re-generate the original results using the adequate
-software versions and retrace changes in the software that lead to
-failure and/or different results.
-
-<<si>>=
-sessionInfo()
-@
-
-We also recommend that users regularly update the packages as well as
-the R itself. This can be done with the \texttt{biocLite} function.
-
-<<update, eval=FALSE>>=
-library("BiocInstaller")
-biocLite()
-@
-
-It is always important to include session information details along
-with a \href{http://adv-r.had.co.nz/Reproducibility.html}{short
-  reproducible example} highlighting the problem or
-\href{https://support.bioconductor.org/}{question} at hand.
-
-The source of this document, including the code necessary to reproduce
-the analyses and figures is available in a public manuscript
-repository on GitHub~\cite{ghrepo}.
-
-\bigskip
-
-\subsection*{Author contributions}
-
-LMB and LG developed the software presented in this workflow. All
-authors wrote and approved the manuscript.
-
-\subsection*{Competing interests}
-
-The authors declare that they have no competing interest.
-
-\subsection*{Grant information}
-
-LMB and CMM are supported by a Wellcome Trust Technology Development
-Grant (grant number 108441/Z/15/Z). KSL is a Wellcome Trust Joint
-Investigator (110170/Z/15/Z). LG is supported by the BBSRC Strategic
-Longer and Larger grant (Award BB/L002817/1).
-
-\subsection*{Acknowledgements}
-
-The authors would like to thank Dr. Stuart Rankin from the High
-Performance Computing Service for his support. Part of this work was
-performed using the Darwin Supercomputer of the University of
-Cambridge High Performance Computing Service, provided by Dell
-Inc. using Strategic Research Infrastructure Funding from the Higher
-Education Funding Council for England and funding from the Science and
-Technology Facilities Council.

### 2018-04-04 bioc-workflow_march2017.tex
addressing reviewers qs

--- 2018-04-03 bioc-workflow_march2017.Rnw
@@
 
-<<install, eval=FALSE>>=
-source("https://bioconductor.org/biocLite.R")
-biocLite(c("MSnase", "pRoloc", "pRolocdata", "pRolocGUI"))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{source}\hlstd{(}\hlstr{"https://bioconductor.org/biocLite.R"}\hlstd{)}
+\hlkwd{biocLite}\hlstd{(}\hlkwd{c}\hlstd{(}\hlstr{"MSnase"}\hlstd{,} \hlstr{"pRoloc"}\hlstd{,} \hlstr{"pRolocdata"}\hlstd{,} \hlstr{"pRolocGUI"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<getFilename, cache = cache>>=
-library("MSnbase")
-extdatadir <- system.file("extdata", package = "pRolocdata")
-csvfile <- dir(extdatadir, full.names = TRUE,
-          pattern = "hyperLOPIT-SIData-ms3-rep12-intersect.csv")
-basename(csvfile)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"MSnbase"}\hlstd{)}
+\hlstd{extdatadir} \hlkwb{<-} \hlkwd{system.file}\hlstd{(}\hlstr{"extdata"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"pRolocdata"}\hlstd{)}
+\hlstd{csvfile} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+          \hlkwc{pattern} \hlstd{=} \hlstr{"hyperLOPIT-SIData-ms3-rep12-intersect.csv"}\hlstd{)}
+\hlkwd{basename}\hlstd{(csvfile)}
+\end{alltt}
+\begin{verbatim}
+## [1] "hyperLOPIT-SIData-ms3-rep12-intersect.csv.gz"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<getEcols>>=
-getEcols(csvfile, split = ",", n = 2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{getEcols}\hlstd{(csvfile,} \hlkwc{split} \hlstd{=} \hlstr{","}\hlstd{,} \hlkwc{n} \hlstd{=} \hlnum{2}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+##  [1] ""                                 
+##  [2] ""                                 
+##  [3] ""                                 
+##  [4] "Experiment 1"                     
+##  [5] "Experiment 2"                     
+##  [6] "Experiment 1"                     
+##  [7] "Experiment 2"                     
+##  [8] "126"                              
+##  [9] "127N"                             
+## [10] "127C"                             
+## [11] "128N"                             
+## [12] "128C"                             
+## [13] "129N"                             
+## [14] "129C"                             
+## [15] "130N"                             
+## [16] "130C"                             
+## [17] "131"                              
+## [18] "126"                              
+## [19] "127N"                             
+## [20] "127C"                             
+## [21] "128N"                             
+## [22] "128C"                             
+## [23] "129N"                             
+## [24] "129C"                             
+## [25] "130N"                             
+## [26] "130C"                             
+## [27] "131"                              
+## [28] "phenoDisco Input"                 
+## [29] "phenoDisco Output"                
+## [30] "Curated phenoDisco Output"        
+## [31] "SVM marker set"                   
+## [32] "SVM classification"               
+## [33] "SVM score"                        
+## [34] "SVM classification (top quartile)"
+## [35] "Final Localization Assignment"    
+## [36] "First localization evidence?"     
+## [37] "Curated Organelles"               
+## [38] "Cytoskeletal Components"          
+## [39] "Trafficking Proteins"             
+## [40] "Protein Complexes"                
+## [41] "Signaling Cascades"               
+## [42] "Oct4 Interactome"                 
+## [43] "Nanog Interactome"                
+## [44] "Sox2 Interactome"                 
+## [45] "Cell Surface Proteins"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<readMSnSet2>>=
-hl <- readMSnSet2(csvfile, ecol = 8:27, fnames = 1, skip = 1)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{readMSnSet2}\hlstd{(csvfile,} \hlkwc{ecol} \hlstd{=} \hlnum{8}\hlopt{:}\hlnum{27}\hlstd{,} \hlkwc{fnames} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{skip} \hlstd{=} \hlnum{1}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
 Below, we display a short summary of the data. The data contains 
-\Sexpr{nrow(hl)} proteins/features common across the 2 biological replicates
-for the respective 2 x 10-plex reporter tags (\Sexpr{ncol(hl)}
+5032 proteins/features common across the 2 biological replicates
+for the respective 2 x 10-plex reporter tags (20
 columns or samples), along with associated feature meta-data such as
@@
 
-<<showhl>>=
-hl
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 5032 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData: none
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+##   fvarLabels: X X.1 ... Cell.Surface.Proteins (25 total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+## Annotation:  
+## - - - Processing information - - -
+##  MSnbase version: 2.5.10
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<assesshl>>=
-exprs(hl)[1:5, ]
-exprs(hl)[c("Q9ERU9", "Q9Z2R6"), c("X126", "X131.1")]
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{exprs}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{5}\hlstd{, ]}
+\end{alltt}
+\begin{verbatim}
+##           X126 X127N X127C X128N X128C X129N X129C X130N X130C  X131
+## Q9JHU4   0.028 0.034 0.024 0.014 0.026 0.045 0.107 0.341 0.059 0.321
+## Q9QXS1-3 0.039 0.134 0.095 0.053 0.084 0.121 0.107 0.128 0.122 0.117
+## Q9ERU9   0.021 0.013 0.014 0.009 0.024 0.054 0.116 0.257 0.209 0.284
+## P26039   0.120 0.255 0.148 0.091 0.135 0.095 0.041 0.057 0.014 0.043
+## Q8BTM8   0.055 0.139 0.078 0.050 0.077 0.098 0.093 0.171 0.079 0.160
+##          X126.1 X127N.1 X127C.1 X128N.1 X128C.1 X129N.1 X129C.1 X130N.1
+## Q9JHU4    0.037   0.064   0.058   0.059   0.067   0.078   0.140   0.208
+## Q9QXS1-3  0.033   0.073   0.074   0.062   0.081   0.142   0.190   0.069
+## Q9ERU9    0.026   0.017   0.023   0.029   0.039   0.071   0.105   0.171
+## P26039    0.111   0.181   0.141   0.144   0.152   0.119   0.075   0.028
+## Q8BTM8    0.062   0.108   0.091   0.086   0.099   0.111   0.117   0.095
+##          X130C.1 X131.1
+## Q9JHU4     0.141  0.147
+## Q9QXS1-3   0.151  0.125
+## Q9ERU9     0.304  0.215
+## P26039     0.017  0.033
+## Q8BTM8     0.144  0.087
+\end{verbatim}
+\begin{alltt}
+\hlkwd{exprs}\hlstd{(hl)[}\hlkwd{c}\hlstd{(}\hlstr{"Q9ERU9"}\hlstd{,} \hlstr{"Q9Z2R6"}\hlstd{),} \hlkwd{c}\hlstd{(}\hlstr{"X126"}\hlstd{,} \hlstr{"X131.1"}\hlstd{)]}
+\end{alltt}
+\begin{verbatim}
+##         X126 X131.1
+## Q9ERU9 0.021  0.215
+## Q9Z2R6 0.563  0.000
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<removefData>>=
-fvarLabels(hl)
-fvarLabels(hl)[1:3] <- c("uniprot.accession", "uniprot.id", "description")
-fvarLabels(hl)[4:6] <- paste0("peptides.expt", 1:3)
-## feature vars 1, 2, and 4 to 6
-fData(hl)[1:4, c(1:2, 4:6)]
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hl)}
+\end{alltt}
+\begin{verbatim}
+##  [1] "X"                                
+##  [2] "X.1"                              
+##  [3] "X.2"                              
+##  [4] "Experiment.1"                     
+##  [5] "Experiment.2"                     
+##  [6] "Experiment.1.1"                   
+##  [7] "Experiment.2.1"                   
+##  [8] "phenoDisco.Input"                 
+##  [9] "phenoDisco.Output"                
+## [10] "Curated.phenoDisco.Output"        
+## [11] "SVM.marker.set"                   
+## [12] "SVM.classification"               
+## [13] "SVM.score"                        
+## [14] "SVM.classification..top.quartile."
+## [15] "Final.Localization.Assignment"    
+## [16] "First.localization.evidence."     
+## [17] "Curated.Organelles"               
+## [18] "Cytoskeletal.Components"          
+## [19] "Trafficking.Proteins"             
+## [20] "Protein.Complexes"                
+## [21] "Signaling.Cascades"               
+## [22] "Oct4.Interactome"                 
+## [23] "Nanog.Interactome"                
+## [24] "Sox2.Interactome"                 
+## [25] "Cell.Surface.Proteins"
+\end{verbatim}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{3}\hlstd{]} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"uniprot.accession"}\hlstd{,} \hlstr{"uniprot.id"}\hlstd{,} \hlstr{"description"}\hlstd{)}
+\hlkwd{fvarLabels}\hlstd{(hl)[}\hlnum{4}\hlopt{:}\hlnum{6}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlstr{"peptides.expt"}\hlstd{,} \hlnum{1}\hlopt{:}\hlnum{3}\hlstd{)}
+\hlcom{## feature vars 1, 2, and 4 to 6}
+\hlkwd{fData}\hlstd{(hl)[}\hlnum{1}\hlopt{:}\hlnum{4}\hlstd{,} \hlkwd{c}\hlstd{(}\hlnum{1}\hlopt{:}\hlnum{2}\hlstd{,} \hlnum{4}\hlopt{:}\hlnum{6}\hlstd{)]}
+\end{alltt}
+\begin{verbatim}
+##          uniprot.accession  uniprot.id peptides.expt1 peptides.expt2
+## Q9JHU4              Q9JHU4 DYHC1_MOUSE            175            166
+## Q9QXS1-3          Q9QXS1-3  PLEC_MOUSE            123            150
+## Q9ERU9              Q9ERU9  RBP2_MOUSE            101             90
+## P26039              P26039  TLN1_MOUSE            101             94
+##          peptides.expt3
+## Q9JHU4              322
+## Q9QXS1-3            174
+## Q9ERU9              181
+## P26039              167
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<pdata>>=
-pData(hl)$Replicate <- rep(1:2, each = 10)
-pData(hl)$Tag <- sub("\\.1$", "", sub("^X", "", sampleNames(hl)))
-pData(hl)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Replicate} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{1}\hlopt{:}\hlnum{2}\hlstd{,} \hlkwc{each} \hlstd{=} \hlnum{10}\hlstd{)}
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Tag} \hlkwb{<-} \hlkwd{sub}\hlstd{(}\hlstr{"\textbackslash{}\textbackslash{}.1$"}\hlstd{,} \hlstr{""}\hlstd{,} \hlkwd{sub}\hlstd{(}\hlstr{"^X"}\hlstd{,} \hlstr{""}\hlstd{,} \hlkwd{sampleNames}\hlstd{(hl)))}
+\hlkwd{pData}\hlstd{(hl)}
+\end{alltt}
+\begin{verbatim}
+##         Replicate  Tag
+## X126            1  126
+## X127N           1 127N
+## X127C           1 127C
+## X128N           1 128N
+## X128C           1 128C
+## X129N           1 129N
+## X129C           1 129C
+## X130N           1 130N
+## X130C           1 130C
+## X131            1  131
+## X126.1          2  126
+## X127N.1         2 127N
+## X127C.1         2 127C
+## X128N.1         2 128N
+## X128C.1         2 128C
+## X129N.1         2 129N
+## X129C.1         2 129C
+## X130N.1         2 130N
+## X130C.1         2 130C
+## X131.1          2  131
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<morepdata>>=
-expinfo <- dir(extdatadir, full.names = TRUE,
-               pattern = "hyperLOPIT-SIData-fraction-info.csv")
-
-fracinfo <- read.csv(expinfo, row.names=1, skip = 2, 
-                     header = FALSE, stringsAsFactors = FALSE)
-
-pData(hl)$Gradient.Fraction <- c(fracinfo[, 1], fracinfo[, 2])
-pData(hl)$Iodixonal.Density <- c(fracinfo[, 4], fracinfo[, 5])
-pData(hl)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{expinfo} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+               \hlkwc{pattern} \hlstd{=} \hlstr{"hyperLOPIT-SIData-fraction-info.csv"}\hlstd{)}
+
+\hlstd{fracinfo} \hlkwb{<-} \hlkwd{read.csv}\hlstd{(expinfo,} \hlkwc{row.names}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{skip} \hlstd{=} \hlnum{2}\hlstd{,}
+                     \hlkwc{header} \hlstd{=} \hlnum{FALSE}\hlstd{,} \hlkwc{stringsAsFactors} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Gradient.Fraction} \hlkwb{<-} \hlkwd{c}\hlstd{(fracinfo[,} \hlnum{1}\hlstd{], fracinfo[,} \hlnum{2}\hlstd{])}
+\hlkwd{pData}\hlstd{(hl)}\hlopt{$}\hlstd{Iodixonal.Density} \hlkwb{<-} \hlkwd{c}\hlstd{(fracinfo[,} \hlnum{4}\hlstd{], fracinfo[,} \hlnum{5}\hlstd{])}
+\hlkwd{pData}\hlstd{(hl)}
+\end{alltt}
+\begin{verbatim}
+##         Replicate  Tag Gradient.Fraction Iodixonal.Density
+## X126            1  126           Cytosol               0.0
+## X127N           1 127N   1 to 6 (pooled)               6.0
+## X127C           1 127C   8 to 9 (pooled)              11.0
+## X128N           1 128N 10 to 11 (pooled)              13.3
+## X128C           1 128C                12              14.6
+## X129N           1 129N                14              17.4
+## X129C           1 129C                16              20.1
+## X130N           1 130N                18              26.8
+## X130C           1 130C         Chromatin                NA
+## X131            1  131                19              34.5
+## X126.1          2  126           Cytosol               0.0
+## X127N.1         2 127N   1 to 6 (pooled)               5.2
+## X127C.1         2 127C   7 to 9 (pooled)              10.0
+## X128N.1         2 128N 10 to 11 (pooled)              12.5
+## X128C.1         2 128C                12              14.0
+## X129N.1         2 129N 14 to 15 (pooled)              17.3
+## X129C.1         2 129C                17              20.9
+## X130N.1         2 130N 18 to 19 (pooled)              24.7
+## X130C.1         2 130C         Chromatin                NA
+## X131.1          2  131                20              31.9
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into \sout{a}\hl{the} same intensity interval (typically 0 and 1) by, for
+profiles into \st{a} \hl{the} same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
 
-<<normsum, eval = FALSE>>= 
-hl <- normalise(hl, method = "sum") 
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{normalise}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"sum"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<normcheck>>=
-summary(rowSums(exprs(hl[, hl$Replicate == 1])))
-summary(rowSums(exprs(hl[, hl$Replicate == 2])))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{1}\hlstd{])))}
+\end{alltt}
+\begin{verbatim}
+##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+##   0.997   0.999   1.000   1.000   1.001   1.003
+\end{verbatim}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(}\hlkwd{rowSums}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{2}\hlstd{])))}
+\end{alltt}
+\begin{verbatim}
+##    Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
+##   0.997   0.999   1.000   1.000   1.001   1.003
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<reps4combine>>=
-library("pRolocdata")
-data(hyperLOPIT2015ms3r1)
-data(hyperLOPIT2015ms3r2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRolocdata"}\hlstd{)}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r1)}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<pkgdata, eval=FALSE>>=
-pRolocdata()
-@
-
-will list the \Sexpr{nrow(pRolocdata()$results)} datasets that are
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{pRolocdata}\hlstd{()}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+will list the 75 datasets that are
 available in \Biocexptpkg{pRolocdata}.
@@
 
-<<samplenames>>=
-identical(sampleNames(hyperLOPIT2015ms3r1), sampleNames(hyperLOPIT2015ms3r2))
-hyperLOPIT2015ms3r1 <- updateSampleNames(hyperLOPIT2015ms3r1, 1)
-hyperLOPIT2015ms3r2 <- updateSampleNames(hyperLOPIT2015ms3r2, 2)
-sampleNames(hyperLOPIT2015ms3r1)
-sampleNames(hyperLOPIT2015ms3r2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{identical}\hlstd{(}\hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r1),} \hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r2))}
+\end{alltt}
+\begin{verbatim}
+## [1] TRUE
+\end{verbatim}
+\begin{alltt}
+\hlstd{hyperLOPIT2015ms3r1} \hlkwb{<-} \hlkwd{updateSampleNames}\hlstd{(hyperLOPIT2015ms3r1,} \hlnum{1}\hlstd{)}
+\hlstd{hyperLOPIT2015ms3r2} \hlkwb{<-} \hlkwd{updateSampleNames}\hlstd{(hyperLOPIT2015ms3r2,} \hlnum{2}\hlstd{)}
+\hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r1)}
+\end{alltt}
+\begin{verbatim}
+##  [1] "X126.1"  "X127N.1" "X127C.1" "X128N.1" "X128C.1" "X129N.1" "X129C.1"
+##  [8] "X130N.1" "X130C.1" "X131.1"
+\end{verbatim}
+\begin{alltt}
+\hlkwd{sampleNames}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
+\begin{verbatim}
+##  [1] "X126.2"  "X127N.2" "X127C.2" "X128N.2" "X128C.2" "X129N.2" "X129C.2"
+##  [8] "X130N.2" "X130C.2" "X131.2"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<fvarnames>>=
-fvarLabels(hyperLOPIT2015ms3r1)
-fvarLabels(hyperLOPIT2015ms3r2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)}
+\end{alltt}
+\begin{verbatim}
+## [1] "EntryName"          "ProteinDescription" "Peptides"          
+## [4] "PSMs"               "ProteinCoverage"    "markers"
+\end{verbatim}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
+\begin{verbatim}
+## [1] "EntryName"          "ProteinDescription" "Peptides"          
+## [4] "PSMs"               "ProteinCoverage"    "markers"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<fvarnames2>>=
-fvarLabels(hyperLOPIT2015ms3r1)[3:5] <- paste0(fvarLabels(hyperLOPIT2015ms3r1)[3:5], 1)
-hyperLOPIT2015ms3r2 <- updateFvarLabels(hyperLOPIT2015ms3r2, "2", sep = "")
-fData(hyperLOPIT2015ms3r1) <- fData(hyperLOPIT2015ms3r1)[1:5]
-fData(hyperLOPIT2015ms3r2) <- fData(hyperLOPIT2015ms3r2)[3:5]
-fvarLabels(hyperLOPIT2015ms3r1)
-fvarLabels(hyperLOPIT2015ms3r2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]} \hlkwb{<-} \hlkwd{paste0}\hlstd{(}\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{],} \hlnum{1}\hlstd{)}
+\hlstd{hyperLOPIT2015ms3r2} \hlkwb{<-} \hlkwd{updateFvarLabels}\hlstd{(hyperLOPIT2015ms3r2,} \hlstr{"2"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
+\hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r1)} \hlkwb{<-} \hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r1)[}\hlnum{1}\hlopt{:}\hlnum{5}\hlstd{]}
+\hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r2)} \hlkwb{<-} \hlkwd{fData}\hlstd{(hyperLOPIT2015ms3r2)[}\hlnum{3}\hlopt{:}\hlnum{5}\hlstd{]}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r1)}
+\end{alltt}
+\begin{verbatim}
+## [1] "EntryName"          "ProteinDescription" "Peptides1"         
+## [4] "PSMs1"              "ProteinCoverage1"
+\end{verbatim}
+\begin{alltt}
+\hlkwd{fvarLabels}\hlstd{(hyperLOPIT2015ms3r2)}
+\end{alltt}
+\begin{verbatim}
+## [1] "Peptides2"        "PSMs2"            "ProteinCoverage2"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<combine, cache = cache>>=
-combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
-combined
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
+\end{alltt}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning: 'combine' is deprecated.\\\#\# Use 'gtable\_combine' instead.\\\#\# See help("{}Deprecated"{})}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[[<-.data.frame`(`*tmp*`, i, value = structure(list(t = numeric(0), : replacement has 0 rows, data has 10}}\begin{alltt}
+\hlstd{combined}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 5032 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData
+##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+##   varMetadata: labelDescription
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+##     total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
+## Annotation:  
+## - - - Processing information - - -
+## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
+## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
+## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
+## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
+##  MSnbase version: 2.5.9
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 data and their inappropriate handling leads to a reduction of
-sub-cellular resolution. \sout{Missing data can be imputated} \hl{We can impute missing data} using
+sub-cellular resolution. \st{Missing data can be imputated} \hl{We can impute missing data} using
 \Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
@@
 
-<<impute, eval=FALSE>>=
-hl <- impute(hl, method = "knn")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
   \centering
-<<namap, out.width=".65\\textwidth">>=
-image2(is.na(combined), col = c("black", "white"),
-       main = "Missing values (white cells) after combining replicates")
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
+       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/namap-1}
+
+\end{knitrout}
 \caption{Heatmap of missing values. Note that the features are
@@
 
-<<filterNA>>=
-combined <- filterNA(combined)
-combined
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
+\hlstd{combined}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 5032 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData
+##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+##   varMetadata: labelDescription
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+##     total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
+## Annotation:  
+## - - - Processing information - - -
+## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
+## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
+## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
+## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
+## Subset [5032,20][5032,20] Tue Apr  3 17:46:47 2018 
+## Removed features with more than 0 NAs: Tue Apr  3 17:46:47 2018 
+## Dropped featureData's levels Tue Apr  3 17:46:48 2018 
+##  MSnbase version: 2.5.9
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
   \centering
-<<qcplot, fig.show = "hold", message=FALSE, warning=FALSE, out.width=".45\\textwidth">>=
-library("pRoloc")
-plot2D(hl, fcol = NULL, col = "black")
-plot2D(hl, method = "hexbin")
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRoloc"}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlkwa{NULL}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"black"}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"hexbin"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.45\textwidth]{figure/qcplot-1} 
+\includegraphics[width=.45\textwidth]{figure/qcplot-2}
+
+\end{knitrout}
 \caption{PCA plot of the mouse stem cell data \texttt{hl}. Each dot
@@
   \centering
-<<qcbx, out.width="\\textwidth", fig.asp=1/2>>=
-par(mfrow = c(1, 2), ## creates a two-panel figure
-    las = 2,         ## axis labels orientation
-    cex.axis = .7)   ## axis label size
-o <- order(hl$Iodixonal.Density)
-plotDist(hl[, o], pcol = "#00000010", xlab = "")
-lines(colMeans(exprs(hl[, o])), col = "red", type = "b")
-boxplot(exprs(hl[, o]))
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{),} \hlcom{## creates a two-panel figure}
+    \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{,}         \hlcom{## axis labels orientation}
+    \hlkwc{cex.axis} \hlstd{=} \hlnum{.7}\hlstd{)}   \hlcom{## axis label size}
+\hlstd{o} \hlkwb{<-} \hlkwd{order}\hlstd{(hl}\hlopt{$}\hlstd{Iodixonal.Density)}
+\hlkwd{plotDist}\hlstd{(hl[, o],} \hlkwc{pcol} \hlstd{=} \hlstr{"#00000010"}\hlstd{,} \hlkwc{xlab} \hlstd{=} \hlstr{""}\hlstd{)}
+\hlkwd{lines}\hlstd{(}\hlkwd{colMeans}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o])),} \hlkwc{col} \hlstd{=} \hlstr{"red"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"b"}\hlstd{)}
+\hlkwd{boxplot}\hlstd{(}\hlkwd{exprs}\hlstd{(hl[, o]))}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/qcbx-1}
+
+\end{knitrout}
 \caption{Protein profiles and distribution of channel intensities. The
@@
 
-<<markers>>=
-## List available marker sets
-pRolocmarkers()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## List available marker sets}
+\hlkwd{pRolocmarkers}\hlstd{()}
+\end{alltt}
+\begin{verbatim}
+## 7 marker lists available:
+## Arabidopsis thaliana [atha]:
+##  Ids: TAIR, 543 markers
+## Drosophila melanogaster [dmel]:
+##  Ids: Uniprot, 179 markers
+## Gallus gallus [ggal]:
+##  Ids: IPI, 102 markers
+## Homo sapiens [hsap]:
+##  Ids: Uniprot, 872 markers
+## Mus musculus [mmus]:
+##  Ids: Uniprot, 937 markers
+## Saccharomyces cerevisiae [scer_sgd]:
+##  Ids: SGD, 259 markers
+## Saccharomyces cerevisiae [scer_uniprot]:
+##  Ids: Uniprot Accession, 259 markers
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<addmarkers>>=
-## Use mouse markers
-mrk <- pRolocmarkers(species = "mmus")
-head(mrk)
-
-## Add mouse markers
-hl <- addMarkers(hl, mrk)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Use mouse markers}
+\hlstd{mrk} \hlkwb{<-} \hlkwd{pRolocmarkers}\hlstd{(}\hlkwc{species} \hlstd{=} \hlstr{"mmus"}\hlstd{)}
+\hlkwd{head}\hlstd{(mrk)}
+\end{alltt}
+\begin{verbatim}
+##                 P26039                 Q6PB66                 P11276 
+##   "Actin cytoskeleton"        "Mitochondrion" "Extracellular matrix" 
+##                 Q6PR54                 Q05793                 P19096 
+##  "Nucleus - Chromatin" "Extracellular matrix"              "Cytosol"
+\end{verbatim}
+\begin{alltt}
+\hlcom{## Add mouse markers}
+\hlstd{hl} \hlkwb{<-} \hlkwd{addMarkers}\hlstd{(hl, mrk)}
+\end{alltt}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Markers in data: 937 out of 5032}}\begin{verbatim}
+## organelleMarkers
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##                      27                      43                      13 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##                      43                      95                      12 
+##    Extracellular matrix         Golgi apparatus                Lysosome 
+##                      10                      27                      33 
+##           Mitochondrion     Nucleus - Chromatin Nucleus - Non-chromatin 
+##                     383                      64                      85 
+##              Peroxisome         Plasma membrane              Proteasome 
+##                      17                      51                      34 
+##                 unknown 
+##                    4095
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 \ref{fig:plotmarkers}).  \hl{This is further highlighted if we plot the
-profiles of these two clusters using the \texttt{plotDist} function
-(Fig \ref{fig:plotDist2}).} \hl{The \texttt{plotDist} function is
+profiles of these two clusters using the \texttt{plotDist} function}
+(Fig \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
 another useful visualisation that relies on marker annotation. It
 allows one to represent the protein profiles occupancy along the
-gradient \sout{using the \texttt{plotDist} function}. While the PCA
+gradient} \st{using the \texttt{plotDist} function}. \hl{While the PCA
 plot enables efficient visualisation of the complete dataset and
@@
 exactly they differ (in terms of peak channels, for example). On
-figure \ref{fig:plotDist2}, we plot the profile of the mitochondrial
+figure} \ref{fig:plotDist2}, \hl{we plot the profile of the mitochondrial
 and peroxisome markers to highlight the differences in profiles
@@
 129C, as represented above along the 7th PC on the PCA plot on
-figure \ref{fig:plotmarkers}.}
+figure} \ref{fig:plotmarkers}.
 
@@
   \centering
-<<plotmarkers, out.width="\\textwidth", fig.width=12, fig.asp=1/2>>=
-par(mfrow = c(1, 2))
-plot2D(hl, main = "pRolocmarkers for mouse")
-addLegend(hl, cex = .6)
-plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{main} \hlstd{=} \hlstr{"pRolocmarkers for mouse"}\hlstd{)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{7}\hlstd{),} \hlkwc{main} \hlstd{=} \hlstr{"Marker resolution along PC 1 and 7"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plotmarkers-1}
+
+\end{knitrout}
 \caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
@@
   \centering
-<<plotDist, out.width=".6\\textwidth", fig.asp=3/5>>=
-hlo <- hl[, order(hl$Iodixonal.Density)]
-plotDist(hlo[fData(hlo)$markers == "Mitochondrion", ],
-         pcol = "purple", fractions = "Tag")
-title(main = "Marker occupancy profiles along the gradient")
-matlines(t(exprs(hlo[fData(hlo)$markers == "Peroxisome", ])),
-         lty = 1, col = "darkblue", type = "l")
-legend("topleft", c("Mitochondrion", "Peroxisome"),
-       lty = 1, col = c("purple", "blue"), bty = "n")
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hlo} \hlkwb{<-} \hlstd{hl[,} \hlkwd{order}\hlstd{(hl}\hlopt{$}\hlstd{Iodixonal.Density)]}
+\hlkwd{plotDist}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Mitochondrion"}\hlstd{, ],}
+         \hlkwc{pcol} \hlstd{=} \hlstr{"purple"}\hlstd{,} \hlkwc{fractions} \hlstd{=} \hlstr{"Tag"}\hlstd{)}
+\hlkwd{title}\hlstd{(}\hlkwc{main} \hlstd{=} \hlstr{"Marker occupancy profiles along the gradient"}\hlstd{)}
+\hlkwd{matlines}\hlstd{(}\hlkwd{t}\hlstd{(}\hlkwd{exprs}\hlstd{(hlo[}\hlkwd{fData}\hlstd{(hlo)}\hlopt{$}\hlstd{markers} \hlopt{==} \hlstr{"Peroxisome"}\hlstd{, ])),}
+         \hlkwc{lty} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{col} \hlstd{=} \hlstr{"darkblue"}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"l"}\hlstd{)}
+\hlkwd{legend}\hlstd{(}\hlstr{"topleft"}\hlstd{,} \hlkwd{c}\hlstd{(}\hlstr{"Mitochondrion"}\hlstd{,} \hlstr{"Peroxisome"}\hlstd{),}
+       \hlkwc{lty} \hlstd{=} \hlnum{1}\hlstd{,} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"purple"}\hlstd{,} \hlstr{"blue"}\hlstd{),} \hlkwc{bty} \hlstd{=} \hlstr{"n"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.6\textwidth]{figure/plotDist-1}
+
+\end{knitrout}
   \caption{Mitochondrion and peroxisome protein profiles.}
@@
   \centering
-<<plotmarkers3d, eval=FALSE>>=
-plot3D(hl, dims = c(1, 2, 7))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot3D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{,} \hlnum{7}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
   \includegraphics[width=.55\textwidth]{./Figures/plot3d.pdf}
@@
 
-<<removeGA>>=
-hl <- fDataToUnknown(hl, from = "Golgi apparatus", to = "unknown")
-getMarkers(hl)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{fDataToUnknown}\hlstd{(hl,} \hlkwc{from} \hlstd{=} \hlstr{"Golgi apparatus"}\hlstd{,} \hlkwc{to} \hlstd{=} \hlstr{"unknown"}\hlstd{)}
+\hlkwd{getMarkers}\hlstd{(hl)}
+\end{alltt}
+\begin{verbatim}
+## organelleMarkers
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##                      27                      43                      13 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##                      43                      95                      12 
+##    Extracellular matrix                Lysosome           Mitochondrion 
+##                      10                      33                     383 
+##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+##                      64                      85                      17 
+##         Plasma membrane              Proteasome                 unknown 
+##                      51                      34                    4122
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 object \hl{and then use the \textit{highlightOnPlot} function to 
-visualise these}. For example, if we wanted to highlight \sout{a} 
+visualise these}. For example, if we wanted to highlight \st{a} 
 proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, 
@@
 we \hl{would call the following code:}
-\sout{would create a first create a \texttt{FeaturesOfInterest} object, and
+\st{would create a first create a \texttt{FeaturesOfInterest} object, and
 subsequently highlight their location on a PCA plot with the
@@
   \centering
-<<foi, fig.width=10, out.width=".8\\textwidth">>=
-prots <- c("Q8CG48", "Q8CG47", "Q8K2Z4", "Q8C156")
-foi13s <- FeaturesOfInterest(description = "13S consensin proteins",
-                             fnames = prots, 
-                             object = hl)
-foi13s
-
-plot2D(hl)
-addLegend(hl, cex = .6)
-highlightOnPlot(hl, foi13s)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{prots} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"Q8CG48"}\hlstd{,} \hlstr{"Q8CG47"}\hlstd{,} \hlstr{"Q8K2Z4"}\hlstd{,} \hlstr{"Q8C156"}\hlstd{)}
+\hlstd{foi13s} \hlkwb{<-} \hlkwd{FeaturesOfInterest}\hlstd{(}\hlkwc{description} \hlstd{=} \hlstr{"13S consensin proteins"}\hlstd{,}
+                             \hlkwc{fnames} \hlstd{= prots,}
+                             \hlkwc{object} \hlstd{= hl)}
+\hlstd{foi13s}
+\end{alltt}
+\begin{verbatim}
+## Traceable object of class "FeaturesOfInterest"
+##  Created on Tue Apr  3 17:46:49 2018 
+##  Description:
+##   13S consensin proteins
+##  4 features of interest:
+##    Q8CG48, Q8CG47, Q8K2Z4, Q8C156
+\end{verbatim}
+\begin{alltt}
+\hlkwd{plot2D}\hlstd{(hl)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
+\hlkwd{highlightOnPlot}\hlstd{(hl, foi13s)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/foi-1}
+
+\end{knitrout}
   \caption{Highlighting protein features of interest.}
@@
 
-<<pdtab, echo=FALSE, eval = TRUE, results = 'asis'>>=
-i <- c(3, 13)
-xtable(pData(hl)[i, ], 
-       caption = "Differences in gradient fraction pooling.", 
-       label = "tab:pdtab")
-@
+% latex table generated in R 3.5.0 by xtable 1.8-2 package
+% Tue Apr  3 17:46:49 2018
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrllr}
+  \hline
+ & Replicate & Tag & Gradient.Fraction & Iodixonal.Density \\ 
+  \hline
+X127C &   1 & 127C & 8 to 9 (pooled) & 11.00 \\ 
+  X127C.1 &   2 & 127C & 7 to 9 (pooled) & 10.00 \\ 
+   \hline
+\end{tabular}
+\caption{Differences in gradient fraction pooling.} 
+\label{tab:pdtab}
+\end{table}
 
@@
   \centering
-<<plot2Drep, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>= 
-par(mfrow = c(1, 2)) 
-plot2D(hl[, hl$Replicate == 1], main = "Replicate 1")
-plot2D(hl[, hl$Replicate == 2], main = "Replicate 2", mirrorX = TRUE)
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{1}\hlstd{],} \hlkwc{main} \hlstd{=} \hlstr{"Replicate 1"}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl[, hl}\hlopt{$}\hlstd{Replicate} \hlopt{==} \hlnum{2}\hlstd{],} \hlkwc{main} \hlstd{=} \hlstr{"Replicate 2"}\hlstd{,} \hlkwc{mirrorX} \hlstd{=} \hlnum{TRUE}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plot2Drep-1}
+
+\end{knitrout}
   \caption{PCA plots of replicates 1 and 2}
@@
 
-<<mainapp, eval = FALSE, echo = TRUE>>=
-library("pRolocGUI")
-pRolocVis(hl)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRolocGUI"}\hlstd{)}
+\hlkwd{pRolocVis}\hlstd{(hl)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<compareapp, eval = FALSE, echo = TRUE>>=
-data(hyperLOPIT2015ms3r1)
-data(hyperLOPIT2015ms3r2)
-hllst <- MSnSetList(list(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))
-pRolocVis(hllst, app = "compare")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r1)}
+\hlkwd{data}\hlstd{(hyperLOPIT2015ms3r2)}
+\hlstd{hllst} \hlkwb{<-} \hlkwd{MSnSetList}\hlstd{(}\hlkwd{list}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2))}
+\hlkwd{pRolocVis}\hlstd{(hllst,} \hlkwc{app} \hlstd{=} \hlstr{"compare"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<displayPDmarkers>>=
-getMarkers(hl, fcol = "phenoDisco.Input")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## organelleMarkers
+##                          40S Ribosome 
+##                                    26 
+##                          60S Ribosome 
+##                                    43 
+## Endoplasmic reticulum/Golgi apparatus 
+##                                    76 
+##                         Mitochondrion 
+##                                   261 
+##                       Plasma membrane 
+##                                    50 
+##                            Proteasome 
+##                                    34 
+##                               unknown 
+##                                  4542
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<runPD, eval = FALSE>>=
-## As per Christoforou et al (2016), 
-hl <- phenoDisco(hl, fcol = "phenoDisco.Input", times = 200, GS = 60)
-@
-
-\sout{Note: We do not evaluate this code chunk in this document as the
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## As per Christoforou et al (2016), }
+\hlstd{hl} \hlkwb{<-} \hlkwd{phenoDisco}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,} \hlkwc{times} \hlstd{=} \hlnum{200}\hlstd{,} \hlkwc{GS} \hlstd{=} \hlnum{60}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\st{Note: We do not evaluate this code chunk in this document as the
 algorithm is computationally intensive and best parallelised over
@@
 
-<<loadPDres, echo=FALSE>>=
-f0 <- dir(extdatadir, full.names = TRUE,
-          pattern = "bpw-pdres.rds")
-pdres <- readRDS(f0)
-hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
-setStockcol(NULL)
-setStockcol(paste0(getStockcol(), 70))
-@
-
-<<showPDres>>=
-hl
-getMarkers(hl, fcol = "pd")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 5032 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData
+##   sampleNames: X126 X127N ... X131.1 (20 total)
+##   varLabels: Replicate Tag Gradient.Fraction Iodixonal.Density
+##   varMetadata: labelDescription
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
+##   fvarLabels: uniprot.accession uniprot.id ... pd (27 total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+## Annotation:  
+## - - - Processing information - - -
+## Added markers from  'mrk' marker vector. Tue Apr  3 17:46:49 2018 
+## Added markers from  'pdres' marker vector. Tue Apr  3 17:46:50 2018 
+##  MSnbase version: 2.5.10
+\end{verbatim}
+\begin{alltt}
+\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+## organelleMarkers
+##                          40S Ribosome 
+##                                   106 
+##                          60S Ribosome 
+##                                    95 
+## Endoplasmic reticulum/Golgi apparatus 
+##                                   393 
+##                         Mitochondrion 
+##                                   525 
+##                           Phenotype 1 
+##                                   300 
+##                           Phenotype 2 
+##                                   253 
+##                           Phenotype 3 
+##                                   203 
+##                           Phenotype 4 
+##                                    74 
+##                           Phenotype 5 
+##                                    91 
+##                       Plasma membrane 
+##                                   421 
+##                            Proteasome 
+##                                    92 
+##                               unknown 
+##                                  2479
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
   \centering
-<<plotPDres, fig.width=12, out.width="\\textwidth", fig.asp=1/2>>=
-## Re-order the colours for the phenoDisco output
-cl <- getMarkerClasses(hl, "pd")
-cols <- getStockcol()[seq(cl)]
-ind <- grep("Pheno", cl, invert = TRUE)
-cols[ind] <- getStockcol()[seq(cl)][1:length(ind)]
-cols[-ind] <- getStockcol()[seq(cl)][(length(ind) + 1):length(cl)]
-
-## Plot the input and output
-par(mfrow = c(1, 2))
-plot2D(hl, fcol = "phenoDisco.Input", 
-       main = "phenoDisco input markers", col = getStockcol()[1:6])
-addLegend(hl, fcol = "phenoDisco.Input", cex = .7)
-plot2D(hl, fcol = "pd", main = "phenoDisco output", col = cols)
-addLegend(hl, fcol = "pd", cex = .7, col = cols)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Re-order the colours for the phenoDisco output}
+\hlstd{cl} \hlkwb{<-} \hlkwd{getMarkerClasses}\hlstd{(hl,} \hlstr{"pd"}\hlstd{)}
+\hlstd{cols} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)]}
+\hlstd{ind} \hlkwb{<-} \hlkwd{grep}\hlstd{(}\hlstr{"Pheno"}\hlstd{, cl,} \hlkwc{invert} \hlstd{=} \hlnum{TRUE}\hlstd{)}
+\hlstd{cols[ind]} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)][}\hlnum{1}\hlopt{:}\hlkwd{length}\hlstd{(ind)]}
+\hlstd{cols[}\hlopt{-}\hlstd{ind]} \hlkwb{<-} \hlkwd{getStockcol}\hlstd{()[}\hlkwd{seq}\hlstd{(cl)][(}\hlkwd{length}\hlstd{(ind)} \hlopt{+} \hlnum{1}\hlstd{)}\hlopt{:}\hlkwd{length}\hlstd{(cl)]}
+
+\hlcom{## Plot the input and output}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,}
+       \hlkwc{main} \hlstd{=} \hlstr{"phenoDisco input markers"}\hlstd{,} \hlkwc{col} \hlstd{=} \hlkwd{getStockcol}\hlstd{()[}\hlnum{1}\hlopt{:}\hlnum{6}\hlstd{])}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{.7}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{main} \hlstd{=} \hlstr{"phenoDisco output"}\hlstd{,} \hlkwc{col} \hlstd{= cols)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{.7}\hlstd{,} \hlkwc{col} \hlstd{= cols)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plotPDres-1}
+
+\end{knitrout}
   \caption{Results of the novelty detection algorithm.}
@@
 
-<<writehl, eval=FALSE>>=
-fData(hl)$pd <- as.character(fData(hl)$pd)
-write.exprs(hl, fDataCols = "pd", file = "pd-results.csv", sep = ",")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{pd} \hlkwb{<-} \hlkwd{as.character}\hlstd{(}\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{pd)}
+\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{fDataCols} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{file} \hlstd{=} \hlstr{"pd-results.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<vishl, eval=FALSE>>=
-pRolocVis(hl, fcol = "pd")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{pRolocVis}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<loadpdRes, echo=FALSE>>=
-svmf <- dir(extdatadir, full.names = TRUE,
-              pattern = "bpw-svmopt.rds") 
-params <- readRDS(svmf)
-@
-
-<<getsvmweight>>=
-w <- table(getMarkers(hl, verbose = TRUE))
-w <- 1/w[names(w) != "unknown"]
-@
-
-<<dosvmopt, eval=FALSE>>=
-## 100 rounds of optimisation with five-fold cross-validation
-params <- svmOptimisation(hl, fcol = "markers",
-                          times = 100, xval = 5,
-                          class.weights = w)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{w} \hlkwb{<-} \hlkwd{table}\hlstd{(}\hlkwd{getMarkers}\hlstd{(hl,} \hlkwc{verbose} \hlstd{=} \hlnum{TRUE}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+## organelleMarkers
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##                      27                      43                      13 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##                      43                      95                      12 
+##    Extracellular matrix                Lysosome           Mitochondrion 
+##                      10                      33                     383 
+##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+##                      64                      85                      17 
+##         Plasma membrane              Proteasome                 unknown 
+##                      51                      34                    4122
+\end{verbatim}
+\begin{alltt}
+\hlstd{w} \hlkwb{<-} \hlnum{1}\hlopt{/}\hlstd{w[}\hlkwd{names}\hlstd{(w)} \hlopt{!=} \hlstr{"unknown"}\hlstd{]}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## 100 rounds of optimisation with five-fold cross-validation}
+\hlstd{params} \hlkwb{<-} \hlkwd{svmOptimisation}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{,}
+                          \hlkwc{times} \hlstd{=} \hlnum{100}\hlstd{,} \hlkwc{xval} \hlstd{=} \hlnum{5}\hlstd{,}
+                          \hlkwc{class.weights} \hlstd{= w)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<f1count>>=
-f1Count(params, 0.6)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{f1Count}\hlstd{(params,} \hlnum{0.6}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+##     4  8 16
+## 0.1 1 10 89
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<visualiseOpt, eval=FALSE>>=
-plot(params)
-levelPlot(params)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot}\hlstd{(params)}
+\hlkwd{levelPlot}\hlstd{(params)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
   \centering
-<<visualiseOptHide, echo=FALSE, out.width=".65\\textwidth">>=
-library('gridExtra')
-grid.arrange(plot(params), levelPlot(params), ncol = 2)
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=.65\textwidth]{figure/visualiseOptHide-1}
+
+\end{knitrout}
   \caption{Assessment of the classification model parameter optimisation.}
@@
 
-<<getbestpars>>=
-(best <- getParams(params))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{(best} \hlkwb{<-} \hlkwd{getParams}\hlstd{(params))}
+\end{alltt}
+\begin{verbatim}
+## sigma  cost 
+##   0.1  16.0
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<classify>>=
-hl <- svmClassification(hl, params, class.weights = w, fcol = "markers")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl, params,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning in seq\_len(trpr): first element used of 'length.out' argument}}\end{kframe}
+\end{knitrout}
 
@@
 
-<<classify2, eval = FALSE>>=
-hl <- svmClassification(hl, cost = 16, sigma = 0.1, class.weights = w, fcol = "markers")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{svmClassification}\hlstd{(hl,} \hlkwc{cost} \hlstd{=} \hlnum{16}\hlstd{,} \hlkwc{sigma} \hlstd{=} \hlnum{0.1}\hlstd{,} \hlkwc{class.weights} \hlstd{= w,} \hlkwc{fcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<setColors, include=FALSE>>=
-setStockcol(NULL)
-setStockcol(paste0(getStockcol(), 70))
-par(mfrow = c(1, 1))
-@
-
 \begin{figure}[!ht]
   \centering
-<<plotSVM, out.width=".8\\textwidth">>=
-## set point size of each protein to be inversely proportional to the
-ptsze <- exp(fData(hl)$svm.scores) - 1
-## plot new predictions
-plot2D(hl, fcol = "svm", cex = ptsze)
-addLegend(hl, fcol = "svm", where = "bottomleft", bty = "n", cex = .5)
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## set point size of each protein to be inversely proportional to the}
+\hlstd{ptsze} \hlkwb{<-} \hlkwd{exp}\hlstd{(}\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{svm.scores)} \hlopt{-} \hlnum{1}
+\hlcom{## plot new predictions}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{cex} \hlstd{= ptsze)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{where} \hlstd{=} \hlstr{"bottomleft"}\hlstd{,} \hlkwc{bty} \hlstd{=} \hlstr{"n"}\hlstd{,} \hlkwc{cex} \hlstd{=} \hlnum{.5}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/plotSVM-1}
+
+\end{knitrout}
 \caption{Classification results. Colours indicate class membership and
@@
   \centering
-<<thresholds, out.width=".8\\textwidth">>=
-## First remove the markers
-preds <- unknownMSnSet(hl)
-## Plot a boxplot of the scores of each organelle
-par(oma = c(10.5, 0, 0, 0)) ## sets outer margins
-boxplot(svm.scores ~ svm, data = fData(preds), 
-        ylab = "SVM scores", las = 2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## First remove the markers}
+\hlstd{preds} \hlkwb{<-} \hlkwd{unknownMSnSet}\hlstd{(hl)}
+\hlcom{## Plot a boxplot of the scores of each organelle}
+\hlkwd{par}\hlstd{(}\hlkwc{oma} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{10.5}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{,} \hlnum{0}\hlstd{))} \hlcom{## sets outer margins}
+\hlkwd{boxplot}\hlstd{(svm.scores} \hlopt{~} \hlstd{svm,} \hlkwc{data} \hlstd{=} \hlkwd{fData}\hlstd{(preds),}
+        \hlkwc{ylab} \hlstd{=} \hlstr{"SVM scores"}\hlstd{,} \hlkwc{las} \hlstd{=} \hlnum{2}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.8\textwidth]{figure/thresholds-1}
+
+\end{knitrout}
   \caption{Visualistion of class-specific classification score distribution.}
@@
 
-<<orgquants>>=
-(ts <- orgQuants(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = .5))
-hl <- getPredictions(hl, fcol = "svm", scol = "svm.scores", mcol = "markers", t = ts)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{(ts} \hlkwb{<-} \hlkwd{orgQuants}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{=} \hlnum{.5}\hlstd{))}
+\end{alltt}
+\begin{verbatim}
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##               0.4333396               0.3051261               0.3870872 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##               0.6736435               0.6011995               0.4213771 
+##    Extracellular matrix                Lysosome           Mitochondrion 
+##               0.4135168               0.5581414               0.9497578 
+##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+##               0.7956060               0.7091274               0.3126237 
+##         Plasma membrane              Proteasome 
+##               0.7100209               0.4188587
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##               0.4333396               0.3051261               0.3870872 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##               0.6736435               0.6011995               0.4213771 
+##    Extracellular matrix                Lysosome           Mitochondrion 
+##               0.4135168               0.5581414               0.9497578 
+##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+##               0.7956060               0.7091274               0.3126237 
+##         Plasma membrane              Proteasome 
+##               0.7100209               0.4188587
+\end{verbatim}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{getPredictions}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{,} \hlkwc{t} \hlstd{= ts)}
+\end{alltt}
+\begin{verbatim}
+## ans
+##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
+##                      85                     172                      89 
+##                 Cytosol   Endoplasmic reticulum                Endosome 
+##                     294                     476                     103 
+##    Extracellular matrix                Lysosome           Mitochondrion 
+##                      27                     120                     522 
+##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
+##                     230                     343                      39 
+##         Plasma membrane              Proteasome                 unknown 
+##                     316                     158                    2058
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<orgts>>=
-(ts <- setNames(c(0.612, 0.701, 0.81, 0.92), c("PM", "Mito", "Golgi", "ER")))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{(ts} \hlkwb{<-} \hlkwd{setNames}\hlstd{(}\hlkwd{c}\hlstd{(}\hlnum{0.612}\hlstd{,} \hlnum{0.701}\hlstd{,} \hlnum{0.81}\hlstd{,} \hlnum{0.92}\hlstd{),} \hlkwd{c}\hlstd{(}\hlstr{"PM"}\hlstd{,} \hlstr{"Mito"}\hlstd{,} \hlstr{"Golgi"}\hlstd{,} \hlstr{"ER"}\hlstd{)))}
+\end{alltt}
+\begin{verbatim}
+##    PM  Mito Golgi    ER 
+## 0.612 0.701 0.810 0.920
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<resetpar, echo=FALSE>>=
-invisible(dev.off())
-@
-
 \begin{figure}[!ht]
     \centering
-<<plotres, out.width=".7\\textwidth">>=
-plot2D(hl, fcol = "svm.pred")
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm.pred"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.7\textwidth]{figure/plotres-1}
+
+\end{knitrout}
   \caption{Results of the localisation preductions after thresholding.}
@@
 
-<<classifyApp, eval=FALSE>>=
-library("pRolocGUI")
-pRolocVis(hl, app = "classify", fcol = "svm", scol = "svm.scores", mcol = "markers")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"pRolocGUI"}\hlstd{)}
+\hlkwd{pRolocVis}\hlstd{(hl,} \hlkwc{app} \hlstd{=} \hlstr{"classify"}\hlstd{,} \hlkwc{fcol} \hlstd{=} \hlstr{"svm"}\hlstd{,} \hlkwc{scol} \hlstd{=} \hlstr{"svm.scores"}\hlstd{,} \hlkwc{mcol} \hlstd{=} \hlstr{"markers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<setap, eval=FALSE>>=
-par <- setAnnotationParams(inputs = c("Mus musculus", "UniProt/Swissprot"))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{par} \hlkwb{<-} \hlkwd{setAnnotationParams}\hlstd{(}\hlkwc{inputs} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"Mus musculus"}\hlstd{,} \hlstr{"UniProt/Swissprot"}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<makego, eval=FALSE>>=
-gocc <- makeGoSet(hl, params = par, namespace = "cellular_component")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{gocc} \hlkwb{<-} \hlkwd{makeGoSet}\hlstd{(hl,} \hlkwc{params} \hlstd{= par,} \hlkwc{namespace} \hlstd{=} \hlstr{"cellular_component"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<filtergo, eval=FALSE>>=
-gocc <- filterBinMSnSet(hl)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{gocc} \hlkwb{<-} \hlkwd{filterBinMSnSet}\hlstd{(hl)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<loadgocc, include=FALSE>>=
-gofile <- dir(extdatadir, full.names = TRUE,
-              pattern = "bpw-gocc.rds")
-gocc <- readRDS(gofile)
-@
-
-<<setuptlopt>>=
-## create new markers column for tl markers
-fData(hl)$tlmarkers <- fData(hl)$markers
-fData(gocc)$tlmarkers <- fData(gocc)$markers
-
-## Remove 4 classes
-torm  <- c("40S Ribosome", "60S Ribosome",
-           "Nucleus - Chromatin", 
-           "Nucleus - Non-chromatin")
-for (i in seq(torm)) {
-  hl <- fDataToUnknown(hl, from = torm[i], fcol = "tlmarkers")
-  gocc <- fDataToUnknown(gocc, from = torm[i], fcol = "tlmarkers")
-}
-getMarkerClasses(hl, fcol = "tlmarkers")
-getMarkerClasses(gocc, fcol = "tlmarkers")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## create new markers column for tl markers}
+\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{tlmarkers} \hlkwb{<-} \hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{markers}
+\hlkwd{fData}\hlstd{(gocc)}\hlopt{$}\hlstd{tlmarkers} \hlkwb{<-} \hlkwd{fData}\hlstd{(gocc)}\hlopt{$}\hlstd{markers}
+
+\hlcom{## Remove 4 classes}
+\hlstd{torm}  \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"40S Ribosome"}\hlstd{,} \hlstr{"60S Ribosome"}\hlstd{,}
+           \hlstr{"Nucleus - Chromatin"}\hlstd{,}
+           \hlstr{"Nucleus - Non-chromatin"}\hlstd{)}
+\hlkwa{for} \hlstd{(i} \hlkwa{in} \hlkwd{seq}\hlstd{(torm)) \{}
+  \hlstd{hl} \hlkwb{<-} \hlkwd{fDataToUnknown}\hlstd{(hl,} \hlkwc{from} \hlstd{= torm[i],} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+  \hlstd{gocc} \hlkwb{<-} \hlkwd{fDataToUnknown}\hlstd{(gocc,} \hlkwc{from} \hlstd{= torm[i],} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\hlstd{\}}
+\hlkwd{getMarkerClasses}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+##  [1] "Actin cytoskeleton"    "Cytosol"              
+##  [3] "Endoplasmic reticulum" "Endosome"             
+##  [5] "Extracellular matrix"  "Lysosome"             
+##  [7] "Mitochondrion"         "Peroxisome"           
+##  [9] "Plasma membrane"       "Proteasome"
+\end{verbatim}
+\begin{alltt}
+\hlkwd{getMarkerClasses}\hlstd{(gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\end{alltt}
+\begin{verbatim}
+##  [1] "Actin cytoskeleton"    "Cytosol"              
+##  [3] "Endoplasmic reticulum" "Endosome"             
+##  [5] "Extracellular matrix"  "Lysosome"             
+##  [7] "Mitochondrion"         "Peroxisome"           
+##  [9] "Plasma membrane"       "Proteasome"
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<knnoptfortl, eval=FALSE>>=
-## get best k's
-kpopt <- knnOptimisation(hl, fcol = "tlmarkers")
-kaopt <- knnOptimisation(gocc, fcol = "tlmarkers")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## get best k's}
+\hlstd{kpopt} \hlkwb{<-} \hlkwd{knnOptimisation}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\hlstd{kaopt} \hlkwb{<-} \hlkwd{knnOptimisation}\hlstd{(gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<runtlopts, eval=FALSE>>=
-## Set appropriate parallelisation backend and  
-## number of workers for the tl
-par <- SnowParam(255L, type = "MPI")
-
-## Now peform tl optimisation 
-tlopt <- knntlOptimisation(hl, gocc, fcol = "tlmarkers",
-                           length.out = 4, times = 50,
-                           xval = 5, k = c(3, 3),
-                           BPPARAM = par)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Set appropriate parallelisation backend and  }
+\hlcom{## number of workers for the tl}
+\hlstd{par} \hlkwb{<-} \hlkwd{SnowParam}\hlstd{(}\hlnum{255L}\hlstd{,} \hlkwc{type} \hlstd{=} \hlstr{"MPI"}\hlstd{)}
+
+\hlcom{## Now peform tl optimisation }
+\hlstd{tlopt} \hlkwb{<-} \hlkwd{knntlOptimisation}\hlstd{(hl, gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{,}
+                           \hlkwc{length.out} \hlstd{=} \hlnum{4}\hlstd{,} \hlkwc{times} \hlstd{=} \hlnum{50}\hlstd{,}
+                           \hlkwc{xval} \hlstd{=} \hlnum{5}\hlstd{,} \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{),}
+                           \hlkwc{BPPARAM} \hlstd{= par)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<loadtlres, echo=FALSE>>=
-tlfile <- dir(extdatadir, full.names = TRUE,
-              pattern = "bpw-tlopt.rds")
-tlopt <- readRDS(tlfile)
-colnames(tlopt@results)[4] <- "Endoplasmic reticulum"
-setStockcol(NULL)
-@
-
 \begin{figure}[!ht]
   \centering
-<<plottl, out.width=".6\\textwidth">>=
-plot(tlopt)
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{plot}\hlstd{(tlopt)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.6\textwidth]{figure/plottl-1}
+
+\end{knitrout}
 \caption{Visualisation of the transfer learning parameter optimisation
@@
 
-<<tlclassify, cache = cache>>=
-## best parameters for the 10 classes
-(bestpar <- getParams(tlopt))
-
-## add weights for classes not included in the optimisation
-otherweights <- rep(1, 4)
-names(otherweights) <- c("40S Ribosome", "60S Ribosome", 
-                         "Nucleus - Chromatin", 
-                         "Nucleus - Non-chromatin")
-(bestpar <- c(bestpar, otherweights))
-
-## re-order classes 
-bestpar <- bestpar[getMarkerClasses(hl)]
-
-## Do the classification
-hl <- knntlClassification(hl, gocc, bestTheta = bestpar, k = c(3, 3))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## best parameters for the 10 classes}
+\hlstd{(bestpar} \hlkwb{<-} \hlkwd{getParams}\hlstd{(tlopt))}
+\end{alltt}
+\begin{verbatim}
+##    Actin cytoskeleton               Cytosol Endoplasmic reticulum 
+##             1.0000000             1.0000000             0.6666667 
+##              Endosome  Extracellular matrix              Lysosome 
+##             1.0000000             1.0000000             0.3333333 
+##         Mitochondrion            Peroxisome       Plasma membrane 
+##             1.0000000             1.0000000             0.6666667 
+##            Proteasome 
+##             0.3333333
+\end{verbatim}
+\begin{alltt}
+\hlcom{## add weights for classes not included in the optimisation}
+\hlstd{otherweights} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{4}\hlstd{)}
+\hlkwd{names}\hlstd{(otherweights)} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlstr{"40S Ribosome"}\hlstd{,} \hlstr{"60S Ribosome"}\hlstd{,}
+                         \hlstr{"Nucleus - Chromatin"}\hlstd{,}
+                         \hlstr{"Nucleus - Non-chromatin"}\hlstd{)}
+\hlstd{(bestpar} \hlkwb{<-} \hlkwd{c}\hlstd{(bestpar, otherweights))}
+\end{alltt}
+\begin{verbatim}
+##      Actin cytoskeleton                 Cytosol   Endoplasmic reticulum 
+##               1.0000000               1.0000000               0.6666667 
+##                Endosome    Extracellular matrix                Lysosome 
+##               1.0000000               1.0000000               0.3333333 
+##           Mitochondrion              Peroxisome         Plasma membrane 
+##               1.0000000               1.0000000               0.6666667 
+##              Proteasome            40S Ribosome            60S Ribosome 
+##               0.3333333               1.0000000               1.0000000 
+##     Nucleus - Chromatin Nucleus - Non-chromatin 
+##               1.0000000               1.0000000
+\end{verbatim}
+\begin{alltt}
+\hlcom{## re-order classes }
+\hlstd{bestpar} \hlkwb{<-} \hlstd{bestpar[}\hlkwd{getMarkerClasses}\hlstd{(hl)]}
+
+\hlcom{## Do the classification}
+\hlstd{hl} \hlkwb{<-} \hlkwd{knntlClassification}\hlstd{(hl, gocc,} \hlkwc{bestTheta} \hlstd{= bestpar,} \hlkwc{k} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,} \hlnum{3}\hlstd{))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
   \centering
-<<hclust, out.width=".5\\textwidth", fig.asp = 1.5, out.extra='angle=-90'>>=
-par(mar = c(15, 4, 1, 0)) ## set figure margin
-mrkHClust(hl)
-@  
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mar} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{15}\hlstd{,} \hlnum{4}\hlstd{,} \hlnum{1}\hlstd{,} \hlnum{0}\hlstd{))} \hlcom{## set figure margin}
+\hlkwd{mrkHClust}\hlstd{(hl)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.5\textwidth,angle=-90]{figure/hclust-1}
+
+\end{knitrout}
   \caption{Hierarchical clustering of the average marker profiles summarising the relation between organelles profiles.}
@@
 
-<<writeData, eval = FALSE>>=
-write.exprs(hl, file = "hl.csv", sep = ",", fcol = fvarLabels(hl))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{file} \hlstd{=} \hlstr{"hl.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{,} \hlkwc{fcol} \hlstd{=} \hlkwd{fvarLabels}\hlstd{(hl))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<saveData, eval = FALSE>>=
-## To save the parameters as an R object
-save(params, file = "svmparams.rda")
-
-## To re-load after saving
-load(file = "svmparams.rda")
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## To save the parameters as an R object}
+\hlkwd{save}\hlstd{(params,} \hlkwc{file} \hlstd{=} \hlstr{"svmparams.rda"}\hlstd{)}
+
+\hlcom{## To re-load after saving}
+\hlkwd{load}\hlstd{(}\hlkwc{file} \hlstd{=} \hlstr{"svmparams.rda"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<si>>=
-sessionInfo()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{sessionInfo}\hlstd{()}
+\end{alltt}
+\begin{verbatim}
+## R Under development (unstable) (2018-03-21 r74432)
+## Platform: x86_64-apple-darwin15.6.0 (64-bit)
+## Running under: macOS High Sierra 10.13.3
+## 
+## Matrix products: default
+## BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
+## LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib
+## 
+## locale:
+## [1] en_GB.UTF-8/en_GB.UTF-8/en_GB.UTF-8/C/en_GB.UTF-8/en_GB.UTF-8
+## 
+## attached base packages:
+## [1] stats4    parallel  stats     graphics  grDevices utils     datasets 
+## [8] methods   base     
+## 
+## other attached packages:
+##  [1] pRolocdata_1.17.2    gridExtra_2.3        xtable_1.8-2        
+##  [4] BiocStyle_2.7.8      knitr_1.20           pRoloc_1.19.3       
+##  [7] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     
+## [10] XML_3.98-1.10        AnnotationDbi_1.41.4 IRanges_2.13.28     
+## [13] S4Vectors_0.17.37    MSnbase_2.5.10       ProtGenerics_1.11.0 
+## [16] BiocParallel_1.13.1  mzR_2.13.5           Rcpp_0.12.16        
+## [19] Biobase_2.39.2       BiocGenerics_0.25.3 
+## 
+## loaded via a namespace (and not attached):
+##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
+##   [4] lazyeval_0.2.1        splines_3.5.0         ggvis_0.4.3          
+##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
+##  [10] foreach_1.4.4         BiocInstaller_1.29.4  htmltools_0.3.6      
+##  [13] viridis_0.5.0         gdata_2.18.0          magrittr_1.5         
+##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
+##  [19] limma_3.35.13         recipes_0.1.2         gower_0.1.2          
+##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
+##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.0           
+##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
+##  [31] genefilter_1.61.1     bindr_0.1.1           impute_1.53.0        
+##  [34] survival_2.41-3       iterators_1.0.9       glue_1.2.0           
+##  [37] DRR_0.0.3             gtable_0.2.0          ipred_0.9-6          
+##  [40] zlibbioc_1.25.0       kernlab_0.9-25        ddalpha_1.3.1.1      
+##  [43] prabclus_2.2-6        DEoptimR_1.0-8        scales_0.5.0         
+##  [46] vsn_3.47.1            mvtnorm_1.0-7         DBI_0.8              
+##  [49] viridisLite_0.3.0     progress_1.1.2        foreign_0.8-70       
+##  [52] bit_1.1-12            proxy_0.4-21          mclust_5.4           
+##  [55] preprocessCore_1.41.0 lava_1.6              prodlim_1.6.1        
+##  [58] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
+##  [61] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
+##  [64] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
+##  [67] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
+##  [70] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
+##  [73] munsell_0.4.3         mlbench_2.1-1         tools_3.5.0          
+##  [76] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
+##  [79] evaluate_0.10.1       stringr_1.3.0         yaml_2.1.18          
+##  [82] mzID_1.17.0           ModelMetrics_1.1.0    bit64_0.9-7          
+##  [85] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
+##  [88] dendextend_1.7.0      bindrcpp_0.2          nlme_3.1-136         
+##  [91] whisker_0.3-2         mime_0.5              RcppRoll_0.2.2       
+##  [94] biomaRt_2.35.12       compiler_3.5.0        e1071_1.6-8          
+##  [97] affyio_1.49.1         tibble_1.4.2          stringi_1.1.7        
+## [100] highr_0.6             lattice_0.20-35       trimcluster_0.1-2    
+## [103] Matrix_1.2-12         psych_1.7.8           gbm_2.1.3            
+## [106] pillar_1.2.1          MALDIquant_1.17       bitops_1.0-6         
+## [109] httpuv_1.3.6.2        R6_2.2.2              pcaMethods_1.71.0    
+## [112] affy_1.57.0           hwriter_1.3.2         codetools_0.2-15     
+## [115] MASS_7.3-49           gtools_3.5.0          assertthat_0.2.0     
+## [118] CVST_0.2-1            rprojroot_1.3-2       withr_2.1.2          
+## [121] mnormt_1.5-5          diptest_0.75-7        grid_3.5.0           
+## [124] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
+## [127] class_7.3-14          rmarkdown_1.9         shiny_1.0.5          
+## [130] lubridate_1.7.3       base64enc_0.1-3
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
@@
 
-<<update, eval=FALSE>>=
-library("BiocInstaller")
-biocLite()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{library}\hlstd{(}\hlstr{"BiocInstaller"}\hlstd{)}
+\hlkwd{biocLite}\hlstd{()}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
 

### 2018-04-04 bioc-workflow_march2017.Rnw
fixed macro conflicts in tex

@@
 suppressPackageStartupMessages(library("xtable"))
-suppressPackageStartupMessages(library("gridExtra"))
 suppressPackageStartupMessages(library("MSnbase"))
@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into \sout{a}\hl{the} same intensity interval (typically 0 and 1) by, for
+profiles into \st{a}\hl{the} same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
 
-\sout{Note: We do not evaluate this code chunk in this document as the
+\st{Note: We do not evaluate this code chunk in this document as the
 algorithm is computationally intensive and best parallelised over

### 2018-04-04 bioc-workflow_march2017.tex
fixed macro conflicts in tex

@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into \st{a} \hl{the} same intensity interval (typically 0 and 1) by, for
+profiles into \st{a}\hl{the} same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
-## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
-## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
-## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
+## Combined [5489,10] and [6268,10] MSnSets Wed Apr  4 13:30:04 2018 
+## Subset [6725,20][5032,20] Wed Apr  4 13:30:17 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:30:17 2018 
+## Dropped featureData's levels Wed Apr  4 13:30:17 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:33:06 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:33:06 2018 
+## Dropped featureData's levels Wed Apr  4 13:33:06 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:42:10 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:42:10 2018 
+## Dropped featureData's levels Wed Apr  4 13:42:10 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:45:12 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:45:12 2018 
+## Dropped featureData's levels Wed Apr  4 13:45:12 2018 
 ##  MSnbase version: 2.5.9
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
-## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
-## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
-## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
-## Subset [5032,20][5032,20] Tue Apr  3 17:46:47 2018 
-## Removed features with more than 0 NAs: Tue Apr  3 17:46:47 2018 
-## Dropped featureData's levels Tue Apr  3 17:46:48 2018 
+## Combined [5489,10] and [6268,10] MSnSets Wed Apr  4 13:30:04 2018 
+## Subset [6725,20][5032,20] Wed Apr  4 13:30:17 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:30:17 2018 
+## Dropped featureData's levels Wed Apr  4 13:30:17 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:33:06 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:33:06 2018 
+## Dropped featureData's levels Wed Apr  4 13:33:06 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:42:10 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:42:10 2018 
+## Dropped featureData's levels Wed Apr  4 13:42:10 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:45:12 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:45:12 2018 
+## Dropped featureData's levels Wed Apr  4 13:45:12 2018 
+## Subset [5032,20][5032,20] Wed Apr  4 13:47:14 2018 
+## Removed features with more than 0 NAs: Wed Apr  4 13:47:14 2018 
+## Dropped featureData's levels Wed Apr  4 13:47:14 2018 
 ##  MSnbase version: 2.5.9
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr  3 17:46:49 2018 
+##  Created on Wed Apr  4 13:47:16 2018 
 ##  Description:
@@
 % latex table generated in R 3.5.0 by xtable 1.8-2 package
-% Tue Apr  3 17:46:49 2018
+% Wed Apr  4 13:47:16 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr  3 17:46:49 2018 
-## Added markers from  'pdres' marker vector. Tue Apr  3 17:46:50 2018 
+## Added markers from  'mrk' marker vector. Wed Apr  4 13:47:15 2018 
+## Added markers from  'pdres' marker vector. Wed Apr  4 13:47:17 2018 
 ##  MSnbase version: 2.5.10
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4333396               0.3051261               0.3870872 
+##               0.4369121               0.3042575               0.3807524 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6736435               0.6011995               0.4213771 
+##               0.6850502               0.6093775               0.4324733 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4135168               0.5581414               0.9497578 
+##               0.4261633               0.5908828               0.9491150 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7956060               0.7091274               0.3126237 
+##               0.7944096               0.7083224               0.3140141 
 ##         Plasma membrane              Proteasome 
-##               0.7100209               0.4188587
+##               0.7126035               0.4150477
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4333396               0.3051261               0.3870872 
+##               0.4369121               0.3042575               0.3807524 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6736435               0.6011995               0.4213771 
+##               0.6850502               0.6093775               0.4324733 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4135168               0.5581414               0.9497578 
+##               0.4261633               0.5908828               0.9491150 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7956060               0.7091274               0.3126237 
+##               0.7944096               0.7083224               0.3140141 
 ##         Plasma membrane              Proteasome 
-##               0.7100209               0.4188587
+##               0.7126035               0.4150477
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      89 
+##                      84                     173                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     294                     476                     103 
+##                     297                     468                     102 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     120                     522 
+##                      28                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      39 
+##                     229                     343                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     316                     158                    2058
+##                     322                     159                    2057
 \end{verbatim}
@@
 ## other attached packages:
-##  [1] pRolocdata_1.17.2    gridExtra_2.3        xtable_1.8-2        
-##  [4] BiocStyle_2.7.8      knitr_1.20           pRoloc_1.19.3       
-##  [7] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     
-## [10] XML_3.98-1.10        AnnotationDbi_1.41.4 IRanges_2.13.28     
-## [13] S4Vectors_0.17.37    MSnbase_2.5.10       ProtGenerics_1.11.0 
-## [16] BiocParallel_1.13.1  mzR_2.13.5           Rcpp_0.12.16        
-## [19] Biobase_2.39.2       BiocGenerics_0.25.3 
+##  [1] BiocStyle_2.7.8      gridExtra_2.3        shiny_1.0.5         
+##  [4] pRolocGUI_1.13.1     rgl_0.99.9           pRolocdata_1.17.2   
+##  [7] pRoloc_1.19.3        MLInterfaces_1.59.1  cluster_2.0.6       
+## [10] annotate_1.57.2      XML_3.98-1.10        AnnotationDbi_1.41.4
+## [13] IRanges_2.13.28      S4Vectors_0.17.37    MSnbase_2.5.10      
+## [16] ProtGenerics_1.11.0  BiocParallel_1.13.1  mzR_2.13.5          
+## [19] Rcpp_0.12.16         Biobase_2.39.2       BiocGenerics_0.25.3 
+## [22] xtable_1.8-2         knitr_1.20          
 ## 
@@
 ##  [10] foreach_1.4.4         BiocInstaller_1.29.4  htmltools_0.3.6      
-##  [13] viridis_0.5.0         gdata_2.18.0          magrittr_1.5         
-##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
-##  [19] limma_3.35.13         recipes_0.1.2         gower_0.1.2          
-##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
-##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.0           
-##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
-##  [31] genefilter_1.61.1     bindr_0.1.1           impute_1.53.0        
-##  [34] survival_2.41-3       iterators_1.0.9       glue_1.2.0           
-##  [37] DRR_0.0.3             gtable_0.2.0          ipred_0.9-6          
-##  [40] zlibbioc_1.25.0       kernlab_0.9-25        ddalpha_1.3.1.1      
-##  [43] prabclus_2.2-6        DEoptimR_1.0-8        scales_0.5.0         
-##  [46] vsn_3.47.1            mvtnorm_1.0-7         DBI_0.8              
-##  [49] viridisLite_0.3.0     progress_1.1.2        foreign_0.8-70       
-##  [52] bit_1.1-12            proxy_0.4-21          mclust_5.4           
-##  [55] preprocessCore_1.41.0 lava_1.6              prodlim_1.6.1        
-##  [58] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
-##  [61] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
-##  [64] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
-##  [67] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
-##  [70] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
-##  [73] munsell_0.4.3         mlbench_2.1-1         tools_3.5.0          
-##  [76] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
-##  [79] evaluate_0.10.1       stringr_1.3.0         yaml_2.1.18          
-##  [82] mzID_1.17.0           ModelMetrics_1.1.0    bit64_0.9-7          
-##  [85] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
-##  [88] dendextend_1.7.0      bindrcpp_0.2          nlme_3.1-136         
-##  [91] whisker_0.3-2         mime_0.5              RcppRoll_0.2.2       
-##  [94] biomaRt_2.35.12       compiler_3.5.0        e1071_1.6-8          
-##  [97] affyio_1.49.1         tibble_1.4.2          stringi_1.1.7        
-## [100] highr_0.6             lattice_0.20-35       trimcluster_0.1-2    
-## [103] Matrix_1.2-12         psych_1.7.8           gbm_2.1.3            
-## [106] pillar_1.2.1          MALDIquant_1.17       bitops_1.0-6         
-## [109] httpuv_1.3.6.2        R6_2.2.2              pcaMethods_1.71.0    
-## [112] affy_1.57.0           hwriter_1.3.2         codetools_0.2-15     
-## [115] MASS_7.3-49           gtools_3.5.0          assertthat_0.2.0     
-## [118] CVST_0.2-1            rprojroot_1.3-2       withr_2.1.2          
-## [121] mnormt_1.5-5          diptest_0.75-7        grid_3.5.0           
-## [124] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
-## [127] class_7.3-14          rmarkdown_1.9         shiny_1.0.5          
-## [130] lubridate_1.7.3       base64enc_0.1-3
+##  [13] rsconnect_0.8.8       viridis_0.5.0         gdata_2.18.0         
+##  [16] magrittr_1.5          memoise_1.1.0         doParallel_1.0.11    
+##  [19] sfsmisc_1.1-2         limma_3.35.13         recipes_0.1.2        
+##  [22] gower_0.1.2           rda_1.0.2-2           dimRed_0.1.0         
+##  [25] lpSolve_5.6.13        prettyunits_1.0.2     colorspace_1.3-2     
+##  [28] blob_1.1.0            dplyr_0.7.4           jsonlite_1.5         
+##  [31] RCurl_1.95-4.10       hexbin_1.27.2         genefilter_1.61.1    
+##  [34] bindr_0.1.1           impute_1.53.0         survival_2.41-3      
+##  [37] iterators_1.0.9       glue_1.2.0            DRR_0.0.3            
+##  [40] gtable_0.2.0          ipred_0.9-6           zlibbioc_1.25.0      
+##  [43] kernlab_0.9-25        ddalpha_1.3.1.1       prabclus_2.2-6       
+##  [46] DEoptimR_1.0-8        scales_0.5.0          vsn_3.47.1           
+##  [49] mvtnorm_1.0-7         DBI_0.8               viridisLite_0.3.0    
+##  [52] progress_1.1.2        foreign_0.8-70        bit_1.1-12           
+##  [55] proxy_0.4-21          mclust_5.4            preprocessCore_1.41.0
+##  [58] DT_0.4                lava_1.6              prodlim_1.6.1        
+##  [61] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
+##  [64] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
+##  [67] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
+##  [70] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
+##  [73] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
+##  [76] munsell_0.4.3         mlbench_2.1-1         tools_3.5.0          
+##  [79] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
+##  [82] evaluate_0.10.1       stringr_1.3.0         mzID_1.17.0          
+##  [85] yaml_2.1.18           ModelMetrics_1.1.0    bit64_0.9-7          
+##  [88] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
+##  [91] dendextend_1.7.0      bindrcpp_0.2          nlme_3.1-136         
+##  [94] whisker_0.3-2         mime_0.5              RcppRoll_0.2.2       
+##  [97] biomaRt_2.35.12       compiler_3.5.0        e1071_1.6-8          
+## [100] affyio_1.49.1         tibble_1.4.2          stringi_1.1.7        
+## [103] highr_0.6             lattice_0.20-35       trimcluster_0.1-2    
+## [106] Matrix_1.2-12         psych_1.7.8           gbm_2.1.3            
+## [109] pillar_1.2.1          MALDIquant_1.17       bitops_1.0-6         
+## [112] httpuv_1.3.6.2        R6_2.2.2              pcaMethods_1.71.0    
+## [115] affy_1.57.0           hwriter_1.3.2         codetools_0.2-15     
+## [118] MASS_7.3-49           gtools_3.5.0          assertthat_0.2.0     
+## [121] CVST_0.2-1            rprojroot_1.3-2       withr_2.1.2          
+## [124] mnormt_1.5-5          diptest_0.75-7        grid_3.5.0           
+## [127] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
+## [130] class_7.3-14          rmarkdown_1.9         lubridate_1.7.3      
+## [133] base64enc_0.1-3
 \end{verbatim}

### 2018-04-10 bioc-workflow_march2017.Rnw
Almost all comments now addressed

@@
 While we focus here on a LOPIT-type dataset, these analyses are
-relevant for any quantitative spatial proteomics data, irrespective of
+\hl{relevent} for any quantitative spatial proteomics data, irrespective of
 the fractionation or quantitation (i.e. labelled or label-free)
@@
 <<getFilename, cache = cache>>=
-library("MSnbase")
 extdatadir <- system.file("extdata", package = "pRolocdata")
@@
 
-There are two aspects related to data normalisation that are relevant
+There are two aspects related to data normalisation that are \hl{relevent}
 to spatial proteomics data processing. The first one focuses on
@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into \st{a}\hl{the} same intensity interval (typically 0 and 1) by, for
+profiles into \hl{the} same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
   red dots represent the mean relative intensity for each channel. }
-  \label{fig:qcbx}
+  \label{fig:q  cbx}
 \end{figure}
+
+\pagebreak
 
@@
 
-Having added a the mouse markers to our \texttt{fData} from the
-\texttt{pRolocmarkers}, we can now visualise these annotations along
+Having added the mouse markers to our \texttt{fData} from the
+\texttt{pRolocmarkers}, we can now visualise these annotations on
 the PCA plot using the \texttt{plot2D} function and then use the
 \texttt{addLegend} function to map the marker classes to the
-pre-defined colours. \hl{As mentioned previously, PCA transforms the
+pre-defined colours. \hl{As previously mentioned, PCA transforms the
 original high dimensional data into a set of linearly uncorrelated
-principal components such that the first accounts for as much
+principal components (PCs) such that the first accounts for as much
 variability in the data as possible and each succeeding component in
@@
 be orthogonal to the preceding components. We saw in the previous
-section how this is useful for quality control and checking
+section how visualisation of the PCs is useful for quality control and checking
 organelle seperation. Adding marker definiton allows one to quickly
@@
 calls to the \texttt{plot2D} function, we have omitted the
-\texttt{fcol} argument and use of the default \texttt{"markers"}
-feature variable to annotate the plot. We choose to display PCs 1
+\texttt{fcol} argument and so by default the feature variable named
+\texttt{"markers"} is used to annotate the plot. We choose to display PCs 1
 and 7 to illustrate that while upper principal components explain
@@
 apparent overlap in the visualisation of the two first PCs (Figure
-\ref{fig:plotmarkers}).  \hl{This is further highlighted if we plot the
+\ref{fig:plotmarkers}).
+
+\begin{figure}[!ht]
+  \centering
+<<plotmarkers, out.width="\\textwidth", fig.width=12, fig.asp=1/2>>=
+par(mfrow = c(1, 2))
+plot2D(hl, main = "pRolocmarkers for mouse")
+addLegend(hl, cex = .6)
+plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
+@  
+\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
+  with \texttt{plot2D}.}
+  \label{fig:plotmarkers}
+\end{figure}
+
+\hl{This is further highlighted if we plot the
 profiles of these two clusters using the \texttt{plotDist} function}
-(Fig \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
+(Figure \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
 another useful visualisation that relies on marker annotation. It
 allows one to represent the protein profiles occupancy along the
-gradient} \st{using the \texttt{plotDist} function}. \hl{While the PCA
+gradient. While the PCA
 plot enables efficient visualisation of the complete dataset and
@@
 figure} \ref{fig:plotmarkers}.
-
-\begin{figure}[!ht]
-  \centering
-<<plotmarkers, out.width="\\textwidth", fig.width=12, fig.asp=1/2>>=
-par(mfrow = c(1, 2))
-plot2D(hl, main = "pRolocmarkers for mouse")
-addLegend(hl, cex = .6)
-plot2D(hl, dims = c(1, 7), main = "Marker resolution along PC 1 and 7")
-@  
-\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
-  with \texttt{plot2D}.}
-  \label{fig:plotmarkers}
-\end{figure}
 
@@
 we \hl{would call the following code:}
-\st{would create a first create a \texttt{FeaturesOfInterest} object, and
-subsequently highlight their location on a PCA plot with the
-\texttt{highlightOnPlot} function.}
 
@@
 Direct comparisons of individual channels in replicated experiments
-does not provide an adequate, goal-driven assessment of different
+do\st{es} not provide an adequate, goal-driven assessment of different
 experiments. Indeed, due to the nature of the experiment and gradient
@@
 
-The more relevant comparison unit is not a single channel, but rather
+The more \hl{relevent} comparison unit is not a single channel, but rather
 the complete protein occupancy profiles, which are best visualised 
@@
 
-\st{Note: We do not evaluate this code chunk in this document as the
-algorithm is computationally intensive and best parallelised over
+\hl{The above analysis is comptationally intensive and best parallelised over
 multiple workers. This phenoDisco analysis took ~24 hours to complete
-when parallelised over 40 workers.}
+when parallelised over 40 workers. As such, to avoid executing this
+time consuming analysis users can access the above results which are
+pre-computed and stored along with the \texttt{pRolocdata} package. 
+Please see the appendix to load these results.}
 
@@
 \texttt{times}, \texttt{G} (among other function arguments, see
-\texttt{?phenoDisco}) will heavily influence (increase) the total
-time taken to run the algorithm.  \texttt{phenoDisco} supports
+\texttt{?phenoDisco}) will heavily influence (increase) the total time
+taken to run the algorithm.  \texttt{phenoDisco} supports
 parallelisation and we strongly suggest you make use of a parallel
 processing to run these analyses.
-
-\hl{Please note, in this document we can not evaluate the call to 
-`phenoDisco` in the code chunk above as the algorithm is computationally 
-intensive and best parallelised over multiple workers. This phenoDisco 
-analysis took ~24 hours to complete when parallelised over 40 workers. 
-The ouput of running the \texttt{phenoDisco} algorithm is an
-\texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. We have made the results
-readily available for users to interrogate and get an idea of the type
-of output one may expect from using this function.} The results can be
-displayed by using the \texttt{getMarkers} function. We see that 5 new
-phenotype data clusters were found.
 
@@
 @
+
+The ouput of running the \texttt{phenoDisco} algorithm is an
+\texttt{MSnSet} containing the new data clusters, appended to the
+\texttt{featureData} under the name \texttt{pd}. The results can be
+displayed by using the \texttt{getMarkers} function. We see that 5 new
+phenotype data clusters were found.
 
@@
 \href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\Biocpkg{pRoloc}
-  machine learning techniques vignette}.  We find the general tendancy
+  machine learning techniques vignette}.  We find the general \hl{tendency}
 to be that it is not the choice of classifier, but the improper
@@
 
-<<loadpdRes, echo=FALSE>>=
+<<getsvmweight>>=
+w <- table(getMarkers(hl, verbose = TRUE))
+w <- 1/w[names(w) != "unknown"]
+@
+
+\hl{In the code chunk below we then pass these weights 
+to the \texttt{svmOptimisation} function. Once again, we provide
+the optimisation results for users to load directly if they wish to save
+computational time, please see the appendix to load these results.}
+
+<<loadpdRes, echo=TRUE>>=
 svmf <- dir(extdatadir, full.names = TRUE,
@@
 params <- readRDS(svmf)
-@
-
-<<getsvmweight>>=
-w <- table(getMarkers(hl, verbose = TRUE))
-w <- 1/w[names(w) != "unknown"]
 @
@@
 argument to change the size of each point on the plot to be inversely
-proportional to the SVM score. This gives an initial overview of the
+proportional to the SVM score (Figure \ref{fig:plotSVM}). This gives an initial overview of the
 high scoring localisations from the SVM predictions.
@@
 \end{figure}
+\pagebreak
 
@@
 
+\pagebreak
+
 There are many ways to set thresholds and the choice of method will
@@
 information is not available, one crude method is to set a threshold
-per organelle by extracting the median or 3rd quantile score per
+per organelle by extracting the median or 3rd \hl{quartile} score per
 organelle.  For example, in the code chunk below, we use the
@@
 evidence codes as desired, see \texttt{?makeGoSet} for more details. 
+\hl{(We also provide the pre-computed \texttt{gocc} object for users
+if they wish to load directly, please see the appendix).}
 
@@
 auxiliary data (so 0 is given to the primary source). If we conduct a
-parameter seach and test weights $\theta = {0, 1/3, 2/3, 1}$ for each
+parameter \hl{search} and test weights $\theta = {0, 1/3, 2/3, 1}$ for each
 class, and if we have, for example 10 subcellular niches, this will
@@
 
-In the example below we remove these 4 classes from the marker set,
+In the example below we \hl{first} remove these 4 classes from the marker set
+\hl{and create a new marker set called tlmarkers. Then we}
 re-run the \texttt{knnOptimisation} for each data source and then run
@@
 707 markers and 4$^{10}$ combinations of parameters takes around 76
-hours to run on the University of Cambridge HPC using 256 workers).
-
-To remove the 4 classes and create a new column of markers in the
-feature data called \texttt{tlmarkers} to use for the analysis:
+hours to run on the University of Cambridge HPC using 256 workers. 
+\hl{To save time for users, the results of the following optimisation 
+are pre-computed and provided with the \texttt{pRolocdata} package. 
+Please see the appendix for details on how
+to load these directly}.)
+
+\st{To remove the 4 classes and create a new column of markers in the
+feature data called tlmarkers to use for the analysis:}
 
@@
 
-From examining the parameter seach plots as described in section
+From examining the parameter \hl{search} plots as described in section
 \textit{Optimisation}, we find the best $k$'s for both the primary and
@@
 data. To directly serialise R objects to disk, on can use the standard
-\texttt{save} function, and later reload the object using
-\texttt{save}. For example, to save and then re-load the parameters
+\texttt{save} function, and later reload the object using \st{save}
+\texttt{load}. For example, to save and then re-load the parameters
 from the SVM optimisation,
@@
 
-\section*{Session information}
+\section*{Session information and getting help}
 
@@
 Technology Facilities Council.
+
+\subsection*{Appendix}
+
+\subsubsection{Novelty detection results}
+
+To load the pre-computed novelty detection results please execute the
+following lines of code
+
+<<loadPDres_apx, eval=FALSE>>=
+## Specifying the path to the PD results
+f0 <- dir(extdatadir, full.names = TRUE,
+          pattern = "bpw-pdres.rds")
+## Reading the PD results
+pdres <- readRDS(f0)
+## Adding the PD results to the hl MSnSet
+hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
+@
+
+\subsubsection{SVM optimisation}
+
+To load the results of the SVM optimisation
+
+<<loadpdRes_apx, eval=FALSE>>=
+## Specifying the path to the SVM optimisation results
+svmf <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-svmopt.rds") 
+## Reading the results
+params <- readRDS(svmf)
+@
+
+\subsubsection{Transfer learning results}
+
+To load the GO CC dataset
+
+<<loadgocc_apx, eval=FALSE>>=
+gofile <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-gocc.rds")
+gocc <- readRDS(gofile)
+@
+
+To load the results of the transfer learning
+
+<<loadtlres_apx, eval=FALSE>>=
+tlfile <- dir(extdatadir, full.names = TRUE,
+              pattern = "bpw-tlopt.rds")
+tlopt <- readRDS(tlfile)
+@

### 2018-04-10 bioc-workflow_march2017.tex
Almost all comments now addressed

--- 2018-04-04 bioc-workflow_march2017.tex
@@
 \begin{alltt}
-\hlkwd{library}\hlstd{(}\hlstr{"MSnbase"}\hlstd{)}
 \hlstd{extdatadir} \hlkwb{<-} \hlkwd{system.file}\hlstd{(}\hlstr{"extdata"}\hlstd{,} \hlkwc{package} \hlstd{=} \hlstr{"pRolocdata"}\hlstd{)}
@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into \st{a} \hl{the} same intensity interval (typically 0 and 1) by, for
+profiles into \hl{the} same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
-## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
-## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
-## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
+## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
+## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
+## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
+## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
+## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
+## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
+## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
+## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
+## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
+## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
 ##  MSnbase version: 2.5.9
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Mon Mar 26 16:47:45 2018 
-## Subset [6725,20][5032,20] Mon Mar 26 16:48:04 2018 
-## Removed features with more than 0 NAs: Mon Mar 26 16:48:04 2018 
-## Dropped featureData's levels Mon Mar 26 16:48:04 2018 
-## Subset [5032,20][5032,20] Tue Apr  3 17:46:47 2018 
-## Removed features with more than 0 NAs: Tue Apr  3 17:46:47 2018 
-## Dropped featureData's levels Tue Apr  3 17:46:48 2018 
+## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
+## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
+## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
+## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
+## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
+## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
+## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
+## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
+## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
+## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
+## Subset [5032,20][5032,20] Tue Apr 10 16:52:54 2018 
+## Removed features with more than 0 NAs: Tue Apr 10 16:52:54 2018 
+## Dropped featureData's levels Tue Apr 10 16:52:54 2018 
 ##  MSnbase version: 2.5.9
@@
 apparent overlap in the visualisation of the two first PCs (Figure
-\ref{fig:plotmarkers}).  \hl{This is further highlighted if we plot the
+\ref{fig:plotmarkers}).
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{main} \hlstd{=} \hlstr{"pRolocmarkers for mouse"}\hlstd{)}
+\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
+\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{7}\hlstd{),} \hlkwc{main} \hlstd{=} \hlstr{"Marker resolution along PC 1 and 7"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=\textwidth]{figure/plotmarkers-1}
+
+\end{knitrout}
+\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
+  with \texttt{plot2D}.}
+  \label{fig:plotmarkers}
+\end{figure}
+
+\hl{This is further highlighted if we plot the
 profiles of these two clusters using the \texttt{plotDist} function}
-(Fig \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
+(Figure \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
 another useful visualisation that relies on marker annotation. It
 allows one to represent the protein profiles occupancy along the
-gradient} \st{using the \texttt{plotDist} function}. \hl{While the PCA
+gradient. While the PCA
 plot enables efficient visualisation of the complete dataset and
@@
 figure} \ref{fig:plotmarkers}.
-
-\begin{figure}[!ht]
-  \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{))}
-\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{main} \hlstd{=} \hlstr{"pRolocmarkers for mouse"}\hlstd{)}
-\hlkwd{addLegend}\hlstd{(hl,} \hlkwc{cex} \hlstd{=} \hlnum{.6}\hlstd{)}
-\hlkwd{plot2D}\hlstd{(hl,} \hlkwc{dims} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{7}\hlstd{),} \hlkwc{main} \hlstd{=} \hlstr{"Marker resolution along PC 1 and 7"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=\textwidth]{figure/plotmarkers-1}
-
-\end{knitrout}
-\caption{Annotated PCA plots of the \texttt{hl} dataset, as produced
-  with \texttt{plot2D}.}
-  \label{fig:plotmarkers}
-\end{figure}
 
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr  3 17:46:49 2018 
+##  Created on Tue Apr 10 16:52:56 2018 
 ##  Description:
@@
 % latex table generated in R 3.5.0 by xtable 1.8-2 package
-% Tue Apr  3 17:46:49 2018
+% Tue Apr 10 16:52:56 2018
 \begin{table}[ht]
@@
 \texttt{times}, \texttt{G} (among other function arguments, see
-\texttt{?phenoDisco}) will heavily influence (increase) the total
-time taken to run the algorithm.  \texttt{phenoDisco} supports
+\texttt{?phenoDisco}) will heavily influence (increase) the total time
+taken to run the algorithm.  \texttt{phenoDisco} supports
 parallelisation and we strongly suggest you make use of a parallel
@@
 
-\hl{Please note, in this document we can not evaluate the call to 
-`phenoDisco` in the code chunk above as the algorithm is computationally 
-intensive and best parallelised over multiple workers. This phenoDisco 
-analysis took ~24 hours to complete when parallelised over 40 workers. 
 The ouput of running the \texttt{phenoDisco} algorithm is an
 \texttt{MSnSet} containing the new data clusters, appended to the
-\texttt{featureData} under the name \texttt{pd}. We have made the results
-readily available for users to interrogate and get an idea of the type
-of output one may expect from using this function.} The results can be
+\texttt{featureData} under the name \texttt{pd}. The results can be
 displayed by using the \texttt{getMarkers} function. We see that 5 new
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr  3 17:46:49 2018 
-## Added markers from  'pdres' marker vector. Tue Apr  3 17:46:50 2018 
+## Added markers from  'mrk' marker vector. Tue Apr 10 16:52:55 2018 
+## Added markers from  'pdres' marker vector. Tue Apr 10 16:52:56 2018 
 ##  MSnbase version: 2.5.10
@@
 
+\hl{In the code chunk below we then pass these weights 
+to the \texttt{svmOptimisation} function. Once again, we provide
+the optimisation results for users to load directly if they wish to save
+computational time, please see the appendix to load these results.}
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{svmf} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-svmopt.rds"}\hlstd{)}
+\hlstd{params} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(svmf)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
 \begin{knitrout}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4333396               0.3051261               0.3870872 
+##               0.4323379               0.3037630               0.3822328 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6736435               0.6011995               0.4213771 
+##               0.6930642               0.6014878               0.4426545 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4135168               0.5581414               0.9497578 
+##               0.4224179               0.5978957               0.9497788 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7956060               0.7091274               0.3126237 
+##               0.7950203               0.7117761               0.3147679 
 ##         Plasma membrane              Proteasome 
-##               0.7100209               0.4188587
+##               0.7042806               0.4150910
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4333396               0.3051261               0.3870872 
+##               0.4323379               0.3037630               0.3822328 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6736435               0.6011995               0.4213771 
+##               0.6930642               0.6014878               0.4426545 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4135168               0.5581414               0.9497578 
+##               0.4224179               0.5978957               0.9497788 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7956060               0.7091274               0.3126237 
+##               0.7950203               0.7117761               0.3147679 
 ##         Plasma membrane              Proteasome 
-##               0.7100209               0.4188587
+##               0.7042806               0.4150910
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     172                      89 
+##                      85                     170                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     294                     476                     103 
+##                     301                     478                      99 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     120                     522 
+##                      27                     123                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      39 
+##                     230                     341                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     316                     158                    2058
+##                     322                     154                    2057
 \end{verbatim}
@@
 ## other attached packages:
-##  [1] pRolocdata_1.17.2    gridExtra_2.3        xtable_1.8-2        
-##  [4] BiocStyle_2.7.8      knitr_1.20           pRoloc_1.19.3       
-##  [7] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     
-## [10] XML_3.98-1.10        AnnotationDbi_1.41.4 IRanges_2.13.28     
-## [13] S4Vectors_0.17.37    MSnbase_2.5.10       ProtGenerics_1.11.0 
-## [16] BiocParallel_1.13.1  mzR_2.13.5           Rcpp_0.12.16        
-## [19] Biobase_2.39.2       BiocGenerics_0.25.3 
+##  [1] gridExtra_2.3        pRolocdata_1.17.3    pRoloc_1.19.3       
+##  [4] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     
+##  [7] XML_3.98-1.10        AnnotationDbi_1.41.4 IRanges_2.13.28     
+## [10] S4Vectors_0.17.37    MSnbase_2.5.10       ProtGenerics_1.11.0 
+## [13] BiocParallel_1.13.1  mzR_2.13.5           Rcpp_0.12.16        
+## [16] Biobase_2.39.2       BiocGenerics_0.25.3  xtable_1.8-2        
+## [19] BiocStyle_2.7.8      knitr_1.20          
 ## 
@@
 ##  [76] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
-##  [79] evaluate_0.10.1       stringr_1.3.0         yaml_2.1.18          
-##  [82] mzID_1.17.0           ModelMetrics_1.1.0    bit64_0.9-7          
+##  [79] evaluate_0.10.1       stringr_1.3.0         mzID_1.17.0          
+##  [82] yaml_2.1.18           ModelMetrics_1.1.0    bit64_0.9-7          
 ##  [85] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
@@
 ## [124] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
-## [127] class_7.3-14          rmarkdown_1.9         shiny_1.0.5          
-## [130] lubridate_1.7.3       base64enc_0.1-3
+## [127] class_7.3-14          rmarkdown_1.9         lubridate_1.7.3      
+## [130] shiny_1.0.5           base64enc_0.1-3
 \end{verbatim}
@@
 Technology Facilities Council.
+
+\subsection*{Appendix}
+
+\subsubsection{Novelty detection results}
+
+To load the pre-computed novelty detection results please execute the
+following lines of code
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Specifying the path to the PD results}
+\hlstd{f0} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+          \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-pdres.rds"}\hlstd{)}
+\hlcom{## Reading the PD results}
+\hlstd{pdres} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(f0)}
+\hlcom{## Adding the PD results to the hl MSnSet}
+\hlstd{hl} \hlkwb{<-} \hlkwd{addMarkers}\hlstd{(hl, pdres,} \hlkwc{mcol} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{verbose} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\subsubsection{SVM optimisation}
+
+To load the results of the SVM optimisation
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlcom{## Specifying the path to the SVM optimisation results}
+\hlstd{svmf} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-svmopt.rds"}\hlstd{)}
+\hlcom{## Reading the results}
+\hlstd{params} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(svmf)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\subsubsection{Transfer learning results}
+
+To load the GO CC dataset
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{gofile} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-gocc.rds"}\hlstd{)}
+\hlstd{gocc} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(gofile)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+To load the results of the transfer learning
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{tlfile} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
+              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-tlopt.rds"}\hlstd{)}
+\hlstd{tlopt} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(tlfile)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}

### 2018-04-11 bioc-workflow_march2017.Rnw
all reviewers comments addressed

@@
 functionality to become available in the R session; this is done with
-the \texttt{library} function e.g.  to load the \Biocpkg{pRoloc} one
+the \texttt{library} function e.g.  to load the \Biocpkg{pRoloc} package one
 would type \texttt{library("pRoloc")} after installation.
@@
 While we focus here on a LOPIT-type dataset, these analyses are
-\hl{relevent} for any quantitative spatial proteomics data, irrespective of
-the fractionation or quantitation (i.e. labelled or label-free)
-methods.
+\hl{relevent} for any quantitative spatial proteomics data,  
+irrespective of the fractionation (i.e. density gradient or 
+differential centrifugation \cite{Itzhak:2016}) or quantitation 
+(i.e. labelled or label-free) methods.
 
@@
 to the identification search (confidence scores, number of peptides,
-\ldots) as well as annotation about know sub-cellular location (see in
+\ldots) as well as annotation about known sub-cellular location (see in
 particular the \textit{Markers} section) and results from data
@@
 are only needed here to locate a file in a package; in a more general
-use case, the user would defined the \texttt{csvfile} variable
+use case, the user would define the \texttt{csvfile} variable
 containing the file name of interest directly.
@@
 <<combine, cache = cache>>=
-combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
+combined <- MSnbase::combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
 combined
@@
 exactly they differ (in terms of peak channels, for example). On
-figure} \ref{fig:plotDist2}, \hl{we plot the profile of the mitochondrial
-and peroxisome markers to highlight the differences in profiles
-between these two sets of markers in the channels labelled with tag
-129C, as represented above along the 7th PC on the PCA plot on
-figure} \ref{fig:plotmarkers}.
+figure} \ref{fig:plotDist2}, \hl{we plot the profiles of the 
+mitochondrial and peroxisome markers to highlight the differences 
+in the channels labelled with tag 129C, also represented above  
+along the 7th PC on the PCA plot on figure} \ref{fig:plotmarkers}.
 
@@
 \texttt{plot3D} function (Figure \ref{fig:plotmarkers3d}). When
-produced interactively, the plot can be rotated and zoomined using the
+produced interactively, the plot can be rotated and zoomed using the
 mouse.
@@
 
+\pagebreak
+
 \subsection*{Features of interest}
@@
 object \hl{and then use the \textit{highlightOnPlot} function to 
-visualise these}. For example, if we wanted to highlight \st{a} 
-proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, 
-which are some of the proteins known to form part of the 13S condensin complex, 
-we \hl{would call the following code:}
+visualise these}. For example, if we wanted to highlight 
+proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and 
+Q8C156, which are some of the proteins known to form part of the 
+13S condensin complex, we \hl{would call the following code:}
 
@@
 
+\pagebreak
+
 \section*{Replication}
@@
 Direct comparisons of individual channels in replicated experiments
-do\st{es} not provide an adequate, goal-driven assessment of different
+do not provide an adequate, goal-driven assessment of different
 experiments. Indeed, due to the nature of the experiment and gradient
@@
 @
+
+\pagebreak
 
@@
 
-\clearpage
+\pagebreak
 
@@
 
-\hl{The above analysis is comptationally intensive and best parallelised over
+\hl{The above analysis is computationally intensive and best parallelised over
 multiple workers. This phenoDisco analysis took ~24 hours to complete
-when parallelised over 40 workers. As such, to avoid executing this
-time consuming analysis users can access the above results which are
+when parallelised over 40 workers. As such, in the interest of time 
+users can access the above results which are
 pre-computed and stored along with the \texttt{pRolocdata} package. 
@@
 
+We can plot the results using the \texttt{plot2D} function (Figure
+\ref{fig:plotPDres}).
+
 \begin{figure}[!ht]
@@
 
-We can plot the results using the \texttt{plot2D} function (Figure
-\ref{fig:plotPDres}). The five new phenotype data clusters can be
+\pagebreak
+
+The five new phenotype data clusters can be
 extracted and examined.  In the code chunk below we write the results
@@
 \texttt{featureData} to write.
+
+\pagebreak
 
@@
 
-<<loadpdRes, echo=TRUE>>=
+<<loadpdRes, echo=FALSE>>=
 svmf <- dir(extdatadir, full.names = TRUE,
@@
 
+\pagebreak
+
 By using the function \texttt{getParams} we can extract the best set
@@
 
-<<classify>>=
+<<classify, warning = FALSE>>=
 hl <- svmClassification(hl, params, class.weights = w, fcol = "markers")
@@
 
-<<classify2, eval = FALSE>>=
+<<classify2, eval = FALSE, warning = FALSE>>=
 hl <- svmClassification(hl, cost = 16, sigma = 0.1, class.weights = w, fcol = "markers")
@@
 \end{figure}
+
+\pagebreak
 
@@
 Please see the appendix for details on how
-to load these directly}.)
-
-\st{To remove the 4 classes and create a new column of markers in the
-feature data called tlmarkers to use for the analysis:}
+to load these directly}.)q
 
@@
 good use of valuable prior knowledge, and in our experience
-unsupervised clustering can be extremely difficult due to poor
-estimates of the number of clusters that may appear in the data.
+unsupervised clustering can be extremely difficult due to \hl{
+(i) the loose definition of what constitutes a cluster (for example 
+whether it is defined by the quantitative data or the localisation 
+information), (ii) the influence of the algorithm assumption on the 
+cluster identification (for example parametric or non-parametric) 
+and (iii) poor estimates of the number of clusters that may appear 
+in the data.}
 
@@
 
-\subsubsection{Novelty detection results}
+\subsubsection*{Novelty detection results}
 
@@
 
-\subsubsection{SVM optimisation}
+\subsubsection*{SVM optimisation}
 
@@
 
-\subsubsection{Transfer learning results}
+\subsubsection*{Transfer learning results}
 

### 2018-04-11 bioc-workflow-resubmit.Rnw
verison for resubmission

--- 2018-04-11 bioc-workflow_march2017.Rnw
@@
 While we focus here on a LOPIT-type dataset, these analyses are
-\hl{relevent} for any quantitative spatial proteomics data,  
+relevent for any quantitative spatial proteomics data,  
 irrespective of the fractionation (i.e. density gradient or 
@@
 
-There are two aspects related to data normalisation that are \hl{relevent}
+There are two aspects related to data normalisation that are relevent
 to spatial proteomics data processing. The first one focuses on
@@
 spatial proteomics data, is scaling all the organelle-specific
-profiles into \hl{the} same intensity interval (typically 0 and 1) by, for
+profiles into the same intensity interval (typically 0 and 1) by, for
 example, dividing each intensity by the sum of the intensities for
@@
 data and their inappropriate handling leads to a reduction of
-sub-cellular resolution. \st{Missing data can be imputated} \hl{We can impute missing data} using
+sub-cellular resolution. We can impute missing data using
 \Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
@@
 \caption{Heatmap of missing values. Note that the features are
-  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
+  re-ordered to highlight clusters of proteins with similar numbers of
   missing values.}
@@
 \texttt{addLegend} function to map the marker classes to the
-pre-defined colours. \hl{As previously mentioned, PCA transforms the
+pre-defined colours. As previously mentioned, PCA transforms the
 original high dimensional data into a set of linearly uncorrelated
@@
 careful though as different organelles may be resolved in different
-dimensions.} For example, we can display the data along the first and
+dimensions. For example, we can display the data along the first and
 seventh PCs using the \texttt{dims} argument. Note that in these
@@
 
-\hl{This is further highlighted if we plot the
-profiles of these two clusters using the \texttt{plotDist} function}
-(Figure \ref{fig:plotDist2}). \hl{The \texttt{plotDist} function is
+This is further highlighted if we plot the
+profiles of these two clusters using the \texttt{plotDist} function
+(Figure \ref{fig:plotDist2}). The \texttt{plotDist} function is
 another useful visualisation that relies on marker annotation. It
@@
 exactly they differ (in terms of peak channels, for example). On
-figure} \ref{fig:plotDist2}, \hl{we plot the profiles of the 
+figure \ref{fig:plotDist2}, we plot the profiles of the 
 mitochondrial and peroxisome markers to highlight the differences 
 in the channels labelled with tag 129C, also represented above  
-along the 7th PC on the PCA plot on figure} \ref{fig:plotmarkers}.
+along the 7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
 
@@
 experiments) they would first create a \texttt{FeaturesOfInterest}
-object \hl{and then use the \textit{highlightOnPlot} function to 
-visualise these}. For example, if we wanted to highlight 
+object and then use the \textit{highlightOnPlot} function to 
+visualise these. For example, if we wanted to highlight 
 proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and 
 Q8C156, which are some of the proteins known to form part of the 
-13S condensin complex, we \hl{would call the following code:}
+13S condensin complex, we would call the following code:
 
@@
 
-The more \hl{relevent} comparison unit is not a single channel, but rather
+The more relevent comparison unit is not a single channel, but rather
 the complete protein occupancy profiles, which are best visualised 
@@
 
-\hl{The above analysis is computationally intensive and best parallelised over
+The above analysis is computationally intensive and best parallelised over
 multiple workers. This phenoDisco analysis took ~24 hours to complete
@@
 pre-computed and stored along with the \texttt{pRolocdata} package. 
-Please see the appendix to load these results.}
+Please see the appendix to load these results.
 
@@
 \href{http://bioconductor.org/packages/release/bioc/vignettes/pRoloc/inst/doc/pRoloc-ml.pdf}{\Biocpkg{pRoloc}
-  machine learning techniques vignette}.  We find the general \hl{tendency}
+  machine learning techniques vignette}.  We find the general tendency
 to be that it is not the choice of classifier, but the improper
@@
 
-\hl{In the code chunk below we then pass these weights 
+In the code chunk below we then pass these weights 
 to the \texttt{svmOptimisation} function. Once again, we provide
 the optimisation results for users to load directly if they wish to save
-computational time, please see the appendix to load these results.}
+computational time, please see the appendix to load these results.
 
@@
 information is not available, one crude method is to set a threshold
-per organelle by extracting the median or 3rd \hl{quartile} score per
+per organelle by extracting the median or 3rd quartile score per
 organelle.  For example, in the code chunk below, we use the
@@
 evidence codes as desired, see \texttt{?makeGoSet} for more details. 
-\hl{(We also provide the pre-computed \texttt{gocc} object for users
-if they wish to load directly, please see the appendix).}
+(We also provide the pre-computed \texttt{gocc} object for users
+if they wish to load directly, please see the appendix).
 
@@
 auxiliary data (so 0 is given to the primary source). If we conduct a
-parameter \hl{search} and test weights $\theta = {0, 1/3, 2/3, 1}$ for each
+parameter search and test weights $\theta = {0, 1/3, 2/3, 1}$ for each
 class, and if we have, for example 10 subcellular niches, this will
@@
 
-In the example below we \hl{first} remove these 4 classes from the marker set
-\hl{and create a new marker set called tlmarkers. Then we}
+In the example below we first remove these 4 classes from the marker set
+and create a new marker set called tlmarkers. Then we
 re-run the \texttt{knnOptimisation} for each data source and then run
@@
 hours to run on the University of Cambridge HPC using 256 workers. 
-\hl{To save time for users, the results of the following optimisation 
+To save time for users, the results of the following optimisation 
 are pre-computed and provided with the \texttt{pRolocdata} package. 
 Please see the appendix for details on how
-to load these directly}.)q
+to load these directly.
 
@@
 
-From examining the parameter \hl{search} plots as described in section
+From examining the parameter search plots as described in section
 \textit{Optimisation}, we find the best $k$'s for both the primary and
@@
 good use of valuable prior knowledge, and in our experience
-unsupervised clustering can be extremely difficult due to \hl{
+unsupervised clustering can be extremely difficult due to
 (i) the loose definition of what constitutes a cluster (for example 
@@
 and (iii) poor estimates of the number of clusters that may appear 
-in the data.}
+in the data.
 
@@
 data. To directly serialise R objects to disk, on can use the standard
-\texttt{save} function, and later reload the object using \st{save}
+\texttt{save} function, and later reload the object using
 \texttt{load}. For example, to save and then re-load the parameters

### 2018-04-11 bioc-workflow-resubmit.tex
verison for resubmission

--- 2018-04-10 bioc-workflow_march2017.tex
@@
 While we focus here on a LOPIT-type dataset, these analyses are
-\hl{relevent} for any quantitative spatial proteomics data, irrespective of
-the fractionation or quantitation (i.e. labelled or label-free)
-methods.
+relevent for any quantitative spatial proteomics data,  
+irrespective of the fractionation (i.e. density gradient or 
+differential centrifugation \cite{Itzhak:2016}) or quantitation 
+(i.e. labelled or label-free) methods.
 
@@
 \begin{alltt}
-\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
-\end{alltt}
-
-{\ttfamily\noindent\color{warningcolor}{\#\# Warning: 'combine' is deprecated.\\\#\# Use 'gtable\_combine' instead.\\\#\# See help("{}Deprecated"{})}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[[<-.data.frame`(`*tmp*`, i, value = structure(list(t = numeric(0), : replacement has 0 rows, data has 10}}\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlstd{MSnbase}\hlopt{::}\hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
+\hlstd{combined}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 6725 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData
+##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+##   varMetadata: labelDescription
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
+##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+##     total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
+## Annotation:  
+## - - - Processing information - - -
+## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:12:49 2018 
+##  MSnbase version: 2.5.9
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+More details about combining data are given in the dedicated
+\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
+\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
+  vignette}.
+
+\subsubsection*{Missing data}
+
+Missing data are a recurrent issue in mass spectrometry applications,
+and should be addressed independently of this workflow
+\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
+described how a high content in missing values in spatial proteomics
+data and their inappropriate handling leads to a reduction of
+sub-cellular resolution. We can impute missing data using
+\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
+the imputation method is then determined by a \texttt{methods}
+parameter (see \texttt{?impute} for available options). To impute
+missing values using nearest neighbour imputation, one would
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+In our particular case, missing values are indicative of protein
+groups that were not acquired in both replicates (Figure
+\ref{fig:namap}).
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
+       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/namap-1}
+
+\end{knitrout}
+\caption{Heatmap of missing values. Note that the features are
+  re-ordered to highlight clusters of proteins with similar numbers of
+  missing values.}
+  \label{fig:namap}
+\end{figure}
+
+We prefer to remove proteins that were not assayed in both replicated
+experiments. This is done with the \texttt{filterNA} function that
+removes features (proteins) that contain more than a certain
+proportion (default is 0) missing values. The \textit{Processing
+  information} section summarises how combining and filtering missing
+values (subsetting) changed the dimensions of the data.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
 \hlstd{combined}
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
-## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
-## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
-## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
-## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
-## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
-## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
-## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
-## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
-## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
-##  MSnbase version: 2.5.9
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-More details about combining data are given in the dedicated
-\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
-\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
-  vignette}.
-
-\subsubsection*{Missing data}
-
-Missing data are a recurrent issue in mass spectrometry applications,
-and should be addressed independently of this workflow
-\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
-described how a high content in missing values in spatial proteomics
-data and their inappropriate handling leads to a reduction of
-sub-cellular resolution. \st{Missing data can be imputated} \hl{We can impute missing data} using
-\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
-the imputation method is then determined by a \texttt{methods}
-parameter (see \texttt{?impute} for available options). To impute
-missing values using nearest neighbour imputation, one would
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-In our particular case, missing values are indicative of protein
-groups that were not acquired in both replicates (Figure
-\ref{fig:namap}).
-
-\begin{figure}[!ht]
-  \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
-       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=.65\textwidth]{figure/namap-1}
-
-\end{knitrout}
-\caption{Heatmap of missing values. Note that the features are
-  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
-  missing values.}
-  \label{fig:namap}
-\end{figure}
-
-We prefer to remove proteins that were not assayed in both replicated
-experiments. This is done with the \texttt{filterNA} function that
-removes features (proteins) that contain more than a certain
-proportion (default is 0) missing values. The \textit{Processing
-  information} section summarises how combining and filtering missing
-values (subsetting) changed the dimensions of the data.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
-\hlstd{combined}
-\end{alltt}
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 5032 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData
-##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
-##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
-##   varMetadata: labelDescription
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
-##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
-##     total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-##   pubMedIds: 26754106 
-## Annotation:  
-## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
-## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
-## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
-## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
-## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
-## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
-## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
-## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
-## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
-## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:52:54 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:52:54 2018 
-## Dropped featureData's levels Tue Apr 10 16:52:54 2018 
+## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:12:49 2018 
+## Subset [6725,20][5032,20] Wed Apr 11 14:12:50 2018 
+## Removed features with more than 0 NAs: Wed Apr 11 14:12:50 2018 
+## Dropped featureData's levels Wed Apr 11 14:12:50 2018 
 ##  MSnbase version: 2.5.9
@@
 exactly they differ (in terms of peak channels, for example). On
-figure} \ref{fig:plotDist2}, \hl{we plot the profile of the mitochondrial
-and peroxisome markers to highlight the differences in profiles
-between these two sets of markers in the channels labelled with tag
-129C, as represented above along the 7th PC on the PCA plot on
-figure} \ref{fig:plotmarkers}.
+figure \ref{fig:plotDist2}, we plot the profiles of the 
+mitochondrial and peroxisome markers to highlight the differences 
+in the channels labelled with tag 129C, also represented above  
+along the 7th PC on the PCA plot on figure \ref{fig:plotmarkers}.
 
@@
 experiments) they would first create a \texttt{FeaturesOfInterest}
-object \hl{and then use the \textit{highlightOnPlot} function to 
-visualise these}. For example, if we wanted to highlight \st{a} 
-proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and Q8C156, 
-which are some of the proteins known to form part of the 13S condensin complex, 
-we \hl{would call the following code:}
+object and then use the \textit{highlightOnPlot} function to 
+visualise these. For example, if we wanted to highlight 
+proteins with the accession numbers Q8CG48, Q8CG47, Q8K2Z4, and 
+Q8C156, which are some of the proteins known to form part of the 
+13S condensin complex, we would call the following code:
 
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 10 16:52:56 2018 
+##  Created on Wed Apr 11 14:12:52 2018 
 ##  Description:
@@
 % latex table generated in R 3.5.0 by xtable 1.8-2 package
-% Tue Apr 10 16:52:56 2018
+% Wed Apr 11 14:12:52 2018
 \begin{table}[ht]
@@
 
-The more \hl{relevent} comparison unit is not a single channel, but rather
+\pagebreak
+
+The more relevent comparison unit is not a single channel, but rather
 the complete protein occupancy profiles, which are best visualised 
@@
 
-\hl{The above analysis is comptationally intensive and best parallelised over
+The above analysis is computationally intensive and best parallelised over
 multiple workers. This phenoDisco analysis took ~24 hours to complete
-when parallelised over 40 workers. As such, to avoid executing this
-time consuming analysis users can access the above results which are
+when parallelised over 40 workers. As such, in the interest of time 
+users can access the above results which are
 pre-computed and stored along with the \texttt{pRolocdata} package. 
-Please see the appendix to load these results.}
+Please see the appendix to load these results.
 
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 10 16:52:55 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 10 16:52:56 2018 
+## Added markers from  'mrk' marker vector. Wed Apr 11 14:12:52 2018 
+## Added markers from  'pdres' marker vector. Wed Apr 11 14:12:53 2018 
 ##  MSnbase version: 2.5.10
@@
 
-\hl{In the code chunk below we then pass these weights 
+In the code chunk below we then pass these weights 
 to the \texttt{svmOptimisation} function. Once again, we provide
 the optimisation results for users to load directly if they wish to save
-computational time, please see the appendix to load these results.}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{svmf} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
-              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-svmopt.rds"}\hlstd{)}
-\hlstd{params} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(svmf)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
+computational time, please see the appendix to load these results.
 
@@
 \end{alltt}
-
-{\ttfamily\noindent\color{warningcolor}{\#\# Warning in seq\_len(trpr): first element used of 'length.out' argument}}\end{kframe}
+\end{kframe}
 \end{knitrout}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4323379               0.3037630               0.3822328 
+##               0.4336250               0.3041097               0.3895958 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6930642               0.6014878               0.4426545 
+##               0.6855842               0.5981495               0.4333482 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4224179               0.5978957               0.9497788 
+##               0.4243711               0.5982753               0.9494273 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950203               0.7117761               0.3147679 
+##               0.7951711               0.7080911               0.3141263 
 ##         Plasma membrane              Proteasome 
-##               0.7042806               0.4150910
+##               0.7170268               0.4188616
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4323379               0.3037630               0.3822328 
+##               0.4336250               0.3041097               0.3895958 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6930642               0.6014878               0.4426545 
+##               0.6855842               0.5981495               0.4333482 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4224179               0.5978957               0.9497788 
+##               0.4243711               0.5982753               0.9494273 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950203               0.7117761               0.3147679 
+##               0.7951711               0.7080911               0.3141263 
 ##         Plasma membrane              Proteasome 
-##               0.7042806               0.4150910
+##               0.7170268               0.4188616
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      84 
+##                      85                     171                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     301                     478                      99 
+##                     294                     474                     107 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     523 
+##                      26                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     341                      38 
+##                     230                     344                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     154                    2057
+##                     310                     159                    2058
 \end{verbatim}
@@
 hours to run on the University of Cambridge HPC using 256 workers. 
-\hl{To save time for users, the results of the following optimisation 
+To save time for users, the results of the following optimisation 
 are pre-computed and provided with the \texttt{pRolocdata} package. 
 Please see the appendix for details on how
-to load these directly}.)
-
-\st{To remove the 4 classes and create a new column of markers in the
-feature data called tlmarkers to use for the analysis:}
+to load these directly.
 
@@
 good use of valuable prior knowledge, and in our experience
-unsupervised clustering can be extremely difficult due to poor
-estimates of the number of clusters that may appear in the data.
+unsupervised clustering can be extremely difficult due to
+(i) the loose definition of what constitutes a cluster (for example 
+whether it is defined by the quantitative data or the localisation 
+information), (ii) the influence of the algorithm assumption on the 
+cluster identification (for example parametric or non-parametric) 
+and (iii) poor estimates of the number of clusters that may appear 
+in the data.
 
@@
 ## other attached packages:
-##  [1] gridExtra_2.3        pRolocdata_1.17.3    pRoloc_1.19.3       
+##  [1] gridExtra_2.3        pRolocdata_1.17.4    pRoloc_1.19.3       
 ##  [4] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     

### 2018-04-11 replies-to-reviewers.tex
reply to reviewers updated

--- 2018-03-26 replies-to-reviewers.tex
@@
 
-We thank reviewer 1 for their comments. Please find our responses to
-these inset below.
+\begin{quote} \textcolor{gray} {Breckels et al.
+have written a very nice piece on analysing appropriate proteomics data for subcellular localisation.
+I particularly like the "workshop characteristics" of the text.
+Which allows a novice, but interested reader to work through the analysis stepwise and reproduce the results described therein.
+The authors took great care in keeping this ideal up during their text and this is also where I have put my greatest reservation to the manuscript in its present form - since a reader cannot work through the code presented in the manuscript, since there at at least two situations where a readily available HPC and quite some time is required.
+This kind of leaves a dent in my impression - however, given this can be resolved as well as some typos - the workflow report is superb.} \end{quote}
 
-\begin{quote} \textcolor{gray}{
-Next to reducing the dimensions of data for visualisation, PCA also offers a way to understand how the variability is distributed across the multidimensional data by providing linear combinations of the variables which then constitute the actual PCs.
+We thank reviewer 1 for their comments. Please find our responses to these inset below. \newline \newline
+
+\textbf{Major comments:} \\
+
+\textcolor{gray}{Next to reducing the dimensions of data for visualisation, PCA also offers a way to understand how the variability is distributed across the multidimensional data by providing linear combinations of the variables which then constitute the actual PCs.
 On that note it would be nice to mention this in Visualising markers section on page 16, where PC7 explains not much variability but due to the correct weighing of the variables we do get a separation between mitochondrial and peroxisome.
 This then can be further motivated with Figure 9 - where we probably can see that the weights for the fractions where the two localisations differ are larger than otherwise.}
-\end{quote}
 
-We have added a paragraph reminding users what PCA does in the Visualising Markers section to motivate the choice of looking at PC's 1 and 7.
-Figure 9 (now Figure 8) and corresponding code has been moved to this section and explanation of the plotDist function to lead on from the reference to marker profiles and separation.
+We have added a paragraph to the 'Visualising markers' section of the manuscript reiterating the purpose of PCA and motivating the choice of looking at PC's 1 and 7.
+Figure 9 now follows on from this (now Figure 8), along with the corresponding code and an explanation of the \texttt{plotDist} function.
+\\
 
-\begin{quote}  \textcolor{gray}{
-}
-\end{quote}
+\textcolor{gray}{I was unable to reproduce Figure 13 comparing the two MSnSets.
+While I was able to look at each set separately using pRolocVis(hllst@x[[I]]), where i is 1 or 2, I only got an error using the code from the manuscript.
+When using ?remap=FALSE?
+it actually works, but since this makes barely sense it is of no use - but just as a hint at debugging it.}
 
-We thank the reviewer for their constructive and positive comments.
+We can not reproduce this error. Have you updated to the latest version of R and the latest version of \texttt{pRolocGUI}? If you still get this error message could you please post this as an issue \\
+(https://github.com/ComputationalProteomicsUnit/pRolocGUI/issues) on the \texttt{pRolocGUI} Github page along with your \texttt{SessionInfo()} and we will certainly attempt to solve this. \\
+
+\textcolor{gray} {You really need to make the results from the phenoDisco classification available too.
+It is super disappointing that one cannot continue reproducing the code from page 23 on, because it takes 24 hours to compute it using 40 cores?}
+
+The results are already available as a \texttt{RDS} file and stored in \texttt{pRolocdata} for users. This is what is called in the manuscript under the hood: \\
+
+\small{
+\begin{lstlisting}[language=R] 
+> f0 <- dir(extdatadir, full.names = TRUE, pattern = "bpw-pdres.rds")
+> pdres <- readRDS(f0)
+> hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
+\end{lstlisting}}
+
+We have made this code available in the manuscript in an appendix so users can continue to produce the exact plots as they see in this workflow. \\
+
+\textcolor{gray} {The above comment is of course also true for the KNN TL Optimisation on page 33 - this needs to be downloadable, since not everyone has access to Cambridge?s HPC and probably even less have 76 hours to spare.}
+
+The same as for the \texttt{phenoDisco} analysis and \texttt{svm}, the TL results are stored as a \texttt{RDS} in \texttt{pRolocdata} and are loaded in the background.
+We have added the code required to the appendix so that users can load the results directly.
+\\
+
+\textcolor{gray} {Your comment on the increase suitability of classification instead of clustering (when additional information on classes is available) at the bottom of page 35 could be more pronounced - for educational reasons.}
+
+To address the above comment on suitability we have added a few additional points on the challenges of using clustering for this type of data. \\
+
+We generally find supervised learning more suited to the task of protein localisation prediction in which we use high-quality curated marker proteins to build a classifier, instead of using an entirely unsupervised approach to look for clusters and then look for enrichment of organelles and complexes.
+In the latter we do not make good use of valuable prior knowledge, and in our experience unsupervised clustering can be extremely difficult due to (i) the loose definition of what constitutes a cluster (for example whether it is defined by the quantitative data or the localisation information), (ii) the influence of the algorithm assumption on the cluster identification (for example parametric or non-parametric) and (iii) poor estimates of the number of clusters that may appear in the data.
+\newline \newline
+
+\textbf{Minor comments:} \\
+\textcolor{gray} {I was not able to naively reproduce the workflow from the R commands in the article due to an error installing pRolocdata on a Windows machine. On OS X it was smooth.}
+
+We didn't experience any problems with installing \texttt{pRolocdata} on Windows.
+If you re-try the installation and please let us know if you still have any issues by opening an issue (https://github.com/lgatto/pRolocdata/issues) or posting on the Bioconductor Support site (https://support.bioconductor.org).
+\\
+
+\textcolor{gray} {On page 10 line 2 there is a ?to? missing.}
+
+In this version we currently can't find the missing 'to'.  \\
+
+\textcolor{gray} {I never came across the verb imputate in the context of missing values, I guess the proper term is impute.}
+
+This has been changed to read "We can impute missing data..." \\
+
+\textcolor{gray} {On page 11 the image2 function is called after the filterNA function a couple of lines above.
+This however would result in an only black heat map (since there are no more missing).
+The image2 function should be called before the filterNA function.
+Since the reader does not see the chunk options, it could be puzzling.}
+
+This was an editing mistake and has now been rectified. \\
+
+\textcolor{gray} {For completeness sake there should also be an install.packages(c("hexbin", "rgl")) somewhere to generate the second PCA-plot and the 3D plot.
+Moreover, Mac users will need to install xquartz to use rgl properly.}
+
+A footnote has been added here to tell users that the package \texttt{rgl} may need to be installed with \texttt{install.packages("rgl")} and mac users may need to install xquartz if it's not already installed.
+\\
+
+\textcolor{gray} {On page 14 the plotting code chunk is off track - in the middle of the marker sets output.}
+
+This has now been rectified. \\
+
+\textcolor{gray} {On page 18: ...wanted to highlight a proteins with the ... -> lose the a and later in the sentence there is a 'create a' too many.}
+
+These typos have been rectified. \\
+
+\textcolor{gray} {Direct comparisons of individual channels in replicated experiments do not provide?}
+
+This typo has been rectified. \\
+
+\textcolor{gray} {You may want to consider adding a layout(1) or similar, after changing the mfrow argument of the parameters to accommodate 2 panels, such that the uncanny reader does not get confused.}
+
+We would prefer to keep the code as it is and not introduce more noise with calls to other functions such as `layout`.
+The workflow is not aimed at teaching R.
+Users should have some basic knowledge of R before tackling this tutorial.
+\\
+
+\textcolor{gray} {I would prefer links to referred sections of the text, but that may be personal taste?}
+
+This is a comment for f1000. We can not control the linking of sections in the final version. \\
+
+\textcolor{gray} {Page 23: One should note that the decreasing the GS, and increasing the ...at least one the too many, probably two.}
+
+We have reworded this sentence as requested. \\
+
+\textcolor{gray} {On page 25: We find the general tendancy to be that it is not the choice ...tendency?}
+
+This typo has been rectified. \\
+
+\textcolor{gray} {On page 28 you refer to '...the code chunk below...' for Figure 17, however, the following code chunk is generating Figure 16 (which is above and btw not referenced in the text).
+Maybe force your figures a little to float where you want them/refer to them.}
+
+We have now referenced Figure 16 in the text and made sure that the code chunks and figures follow inline where they are referenced in the text. \\
+
+\textcolor{gray} {On page 28: ...by extracting the median or 3rd quantile score per organelle? do you mean quartile? Otherwise I do not follow.}
+
+Thank you, yes this is a typo and has now been changed to 'quartile'. \\
+
+\textcolor{gray} {On page 32: package to query the relevent database ? relevant?}
+
+This typo has been rectified. \\
+
+\textcolor{gray} {On page 32 - there is something wrong with this sentence: To remove the 4 classes and create a new column of markers in the feature data called tlmarkers to use for the analysis:}
+
+This sentence is not needed here and so it has now been removed as it essentially reiterates what is said in the above paragraph. \\
+
+\textcolor{gray} {On page 34: From examining the parameter seach plots as described in section Optimisation? search!}
+
+This typo has been rectified.  \\
+
+\textcolor{gray} {On page 36: ?and later reload the object using save. -> that would be ?load? then!}
+
+This typo has been rectified. \\
+
+\textcolor{gray} {On page 38 -  I fully agree with the following sentence, but right after the updating comment it kind of seems ?misplaced?? Maybe add a title like ?Getting help??}
+
+We have changed the title of this section to 'Session information and getting help' to clarify this section of the tutorial. \\
 
@@
 
-\begin{quote}\textcolor{gray}{ }
-\end{quote}
+\begin{quote} \textcolor{gray} {This manuscript describes a Bioconductor workflow for analyzing subcellular proteomics data.
+It is very detailed and comprehensive and will be useful for others in the field.} \end{quote}
+
+Many thanks for your comments, we have responded to them inset below. \\ \newline
+
+\textbf{A few comments:}
+
+\textcolor{gray} {Some clearer statement early on would help to clarify for readers what types of data this works with.
+I know that the authors indicate that the example they use is 10-plex TMT and that it can be used with label-free or other labels, but that is not what I am referring to.
+Rather, structure of the experiment.
+That is, that one needs systematic quantitative data on all the different relevant fractions from a cell, as opposed to someone who perhaps did a differential centrifugation experiment to isolate a couple fractions and then wants to apply this (my understanding is that this latter example would not be usable).}
+
+From a purely technical point of view, we need a data matrix with features (typically proteins or protein groups) along the rows and sub-cellular fractions along the columns.
+Given the requirement for complete (or near-complete) quantitative vectors along all fractions to assure best results, a data set that would only contain one quantitation value per fraction would not work.
+However, separation using differential centrifugation, or any separation that generates  rganelle-specific separation profiles is a good fit for pRoloc.
+For instance, the type of data generated by methods such as described in Itzhak et al.
+(2016) [1] are well fitted for our software: \\
+
+\small{
+\begin{lstlisting}[language=R] 
+> library(pRoloc)
+> library(pRolocdata)
+> data(itzhak2016stcSILAC)
+> ## 6 combined replicates of 5 fractions each
+> dim(itzhak2016stcSILAC)
+[1] 5265   30
+> plot2D(itzhak2016stcSILAC)
+\end{lstlisting}}
+
+If there were only very few fractions (say cytosol, nucleus and plasma membrane), the infrastructure could still be used, although for less benefits. \\
+
+\textcolor{gray} {How do the authors recommend collapsing replicates?
+This could be covered in the section dedicated to the Compare function.
+Two replicates will (almost) never agree 100\% so how are discrepancies handled?}
+
+Currently, we recommend to visualise different replicates on their own, to confirm that they are of sufficient quality, and then combine them, retaining the proteins that have been quantified over all replicated experiments.
+This allows to obtain localisation information over all replicated data.
+We however do not explicitly assess the variability using this approach.
+This could be done by analysing replicated independently and then compare the coherence of the classification results.
+Proteins that are only observed in some replicates could be rescued by repeating the analysis using only the relevant (possibly unique) replicate(s).
+\\ \newline
+
+\small{References
+[1] Itzhak DN, Tyanova S, Cox J, Borner GH.
+Global, quantitative and dynamic mapping of protein subcellular localization.
+Elife.
+2016 Jun 9;5.
+pii: e16950.
+doi: 10.7554/eLife.16950.
+PubMed PMID: 27278775; PubMed Central PMCID: PMC4959882.}
 

### 2018-04-16 replies-to-reviewers.tex
typos

@@
 We didn't experience any problems with installing \texttt{pRolocdata} on Windows.
-If you re-try the installation and please let us know if you still have any issues by opening an issue (https://github.com/lgatto/pRolocdata/issues) or posting on the Bioconductor Support site (https://support.bioconductor.org).
+If you re-try the installation and please let us know if you still have any issues by opening an issue (https://github.com/lgatto/pRolocdata/issues) or by posting this issue on the Bioconductor Support site (https://support.bioconductor.org).
 \\
@@
 
-Currently, we recommend to visualise different replicates on their own, to confirm that they are of sufficient quality, and then combine them, retaining the proteins that have been quantified over all replicated experiments.
+Currently, we recommend to visualise different replicates on their own, to confirm that they are of sufficient quality, and then combine them, retaining the proteins that have been quantified over all replicatedd experiments.
 This allows to obtain localisation information over all replicated data.

### 2018-04-16 bioc-workflow-resubmit.tex
don't load and attach gridExtra

@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.5.10
+##  MSnbase version: 2.4.2
 \end{verbatim}
@@
 
-will list the 75 datasets that are
+will list the 62 datasets that are
 available in \Biocexptpkg{pRolocdata}.
@@
 ## experimentData: use 'experimentData(object)'
-##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:12:49 2018 
-##  MSnbase version: 2.5.9
+## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:26:11 2018 
+##  MSnbase version: 1.21.7
 \end{verbatim}
@@
 ## experimentData: use 'experimentData(object)'
-##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:12:49 2018 
-## Subset [6725,20][5032,20] Wed Apr 11 14:12:50 2018 
-## Removed features with more than 0 NAs: Wed Apr 11 14:12:50 2018 
-## Dropped featureData's levels Wed Apr 11 14:12:50 2018 
-##  MSnbase version: 2.5.9
+## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:26:11 2018 
+## Subset [6725,20][5032,20] Mon Apr 16 15:26:12 2018 
+## Removed features with more than 0 NAs: Mon Apr 16 15:26:12 2018 
+## Dropped featureData's levels Mon Apr 16 15:26:12 2018 
+##  MSnbase version: 1.21.7
 \end{verbatim}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Wed Apr 11 14:12:52 2018 
+##  Created on Mon Apr 16 15:26:14 2018 
 ##  Description:
@@
 
-% latex table generated in R 3.5.0 by xtable 1.8-2 package
-% Wed Apr 11 14:12:52 2018
+% latex table generated in R 3.4.4 by xtable 1.8-2 package
+% Mon Apr 16 15:26:14 2018
 \begin{table}[ht]
@@
 \begin{alltt}
-\hlcom{## As per Christoforou et al (2016), }
+\hlcom{## As per Christoforou et al (2016),}
 \hlstd{hl} \hlkwb{<-} \hlkwd{phenoDisco}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"phenoDisco.Input"}\hlstd{,} \hlkwc{times} \hlstd{=} \hlnum{200}\hlstd{,} \hlkwc{GS} \hlstd{=} \hlnum{60}\hlstd{)}
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Apr 11 14:12:52 2018 
-## Added markers from  'pdres' marker vector. Wed Apr 11 14:12:53 2018 
-##  MSnbase version: 2.5.10
+## Added markers from  'mrk' marker vector. Mon Apr 16 15:26:13 2018 
+## Added markers from  'pdres' marker vector. Mon Apr 16 15:26:14 2018 
+##  MSnbase version: 2.4.2
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4336250               0.3041097               0.3895958 
+##               0.4355329               0.3041412               0.3939341 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6855842               0.5981495               0.4333482 
+##               0.6838362               0.6082860               0.4344442 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4243711               0.5982753               0.9494273 
+##               0.4147867               0.5994097               0.9498454 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951711               0.7080911               0.3141263 
+##               0.7961992               0.7100991               0.3171891 
 ##         Plasma membrane              Proteasome 
-##               0.7170268               0.4188616
+##               0.7093459               0.4225532
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4336250               0.3041097               0.3895958 
+##               0.4355329               0.3041412               0.3939341 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6855842               0.5981495               0.4333482 
+##               0.6838362               0.6082860               0.4344442 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4243711               0.5982753               0.9494273 
+##               0.4147867               0.5994097               0.9498454 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951711               0.7080911               0.3141263 
+##               0.7961992               0.7100991               0.3171891 
 ##         Plasma membrane              Proteasome 
-##               0.7170268               0.4188616
+##               0.7093459               0.4225532
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      89 
+##                      85                     170                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     294                     474                     107 
+##                     299                     476                      92 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     124                     523 
+##                      27                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      38 
+##                     229                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     310                     159                    2058
+##                     322                     156                    2057
 \end{verbatim}
@@
 \begin{alltt}
-\hlcom{## Set appropriate parallelisation backend and  }
+\hlcom{## Set appropriate parallelisation backend and}
 \hlcom{## number of workers for the tl}
@@
 
-\hlcom{## Now peform tl optimisation }
+\hlcom{## Now peform tl optimisation}
 \hlstd{tlopt} \hlkwb{<-} \hlkwd{knntlOptimisation}\hlstd{(hl, gocc,} \hlkwc{fcol} \hlstd{=} \hlstr{"tlmarkers"}\hlstd{,}
@@
 \begin{alltt}
-\hlcom{## re-order classes }
+\hlcom{## re-order classes}
 \hlstd{bestpar} \hlkwb{<-} \hlstd{bestpar[}\hlkwd{getMarkerClasses}\hlstd{(hl)]}
@@
 \begin{verbatim}
-## R Under development (unstable) (2018-03-21 r74432)
-## Platform: x86_64-apple-darwin15.6.0 (64-bit)
-## Running under: macOS High Sierra 10.13.3
+## R version 3.4.4 Patched (2018-03-19 r74516)
+## Platform: x86_64-pc-linux-gnu (64-bit)
+## Running under: Ubuntu 14.04.5 LTS
 ## 
 ## Matrix products: default
-## BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
-## LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib
+## BLAS: /usr/lib/atlas-base/atlas/libblas.so.3.0
+## LAPACK: /usr/lib/lapack/liblapack.so.3.0
 ## 
 ## locale:
-## [1] en_GB.UTF-8/en_GB.UTF-8/en_GB.UTF-8/C/en_GB.UTF-8/en_GB.UTF-8
+##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
+##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
+##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
+##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
+##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
+## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
 ## 
 ## attached base packages:
-## [1] stats4    parallel  stats     graphics  grDevices utils     datasets 
-## [8] methods   base     
+## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
+## [8] datasets  base     
 ## 
 ## other attached packages:
-##  [1] gridExtra_2.3        pRolocdata_1.17.4    pRoloc_1.19.3       
-##  [4] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     
-##  [7] XML_3.98-1.10        AnnotationDbi_1.41.4 IRanges_2.13.28     
-## [10] S4Vectors_0.17.37    MSnbase_2.5.10       ProtGenerics_1.11.0 
-## [13] BiocParallel_1.13.1  mzR_2.13.5           Rcpp_0.12.16        
-## [16] Biobase_2.39.2       BiocGenerics_0.25.3  xtable_1.8-2        
-## [19] BiocStyle_2.7.8      knitr_1.20          
+##  [1] pRolocdata_1.16.0    pRoloc_1.18.0        MLInterfaces_1.58.1 
+##  [4] cluster_2.0.7-1      annotate_1.56.2      XML_3.98-1.10       
+##  [7] AnnotationDbi_1.40.0 IRanges_2.12.0       S4Vectors_0.16.0    
+## [10] MSnbase_2.4.2        ProtGenerics_1.10.0  BiocParallel_1.12.0 
+## [13] mzR_2.12.0           Rcpp_0.12.16         Biobase_2.38.0      
+## [16] BiocGenerics_0.24.0  xtable_1.8-2         BiocStyle_2.6.1     
+## [19] knitr_1.20          
 ## 
@@
 ##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
-##   [4] lazyeval_0.2.1        splines_3.5.0         ggvis_0.4.3          
+##   [4] lazyeval_0.2.1        splines_3.4.4         ggvis_0.4.3          
 ##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
-##  [10] foreach_1.4.4         BiocInstaller_1.29.4  htmltools_0.3.6      
-##  [13] viridis_0.5.0         gdata_2.18.0          magrittr_1.5         
+##  [10] foreach_1.4.4         BiocInstaller_1.28.0  htmltools_0.3.6      
+##  [13] viridis_0.5.1         gdata_2.18.0          magrittr_1.5         
 ##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
-##  [19] limma_3.35.13         recipes_0.1.2         gower_0.1.2          
+##  [19] limma_3.34.9          recipes_0.1.2         gower_0.1.2          
 ##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
-##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.0           
+##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.1           
 ##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
-##  [31] genefilter_1.61.1     bindr_0.1.1           impute_1.53.0        
-##  [34] survival_2.41-3       iterators_1.0.9       glue_1.2.0           
-##  [37] DRR_0.0.3             gtable_0.2.0          ipred_0.9-6          
-##  [40] zlibbioc_1.25.0       kernlab_0.9-25        ddalpha_1.3.1.1      
-##  [43] prabclus_2.2-6        DEoptimR_1.0-8        scales_0.5.0         
-##  [46] vsn_3.47.1            mvtnorm_1.0-7         DBI_0.8              
-##  [49] viridisLite_0.3.0     progress_1.1.2        foreign_0.8-70       
-##  [52] bit_1.1-12            proxy_0.4-21          mclust_5.4           
-##  [55] preprocessCore_1.41.0 lava_1.6              prodlim_1.6.1        
-##  [58] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
-##  [61] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
-##  [64] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
-##  [67] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
-##  [70] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
-##  [73] munsell_0.4.3         mlbench_2.1-1         tools_3.5.0          
-##  [76] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
-##  [79] evaluate_0.10.1       stringr_1.3.0         mzID_1.17.0          
-##  [82] yaml_2.1.18           ModelMetrics_1.1.0    bit64_0.9-7          
-##  [85] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
-##  [88] dendextend_1.7.0      bindrcpp_0.2          nlme_3.1-136         
-##  [91] whisker_0.3-2         mime_0.5              RcppRoll_0.2.2       
-##  [94] biomaRt_2.35.12       compiler_3.5.0        e1071_1.6-8          
-##  [97] affyio_1.49.1         tibble_1.4.2          stringi_1.1.7        
-## [100] highr_0.6             lattice_0.20-35       trimcluster_0.1-2    
-## [103] Matrix_1.2-12         psych_1.7.8           gbm_2.1.3            
-## [106] pillar_1.2.1          MALDIquant_1.17       bitops_1.0-6         
-## [109] httpuv_1.3.6.2        R6_2.2.2              pcaMethods_1.71.0    
-## [112] affy_1.57.0           hwriter_1.3.2         codetools_0.2-15     
-## [115] MASS_7.3-49           gtools_3.5.0          assertthat_0.2.0     
-## [118] CVST_0.2-1            rprojroot_1.3-2       withr_2.1.2          
-## [121] mnormt_1.5-5          diptest_0.75-7        grid_3.5.0           
-## [124] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
-## [127] class_7.3-14          rmarkdown_1.9         lubridate_1.7.3      
-## [130] shiny_1.0.5           base64enc_0.1-3
+##  [31] genefilter_1.60.0     bindr_0.1.1           impute_1.52.0        
+##  [34] DRR_0.0.3             survival_2.41-3       iterators_1.0.9      
+##  [37] glue_1.2.0            gtable_0.2.0          ipred_0.9-6          
+##  [40] zlibbioc_1.24.0       kernlab_0.9-25        ddalpha_1.3.2        
+##  [43] prabclus_2.2-6        DEoptimR_1.0-8        abind_1.4-5          
+##  [46] scales_0.5.0.9000     vsn_3.46.0            mvtnorm_1.0-7        
+##  [49] DBI_0.8               viridisLite_0.3.0     progress_1.1.2       
+##  [52] magic_1.5-8           foreign_0.8-69        bit_1.1-12           
+##  [55] proxy_0.4-22          mclust_5.4            preprocessCore_1.40.0
+##  [58] lava_1.6.1            prodlim_1.6.1         sampling_2.8         
+##  [61] htmlwidgets_1.0       httr_1.3.1            threejs_0.3.1        
+##  [64] FNN_1.1               RColorBrewer_1.1-2    fpc_2.1-11           
+##  [67] modeltools_0.2-21     pkgconfig_2.0.1       flexmix_2.3-14       
+##  [70] nnet_7.3-12           caret_6.0-79          tidyselect_0.2.4     
+##  [73] rlang_0.2.0           reshape2_1.4.3        munsell_0.4.3        
+##  [76] mlbench_2.1-1         tools_3.4.4           msdata_0.18.0        
+##  [79] RSQLite_2.1.0         pls_2.6-0             broom_0.4.4          
+##  [82] evaluate_0.10.1       geometry_0.3-6        stringr_1.3.0        
+##  [85] mzID_1.16.0           yaml_2.1.18           ModelMetrics_1.1.0   
+##  [88] bit64_0.9-7           robustbase_0.92-8     randomForest_4.6-14  
+##  [91] purrr_0.2.4           dendextend_1.7.0      bindrcpp_0.2.2       
+##  [94] nlme_3.1-137          whisker_0.3-2         mime_0.5             
+##  [97] RcppRoll_0.2.2        biomaRt_2.34.2        compiler_3.4.4       
+## [100] e1071_1.6-8          
+##  [ reached getOption("max.print") -- omitted 37 entries ]
 \end{verbatim}

### 2018-04-16 bioc-workflow-annotated.tex

--- 2018-04-10 bioc-workflow_march2017.tex
@@
 \begin{alltt}
-\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
-\end{alltt}
-
-{\ttfamily\noindent\color{warningcolor}{\#\# Warning: 'combine' is deprecated.\\\#\# Use 'gtable\_combine' instead.\\\#\# See help("{}Deprecated"{})}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[[<-.data.frame`(`*tmp*`, i, value = structure(list(t = numeric(0), : replacement has 0 rows, data has 10}}\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlstd{MSnbase}\hlopt{::}\hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
+\hlstd{combined}
+\end{alltt}
+\begin{verbatim}
+## MSnSet (storageMode: lockedEnvironment)
+## assayData: 6725 features, 20 samples 
+##   element names: exprs 
+## protocolData: none
+## phenoData
+##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
+##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
+##   varMetadata: labelDescription
+## featureData
+##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2Y3-3 (6725 total)
+##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
+##     total)
+##   fvarMetadata: labelDescription
+## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
+## Annotation:  
+## - - - Processing information - - -
+## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:03:28 2018 
+##  MSnbase version: 2.5.9
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+More details about combining data are given in the dedicated
+\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
+\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
+  vignette}.
+
+\subsubsection*{Missing data}
+
+Missing data are a recurrent issue in mass spectrometry applications,
+and should be addressed independently of this workflow
+\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
+described how a high content in missing values in spatial proteomics
+data and their inappropriate handling leads to a reduction of
+sub-cellular resolution. \st{Missing data can be imputated} \hl{We can impute missing data} using
+\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
+the imputation method is then determined by a \texttt{methods}
+parameter (see \texttt{?impute} for available options). To impute
+missing values using nearest neighbour imputation, one would
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+In our particular case, missing values are indicative of protein
+groups that were not acquired in both replicates (Figure
+\ref{fig:namap}).
+
+\begin{figure}[!ht]
+  \centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
+       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
+\end{alltt}
+\end{kframe}
+\includegraphics[width=.65\textwidth]{figure/namap-1}
+
+\end{knitrout}
+\caption{Heatmap of missing values. Note that the features are
+  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
+  missing values.}
+  \label{fig:namap}
+\end{figure}
+
+We prefer to remove proteins that were not assayed in both replicated
+experiments. This is done with the \texttt{filterNA} function that
+removes features (proteins) that contain more than a certain
+proportion (default is 0) missing values. The \textit{Processing
+  information} section summarises how combining and filtering missing
+values (subsetting) changed the dimensions of the data.
+
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
 \hlstd{combined}
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
-## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
-## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
-## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
-## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
-## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
-## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
-## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
-## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
-## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
-##  MSnbase version: 2.5.9
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-More details about combining data are given in the dedicated
-\textit{Combining MSnSet instances} section of the \Biocpkg{MSnbase}
-\href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-demo.pdf}{tutorial
-  vignette}.
-
-\subsubsection*{Missing data}
-
-Missing data are a recurrent issue in mass spectrometry applications,
-and should be addressed independently of this workflow
-\cite{Webb-Robertson:2015,Lazar:2016}. In \cite{Gatto:2014b}, we have
-described how a high content in missing values in spatial proteomics
-data and their inappropriate handling leads to a reduction of
-sub-cellular resolution. \st{Missing data can be imputated} \hl{We can impute missing data} using
-\Biocpkg{MSnbase}'s \texttt{impute} function. The method underlying
-the imputation method is then determined by a \texttt{methods}
-parameter (see \texttt{?impute} for available options). To impute
-missing values using nearest neighbour imputation, one would
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{hl} \hlkwb{<-} \hlkwd{impute}\hlstd{(hl,} \hlkwc{method} \hlstd{=} \hlstr{"knn"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-In our particular case, missing values are indicative of protein
-groups that were not acquired in both replicates (Figure
-\ref{fig:namap}).
-
-\begin{figure}[!ht]
-  \centering
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{image2}\hlstd{(}\hlkwd{is.na}\hlstd{(combined),} \hlkwc{col} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"black"}\hlstd{,} \hlstr{"white"}\hlstd{),}
-       \hlkwc{main} \hlstd{=} \hlstr{"Missing values (white cells) after combining replicates"}\hlstd{)}
-\end{alltt}
-\end{kframe}
-\includegraphics[width=.65\textwidth]{figure/namap-1}
-
-\end{knitrout}
-\caption{Heatmap of missing values. Note that the features are
-  re-ordered to highlight \hl{clusters} of proteins with similar numbers of
-  missing values.}
-  \label{fig:namap}
-\end{figure}
-
-We prefer to remove proteins that were not assayed in both replicated
-experiments. This is done with the \texttt{filterNA} function that
-removes features (proteins) that contain more than a certain
-proportion (default is 0) missing values. The \textit{Processing
-  information} section summarises how combining and filtering missing
-values (subsetting) changed the dimensions of the data.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{combined} \hlkwb{<-} \hlkwd{filterNA}\hlstd{(combined)}
-\hlstd{combined}
-\end{alltt}
-\begin{verbatim}
-## MSnSet (storageMode: lockedEnvironment)
-## assayData: 5032 features, 20 samples 
-##   element names: exprs 
-## protocolData: none
-## phenoData
-##   sampleNames: X126.1 X127N.1 ... X131.2 (20 total)
-##   varLabels: Replicate TMT.Reagent ... Iodixonal.Density (5 total)
-##   varMetadata: labelDescription
-## featureData
-##   featureNames: Q9JHU4 Q9QXS1-3 ... Q9Z2R6 (5032 total)
-##   fvarLabels: EntryName ProteinDescription ... ProteinCoverage2 (8
-##     total)
-##   fvarMetadata: labelDescription
-## experimentData: use 'experimentData(object)'
-##   pubMedIds: 26754106 
-## Annotation:  
-## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Tue Apr 10 11:36:36 2018 
-## Subset [6725,20][5032,20] Tue Apr 10 11:36:36 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 11:36:36 2018 
-## Dropped featureData's levels Tue Apr 10 11:36:36 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 14:00:41 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 14:00:41 2018 
-## Dropped featureData's levels Tue Apr 10 14:00:41 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:11:22 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:11:22 2018 
-## Dropped featureData's levels Tue Apr 10 16:11:22 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:17:52 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:17:52 2018 
-## Dropped featureData's levels Tue Apr 10 16:17:52 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:25:51 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:25:51 2018 
-## Dropped featureData's levels Tue Apr 10 16:25:51 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:29:26 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:29:26 2018 
-## Dropped featureData's levels Tue Apr 10 16:29:27 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:35:47 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:35:47 2018 
-## Dropped featureData's levels Tue Apr 10 16:35:47 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:47:20 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:47:20 2018 
-## Dropped featureData's levels Tue Apr 10 16:47:20 2018 
-## Subset [5032,20][5032,20] Tue Apr 10 16:52:54 2018 
-## Removed features with more than 0 NAs: Tue Apr 10 16:52:54 2018 
-## Dropped featureData's levels Tue Apr 10 16:52:54 2018 
+## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:03:28 2018 
+## Subset [6725,20][5032,20] Wed Apr 11 14:03:28 2018 
+## Removed features with more than 0 NAs: Wed Apr 11 14:03:28 2018 
+## Dropped featureData's levels Wed Apr 11 14:03:28 2018 
 ##  MSnbase version: 2.5.9
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 10 16:52:56 2018 
+##  Created on Wed Apr 11 14:03:30 2018 
 ##  Description:
@@
 % latex table generated in R 3.5.0 by xtable 1.8-2 package
-% Tue Apr 10 16:52:56 2018
+% Wed Apr 11 14:03:30 2018
 \begin{table}[ht]
@@
 
+\pagebreak
+
 The more \hl{relevent} comparison unit is not a single channel, but rather
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 10 16:52:55 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 10 16:52:56 2018 
+## Added markers from  'mrk' marker vector. Wed Apr 11 14:03:29 2018 
+## Added markers from  'pdres' marker vector. Wed Apr 11 14:03:30 2018 
 ##  MSnbase version: 2.5.10
@@
 \begin{alltt}
-\hlstd{svmf} \hlkwb{<-} \hlkwd{dir}\hlstd{(extdatadir,} \hlkwc{full.names} \hlstd{=} \hlnum{TRUE}\hlstd{,}
-              \hlkwc{pattern} \hlstd{=} \hlstr{"bpw-svmopt.rds"}\hlstd{)}
-\hlstd{params} \hlkwb{<-} \hlkwd{readRDS}\hlstd{(svmf)}
-\end{alltt}
-\end{kframe}
-\end{knitrout}
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
 \hlcom{## 100 rounds of optimisation with five-fold cross-validation}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4323379               0.3037630               0.3822328 
+##               0.4318484               0.3059444               0.3846145 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6930642               0.6014878               0.4426545 
+##               0.6952120               0.6096139               0.4308645 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4224179               0.5978957               0.9497788 
+##               0.4099447               0.5950343               0.9499410 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950203               0.7117761               0.3147679 
+##               0.7954586               0.7096376               0.3119884 
 ##         Plasma membrane              Proteasome 
-##               0.7042806               0.4150910
+##               0.7244522               0.4199965
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4323379               0.3037630               0.3822328 
+##               0.4318484               0.3059444               0.3846145 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6930642               0.6014878               0.4426545 
+##               0.6952120               0.6096139               0.4308645 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4224179               0.5978957               0.9497788 
+##               0.4099447               0.5950343               0.9499410 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7950203               0.7117761               0.3147679 
+##               0.7954586               0.7096376               0.3119884 
 ##         Plasma membrane              Proteasome 
-##               0.7042806               0.4150910
+##               0.7244522               0.4199965
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      84 
+##                      85                     173                      87 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     301                     478                      99 
+##                     298                     472                     103 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     523 
+##                      26                     124                     521 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     341                      38 
+##                     229                     346                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     154                    2057
+##                     317                     155                    2058
 \end{verbatim}

### 2018-04-16 bioc-workflow-resubmit.tex
fix img label

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:26:11 2018 
+## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:33:11 2018 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:26:11 2018 
-## Subset [6725,20][5032,20] Mon Apr 16 15:26:12 2018 
-## Removed features with more than 0 NAs: Mon Apr 16 15:26:12 2018 
-## Dropped featureData's levels Mon Apr 16 15:26:12 2018 
+## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:33:11 2018 
+## Subset [6725,20][5032,20] Mon Apr 16 15:33:12 2018 
+## Removed features with more than 0 NAs: Mon Apr 16 15:33:12 2018 
+## Dropped featureData's levels Mon Apr 16 15:33:12 2018 
 ##  MSnbase version: 1.21.7
@@
   red dots represent the mean relative intensity for each channel. }
-  \label{fig:q  cbx}
+  \label{fig:qcbx}
 \end{figure}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Mon Apr 16 15:26:14 2018 
+##  Created on Mon Apr 16 15:33:14 2018 
 ##  Description:
@@
 % latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Mon Apr 16 15:26:14 2018
+% Mon Apr 16 15:33:14 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Mon Apr 16 15:26:13 2018 
-## Added markers from  'pdres' marker vector. Mon Apr 16 15:26:14 2018 
+## Added markers from  'mrk' marker vector. Mon Apr 16 15:33:13 2018 
+## Added markers from  'pdres' marker vector. Mon Apr 16 15:33:14 2018 
 ##  MSnbase version: 2.4.2
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4355329               0.3041412               0.3939341 
+##               0.4371207               0.3057074               0.3881852 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6838362               0.6082860               0.4344442 
+##               0.6925752               0.6021888               0.4265610 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4147867               0.5994097               0.9498454 
+##               0.4262321               0.5923175               0.9498363 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7961992               0.7100991               0.3171891 
+##               0.7961720               0.7109965               0.3179536 
 ##         Plasma membrane              Proteasome 
-##               0.7093459               0.4225532
+##               0.7201960               0.4170372
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4355329               0.3041412               0.3939341 
+##               0.4371207               0.3057074               0.3881852 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6838362               0.6082860               0.4344442 
+##               0.6925752               0.6021888               0.4265610 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4147867               0.5994097               0.9498454 
+##               0.4262321               0.5923175               0.9498363 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7961992               0.7100991               0.3171891 
+##               0.7961720               0.7109965               0.3179536 
 ##         Plasma membrane              Proteasome 
-##               0.7093459               0.4225532
+##               0.7201960               0.4170372
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      89 
+##                      84                     171                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     299                     476                      92 
+##                     294                     475                      97 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     523 
+##                      29                     124                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     344                      39 
+##                     229                     343                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     322                     156                    2057
+##                     317                     161                    2057
 \end{verbatim}

### 2018-04-17 bioc-workflow-resubmit.Rnw
don't load and attach gridExtra
fix img label
small changes to intro

@@
 
-Quantitative mass spectrometry based spatial proteomics involves
+Quantitative mass spectrometry-based spatial proteomics involves
 elaborate, expensive and time consuming experimental protocols and
@@
 \cite{Gatto:2010} for a review, and
-\cite{hyper,Itzhak:2016,Jean_Beltran:2016} for recent
-examples). However, data analysis is as critical as data production
-for reliable and insightful biological interpretation. Here, we walk
-the reader through a typical pipeline for the analysis of such data
-using several Bioconductor~\cite{Huber:2015} packages for the R
-statistical programming environment.
+\cite{hyper,Itzhak:2016,Jean_Beltran:2016,Itzhak:2017,Hirst:2018} for
+recent examples). However, data analysis is as critical as data
+production for reliable and insightful biological
+interpretation. Here, we walk the reader through a typical pipeline
+for the analysis of such data using several
+Bioconductor~\cite{Huber:2015} packages for the R statistical
+programming environment.
 
@@
 
-Installation of Bioconductor package is documented in detail on the
+Installation of Bioconductor packages is documented in detail on the
 \href{http://bioconductor.org/install/#install-bioconductor-packages}{Bioconductor
@@
 would type \texttt{library("pRoloc")} after installation.
+
+\bigskip
 
@@
 <<visualiseOptHide, echo=FALSE, out.width=".65\\textwidth">>=
-library('gridExtra')
-grid.arrange(plot(params), levelPlot(params), ncol = 2)
+gridExtra::grid.arrange(plot(params), levelPlot(params), ncol = 2)
 @

### 2018-04-17 bioc-workflow-resubmit.tex
small changes to intro

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:33:11 2018 
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 08:39:23 2018 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Mon Apr 16 15:33:11 2018 
-## Subset [6725,20][5032,20] Mon Apr 16 15:33:12 2018 
-## Removed features with more than 0 NAs: Mon Apr 16 15:33:12 2018 
-## Dropped featureData's levels Mon Apr 16 15:33:12 2018 
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 08:39:23 2018 
+## Subset [6725,20][5032,20] Tue Apr 17 08:39:24 2018 
+## Removed features with more than 0 NAs: Tue Apr 17 08:39:24 2018 
+## Dropped featureData's levels Tue Apr 17 08:39:24 2018 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Mon Apr 16 15:33:14 2018 
+##  Created on Tue Apr 17 08:39:26 2018 
 ##  Description:
@@
 % latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Mon Apr 16 15:33:14 2018
+% Tue Apr 17 08:39:26 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Mon Apr 16 15:33:13 2018 
-## Added markers from  'pdres' marker vector. Mon Apr 16 15:33:14 2018 
+## Added markers from  'mrk' marker vector. Tue Apr 17 08:39:25 2018 
+## Added markers from  'pdres' marker vector. Tue Apr 17 08:39:26 2018 
 ##  MSnbase version: 2.4.2
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4371207               0.3057074               0.3881852 
+##               0.4363460               0.3023446               0.3833708 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6925752               0.6021888               0.4265610 
+##               0.6846500               0.5947855               0.4246911 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4262321               0.5923175               0.9498363 
+##               0.4302963               0.5753825               0.9498381 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7961720               0.7109965               0.3179536 
+##               0.7947918               0.7075039               0.3176585 
 ##         Plasma membrane              Proteasome 
-##               0.7201960               0.4170372
+##               0.7124225               0.4149409
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4371207               0.3057074               0.3881852 
+##               0.4363460               0.3023446               0.3833708 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6925752               0.6021888               0.4265610 
+##               0.6846500               0.5947855               0.4246911 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4262321               0.5923175               0.9498363 
+##               0.4302963               0.5753825               0.9498381 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7961720               0.7109965               0.3179536 
+##               0.7947918               0.7075039               0.3176585 
 ##         Plasma membrane              Proteasome 
-##               0.7201960               0.4170372
+##               0.7124225               0.4149409
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      84                     171                      89 
+##                      86                     170                      84 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     294                     475                      97 
+##                     296                     474                      95 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      29                     124                     523 
+##                      28                     123                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     343                      39 
+##                     230                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     317                     161                    2057
+##                     326                     157                    2057
 \end{verbatim}

### 2018-04-17 bioc-workflow-resubmit.Rnw
more changes to pre-processing section; should now also build on travis

@@
 
-To make use of the full functionality of the \Biocpkg{pRoloc} software
-one needs to import their data into R and prepare them as an
+To make use of the full functionality of the \Biocpkg{pRoloc} software,
+users need to import their data into R and prepare them as an
 \texttt{MSnSet}. The \texttt{MSnSet} is a dedicated data structure for
@@
 under \textit{Processing information} in the textual object summary
-when an \texttt{MSnSet}'s name is typed in the R console.
+when an \texttt{MSnSet}'s name is typed in the R console (see example
+below).
 
@@
 \href{http://bioconductor.org/packages/release/bioc/vignettes/MSnbase/inst/doc/MSnbase-io.pdf}{input/output
-  capabilities vignette}. One suggested simple method is to use the
+  capabilities vignette}. The simplest method is to use the
 function \texttt{readMSnSet2}. The function takes a single spreadsheet
@@
 content of that directory and store the path that matches the file
-name of interest in the \texttt{csvfile}. Note that these two lines
+name of interest in the \texttt{csvfile} variable. Note that these two lines
 are only needed here to locate a file in a package; in a more general
@@
 The feature meta-data is stored in the \texttt{fData} slot and can be
-accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2}, automatically,
-everything that is not defined as quantitation data by \texttt{ecol} or the
-feature names by \texttt{fnames} is deposited to the \texttt{fData} slot.
+accessed by \texttt{fData(hl)}. When using \texttt{readMSnSet2}
+everything that is not defined as quantitation data by \texttt{ecol}
+is deposited to the \texttt{fData} slot.
 
@@
 remove the shared annotation. In the first line, we update only the
-feature variable names 3 to 5 (by appending a \texttt{1}) and in the
-second line, we apply the \texttt{updateFvarLabels} function to update
-all feature variable names (by appending a \texttt{2}).
+feature variable names 3 to 5 (by appending a \texttt{1}) of the first
+replicate and in the second line, we apply the
+\texttt{updateFvarLabels} function to update all feature variable
+names (by appending a \texttt{2}) of the second replicate. In lines 3
+and 4, we retain the first 5 feature variables for the first replicate
+and the relevant third to fourth variables for the second replicate.
 
@@
 <<combine, cache = cache>>=
-combined <- MSnbase::combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
+combined <- combine(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)
 combined
@@
 groups that were not acquired in both replicates (Figure
-\ref{fig:namap}).
+\ref{fig:namap}, produced with the \texttt{image2} function).
 
@@
 components ordered according to the amount of variability that they
-describe. The \texttt{plot2D} and \texttt{plot3D} methods in
+describe. The \texttt{plot2D} and \texttt{plot3D} functions in
 \Biocpkg{pRoloc} allows one to plot the principal components (PCs) of
-a dataset against one another, by default the first two components are
+a dataset against one another. By default, the first two components are
 plotted on the x- and y-axis for the \texttt{plot2D} function, and
@@
 functions, that plot the protein profiles occupancy along the gradient
-(we also display the mean channel intensities) and a \texttt{boxplot}
-of the column intensities.  In the two plots displayed on figure
-\ref{fig:qcbx}, we re-order the TMT channels to pair corresponding
-channels in the two replicates (rather than ordering the channels by
-replicate).
+(we also display the mean channel intensities below) and a
+\texttt{boxplot} of the column intensities.  In the two plots
+displayed on figure \ref{fig:qcbx}, we re-order the TMT channels to
+pair corresponding channels in the two replicates (rather than
+ordering the channels by replicate).
 
@@
 \textit{and} condition of interest. Applying this to machine learning
-(ML), and specifically supervised learning, for the task of protein
+(ML), and specifically supervised learning for the task of protein
 localisation prediction, these markers constitute the labelled

### 2018-04-17 bioc-workflow-resubmit.tex
more changes to pre-processing section; should now also build on travis

@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.4.2
+##  MSnbase version: 2.5.12
 \end{verbatim}
@@
 
-will list the 62 datasets that are
+will list the 75 datasets that are
 available in \Biocexptpkg{pRolocdata}.
@@
 \begin{alltt}
-\hlstd{combined} \hlkwb{<-} \hlstd{MSnbase}\hlopt{::}\hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
+\hlstd{combined} \hlkwb{<-} \hlkwd{combine}\hlstd{(hyperLOPIT2015ms3r1, hyperLOPIT2015ms3r2)}
 \hlstd{combined}
@@
 ## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 08:39:23 2018 
-##  MSnbase version: 1.21.7
+## Combined [5489,10] and [6268,10] MSnSets Tue Apr 17 11:44:16 2018 
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 ## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 08:39:23 2018 
-## Subset [6725,20][5032,20] Tue Apr 17 08:39:24 2018 
-## Removed features with more than 0 NAs: Tue Apr 17 08:39:24 2018 
-## Dropped featureData's levels Tue Apr 17 08:39:24 2018 
-##  MSnbase version: 1.21.7
+## Combined [5489,10] and [6268,10] MSnSets Tue Apr 17 11:44:16 2018 
+## Subset [6725,20][5032,20] Tue Apr 17 11:44:17 2018 
+## Removed features with more than 0 NAs: Tue Apr 17 11:44:17 2018 
+## Dropped featureData's levels Tue Apr 17 11:44:17 2018 
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 17 08:39:26 2018 
+##  Created on Tue Apr 17 11:44:19 2018 
 ##  Description:
@@
 
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Apr 17 08:39:26 2018
+% latex table generated in R 3.6.0 by xtable 1.8-2 package
+% Tue Apr 17 11:44:19 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 17 08:39:25 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 17 08:39:26 2018 
-##  MSnbase version: 2.4.2
+## Added markers from  'mrk' marker vector. Tue Apr 17 11:44:18 2018 
+## Added markers from  'pdres' marker vector. Tue Apr 17 11:44:19 2018 
+##  MSnbase version: 2.5.12
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4363460               0.3023446               0.3833708 
+##               0.4368645               0.3031282               0.3889317 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6846500               0.5947855               0.4246911 
+##               0.6760609               0.6075289               0.4353435 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4302963               0.5753825               0.9498381 
+##               0.4221860               0.5895546               0.9499026 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7947918               0.7075039               0.3176585 
+##               0.7952921               0.7096399               0.3106030 
 ##         Plasma membrane              Proteasome 
-##               0.7124225               0.4149409
+##               0.7232975               0.4205260
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4363460               0.3023446               0.3833708 
+##               0.4368645               0.3031282               0.3889317 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6846500               0.5947855               0.4246911 
+##               0.6760609               0.6075289               0.4353435 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4302963               0.5753825               0.9498381 
+##               0.4221860               0.5895546               0.9499026 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7947918               0.7075039               0.3176585 
+##               0.7952921               0.7096399               0.3106030 
 ##         Plasma membrane              Proteasome 
-##               0.7124225               0.4149409
+##               0.7232975               0.4205260
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     170                      84 
+##                      85                     171                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     474                      95 
+##                     299                     477                      99 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      28                     123                     523 
+##                      27                     123                     521 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      39 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     326                     157                    2057
+##                     317                     156                    2057
 \end{verbatim}
@@
 \begin{verbatim}
-## R version 3.4.4 Patched (2018-03-19 r74516)
+## R Under development (unstable) (2018-04-02 r74505)
 ## Platform: x86_64-pc-linux-gnu (64-bit)
@@
 ## attached base packages:
-## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
-## [8] datasets  base     
+## [1] stats4    parallel  stats     graphics  grDevices utils     datasets 
+## [8] methods   base     
 ## 
 ## other attached packages:
-##  [1] pRolocdata_1.16.0    pRoloc_1.18.0        MLInterfaces_1.58.1 
-##  [4] cluster_2.0.7-1      annotate_1.56.2      XML_3.98-1.10       
-##  [7] AnnotationDbi_1.40.0 IRanges_2.12.0       S4Vectors_0.16.0    
-## [10] MSnbase_2.4.2        ProtGenerics_1.10.0  BiocParallel_1.12.0 
-## [13] mzR_2.12.0           Rcpp_0.12.16         Biobase_2.38.0      
-## [16] BiocGenerics_0.24.0  xtable_1.8-2         BiocStyle_2.6.1     
+##  [1] pRolocdata_1.17.3    pRoloc_1.19.4        MLInterfaces_1.59.1 
+##  [4] cluster_2.0.7        annotate_1.57.3      XML_3.98-1.10       
+##  [7] AnnotationDbi_1.41.4 IRanges_2.13.28      S4Vectors_0.17.41   
+## [10] MSnbase_2.5.12       ProtGenerics_1.11.0  BiocParallel_1.13.3 
+## [13] mzR_2.13.6           Rcpp_0.12.16         Biobase_2.39.2      
+## [16] BiocGenerics_0.25.3  xtable_1.8-2         BiocStyle_2.7.8     
 ## [19] knitr_1.20          
@@
 ##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
-##   [4] lazyeval_0.2.1        splines_3.4.4         ggvis_0.4.3          
+##   [4] lazyeval_0.2.1        splines_3.6.0         ggvis_0.4.3          
 ##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
-##  [10] foreach_1.4.4         BiocInstaller_1.28.0  htmltools_0.3.6      
+##  [10] foreach_1.4.4         BiocInstaller_1.29.6  htmltools_0.3.6      
 ##  [13] viridis_0.5.1         gdata_2.18.0          magrittr_1.5         
 ##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
-##  [19] limma_3.34.9          recipes_0.1.2         gower_0.1.2          
+##  [19] limma_3.35.14         recipes_0.1.2         gower_0.1.2          
 ##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
 ##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.1           
-##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
-##  [31] genefilter_1.60.0     bindr_0.1.1           impute_1.52.0        
+##  [28] dplyr_0.7.4.9003      RCurl_1.95-4.10       hexbin_1.27.2        
+##  [31] genefilter_1.61.1     bindr_0.1.1           impute_1.53.0        
 ##  [34] DRR_0.0.3             survival_2.41-3       iterators_1.0.9      
 ##  [37] glue_1.2.0            gtable_0.2.0          ipred_0.9-6          
-##  [40] zlibbioc_1.24.0       kernlab_0.9-25        ddalpha_1.3.2        
-##  [43] prabclus_2.2-6        DEoptimR_1.0-8        abind_1.4-5          
-##  [46] scales_0.5.0.9000     vsn_3.46.0            mvtnorm_1.0-7        
-##  [49] DBI_0.8               viridisLite_0.3.0     progress_1.1.2       
-##  [52] magic_1.5-8           foreign_0.8-69        bit_1.1-12           
-##  [55] proxy_0.4-22          mclust_5.4            preprocessCore_1.40.0
-##  [58] lava_1.6.1            prodlim_1.6.1         sampling_2.8         
-##  [61] htmlwidgets_1.0       httr_1.3.1            threejs_0.3.1        
-##  [64] FNN_1.1               RColorBrewer_1.1-2    fpc_2.1-11           
-##  [67] modeltools_0.2-21     pkgconfig_2.0.1       flexmix_2.3-14       
-##  [70] nnet_7.3-12           caret_6.0-79          tidyselect_0.2.4     
-##  [73] rlang_0.2.0           reshape2_1.4.3        munsell_0.4.3        
-##  [76] mlbench_2.1-1         tools_3.4.4           msdata_0.18.0        
-##  [79] RSQLite_2.1.0         pls_2.6-0             broom_0.4.4          
-##  [82] evaluate_0.10.1       geometry_0.3-6        stringr_1.3.0        
-##  [85] mzID_1.16.0           yaml_2.1.18           ModelMetrics_1.1.0   
-##  [88] bit64_0.9-7           robustbase_0.92-8     randomForest_4.6-14  
-##  [91] purrr_0.2.4           dendextend_1.7.0      bindrcpp_0.2.2       
-##  [94] nlme_3.1-137          whisker_0.3-2         mime_0.5             
-##  [97] RcppRoll_0.2.2        biomaRt_2.34.2        compiler_3.4.4       
-## [100] e1071_1.6-8          
-##  [ reached getOption("max.print") -- omitted 37 entries ]
+##  [40] zlibbioc_1.25.0       kernlab_0.9-25        ddalpha_1.3.1.1      
+##  [43] prabclus_2.2-6        DEoptimR_1.0-8        scales_0.5.0         
+##  [46] vsn_3.47.1            mvtnorm_1.0-7         DBI_0.8              
+##  [49] viridisLite_0.3.0     progress_1.1.2        foreign_0.8-70       
+##  [52] bit_1.1-12            proxy_0.4-22          mclust_5.4           
+##  [55] preprocessCore_1.41.0 lava_1.6.1            prodlim_1.6.1        
+##  [58] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
+##  [61] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
+##  [64] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
+##  [67] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
+##  [70] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
+##  [73] munsell_0.4.3         mlbench_2.1-1         tools_3.6.0          
+##  [76] msdata_0.19.3         RSQLite_2.1.0         pls_2.6-0            
+##  [79] broom_0.4.4           evaluate_0.10.1       stringr_1.3.0        
+##  [82] mzID_1.17.0           yaml_2.1.18           ModelMetrics_1.1.0   
+##  [85] bit64_0.9-7           robustbase_0.92-8     randomForest_4.6-14  
+##  [88] purrr_0.2.4           dendextend_1.7.0      bindrcpp_0.2.2.9000  
+##  [91] nlme_3.1-136          whisker_0.3-2         mime_0.5             
+##  [94] RcppRoll_0.2.2        biomaRt_2.35.13       compiler_3.6.0       
+##  [97] e1071_1.6-8           affyio_1.49.2         tibble_1.4.2         
+## [100] stringi_1.1.7        
+##  [ reached getOption("max.print") -- omitted 33 entries ]
 \end{verbatim}

### 2018-04-17 bioc-workflow-resubmit.Rnw
changes until sml, with travis condif update

@@
 \href{http://bioconductor.org/help/workflows/annotation/Annotation_Resources/}{annotation
-  resouces} or any conversion softwares available online.
+  resources} or any conversion softwares available online.
 
@@
 
-Finally, in addition to \texttt{plot2D}, the \texttt{plot3D} function
-allows to interactively explore a 3-dimensional plot of the data.
-
 \pagebreak
@@
 Q8C156, which are some of the proteins known to form part of the
-13S condensin complex, we would call the following code:
+13S condensin complex, we would call the code displayed
+on Figure \ref{fig:foi}. Users can also create several sets of \texttt{FeaturesOfInterest}
+object and store them in a \texttt{FoICollection}.
 
@@
 
-Users can also create several sets of \texttt{FeaturesOfInterest}
-object and store them in a \texttt{FoICollection}.
+\bigskip
 
@@
 follow the advice in \cite{Trotter:2010} and combine replicated spatial
-proteomics experiments as described above. Indeed, Trotter et
-al. have shown a significant improvement in protein–organelle
+proteomics experiments as described above. Indeed, Trotter \textit{et
+  al.} have shown a significant improvement in protein–organelle
 association upon direct combination of single experiments, in
@@
 application part of the \Biocpkg{pRolocGUI} package is also useful for
-examining replicate experiments (see the next section interactive
-visualisation for details).
+examining replicate experiments (see the next section
+\textit{Interactive visualisation} for details).
 
@@
 analyses allowing one to shed light on data structure and patterns of
-interest. Using the \Biocpkg{pRolocGUI} package we can interactively
+interest. Using the \Biocpkg{pRolocGUI} package, we can interactively
 visualise, explore and interrogate quantitative spatial proteomics
-data. The \Biocpkg{pRolocGUI} package is currently under active
-development and it relies on the \texttt{shiny} framework for
-reactivity and interactivity. The package currently distributes 3
-different GUI's (\textit{main} (default), \textit{compare} or
-\textit{classify}) which are wrapped and launched by the
-\texttt{pRolocVis} function.
+data. The \Biocpkg{pRolocGUI} package relies on the \texttt{shiny}
+framework for reactivity and interactivity. It distributes 3 different
+GUI's (\textit{main} (default), \textit{compare} or \textit{classify})
+which are wrapped and launched by the \texttt{pRolocVis} function.
 
@@
 thresholds for the assignment of proteins to its sub-cellular
-location, as discussed later in the subsection \textit{thresholding}
-in the \textit{supervised machine learning section}.
+location, as discussed later in the subsection \textit{Thresholding}
+in the \textit{Supervised machine learning} section.
 
@@
 are also useful for confirming the presence of known or postulated
-clusters in an unbiased fashion. For example, in \cite{hyper} the
+clusters in an unbiased fashion. For example in \cite{hyper}, the
 \texttt{phenoDisco} algorithm was used to confirm the data-specific presence
@@
 
-As the \texttt{phenoDisco} algorithm is semi-supervised it uses both labelled
-(markers) and unlabelled data to explore the data structure and find
-new sub-cellular data clusters. Thus the first step is to define some
-input labelled data i.e. markers, that the algorithm will use as input
-for the supervised learning aspect of the algorithm. As described in
-\cite{hyper} we define a set of markers to use as input for the analyses
-that cover well-known residents from three distinct organelle
-structures; the mitochondria, plasma membrane and ER, and from three
-well-known and abundant protein complexes; the proteasome and two
-ribosomal subunits, 40S and 60S.  These input markers are stored in
-the \texttt{featureData} column of \texttt{hl} where \texttt{fcol = "phenoDisco.Input"}. We
-can use the convenience accessor function \texttt{getMarkers} to print out a
-table of the markers contained in this marker set. These initial
+As the \texttt{phenoDisco} algorithm is semi-supervised it uses both
+labelled (markers) and unlabelled data to explore the data structure
+and find new sub-cellular data clusters. Thus the first step is to
+define some input labelled data i.e. markers, that the algorithm will
+use as input for the supervised learning aspect of the algorithm. As
+described in \cite{hyper} we define a set of markers to use as input
+for the analyses that cover well-known residents from three distinct
+organelle structures; the mitochondria, plasma membrane and ER, and
+from three well-known and abundant protein complexes; the proteasome
+and two ribosomal subunits, 40S and 60S.  These input markers are
+stored in the \texttt{phenoDisco.Input} \texttt{featureData} column of
+\texttt{hl} and below set by \texttt{fcol = "phenoDisco.Input"}. We
+can use the convenience accessor function \texttt{getMarkers} to print
+out a table of the markers contained in this marker set. These initial
 markers were manually curated using information from the UniProt
@@
 example, if a user is interested in the detection of small complexes
-they may wish to use a small \texttt{GS = 10}, or \texttt{= 20} etc.
+they may wish to use a small \texttt{GS = 10}, or \texttt{GS = 20} etc.
 If they wish to detect larger, more abundant sub-cellular niches a

### 2018-04-17 bioc-workflow-resubmit.tex
changes until sml, with travis condif update

@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.5.12
+##  MSnbase version: 2.4.2
 \end{verbatim}
@@
 ## experimentData: use 'experimentData(object)'
-##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Tue Apr 17 11:44:16 2018 
-##  MSnbase version: 2.5.9
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 13:59:46 2018 
+##  MSnbase version: 1.21.7
 \end{verbatim}
@@
 ## experimentData: use 'experimentData(object)'
-##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Tue Apr 17 11:44:16 2018 
-## Subset [6725,20][5032,20] Tue Apr 17 11:44:17 2018 
-## Removed features with more than 0 NAs: Tue Apr 17 11:44:17 2018 
-## Dropped featureData's levels Tue Apr 17 11:44:17 2018 
-##  MSnbase version: 2.5.9
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 13:59:46 2018 
+## Subset [6725,20][5032,20] Tue Apr 17 13:59:47 2018 
+## Removed features with more than 0 NAs: Tue Apr 17 13:59:47 2018 
+## Dropped featureData's levels Tue Apr 17 13:59:47 2018 
+##  MSnbase version: 1.21.7
 \end{verbatim}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 17 11:44:19 2018 
+##  Created on Tue Apr 17 13:59:49 2018 
 ##  Description:
@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-2 package
-% Tue Apr 17 11:44:19 2018
+% latex table generated in R 3.4.4 by xtable 1.8-2 package
+% Tue Apr 17 13:59:49 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 17 11:44:18 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 17 11:44:19 2018 
-##  MSnbase version: 2.5.12
+## Added markers from  'mrk' marker vector. Tue Apr 17 13:59:48 2018 
+## Added markers from  'pdres' marker vector. Tue Apr 17 13:59:49 2018 
+##  MSnbase version: 2.4.2
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4368645               0.3031282               0.3889317 
+##               0.4340512               0.3058889               0.3932999 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6760609               0.6075289               0.4353435 
+##               0.6866550               0.6085893               0.4267912 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4221860               0.5895546               0.9499026 
+##               0.4088995               0.5917139               0.9495218 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7952921               0.7096399               0.3106030 
+##               0.7946689               0.7107085               0.3134403 
 ##         Plasma membrane              Proteasome 
-##               0.7232975               0.4205260
+##               0.7229096               0.4193182
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4368645               0.3031282               0.3889317 
+##               0.4340512               0.3058889               0.3932999 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6760609               0.6075289               0.4353435 
+##               0.6866550               0.6085893               0.4267912 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4221860               0.5895546               0.9499026 
+##               0.4088995               0.5917139               0.9495218 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7952921               0.7096399               0.3106030 
+##               0.7946689               0.7107085               0.3134403 
 ##         Plasma membrane              Proteasome 
-##               0.7232975               0.4205260
+##               0.7229096               0.4193182
 \end{verbatim}
@@
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     299                     477                      99 
+##                     295                     477                      96 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     521 
+##                      27                     125                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     229                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     317                     156                    2057
+##                     317                     159                    2058
 \end{verbatim}
@@
 \begin{verbatim}
-## R Under development (unstable) (2018-04-02 r74505)
+## R version 3.4.4 Patched (2018-03-19 r74516)
 ## Platform: x86_64-pc-linux-gnu (64-bit)
@@
 ## attached base packages:
-## [1] stats4    parallel  stats     graphics  grDevices utils     datasets 
-## [8] methods   base     
+## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
+## [8] datasets  base     
 ## 
 ## other attached packages:
-##  [1] pRolocdata_1.17.3    pRoloc_1.19.4        MLInterfaces_1.59.1 
-##  [4] cluster_2.0.7        annotate_1.57.3      XML_3.98-1.10       
-##  [7] AnnotationDbi_1.41.4 IRanges_2.13.28      S4Vectors_0.17.41   
-## [10] MSnbase_2.5.12       ProtGenerics_1.11.0  BiocParallel_1.13.3 
-## [13] mzR_2.13.6           Rcpp_0.12.16         Biobase_2.39.2      
-## [16] BiocGenerics_0.25.3  xtable_1.8-2         BiocStyle_2.7.8     
+##  [1] pRolocdata_1.16.0    pRoloc_1.18.0        MLInterfaces_1.58.1 
+##  [4] cluster_2.0.7-1      annotate_1.56.2      XML_3.98-1.10       
+##  [7] AnnotationDbi_1.40.0 IRanges_2.12.0       S4Vectors_0.16.0    
+## [10] MSnbase_2.4.2        ProtGenerics_1.10.0  BiocParallel_1.12.0 
+## [13] mzR_2.12.0           Rcpp_0.12.16         Biobase_2.38.0      
+## [16] BiocGenerics_0.24.0  xtable_1.8-2         BiocStyle_2.6.1     
 ## [19] knitr_1.20          
@@
 ##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
-##   [4] lazyeval_0.2.1        splines_3.6.0         ggvis_0.4.3          
+##   [4] lazyeval_0.2.1        splines_3.4.4         ggvis_0.4.3          
 ##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
-##  [10] foreach_1.4.4         BiocInstaller_1.29.6  htmltools_0.3.6      
+##  [10] foreach_1.4.4         BiocInstaller_1.28.0  htmltools_0.3.6      
 ##  [13] viridis_0.5.1         gdata_2.18.0          magrittr_1.5         
 ##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
-##  [19] limma_3.35.14         recipes_0.1.2         gower_0.1.2          
+##  [19] limma_3.34.9          recipes_0.1.2         gower_0.1.2          
 ##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
 ##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.1           
-##  [28] dplyr_0.7.4.9003      RCurl_1.95-4.10       hexbin_1.27.2        
-##  [31] genefilter_1.61.1     bindr_0.1.1           impute_1.53.0        
+##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
+##  [31] genefilter_1.60.0     bindr_0.1.1           impute_1.52.0        
 ##  [34] DRR_0.0.3             survival_2.41-3       iterators_1.0.9      
 ##  [37] glue_1.2.0            gtable_0.2.0          ipred_0.9-6          
-##  [40] zlibbioc_1.25.0       kernlab_0.9-25        ddalpha_1.3.1.1      
-##  [43] prabclus_2.2-6        DEoptimR_1.0-8        scales_0.5.0         
-##  [46] vsn_3.47.1            mvtnorm_1.0-7         DBI_0.8              
-##  [49] viridisLite_0.3.0     progress_1.1.2        foreign_0.8-70       
-##  [52] bit_1.1-12            proxy_0.4-22          mclust_5.4           
-##  [55] preprocessCore_1.41.0 lava_1.6.1            prodlim_1.6.1        
-##  [58] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
-##  [61] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
-##  [64] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
-##  [67] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
-##  [70] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
-##  [73] munsell_0.4.3         mlbench_2.1-1         tools_3.6.0          
-##  [76] msdata_0.19.3         RSQLite_2.1.0         pls_2.6-0            
-##  [79] broom_0.4.4           evaluate_0.10.1       stringr_1.3.0        
-##  [82] mzID_1.17.0           yaml_2.1.18           ModelMetrics_1.1.0   
-##  [85] bit64_0.9-7           robustbase_0.92-8     randomForest_4.6-14  
-##  [88] purrr_0.2.4           dendextend_1.7.0      bindrcpp_0.2.2.9000  
-##  [91] nlme_3.1-136          whisker_0.3-2         mime_0.5             
-##  [94] RcppRoll_0.2.2        biomaRt_2.35.13       compiler_3.6.0       
-##  [97] e1071_1.6-8           affyio_1.49.2         tibble_1.4.2         
-## [100] stringi_1.1.7        
-##  [ reached getOption("max.print") -- omitted 33 entries ]
+##  [40] zlibbioc_1.24.0       kernlab_0.9-25        ddalpha_1.3.2        
+##  [43] prabclus_2.2-6        DEoptimR_1.0-8        abind_1.4-5          
+##  [46] scales_0.5.0.9000     vsn_3.46.0            mvtnorm_1.0-7        
+##  [49] DBI_0.8               viridisLite_0.3.0     progress_1.1.2       
+##  [52] magic_1.5-8           foreign_0.8-69        bit_1.1-12           
+##  [55] proxy_0.4-22          mclust_5.4            preprocessCore_1.40.0
+##  [58] lava_1.6.1            prodlim_1.6.1         sampling_2.8         
+##  [61] htmlwidgets_1.0       httr_1.3.1            threejs_0.3.1        
+##  [64] FNN_1.1               RColorBrewer_1.1-2    fpc_2.1-11           
+##  [67] modeltools_0.2-21     pkgconfig_2.0.1       flexmix_2.3-14       
+##  [70] nnet_7.3-12           caret_6.0-79          tidyselect_0.2.4     
+##  [73] rlang_0.2.0           reshape2_1.4.3        munsell_0.4.3        
+##  [76] mlbench_2.1-1         tools_3.4.4           msdata_0.18.0        
+##  [79] RSQLite_2.1.0         pls_2.6-0             broom_0.4.4          
+##  [82] evaluate_0.10.1       geometry_0.3-6        stringr_1.3.0        
+##  [85] mzID_1.16.0           yaml_2.1.18           ModelMetrics_1.1.0   
+##  [88] bit64_0.9-7           robustbase_0.92-8     randomForest_4.6-14  
+##  [91] purrr_0.2.4           dendextend_1.7.0      bindrcpp_0.2.2       
+##  [94] nlme_3.1-137          whisker_0.3-2         mime_0.5             
+##  [97] RcppRoll_0.2.2        biomaRt_2.34.2        compiler_3.4.4       
+## [100] e1071_1.6-8          
+##  [ reached getOption("max.print") -- omitted 37 entries ]
 \end{verbatim}

### 2018-04-17 bioc-workflow-resubmit.Rnw
modifications up to transfer learning

@@
 
-In the code chunk below we then pass these weights
-to the \texttt{svmOptimisation} function. Once again, we provide
-the optimisation results for users to load directly if they wish to save
-computational time, please see the appendix to load these results.
+In the code chunk below we then pass these weights to the
+\texttt{svmOptimisation} function. Once again, we provide the
+optimisation results for users to load directly if they wish to save
+computational time (see the appendix for details).
 
@@
 Additionally, when calling \texttt{plot2D} we can use the \texttt{cex}
-argument to change the size of each point on the plot to be inversely
-proportional to the SVM score (Figure \ref{fig:plotSVM}). This gives an initial overview of the
-high scoring localisations from the SVM predictions.
+argument to change the size of each point on the plot to be
+proportional to the SVM score (Figure \ref{fig:plotSVM}). This gives
+an initial overview of the high scoring localisations from the SVM
+predictions.
 
@@
 <<plotSVM, out.width=".8\\textwidth">>=
-## set point size of each protein to be inversely proportional to the
+## set point size of each protein to be proportional to the svm score
 ptsze <- exp(fData(hl)$svm.scores) - 1
@@
 The adjustment of the point size intuitively confers important
-information that is more difficult to define formally (we
+information that is more difficult to define formally (that we
 will address in the next section). The classifier (SVM in our case, but
@@
 labelled training data. We can not guarantee (and do not expect) that
-the whole sub-cellular diversity to be represented in the labelled
+the whole sub-cellular diversity is represented in the labelled
 training data as (1) finding markers that represent the whole
@@
 
+\bigskip
+
 There is also a dedicated interactive application to help users
@@
 then the sliders for defining user-specified scores become active and
-the scores and highlighted on the boxplot by blue dots. For more
+the scores are highlighted on the boxplot by blue dots. For more
 information we refer users to the \Biocpkg{pRolocGUI} tutorial

### 2018-04-17 bioc-workflow-resubmit.tex
modifications up to transfer learning

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 13:59:46 2018 
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:22:42 2018 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 13:59:46 2018 
-## Subset [6725,20][5032,20] Tue Apr 17 13:59:47 2018 
-## Removed features with more than 0 NAs: Tue Apr 17 13:59:47 2018 
-## Dropped featureData's levels Tue Apr 17 13:59:47 2018 
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:22:42 2018 
+## Subset [6725,20][5032,20] Tue Apr 17 15:22:43 2018 
+## Removed features with more than 0 NAs: Tue Apr 17 15:22:43 2018 
+## Dropped featureData's levels Tue Apr 17 15:22:43 2018 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 17 13:59:49 2018 
+##  Created on Tue Apr 17 15:22:45 2018 
 ##  Description:
@@
 % latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Apr 17 13:59:49 2018
+% Tue Apr 17 15:22:45 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 17 13:59:48 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 17 13:59:49 2018 
+## Added markers from  'mrk' marker vector. Tue Apr 17 15:22:45 2018 
+## Added markers from  'pdres' marker vector. Tue Apr 17 15:22:46 2018 
 ##  MSnbase version: 2.4.2
@@
 \begin{alltt}
-\hlcom{## set point size of each protein to be inversely proportional to the}
+\hlcom{## set point size of each protein to be proportional to the svm score}
 \hlstd{ptsze} \hlkwb{<-} \hlkwd{exp}\hlstd{(}\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{svm.scores)} \hlopt{-} \hlnum{1}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4340512               0.3058889               0.3932999 
+##               0.4323939               0.3044486               0.3845760 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6866550               0.6085893               0.4267912 
+##               0.6850826               0.6012822               0.4245089 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4088995               0.5917139               0.9495218 
+##               0.4256415               0.5944635               0.9493494 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946689               0.7107085               0.3134403 
+##               0.7949812               0.7104192               0.3114933 
 ##         Plasma membrane              Proteasome 
-##               0.7229096               0.4193182
+##               0.7229559               0.4157514
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4340512               0.3058889               0.3932999 
+##               0.4323939               0.3044486               0.3845760 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6866550               0.6085893               0.4267912 
+##               0.6850826               0.6012822               0.4245089 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4088995               0.5917139               0.9495218 
+##               0.4256415               0.5944635               0.9493494 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7946689               0.7107085               0.3134403 
+##               0.7949812               0.7104192               0.3114933 
 ##         Plasma membrane              Proteasome 
-##               0.7229096               0.4193182
+##               0.7229559               0.4157514
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     171                      89 
+##                      85                     170                      88 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     295                     477                      96 
+##                     296                     475                     101 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     125                     523 
+##                      27                     124                     524 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     343                      38 
+##                     230                     343                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     317                     159                    2058
+##                     316                     158                    2057
 \end{verbatim}
@@
 \end{figure}
+
+\bigskip
 

### 2018-04-17 bioc-workflow-resubmit.Rnw
update unsup ml

@@
 fData(hl)$pd <- as.character(fData(hl)$pd)
-write.exprs(hl, fDataCols = "pd", file = "pd-results.csv", sep = ",")
+write.exprs(hl, fcol = "pd", file = "pd-results.csv", sep = ",")
 @
@@
 where the data can be projected from a high-dimensional space down to
-two or three dimensions for the purpose of visualisation.
+two or three dimensions.
 
@@
 application uses hierarchical clustering to summarise the relation
-between marker proteins using he \texttt{mrkHClust} function, where
+between marker proteins using the \texttt{mrkHClust} function, where
 the euclidean distance between average class-specific profiles is used
@@
 
+\bigskip
+
 The source of this document, including the code necessary to reproduce

### 2018-04-17 bioc-workflow-resubmit.tex
update unsup ml

@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:22:42 2018 
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:40:33 2018 
 ##  MSnbase version: 1.21.7
@@
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:22:42 2018 
-## Subset [6725,20][5032,20] Tue Apr 17 15:22:43 2018 
-## Removed features with more than 0 NAs: Tue Apr 17 15:22:43 2018 
-## Dropped featureData's levels Tue Apr 17 15:22:43 2018 
+## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:40:33 2018 
+## Subset [6725,20][5032,20] Tue Apr 17 15:40:34 2018 
+## Removed features with more than 0 NAs: Tue Apr 17 15:40:34 2018 
+## Dropped featureData's levels Tue Apr 17 15:40:34 2018 
 ##  MSnbase version: 1.21.7
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 17 15:22:45 2018 
+##  Created on Tue Apr 17 15:40:38 2018 
 ##  Description:
@@
 % latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Apr 17 15:22:45 2018
+% Tue Apr 17 15:40:38 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 17 15:22:45 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 17 15:22:46 2018 
+## Added markers from  'mrk' marker vector. Tue Apr 17 15:40:37 2018 
+## Added markers from  'pdres' marker vector. Tue Apr 17 15:40:39 2018 
 ##  MSnbase version: 2.4.2
@@
 \hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{pd} \hlkwb{<-} \hlkwd{as.character}\hlstd{(}\hlkwd{fData}\hlstd{(hl)}\hlopt{$}\hlstd{pd)}
-\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{fDataCols} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{file} \hlstd{=} \hlstr{"pd-results.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{)}
+\hlkwd{write.exprs}\hlstd{(hl,} \hlkwc{fcol} \hlstd{=} \hlstr{"pd"}\hlstd{,} \hlkwc{file} \hlstd{=} \hlstr{"pd-results.csv"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{","}\hlstd{)}
 \end{alltt}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4323939               0.3044486               0.3845760 
+##               0.4382130               0.3040042               0.3797701 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6850826               0.6012822               0.4245089 
+##               0.6897383               0.6007307               0.4284784 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4256415               0.5944635               0.9493494 
+##               0.4200181               0.5879109               0.9502066 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7949812               0.7104192               0.3114933 
+##               0.7951514               0.7100151               0.3166095 
 ##         Plasma membrane              Proteasome 
-##               0.7229559               0.4157514
+##               0.7050611               0.4169158
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4323939               0.3044486               0.3845760 
+##               0.4382130               0.3040042               0.3797701 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6850826               0.6012822               0.4245089 
+##               0.6897383               0.6007307               0.4284784 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4256415               0.5944635               0.9493494 
+##               0.4200181               0.5879109               0.9502066 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7949812               0.7104192               0.3114933 
+##               0.7951514               0.7100151               0.3166095 
 ##         Plasma membrane              Proteasome 
-##               0.7229559               0.4157514
+##               0.7050611               0.4169158
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     170                      88 
+##                      86                     169                      85 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     475                     101 
+##                     296                     477                      94 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     124                     524 
+##                      27                     123                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     343                      38 
+##                     230                     344                      39 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     316                     158                    2057
+##                     325                     157                    2058
 \end{verbatim}

### 2018-05-16 replies-to-reviewers.md
add replies to reviewers

# A Bioconductor workflow for processing and analysing spatial proteomics data  - Reply to reviewers

## Reviewer 1 - Daniel J. Stekhoven1

> Breckels et al. have written a very nice piece on analysing
> appropriate proteomics data for subcellular localisation. I
> particularly like the "workshop characteristics" of the text. Which
> allows a novice, but interested reader to work through the analysis
> stepwise and reproduce the results described therein. The authors
> took great care in keeping this ideal up during their text and this
> is also where I have put my greatest reservation to the manuscript
> in its present form - since a reader cannot work through the code
> presented in the manuscript, since there at at least two situations
> where a readily available HPC and quite some time is required. This
> kind of leaves a dent in my impression - however, given this can be
> resolved as well as some typos - the workflow report is superb.

We thank reviewer 1 for their comments. Please find our responses to
these inset below.

### Major comments

> Next to reducing the dimensions of data for visualisation, PCA also
> offers a way to understand how the variability is distributed across
> the multidimensional data by providing linear combinations of the
> variables which then constitute the actual PCs. On that note it would
> be nice to mention this in Visualising markers section on page 16,
> where PC7 explains not much variability but due to the correct
> weighing of the variables we do get a separation between
> mitochondrial and peroxisome. This then can be further motivated with
> Figure 9 - where we probably can see that the weights for the
> fractions where the two localisations differ are larger than
> otherwise.

We have added a paragraph to the 'Visualising markers' section of the
manuscript reiterating the purpose of PCA and motivating the choice of
looking at PC's 1 and 7. Figure 9 now follows on from this (now Figure
8), along with the corresponding code and an explanation of the
`plotDist` function.

> I was unable to reproduce Figure 13 comparing the two MSnSets. While
> I was able to look at each set separately using
> `pRolocVis(hllst@x[[I]])`, where i is 1 or 2, I only got an error
> using the code from the manuscript. When using `remap=FALSE` it
> actually works, but since this makes barely sense it is of no use -
> but just as a hint at debugging it.

We can not reproduce this error. Have you updated to the latest
version of R and the latest version of `pRolocGUI`? If you still get
this error message could you please post this as an issue
(https://github.com/ComputationalProteomicsUnit/pRolocGUI/issues) on
the `pRolocGUI` Github page along with your `sessionInfo()` and we
will certainly attempt to solve this.

> You really need to make the results from the phenoDisco
> classification available too. It is super disappointing that one
> cannot continue reproducing the code from page 23 on, because it
> takes 24 hours to compute it using 40 cores?

The results are already available as a `RDS` file and stored in
`pRolocdata` for users. This is what is called in the manuscript under
the hood:

```
f0 <- dir(extdatadir, full.names = TRUE, pattern = "bpw-pdres.rds")
pdres <- readRDS(f0)
hl <- addMarkers(hl, pdres, mcol = "pd", verbose = FALSE)
```

We have made this code available in the manuscript in an appendix so
users can continue to produce the exact plots as they see in this
workflow.

> The above comment is of course also true for the KNN TL Optimisation
> on page 33 - this needs to be downloadable, since not everyone has
> access to Cambridge?s HPC and probably even less have 76 hours to
> spare.

The same as for the `phenoDisco` analysis and `svm`, the TL results
are stored as a `RDS` in `pRolocdata` and are loaded in the
background. We have added the code required to the appendix so that
users can load the results directly.

> Your comment on the increase suitability of classification instead
> of clustering (when additional information on classes is available)
> at the bottom of page 35 could be more pronounced - for educational
> reasons.

To address the above comment on suitability we have added a few
additional points on the challenges of using clustering for this type
of data.

We generally find supervised learning more suited to the task of
protein localisation prediction in which we use high-quality curated
marker proteins to build a classifier, instead of using an entirely
unsupervised approach to look for clusters and then look for
enrichment of organelles and complexes. In the latter we do not make
good use of valuable prior knowledge, and in our experience
unsupervised clustering can be extremely difficult due to (i) the
loose definition of what constitutes a cluster (for example whether it
is defined by the quantitative data or the localisation information),
(ii) the influence of the algorithm assumption on the cluster
identification (for example parametric or non-parametric) and (iii)
poor estimates of the number of clusters that may appear in the
data.

### Minor comments:

> I was not able to naively reproduce the workflow from the R commands
> in the article due to an error installing pRolocdata on a Windows
> machine. On OS X it was smooth.

We didn't experience any Windows-specific problems. If you re-try the
installation and please let us know if you still have any issues by
opening an issue (https://github.com/lgatto/pRolocdata/issues) or by
posting this issue on the Bioconductor Support site
(https://support.bioconductor.org).

> On page 10 line 2 there is a ?to? missing.

In this version we currently can't find the missing 'to'.

> I never came across the verb imputate in the context of missing
> values, I guess the proper term is impute.

This has been changed to read "We can impute missing data..."

> On page 11 the image2 function is called after the filterNA function
> a couple of lines above. This however would result in an only black
> heat map (since there are no more missing). The image2 function
> should be called before the filterNA function. Since the reader does
> not see the chunk options, it could be puzzling.

This was an editing mistake and has now been rectified.

> For completeness sake there should also be an
> install.packages(c("hexbin", "rgl")) somewhere to generate the
> second PCA-plot and the 3D plot. Moreover, Mac users will need to
> install xquartz to use rgl properly.

A footnote has been added here to tell users that the package `rgl`
may need to be installed with `install.packages("rgl")` and mac users
may need to install xquartz if it's not already installed. LG: this
one is still missing.

> On page 14 the plotting code chunk is off track - in the middle of
> the marker sets output.

This has now been rectified.

> On page 18: ...wanted to highlight a proteins with the ... -> lose
> the a and later in the sentence there is a 'create a' too many.

These typos have been rectified.

> Direct comparisons of individual channels in replicated experiments
> do not provide?

This typo has been rectified.

> You may want to consider adding a layout(1) or similar, after
> changing the mfrow argument of the parameters to accommodate 2
> panels, such that the uncanny reader does not get confused.

We would prefer to keep the code as it is and not introduce more noise
with calls to other functions such as `layout`. The workflow is not
aimed at teaching R. Users should have some basic knowledge of R
before tackling this tutorial.

> I would prefer links to referred sections of the text, but that may
> be personal taste?

This is a comment for F1000. We cannot control the linking of sections
in the final version.

> Page 23: One should note that the decreasing the GS, and increasing
> the ...at least one the too many, probably two.

We have reworded this sentence as requested.

> On page 25: We find the general tendancy to be that it is not the
> choice ...tendency?

This typo has been rectified.

> On page 28 you refer to '...the code chunk below...' for Figure 17,
> however, the following code chunk is generating Figure 16 (which is
> above and btw not referenced in the text). Maybe force your figures
> a little to float where you want them/refer to them.

We have now referenced Figure 16 in the text and made sure that the
code chunks and figures follow inline where they are referenced in the
text.

> On page 28: ...by extracting the median or 3rd quantile score per
> organelle? do you mean quartile? Otherwise I do not follow.

Thank you, yes this is a typo and has now been changed to 'quartile'.

> On page 32: package to query the relevent database ? relevant?

This typo has been rectified.

> On page 32 - there is something wrong with this sentence: To remove
> the 4 classes and create a new column of markers in the feature data
> called tlmarkers to use for the analysis:

This sentence is not needed here and so it has now been removed as it
essentially reiterates what is said in the above paragraph.

> On page 34: From examining the parameter seach plots as described in
> section Optimisation? search!

This typo has been rectified.

> On page 36: ?and later reload the object using save. -> that would
> be ?load? then!

This typo has been rectified.

> On page 38 - I fully agree with the following sentence, but right
> after the updating comment it kind of seems misplaced? Maybe add a
> title like Getting help?

We have changed the title of this section to 'Session information and
getting help' to clarify this section of the tutorial.

## Reviewer 2: Leonard J. Foster

> This manuscript describes a Bioconductor workflow for analyzing
> subcellular proteomics data. It is very detailed and comprehensive
> and will be useful for others in the field.

Many thanks for your comments, we have responded to them inset below.

### A few comments:

> Some clearer statement early on would help to clarify for readers
> what types of data this works with. I know that the authors indicate
> that the example they use is 10-plex TMT and that it can be used
> with label-free or other labels, but that is not what I am referring
> to. Rather, structure of the experiment. That is, that one needs
> systematic quantitative data on all the different relevant fractions
> from a cell, as opposed to someone who perhaps did a differential
> centrifugation experiment to isolate a couple fractions and then
> wants to apply this (my understanding is that this latter example
> would not be usable).

From a purely technical point of view, we need a data matrix with
features (typically proteins or protein groups) along the rows and
sub-cellular fractions along the columns. Given the requirement for
complete (or near-complete) quantitative vectors along all fractions
to assure best results, a data set that would only contain one
quantitation value per fraction would not work. However, separation
using differential centrifugation, or any separation that generates
rganelle-specific separation profiles is a good fit for pRoloc. For
instance, the type of data generated by methods such as described in
Itzhak et al. (2016) are well fitted for our software:

```
> library(pRoloc)
> library(pRolocdata)
> data(itzhak2016stcSILAC)
> ## 6 combined replicates of 5 fractions each
> dim(itzhak2016stcSILAC)
[1] 5265   30
> plot2D(itzhak2016stcSILAC)
```

Itzhak DN, Tyanova S, Cox J, Borner GH. Global, quantitative and
dynamic mapping of protein subcellular localization. Elife. 2016 Jun
9;5. pii: e16950.  doi: 10.7554/eLife.16950. PubMed PMID: 27278775;
PubMed Central PMCID: PMC4959882.

If there were only very few fractions (say cytosol, nucleus and plasma
membrane), the infrastructure could still be used, although for less
benefits.

> How do the authors recommend collapsing replicates? This could be
> covered in the section dedicated to the Compare function. Two
> replicates will (almost) never agree 100% so how are discrepancies
> handled?

Currently, we recommend to visualise different replicates on their
own, to confirm that they are of sufficient quality, and then combine
them, retaining the proteins that have been quantified over all
replicatedd experiments. This allows to obtain localisation
information over all replicated data. We however do not explicitly
assess the variability using this approach. This could be done by
analysing replicated independently and then compare the coherence of
the classification results. Proteins that are only observed in some
replicates could be rescued by repeating the analysis using only the
relevant (possibly unique) replicate(s).

### 2018-05-22 bioc-workflow-annotated.tex
make annotated pdf

@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.5.10
+##  MSnbase version: 2.7.1
 \end{verbatim}
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:03:28 2018 
+## Combined [5489,10] and [6268,10] MSnSets Tue May 22 16:06:39 2018 
 ##  MSnbase version: 2.5.9
@@
 ## - - - Processing information - - -
-## Combined [5489,10] and [6268,10] MSnSets Wed Apr 11 14:03:28 2018 
-## Subset [6725,20][5032,20] Wed Apr 11 14:03:28 2018 
-## Removed features with more than 0 NAs: Wed Apr 11 14:03:28 2018 
-## Dropped featureData's levels Wed Apr 11 14:03:28 2018 
+## Combined [5489,10] and [6268,10] MSnSets Tue May 22 16:06:39 2018 
+## Subset [6725,20][5032,20] Tue May 22 16:06:39 2018 
+## Removed features with more than 0 NAs: Tue May 22 16:06:39 2018 
+## Dropped featureData's levels Tue May 22 16:06:39 2018 
 ##  MSnbase version: 2.5.9
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Wed Apr 11 14:03:30 2018 
+##  Created on Tue May 22 16:06:41 2018 
 ##  Description:
@@
 
-% latex table generated in R 3.5.0 by xtable 1.8-2 package
-% Wed Apr 11 14:03:30 2018
+% latex table generated in R 3.6.0 by xtable 1.8-2 package
+% Tue May 22 16:06:41 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Wed Apr 11 14:03:29 2018 
-## Added markers from  'pdres' marker vector. Wed Apr 11 14:03:30 2018 
-##  MSnbase version: 2.5.10
+## Added markers from  'mrk' marker vector. Tue May 22 16:06:40 2018 
+## Added markers from  'pdres' marker vector. Tue May 22 16:06:42 2018 
+##  MSnbase version: 2.7.1
 \end{verbatim}
@@
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# \\\#\# Attaching package: 'gridExtra'}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# The following object is masked from 'package:MSnbase':\\\#\# \\\#\#\ \ \ \  combine}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# The following object is masked from 'package:Biobase':\\\#\# \\\#\#\ \ \ \  combine}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# The following object is masked from 'package:BiocGenerics':\\\#\# \\\#\#\ \ \ \  combine}}\end{kframe}
 \includegraphics[width=.65\textwidth]{figure/visualiseOptHide-1}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4318484               0.3059444               0.3846145 
+##               0.4358428               0.3026164               0.3923006 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6952120               0.6096139               0.4308645 
+##               0.6784764               0.6086725               0.4306162 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4099447               0.5950343               0.9499410 
+##               0.4211113               0.5801971               0.9497222 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7954586               0.7096376               0.3119884 
+##               0.7959319               0.7094645               0.3164993 
 ##         Plasma membrane              Proteasome 
-##               0.7244522               0.4199965
+##               0.7120804               0.4217623
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4318484               0.3059444               0.3846145 
+##               0.4358428               0.3026164               0.3923006 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6952120               0.6096139               0.4308645 
+##               0.6784764               0.6086725               0.4306162 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4099447               0.5950343               0.9499410 
+##               0.4211113               0.5801971               0.9497222 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7954586               0.7096376               0.3119884 
+##               0.7959319               0.7094645               0.3164993 
 ##         Plasma membrane              Proteasome 
-##               0.7244522               0.4199965
+##               0.7120804               0.4217623
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      85                     173                      87 
+##                      86                     170                      89 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     298                     472                     103 
+##                     301                     475                      96 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      26                     124                     521 
+##                      27                     121                     523 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     229                     346                      38 
+##                     230                     344                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     317                     155                    2058
+##                     321                     153                    2058
 \end{verbatim}
@@
 \begin{verbatim}
-## R Under development (unstable) (2018-03-21 r74432)
-## Platform: x86_64-apple-darwin15.6.0 (64-bit)
-## Running under: macOS High Sierra 10.13.3
+## R Under development (unstable) (2018-04-02 r74505)
+## Platform: x86_64-pc-linux-gnu (64-bit)
+## Running under: Ubuntu 14.04.5 LTS
 ## 
 ## Matrix products: default
-## BLAS: /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib
-## LAPACK: /Library/Frameworks/R.framework/Versions/3.5/Resources/lib/libRlapack.dylib
+## BLAS: /usr/lib/atlas-base/atlas/libblas.so.3.0
+## LAPACK: /usr/lib/lapack/liblapack.so.3.0
 ## 
 ## locale:
-## [1] en_GB.UTF-8/en_GB.UTF-8/en_GB.UTF-8/C/en_GB.UTF-8/en_GB.UTF-8
+##  [1] LC_CTYPE=en_GB.UTF-8       LC_NUMERIC=C              
+##  [3] LC_TIME=en_GB.UTF-8        LC_COLLATE=en_GB.UTF-8    
+##  [5] LC_MONETARY=en_GB.UTF-8    LC_MESSAGES=en_GB.UTF-8   
+##  [7] LC_PAPER=en_GB.UTF-8       LC_NAME=C                 
+##  [9] LC_ADDRESS=C               LC_TELEPHONE=C            
+## [11] LC_MEASUREMENT=en_GB.UTF-8 LC_IDENTIFICATION=C       
 ## 
@@
 ## other attached packages:
-##  [1] gridExtra_2.3        pRolocdata_1.17.4    pRoloc_1.19.3       
-##  [4] MLInterfaces_1.59.1  cluster_2.0.6        annotate_1.57.2     
-##  [7] XML_3.98-1.10        AnnotationDbi_1.41.4 IRanges_2.13.28     
-## [10] S4Vectors_0.17.37    MSnbase_2.5.10       ProtGenerics_1.11.0 
-## [13] BiocParallel_1.13.1  mzR_2.13.5           Rcpp_0.12.16        
-## [16] Biobase_2.39.2       BiocGenerics_0.25.3  xtable_1.8-2        
-## [19] BiocStyle_2.7.8      knitr_1.20          
+##  [1] gridExtra_2.3        pRolocdata_1.19.0    pRoloc_1.21.0       
+##  [4] MLInterfaces_1.61.1  cluster_2.0.7-1      annotate_1.59.0     
+##  [7] XML_3.98-1.11        AnnotationDbi_1.43.1 IRanges_2.15.13     
+## [10] S4Vectors_0.19.5     MSnbase_2.7.1        ProtGenerics_1.13.0 
+## [13] BiocParallel_1.15.3  mzR_2.15.1           Rcpp_0.12.17        
+## [16] Biobase_2.41.0       BiocGenerics_0.27.0  xtable_1.8-2        
+## [19] BiocStyle_2.9.2      knitr_1.20          
 ## 
@@
 ##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
-##   [4] lazyeval_0.2.1        splines_3.5.0         ggvis_0.4.3          
+##   [4] lazyeval_0.2.1        splines_3.6.0         ggvis_0.4.3          
 ##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
-##  [10] foreach_1.4.4         BiocInstaller_1.29.4  htmltools_0.3.6      
-##  [13] viridis_0.5.0         gdata_2.18.0          magrittr_1.5         
+##  [10] foreach_1.4.4         BiocInstaller_1.31.1  htmltools_0.3.6      
+##  [13] viridis_0.5.1         gdata_2.18.0          magrittr_1.5         
 ##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
-##  [19] limma_3.35.13         recipes_0.1.2         gower_0.1.2          
+##  [19] limma_3.37.1          recipes_0.1.2         gower_0.1.2          
 ##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
-##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.0           
-##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
-##  [31] genefilter_1.61.1     bindr_0.1.1           impute_1.53.0        
-##  [34] survival_2.41-3       iterators_1.0.9       glue_1.2.0           
-##  [37] DRR_0.0.3             gtable_0.2.0          ipred_0.9-6          
-##  [40] zlibbioc_1.25.0       kernlab_0.9-25        ddalpha_1.3.1.1      
-##  [43] prabclus_2.2-6        DEoptimR_1.0-8        scales_0.5.0         
-##  [46] vsn_3.47.1            mvtnorm_1.0-7         DBI_0.8              
-##  [49] viridisLite_0.3.0     progress_1.1.2        foreign_0.8-70       
-##  [52] bit_1.1-12            proxy_0.4-21          mclust_5.4           
-##  [55] preprocessCore_1.41.0 lava_1.6              prodlim_1.6.1        
-##  [58] sampling_2.8          htmlwidgets_1.0       httr_1.3.1           
-##  [61] threejs_0.3.1         FNN_1.1               RColorBrewer_1.1-2   
-##  [64] fpc_2.1-11            modeltools_0.2-21     pkgconfig_2.0.1      
-##  [67] flexmix_2.3-14        nnet_7.3-12           caret_6.0-78         
-##  [70] tidyselect_0.2.4      rlang_0.2.0           reshape2_1.4.3       
-##  [73] munsell_0.4.3         mlbench_2.1-1         tools_3.5.0          
-##  [76] RSQLite_2.0           pls_2.6-0             broom_0.4.3          
-##  [79] evaluate_0.10.1       stringr_1.3.0         mzID_1.17.0          
-##  [82] yaml_2.1.18           ModelMetrics_1.1.0    bit64_0.9-7          
-##  [85] robustbase_0.92-8     randomForest_4.6-12   purrr_0.2.4          
-##  [88] dendextend_1.7.0      bindrcpp_0.2          nlme_3.1-136         
-##  [91] whisker_0.3-2         mime_0.5              RcppRoll_0.2.2       
-##  [94] biomaRt_2.35.12       compiler_3.5.0        e1071_1.6-8          
-##  [97] affyio_1.49.1         tibble_1.4.2          stringi_1.1.7        
-## [100] highr_0.6             lattice_0.20-35       trimcluster_0.1-2    
-## [103] Matrix_1.2-12         psych_1.7.8           gbm_2.1.3            
-## [106] pillar_1.2.1          MALDIquant_1.17       bitops_1.0-6         
-## [109] httpuv_1.3.6.2        R6_2.2.2              pcaMethods_1.71.0    
-## [112] affy_1.57.0           hwriter_1.3.2         codetools_0.2-15     
-## [115] MASS_7.3-49           gtools_3.5.0          assertthat_0.2.0     
-## [118] CVST_0.2-1            rprojroot_1.3-2       withr_2.1.2          
-## [121] mnormt_1.5-5          diptest_0.75-7        grid_3.5.0           
-## [124] rpart_4.1-13          timeDate_3043.102     tidyr_0.8.0          
-## [127] class_7.3-14          rmarkdown_1.9         lubridate_1.7.3      
-## [130] shiny_1.0.5           base64enc_0.1-3
+##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.1           
+##  [28] dplyr_0.7.5           RCurl_1.95-4.10       hexbin_1.27.2        
+##  [31] genefilter_1.63.0     bindr_0.1.1           impute_1.55.0        
+##  [34] DRR_0.0.3             survival_2.42-3.1     iterators_1.0.9      
+##  [37] glue_1.2.0            gtable_0.2.0          ipred_0.9-6          
+##  [40] zlibbioc_1.27.0       ddalpha_1.3.3         kernlab_0.9-26       
+##  [43] prabclus_2.2-6        DEoptimR_1.0-8        abind_1.4-5          
+##  [46] scales_0.5.0          vsn_3.49.0            mvtnorm_1.0-7        
+##  [49] DBI_1.0.0             viridisLite_0.3.0     progress_1.1.2       
+##  [52] magic_1.5-8           foreign_0.8-70        bit_1.1-13           
+##  [55] proxy_0.4-22          mclust_5.4            preprocessCore_1.43.0
+##  [58] lava_1.6.1            prodlim_2018.04.18    sampling_2.8         
+##  [61] htmlwidgets_1.2       httr_1.3.1            threejs_0.3.1        
+##  [64] FNN_1.1               RColorBrewer_1.1-2    fpc_2.1-11           
+##  [67] modeltools_0.2-21     pkgconfig_2.0.1       flexmix_2.3-14       
+##  [70] nnet_7.3-12           caret_6.0-79          tidyselect_0.2.4     
+##  [73] rlang_0.2.0           reshape2_1.4.3        later_0.7.2          
+##  [76] munsell_0.4.3         mlbench_2.1-1         tools_3.6.0          
+##  [79] msdata_0.21.0         RSQLite_2.1.1         pls_2.6-0            
+##  [82] broom_0.4.4           geometry_0.3-6        evaluate_0.10.1      
+##  [85] stringr_1.3.1         mzID_1.19.0           yaml_2.1.19          
+##  [88] ModelMetrics_1.1.0    bit64_0.9-7           robustbase_0.93-0    
+##  [91] randomForest_4.6-14   purrr_0.2.4           dendextend_1.8.0     
+##  [94] bindrcpp_0.2.2        nlme_3.1-137          whisker_0.3-2        
+##  [97] mime_0.5              RcppRoll_0.2.2        biomaRt_2.37.1       
+## [100] compiler_3.6.0       
+##  [ reached getOption("max.print") -- omitted 37 entries ]
 \end{verbatim}

### 2018-05-22 bioc-workflow-resubmit.tex
make annotated pdf

@@
 ## - - - Processing information - - -
-##  MSnbase version: 2.4.2
+##  MSnbase version: 2.7.1
 \end{verbatim}
@@
 ## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:40:33 2018 
-##  MSnbase version: 1.21.7
+## Combined [5489,10] and [6268,10] MSnSets Tue May 22 16:07:39 2018 
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 ## experimentData: use 'experimentData(object)'
+##   pubMedIds: 26754106 
 ## Annotation:  
 ## - - - Processing information - - -
-## Combined [6725,20] and [6268,10] MSnSets Tue Apr 17 15:40:33 2018 
-## Subset [6725,20][5032,20] Tue Apr 17 15:40:34 2018 
-## Removed features with more than 0 NAs: Tue Apr 17 15:40:34 2018 
-## Dropped featureData's levels Tue Apr 17 15:40:34 2018 
-##  MSnbase version: 1.21.7
+## Combined [5489,10] and [6268,10] MSnSets Tue May 22 16:07:39 2018 
+## Subset [6725,20][5032,20] Tue May 22 16:07:40 2018 
+## Removed features with more than 0 NAs: Tue May 22 16:07:40 2018 
+## Dropped featureData's levels Tue May 22 16:07:40 2018 
+##  MSnbase version: 2.5.9
 \end{verbatim}
@@
 ## Traceable object of class "FeaturesOfInterest"
-##  Created on Tue Apr 17 15:40:38 2018 
+##  Created on Tue May 22 16:07:42 2018 
 ##  Description:
@@
 
-% latex table generated in R 3.4.4 by xtable 1.8-2 package
-% Tue Apr 17 15:40:38 2018
+% latex table generated in R 3.6.0 by xtable 1.8-2 package
+% Tue May 22 16:07:42 2018
 \begin{table}[ht]
@@
 ## - - - Processing information - - -
-## Added markers from  'mrk' marker vector. Tue Apr 17 15:40:37 2018 
-## Added markers from  'pdres' marker vector. Tue Apr 17 15:40:39 2018 
-##  MSnbase version: 2.4.2
+## Added markers from  'mrk' marker vector. Tue May 22 16:07:41 2018 
+## Added markers from  'pdres' marker vector. Tue May 22 16:07:42 2018 
+##  MSnbase version: 2.7.1
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4382130               0.3040042               0.3797701 
+##               0.4343633               0.3047074               0.3832276 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6897383               0.6007307               0.4284784 
+##               0.6910080               0.6085463               0.4233563 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4200181               0.5879109               0.9502066 
+##               0.4220800               0.5889173               0.9497770 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951514               0.7100151               0.3166095 
+##               0.7936060               0.7103170               0.3143802 
 ##         Plasma membrane              Proteasome 
-##               0.7050611               0.4169158
+##               0.7159213               0.4144325
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##               0.4382130               0.3040042               0.3797701 
+##               0.4343633               0.3047074               0.3832276 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##               0.6897383               0.6007307               0.4284784 
+##               0.6910080               0.6085463               0.4233563 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##               0.4200181               0.5879109               0.9502066 
+##               0.4220800               0.5889173               0.9497770 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##               0.7951514               0.7100151               0.3166095 
+##               0.7936060               0.7103170               0.3143802 
 ##         Plasma membrane              Proteasome 
-##               0.7050611               0.4169158
+##               0.7159213               0.4144325
 \end{verbatim}
@@
 ##            40S Ribosome            60S Ribosome      Actin cytoskeleton 
-##                      86                     169                      85 
+##                      84                     172                      85 
 ##                 Cytosol   Endoplasmic reticulum                Endosome 
-##                     296                     477                      94 
+##                     296                     476                     103 
 ##    Extracellular matrix                Lysosome           Mitochondrion 
-##                      27                     123                     522 
+##                      25                     124                     522 
 ##     Nucleus - Chromatin Nucleus - Non-chromatin              Peroxisome 
-##                     230                     344                      39 
+##                     230                     342                      38 
 ##         Plasma membrane              Proteasome                 unknown 
-##                     325                     157                    2058
+##                     320                     157                    2058
 \end{verbatim}
@@
 ## attached base packages:
-## [1] stats4    parallel  methods   stats     graphics  grDevices utils    
-## [8] datasets  base     
+## [1] stats4    parallel  stats     graphics  grDevices utils     datasets 
+## [8] methods   base     
 ## 
 ## other attached packages:
-##  [1] pRolocdata_1.16.0    pRoloc_1.18.0        MLInterfaces_1.58.1 
-##  [4] cluster_2.0.7-1      annotate_1.56.2      XML_3.98-1.10       
-##  [7] AnnotationDbi_1.40.0 IRanges_2.12.0       S4Vectors_0.16.0    
-## [10] MSnbase_2.4.2        ProtGenerics_1.10.0  BiocParallel_1.12.0 
-## [13] mzR_2.12.0           Rcpp_0.12.16         Biobase_2.38.0      
-## [16] BiocGenerics_0.24.0  xtable_1.8-2         BiocStyle_2.6.1     
+##  [1] pRolocdata_1.19.0    pRoloc_1.21.0        MLInterfaces_1.61.1 
+##  [4] cluster_2.0.7-1      annotate_1.59.0      XML_3.98-1.11       
+##  [7] AnnotationDbi_1.43.1 IRanges_2.15.13      S4Vectors_0.19.5    
+## [10] MSnbase_2.7.1        ProtGenerics_1.13.0  BiocParallel_1.15.3 
+## [13] mzR_2.15.1           Rcpp_0.12.17         Biobase_2.41.0      
+## [16] BiocGenerics_0.27.0  xtable_1.8-2         BiocStyle_2.9.2     
 ## [19] knitr_1.20          
@@
 ##   [1] backports_1.1.2       plyr_1.8.4            igraph_1.2.1         
-##   [4] lazyeval_0.2.1        splines_3.4.4         ggvis_0.4.3          
+##   [4] lazyeval_0.2.1        splines_3.6.0         ggvis_0.4.3          
 ##   [7] crosstalk_1.0.0       ggplot2_2.2.1         digest_0.6.15        
-##  [10] foreach_1.4.4         BiocInstaller_1.28.0  htmltools_0.3.6      
+##  [10] foreach_1.4.4         BiocInstaller_1.31.1  htmltools_0.3.6      
 ##  [13] viridis_0.5.1         gdata_2.18.0          magrittr_1.5         
 ##  [16] memoise_1.1.0         doParallel_1.0.11     sfsmisc_1.1-2        
-##  [19] limma_3.34.9          recipes_0.1.2         gower_0.1.2          
+##  [19] limma_3.37.1          recipes_0.1.2         gower_0.1.2          
 ##  [22] rda_1.0.2-2           dimRed_0.1.0          lpSolve_5.6.13       
 ##  [25] prettyunits_1.0.2     colorspace_1.3-2      blob_1.1.1           
-##  [28] dplyr_0.7.4           RCurl_1.95-4.10       hexbin_1.27.2        
-##  [31] genefilter_1.60.0     bindr_0.1.1           impute_1.52.0        
-##  [34] DRR_0.0.3             survival_2.41-3       iterators_1.0.9      
+##  [28] dplyr_0.7.5           RCurl_1.95-4.10       hexbin_1.27.2        
+##  [31] genefilter_1.63.0     bindr_0.1.1           impute_1.55.0        
+##  [34] DRR_0.0.3             survival_2.42-3.1     iterators_1.0.9      
 ##  [37] glue_1.2.0            gtable_0.2.0          ipred_0.9-6          
-##  [40] zlibbioc_1.24.0       kernlab_0.9-25        ddalpha_1.3.2        
+##  [40] zlibbioc_1.27.0       ddalpha_1.3.3         kernlab_0.9-26       
 ##  [43] prabclus_2.2-6        DEoptimR_1.0-8        abind_1.4-5          
-##  [46] scales_0.5.0.9000     vsn_3.46.0            mvtnorm_1.0-7        
-##  [49] DBI_0.8               viridisLite_0.3.0     progress_1.1.2       
-##  [52] magic_1.5-8           foreign_0.8-69        bit_1.1-12           
-##  [55] proxy_0.4-22          mclust_5.4            preprocessCore_1.40.0
-##  [58] lava_1.6.1            prodlim_1.6.1         sampling_2.8         
-##  [61] htmlwidgets_1.0       httr_1.3.1            threejs_0.3.1        
+##  [46] scales_0.5.0          vsn_3.49.0            mvtnorm_1.0-7        
+##  [49] DBI_1.0.0             viridisLite_0.3.0     progress_1.1.2       
+##  [52] magic_1.5-8           foreign_0.8-70        bit_1.1-13           
+##  [55] proxy_0.4-22          mclust_5.4            preprocessCore_1.43.0
+##  [58] lava_1.6.1            prodlim_2018.04.18    sampling_2.8         
+##  [61] htmlwidgets_1.2       httr_1.3.1            threejs_0.3.1        
 ##  [64] FNN_1.1               RColorBrewer_1.1-2    fpc_2.1-11           
@@
 ##  [70] nnet_7.3-12           caret_6.0-79          tidyselect_0.2.4     
-##  [73] rlang_0.2.0           reshape2_1.4.3        munsell_0.4.3        
-##  [76] mlbench_2.1-1         tools_3.4.4           msdata_0.18.0        
-##  [79] RSQLite_2.1.0         pls_2.6-0             broom_0.4.4          
-##  [82] evaluate_0.10.1       geometry_0.3-6        stringr_1.3.0        
-##  [85] mzID_1.16.0           yaml_2.1.18           ModelMetrics_1.1.0   
-##  [88] bit64_0.9-7           robustbase_0.92-8     randomForest_4.6-14  
-##  [91] purrr_0.2.4           dendextend_1.7.0      bindrcpp_0.2.2       
-##  [94] nlme_3.1-137          whisker_0.3-2         mime_0.5             
-##  [97] RcppRoll_0.2.2        biomaRt_2.34.2        compiler_3.4.4       
-## [100] e1071_1.6-8          
-##  [ reached getOption("max.print") -- omitted 37 entries ]
+##  [73] rlang_0.2.0           reshape2_1.4.3        later_0.7.2          
+##  [76] munsell_0.4.3         mlbench_2.1-1         tools_3.6.0          
+##  [79] msdata_0.21.0         RSQLite_2.1.1         pls_2.6-0            
+##  [82] broom_0.4.4           geometry_0.3-6        evaluate_0.10.1      
+##  [85] stringr_1.3.1         mzID_1.19.0           yaml_2.1.19          
+##  [88] ModelMetrics_1.1.0    bit64_0.9-7           robustbase_0.93-0    
+##  [91] randomForest_4.6-14   purrr_0.2.4           dendextend_1.8.0     
+##  [94] bindrcpp_0.2.2        nlme_3.1-137          whisker_0.3-2        
+##  [97] mime_0.5              RcppRoll_0.2.2        biomaRt_2.37.1       
+## [100] compiler_3.6.0       
+##  [ reached getOption("max.print") -- omitted 38 entries ]
 \end{verbatim}

### 2018-05-22 replies-to-reviewers.Rmd
update replies

--- 2018-05-16 replies-to-reviewers.md
@@
 
-## Reviewer 1 - Daniel J. Stekhoven1
+## Reviewer 1 - Daniel J. Stekhoven
 
@@
 > on page 33 - this needs to be downloadable, since not everyone has
-> access to Cambridge?s HPC and probably even less have 76 hours to
+> access to Cambridge's HPC and probably even less have 76 hours to
 > spare.
@@
 
-> On page 10 line 2 there is a ?to? missing.
+> On page 10 line 2 there is a *to* missing.
 
@@
 
-> On page 32: package to query the relevent database ? relevant?
+> On page 32: package to query the relevent database *relevant*
 
@@
 > On page 34: From examining the parameter seach plots as described in
-> section Optimisation? search!
+> section Optimisation... search!
 
@@
 
-> On page 36: ?and later reload the object using save. -> that would
-> be ?load? then!
+> On page 36: and later reload the object using save. -> that would be
+> `load` then!
 
@@
 instance, the type of data generated by methods such as described in
-Itzhak et al. (2016) are well fitted for our software:
+Itzhak et al. (2016)[^1] are well fitted for our software:
 
-```
-> library(pRoloc)
-> library(pRolocdata)
-> data(itzhak2016stcSILAC)
-> ## 6 combined replicates of 5 fractions each
-> dim(itzhak2016stcSILAC)
-[1] 5265   30
-> plot2D(itzhak2016stcSILAC)
+```{r plot2d, message = FALSE, fig.width = 6, fig.height = 6}
+library(pRoloc)
+library(pRolocdata)
+data(itzhak2016stcSILAC)
+## 6 combined replicates of 5 fractions each
+dim(itzhak2016stcSILAC)
+plot2D(itzhak2016stcSILAC)
+addLegend(itzhak2016stcSILAC, where = "topright", cex = .7)
 ```
 
-Itzhak DN, Tyanova S, Cox J, Borner GH. Global, quantitative and
-dynamic mapping of protein subcellular localization. Elife. 2016 Jun
-9;5. pii: e16950.  doi: 10.7554/eLife.16950. PubMed PMID: 27278775;
-PubMed Central PMCID: PMC4959882.
+[^1]: Itzhak DN, Tyanova S, Cox J, Borner GH.
+Global, quantitative and dynamic mapping of protein subcellular localization.
+Elife.
+2016 Jun 9;5.
+pii: e16950.
+doi: 10.7554/eLife.16950.
+PubMed PMID: 27278775; PubMed Central PMCID: PMC4959882.
 

### 2018-05-23 replies-to-reviewers.Rmd
fix typo in replies

@@
 using differential centrifugation, or any separation that generates
-rganelle-specific separation profiles is a good fit for pRoloc. For
+organelle-specific separation profiles is a good fit for pRoloc. For
 instance, the type of data generated by methods such as described in
